### 2023-07-05 paper.tex
Update on Overleaf.

\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}
\usepackage[utf8]{inputenc} % für Umlaute
\usepackage{fontenc}

% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%%%% ---- Useful packages

%%% --- Figures and Graphics
\usepackage{graphicx}
\usepackage{subcaption}

%%% --- Colors
\usepackage{color}
\usepackage[dvpinames]{xcolor}

%%% --- Tabulars
\usepackage{booktabs}

%%% --- Mathematics
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools}
\usepackage{bm} % bold mathematics
\usepackage{bbm} % for blackboard 1
\usepackage[linesnumbered,lined,algo2e,figure,boxed]{algorithm2e}
\usepackage[short]{optidef} % for aligning linear programs

%%% --- Other
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
%\usepackage[hidelinks]{hyperref}
\usepackage[style=authoryear,backend=bibtex]{biblatex}
%\addbibresource{library.bib}
\usepackage{csquotes}
\usepackage[shortcuts]{extdash} % For Hyphens in English
%\usepackage{acro}
% \DeclareAcronym{pdf}{short=PDF, long={probability distribution function}}

%%% --- Commenting, Todonotes
\usepackage{comment, todonotes}
\newcommand{\ar}{$\rightarrow$}
\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}

%%%% ---- Own Commands
%%% --- Theorem Settings
\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}
\newtheorem{exa}{Example}
\newtheorem{rem}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

%%% --- Math Commands

\newcommand{\lag}[1][l]{\Delta_{#1}}
\DeclarePairedDelimiter{\abs}\lvert\rvert

%%% --- Other Commands
\renewcommand{\arraystretch}{1.2}

\title{Trending in Nowcasting}
\author{Oliver Grothe, Bolin Liu, Jonas Rieger}

\begin{document}
\maketitle

%\begin{abstract}
%Your abstract.
%\end{abstract}

\input{sections/01-introduction.tex}

\section{Notation}
\begin{itemize}
  \item Sei $T = \{1, 2, \dots\}$ die Zeitindexmenge, wobei für jeden Zeitpunkt $t \in T$ eine Realisierung vorliegt (die ggf. im Nachhinein veröffentlicht wird).
  \item Sei $(Y_t)_{t \in T}$ die Zeitreihe der Realisierungen, dabei steht $t$ für den Zeitpunkt, auf den sich der Wert bezieht, nicht den Veröffentlichungszeitpunkt.
  \item Sei $K = \{1, 2, \dots\}$ die Menge der Nowcaster.
	\begin{itemize}
	  \item Sei $X_{t \lvert \tau}^k (k \in K, t \in T, \tau \in T_t^k)$ der Nowcast von $k$ bezüglich des Zeitpunktes $t$, der am Zeitpunkt $\tau$ \textbf{veröffentlicht} wird (oder: berechnet / dessen Informationen sich auf Zeitpunkt $\tau$ beziehen).
	  \item Sei $g: \mathbb{R}^{\lvert T_t^k \lvert} \rightarrow \mathbb{R}$ eine Aggregationsfunktion, die alle Nowcasts bezüglich eines Zeitpunkts zu einem Nowcast zusammenfasst ($X_t^k \coloneqq g((X_{t \lvert \tau}^k)_{\tau \in T_t^k})$). Falls jeweils nur ein Nowcast veröffentlich wird, wähle $g(x) = x$.
	  \item Sei $(X_t^k)_{t \in T}$ die Zeitreihe der aggregierten Nowcasts.
	\end{itemize}
  \item Sei $\lag[l]$ der lag-$l$-Operator, der für eine Zeitreihe $(Z_t)_{t \in T}$ definiert wird durch 
		\begin{equation}
			\lag[l]Z_t = Z_{t+l} - Z_t \quad (t \in T)
		\end{equation} 
  \item Der Anteil an konkordanten Punkten für Realisierungen $y = (y_1, y_2, \dots, y_n)$ und $(x_1, x_2, \dots, x_n)$ bezüglich des Lag $l$ ergibt sich dann durch
	\begin{equation}
  		k (x, y; l, \epsilon) = \frac{\sum_{t}^{n-l} k^s (\lag y_t, \lag x_t; l, \epsilon)}{\sum_{t}^{n-l} k^\epsilon (\lag y_t, \lag x_t)}
	\end{equation}
	wobei $k^s$ die Indikatorfunktion für die Konkordanz zweier Werte außerhalb der exclusion area ist \todo{stimmt Konkordanz hier?}
	\begin{equation}
  		k^s (x, y; \epsilon) \coloneqq \begin{cases}
  			1 &, \text{falls}\ xy > 0\ \text{und}\ k^\epsilon(x, y; \epsilon) = 1\\
  			0 &, \text{sonst}
  		\end{cases}
	\end{equation}
	und $k^\epsilon$ die Indikatorfunktion für Punkte außerhalb der exclusion area ist
	\begin{equation}
  		k^\epsilon (x, y; \epsilon) \coloneqq \begin{cases}
  			1 &, \text{falls} \ \abs{x} > \epsilon \text{oder} \ \abs{y} > \epsilon \\
  			0 &, \text{sonst}
  		\end{cases}
	\end{equation}
\end{itemize}

\section{Statistische Modellierung}

\begin{itemize}
  \item Sei $l$ der zeitliche Abstand, mit dem Realisationen verfügbar werden
  \item $Y_{t + l} = Y_t + \lag Y_t$
  \item Die Nowcast schätzen (fehlerbehaftet) $\lag Y_t$ mithilfe von Daten, die zum Zeitpunkt $\tau \in T_t^k$ verfügbar sind, und dem neuesten, bekannten Wert $Y_t$:
	\begin{equation}
  		X_{t+l \lvert \tau}^k = Y_t + (\widehat{\lag Y_t})_{t \lvert \tau \in T_t^k}^k,
	\end{equation}
	\todo{Stimmt das so?}
	\todo{Ist festes $l$ zu starke Annahme?}
	und damit 
		\begin{equation}
  			\lag Y_t = (\widehat{\lag Y_t})_{t \lvert \tau \in T_t^k}^k + \varepsilon.
		\end{equation}

\end{itemize}

\section{Trending}

\begin{itemize}
  \item \enquote{Schwaches} Trending für lag $l$:
   \begin{equation}
  	P(\lag Y_t \widehat{(\lag Y_t)^k} > 0 \:\lvert\: \abs{\lag Y_t} > \epsilon) > P(\lag Y_t \widehat{(\lag Y_t)^k} < 0 \:\lvert\: \abs{\lag Y_t} > \epsilon)
   \end{equation}

\end{itemize}

%\printbibliography

\end{document}

### 2023-07-19 paper.tex
Update on Overleaf.

@@
 
+%%% --- ToDo List
+\usepackage{enumitem}
+\newlist{todolist}{itemize}{2}
+\setlist[todolist]{label=$\square$}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
+\hspace{-2.5pt}}
+\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}
+
 \title{Trending in Nowcasting}
@@
 %\end{abstract}
+
+\section{ToDos}
+
+\begin{todolist}
+\item[\done] Vorlage erstellen (Jonas)
+\item[\done] Schätzer einarbeiten (Jonas)
+\item[\done] Alternativvorschlag Notation (Bolin)
+\item[\done] Aufschreiben der Diskussionsergebnisse (Bolin): Grundproblems, Anforderungen und Beispielmaß
+\item Eigenschaften des Beispiel-Maßes untersuchen/ angeben (Jonas, Bolin ) 
+\end{todolist}
 
@@
 	\end{equation}
-	wobei $k^s$ die Indikatorfunktion für die Konkordanz zweier Werte außerhalb der exclusion area ist \todo{stimmt Konkordanz hier?}
+	wobei $k^s$ die Indikatorfunktion für die Konkordanz zweier Werte außerhalb der exclusion area ist
 	\begin{equation}
@@
 
+\section{Trending-Problem}
+\subsection{Grundproblem}
+Gegeben sind zwei Zeitreihen $(x_t)$ und $(y_t)$, $t\in \{t_0,...,t_N\}$.
+Dabei kann $(x_t)$ die zeitliche Entwicklung einer Zielgröße und $(y_t)$ die zeitlichen Schätzungen eines Nowcasters darstellen.
+Uns interessiert, ob die Entwicklungen der beiden Zeitreihen den gleichen Trend aufweisen bzw.
+ob der Nowcaster die Entwicklung der zu schätzenden Größe mit dem richtigen Trend schätzen kann.
+Es sollte ein Maß konstruiert werden, das beschreibt, wie gut zwei Zeitreihen den gleichen Trend aufweisen.
+\subsection{Anforderungen an einem Maß/ Intuitionen}
+
+Zwei Zeitreihen weisen den gleichen Trend auf, wenn für zwei beliebige Zeitpunkte $t$ und $\tau$ gilt: $(x_t-x_\tau)(y_t-y_\tau)>0$ oder $x_t=x_\tau \land y_t = y_\tau$.
+
+\todo{Kann man theoretisch nicht einfach mit Rang-korrelations-koeffizient Trending überprüfen?}
+\subsection{Ein einfaches Maß}
+Sei $\Delta_{t,\tau}^{x}:=x_\tau-x_t$.
+\begin{equation}
+  		R_{t,\tau}^{x,y} \coloneqq 
+        \begin{cases}
+  		    \frac{\Delta_{t,\tau}^{x}\Delta_{t,\tau}^{y}}   {\vert\Delta_{t,\tau}^{x}\vert\vert\Delta_{t,\tau}^{y}\vert } &, \text{falls} \Delta_{t,\tau}^{x}\neq 0 \land \Delta_{t,\tau}^{y}\neq0\\
+           1 &, \text{falls} \Delta_{t,\tau}^{x}=\Delta_{t,\tau}^{y}=0\\
+            -1 &, \text{sonst} 
+  	\end{cases}
+	\end{equation}
+Beispiel-Maß: $S(\bold{x},\bold{y})=\sum_{k=1}^{k^*}\frac{w_k}{N+1-k}\sum_{j=0}^{N-k}R_{t_j,t_{j+k}}^{x,y},$
+wobei $w_k, k=1,...,k^*$ eine Gewichtung für das Trending-Verhalten in verschiedenen Zeitdistanzen darstellt. Dabei erfüllt $w_k$ die folgenden Anforderungen:
+
+\begin{itemize}
+    \item $w_k$ ist streng monoton fallend in $k$. Hintergrund: je klein der betrachtete Zeitabstand ist, desto größer ist hier der Einfluss des Rauschens
+    \item $\sum_{k=1}^{k^*}w_k=1$
+\end{itemize}
+
+\subsection{Eigenschaften des vorgeschlagenen Maßes}
+\begin{itemize}
+    \item Seien $(x_n)$ und $(y_n)$ zwei Realisierungen eines Random Walk -Modells.\\
+    Hypothese: $S((x_n),(y_n))=0$. \\
+    Intuition:  Zwei Random Walks $(x_n)$ und $(y_n)$ zeigen kein gemeinsames Tendenzverhalten auf. 
+\end{itemize}
+
+ \subsection{Alternative Notation}
+
+\begin{itemize}
+	\item Seien $x_1, x_2, \dots, x_N$ die (aggregierten) Nowcasts; $y_1, x_2, \dots, y_N$ die realisierten Werte
+	\item Gewichteter Anteil an konkordanten Differenzen: 
+	\begin{equation}
+		\sum_{l = 1}^{l^*} \omega_l \sum_{t=1}^{N - l} \frac{\lag x_i \lag y_i}{\abs{\lag x_i} \abs{\lag{y_i}}}, 
+	\end{equation}
+	wobei $\infty / \infty \coloneqq 1$ und $\infty / - \infty \coloneqq -1$.
+	\item Evtl. geschicktere Lokalisierung (0: \enquote{perfekte} Gegenläufigkeit, 1: \enquote{Perfektes} Trending bis Lag $l^*$):
+	\begin{equation}
+		\sum_{l = 1}^{l^*} \omega_l \sum_{t=1}^{N - l} \frac{1}{2 (N-l)} \left( \frac{\lag x_i \lag y_i}{\abs{\lag x_i} \abs{\lag{y_i}}} + (N-l) \right)
+	\end{equation}
+\end{itemize}
+
 \section{Statistische Modellierung}
-
+\subsection{Modellierung}
 \begin{itemize}
@@
 	\begin{equation}
-  		X_{t+l \lvert \tau}^k = Y_t + (\widehat{\lag Y_t})_{t \lvert \tau \in T_t^k}^k,
+  		X_{t \lvert \tau}^k = Y_{\tau - l} + (\widehat{\lag[t-\tau+l] Y_t})_{t \lvert \tau}^k,
 	\end{equation}
-	\todo{Stimmt das so?}
-	\todo{Ist festes $l$ zu starke Annahme?}
-	und damit 
+	wobei
 		\begin{equation}
-  			\lag Y_t = (\widehat{\lag Y_t})_{t \lvert \tau \in T_t^k}^k + \varepsilon.
+  			\lag[t-\tau+l] Y_t =  (\widehat{\lag[t-\tau+l] Y_t})_{t \lvert \tau}^k + \varepsilon.
 		\end{equation}
@@
 \end{itemize}
-
-\section{Trending}
+\subsection{Weitere Ideen zu Trending}
 
 \begin{itemize}
-  \item \enquote{Schwaches} Trending für lag $l$:
-   \begin{equation}
-  	P(\lag Y_t \widehat{(\lag Y_t)^k} > 0 \:\lvert\: \abs{\lag Y_t} > \epsilon) > P(\lag Y_t \widehat{(\lag Y_t)^k} < 0 \:\lvert\: \abs{\lag Y_t} > \epsilon)
-   \end{equation}
-
+  \item \enquote{Schwaches} Trending für lag $l$: Wahrscheinlichkeit, in die gleiche Richtung zu zeigen ist größer als Wahrscheinlichkeit in die falsche Richtung zu zeigen (für lag $l$)
+  \item \enquote{Downwards}-Trending: Trends in die negative Richtung werden erkannt, in die positive Richtung nicht
+  \item \enquote{Upwards}-Trending: Trends in die positive Richtung werden erkannt, in die negative nicht
+  \item Nicht einfach nur 0-1-Kodierung für gleiches Vorzeichen von $\lag x$ und $\lag y$, sondern Gewichtung des $\mathbb{R}^2$, sodass zum Beispiel Punkte nahe der Achsen weniger Gewicht bekommen als solche nahe der Winkelhalbierenden
+  \item PCA auf $\lag x$ und $\lag y$, Bestrafung der zweiten Komponente (Abweichung von Gerade): Bestrafung würde aber mehr umfassen, als lediglich den Trend, sondern würde Abweichung von Linearität betreffen
+  \item Falls wir Exclusion Area wollen und $\lag Y_t$ heteroskedastisch: relative Werte betrachten?
 \end{itemize}

### 2023-12-04 paper.tex
New outline of paper

--- 2023-07-05 paper.tex
@@
-\documentclass{article}
+\documentclass[oneside]{article}
 
@@
 \usepackage{subcaption}
+\usepackage{tikz}
 
 %%% --- Colors
-\usepackage{color}
-\usepackage[dvpinames]{xcolor}
+\usepackage[]{xcolor}
 
@@
 \usepackage{booktabs}
+\usepackage{multirow} % cells over several rows
+\usepackage{tabularx}
 
@@
 %%% --- Commenting, Todonotes
-\usepackage{comment, todonotes}
+\usepackage{comment, todonotes, soul}
 \newcommand{\ar}{$\rightarrow$}
 \presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
-\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}
+%\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}
 
@@
 \DeclarePairedDelimiter{\abs}\lvert\rvert
+\DeclareMathOperator{\sign}{sign}
+\newcommand{\tmax}{\bar{t}}
+\newcommand{\ind}[1]{\mathbbm{1}\{#1\}}
+\newcommand{\ydiff}{\tilde{y}}
+\newcommand{\ydifft}{\tilde{y}^*}
+\newcommand{\xdiff}{\tilde{x}}
+\newcommand{\xdifft}{\tilde{x}^*}
+\newcommand{\Ydiff}{\tilde{Y}}
+\newcommand{\Xdiff}{\tilde{X}}
+\newcommand{\Prob}[1]{P(#1)}
+\newcommand{\mprob}{\tilde{m}}
 
@@
 
-\title{Trending in Nowcasting}
+%%% --- ToDo List
+\usepackage{enumitem}
+\newlist{todolist}{itemize}{2}
+\setlist[todolist]{label=$\square$}
+\usepackage{pifont}
+\newcommand{\cmark}{\ding{51}}%
+\newcommand{\xmark}{\ding{55}}%
+\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
+\hspace{-2.5pt}}
+\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}
+
+\title{Assessment and Optimization of Nowcast Measures for Trend Detection}
 \author{Oliver Grothe, Bolin Liu, Jonas Rieger}
@@
 
-%\begin{abstract}
-%Your abstract.
-%\end{abstract}
+\begin{abstract}
+Existing and new measures for the capability of capturing the trend of nowcasts are presented, evaluated and compared on synthetic and real-world data.
+\end{abstract}
 
-\input{sections/01-introduction.tex}
+\section{Introduction}
 
-\section{Notation}
 \begin{itemize}
-  \item Sei $T = \{1, 2, \dots\}$ die Zeitindexmenge, wobei für jeden Zeitpunkt $t \in T$ eine Realisierung vorliegt (die ggf. im Nachhinein veröffentlicht wird).
-  \item Sei $(Y_t)_{t \in T}$ die Zeitreihe der Realisierungen, dabei steht $t$ für den Zeitpunkt, auf den sich der Wert bezieht, nicht den Veröffentlichungszeitpunkt.
-  \item Sei $K = \{1, 2, \dots\}$ die Menge der Nowcaster.
+	\item Trending is an important feature of nowcasts to be identify changes in the situation before the actual quantity can be measured
+	\item Examples: Covid, BIP, Medicine, ... 
+\end{itemize}
+
+\section{Trend measures for nowcasts}
+
+Notation
 \begin{itemize}
-	  \item Sei $X_{t \lvert \tau}^k (k \in K, t \in T, \tau \in T_t^k)$ der Nowcast von $k$ bezüglich des Zeitpunktes $t$, der am Zeitpunkt $\tau$ \textbf{veröffentlicht} wird (oder: berechnet / dessen Informationen sich auf Zeitpunkt $\tau$ beziehen).
-	  \item Sei $g: \mathbb{R}^{\lvert T_t^k \lvert} \rightarrow \mathbb{R}$ eine Aggregationsfunktion, die alle Nowcasts bezüglich eines Zeitpunkts zu einem Nowcast zusammenfasst ($X_t^k \coloneqq g((X_{t \lvert \tau}^k)_{\tau \in T_t^k})$).
-Falls jeweils nur ein Nowcast veröffentlich wird, wähle $g(x) = x$.
-	  \item Sei $(X_t^k)_{t \in T}$ die Zeitreihe der aggregierten Nowcasts.
+  \item Let $y_1, \ldots, y_T$ be the values of the nowcasted quantity.
+  \item Let $x_t^k$ be the $k$-th nowcast for time $t = 1, \ldots, T$ (possibly after some aggregation if several raw nowcasts apply).
+  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change occurred between time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - y_{t-1}$ the change nowcaster $k$ issued between the time steps. 
+  \item Then, $\xdiff_t^k$ can be seen as point forecast of $\ydiff_t$. 
 \end{itemize}
-  \item Sei $\lag[l]$ der lag-$l$-Operator, der für eine Zeitreihe $(Z_t)_{t \in T}$ definiert wird durch 
+
+\subsection{What is the trend of a nowcast?}
+
+Trend: Issuing the right \enquote{direction} of change between two time steps
+
+\begin{itemize}
+  \item Translate into mathematical notation
+  \item Illustrative Example: Two nowcasts of quantity of interest, where one jumps up and down and one points in right direction with the same rmse
+\end{itemize}
+
+\subsection{Measures}
+
+%\begin{table}
+%	\centering
+%	\begin{tabular}{l l c c}
+%	\toprule
+%	& & \multicolumn{2}{c}{Observed trend} \\
+%	& & $\ydiff_t > 0$ & $\ydiff_t < 0$ \\	\cline{3-4}
+%	\multirow{2}{*}{Forecasted trend} & $\xdiff_t > 0$ & hit & false alarm\\
+%	& $\xdiff_t < 0$ & miss & correct negative \\
+%	\bottomrule
+%\end{tabular}
+%\end{table}
+
+Non-probabilistic measures:
+
+\begin{itemize}
+  \item Accuracy: \begin{equation}
+  	m_{\text{acc}} = \frac{\sum_{t=2}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=2}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-1}
+\end{equation}
+	\item Capability of detecting a positive trend (probability of detection)
 	\begin{equation}
-			\lag[l]Z_t = Z_{t+l} - Z_t \quad (t \in T)
+  		m_{\text{py}} = \frac{\sum_{t=2}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=2}^T \ind{\ydiff_t > 0}}
 	\end{equation}
-  \item Der Anteil an konkordanten Punkten für Realisierungen $y = (y_1, y_2, \dots, y_n)$ und $(x_1, x_2, \dots, x_n)$ bezüglich des Lag $l$ ergibt sich dann durch
+	\item False alarm rate \hl{Name does not fit} \begin{equation}
+  m_{\text{far}} = \frac{\sum_{t=2}^T \ind{\ydiff_t < 0, \xdiff_t > 0}}{\sum_{t=2}^T \ind{\xdiff_t > 0}}
+\end{equation}
+maybe use \enquote{positive} measure instead, i.e.,
 \begin{equation}
-  		k (x, y; l, \epsilon) = \frac{\sum_{t}^{n-l} k^s (\lag y_t, \lag x_t; l, \epsilon)}{\sum_{t}^{n-l} k^\epsilon (\lag y_t, \lag x_t)}
+  m_{\text{px}} = \frac{\sum_{t=2}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=2}^T \ind{\xdiff_t > 0}}  
 \end{equation}
-	wobei $k^s$ die Indikatorfunktion für die Konkordanz zweier Werte außerhalb der exclusion area ist \todo{stimmt Konkordanz hier?}
+	\item same measures for negative direction:
+	\begin{align}
+		m_{\text{ny}} &=  \frac{\sum_{t=2}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=2}^T \ind{\ydiff_t < 0}} \\
+		m_{\text{nx}} &=  \frac{\sum_{t=2}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=2}^T \ind{\xdiff_t < 0}}
+	\end{align}
+\end{itemize}
+
+Resulting measures:
+\begin{itemize}
+  \item Scoring rules for probabilities
+
+\begin{itemize}
+  \item Quadratic/Brier Score for $k$ possible outcomes: \begin{equation}
+  S((p_1, \dots, p_k), i) = 2 p_i - \sum_{l=1}^k p_i^2 - 1;
+\end{equation}
+for dichotomous outcomes ($p_1 = \hat{P}(\ydiff_t > 0)$: probability for $\ydiff_t > 0$):
 \begin{equation}
-  		k^s (x, y; \epsilon) \coloneqq \begin{cases}
-  			1 &, \text{falls}\ xy > 0\ \text{und}\ k^\epsilon(x, y; \epsilon) = 1\\
-  			0 &, \text{sonst}
-  		\end{cases}
-	\end{equation}
-	und $k^\epsilon$ die Indikatorfunktion für Punkte außerhalb der exclusion area ist
-	\begin{equation}
-  		k^\epsilon (x, y; \epsilon) \coloneqq \begin{cases}
-  			1 &, \text{falls} \ \abs{x} > \epsilon \text{oder} \ \abs{y} > \epsilon \\
-  			0 &, \text{sonst}
-  		\end{cases}
+  S(p_1, \ydiff_t) = p_1 \ind{\ydiff_t > 0} + (1 - p_1) (1 - \ind{\ydiff_t > 0}) + p_1^2 + (1-p_1)^2 - 1
 \end{equation}
 \end{itemize}
-
-\section{Statistische Modellierung}
-
+\item Probabilistic versions of non-probabilistic measures:
 \begin{itemize}
-  \item Sei $l$ der zeitliche Abstand, mit dem Realisationen verfügbar werden
-  \item $Y_{t + l} = Y_t + \lag Y_t$
-  \item Die Nowcast schätzen (fehlerbehaftet) $\lag Y_t$ mithilfe von Daten, die zum Zeitpunkt $\tau \in T_t^k$ verfügbar sind, und dem neuesten, bekannten Wert $Y_t$:
+  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects:
+  	\begin{align}
+  		\ydiff_t &= \ydifft + \varepsilon_t^y \\
+  		\xdiff_t &= \xdifft + \varepsilon_t^x
+	\end{align}
+\item  Replace indicators in the above formulations by the corresponding probabilities:
+\begin{align}
+		\mprob_{\text{acc}} &= \frac{\sum_{t=2}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-1}  \\
+   \mprob_{\text{py}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=2}^T \Prob{\ydifft_t > 0}} = \sum_{t=2}^T \frac{\Prob{\ydifft_t > 0, \xdifft_t > 0}}{\Prob{\ydifft_t > 0}} = \sum_{t=2}^T \Prob{\xdifft_t > 0 | \ydifft_t > 0} \\
+    \mprob_{\text{px}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=2}^T \Prob{\xdifft_t > 0}} = \sum_{t=2}^T \Prob{\ydifft_t > 0 | \xdifft_t > 0} \\
+    \mprob_{\text{ny}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=2}^T \Prob{\ydifft_t < 0}} = \sum_{t=2}^T \Prob{\xdifft_t < 0 | \ydifft_t < 0} \\
+    \mprob_{\text{px}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=2}^T \Prob{\xdifft_t < 0}} = \sum_{t=2}^T \Prob{\ydifft_t < 0 | \xdifft_t < 0}
+\end{align}
+\item Simple error model:
+  \begin{align}
+	  \varepsilon_Y \sim N(0, \sigma_Y) \\
+	  \varepsilon_X \sim N(0, \sigma_X)
+  \end{align}
+  yields, e.g.,
   	\begin{equation}
-  		X_{t+l \lvert \tau}^k = Y_t + (\widehat{\lag Y_t})_{t \lvert \tau \in T_t^k}^k,
+  		\mprob_{\text{acc}} = \frac{\sum_{t=2}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{(\ydiff_t, \xdiff_t); (\sigma_y, \sigma_x)}( 0, 0) \big) }{T-1}, 
 	\end{equation}
-	\todo{Stimmt das so?}
-	\todo{Ist festes $l$ zu starke Annahme?}
-	und damit 
-		\begin{equation}
-  			\lag Y_t = (\widehat{\lag Y_t})_{t \lvert \tau \in T_t^k}^k + \varepsilon.
-		\end{equation}
-
+	where $\Phi$ denotes a (possibly multivariate) normal distribution
+\end{itemize}
 \end{itemize}
 
-\section{Trending}
+\subsection{Confidence intervals for the measures}
 
-\begin{itemize}
-  \item \enquote{Schwaches} Trending für lag $l$:
+\section{Simulation studies}
+
+\subsection{Different trending abilities for same rmse and mae} \label{sec:simulation_rmse_mae}
+
+Let the quantity of interest 
 \begin{equation}
-  	P(\lag Y_t \widehat{(\lag Y_t)^k} > 0 \:\lvert\: \abs{\lag Y_t} > \epsilon) > P(\lag Y_t \widehat{(\lag Y_t)^k} < 0 \:\lvert\: \abs{\lag Y_t} > \epsilon)
+  y_t = a \sin(2 \pi t / T) + \varepsilon_t^y, \ t = 1, \dots, T
 \end{equation}
+with $\varepsilon_t^y \stackrel{\text{iid}}{\sim} N(0, \sigma_y)$ and 
+\begin{equation}
+  \ydiff_t = y_t - y_{t-1}.
+\end{equation}
+The nowcasts are characterised by 
+\begin{align}
+	\xdiff_t^1 &= e^x_t \ydiff_t \\
+	\xdiff_t^2 &= \begin{cases}
+		e^x_t \ydiff_t &, \text{if}\ q^2_t = 0 \lor e^x_t > 0\\
+		(| e^x_t | + 2) \ydiff_t &, \text{if}\ q^2_t = 1 \land e^x_t < 0
+	\end{cases} \\
+	\xdiff_t^3 &= \begin{cases}
+		e^x_t \ydiff_t &, \text{if}\ (q^{3, \text{down}}_t = 0 \lor e^x_t > 0) \land \ydiff_t \geq 0\\
+		(| e^x_t | + 2) \ydiff_t &, \text{if}\ (q^{3, \text{down}}_t = 1 \land e^x_t < 0) \land \ydiff_t \geq 0 \\
+		e^x_t \ydiff_t &, \text{if}\ (q^{3, \text{up}}_t = 0 \lor e^x_t > 0) \land \ydiff_t < 0\\
+		(| e^x_t | + 2) \ydiff_t &, \text{if}\ (q^{3, \text{up}}_t = 1 \land e^x_t < 0) \land \ydiff_t < 0
+	\end{cases}
+\end{align}
+where $q^2_t \sim \text{Ber}(p_2)$ and $p_2$ denotes the probability that the direction of nowcast 2 is corrected and for nowcast 3 with $q^{3, \text{up}}_t \sim \text{Ber}(p_3^{\text{up}})$, $q^{3, \text{down}}_t \sim \text{Ber}(p_3^{\text{down}})$ the probabilities of direction correction depend on the direction of change.
 
-\end{itemize}
+\begin{figure}
+  \centering
+  \includegraphics{plots/simulation_same_rmse_mae/time_series.pdf}
+  \caption{Realisation of Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.8, 0.4)$}
+  \label{fig:simulation_rmse_mae_ts}
+\end{figure}
 
-%\printbibliography
+\begin{figure}
+  \centering
+  \includegraphics{plots/simulation_same_rmse_mae/4q_plot.pdf}
+  \caption{4Q-Plot for Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.8, 0.4)$.}
+  \label{fig:simulation_rmse_mae_4q}
+\end{figure}
+
+\hl{Add results}
+
+\section{Application on disease and economic data}
+
+\section{Conclusion}
 

### 2023-12-05 paper.tex
Update draft, correct prob. measures

@@
 	\item Trending is an important feature of nowcasts to be identify changes in the situation before the actual quantity can be measured
+	\item What scores are currently used?
 	\item Examples: Covid, BIP, Medicine, ... 
@@
   \item Translate into mathematical notation
-  \item Illustrative Example: Two nowcasts of quantity of interest, where one jumps up and down and one points in right direction with the same rmse
+  \item Illustrative Example: Two nowcasts of quantity of interest, where one jumps up and down and one points in right direction with the same rmse; small subset of Simulation~\ref{sec:simulation_rmse_mae}; Plot time series \& time series of differences
+\end{itemize}
+
+\subsection{Graphical methods}
+
+\begin{itemize}
+  \item 4Q-Plot
+  \item Andere Medizin-Plots mit Verweis auf Grothe-Paper
 \end{itemize}
@@
 		\mprob_{\text{acc}} &= \frac{\sum_{t=2}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-1}  \\
-   \mprob_{\text{py}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=2}^T \Prob{\ydifft_t > 0}} = \sum_{t=2}^T \frac{\Prob{\ydifft_t > 0, \xdifft_t > 0}}{\Prob{\ydifft_t > 0}} = \sum_{t=2}^T \Prob{\xdifft_t > 0 | \ydifft_t > 0} \\
-    \mprob_{\text{px}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=2}^T \Prob{\xdifft_t > 0}} = \sum_{t=2}^T \Prob{\ydifft_t > 0 | \xdifft_t > 0} \\
-    \mprob_{\text{ny}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=2}^T \Prob{\ydifft_t < 0}} = \sum_{t=2}^T \Prob{\xdifft_t < 0 | \ydifft_t < 0} \\
-    \mprob_{\text{px}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=2}^T \Prob{\xdifft_t < 0}} = \sum_{t=2}^T \Prob{\ydifft_t < 0 | \xdifft_t < 0}
+   \mprob_{\text{py}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=2}^T \Prob{\ydifft_t > 0}} \\
+    \mprob_{\text{px}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=2}^T \Prob{\xdifft_t > 0}} \\
+    \mprob_{\text{ny}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=2}^T \Prob{\ydifft_t < 0}} \\
+    \mprob_{\text{px}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=2}^T \Prob{\xdifft_t < 0}} 
 \end{align}
@@
 
-\subsection{Confidence intervals for the measures}
+\subsection{Confidence intervals and missing data}
+
+\begin{itemize}
+  \item Bootstrapping: Describe approach (Simple bootstrap or multiplier?)
+  \item Missing data
+\end{itemize}
 
@@
 
+\subsection{Effect of increasing sample size?}
+
+\subsection{Simulation: Bootstrap}
+
+Question: How well does bootstrap approach perform?
+
+\begin{itemize}
+  \item 
+\end{itemize}
+
 \section{Application on disease and economic data}
+
+\subsection{Nowcasts for the COVID infections in Germany}
+
+\begin{itemize}
+  \item Describe QoI and Nowcasts; which nowcasts do we consider? Which data is available when?
+  \item Plot data
+\end{itemize}
+
+\subsection{Nowcasting of GDP}
 

### 2023-12-05 paper.tex
Update on Overleaf.

@@
 \begin{abstract}
-Existing and new measures for the capability of capturing the trend of nowcasts are presented, evaluated and compared on synthetic and real-world data.
+Existing and new measures for the capability of capturing the trend of nowcasts are presented, evaluated, and compared on synthetic and real-world data.
 \end{abstract}
@@
 \section{Introduction}
+
+Nicht nur Nowcast, sondern auch Measurement und Forecasts; in Introduction Unterschiede zwischen den Feldern; mehr Literatur; Einleitung ausbauen
 
@@
   \item Let $x_t^k$ be the $k$-th nowcast for time $t = 1, \ldots, T$ (possibly after some aggregation if several raw nowcasts apply).
-  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change occurred between time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - y_{t-1}$ the change nowcaster $k$ issued between the time steps. 
+  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - y_{t-1}$ the change nowcaster $k$ issued between the time steps. 
   \item Then, $\xdiff_t^k$ can be seen as point forecast of $\ydiff_t$. 
@@
 
-Non-probabilistic measures:
+Non-probabilistic measures: Summation anpassen; sodass rolling windows einfacher; untere Summationsgrenze $T - l$ ($l$ rolling window Größe); statt tilde besser delta; Nochmal anschauen, was unbeobachtbar und was beobachtbar ist; "verrauschtes" echtes kann auch daher kommen, dass ich nicht unbedingt $x - y_{t-1}$ rechnen kann
 
@@
 \begin{itemize}
-  \item Quadratic/Brier Score for $k$ possible outcomes: \begin{equation}
+  \item Quadratic/Brier Score for $k$ possible outcomes \hl{Erst mal rauslassen; wäre anderes setting: probabilistischer Nowcast}: \begin{equation}
   S((p_1, \dots, p_k), i) = 2 p_i - \sum_{l=1}^k p_i^2 - 1;
@@
 
-\subsection{Effect of increasing sample size?}
-
 \subsection{Simulation: Bootstrap}
@@
 \begin{itemize}
-  \item 
+  \item Auf jeden Fall
 \end{itemize}
+
+\subsection{Effect of increasing sample size?}
 

### 2023-12-12 paper.tex
Update on Overleaf.

@@
 \newcommand{\ind}[1]{\mathbbm{1}\{#1\}}
-\newcommand{\ydiff}{\tilde{y}}
-\newcommand{\ydifft}{\tilde{y}^*}
-\newcommand{\xdiff}{\tilde{x}}
-\newcommand{\xdifft}{\tilde{x}^*}
-\newcommand{\Ydiff}{\tilde{Y}}
-\newcommand{\Xdiff}{\tilde{X}}
+\newcommand{\ydiff}{\overset{\triangle}{y}}
+\newcommand{\ydifft}{\overset{\triangle}{y}^*}
+\newcommand{\xdiff}{\overset{\triangle}{x}}
+\newcommand{\xdifft}{\overset{\triangle}{y}^*}
+\newcommand{\Ydiff}{\overset{\triangle}{Y}}
+\newcommand{\Xdiff}{\overset{\triangle}{X}}
 \newcommand{\Prob}[1]{P(#1)}
@@
 
+\subsection*{Offene Punkte}
+
+\begin{itemize}
+    \item was ist mit dem Trendinghorizont, also welche Zeitinkremente sind zu nehmen.
+Da würde sich irgendwie ein Plot Maß gegen Inkrement anbieten.
+Dann würden sich für unterschiedliche Inkremente unterschiedliche Trendingcapabilities ergeben (gabe es nicht so einen schonmal bei irgendeinem Treffen?)
+    \item Wahl des Rauschens; haben wir da irgendeine Methode aus den Daten das Sigma zu bekommen?
+    \item Nicht nur Nowcast, sondern auch Measurement und Forecasts; in Introduction Unterschiede zwischen den Feldern; mehr Literatur; Einleitung ausbauen
+\end{itemize}
+
 \section{Introduction}
-
-Nicht nur Nowcast, sondern auch Measurement und Forecasts; in Introduction Unterschiede zwischen den Feldern; mehr Literatur; Einleitung ausbauen
 
@@
 \begin{itemize}
-  \item Scoring rules for probabilities
+  % \item Scoring rules for probabilities
 
-\begin{itemize}
-  \item Quadratic/Brier Score for $k$ possible outcomes \hl{Erst mal rauslassen; wäre anderes setting: probabilistischer Nowcast}: \begin{equation}
-  S((p_1, \dots, p_k), i) = 2 p_i - \sum_{l=1}^k p_i^2 - 1;
-\end{equation}
-for dichotomous outcomes ($p_1 = \hat{P}(\ydiff_t > 0)$: probability for $\ydiff_t > 0$):
-\begin{equation}
-  S(p_1, \ydiff_t) = p_1 \ind{\ydiff_t > 0} + (1 - p_1) (1 - \ind{\ydiff_t > 0}) + p_1^2 + (1-p_1)^2 - 1
-\end{equation}
-\end{itemize}
+% \begin{itemize}
+%   \item Quadratic/Brier Score for $k$ possible outcomes \hl{Erst mal rauslassen; wäre anderes setting: probabilistischer Nowcast}: \begin{equation}
+%   S((p_1, \dots, p_k), i) = 2 p_i - \sum_{l=1}^k p_i^2 - 1;
+% \end{equation}
+% for dichotomous outcomes ($p_1 = \hat{P}(\ydiff_t > 0)$: probability for $\ydiff_t > 0$):
+% \begin{equation}
+%   S(p_1, \ydiff_t) = p_1 \ind{\ydiff_t > 0} + (1 - p_1) (1 - \ind{\ydiff_t > 0}) + p_1^2 + (1-p_1)^2 - 1
+% \end{equation}
+% \end{itemize}
 \item Probabilistic versions of non-probabilistic measures:

### 2023-12-12 paper.tex
Update plots

@@
 \newcommand{\mprob}{\tilde{m}}
+\newcommand{\Ber}{\text{Ber}}
+\newcommand{\SBer}{\text{SBer}}
 
@@
 
+\hl{How to determine $\sigma$?}
+
 \subsection{Confidence intervals and missing data}
@@
 \begin{itemize}
-  \item Bootstrapping: Describe approach (Simple bootstrap or multiplier?)
+  \item Bootstrapping: Compare percentile and BCa approach?
+  \item Bootstrapping: What is the \enquote{true} value?
   \item Missing data
@@
   \centering
-  \includegraphics{plots/simulation_same_rmse_mae/4q_plot.pdf}
+  \begin{subfigure}{.32\textwidth}
+  	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_1}
+  \end{subfigure}
+  \begin{subfigure}{.32\textwidth}
+  	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_2}
+  \end{subfigure}
+    \begin{subfigure}{.32\textwidth}
+  	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_3}
+  \end{subfigure}
   \caption{4Q-Plot for Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.8, 0.4)$.}
@@
 
-Question: How well does bootstrap approach perform?
+Simulation setup:
 
 \begin{itemize}
-  \item Auf jeden Fall
+  \item $T = 10,000$
+  \item $\ydiff_t \sim N(0, 1)$ 
+  \item $\xdiff_t^{1, k} = b_t  \ydiff_t, b_t \sim \SBer(k)$ (symmetric Bernoulli)
+  \item \hl{(or both + random (non-systematic) noise?)}
+\end{itemize}
+
+Analysis:
+\begin{itemize}
+\item Do for $l = 1, \dots, 1,00$:
+\begin{itemize}
+  \item Sample $\ydiff, \xdiff$
+  \item Compute bootstrap confidence intervals $(b_{low}, b_{up})_l$ using $B = 10,000$
+\end{itemize}
+  \item Visualize all intervals with real value (confidence interval should cover real value in $\alpha$ cases)
 \end{itemize}

### 2023-12-12 paper.tex
Update on Overleaf.

@@
 %%% --- Mathematics
-\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools}
+\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools, mathabx}
 \usepackage{bm} % bold mathematics
@@
 %%% --- Commenting, Todonotes
-\usepackage{comment, todonotes, soul}
+\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
+\usepackage{comment, soul}
 \newcommand{\ar}{$\rightarrow$}
@@
 %\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}
+
+\usepackage{xargs} % Use optional arguements in commands
+\newcommandx{\unsure}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
+\newcommandx{\missing}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+\newcommandx{\change}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
+\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
+\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
+\usepackage{marginnote}
+\let\marginpar\marginnote
 
@@
 \newcommand{\ind}[1]{\mathbbm{1}\{#1\}}
-\newcommand{\ydiff}{\overset{\triangle}{y}}
-\newcommand{\ydifft}{\overset{\triangle}{y}^*}
-\newcommand{\xdiff}{\overset{\triangle}{x}}
-\newcommand{\xdifft}{\overset{\triangle}{y}^*}
-\newcommand{\Ydiff}{\overset{\triangle}{Y}}
-\newcommand{\Xdiff}{\overset{\triangle}{X}}
+\newcommand{\ydiff}{\overset{\smalltriangleup}{y}}
+\newcommand{\ydifft}{\overset{\smalltriangleup}{y}^*}
+\newcommand{\xdiff}{\overset{\smalltriangleup}{x}}
+\newcommand{\xdifft}{\overset{\smalltriangleup}{x}^*}
+\newcommand{\Ydiff}{\overset{\smalltriangleup}{Y}}
+\newcommand{\Xdiff}{\overset{\smalltriangleup}{X}}
 \newcommand{\Prob}[1]{P(#1)}
@@
     \item Wahl des Rauschens; haben wir da irgendeine Methode aus den Daten das Sigma zu bekommen?
+    \item Was ist mit anderen Verteilungen?
     \item Nicht nur Nowcast, sondern auch Measurement und Forecasts; in Introduction Unterschiede zwischen den Feldern; mehr Literatur; Einleitung ausbauen
+\end{itemize}
+
+\listoftodos
+
+Für nächstes Treffen:
+\begin{itemize}
+    \item Wie wird $\sigma$ geschätzt? Literaturrecherche, wo sonst noch \enquote{beliebig} festgelegt wird?
 \end{itemize}
@@
 
+\todo[inline]{$\xdiff_t^k = x_t - y_{t-1}$ oder $x_{t-1}$? Evtl. mit Bezug zu Verzögerung des Bekanntwerdens von $y_{t-1}$.}
+
 \subsection{What is the trend of a nowcast?}
@@
 \subsection{Graphical methods}
+
+Bolin
 
@@
 
-Non-probabilistic measures: Summation anpassen; sodass rolling windows einfacher; untere Summationsgrenze $T - l$ ($l$ rolling window Größe); statt tilde besser delta; Nochmal anschauen, was unbeobachtbar und was beobachtbar ist; "verrauschtes" echtes kann auch daher kommen, dass ich nicht unbedingt $x - y_{t-1}$ rechnen kann
+Non-probabilistic measures: Summation anpassen; sodass rolling windows einfacher; untere Summationsgrenze $T - l$ ($l$ rolling window Größe); Nochmal anschauen, was unbeobachtbar und was beobachtbar ist; "verrauschtes" echtes kann auch daher kommen, dass ich nicht unbedingt $x - y_{t-1}$ rechnen kann
 
@@
 
-\hl{How to determine $\sigma$?}
+\todo[inline]{Annahmen an $\varepsilon$: white noise; wie kommt man auf den Wert?
+Ideen: Expertenwissen, Anteil an Gesamtvarianz der Zeitreihen (z.B.
+10 \%), \textbf{Literaturrecherche zu wie das sonst so gemacht werden könnte}}
 
 \subsection{Confidence intervals and missing data}
+
+Bolin
 
@@
 where $q^2_t \sim \text{Ber}(p_2)$ and $p_2$ denotes the probability that the direction of nowcast 2 is corrected and for nowcast 3 with $q^{3, \text{up}}_t \sim \text{Ber}(p_3^{\text{up}})$, $q^{3, \text{down}}_t \sim \text{Ber}(p_3^{\text{down}})$ the probabilities of direction correction depend on the direction of change.
+$e^x_t \sim N(0, 1)$
 
@@
 
-\hl{Add results}
+\todo[inline]{Add results}
 
@@
   \item $\xdiff_t^{1, k} = b_t  \ydiff_t, b_t \sim \SBer(k)$ (symmetric Bernoulli)
-  \item \hl{(or both + random (non-systematic) noise?)}
 \end{itemize}
+\unsure[inline]{(or both + random (non-systematic) noise?)}
 
@@
 \begin{itemize}
-\item Do for $l = 1, \dots, 1,00$:
+\item Do for $l = 1, \dots, 100$:
 \begin{itemize}

### 2023-12-15 paper.tex
Update on Overleaf.

@@
 \usepackage[style=authoryear,backend=bibtex]{biblatex}
-%\addbibresource{library.bib}
+\addbibresource{library.bib}
 \usepackage{csquotes}
@@
 %%% --- Commenting, Todonotes
-\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
+\usepackage[colorinlistoftodos,prependcaption]{todonotes}
 \usepackage{comment, soul}
@@
 \newcommand{\ind}[1]{\mathbbm{1}\{#1\}}
-\newcommand{\ydiff}{\overset{\smalltriangleup}{y}}
-\newcommand{\ydifft}{\overset{\smalltriangleup}{y}^*}
-\newcommand{\xdiff}{\overset{\smalltriangleup}{x}}
-\newcommand{\xdifft}{\overset{\smalltriangleup}{x}^*}
-\newcommand{\Ydiff}{\overset{\smalltriangleup}{Y}}
-\newcommand{\Xdiff}{\overset{\smalltriangleup}{X}}
+\newcommand{\ydiff}{D y}
+\newcommand{\ydifft}{Dy^\star}
+\newcommand{\xdiff}{Dx}
+\newcommand{\xdifft}{Dx^\star}
+\newcommand{\Ydiff}{DY}
+\newcommand{\Xdiff}{DX}
 \newcommand{\Prob}[1]{P(#1)}
@@
 \begin{itemize}
-    \item Wie wird $\sigma$ geschätzt? Literaturrecherche, wo sonst noch \enquote{beliebig} festgelegt wird?
+    \item Formale Fragestellung
+    \item Wie wird $\sigma$ geschätzt? Filter vs. Empirie
+    \item Kondidenzinterval
+
 \end{itemize}
@@
   \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - y_{t-1}$ the change nowcaster $k$ issued between the time steps.
-  \item Then, $\xdiff_t^k$ can be seen as point forecast of $\ydiff_t$. 
+If $y_{t-1}$ is not available at time $t$, it can be reasonable to use an update nowcasts for time $t-1$.
+  \item Then, $\xdiff_t^k$ can be seen as point forecast of nowcaster $k$ for $\ydiff_t$. 
 \end{itemize}
-
-\todo[inline]{$\xdiff_t^k = x_t - y_{t-1}$ oder $x_{t-1}$? Evtl. mit Bezug zu Verzögerung des Bekanntwerdens von $y_{t-1}$.}
 
@@
 \subsection{Graphical methods}
-
-Bolin
 
@@
 
-Non-probabilistic measures: Summation anpassen; sodass rolling windows einfacher; untere Summationsgrenze $T - l$ ($l$ rolling window Größe); Nochmal anschauen, was unbeobachtbar und was beobachtbar ist; "verrauschtes" echtes kann auch daher kommen, dass ich nicht unbedingt $x - y_{t-1}$ rechnen kann
-
 \begin{itemize}
   \item Accuracy: \begin{equation}
-  	m_{\text{acc}} = \frac{\sum_{t=2}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=2}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-1}
+  	m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-1}
 \end{equation}
@@
 	\begin{equation}
-  		m_{\text{py}} = \frac{\sum_{t=2}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=2}^T \ind{\ydiff_t > 0}}
+  		m_{\text{py}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\ydiff_t > 0}}
 	\end{equation}
 	\item False alarm rate \hl{Name does not fit} \begin{equation}
-  m_{\text{far}} = \frac{\sum_{t=2}^T \ind{\ydiff_t < 0, \xdiff_t > 0}}{\sum_{t=2}^T \ind{\xdiff_t > 0}}
+  m_{\text{far}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}
 \end{equation}
@@
 \begin{equation}
-  m_{\text{px}} = \frac{\sum_{t=2}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=2}^T \ind{\xdiff_t > 0}}  
+  m_{\text{px}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}  
 \end{equation}
@@
 	\begin{align}
-		m_{\text{ny}} &=  \frac{\sum_{t=2}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=2}^T \ind{\ydiff_t < 0}} \\
-		m_{\text{nx}} &=  \frac{\sum_{t=2}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=2}^T \ind{\xdiff_t < 0}}
+		m_{\text{ny}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} \\
+		m_{\text{nx}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\xdiff_t < 0}}
 	\end{align}
+ \item Bezug auf Consusionsmatrix 
 \end{itemize}
@@
 % \end{itemize}
-\item Probabilistic versions of non-probabilistic measures:
+\item Probabilistic versions of non-probabilistic measures; small sample characteristic of $m_{acc}$ can be driven by random effects near the axis; therefore probabilistic \enquote{correction}:
 \begin{itemize}
-  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects:
-  	\begin{align}
+  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$:
+  	\begin{align}\label{additive error decomposition}
   		\ydiff_t &= \ydifft + \varepsilon_t^y \\
@@
 	\end{align}
+ \item \hl{Wouldn't it be more natural to model the original time series with a noise term?} (Qualitatively, it would yield the same results)
+ \begin{align}
+     y_t &= y_t^\star + \epsilon_t^y \\
+     x_t &= x_t^\star + \epsilon_t^x \\
+     \ydiff_t &= y_t^\star + \epsilon_t^y - (y_{t-1}^\star + \epsilon_{t-1}^y) = y_t^\star - y_{t-1}^\star + \underbrace{\epsilon_t^y - \epsilon_{t-1}^y}_{\sim N(0, \sqrt{2} \sigma_y)} \\
+     \xdiff_t &= x_t^\star + \epsilon_t^x - (x_{t-1}^\star + \epsilon_{t-1}^x) = x_t^\star - x_{t-1}^\star + \underbrace{\epsilon_t^x - \epsilon_{t-1}^x}_{\sim N(0, \sqrt{2} \sigma_x)}
+ \end{align}
+ Advantage of the approach: For different lags, no new standard deviation has to be estimated.\\
+ Disadvantage: Theoretically, there is serial correlation of the errors.
 \item  Replace indicators in the above formulations by the corresponding probabilities:
 \begin{align}
-		\mprob_{\text{acc}} &= \frac{\sum_{t=2}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-1}  \\
-   \mprob_{\text{py}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=2}^T \Prob{\ydifft_t > 0}} \\
-    \mprob_{\text{px}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=2}^T \Prob{\xdifft_t > 0}} \\
-    \mprob_{\text{ny}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=2}^T \Prob{\ydifft_t < 0}} \\
-    \mprob_{\text{px}} &= \frac{\sum_{t=2}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=2}^T \Prob{\xdifft_t < 0}} 
+		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-1}  \\
+   \mprob_{\text{py}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t > 0}} \\
+    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t > 0}} \\
+    \mprob_{\text{ny}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t < 0}} \\
+    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t < 0}} 
 \end{align} 
@@
   	\begin{equation}
-  		\mprob_{\text{acc}} = \frac{\sum_{t=2}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{(\ydiff_t, \xdiff_t); (\sigma_y, \sigma_x)}( 0, 0) \big) }{T-1}, 
+  		\mprob_{\text{acc}} = \frac{1}{T-1} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
 	\end{equation}
@@
 \end{itemize}
+\item For $\sigma_x \sigma_y \rightarrow 0$, the probabilistic measures approach the count-based ones
+\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic measures approach 0.5 as all probabilities are approximately 0.5
+\item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
 \end{itemize}
 
-\todo[inline]{Annahmen an $\varepsilon$: white noise; wie kommt man auf den Wert?
-Ideen: Expertenwissen, Anteil an Gesamtvarianz der Zeitreihen (z.B.
-10 \%), \textbf{Literaturrecherche zu wie das sonst so gemacht werden könnte}}
+\subsection{Estimating the standard deviation of white noise}
+
+First, sometimes variance can be determined heuristically directly from domain knowledge. 
+However, to determine the standard deviation of white noise using only the time series data, one can basically follow two approaches, time domain approach or frequency domain approach.
+For the first approach, a model assumption (e.g.
+ARIMA models) is specified and the residual analysis technique is applied [\cite{chitturi1974distribution}].
+In the frequency domain approach, the time series are first decomposed using spectral methods such as (discrete) Fourier transformation [\cite{sauer1992noise}] or wavelet transformation [\cite{heil1989continuous,donoho1994ideal,jansen2012noise}].
+After reconstructing the time series, the residuals can be formed in order to determine the standard deviation of white noise.
+For the use of wavelet, it is particularly convenient that the standard deviation of the white noise can be estimated by the standard deviation of the detail coefficients of the last level, e.g., if the wavelet family db1 is used.
+
+For the noisy signal in the form of \ref{eq: simulation},  there are the following results (cf. pic \ref{fig:estimation std})
+
+\begin{figure}
+  \centering
+\includegraphics[scale=0.5]{plots/std_estimate.png}
+  \caption{Estimation of std with different methods.}
+  \label{fig:estimation std}
+\end{figure}
 
 \subsection{Confidence intervals and missing data}
-
-Bolin
 
@@
 Let the quantity of interest 
-\begin{equation}
+\begin{equation}\label{eq: simulation}
   y_t = a \sin(2 \pi t / T) + \varepsilon_t^y, \ t = 1, \dots, T
@@
 \end{equation}
+
+\todo[inline]{Anpassung der Simulationssettings an Gl. \ref{additive error decomposition} (?)}
 The nowcasts are characterised by 
 \begin{align}
-	\xdiff_t^1 &= e^x_t \ydiff_t \\
-	\xdiff_t^2 &= \begin{cases}
-		e^x_t \ydiff_t &, \text{if}\ q^2_t = 0 \lor e^x_t > 0\\
-		(| e^x_t | + 2) \ydiff_t &, \text{if}\ q^2_t = 1 \land e^x_t < 0
+	(\xdifft_t)^1 &= e^x_t \ydifft_t \\
+	(\xdifft_t)^2 &= \begin{cases}
+		e^x_t \ydifft_t &, \text{if}\ q^2_t = 0 \lor e^x_t \geq 0\\
+		(| e^x_t | + 2) \ydifft_t &, \text{if}\ q^2_t = 1 \land e^x_t < 0
 	\end{cases} \\
-	\xdiff_t^3 &= \begin{cases}
-		e^x_t \ydiff_t &, \text{if}\ (q^{3, \text{down}}_t = 0 \lor e^x_t > 0) \land \ydiff_t \geq 0\\
-		(| e^x_t | + 2) \ydiff_t &, \text{if}\ (q^{3, \text{down}}_t = 1 \land e^x_t < 0) \land \ydiff_t \geq 0 \\
-		e^x_t \ydiff_t &, \text{if}\ (q^{3, \text{up}}_t = 0 \lor e^x_t > 0) \land \ydiff_t < 0\\
-		(| e^x_t | + 2) \ydiff_t &, \text{if}\ (q^{3, \text{up}}_t = 1 \land e^x_t < 0) \land \ydiff_t < 0
+	(\xdifft_t)^3 &= \begin{cases}
+		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 0 \lor e^x_t > 0) \land \ydifft_t \geq 0\\
+		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 1 \land e^x_t < 0) \land \ydifft_t \geq 0 \\
+		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 0 \lor e^x_t > 0) \land \ydifft_t < 0\\
+		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 1 \land e^x_t < 0) \land \ydifft_t < 0
 	\end{cases}
 \end{align}
-where $q^2_t \sim \text{Ber}(p_2)$ and $p_2$ denotes the probability that the direction of nowcast 2 is corrected and for nowcast 3 with $q^{3, \text{up}}_t \sim \text{Ber}(p_3^{\text{up}})$, $q^{3, \text{down}}_t \sim \text{Ber}(p_3^{\text{down}})$ the probabilities of direction correction depend on the direction of change.
-$e^x_t \sim N(0, 1)$
+where $q^2_t \sim \text{Ber}(2p_2 - 1)$ and $p_2 > 0.5$ denotes the probability that the direction of nowcast 2 is correct.
+For nowcast 3 with $q^{3, \text{up}}_t \sim \text{Ber}(2p_3^{\text{up}}-1)$, $q^{3, \text{down}}_t \sim \text{Ber}(2p_3^{\text{down}}-1)$ the probabilities of direction correction depend on the direction of change.
+Additionally, let $e^x_t \sim N(0, 1)$.
 
@@
 
-\todo[inline]{Add results}
-
+\todo[inline]{Weiteres Beispiel für viele richtigen Punkte sehr nah an den Entscheidungsgrenzen.}
 \subsection{Simulation: Bootstrap}
@@
 \section{Conclusion}
+\medskip
+\printbibliography
 

### 2023-12-15 paper.tex
Update on Overleaf.

@@
   \includegraphics{plots/simulation_same_rmse_mae/time_series.pdf}
-  \caption{Realisation of Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.8, 0.4)$}
+  \caption{Realisation of Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$}
   \label{fig:simulation_rmse_mae_ts}
@@
   \end{subfigure}
-  \caption{4Q-Plot for Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.8, 0.4)$.}
+  \caption{4Q-Plot for Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$.}
   \label{fig:simulation_rmse_mae_4q}

### 2023-12-16 paper.tex
Update on Overleaf.

@@
 
-Trend: Issuing the right \enquote{direction} of change between two time steps
+Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$; $P(\ydiff_t \cdot \xdiff_t > 0)$
 
@@
 \item For $\sigma_x \sigma_y \rightarrow 0$, the probabilistic measures approach the count-based ones
-\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic measures approach 0.5 as all probabilities are approximately 0.5
+\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic accuracy approaches   0.5 as all probabilities are approximately 0.5
 \item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
@@
 
+% \begin{figure}
+%   \centering
+% \includegraphics[scale=0.5]{plots/std_estimate.png}
+%   \caption{Estimation of std with different methods.}
+%   \label{fig:estimation std}
+% \end{figure}
+
 \begin{figure}
   \centering
-\includegraphics[scale=0.5]{plots/std_estimate.png}
+    \begin{subfigure}{0.24\textwidth}
+        \includegraphics{plots/std estimation/std_estimation_boxplot_25.pdf}
+    \end{subfigure}
+    \begin{subfigure}{0.24\textwidth}
+        \includegraphics{plots/std estimation/std_estimation_boxplot_50.pdf}
+    \end{subfigure}
+    \begin{subfigure}{0.24\textwidth}
+        \includegraphics{plots/std estimation/std_estimation_boxplot_100.pdf}
+    \end{subfigure}
+    \begin{subfigure}{0.24\textwidth}
+        \includegraphics{plots/std estimation/std_estimation_boxplot_1000.pdf}
+    \end{subfigure}
   \caption{Estimation of std with different methods.}
@@
 
+\appendix
+\section{Illustrative examples for probabilistic measures}
+
 \end{document}

### 2023-12-16 paper.tex
Update on Overleaf.

@@
 
-\listoftodos
-
 Für nächstes Treffen:
@@
     \item Kondidenzinterval     
+\end{itemize}
 
-\end{itemize}
+\listoftodos
 
@@
   \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - y_{t-1}$ the change nowcaster $k$ issued between the time steps.
-If $y_{t-1}$ is not available at time $t$, it can be reasonable to use an update nowcasts for time $t-1$.
+If $y_{t-1}$ is not available at time $t$, it can be reasonable to use an updated nowcast for time $t-1$.
+(Move this point to a place after error model, as the error model can account for the error made at this point)
   \item Then, $\xdiff_t^k$ can be seen as point forecast of nowcaster $k$ for $\ydiff_t$. 
@@
 
-Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$; $P(\ydiff_t \cdot \xdiff_t > 0)$
+Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$
 
@@
 	\end{align}
- \item Bezug auf Consusionsmatrix 
+ \item Bezug auf Confusionsmatrix 
 \end{itemize}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics{plots/illustrative_examples/plots_4q_and_mtilde.pdf}
+    \caption{Various examplary courses of the measures $\tilde{m}$ for different configurations of $\ydiff$ and $\xdiff$}
+    \label{fig:example_mtilde}
+\end{figure}
+
 \end{document}

### 2023-12-18 paper.tex
Update on Overleaf.

@@
     \item Nicht nur Nowcast, sondern auch Measurement und Forecasts; in Introduction Unterschiede zwischen den Feldern; mehr Literatur; Einleitung ausbauen
+    \item Messen die Maße das richtige? Sollte man andere Maße nutzen? Braucht man Korrekturterm?
+    \item Indikatoren oder Wahrscheinlichkeiten?
+    \item Wahrscheinlichkeit von der Pike herleiten
 \end{itemize}
@@
 	\item Examples: Covid, BIP, Medicine, ... 
+ \item nowcasting schön aufschreiben, dass auch für Nicht-Nowcaster klar wird, warum Nowcasting speziell ist
 \end{itemize}
@@
 If $y_{t-1}$ is not available at time $t$, it can be reasonable to use an updated nowcast for time $t-1$.
-(Move this point to a place after error model, as the error model can account for the error made at this point)
+(Move this point to a place after error model, as the error model can account for the error made at this point) \todo{Kaligraphisches $x$, um vorzubereiten, dass es anderes sein könnte}
   \item Then, $\xdiff_t^k$ can be seen as point forecast of nowcaster $k$ for $\ydiff_t$. 
@@
 
-Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$
+Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$ (Problem: mit kleinen Fehlern; Hinweis Exclusion area)
 
@@
   \item Accuracy: \begin{equation}
-  	m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-1}
+  	m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-w}
 \end{equation}
@@
 \begin{itemize}
-  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$:
+  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$ (\hl{this, no additional complexity}):
   	\begin{align}\label{additive error decomposition}
@@
  Advantage of the approach: For different lags, no new standard deviation has to be estimated.\\
- Disadvantage: Theoretically, there is serial correlation of the errors.
+ Disadvantage: Theoretically, there is serial correlation of the errors. (Wie behandle ich die Autokorrelation; wie rausrechnen)
 \item  Replace indicators in the above formulations by the corresponding probabilities:
 \begin{align}
-		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-1}  \\
+		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-w}  \\
    \mprob_{\text{py}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t > 0}} \\
@@
   	\begin{equation}
-  		\mprob_{\text{acc}} = \frac{1}{T-1} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
+  		\mprob_{\text{acc}} = \frac{1}{T-w} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
 	\end{equation}
@@
 
-% \begin{figure}
+%\begin{figure}
 %  \centering
-% \includegraphics[scale=0.5]{plots/std_estimate.png}
+%    \begin{subfigure}{0.24\textwidth}
+%        \includegraphics{plots/std estimation/std_estimation_boxplot_25.pdf}
+%    \end{subfigure}
+%    \begin{subfigure}{0.24\textwidth}
+%        \includegraphics{plots/std estimation/std_estimation_boxplot_50.pdf}
+%    \end{subfigure}
+%    \begin{subfigure}{0.24\textwidth}
+%        \includegraphics{plots/std estimation/std_estimation_boxplot_100.pdf}
+%    \end{subfigure}
+%    \begin{subfigure}{0.24\textwidth}
+%        \includegraphics{plots/std estimation/std_estimation_boxplot_1000.pdf}
+%    \end{subfigure}
 %  \caption{Estimation of std with different methods.}
 %  \label{fig:estimation std}
-% \end{figure}
+%\end{figure}%
 
@@
   \centering
-    \begin{subfigure}{0.24\textwidth}
-        \includegraphics{plots/std estimation/std_estimation_boxplot_25.pdf}
+    \begin{subfigure}{0.48\textwidth}
+        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_25.pdf}
     \end{subfigure}
-    \begin{subfigure}{0.24\textwidth}
-        \includegraphics{plots/std estimation/std_estimation_boxplot_50.pdf}
+    \begin{subfigure}{0.48\textwidth}
+        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_50.pdf}
     \end{subfigure}
-    \begin{subfigure}{0.24\textwidth}
-        \includegraphics{plots/std estimation/std_estimation_boxplot_100.pdf}
+
+    \begin{subfigure}{0.48\textwidth}
+        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_100.pdf}
     \end{subfigure}
-    \begin{subfigure}{0.24\textwidth}
-        \includegraphics{plots/std estimation/std_estimation_boxplot_1000.pdf}
+    \begin{subfigure}{0.48\textwidth}
+        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_1000.pdf}
     \end{subfigure}
+
   \caption{Estimation of std with different methods.}

### 2023-12-19 paper.tex
Add new weighted estimator; remove whitespace in folder

@@
 \newcommand{\SBer}{\text{SBer}}
+\newcommand{\cond}{\:\lvert\:}
 
@@
 
+\hl{Idea: Weighted Indicator, accounting for the likeliness of a point's location being driven by noise}
+
+\begin{itemize}
+  \item Weights $\omega_t$ for $(\ydiff_t, \xdiff_t)$ could be determined by error model and corresponds to the maximal likeliness of the point belonging to the different quadrants ($Q_1, \dots, Q_4$); thus, point that have less explanatory power for assessing the trending ability becomes less weight:
+ 	\begin{equation}
+  		\omega_t = \max_{i \in [4]} {P\big( (\ydifft_t, \xdifft_t) \in Q_i \cond \ydiff_t, \xdiff_t \big)} 
+	\end{equation}
+	\item E.g., the accuracy becomes
+	\begin{equation}
+  m_{acc, \omega} = \frac{\sum_{t=T-w}^T \omega_t \ind{\ydiff_t \cdot \xdiff_t > 0} }{\sum_{t=T-w}^T \omega_t}
+\end{equation}
+	\item Thus, $\omega_t \in [0.25, 1]$. Maybe it would make sense to \enquote{ignore} values without explanatory power and use $\omega_t^\star = 4/3 \ \omega_t \in [0, 1]$ \hl{to discuss}
+	\item Figure~\ref{fig:comparison_estimators} shows the resulting estimates for the simulation study comparing the pure indicator, the probability-version and the weighted indicator estimation; the data is generated (in a very simpel model) by
+	\begin{align}
+		\ydifft_t &\sim N(0, 2) \\
+		\xdifft_t &= (2 b_t - 1) \cdot \ydiff_t , b_t \sim \Ber (k) \\
+		\ydiff_t &= \ydifft_t + \varepsilon_t^y, \varepsilon_t^y \sim N(0, 0.1) \\
+		\xdiff_t &= \xdifft_t + \varepsilon_t^x, \varepsilon_t^x \sim N(0, 0.1) \\
+	\end{align}
+\end{itemize}
+
+\begin{figure}
+  \centering
+	\includegraphics{plots/unbiasedness/boxplot_k_0.1.pdf}
+	\includegraphics{plots/unbiasedness/boxplot_k_0.3.pdf}
+	\includegraphics{plots/unbiasedness/boxplot_k_0.5.pdf}
+	\includegraphics{plots/unbiasedness/boxplot_k_0.7.pdf}
+	\includegraphics{plots/unbiasedness/boxplot_k_0.9.pdf}			
+  \caption{Simulations results comparing the three types of estimators for the accuracy in a trending setting}
+  \label{fig:comparison_estimators}
+\end{figure}
+
 Resulting measures:
@@
 	\end{align}
- \item \hl{Wouldn't it be more natural to model the original time series with a noise term?} (Qualitatively, it would yield the same results)
+ \item Wouldn't it be more natural to model the original time series with a noise term? (Qualitatively, it would yield the same results)
  \begin{align}
@@
     \begin{subfigure}{0.48\textwidth}
-        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_25.pdf}
+        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_25.pdf}
     \end{subfigure}
     \begin{subfigure}{0.48\textwidth}
-        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_50.pdf}
+        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_50.pdf}
     \end{subfigure}
@@
     \begin{subfigure}{0.48\textwidth}
-        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_100.pdf}
+        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_100.pdf}
     \end{subfigure}
     \begin{subfigure}{0.48\textwidth}
-        \includegraphics[width=\textwidth]{plots/std estimation/std_estimation_boxplot_1000.pdf}
+        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_1000.pdf}
     \end{subfigure}

### 2024-01-10 paper.tex
Update on Overleaf.

@@
 \begin{itemize}
-    \item was ist mit dem Trendinghorizont, also welche Zeitinkremente sind zu nehmen.
-Da würde sich irgendwie ein Plot Maß gegen Inkrement anbieten.
-Dann würden sich für unterschiedliche Inkremente unterschiedliche Trendingcapabilities ergeben (gabe es nicht so einen schonmal bei irgendeinem Treffen?)
-    \item Wahl des Rauschens; haben wir da irgendeine Methode aus den Daten das Sigma zu bekommen?
-    \item Was ist mit anderen Verteilungen?
-    \item Nicht nur Nowcast, sondern auch Measurement und Forecasts; in Introduction Unterschiede zwischen den Feldern; mehr Literatur; Einleitung ausbauen
-    \item Messen die Maße das richtige? Sollte man andere Maße nutzen? Braucht man Korrekturterm?
-    \item Indikatoren oder Wahrscheinlichkeiten?
-    \item Wahrscheinlichkeit von der Pike herleiten
-\end{itemize}
-
-Für nächstes Treffen:
-\begin{itemize}
-    \item Formale Fragestellung
-    \item Wie wird $\sigma$ geschätzt? Filter vs. Empirie
-    \item Kondidenzinterval     
+    \item Welche Form sollte das Maß haben? Mögliche Wahlen: Indikator, Probabilistisch, Weighted (1), Weighted (2), Indikator mit Korrektur, ... (Siehe Tabelle \ref{tab:comparison-measures})
+    \item Schätzung/Wahl des Rauschparameters $\sigma$; haben wir da irgendeine Methode aus den Daten das Sigma zu bekommen? Filter vs. Empirie 
+    \item Gehen wir auf andere Verteilungen für den Fehler ein?
+    \item Konfidenzintervalle
 \end{itemize}
@@
 	\item Examples: Covid, BIP, Medicine, ... 
- \item nowcasting schön aufschreiben, dass auch für Nicht-Nowcaster klar wird, warum Nowcasting speziell ist
+ \item Nice an comprehensive introduction to nowcasting, so that it becomes clear for non-nowcasters what is special about nowcasting
+ \item Measurements and Forecasting; What are the differences to nowcasting from a modeling point of view?
 \end{itemize}
@@
   \item Let $x_t^k$ be the $k$-th nowcast for time $t = 1, \ldots, T$ (possibly after some aggregation if several raw nowcasts apply).
-  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - y_{t-1}$ the change nowcaster $k$ issued between the time steps.
-If $y_{t-1}$ is not available at time $t$, it can be reasonable to use an updated nowcast for time $t-1$.
-(Move this point to a place after error model, as the error model can account for the error made at this point) \todo{Kaligraphisches $x$, um vorzubereiten, dass es anderes sein könnte}
-  \item Then, $\xdiff_t^k$ can be seen as point forecast of nowcaster $k$ for $\ydiff_t$. 
+  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - \mathcal{x}_{t-1}$ the change nowcaster $k$ issued between the time steps.
+Let us first assume that $y_{t-1}$ is available at time $t$ so that it can be used for the evaluation.
+  In Section \hl{error-model-section}, we expand on other reasonable choices for $\mathcal{x}_{t-1}$.
+  \item Then, $\xdiff_t^k$ can be seen as the point forecast of nowcaster $k$ for $\ydiff_t$. 
 \end{itemize}
@@
 
+\begin{table}[]
+    \centering
+    \begin{tabular}{l l p{4cm} p{4cm} p{4cm}}
+    \toprule
+    name & abbr. & accuracy & + & - \\ \midrule
+    indicator & i & $\frac{1}{T-1} \sum \ind{x_t y_t > 0}$ &  well-known, intuitive & what about points near (0,0) \\
+    Probabilistic & p & $\frac{1}{T-1} \sum P{(x_t y_t > 0)}$ & small variance, weighting due to probabilities & unintuitive, asymptotics unclear, driven towards 0.5, no standard estimation for $\sigma$  \\
+    weighted 1 (*) & w1 & {\small{$\sum w_i \ind{x_t y_t > 0} / \sum w_i$,\newline $w_i = \max_j\{P(y_i, x_i) \in Q_j \}$}} & intuitive, implicit exclusion area, natural extension & all points gain weight, estimation of $\sigma$.
+\\
+    weighted 2 (*) & w2 & {\small {$\sum w_i \ind{x_t y_t > 0} / \sum w_i$,\newline $w_i = \max_j\{P(y_i, x_i) \in Q_j \} - 0.25$}} & as w1, points near (0,0) do not gain weight & \\ \bottomrule
+    \end{tabular}
+    \caption{Comparison between different measures for accuracy. }
+    \label{tab:comparison-measures}
+\end{table}
+
+\unsure[inline]{Anpassung der gewichteten Formel}
+(*): The weighted formula can also be changed by using the signum function $\sign$ instead of $\ind$ to take into account the influence of negative trending.\\
+
+Natural evaluation method for 4Q-plot, as known from dichotomous event evaluation, evaluation of classification (confusion matrix)
+
 \begin{itemize}
@@
 \end{equation}
-	\item Capability of detecting a positive trend (probability of detection)
+	\item Capability of detecting a positive trend (probability of detection)/Sensitivity
 	\begin{equation}
@@
 	\end{equation}
-	\item False alarm rate \hl{Name does not fit} \begin{equation}
-  m_{\text{far}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}
-\end{equation}
-maybe use \enquote{positive} measure instead, i.e.,
+	\item positive predictive values
 \begin{equation}
@@
 \end{equation}
-	\item same measures for negative direction:
+	\item Specificity
 	\begin{align}
-		m_{\text{ny}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} \\
+		m_{\text{ny}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} 
+  \end{align}
+  \item Negative predictive value
+	\begin{align}
 		m_{\text{nx}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\xdiff_t < 0}}
 	\end{align}
- \item Bezug auf Confusionsmatrix 
 \end{itemize}
 
-\hl{Idea: Weighted Indicator, accounting for the likeliness of a point's location being driven by noise}
+Different setting here: We do not only have binary predictions and realization, but we do know the real values behind; intuitive evaluation of a 4Q-plot assigns a greater explanatory power to point far away from (0,0), as small deviations from (0,0) could be driven by noise.
 
+Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$ (\hl{this, no additional complexity}):
+  	\begin{align}\label{additive error decomposition}
+  		\ydiff_t &= \ydifft + \varepsilon_t^y \\
+  		\xdiff_t &= \xdifft + \varepsilon_t^x
+	\end{align}
+
+Extension 1: 
 \begin{itemize}
-  \item Weights $\omega_t$ for $(\ydiff_t, \xdiff_t)$ could be determined by error model and corresponds to the maximal likeliness of the point belonging to the different quadrants ($Q_1, \dots, Q_4$); thus, point that have less explanatory power for assessing the trending ability becomes less weight:
+  \item 
+Adapt the measures to be weighted by the explanatory power of the points:\\
+Weights $\omega_t$ for $(\ydiff_t, \xdiff_t)$ could be determined by error model and corresponds to the maximal likeliness of the point belonging to the different quadrants ($Q_1, \dots, Q_4$); thus, point that have less explanatory power for assessing the trending ability becomes less weight:
  	\begin{equation}
@@
 \end{equation}
-	\item Thus, $\omega_t \in [0.25, 1]$. Maybe it would make sense to \enquote{ignore} values without explanatory power and use $\omega_t^\star = 4/3 \ \omega_t \in [0, 1]$ \hl{to discuss}
+	\item Thus, $\omega_t \in [0.25, 1]$. Maybe it would make sense to \enquote{ignore} values without explanatory power and use $\omega_t^\star = 4/3 \ \omega_t \in [0, 1]$ \hl{w2}
 	\item Figure~\ref{fig:comparison_estimators} shows the resulting estimates for the simulation study comparing the pure indicator, the probability-version and the weighted indicator estimation; the data is generated (in a very simpel model) by
@@
 
-Resulting measures:
+Probabilistic extension:
 \begin{itemize}
-  % \item Scoring rules for probabilities
-
-% \begin{itemize}
-%   \item Quadratic/Brier Score for $k$ possible outcomes \hl{Erst mal rauslassen; wäre anderes setting: probabilistischer Nowcast}: \begin{equation}
-%   S((p_1, \dots, p_k), i) = 2 p_i - \sum_{l=1}^k p_i^2 - 1;
-% \end{equation}
-% for dichotomous outcomes ($p_1 = \hat{P}(\ydiff_t > 0)$: probability for $\ydiff_t > 0$):
-% \begin{equation}
-%   S(p_1, \ydiff_t) = p_1 \ind{\ydiff_t > 0} + (1 - p_1) (1 - \ind{\ydiff_t > 0}) + p_1^2 + (1-p_1)^2 - 1
-% \end{equation}
-% \end{itemize}
 \item Probabilistic versions of non-probabilistic measures; small sample characteristic of $m_{acc}$ can be driven by random effects near the axis; therefore probabilistic \enquote{correction}:
@@
 \item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
+\end{itemize}
+
+\unsure[inline]{Diskussion der Annahme der additiven Fehler-Dekomposition}
+
+Nach Annahme der additiven Fehler-Dekomposition ist es ohne weiteres nicht möglich, die probabilistische Formel, insbesondere den Ausdruck $P(Dy^*\cdot Dx^* > 0|Dx=Dx_t, Dy=Dy_t) $ auszurechnen, da $(Dx,Dy)$ und $(Dx^*,Dy^*)$ abhängig sind
+
+\subsection{Gewichtete Formel }
+\unsure[inline]{Gewichtete Formel mit einer anderen Modelannahme und storyline}
+
+Motivation: Die Punkte $(Dx_t, Dy_t)$, die nahe an den Entscheidungsgrenzen (Ursprung und Koordinatenachsen) liegen, sollen für die Berechnung des Trending-Maßes nur begrenzt aussagekräftig sein, da es unsicher ist, ob solche Punkte wirklich zu den zugehörigen Quadranten (1,3 oder 2,4) gehören.\\\\
+Idee: Sicherheit der Zugehörigkeit als Gewichtung in die Indikatorformel einführen.\\\\
+Annahme: Der wahre Wert eines beobachteten Wertepaares $(Dx_t,Dy_t)$ unterliegt einer Normalverteilung, die durch den Mittelpunkt $(Dx_t,Dy_t)$ und die Kovarianz  beschrieben 
+$\begin{pmatrix} \sigma_x & 0 \\
+0 & \sigma_y 
+\end{pmatrix}$ wird.\\\\
+Berechnung der Gewichte als Sicherheit der Zugehörigkeit: 
+\[ 
+w_t = 
+\begin{cases} 
+  P(Dx_t \cdot Dy_t > 0) = \left(1 - \Phi\left(\frac{-Dx_t}{\sigma_X}\right)\right)\left(1 - \Phi\left(\frac{-Dy_t}{\sigma_Y}\right)\right) + \Phi\left(\frac{-Dx_t}{\sigma_X}\right) \Phi\left(\frac{-Dy_t}{\sigma_Y}\right) & \text{wenn } Dx_t \cdot Dy_t > 0,  \\
+  P(Dx_t \cdot Dy_t < 0) = \left(1 - \Phi\left(\frac{-Dx_t}{\sigma_X}\right)\right) \Phi\left(\frac{-Dy_t}{\sigma_Y}\right) + \Phi\left(\frac{-Dx_t}{\sigma_X}\right) \left(1 - \Phi\left(\frac{-Dy_t}{\sigma_Y}\right)\right)  & \text{wenn } Dx_t \cdot Dy_t < 0,  \\
+  0 & \text{sonst} 
+\end{cases}
+\]\\\\
+Dies ermöglicht eine kanonische Erweiterung der Indikatorformel:
+
+\[m_{\text{acc}} = \frac{\sum_{t=T-w}^T \omega_t \cdot \text{sgn}(Dx_t\cdot Dy_t)}{T-w}\]\\\\
+Anderer Normierungsfaktor: z.B.
+$\sum_t w_t$ möglich, aber folgendes Problem: Wenn alle Punkte z.B.
+im ersten Quadranten liegen, dann verursacht der Normierungsfaktor $\sum_t w_t$ immer das Maß $1$, was in manchen Fällen nicht der Intuition entspricht, z.B.
+wenn die Punkte nahe an der Entscheidungsgrenze liegen (in diesem Fall würde man intuitiv eigentlich eine leicht positive Zahl erwarten).
+\\\\
+Bestimmung der Unsicherheit: $\sigma_x$ und $\sigma_y$ geben die Schätzung der Unsicherheit wieder, d.h.
+wie "weit" der wahre Punkt vom beobachteten Punkt im Vier-Quadranten-Plot entfernt sein kann.
+Es ist schwierig, die Standardabweichungen anhand der Daten zu bestimmen, da die beobachtbaren Daten diese Information nicht enthalten.
+Eine empirische Methode besteht z.
+B.
+darin, die empirische Varianz von $Dx$ und $Dy$ als Referenz zu nehmen und daraus Werte zu berechnen (z.
+B.
+10 $\%$ der Varianz von $Dx$ und $Dy$).\\\\
+Eine weitre Möglichkeit ist die Funktion $(\sigma_x,\sigma_y)\mapsto m_{\text{acc}}(\sigma_x,\sigma_y)$ zu betrachten und robuste Aussage (z.B. durch Optimierung) zu treffen.
+
+\begin{itemize}
+     \item Maximum der Funktion: Wie viel Trending ist unter Berücksichtigung der Unsicherheit maximal möglich.
+    \item Minimum der Funktion: Wie viel Trend ist unter Berücksichtigung der Unsicherheit minimal möglich.
+    \item Erwartungswert: Hier kann eine zusätzliche Annahme über die Verteilung der Unsicherheit getroffen werden.Oder man berechnet für einen hinreichend großen Bereich von $\sigma_x$ und $\sigma_y$ den Mittelwert der aus verschiedenen $\sigma_x$ und $\sigma_y$ resultierenden Maße.
 \end{itemize}
@@
 
-  \caption{Estimation of std with different methods.}
+  \caption{Estimation of std using different methods.}
   \label{fig:estimation std}
@@
 
+\subsection{Trending over various time lags}
+
+\begin{itemize}
+    \item Introduce plot over several time lags
+    \item Question: Re-estimate noise for every lag? 
+\end{itemize}
+
 \section{Simulation studies}
+
+\subsection{Bias-Variance tradeoff}
+
+Simulation:
+\begin{enumerate}
+    \item For $T = 50, 100, 500, 1000$, \verb|dist| in N(0,1), N(0,5), N(0,10), N(1,1), N(1,5), N(1,10), Chi-squared(3), Chi-squared(7), Chi-squared(9)
+    \begin{enumerate}
+        \item For $k \in \{0.05, \dots, 0.95\}$:
+        \begin{enumerate}
+            \item Sample $T$ samples from $\ydifft$ following \verb|dist|
+            \item For $t = 1, \dots, T$: Set $\xdifft_t = \ydifft_t$ with probability $k$, otherwise set $\ydifft_t = - \ydifft_t$
+            \item Set $\ydiff_t = \ydifft + N(0,0.1)$, $\xdiff_t = \xdifft + N(0,0.1)$
+            \item Compute all trending measures
+            \item Repeat 1,000 times
+        \end{enumerate}
+        \item Plot the trending estimates over $k$.
+    \end{enumerate}
+\end{enumerate}
 
@@
 
-\todo[inline]{Anpassung der Simulationssettings an Gl. \ref{additive error decomposition} (?)}
+\todo[inline]{Anpassung der Simulationssettings an Gl. \ref{additive error decomposition}; Beschreibung Verbessern}
 The nowcasts are characterised by 

### 2024-01-13 paper.tex
Update on Overleaf.

@@
 
+%%% --- Colors
+\usepackage[dvipsnames]{xcolor}
+
 %%% --- Figures and Graphics
@@
 \usepackage{tikz}
-
-%%% --- Colors
-\usepackage[]{xcolor}
 
@@
 \begin{itemize}
-    \item Welche Form sollte das Maß haben? Mögliche Wahlen: Indikator, Probabilistisch, Weighted (1), Weighted (2), Indikator mit Korrektur, ... (Siehe Tabelle \ref{tab:comparison-measures})
-    \item Schätzung/Wahl des Rauschparameters $\sigma$; haben wir da irgendeine Methode aus den Daten das Sigma zu bekommen? Filter vs. Empirie 
+    \item Welche Form sollte das Maß haben? Mögliche Wahlen: Indikator, Probabilistisch, verschiedene Gewichtungen, ... (Siehe Tabelle \ref{tab:comparison-measures})
+    \item Indikator oder Signum?
+    \item Schätzung/Wahl des Rauschparameters $\sigma$ bei Maßen außer Indikator
     \item Gehen wir auf andere Verteilungen für den Fehler ein?
@@
 
+Comparsion and outline of various ideas for measures; overview in Table~\ref{tab:comparison-measures}:
+\begin{itemize}
+    \item Subsection~\ref{sec:equal-weight} Equal-weighted indicator
+    \item Subsection~\ref{subsec:w1} Weighted formula by explanatory power of points
+    \item Subsection~\ref{subsec:w2} Combined assessment of trend and its estimation uncertainty with weighting by explanatory power of points
+    \item A comparison of the Measures of Subsection~\ref{subsec:w1} and Subsection~\ref{subsec:w2} is given in \ref{subsec:w1-w2}
+    \item Subsection~\ref{subsec:probabilistic} Probabilistic version of estimation (as historic version)
+\end{itemize}
+
 %\begin{table}
@@
     Probabilistic & p & $\frac{1}{T-1} \sum P{(x_t y_t > 0)}$ & small variance, weighting due to probabilities & unintuitive, asymptotics unclear, driven towards 0.5, no standard estimation for $\sigma$  \\
-    weighted 1 (*) & w1 & {\small{$\sum w_i \ind{x_t y_t > 0} / \sum w_i$,\newline $w_i = \max_j\{P(y_i, x_i) \in Q_j \}$}} & intuitive, implicit exclusion area, natural extension & all points gain weight, estimation of $\sigma$.
-\\
-    weighted 2 (*) & w2 & {\small {$\sum w_i \ind{x_t y_t > 0} / \sum w_i$,\newline $w_i = \max_j\{P(y_i, x_i) \in Q_j \} - 0.25$}} & as w1, points near (0,0) do not gain weight & \\ \bottomrule
+    weighted 1 (*) & w1 & {\small {$\sum w_i \ind{x_t y_t > 0} / \sum w_i$,\newline $w_i = \max_j\{P(y_i, x_i) \in Q_j \} - 0.25$}} & intuitive, natural extension, implicit exclusion area as points near (0,0) do not gain weight & \\ \bottomrule
     \end{tabular}
-    \caption{Comparison between different measures for accuracy. }
+    \caption{Comparison between different measures for accuracy. For the weighted measures, the indicator could be replaced by the signum function of the product.}
     \label{tab:comparison-measures}
@@
 
-\unsure[inline]{Anpassung der gewichteten Formel}
-(*): The weighted formula can also be changed by using the signum function $\sign$ instead of $\ind$ to take into account the influence of negative trending.\\
+\subsubsection{Equal-weighted indicators} \label{sec:equal-weight}
 
-Natural evaluation method for 4Q-plot, as known from dichotomous event evaluation, evaluation of classification (confusion matrix)
+Natural evaluation method for 4Q-plot, as known from dichotomous event evaluation, evaluation of classification (confusion matrix):
 
-\begin{itemize}
-  \item Accuracy: \begin{equation}
-  	m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-w}
-\end{equation}
-	\item Capability of detecting a positive trend (probability of detection)/Sensitivity
-	\begin{equation}
-  		m_{\text{py}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\ydiff_t > 0}}
-	\end{equation}
-	\item positive predictive values
-\begin{equation}
-  m_{\text{px}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}  
-\end{equation}
-	\item Specificity
-	\begin{align}
-		m_{\text{ny}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} 
-  \end{align}
-  \item Negative predictive value
-	\begin{align}
-		m_{\text{nx}} &=  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\xdiff_t < 0}}
-	\end{align}
-\end{itemize}
+\begin{tabularx}{\textwidth}{p{5cm} X}
+     Accuracy & $m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-w}$ \\
+     Capability of detecting a positive trend (probability of detection)/Sensitivity & $m_{\text{py}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\ydiff_t > 0}}$ \\
+     Positive predictive values & $ m_{\text{px}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}$ \\
+     Specificity & $m_{\text{ny}} =  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} $ \\
+     Negative predictive value & $m_{\text{nx}} =  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\xdiff_t < 0}}$
+\end{tabularx}
 
@@
 
-Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$ (\hl{this, no additional complexity}):
-  	\begin{align}\label{additive error decomposition}
-  		\ydiff_t &= \ydifft + \varepsilon_t^y \\
-  		\xdiff_t &= \xdifft + \varepsilon_t^x
-	\end{align}
+% Extension 1: 
+% \begin{itemize}
+%   \item 
+% Adapt the measures to be weighted by the explanatory power of the points:\\
+% Weights $\omega_t$ for $(\ydiff_t, \xdiff_t)$ could be determined by error model and corresponds to the maximal likeliness of the point belonging to the different quadrants ($Q_1, \dots, Q_4$); thus, point that have less explanatory power for assessing the trending ability becomes less weight:
+%  	\begin{equation}
+%   		\omega_t = \max_{i \in [4]} {P\big( (\ydifft_t, \xdifft_t) \in Q_i \cond \ydiff_t, \xdiff_t \big)} 
+% 	\end{equation}
+% 	\item E.g., the accuracy becomes
+% 	\begin{equation}
+%   m_{acc, \omega} = \frac{\sum_{t=T-w}^T \omega_t \ind{\ydiff_t \cdot \xdiff_t > 0} }{\sum_{t=T-w}^T \omega_t}
+% \end{equation}
+% 	\item Thus, $\omega_t \in [0.25, 1]$. Maybe it would make sense to \enquote{ignore} values without explanatory power and use $\omega_t^\star = 4/3 \ \omega_t \in [0, 1]$ \hl{w2}
+% 	\item Figure~\ref{fig:comparison_estimators} shows the resulting estimates for the simulation study comparing the pure indicator, the probability-version and the weighted indicator estimation; the data is generated (in a very simpel model) by
+% 	\begin{align}
+% 		\ydifft_t &\sim N(0, 2) \\
+% 		\xdifft_t &= (2 b_t - 1) \cdot \ydiff_t , b_t \sim \Ber (k) \\
+% 		\ydiff_t &= \ydifft_t + \varepsilon_t^y, \varepsilon_t^y \sim N(0, 0.1) \\
+% 		\xdiff_t &= \xdifft_t + \varepsilon_t^x, \varepsilon_t^x \sim N(0, 0.1) \\
+% 	\end{align}
+% \end{itemize}
 
-Extension 1: 
-\begin{itemize}
-  \item 
-Adapt the measures to be weighted by the explanatory power of the points:\\
-Weights $\omega_t$ for $(\ydiff_t, \xdiff_t)$ could be determined by error model and corresponds to the maximal likeliness of the point belonging to the different quadrants ($Q_1, \dots, Q_4$); thus, point that have less explanatory power for assessing the trending ability becomes less weight:
+% \begin{figure}
+%   \centering
+% 	\includegraphics{plots/unbiasedness/boxplot_k_0.1.pdf}
+% 	\includegraphics{plots/unbiasedness/boxplot_k_0.3.pdf}
+% 	\includegraphics{plots/unbiasedness/boxplot_k_0.5.pdf}
+% 	\includegraphics{plots/unbiasedness/boxplot_k_0.7.pdf}
+% 	\includegraphics{plots/unbiasedness/boxplot_k_0.9.pdf}			
+%   \caption{Simulations results comparing the three types of estimators for the accuracy in a trending setting}
+%   \label{fig:comparison_estimators}
+% \end{figure}
+
+\subsubsection{Weighted measure by accounting for explanatory power of the points}\label{subsec:w1}
+
+Unlike the situation in dichotomous forecast evaluation or binary classification, the points $(\ydiff_t, \xdiff_t)$ are not binary.
+For points with small $\ydiff_t$ and $\xdiff_t$, the sign of their product is more likely to be driven by noise and not a systematic relationship.
+Weighting the observations by their likelihood to be driven by noise makes the estimation more robust and does not decrease the trending ability because of times of small and nonsystematic changes. 
+A simple method to account for the uncertainty in the measurement of $\ydiff_t$ and $\xdiff_t$ is to assume that the true values $\ydifft_t$ and $\xdifft$ follow (independent) Gaussian distributions around their observed counterparts with standard deviation $\sigma_y$ and $\sigma_y$, respectively.
+Then, the estimation can be weighted by how clearly a point can be assigned to one of the quadrants:
 \begin{equation}
-  		\omega_t = \max_{i \in [4]} {P\big( (\ydifft_t, \xdifft_t) \in Q_i \cond \ydiff_t, \xdiff_t \big)} 
+    m_{\text{acc}}^1 = \frac{\sum_{t=T-w}^T \omega_t \ind{\ydiff_t \xdiff_t > 0} }{\sum w_t}
 \end{equation}
-	\item E.g., the accuracy becomes
-	\begin{equation}
-  m_{acc, \omega} = \frac{\sum_{t=T-w}^T \omega_t \ind{\ydiff_t \cdot \xdiff_t > 0} }{\sum_{t=T-w}^T \omega_t}
-\end{equation}
-	\item Thus, $\omega_t \in [0.25, 1]$. Maybe it would make sense to \enquote{ignore} values without explanatory power and use $\omega_t^\star = 4/3 \ \omega_t \in [0, 1]$ \hl{w2}
-	\item Figure~\ref{fig:comparison_estimators} shows the resulting estimates for the simulation study comparing the pure indicator, the probability-version and the weighted indicator estimation; the data is generated (in a very simpel model) by
-	\begin{align}
-		\ydifft_t &\sim N(0, 2) \\
-		\xdifft_t &= (2 b_t - 1) \cdot \ydiff_t , b_t \sim \Ber (k) \\
-		\ydiff_t &= \ydifft_t + \varepsilon_t^y, \varepsilon_t^y \sim N(0, 0.1) \\
-		\xdiff_t &= \xdifft_t + \varepsilon_t^x, \varepsilon_t^x \sim N(0, 0.1) \\
-	\end{align}
-\end{itemize}
+with $w_t = \max_{i \in [4]} P((\ydiff_t \xdiff_t) \in Q_i) - 0.25$.
+Thus, the $w_t$ are in $[0, 0.75]$ and points very close to $(0, 0)$ do not have any weight in assessing the trend. 
+Due to the typical weight structure, $m_{\text{acc}}^1$ is in $[0,1]$ and the bounds are attained if and only if all points are concordant or disconcordant, respectively.
 
-\begin{figure}
-  \centering
-	\includegraphics{plots/unbiasedness/boxplot_k_0.1.pdf}
-	\includegraphics{plots/unbiasedness/boxplot_k_0.3.pdf}
-	\includegraphics{plots/unbiasedness/boxplot_k_0.5.pdf}
-	\includegraphics{plots/unbiasedness/boxplot_k_0.7.pdf}
-	\includegraphics{plots/unbiasedness/boxplot_k_0.9.pdf}			
-  \caption{Simulations results comparing the three types of estimators for the accuracy in a trending setting}
-  \label{fig:comparison_estimators}
-\end{figure}
+Instead of using the indicator of concordance, one could also use the signum function of the product of $(\ydiff_t, \xdiff_t)$ to scale the result to $(-1, 1)$.
 
-Probabilistic extension:
-\begin{itemize}
-\item Probabilistic versions of non-probabilistic measures; small sample characteristic of $m_{acc}$ can be driven by random effects near the axis; therefore probabilistic \enquote{correction}:
-\begin{itemize}
-  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$ (\hl{this, no additional complexity}):
-  	\begin{align}\label{additive error decomposition}
-  		\ydiff_t &= \ydifft + \varepsilon_t^y \\
-  		\xdiff_t &= \xdifft + \varepsilon_t^x
-	\end{align}
- \item Wouldn't it be more natural to model the original time series with a noise term? (Qualitatively, it would yield the same results)
- \begin{align}
-     y_t &= y_t^\star + \epsilon_t^y \\
-     x_t &= x_t^\star + \epsilon_t^x \\
-     \ydiff_t &= y_t^\star + \epsilon_t^y - (y_{t-1}^\star + \epsilon_{t-1}^y) = y_t^\star - y_{t-1}^\star + \underbrace{\epsilon_t^y - \epsilon_{t-1}^y}_{\sim N(0, \sqrt{2} \sigma_y)} \\
-     \xdiff_t &= x_t^\star + \epsilon_t^x - (x_{t-1}^\star + \epsilon_{t-1}^x) = x_t^\star - x_{t-1}^\star + \underbrace{\epsilon_t^x - \epsilon_{t-1}^x}_{\sim N(0, \sqrt{2} \sigma_x)}
- \end{align}
- Advantage of the approach: For different lags, no new standard deviation has to be estimated.\\
- Disadvantage: Theoretically, there is serial correlation of the errors. (Wie behandle ich die Autokorrelation; wie rausrechnen)
-\item  Replace indicators in the above formulations by the corresponding probabilities:
-\begin{align}
-		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-w}  \\
-   \mprob_{\text{py}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t > 0}} \\
-    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t > 0}} \\
-    \mprob_{\text{ny}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t < 0}} \\
-    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t < 0}} 
-\end{align} 
-\item Simple error model:
-  \begin{align}
-	  \varepsilon_Y \sim N(0, \sigma_Y) \\
-	  \varepsilon_X \sim N(0, \sigma_X)
-  \end{align}
-  yields, e.g.,
-  	\begin{equation}
-  		\mprob_{\text{acc}} = \frac{1}{T-w} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
-	\end{equation}
-	where $\Phi$ denotes a (possibly multivariate) normal distribution
-\end{itemize}
-\item For $\sigma_x \sigma_y \rightarrow 0$, the probabilistic measures approach the count-based ones
-\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic accuracy approaches   0.5 as all probabilities are approximately 0.5
-\item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
-\end{itemize}
-
-\unsure[inline]{Diskussion der Annahme der additiven Fehler-Dekomposition}
-
-Nach Annahme der additiven Fehler-Dekomposition ist es ohne weiteres nicht möglich, die probabilistische Formel, insbesondere den Ausdruck $P(Dy^*\cdot Dx^* > 0|Dx=Dx_t, Dy=Dy_t) $ auszurechnen, da $(Dx,Dy)$ und $(Dx^*,Dy^*)$ abhängig sind
-
-\subsection{Gewichtete Formel }
+\subsubsection{Combined measure of trending and its uncertainty} \label{subsec:w2}
 \unsure[inline]{Gewichtete Formel mit einer anderen Modelannahme und storyline}
@@
     \item Erwartungswert: Hier kann eine zusätzliche Annahme über die Verteilung der Unsicherheit getroffen werden.Oder man berechnet für einen hinreichend großen Bereich von $\sigma_x$ und $\sigma_y$ den Mittelwert der aus verschiedenen $\sigma_x$ und $\sigma_y$ resultierenden Maße.
+\end{itemize}
+
+\newpage
+\subsubsection{Comparison of weighted measures using examples} \label{subsec:w1-w2}
+
+\begin{tabularx}{\textwidth}{c l l l}
+     & Indicator & Weighted 1 & Weighted 2 \\
+     Verrauschte Daten & & & \\
+     Konzentration in Q1/Q3, aber viel an Entscheidungsgrenzen, kein Noise & & & \\
+     $\ydiff_t = \xdiff_t$ und $\ydiff_t \sim N(0, 1)$ & & & \\
+
+\end{tabularx}
+
+\newpage
+\subsubsection{Probabilistic measure (wrong derivation)} \label{subsec:probabilistic}
+
+\improvement[inline]{Due to the additive structure, $(Dx,Dy)$ and $(Dx^*,Dy^*)$ are not independet, and $P(Dy^*\cdot Dx^* > 0|Dx=Dx_t, Dy=Dy_t) $ is difficult to compute}
+
+\begin{itemize}
+\item Probabilistic versions of non-probabilistic measures; small sample characteristic of $m_{acc}$ can be driven by random effects near the axis; therefore probabilistic \enquote{correction}:
+\begin{itemize}
+  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$ (\hl{this, no additional complexity}):
+  	\begin{align}\label{additive error decomposition}
+  		\ydiff_t + \varepsilon_t^y &= \ydifft  \\
+  		\xdiff_t + \varepsilon_t^x &= \xdifft 
+	\end{align}
+\item  Replace indicators in the above formulations by the corresponding probabilities:
+\begin{align}
+		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-w}  \\
+   \mprob_{\text{py}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t > 0}} \\
+    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t > 0}} \\
+    \mprob_{\text{ny}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t < 0}} \\
+    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t < 0}} 
+\end{align} 
+\item Simple error model:
+  \begin{align}
+	  \varepsilon_Y \sim N(0, \sigma_Y) \\
+	  \varepsilon_X \sim N(0, \sigma_X)
+  \end{align}
+  yields, e.g.,
+  	\begin{equation}
+  		\mprob_{\text{acc}} = \frac{1}{T-w} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
+	\end{equation}
+	where $\Phi$ denotes a (possibly multivariate) normal distribution
+\end{itemize}
+\item For $\sigma_x \sigma_y \rightarrow 0$, the probabilistic measures approach the count-based ones
+\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic accuracy approaches   0.5 as all probabilities are approximately 0.5
+\item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
 \end{itemize}

### 2024-01-13 paper.tex
Update on Overleaf.

@@
      & Indicator & Weighted 1 & Weighted 2 \\
-     Verrauschte Daten & & & \\
+     \includegraphics{plots/simulation_com} & & & \\
      Konzentration in Q1/Q3, aber viel an Entscheidungsgrenzen, kein Noise & & & \\

### 2024-01-13 paper.tex
Update on Overleaf.

--- 2024-01-13 paper.tex
@@
 \usepackage{tabularx}
+\usepackage{tabularray}
 
@@
 
-\begin{tabularx}{\textwidth}{c l l l}
-     & Indicator & Weighted 1 & Weighted 2 \\
-     Verrauschte Daten & & & \\
-     Konzentration in Q1/Q3, aber viel an Entscheidungsgrenzen, kein Noise & & & \\
-     $\ydiff_t = \xdiff_t$ und $\ydiff_t \sim N(0, 1)$ & & & \\
+Compare the resulting trending assessments for the indicator formula and the weighted schemes. The horizontal axis displays $\ydiff$; the vertical axis corresponds to $\xdiff$. 
+The plots show 200 points, but the estimations are computed on 10,000 points.
 
-\end{tabularx}
+\noindent
+\begin{tblr}{
+  colspec = {Q[h]X[-1,m]X[-1,m]X[-1,m]|X[m]},
+  width=\textwidth,
+  stretch = 0,
+  rowsep = 6pt,
+}
+     & Indicator & Weighted 1 & Weighted 2 & Remarks \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7705 & 0.7906 & &  \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_no_noise.pdf} & 1 & 1 &  & {\scriptsize$\xdiff = \xdifft = 0.05 \ydifft = 0.05 \ydiff$}\newline critique on w1: slight deviation of $\xdiff$ from 0 (near decision boundary), nevertheless, full trending ability.
+\\
+     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_noise.pdf} & 0.6485 & 0.6835 & \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/perfect_normal.pdf} & 1 & 1 & & critique on w2: $\xdiff = \ydiff$, but not trending ability 1\\
+
+     \includegraphics{plots/simulation_comparison_weighted_measures/sure_unsure_combination.pdf} & 0.6152 & 0.6932 & & \\
+\end{tblr}
 

### 2024-01-13 paper.tex
Update on Overleaf.

@@
      & Indicator & Weighted 1 & Weighted 2 & Remarks \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7705 & 0.7906 & &  \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7748 & 0.7876 & &  \\
      \includegraphics{plots/simulation_comparison_weighted_measures/concordant_no_noise.pdf} & 1 & 1 &  & {\scriptsize$\xdiff = \xdifft = 0.05 \ydifft = 0.05 \ydiff$}\newline critique on w1: slight deviation of $\xdiff$ from 0 (near decision boundary), nevertheless, full trending ability.
 \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_noise.pdf} & 0.6485 & 0.6835 & \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_noise.pdf} & 0.6466 & 0.6823 & \\
      \includegraphics{plots/simulation_comparison_weighted_measures/perfect_normal.pdf} & 1 & 1 & & critique on w2: $\xdiff = \ydiff$, but not trending ability 1\\

### 2024-01-13 paper.tex
Update on Overleaf.

@@
 Unlike the situation in dichotomous forecast evaluation or binary classification, the points $(\ydiff_t, \xdiff_t)$ are not binary.
-For points with small $\ydiff_t$ and $\xdiff_t$, the sign of their product is more likely to be driven by noise and not a systematic relationship.
-Weighting the observations by their likelihood to be driven by noise makes the estimation more robust and does not decrease the trending ability because of times of small and nonsystematic changes. 
+For points with small $\ydiff_t$ and $\xdiff_t$, the sign of their product is more likely to be driven by noise or a nonsystematic relationship.
+Weighting the observations by their likelihood to be driven by noise makes the estimation more robust and prevents a decrease of the trending ability because of periods of small and nonsystematic changes.
 A simple method to account for the uncertainty in the measurement of $\ydiff_t$ and $\xdiff_t$ is to assume that the true values $\ydifft_t$ and $\xdifft$ follow (independent) Gaussian distributions around their observed counterparts with standard deviation $\sigma_y$ and $\sigma_y$, respectively.
@@
 Instead of using the indicator of concordance, one could also use the signum function of the product of $(\ydiff_t, \xdiff_t)$ to scale the result to $(-1, 1)$.
+Consistency to the accuracy measure from binary forecasting or classification argues for the indicator function.
+Similarly, the weights could be determined as the maximum probability of the true observation lying in $Q_1 \cup Q_3$ or $Q_2 \cup Q_4$.
+
+The estimation uncertainty of $m_{\text{acc}}^1$ can be determined by computing the confidence intervals of the estimation using bootstrapping or by resampling using the Gaussian distribution assumption.
 
@@
 
-\subsection{Bias-Variance tradeoff}
+\todo{Discuss aims of simulation studies, appropriate study designs and implement}
 
-Simulation:
-\begin{enumerate}
-    \item For $T = 50, 100, 500, 1000$, \verb|dist| in N(0,1), N(0,5), N(0,10), N(1,1), N(1,5), N(1,10), Chi-squared(3), Chi-squared(7), Chi-squared(9)
-    \begin{enumerate}
-        \item For $k \in \{0.05, \dots, 0.95\}$:
-        \begin{enumerate}
-            \item Sample $T$ samples from $\ydifft$ following \verb|dist|
-            \item For $t = 1, \dots, T$: Set $\xdifft_t = \ydifft_t$ with probability $k$, otherwise set $\ydifft_t = - \ydifft_t$
-            \item Set $\ydiff_t = \ydifft + N(0,0.1)$, $\xdiff_t = \xdifft + N(0,0.1)$
-            \item Compute all trending measures
-            \item Repeat 1,000 times
-        \end{enumerate}
-        \item Plot the trending estimates over $k$.
-    \end{enumerate}
-\end{enumerate}
+% \subsection{Bias-Variance tradeoff}
 
-\subsection{Different trending abilities for same rmse and mae} \label{sec:simulation_rmse_mae}
+% Simulation:
+% \begin{enumerate}
+%     \item For $T = 50, 100, 500, 1000$, \verb|dist| in N(0,1), N(0,5), N(0,10), N(1,1), N(1,5), N(1,10), Chi-squared(3), Chi-squared(7), Chi-squared(9)
+%     \begin{enumerate}
+%         \item For $k \in \{0.05, \dots, 0.95\}$:
+%         \begin{enumerate}
+%             \item Sample $T$ samples from $\ydifft$ following \verb|dist|
+%             \item For $t = 1, \dots, T$: Set $\xdifft_t = \ydifft_t$ with probability $k$, otherwise set $\ydifft_t = - \ydifft_t$
+%             \item Set $\ydiff_t = \ydifft + N(0,0.1)$, $\xdiff_t = \xdifft + N(0,0.1)$
+%             \item Compute all trending measures
+%             \item Repeat 1,000 times
+%         \end{enumerate}
+%         \item Plot the trending estimates over $k$.
+%     \end{enumerate}
+% \end{enumerate}
 
-Let the quantity of interest 
-\begin{equation}\label{eq: simulation}
-  y_t = a \sin(2 \pi t / T) + \varepsilon_t^y, \ t = 1, \dots, T
-\end{equation}
-with $\varepsilon_t^y \stackrel{\text{iid}}{\sim} N(0, \sigma_y)$ and 
-\begin{equation}
-  \ydiff_t = y_t - y_{t-1}. 
-\end{equation}
+% \subsection{Different trending abilities for same rmse and mae} \label{sec:simulation_rmse_mae}
 
-\todo[inline]{Anpassung der Simulationssettings an Gl. \ref{additive error decomposition}; Beschreibung Verbessern}
-The nowcasts are characterised by 
-\begin{align}
-	(\xdifft_t)^1 &= e^x_t \ydifft_t \\
-	(\xdifft_t)^2 &= \begin{cases}
-		e^x_t \ydifft_t &, \text{if}\ q^2_t = 0 \lor e^x_t \geq 0\\
-		(| e^x_t | + 2) \ydifft_t &, \text{if}\ q^2_t = 1 \land e^x_t < 0
-	\end{cases} \\
-	(\xdifft_t)^3 &= \begin{cases}
-		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 0 \lor e^x_t > 0) \land \ydifft_t \geq 0\\
-		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 1 \land e^x_t < 0) \land \ydifft_t \geq 0 \\
-		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 0 \lor e^x_t > 0) \land \ydifft_t < 0\\
-		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 1 \land e^x_t < 0) \land \ydifft_t < 0
-	\end{cases}
-\end{align}
-where $q^2_t \sim \text{Ber}(2p_2 - 1)$ and $p_2 > 0.5$ denotes the probability that the direction of nowcast 2 is correct.
-For nowcast 3 with $q^{3, \text{up}}_t \sim \text{Ber}(2p_3^{\text{up}}-1)$, $q^{3, \text{down}}_t \sim \text{Ber}(2p_3^{\text{down}}-1)$ the probabilities of direction correction depend on the direction of change.
+% Let the quantity of interest 
+% \begin{equation}\label{eq: simulation}
+%   y_t = a \sin(2 \pi t / T) + \varepsilon_t^y, \ t = 1, \dots, T
+% \end{equation}
+% with $\varepsilon_t^y \stackrel{\text{iid}}{\sim} N(0, \sigma_y)$ and 
+% \begin{equation}
+%   \ydiff_t = y_t - y_{t-1}. 
+% \end{equation}
+
+% \todo[inline]{Anpassung der Simulationssettings an Gl. \ref{additive error decomposition}; Beschreibung Verbessern}
+% The nowcasts are characterised by 
+% \begin{align}
+% 	(\xdifft_t)^1 &= e^x_t \ydifft_t \\
+% 	(\xdifft_t)^2 &= \begin{cases}
+% 		e^x_t \ydifft_t &, \text{if}\ q^2_t = 0 \lor e^x_t \geq 0\\
+% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ q^2_t = 1 \land e^x_t < 0
+% 	\end{cases} \\
+% 	(\xdifft_t)^3 &= \begin{cases}
+% 		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 0 \lor e^x_t > 0) \land \ydifft_t \geq 0\\
+% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 1 \land e^x_t < 0) \land \ydifft_t \geq 0 \\
+% 		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 0 \lor e^x_t > 0) \land \ydifft_t < 0\\
+% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 1 \land e^x_t < 0) \land \ydifft_t < 0
+% 	\end{cases}
+% \end{align}
+% where $q^2_t \sim \text{Ber}(2p_2 - 1)$ and $p_2 > 0.5$ denotes the probability that the direction of nowcast 2 is correct.
+% For nowcast 3 with $q^{3, \text{up}}_t \sim \text{Ber}(2p_3^{\text{up}}-1)$, $q^{3, \text{down}}_t \sim \text{Ber}(2p_3^{\text{down}}-1)$ the probabilities of direction correction depend on the direction of change.
 Additionally, let $e^x_t \sim N(0, 1)$.
 
-\begin{figure}
-  \centering
-  \includegraphics{plots/simulation_same_rmse_mae/time_series.pdf}
-  \caption{Realisation of Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$}
-  \label{fig:simulation_rmse_mae_ts}
-\end{figure}
+% \begin{figure}
+%   \centering
+%   \includegraphics{plots/simulation_same_rmse_mae/time_series.pdf}
+%   \caption{Realisation of Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$}
+%   \label{fig:simulation_rmse_mae_ts}
+% \end{figure}
 
-\begin{figure}
-  \centering
-  \begin{subfigure}{.32\textwidth}
-  	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_1}
-  \end{subfigure}
-  \begin{subfigure}{.32\textwidth}
-  	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_2}
-  \end{subfigure}
-    \begin{subfigure}{.32\textwidth}
-  	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_3}
-  \end{subfigure}
-  \caption{4Q-Plot for Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$.}
-  \label{fig:simulation_rmse_mae_4q}
-\end{figure}
+% \begin{figure}
+%   \centering
+%   \begin{subfigure}{.32\textwidth}
+%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_1}
+%   \end{subfigure}
+%   \begin{subfigure}{.32\textwidth}
+%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_2}
+%   \end{subfigure}
+%     \begin{subfigure}{.32\textwidth}
+%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_3}
+%   \end{subfigure}
+%   \caption{4Q-Plot for Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$.}
+%   \label{fig:simulation_rmse_mae_4q}
+% \end{figure}
 
-\todo[inline]{Weiteres Beispiel für viele richtigen Punkte sehr nah an den Entscheidungsgrenzen.}
-\subsection{Simulation: Bootstrap}
+% \todo[inline]{Weiteres Beispiel für viele richtigen Punkte sehr nah an den Entscheidungsgrenzen.}
+% \subsection{Simulation: Bootstrap}
 
-Simulation setup:
+% Simulation setup:
 
-\begin{itemize}
-  \item $T = 10,000$
-  \item $\ydiff_t \sim N(0, 1)$ 
-  \item $\xdiff_t^{1, k} = b_t  \ydiff_t, b_t \sim \SBer(k)$ (symmetric Bernoulli)
-\end{itemize}
-\unsure[inline]{(or both + random (non-systematic) noise?)}
+% \begin{itemize}
+%   \item $T = 10,000$
+%   \item $\ydiff_t \sim N(0, 1)$ 
+%   \item $\xdiff_t^{1, k} = b_t  \ydiff_t, b_t \sim \SBer(k)$ (symmetric Bernoulli)
+% \end{itemize}
+% \unsure[inline]{(or both + random (non-systematic) noise?)}
 
-Analysis:
-\begin{itemize}
-\item Do for $l = 1, \dots, 100$:
-\begin{itemize}
-  \item Sample $\ydiff, \xdiff$
-  \item Compute bootstrap confidence intervals $(b_{low}, b_{up})_l$ using $B = 10,000$
-\end{itemize}
-  \item Visualize all intervals with real value (confidence interval should cover real value in $\alpha$ cases)
-\end{itemize}
+% Analysis:
+% \begin{itemize}
+% \item Do for $l = 1, \dots, 100$:
+% \begin{itemize}
+%   \item Sample $\ydiff, \xdiff$
+%   \item Compute bootstrap confidence intervals $(b_{low}, b_{up})_l$ using $B = 10,000$
+% \end{itemize}
+%   \item Visualize all intervals with real value (confidence interval should cover real value in $\alpha$ cases)
+% \end{itemize}
 
@@
 \appendix
-\section{Illustrative examples for probabilistic measures}
+% \section{Illustrative examples for probabilistic measures}
 
-\begin{figure}
-    \centering
-    \includegraphics{plots/illustrative_examples/plots_4q_and_mtilde.pdf}
-    \caption{Various examplary courses of the measures $\tilde{m}$ for different configurations of $\ydiff$ and $\xdiff$}
-    \label{fig:example_mtilde}
-\end{figure}
+% \begin{figure}
+%     \centering
+%     \includegraphics{plots/illustrative_examples/plots_4q_and_mtilde.pdf}
+%     \caption{Various examplary courses of the measures $\tilde{m}$ for different configurations of $\ydiff$ and $\xdiff$}
+%     \label{fig:example_mtilde}
+% \end{figure}
 

### 2024-01-15 paper.tex
Update on Overleaf.

@@
 \subsubsection{Combined measure of trending and its uncertainty} \label{subsec:w2}
-\unsure[inline]{Gewichtete Formel mit einer anderen Modelannahme und storyline}
+The points $(Dx_t, Dy_t)$, which are close to the decision boundaries (origin and coordinate axes), should only be of limited informative value for the calculation of the trending measure, as it is uncertain whether such points really belong to the associated quadrants (1,3 or 2,4).
 
-Motivation: Die Punkte $(Dx_t, Dy_t)$, die nahe an den Entscheidungsgrenzen (Ursprung und Koordinatenachsen) liegen, sollen für die Berechnung des Trending-Maßes nur begrenzt aussagekräftig sein, da es unsicher ist, ob solche Punkte wirklich zu den zugehörigen Quadranten (1,3 oder 2,4) gehören.\\\\
-Idee: Sicherheit der Zugehörigkeit als Gewichtung in die Indikatorformel einführen.\\\\
-Annahme: Der wahre Wert eines beobachteten Wertepaares $(Dx_t,Dy_t)$ unterliegt einer Normalverteilung, die durch den Mittelpunkt $(Dx_t,Dy_t)$ und die Kovarianz  beschrieben 
+The certainty of belonging can be introduced as a weighting in the indicator formula.
+The simplest assumption is that the true value of an observed pair of values $(Dx_t,Dy_t)$ is subject to a normal distribution described by the centre $(Dx_t,Dy_t)$ and the covariance
 $\begin{pmatrix} \sigma_x & 0 \\
 0 & \sigma_y 
-\end{pmatrix}$ wird.\\\\
-Berechnung der Gewichte als Sicherheit der Zugehörigkeit: 
+\end{pmatrix}$.
+
+The weights as certainty of affiliation can be calculated as follows:
 \[ 
@@
 \end{cases}
-\]\\\\
-Dies ermöglicht eine kanonische Erweiterung der Indikatorformel:
+\]
+
+This enables a canonical extension of the indicator formula:
 
 \[m_{\text{acc}} = \frac{\sum_{t=T-w}^T \omega_t \cdot \text{sgn}(Dx_t\cdot Dy_t)}{T-w}\]\\\\
-Anderer Normierungsfaktor: z.B.
-$\sum_t w_t$ möglich, aber folgendes Problem: Wenn alle Punkte z.B.
-im ersten Quadranten liegen, dann verursacht der Normierungsfaktor $\sum_t w_t$ immer das Maß $1$, was in manchen Fällen nicht der Intuition entspricht, z.B.
-wenn die Punkte nahe an der Entscheidungsgrenze liegen (in diesem Fall würde man intuitiv eigentlich eine leicht positive Zahl erwarten).
-\\\\
-Bestimmung der Unsicherheit: $\sigma_x$ und $\sigma_y$ geben die Schätzung der Unsicherheit wieder, d.h.
-wie "weit" der wahre Punkt vom beobachteten Punkt im Vier-Quadranten-Plot entfernt sein kann.
-Es ist schwierig, die Standardabweichungen anhand der Daten zu bestimmen, da die beobachtbaren Daten diese Information nicht enthalten.
-Eine empirische Methode besteht z.
-B.
-darin, die empirische Varianz von $Dx$ und $Dy$ als Referenz zu nehmen und daraus Werte zu berechnen (z.
-B.
-10 $\%$ der Varianz von $Dx$ und $Dy$).\\\\
-Eine weitre Möglichkeit ist die Funktion $(\sigma_x,\sigma_y)\mapsto m_{\text{acc}}(\sigma_x,\sigma_y)$ zu betrachten und robuste Aussage (z.B. durch Optimierung) zu treffen.
+Other normalisation factor: e.g.
+$\sum_t w_t$ possible, but the following problem: If all points lie in the first quadrant, for example, then the normalisation factor $\sum_t w_t$ always causes the measure $1$, which in some cases does not correspond to intuition, e.g.
+if the points lie close to the decision limit (in this case one would intuitively expect a slightly positive number).
+
+Determination of the uncertainty: $\sigma_x$ and $\sigma_y$ reflect the estimate of the uncertainty, i.e.
+how "far" the true point can be from the observed point in the four-quadrant plot.
+It is difficult to determine the standard deviations from the data, as the observable data do not contain this information.
+One empirical method, for example, is to take the empirical variance of $Dx$ and $Dy$ as a reference and calculate values from it (e.g.
+10 $\%$ of the variance of $Dx$ and $Dy$).\\\\
+Another possibility is to look at the function $(\sigma_x,\sigma_y)\mapsto m_{\text{acc}}(\sigma_x,\sigma_y)$ and make a robust statement (e.g. by optimisation).
 
 \begin{itemize}
-     \item Maximum der Funktion: Wie viel Trending ist unter Berücksichtigung der Unsicherheit maximal möglich.
-    \item Minimum der Funktion: Wie viel Trend ist unter Berücksichtigung der Unsicherheit minimal möglich.
-    \item Erwartungswert: Hier kann eine zusätzliche Annahme über die Verteilung der Unsicherheit getroffen werden.Oder man berechnet für einen hinreichend großen Bereich von $\sigma_x$ und $\sigma_y$ den Mittelwert der aus verschiedenen $\sigma_x$ und $\sigma_y$ resultierenden Maße.
+    \item Maximum of the function: How much trending is maximally possible taking into account the uncertainty.
+    \item Minimum of the function: What is the minimum possible trend taking into account the uncertainty.
+    \item Expected value: An additional assumption about the distribution of the uncertainty can be made here, or the mean value of the measures resulting from different $\sigma_x$ and $\sigma_y$ can be calculated for a sufficiently large range of $\sigma_x$ and $\sigma_y$.
 \end{itemize}

### 2024-01-16 paper.tex
Update on Overleaf.

@@
      & Indicator & Weighted 1 & Weighted 2 & Remarks \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7748 & 0.7876 & &  \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_no_noise.pdf} & 1 & 1 &  & {\scriptsize$\xdiff = \xdifft = 0.05 \ydifft = 0.05 \ydiff$}\newline critique on w1: slight deviation of $\xdiff$ from 0 (near decision boundary), nevertheless, full trending ability.
+     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7672 & 0.7836 & 0.7694 &  \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_no_noise.pdf} & 1 & 1 & 0.8235 & {\scriptsize$\xdiff = \xdifft = 0.05 \ydifft = 0.05 \ydiff$}\newline critique on w1: slight deviation of $\xdiff$ from 0 (near decision boundary), nevertheless, full trending ability.
 \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_noise.pdf} & 0.6466 & 0.6823 & \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/perfect_normal.pdf} & 1 & 1 & & critique on w2: $\xdiff = \ydiff$, but not trending ability 1\\
+     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_noise.pdf} & 0.6466 & 0.6823 & 0.6266 & \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/perfect_normal.pdf} & 1 & 1 & 0.9779 & critique on w2: $\xdiff = \ydiff$, but not trending ability 1\\
 
-     \includegraphics{plots/simulation_comparison_weighted_measures/sure_unsure_combination.pdf} & 0.6152 & 0.6932 & & \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/sure_unsure_combination.pdf} & 0.6152 & 0.6932 & 0.6389 & \\
 \end{tblr}

### 2024-01-16 paper.tex
Update on Overleaf.

@@
 \begin{itemize}
-    \item Welche Form sollte das Maß haben? Mögliche Wahlen: Indikator, Probabilistisch, verschiedene Gewichtungen, ... (Siehe Tabelle \ref{tab:comparison-measures})
+    \item Welche Anforderungen sollte das Maß erfüllen
+    \item Welche Gewichtung sollte im Maß benutzt werden?
+    \item Steht im Nenner die Summe der Gewichte oder die Anzahl der Beobachtungen?
     \item Indikator oder Signum?
+    \item Punkte für die weitere Diskussion:
+    \begin{itemize}
         \item Schätzung/Wahl des Rauschparameters $\sigma$ bei Maßen außer Indikator
@@
         \item Konfidenzintervalle
+    \end{itemize}
 \end{itemize}
@@
     \item Subsection~\ref{subsec:probabilistic} Probabilistic version of estimation (as historic version)
+\end{itemize}
+
+Requirements for a reasonable measure:
+\begin{itemize}
+    \item $\ydiff = \xdiff$: trending measure should be 1
+    \item $\ydiff = - \xdiff$: trending measure should be 0 or -1
+    \item Points \enquote{close} to the axis should be treated differently than points that are far away from axis
+    \item (to be discussed) scalar multiplication of $y$ or $x$ should not change the trending measure
 \end{itemize}
@@
 with $w_t = \max_{i \in [4]} P((\ydiff_t \xdiff_t) \in Q_i) - 0.25$.
-Thus, the $w_t$ are in $[0, 0.75]$ and points very close to $(0, 0)$ do not have any weight in assessing the trend. 
+Thus, $w_t$ are in $[0, 0.75]$ and points very close to $(0, 0)$ do not have any weight in assessing the trend. 
 Due to the typical weight structure, $m_{\text{acc}}^1$ is in $[0,1]$ and the bounds are attained if and only if all points are concordant or disconcordant, respectively.
@@
 Instead of using the indicator of concordance, one could also use the signum function of the product of $(\ydiff_t, \xdiff_t)$ to scale the result to $(-1, 1)$.
-Consistency to the accuracy measure from binary forecasting or classification argues for the indicator function.
+Consistency to the accuracy measure from binary forecasting or classification is an argument for the indicator function.
 Similarly, the weights could be determined as the maximum probability of the true observation lying in $Q_1 \cup Q_3$ or $Q_2 \cup Q_4$.
@@
 \subsubsection{Combined measure of trending and its uncertainty} \label{subsec:w2}
-The points $(Dx_t, Dy_t)$, which are close to the decision boundaries (origin and coordinate axes), should only be of limited informative value for the calculation of the trending measure, as it is uncertain whether such points really belong to the associated quadrants (1,3 or 2,4).
+The points $(Dx_t, Dy_t)$, which are close to the decision boundaries (origin and coordinate axes), should only have a limited informative value for the calculation of the trending measure, as it is uncertain whether such points really belong to the associated quadrants (1,3 or 2,4).
 
 The certainty of belonging can be introduced as a weighting in the indicator formula.
-The simplest assumption is that the true value of an observed pair of values $(Dx_t,Dy_t)$ is subject to a normal distribution described by the centre $(Dx_t,Dy_t)$ and the covariance
+The simplest assumption is that the true value of an observed pair of values $(Dx_t,Dy_t)$ is subject to a normal distribution described by the center $(Dx_t,Dy_t)$ and the covariance
 $\begin{pmatrix} \sigma_x & 0 \\
@@
      & Indicator & Weighted 1 & Weighted 2 & Remarks \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7672 & 0.7836 & 0.7694 &  \\
+     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7661 & 0.7801 & 0.7605 &  \\
      \includegraphics{plots/simulation_comparison_weighted_measures/concordant_no_noise.pdf} & 1 & 1 & 0.8235 & {\scriptsize$\xdiff = \xdifft = 0.05 \ydifft = 0.05 \ydiff$}\newline critique on w1: slight deviation of $\xdiff$ from 0 (near decision boundary), nevertheless, full trending ability.
@@
      \includegraphics{plots/simulation_comparison_weighted_measures/perfect_normal.pdf} & 1 & 1 & 0.9779 & critique on w2: $\xdiff = \ydiff$, but not trending ability 1\\
-
-     \includegraphics{plots/simulation_comparison_weighted_measures/sure_unsure_combination.pdf} & 0.6152 & 0.6932 & 0.6389 & \\
+    \includegraphics{plots/simulation_comparison_weighted_measures/sure_unsure_combination.pdf} & 0.7460 & 0.8429 & 0.7360 & \\
 \end{tblr}

### 2024-01-23 paper.tex
Update on Overleaf.

--- 2024-01-15 paper.tex
@@
 
-\subsection*{Offene Punkte}
-
-\begin{itemize}
-    \item Welche Form sollte das Maß haben? Mögliche Wahlen: Indikator, Probabilistisch, verschiedene Gewichtungen, ... (Siehe Tabelle \ref{tab:comparison-measures})
-    \item Indikator oder Signum?
-    \item Schätzung/Wahl des Rauschparameters $\sigma$ bei Maßen außer Indikator
-    \item Gehen wir auf andere Verteilungen für den Fehler ein?
-    \item Konfidenzintervalle
-\end{itemize}
-
 \listoftodos
@@
 \begin{itemize}
-	\item Trending is an important feature of nowcasts to be identify changes in the situation before the actual quantity can be measured
-	\item What scores are currently used?
-	\item Examples: Covid, BIP, Medicine, ... 
- \item Nice an comprehensive introduction to nowcasting, so that it becomes clear for non-nowcasters what is special about nowcasting
- \item Measurements and Forecasting; What are the differences to nowcasting from a modeling point of view?
+    \item 
 \end{itemize}
 
-\section{Trend measures for nowcasts}
+\section{Trending assessment}
 
-Notation
 \begin{itemize}
-  \item Let $y_1, \ldots, y_T$ be the values of the nowcasted quantity.
-  \item Let $x_t^k$ be the $k$-th nowcast for time $t = 1, \ldots, T$ (possibly after some aggregation if several raw nowcasts apply).
-  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - \mathcal{x}_{t-1}$ the change nowcaster $k$ issued between the time steps.
-Let us first assume that $y_{t-1}$ is available at time $t$ so that it can be used for the evaluation.
-  In Section \hl{error-model-section}, we expand on other reasonable choices for $\mathcal{x}_{t-1}$.
-  \item Then, $\xdiff_t^k$ can be seen as the point forecast of nowcaster $k$ for $\ydiff_t$. 
+    \item Es interessiert uns $P(x y) > 0$
+    \item 
 \end{itemize}
 
-\subsection{What is the trend of a nowcast?}
-
-Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$ (Problem: mit kleinen Fehlern; Hinweis Exclusion area)
-
-\begin{itemize}
-  \item Translate into mathematical notation
-  \item Illustrative Example: Two nowcasts of quantity of interest, where one jumps up and down and one points in right direction with the same rmse; small subset of Simulation~\ref{sec:simulation_rmse_mae}; Plot time series \& time series of differences
-\end{itemize}
-
-\subsection{Graphical methods}
-
-\begin{itemize}
-  \item 4Q-Plot
-  \item Andere Medizin-Plots mit Verweis auf Grothe-Paper
-\end{itemize}
-
-\subsection{Measures}
-
-Comparsion and outline of various ideas for measures; overview in Table~\ref{tab:comparison-measures}:
-\begin{itemize}
-    \item Subsection~\ref{sec:equal-weight} Equal-weighted indicator
-    \item Subsection~\ref{subsec:w1} Weighted formula by explanatory power of points
-    \item Subsection~\ref{subsec:w2} Combined assessment of trend and its estimation uncertainty with weighting by explanatory power of points
-    \item A comparison of the Measures of Subsection~\ref{subsec:w1} and Subsection~\ref{subsec:w2} is given in \ref{subsec:w1-w2}
-    \item Subsection~\ref{subsec:probabilistic} Probabilistic version of estimation (as historic version)
-\end{itemize}
-
-%\begin{table}
-%	\centering
-%	\begin{tabular}{l l c c}
-%	\toprule
-%	& & \multicolumn{2}{c}{Observed trend} \\
-%	& & $\ydiff_t > 0$ & $\ydiff_t < 0$ \\	\cline{3-4}
-%	\multirow{2}{*}{Forecasted trend} & $\xdiff_t > 0$ & hit & false alarm\\
-%	& $\xdiff_t < 0$ & miss & correct negative \\
-%	\bottomrule
-%\end{tabular}
-%\end{table}
-
-\begin{table}[]
-    \centering
-    \begin{tabular}{l l p{4cm} p{4cm} p{4cm}}
-    \toprule
-    name & abbr. & accuracy & + & - \\ \midrule
-    indicator & i & $\frac{1}{T-1} \sum \ind{x_t y_t > 0}$ &  well-known, intuitive & what about points near (0,0) \\
-    Probabilistic & p & $\frac{1}{T-1} \sum P{(x_t y_t > 0)}$ & small variance, weighting due to probabilities & unintuitive, asymptotics unclear, driven towards 0.5, no standard estimation for $\sigma$  \\
-    weighted 1 (*) & w1 & {\small {$\sum w_i \ind{x_t y_t > 0} / \sum w_i$,\newline $w_i = \max_j\{P(y_i, x_i) \in Q_j \} - 0.25$}} & intuitive, natural extension, implicit exclusion area as points near (0,0) do not gain weight & \\ \bottomrule
-    \end{tabular}
-    \caption{Comparison between different measures for accuracy. For the weighted measures, the indicator could be replaced by the signum function of the product.}
-    \label{tab:comparison-measures}
-\end{table}
-
-\subsubsection{Equal-weighted indicators} \label{sec:equal-weight}
-
-Natural evaluation method for 4Q-plot, as known from dichotomous event evaluation, evaluation of classification (confusion matrix):
-
-\begin{tabularx}{\textwidth}{p{5cm} X}
-     Accuracy & $m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-w}$ \\
-     Capability of detecting a positive trend (probability of detection)/Sensitivity & $m_{\text{py}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\ydiff_t > 0}}$ \\
-     Positive predictive values & $ m_{\text{px}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}$ \\
-     Specificity & $m_{\text{ny}} =  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} $ \\
-     Negative predictive value & $m_{\text{nx}} =  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\xdiff_t < 0}}$
-\end{tabularx}
-
-Different setting here: We do not only have binary predictions and realization, but we do know the real values behind; intuitive evaluation of a 4Q-plot assigns a greater explanatory power to point far away from (0,0), as small deviations from (0,0) could be driven by noise.
-
-% Extension 1: 
-% \begin{itemize}
-%   \item 
-% Adapt the measures to be weighted by the explanatory power of the points:\\
-% Weights $\omega_t$ for $(\ydiff_t, \xdiff_t)$ could be determined by error model and corresponds to the maximal likeliness of the point belonging to the different quadrants ($Q_1, \dots, Q_4$); thus, point that have less explanatory power for assessing the trending ability becomes less weight:
-%  	\begin{equation}
-%   		\omega_t = \max_{i \in [4]} {P\big( (\ydifft_t, \xdifft_t) \in Q_i \cond \ydiff_t, \xdiff_t \big)} 
-% 	\end{equation}
-% 	\item E.g., the accuracy becomes
-% 	\begin{equation}
-%   m_{acc, \omega} = \frac{\sum_{t=T-w}^T \omega_t \ind{\ydiff_t \cdot \xdiff_t > 0} }{\sum_{t=T-w}^T \omega_t}
-% \end{equation}
-% 	\item Thus, $\omega_t \in [0.25, 1]$. Maybe it would make sense to \enquote{ignore} values without explanatory power and use $\omega_t^\star = 4/3 \ \omega_t \in [0, 1]$ \hl{w2}
-% 	\item Figure~\ref{fig:comparison_estimators} shows the resulting estimates for the simulation study comparing the pure indicator, the probability-version and the weighted indicator estimation; the data is generated (in a very simpel model) by
-% 	\begin{align}
-% 		\ydifft_t &\sim N(0, 2) \\
-% 		\xdifft_t &= (2 b_t - 1) \cdot \ydiff_t , b_t \sim \Ber (k) \\
-% 		\ydiff_t &= \ydifft_t + \varepsilon_t^y, \varepsilon_t^y \sim N(0, 0.1) \\
-% 		\xdiff_t &= \xdifft_t + \varepsilon_t^x, \varepsilon_t^x \sim N(0, 0.1) \\
-% 	\end{align}
-% \end{itemize}
-
-% \begin{figure}
-%   \centering
-% 	\includegraphics{plots/unbiasedness/boxplot_k_0.1.pdf}
-% 	\includegraphics{plots/unbiasedness/boxplot_k_0.3.pdf}
-% 	\includegraphics{plots/unbiasedness/boxplot_k_0.5.pdf}
-% 	\includegraphics{plots/unbiasedness/boxplot_k_0.7.pdf}
-% 	\includegraphics{plots/unbiasedness/boxplot_k_0.9.pdf}			
-%   \caption{Simulations results comparing the three types of estimators for the accuracy in a trending setting}
-%   \label{fig:comparison_estimators}
-% \end{figure}
-
-\subsubsection{Weighted measure by accounting for explanatory power of the points}\label{subsec:w1}
-
-Unlike the situation in dichotomous forecast evaluation or binary classification, the points $(\ydiff_t, \xdiff_t)$ are not binary.
-For points with small $\ydiff_t$ and $\xdiff_t$, the sign of their product is more likely to be driven by noise or a nonsystematic relationship.
-Weighting the observations by their likelihood to be driven by noise makes the estimation more robust and prevents a decrease of the trending ability because of periods of small and nonsystematic changes.
-A simple method to account for the uncertainty in the measurement of $\ydiff_t$ and $\xdiff_t$ is to assume that the true values $\ydifft_t$ and $\xdifft$ follow (independent) Gaussian distributions around their observed counterparts with standard deviation $\sigma_y$ and $\sigma_y$, respectively.
-Then, the estimation can be weighted by how clearly a point can be assigned to one of the quadrants:
-\begin{equation}
-    m_{\text{acc}}^1 = \frac{\sum_{t=T-w}^T \omega_t \ind{\ydiff_t \xdiff_t > 0} }{\sum w_t}
-\end{equation}
-with $w_t = \max_{i \in [4]} P((\ydiff_t \xdiff_t) \in Q_i) - 0.25$.
-Thus, the $w_t$ are in $[0, 0.75]$ and points very close to $(0, 0)$ do not have any weight in assessing the trend. 
-Due to the typical weight structure, $m_{\text{acc}}^1$ is in $[0,1]$ and the bounds are attained if and only if all points are concordant or disconcordant, respectively.
-
-Instead of using the indicator of concordance, one could also use the signum function of the product of $(\ydiff_t, \xdiff_t)$ to scale the result to $(-1, 1)$.
-Consistency to the accuracy measure from binary forecasting or classification argues for the indicator function.
-Similarly, the weights could be determined as the maximum probability of the true observation lying in $Q_1 \cup Q_3$ or $Q_2 \cup Q_4$.
-
-The estimation uncertainty of $m_{\text{acc}}^1$ can be determined by computing the confidence intervals of the estimation using bootstrapping or by resampling using the Gaussian distribution assumption.
-
-\subsubsection{Combined measure of trending and its uncertainty} \label{subsec:w2}
-The points $(Dx_t, Dy_t)$, which are close to the decision boundaries (origin and coordinate axes), should only be of limited informative value for the calculation of the trending measure, as it is uncertain whether such points really belong to the associated quadrants (1,3 or 2,4).
-
-The certainty of belonging can be introduced as a weighting in the indicator formula.
-The simplest assumption is that the true value of an observed pair of values $(Dx_t,Dy_t)$ is subject to a normal distribution described by the centre $(Dx_t,Dy_t)$ and the covariance
-$\begin{pmatrix} \sigma_x & 0 \\
-0 & \sigma_y 
-\end{pmatrix}$.
-
-The weights as certainty of affiliation can be calculated as follows:
-\[ 
-w_t = 
-\begin{cases} 
-  P(Dx_t \cdot Dy_t > 0) = \left(1 - \Phi\left(\frac{-Dx_t}{\sigma_X}\right)\right)\left(1 - \Phi\left(\frac{-Dy_t}{\sigma_Y}\right)\right) + \Phi\left(\frac{-Dx_t}{\sigma_X}\right) \Phi\left(\frac{-Dy_t}{\sigma_Y}\right) & \text{wenn } Dx_t \cdot Dy_t > 0,  \\
-  P(Dx_t \cdot Dy_t < 0) = \left(1 - \Phi\left(\frac{-Dx_t}{\sigma_X}\right)\right) \Phi\left(\frac{-Dy_t}{\sigma_Y}\right) + \Phi\left(\frac{-Dx_t}{\sigma_X}\right) \left(1 - \Phi\left(\frac{-Dy_t}{\sigma_Y}\right)\right)  & \text{wenn } Dx_t \cdot Dy_t < 0,  \\
-  0 & \text{sonst} 
-\end{cases}
-\]
-
-This enables a canonical extension of the indicator formula:
-
-\[m_{\text{acc}} = \frac{\sum_{t=T-w}^T \omega_t \cdot \text{sgn}(Dx_t\cdot Dy_t)}{T-w}\]\\\\
-Other normalisation factor: e.g.
-$\sum_t w_t$ possible, but the following problem: If all points lie in the first quadrant, for example, then the normalisation factor $\sum_t w_t$ always causes the measure $1$, which in some cases does not correspond to intuition, e.g.
-if the points lie close to the decision limit (in this case one would intuitively expect a slightly positive number).
-
-Determination of the uncertainty: $\sigma_x$ and $\sigma_y$ reflect the estimate of the uncertainty, i.e.
-how "far" the true point can be from the observed point in the four-quadrant plot.
-It is difficult to determine the standard deviations from the data, as the observable data do not contain this information.
-One empirical method, for example, is to take the empirical variance of $Dx$ and $Dy$ as a reference and calculate values from it (e.g.
-10 $\%$ of the variance of $Dx$ and $Dy$).\\\\
-Another possibility is to look at the function $(\sigma_x,\sigma_y)\mapsto m_{\text{acc}}(\sigma_x,\sigma_y)$ and make a robust statement (e.g. by optimisation).
-
-\begin{itemize}
-    \item Maximum of the function: How much trending is maximally possible taking into account the uncertainty.
-    \item Minimum of the function: What is the minimum possible trend taking into account the uncertainty.
-    \item Expected value: An additional assumption about the distribution of the uncertainty can be made here, or the mean value of the measures resulting from different $\sigma_x$ and $\sigma_y$ can be calculated for a sufficiently large range of $\sigma_x$ and $\sigma_y$.
-\end{itemize}
-
-\newpage
-\subsubsection{Comparison of weighted measures using examples} \label{subsec:w1-w2}
-
-Compare the resulting trending assessments for the indicator formula and the weighted schemes. The horizontal axis displays $\ydiff$; the vertical axis corresponds to $\xdiff$. 
-The plots show 200 points, but the estimations are computed on 10,000 points.
-
-\noindent
-\begin{tblr}{
-  colspec = {Q[h]X[-1,m]X[-1,m]X[-1,m]|X[m]},
-  width=\textwidth,
-  stretch = 0,
-  rowsep = 6pt,
-}
-     & Indicator & Weighted 1 & Weighted 2 & Remarks \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7748 & 0.7876 & &  \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_no_noise.pdf} & 1 & 1 &  & {\scriptsize$\xdiff = \xdifft = 0.05 \ydifft = 0.05 \ydiff$}\newline critique on w1: slight deviation of $\xdiff$ from 0 (near decision boundary), nevertheless, full trending ability.
-\\
-     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_noise.pdf} & 0.6466 & 0.6823 & \\
-     \includegraphics{plots/simulation_comparison_weighted_measures/perfect_normal.pdf} & 1 & 1 & & critique on w2: $\xdiff = \ydiff$, but not trending ability 1\\
-
-     \includegraphics{plots/simulation_comparison_weighted_measures/sure_unsure_combination.pdf} & 0.6152 & 0.6932 & & \\
-\end{tblr}
-
-\newpage
-\subsubsection{Probabilistic measure (wrong derivation)} \label{subsec:probabilistic}
-
-\improvement[inline]{Due to the additive structure, $(Dx,Dy)$ and $(Dx^*,Dy^*)$ are not independet, and $P(Dy^*\cdot Dx^* > 0|Dx=Dx_t, Dy=Dy_t) $ is difficult to compute}
-
-\begin{itemize}
-\item Probabilistic versions of non-probabilistic measures; small sample characteristic of $m_{acc}$ can be driven by random effects near the axis; therefore probabilistic \enquote{correction}:
-\begin{itemize}
-  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$ (\hl{this, no additional complexity}):
-  	\begin{align}\label{additive error decomposition}
-  		\ydiff_t + \varepsilon_t^y &= \ydifft  \\
-  		\xdiff_t + \varepsilon_t^x &= \xdifft 
-	\end{align}
-\item  Replace indicators in the above formulations by the corresponding probabilities:
-\begin{align}
-		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-w}  \\
-   \mprob_{\text{py}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t > 0}} \\
-    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t > 0}} \\
-    \mprob_{\text{ny}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t < 0}} \\
-    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t < 0}} 
-\end{align} 
-\item Simple error model:
-  \begin{align}
-	  \varepsilon_Y \sim N(0, \sigma_Y) \\
-	  \varepsilon_X \sim N(0, \sigma_X)
-  \end{align}
-  yields, e.g.,
-  	\begin{equation}
-  		\mprob_{\text{acc}} = \frac{1}{T-w} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
-	\end{equation}
-	where $\Phi$ denotes a (possibly multivariate) normal distribution
-\end{itemize}
-\item For $\sigma_x \sigma_y \rightarrow 0$, the probabilistic measures approach the count-based ones
-\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic accuracy approaches   0.5 as all probabilities are approximately 0.5
-\item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
-\end{itemize}
-
-\subsection{Estimating the standard deviation of white noise}
-
-First, sometimes variance can be determined heuristically directly from domain knowledge. 
-However, to determine the standard deviation of white noise using only the time series data, one can basically follow two approaches, time domain approach or frequency domain approach.
-For the first approach, a model assumption (e.g.
-ARIMA models) is specified and the residual analysis technique is applied [\cite{chitturi1974distribution}].
-In the frequency domain approach, the time series are first decomposed using spectral methods such as (discrete) Fourier transformation [\cite{sauer1992noise}] or wavelet transformation [\cite{heil1989continuous,donoho1994ideal,jansen2012noise}].
-After reconstructing the time series, the residuals can be formed in order to determine the standard deviation of white noise.
-For the use of wavelet, it is particularly convenient that the standard deviation of the white noise can be estimated by the standard deviation of the detail coefficients of the last level, e.g., if the wavelet family db1 is used.
-
-For the noisy signal in the form of \ref{eq: simulation},  there are the following results (cf. pic \ref{fig:estimation std})
-
-%\begin{figure}
-%  \centering
-%    \begin{subfigure}{0.24\textwidth}
-%        \includegraphics{plots/std estimation/std_estimation_boxplot_25.pdf}
-%    \end{subfigure}
-%    \begin{subfigure}{0.24\textwidth}
-%        \includegraphics{plots/std estimation/std_estimation_boxplot_50.pdf}
-%    \end{subfigure}
-%    \begin{subfigure}{0.24\textwidth}
-%        \includegraphics{plots/std estimation/std_estimation_boxplot_100.pdf}
-%    \end{subfigure}
-%    \begin{subfigure}{0.24\textwidth}
-%        \includegraphics{plots/std estimation/std_estimation_boxplot_1000.pdf}
-%    \end{subfigure}
-%  \caption{Estimation of std with different methods.}
-%  \label{fig:estimation std}
-%\end{figure}%
-
-\begin{figure}
-  \centering
-    \begin{subfigure}{0.48\textwidth}
-        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_25.pdf}
-    \end{subfigure}
-    \begin{subfigure}{0.48\textwidth}
-        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_50.pdf}
-    \end{subfigure}
-
-    \begin{subfigure}{0.48\textwidth}
-        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_100.pdf}
-    \end{subfigure}
-    \begin{subfigure}{0.48\textwidth}
-        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_1000.pdf}
-    \end{subfigure}
-
-  \caption{Estimation of std using different methods.}
-  \label{fig:estimation std}
-\end{figure}
-
-\subsection{Confidence intervals and missing data}
-
-\begin{itemize}
-  \item Bootstrapping: Compare percentile and BCa approach?
-  \item Bootstrapping: What is the \enquote{true} value?
-  \item Missing data
-\end{itemize}
-
-\subsection{Trending over various time lags}
-
-\begin{itemize}
-    \item Introduce plot over several time lags
-    \item Question: Re-estimate noise for every lag? 
-\end{itemize}
-
-\section{Simulation studies}
-
-\todo{Discuss aims of simulation studies, appropriate study designs and implement}
-
-% \subsection{Bias-Variance tradeoff}
-
-% Simulation:
-% \begin{enumerate}
-%     \item For $T = 50, 100, 500, 1000$, \verb|dist| in N(0,1), N(0,5), N(0,10), N(1,1), N(1,5), N(1,10), Chi-squared(3), Chi-squared(7), Chi-squared(9)
-%     \begin{enumerate}
-%         \item For $k \in \{0.05, \dots, 0.95\}$:
-%         \begin{enumerate}
-%             \item Sample $T$ samples from $\ydifft$ following \verb|dist|
-%             \item For $t = 1, \dots, T$: Set $\xdifft_t = \ydifft_t$ with probability $k$, otherwise set $\ydifft_t = - \ydifft_t$
-%             \item Set $\ydiff_t = \ydifft + N(0,0.1)$, $\xdiff_t = \xdifft + N(0,0.1)$
-%             \item Compute all trending measures
-%             \item Repeat 1,000 times
-%         \end{enumerate}
-%         \item Plot the trending estimates over $k$.
-%     \end{enumerate}
-% \end{enumerate}
-
-% \subsection{Different trending abilities for same rmse and mae} \label{sec:simulation_rmse_mae}
-
-% Let the quantity of interest 
-% \begin{equation}\label{eq: simulation}
-%   y_t = a \sin(2 \pi t / T) + \varepsilon_t^y, \ t = 1, \dots, T
-% \end{equation}
-% with $\varepsilon_t^y \stackrel{\text{iid}}{\sim} N(0, \sigma_y)$ and 
-% \begin{equation}
-%   \ydiff_t = y_t - y_{t-1}. 
-% \end{equation}
-
-% \todo[inline]{Anpassung der Simulationssettings an Gl. \ref{additive error decomposition}; Beschreibung Verbessern}
-% The nowcasts are characterised by 
-% \begin{align}
-% 	(\xdifft_t)^1 &= e^x_t \ydifft_t \\
-% 	(\xdifft_t)^2 &= \begin{cases}
-% 		e^x_t \ydifft_t &, \text{if}\ q^2_t = 0 \lor e^x_t \geq 0\\
-% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ q^2_t = 1 \land e^x_t < 0
-% 	\end{cases} \\
-% 	(\xdifft_t)^3 &= \begin{cases}
-% 		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 0 \lor e^x_t > 0) \land \ydifft_t \geq 0\\
-% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 1 \land e^x_t < 0) \land \ydifft_t \geq 0 \\
-% 		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 0 \lor e^x_t > 0) \land \ydifft_t < 0\\
-% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 1 \land e^x_t < 0) \land \ydifft_t < 0
-% 	\end{cases}
-% \end{align}
-% where $q^2_t \sim \text{Ber}(2p_2 - 1)$ and $p_2 > 0.5$ denotes the probability that the direction of nowcast 2 is correct.
-% For nowcast 3 with $q^{3, \text{up}}_t \sim \text{Ber}(2p_3^{\text{up}}-1)$, $q^{3, \text{down}}_t \sim \text{Ber}(2p_3^{\text{down}}-1)$ the probabilities of direction correction depend on the direction of change.
-Additionally, let $e^x_t \sim N(0, 1)$.
-
-% \begin{figure}
-%   \centering
-%   \includegraphics{plots/simulation_same_rmse_mae/time_series.pdf}
-%   \caption{Realisation of Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$}
-%   \label{fig:simulation_rmse_mae_ts}
-% \end{figure}
-
-% \begin{figure}
-%   \centering
-%   \begin{subfigure}{.32\textwidth}
-%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_1}
-%   \end{subfigure}
-%   \begin{subfigure}{.32\textwidth}
-%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_2}
-%   \end{subfigure}
-%     \begin{subfigure}{.32\textwidth}
-%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_3}
-%   \end{subfigure}
-%   \caption{4Q-Plot for Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$.}
-%   \label{fig:simulation_rmse_mae_4q}
-% \end{figure}
-
-% \todo[inline]{Weiteres Beispiel für viele richtigen Punkte sehr nah an den Entscheidungsgrenzen.}
-% \subsection{Simulation: Bootstrap}
-
-% Simulation setup:
-
-% \begin{itemize}
-%   \item $T = 10,000$
-%   \item $\ydiff_t \sim N(0, 1)$ 
-%   \item $\xdiff_t^{1, k} = b_t  \ydiff_t, b_t \sim \SBer(k)$ (symmetric Bernoulli)
-% \end{itemize}
-% \unsure[inline]{(or both + random (non-systematic) noise?)}
-
-% Analysis:
-% \begin{itemize}
-% \item Do for $l = 1, \dots, 100$:
-% \begin{itemize}
-%   \item Sample $\ydiff, \xdiff$
-%   \item Compute bootstrap confidence intervals $(b_{low}, b_{up})_l$ using $B = 10,000$
-% \end{itemize}
-%   \item Visualize all intervals with real value (confidence interval should cover real value in $\alpha$ cases)
-% \end{itemize}
-
-\subsection{Effect of increasing sample size?}
-
-\section{Application on disease and economic data}
-
-\subsection{Nowcasts for the COVID infections in Germany}
-
-\begin{itemize}
-  \item Describe QoI and Nowcasts; which nowcasts do we consider? Which data is available when?
-  \item Plot data
-\end{itemize}
-
-\subsection{Nowcasting of GDP}
-
-\section{Conclusion}
-\medskip
-\printbibliography
-
-\appendix
-% \section{Illustrative examples for probabilistic measures}
-
-% \begin{figure}
-%     \centering
-%     \includegraphics{plots/illustrative_examples/plots_4q_and_mtilde.pdf}
-%     \caption{Various examplary courses of the measures $\tilde{m}$ for different configurations of $\ydiff$ and $\xdiff$}
-%     \label{fig:example_mtilde}
-% \end{figure}
-
 \end{document}

### 2024-01-23 paper_20240123.tex
Update on Overleaf.

\documentclass[oneside]{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}
\usepackage[utf8]{inputenc} % für Umlaute
\usepackage{fontenc}

% Set page size and margins
% Replace `letterpaper' with`a4paper' for UK/EU standard size
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%%%% ---- Useful packages

%%% --- Colors
\usepackage[dvipsnames]{xcolor}

%%% --- Figures and Graphics
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{tikz}

%%% --- Tabulars
\usepackage{booktabs}
\usepackage{multirow} % cells over several rows
\usepackage{tabularx}
\usepackage{tabularray}

%%% --- Mathematics
\usepackage{amsmath,amsfonts,amssymb,amsthm,mathtools, mathabx}
\usepackage{bm} % bold mathematics
\usepackage{bbm} % for blackboard 1
\usepackage[linesnumbered,lined,algo2e,figure,boxed]{algorithm2e}
\usepackage[short]{optidef} % for aligning linear programs

%%% --- Other
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
%\usepackage[hidelinks]{hyperref}
\usepackage[style=authoryear,backend=bibtex]{biblatex}
\addbibresource{library.bib}
\usepackage{csquotes}
\usepackage[shortcuts]{extdash} % For Hyphens in English
%\usepackage{acro}
% \DeclareAcronym{pdf}{short=PDF, long={probability distribution function}}

%%% --- Commenting, Todonotes
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
\usepackage{comment, soul}
\newcommand{\ar}{$\rightarrow$}
\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
%\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}

\usepackage{xargs} % Use optional arguements in commands
\newcommandx{\unsure}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\missing}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
\usepackage{marginnote}
\let\marginpar\marginnote

%%%% ---- Own Commands
%%% --- Theorem Settings
\theoremstyle{plain}% Theorem-like structures provided by amsthm.sty
\newtheorem{theorem}{Theorem}
\newtheorem{exa}{Example}
\newtheorem{rem}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{example}{Example}

%%% --- Math Commands

\newcommand{\lag}[1][l]{\Delta_{#1}}
\DeclarePairedDelimiter{\abs}\lvert\rvert
\DeclareMathOperator{\sign}{sign}
\newcommand{\tmax}{\bar{t}}
\newcommand{\ind}[1]{\mathbbm{1}\{#1\}}
\newcommand{\ydiff}{D y}
\newcommand{\ydifft}{Dy^\star}
\newcommand{\xdiff}{Dx}
\newcommand{\xdifft}{Dx^\star}
\newcommand{\Ydiff}{DY}
\newcommand{\Xdiff}{DX}
\newcommand{\Prob}[1]{P(#1)}
\newcommand{\mprob}{\tilde{m}}
\newcommand{\Ber}{\text{Ber}}
\newcommand{\SBer}{\text{SBer}}
\newcommand{\cond}{\:\lvert\:}

%%% --- Other Commands
\renewcommand{\arraystretch}{1.2}

%%% --- ToDo List
\usepackage{enumitem}
\newlist{todolist}{itemize}{2}
\setlist[todolist]{label=$\square$}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
\hspace{-2.5pt}}
\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}

\title{Assessment and Optimization of Nowcast Measures for Trend Detection}
\author{Oliver Grothe, Bolin Liu, Jonas Rieger}

\begin{document}
\maketitle

\begin{abstract}
Existing and new measures for the capability of capturing the trend of nowcasts are presented, evaluated, and compared on synthetic and real-world data.
\end{abstract}

\subsection*{Offene Punkte}

\begin{itemize}
    \item Welche Anforderungen sollte das Maß erfüllen
    \item Welche Gewichtung sollte im Maß benutzt werden?
    \item Steht im Nenner die Summe der Gewichte oder die Anzahl der Beobachtungen?
    \item Indikator oder Signum?
    \item Punkte für die weitere Diskussion:
    \begin{itemize}
        \item Schätzung/Wahl des Rauschparameters $\sigma$ bei Maßen außer Indikator
        \item Gehen wir auf andere Verteilungen für den Fehler ein?
        \item Konfidenzintervalle
    \end{itemize}
    \item Was passiert mit missing values?
\end{itemize}

\listoftodos

\section{Introduction}

\begin{itemize}
	\item Trending is an important feature of nowcasts to be identify changes in the situation before the actual quantity can be measured
	\item What scores are currently used?
	\item Examples: Covid, BIP, Medicine, ... 
 \item Nice an comprehensive introduction to nowcasting, so that it becomes clear for non-nowcasters what is special about nowcasting
 \item Measurements and Forecasting; What are the differences to nowcasting from a modeling point of view?
\end{itemize}

\section{Trend measures for nowcasts}

Notation
\begin{itemize}
  \item Let $y_1, \ldots, y_T$ be the values of the nowcasted quantity.
  \item Let $x_t^k$ be the $k$-th nowcast for time $t = 1, \ldots, T$ (possibly after some aggregation if several raw nowcasts apply).
  \item Let $\ydiff_t = y_t - y_{t-1}$ be the change that occurred between the time steps $t-1$ and $t$ and $\xdiff_t^k = x_t - \mathcal{x}_{t-1}$ the change nowcaster $k$ issued between the time steps. Let us first assume that $y_{t-1}$ is available at time $t$ so that it can be used for the evaluation. 
  In Section \hl{error-model-section}, we expand on other reasonable choices for $\mathcal{x}_{t-1}$.
  \item Then, $\xdiff_t^k$ can be seen as the point forecast of nowcaster $k$ for $\ydiff_t$. 
\end{itemize}

\subsection{What is the trend of a nowcast?}

Trend: Issuing the right \enquote{direction} of change between two time steps; statistical consistency of $\sign(\ydiff_t)$ and $\sign(\xdiff_t)$ (Problem: mit kleinen Fehlern; Hinweis Exclusion area)

\begin{itemize}
  \item Translate into mathematical notation
  \item Illustrative Example: Two nowcasts of quantity of interest, where one jumps up and down and one points in right direction with the same rmse; small subset of Simulation~\ref{sec:simulation_rmse_mae}; Plot time series \& time series of differences
\end{itemize}

\subsection{Graphical methods}

\begin{itemize}
  \item 4Q-Plot
  \item Andere Medizin-Plots mit Verweis auf Grothe-Paper
\end{itemize}

\subsection{Measures}

Comparsion and outline of various ideas for measures; overview in Table~\ref{tab:comparison-measures}:
\begin{itemize}
    \item Subsection~\ref{sec:equal-weight} Equal-weighted indicator
    \item Subsection~\ref{subsec:w1} Weighted formula by explanatory power of points
    \item Subsection~\ref{subsec:w2} Combined assessment of trend and its estimation uncertainty with weighting by explanatory power of points
    \item A comparison of the Measures of Subsection~\ref{subsec:w1} and Subsection~\ref{subsec:w2} is given in \ref{subsec:w1-w2}
    \item Subsection~\ref{subsec:probabilistic} Probabilistic version of estimation (as historic version)
\end{itemize}

Requirements for a reasonable measure:
\begin{itemize}
    \item $\ydiff = \xdiff$: trending measure should be 1
    \item $\ydiff = - \xdiff$: trending measure should be 0 or -1
    \item Points \enquote{close} to the axis should be treated differently than points that are far away from axis
    \item (to be discussed) scalar multiplication of $y$ or $x$ should not change the trending measure
\end{itemize}

%\begin{table}
%	\centering
%	\begin{tabular}{l l c c}
%	\toprule
%	& & \multicolumn{2}{c}{Observed trend} \\
%	& & $\ydiff_t > 0$ & $\ydiff_t < 0$ \\	\cline{3-4}
%	\multirow{2}{*}{Forecasted trend} & $\xdiff_t > 0$ & hit & false alarm\\
%	& $\xdiff_t < 0$ & miss & correct negative \\
%	\bottomrule
%\end{tabular}
%\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{l l p{4cm} p{4cm} p{4cm}}
    \toprule
    name & abbr. & accuracy & + & - \\ \midrule
    indicator & i & $\frac{1}{T-1} \sum \ind{x_t y_t > 0}$ &  well-known, intuitive & what about points near (0,0) \\
    Probabilistic & p & $\frac{1}{T-1} \sum P{(x_t y_t > 0)}$ & small variance, weighting due to probabilities & unintuitive, asymptotics unclear, driven towards 0.5, no standard estimation for $\sigma$  \\
    weighted 1 (*) & w1 & {\small {$\sum w_i \ind{x_t y_t > 0} / \sum w_i$,\newline $w_i = \max_j\{P(y_i, x_i) \in Q_j \} - 0.25$}} & intuitive, natural extension, implicit exclusion area as points near (0,0) do not gain weight & \\ \bottomrule
    \end{tabular}
    \caption{Comparison between different measures for accuracy. For the weighted measures, the indicator could be replaced by the signum function of the product.}
    \label{tab:comparison-measures}
\end{table}

\subsubsection{Equal-weighted indicators} \label{sec:equal-weight}

Natural evaluation method for 4Q-plot, as known from dichotomous event evaluation, evaluation of classification (confusion matrix):

\begin{tabularx}{\textwidth}{p{5cm} X}
     Accuracy & $m_{\text{acc}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0} + \sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{T-w}$ \\
     Capability of detecting a positive trend (probability of detection)/Sensitivity & $m_{\text{py}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\ydiff_t > 0}}$ \\
     Positive predictive values & $ m_{\text{px}} = \frac{\sum_{t=T-w}^T \ind{\ydiff_t > 0, \xdiff_t > 0}}{\sum_{t=T-w}^T \ind{\xdiff_t > 0}}$ \\
     Specificity & $m_{\text{ny}} =  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\ydiff_t < 0}} $ \\
     Negative predictive value & $m_{\text{nx}} =  \frac{\sum_{t=T-w}^T \ind{\ydiff_t < 0, \xdiff_t < 0}}{\sum_{t=T-w}^T \ind{\xdiff_t < 0}}$
\end{tabularx}

Different setting here: We do not only have binary predictions and realization, but we do know the real values behind; intuitive evaluation of a 4Q-plot assigns a greater explanatory power to point far away from (0,0), as small deviations from (0,0) could be driven by noise.

% Extension 1: 
% \begin{itemize}
%   \item 
% Adapt the measures to be weighted by the explanatory power of the points:\\
% Weights $\omega_t$ for $(\ydiff_t, \xdiff_t)$ could be determined by error model and corresponds to the maximal likeliness of the point belonging to the different quadrants ($Q_1, \dots, Q_4$); thus, point that have less explanatory power for assessing the trending ability becomes less weight:
%  	\begin{equation}
%   		\omega_t = \max_{i \in [4]} {P\big( (\ydifft_t, \xdifft_t) \in Q_i \cond \ydiff_t, \xdiff_t \big)} 
% 	\end{equation}
% 	\item E.g., the accuracy becomes
% 	\begin{equation}
%   m_{acc, \omega} = \frac{\sum_{t=T-w}^T \omega_t \ind{\ydiff_t \cdot \xdiff_t > 0} }{\sum_{t=T-w}^T \omega_t}
% \end{equation}
% 	\item Thus, $\omega_t \in [0.25, 1]$. Maybe it would make sense to \enquote{ignore} values without explanatory power and use $\omega_t^\star = 4/3 \ \omega_t \in [0, 1]$ \hl{w2}
% 	\item Figure~\ref{fig:comparison_estimators} shows the resulting estimates for the simulation study comparing the pure indicator, the probability-version and the weighted indicator estimation; the data is generated (in a very simpel model) by
% 	\begin{align}
% 		\ydifft_t &\sim N(0, 2) \\
% 		\xdifft_t &= (2 b_t - 1) \cdot \ydiff_t , b_t \sim \Ber (k) \\
% 		\ydiff_t &= \ydifft_t + \varepsilon_t^y, \varepsilon_t^y \sim N(0, 0.1) \\
% 		\xdiff_t &= \xdifft_t + \varepsilon_t^x, \varepsilon_t^x \sim N(0, 0.1) \\
% 	\end{align}
% \end{itemize}

% \begin{figure}
%   \centering
% 	\includegraphics{plots/unbiasedness/boxplot_k_0.1.pdf}
% 	\includegraphics{plots/unbiasedness/boxplot_k_0.3.pdf}
% 	\includegraphics{plots/unbiasedness/boxplot_k_0.5.pdf}
% 	\includegraphics{plots/unbiasedness/boxplot_k_0.7.pdf}
% 	\includegraphics{plots/unbiasedness/boxplot_k_0.9.pdf}			
%   \caption{Simulations results comparing the three types of estimators for the accuracy in a trending setting}
%   \label{fig:comparison_estimators}
% \end{figure}

\subsubsection{Weighted measure by accounting for explanatory power of the points}\label{subsec:w1}

Unlike the situation in dichotomous forecast evaluation or binary classification, the points $(\ydiff_t, \xdiff_t)$ are not binary.
For points with small $\ydiff_t$ and $\xdiff_t$, the sign of their product is more likely to be driven by noise or a nonsystematic relationship.
Weighting the observations by their likelihood to be driven by noise makes the estimation more robust and prevents a decrease of the trending ability because of periods of small and nonsystematic changes. 
A simple method to account for the uncertainty in the measurement of $\ydiff_t$ and $\xdiff_t$ is to assume that the true values $\ydifft_t$ and $\xdifft$ follow (independent) Gaussian distributions around their observed counterparts with standard deviation $\sigma_y$ and $\sigma_y$, respectively.
Then, the estimation can be weighted by how clearly a point can be assigned to one of the quadrants:
\begin{equation}
    m_{\text{acc}}^1 = \frac{\sum_{t=T-w}^T \omega_t \ind{\ydiff_t \xdiff_t > 0} }{\sum w_t}
\end{equation}
with $w_t = \max_{i \in [4]} P((\ydiff_t \xdiff_t) \in Q_i) - 0.25$.
Thus, $w_t$ are in $[0, 0.75]$ and points very close to $(0, 0)$ do not have any weight in assessing the trend. 
Due to the typical weight structure, $m_{\text{acc}}^1$ is in $[0,1]$ and the bounds are attained if and only if all points are concordant or disconcordant, respectively.

Instead of using the indicator of concordance, one could also use the signum function of the product of $(\ydiff_t, \xdiff_t)$ to scale the result to $(-1, 1)$.
Consistency to the accuracy measure from binary forecasting or classification is an argument for the indicator function.
Similarly, the weights could be determined as the maximum probability of the true observation lying in $Q_1 \cup Q_3$ or $Q_2 \cup Q_4$.

The estimation uncertainty of $m_{\text{acc}}^1$ can be determined by computing the confidence intervals of the estimation using bootstrapping or by resampling using the Gaussian distribution assumption.

\subsubsection{Combined measure of trending and its uncertainty} \label{subsec:w2}
The points $(Dx_t, Dy_t)$, which are close to the decision boundaries (origin and coordinate axes), should only have a limited informative value for the calculation of the trending measure, as it is uncertain whether such points really belong to the associated quadrants (1,3 or 2,4).

The certainty of belonging can be introduced as a weighting in the indicator formula. The simplest assumption is that the true value of an observed pair of values $(Dx_t,Dy_t)$ is subject to a normal distribution described by the center $(Dx_t,Dy_t)$ and the covariance
$\begin{pmatrix} \sigma_x & 0 \\
0 & \sigma_y 
\end{pmatrix}$.

The weights as certainty of affiliation can be calculated as follows:
\[ 
w_t = 
\begin{cases} 
  P(Dx_t \cdot Dy_t > 0) = \left(1 - \Phi\left(\frac{-Dx_t}{\sigma_X}\right)\right)\left(1 - \Phi\left(\frac{-Dy_t}{\sigma_Y}\right)\right) + \Phi\left(\frac{-Dx_t}{\sigma_X}\right) \Phi\left(\frac{-Dy_t}{\sigma_Y}\right) & \text{wenn } Dx_t \cdot Dy_t > 0,  \\
  P(Dx_t \cdot Dy_t < 0) = \left(1 - \Phi\left(\frac{-Dx_t}{\sigma_X}\right)\right) \Phi\left(\frac{-Dy_t}{\sigma_Y}\right) + \Phi\left(\frac{-Dx_t}{\sigma_X}\right) \left(1 - \Phi\left(\frac{-Dy_t}{\sigma_Y}\right)\right)  & \text{wenn } Dx_t \cdot Dy_t < 0,  \\
  0 & \text{sonst} 
\end{cases}
\]

This enables a canonical extension of the indicator formula:

\[m_{\text{acc}} = \frac{\sum_{t=T-w}^T \omega_t \cdot \text{sgn}(Dx_t\cdot Dy_t)}{T-w}\]\\\\
Other normalisation factor: e.g. $\sum_t w_t$ possible, but the following problem: If all points lie in the first quadrant, for example, then the normalisation factor $\sum_t w_t$ always causes the measure $1$, which in some cases does not correspond to intuition, e.g. if the points lie close to the decision limit (in this case one would intuitively expect a slightly positive number).

Determination of the uncertainty: $\sigma_x$ and $\sigma_y$ reflect the estimate of the uncertainty, i.e. how "far" the true point can be from the observed point in the four-quadrant plot. It is difficult to determine the standard deviations from the data, as the observable data do not contain this information. One empirical method, for example, is to take the empirical variance of $Dx$ and $Dy$ as a reference and calculate values from it (e.g. 10 $\%$ of the variance of $Dx$ and $Dy$).\\\\
Another possibility is to look at the function $(\sigma_x,\sigma_y)\mapsto m_{\text{acc}}(\sigma_x,\sigma_y)$ and make a robust statement (e.g. by optimisation).

\begin{itemize}
    \item Maximum of the function: How much trending is maximally possible taking into account the uncertainty.
    \item Minimum of the function: What is the minimum possible trend taking into account the uncertainty.
    \item Expected value: An additional assumption about the distribution of the uncertainty can be made here, or the mean value of the measures resulting from different $\sigma_x$ and $\sigma_y$ can be calculated for a sufficiently large range of $\sigma_x$ and $\sigma_y$.
\end{itemize}

\newpage
\subsubsection{Comparison of weighted measures using examples} \label{subsec:w1-w2}

Compare the resulting trending assessments for the indicator formula and the weighted schemes. The horizontal axis displays $\ydiff$; the vertical axis corresponds to $\xdiff$. 
The plots show 200 points, but the estimations are computed on 10,000 points.

\noindent
\begin{tblr}{
  colspec = {Q[h]X[-1,m]X[-1,m]X[-1,m]|X[m]},
  width=\textwidth,
  stretch = 0,
  rowsep = 6pt,
}
     & Indicator & Weighted 1 & Weighted 2 & Remarks \\
     \includegraphics{plots/simulation_comparison_weighted_measures/simple.pdf} & 0.7661 & 0.7801 & 0.7605 &  \\
     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_no_noise.pdf} & 1 & 1 & 0.8235 & {\scriptsize$\xdiff = \xdifft = 0.05 \ydifft = 0.05 \ydiff$}\newline critique on w1: slight deviation of $\xdiff$ from 0 (near decision boundary), nevertheless, full trending ability. \\
     \includegraphics{plots/simulation_comparison_weighted_measures/concordant_noise.pdf} & 0.6466 & 0.6823 & 0.6266 & \\
     \includegraphics{plots/simulation_comparison_weighted_measures/perfect_normal.pdf} & 1 & 1 & 0.9779 & critique on w2: $\xdiff = \ydiff$, but not trending ability 1\\
    \includegraphics{plots/simulation_comparison_weighted_measures/sure_unsure_combination.pdf} & 0.7460 & 0.8429 & 0.7360 & \\
\end{tblr}

\newpage
\subsubsection{Probabilistic measure (wrong derivation)} \label{subsec:probabilistic}

\improvement[inline]{Due to the additive structure, $(Dx,Dy)$ and $(Dx^*,Dy^*)$ are not independet, and $P(Dy^*\cdot Dx^* > 0|Dx=Dx_t, Dy=Dy_t) $ is difficult to compute}

\begin{itemize}
\item Probabilistic versions of non-probabilistic measures; small sample characteristic of $m_{acc}$ can be driven by random effects near the axis; therefore probabilistic \enquote{correction}:
\begin{itemize}
  \item Assume an additive error decomposition for both observation and nowcast to account for non-systematic and short-term (i.e., intra-day) effects or for effects on inaccurate representation of $y_{t-1}$ in the computation of $Dx_t$ (\hl{this, no additional complexity}):
  	\begin{align}\label{additive error decomposition}
  		\ydiff_t + \varepsilon_t^y &= \ydifft  \\
  		\xdiff_t + \varepsilon_t^x &= \xdifft 
	\end{align}
\item  Replace indicators in the above formulations by the corresponding probabilities:
\begin{align}
		\mprob_{\text{acc}} &= \frac{\sum_{t=T-w}^T \Prob{ \ydifft_t > 0, \xdifft_t > 0} + \sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{T-w}  \\
   \mprob_{\text{py}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t > 0}} \\
    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t > 0, \xdifft_t > 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t > 0}} \\
    \mprob_{\text{ny}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\ydifft_t < 0}} \\
    \mprob_{\text{px}} &= \frac{\sum_{t=T-w}^T \Prob{\ydifft_t < 0, \xdifft_t < 0}}{\sum_{t=T-w}^T \Prob{\xdifft_t < 0}} 
\end{align} 
\item Simple error model:
  \begin{align}
	  \varepsilon_Y \sim N(0, \sigma_Y) \\
	  \varepsilon_X \sim N(0, \sigma_X)
  \end{align}
  yields, e.g.,
  	\begin{equation}
  		\mprob_{\text{acc}} = \frac{1}{T-w} \sum_{t=T-w}^T  \big( 1 - \Phi_{\ydiff_t, \sigma_y}(0) - \Phi_{\xdiff_t, \sigma_x} (0) + 2 \Phi_{\xdiff_t,\sigma_x}( 0)\Phi_{\ydiff_t,\sigma_y}( 0) \big), 
	\end{equation}
	where $\Phi$ denotes a (possibly multivariate) normal distribution
\end{itemize}
\item For $\sigma_x \sigma_y \rightarrow 0$, the probabilistic measures approach the count-based ones
\item For $\sigma_x \sigma_y \rightarrow \infty$, the probabilistic accuracy approaches   0.5 as all probabilities are approximately 0.5
\item Lag-$l$-measures can easily be obtained by using the $D^l$-difference, i.e., $D^l y_t = y_t - y_{t-l}$
\end{itemize}

\subsection{Estimating the standard deviation of white noise}

First, sometimes variance can be determined heuristically directly from domain knowledge. 
However, to determine the standard deviation of white noise using only the time series data, one can basically follow two approaches, time domain approach or frequency domain approach. For the first approach, a model assumption (e.g. ARIMA models) is specified and the residual analysis technique is applied [\cite{chitturi1974distribution}]. 
In the frequency domain approach, the time series are first decomposed using spectral methods such as (discrete) Fourier transformation [\cite{sauer1992noise}] or wavelet transformation [\cite{heil1989continuous,donoho1994ideal,jansen2012noise}]. After reconstructing the time series, the residuals can be formed in order to determine the standard deviation of white noise. For the use of wavelet, it is particularly convenient that the standard deviation of the white noise can be estimated by the standard deviation of the detail coefficients of the last level, e.g., if the wavelet family db1 is used.

For the noisy signal in the form of \ref{eq: simulation},  there are the following results (cf. pic \ref{fig:estimation std})

%\begin{figure}
%  \centering
%    \begin{subfigure}{0.24\textwidth}
%        \includegraphics{plots/std estimation/std_estimation_boxplot_25.pdf}
%    \end{subfigure}
%    \begin{subfigure}{0.24\textwidth}
%        \includegraphics{plots/std estimation/std_estimation_boxplot_50.pdf}
%    \end{subfigure}
%    \begin{subfigure}{0.24\textwidth}
%        \includegraphics{plots/std estimation/std_estimation_boxplot_100.pdf}
%    \end{subfigure}
%    \begin{subfigure}{0.24\textwidth}
%        \includegraphics{plots/std estimation/std_estimation_boxplot_1000.pdf}
%    \end{subfigure}
%  \caption{Estimation of std with different methods.}
%  \label{fig:estimation std}
%\end{figure}%

\begin{figure}
  \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_25.pdf}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_50.pdf}
    \end{subfigure}

    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_100.pdf}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{plots/std_estimation/std_estimation_boxplot_1000.pdf}
    \end{subfigure}

  \caption{Estimation of std using different methods.}
  \label{fig:estimation std}
\end{figure}

\subsection{Confidence intervals and missing data}

\begin{itemize}
  \item Bootstrapping: Compare percentile and BCa approach?
  \item Bootstrapping: What is the \enquote{true} value?
  \item Missing data
\end{itemize}

\subsection{Trending over various time lags}

\begin{itemize}
    \item Introduce plot over several time lags
    \item Question: Re-estimate noise for every lag? 
\end{itemize}

\section{Simulation studies}

\todo{Discuss aims of simulation studies, appropriate study designs and implement}

% \subsection{Bias-Variance tradeoff}

% Simulation:
% \begin{enumerate}
%     \item For $T = 50, 100, 500, 1000$, \verb|dist| in N(0,1), N(0,5), N(0,10), N(1,1), N(1,5), N(1,10), Chi-squared(3), Chi-squared(7), Chi-squared(9)
%     \begin{enumerate}
%         \item For $k \in \{0.05, \dots, 0.95\}$:
%         \begin{enumerate}
%             \item Sample $T$ samples from $\ydifft$ following \verb|dist|
%             \item For $t = 1, \dots, T$: Set $\xdifft_t = \ydifft_t$ with probability $k$, otherwise set $\ydifft_t = - \ydifft_t$
%             \item Set $\ydiff_t = \ydifft + N(0,0.1)$, $\xdiff_t = \xdifft + N(0,0.1)$
%             \item Compute all trending measures
%             \item Repeat 1,000 times
%         \end{enumerate}
%         \item Plot the trending estimates over $k$.
%     \end{enumerate}
% \end{enumerate}

% \subsection{Different trending abilities for same rmse and mae} \label{sec:simulation_rmse_mae}

% Let the quantity of interest 
% \begin{equation}\label{eq: simulation}
%   y_t = a \sin(2 \pi t / T) + \varepsilon_t^y, \ t = 1, \dots, T
% \end{equation}
% with $\varepsilon_t^y \stackrel{\text{iid}}{\sim} N(0, \sigma_y)$ and 
% \begin{equation}
%   \ydiff_t = y_t - y_{t-1}. 
% \end{equation}

% \todo[inline]{Anpassung der Simulationssettings an Gl. \ref{additive error decomposition}; Beschreibung Verbessern}
% The nowcasts are characterised by 
% \begin{align}
% 	(\xdifft_t)^1 &= e^x_t \ydifft_t \\
% 	(\xdifft_t)^2 &= \begin{cases}
% 		e^x_t \ydifft_t &, \text{if}\ q^2_t = 0 \lor e^x_t \geq 0\\
% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ q^2_t = 1 \land e^x_t < 0
% 	\end{cases} \\
% 	(\xdifft_t)^3 &= \begin{cases}
% 		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 0 \lor e^x_t > 0) \land \ydifft_t \geq 0\\
% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{down}}_t = 1 \land e^x_t < 0) \land \ydifft_t \geq 0 \\
% 		e^x_t \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 0 \lor e^x_t > 0) \land \ydifft_t < 0\\
% 		(| e^x_t | + 2) \ydifft_t &, \text{if}\ (q^{3, \text{up}}_t = 1 \land e^x_t < 0) \land \ydifft_t < 0
% 	\end{cases}
% \end{align}
% where $q^2_t \sim \text{Ber}(2p_2 - 1)$ and $p_2 > 0.5$ denotes the probability that the direction of nowcast 2 is correct.
% For nowcast 3 with $q^{3, \text{up}}_t \sim \text{Ber}(2p_3^{\text{up}}-1)$, $q^{3, \text{down}}_t \sim \text{Ber}(2p_3^{\text{down}}-1)$ the probabilities of direction correction depend on the direction of change. Additionally, let $e^x_t \sim N(0, 1)$.

% \begin{figure}
%   \centering
%   \includegraphics{plots/simulation_same_rmse_mae/time_series.pdf}
%   \caption{Realisation of Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$}
%   \label{fig:simulation_rmse_mae_ts}
% \end{figure}

% \begin{figure}
%   \centering
%   \begin{subfigure}{.32\textwidth}
%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_1}
%   \end{subfigure}
%   \begin{subfigure}{.32\textwidth}
%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_2}
%   \end{subfigure}
%     \begin{subfigure}{.32\textwidth}
%   	\includegraphics{plots/simulation_same_rmse_mae/4q_plot_3}
%   \end{subfigure}
%   \caption{4Q-Plot for Simulation~\ref{sec:simulation_rmse_mae} with $T = 100$, $\sigma_y=0.1$, $a = 10$, $p_2 = 0.8$, and $p_3 = (0.9, 0.5)$.}
%   \label{fig:simulation_rmse_mae_4q}
% \end{figure}

% \todo[inline]{Weiteres Beispiel für viele richtigen Punkte sehr nah an den Entscheidungsgrenzen.}
% \subsection{Simulation: Bootstrap}

% Simulation setup:

% \begin{itemize}
%   \item $T = 10,000$
%   \item $\ydiff_t \sim N(0, 1)$ 
%   \item $\xdiff_t^{1, k} = b_t  \ydiff_t, b_t \sim \SBer(k)$ (symmetric Bernoulli)
% \end{itemize}
% \unsure[inline]{(or both + random (non-systematic) noise?)}

% Analysis:
% \begin{itemize}
% \item Do for $l = 1, \dots, 100$:
% \begin{itemize}
%   \item Sample $\ydiff, \xdiff$
%   \item Compute bootstrap confidence intervals $(b_{low}, b_{up})_l$ using $B = 10,000$
% \end{itemize}
%   \item Visualize all intervals with real value (confidence interval should cover real value in $\alpha$ cases)
% \end{itemize}

\subsection{Effect of increasing sample size?}

\section{Application on disease and economic data}

\subsection{Nowcasts for the COVID infections in Germany}

\begin{itemize}
  \item Describe QoI and Nowcasts; which nowcasts do we consider? Which data is available when?
  \item Plot data
\end{itemize}

\subsection{Nowcasting of GDP}

\section{Conclusion}
\medskip
\printbibliography

\appendix
% \section{Illustrative examples for probabilistic measures}

% \begin{figure}
%     \centering
%     \includegraphics{plots/illustrative_examples/plots_4q_and_mtilde.pdf}
%     \caption{Various examplary courses of the measures $\tilde{m}$ for different configurations of $\ydiff$ and $\xdiff$}
%     \label{fig:example_mtilde}
% \end{figure}

\end{document}

### 2024-02-02 sections/01-introduction.tex
Update on Overleaf.

\begin{itemize}
    \item Aktualität/Wichtigkeit herausstellen: 
    \begin{itemize}
        \item zwei Nowcasts mit gleichem MSE, aber unterschiedlichem Trending $\rightarrow$ Vergleich zwischen Nowcasts
    \end{itemize}
    \item Einordnung in Literatur
    \item Contributions
    \begin{itemize}
        \item Argue, why trending is important and why other measures do not \enquote{detect} it
        \item (Formalisation of Trending Ability)
        \item Evaluation and review of existing Approaches (both measures and graphical assessment)
        \item Development of new measure and graphical method to assess trending for measurements, nowcasts and forecasts
        \item Application to various data examples from practice
        \item Publish ready-to-use code
    \end{itemize}
    \item Überblick über Paper
\end{itemize}

### 2024-02-02 sections/03-trending.tex
Update on Overleaf.

\begin{itemize}
    \item Notation klären (hier oder oben?)
    \item Trending mathematisch $P(x y) > 0$ / Statistical consistency of $sgn(x)$ mit $sgn(y)$
    \item Existierende Ansätze aus Literatur: Accuracy, 4Q-Plot (inklusive Agreement-Maße, Kreuztabellen, Measurement Agreement)
    \item Kleine Beobachtungen interessieren uns weniger als große; verrauschen vor allem Ergebnis
    \item Daten mit Rauschen: Man kann kleine Beobachtungen (zu klären, was das genau heißt) ausschließen: man schätzt $P(XY > 0 | \abs (X) > \epsilon$ $\rightarrow$ kleiner Exkurs, warum Gewichtung schwierig zu interpretieren ist?
    \item Grafische Evaluation von $P(XY > 0 | X)$ mit bedingter KDE
    \begin{itemize}
        \item Vorteil: Im Gegensatz zum 4Q-Plot kann man mehrere Nowcasts in einem Schaubild vergleichen
        \item Vorteil: Ermöglicht Analyse von beliebigen Bereichen separat, im Gegensatz zu festgeschriebenen Maßen
        \item Vorteil: man macht Asymmetrien und lokales Verhalten sichtbar 
        \item Begründen, warum andere "klassische" Maße nicht funktionieren
    \end{itemize}
    \item Trending als rolling estimate über die Zeit
    \item Was ist mit missing values?
\end{itemize}

Zu diskutieren:
\begin{itemize}
    \item Konfidenzintervalle
    \item Weitere Maße aus dem Kreuztabellen/dichotomes Forecasting: sensitivity, specitivity, ...
\end{itemize}

\subsection{Beispiele (simple, illustrative Beispiele)}

### 2024-02-02 sections/04-application.tex
Update on Overleaf.

\begin{itemize}
    \item Drei Anwendungen, jeweils eine aus den Bereichen Messwerte, Nowcasting, Forecasting
    \item Messwerte: Daten Medizinmesswerte
    \item Nowcasting: Covid-Nowcasting-Hub
    \begin{itemize}
        \item Daten sind schon aufbereitet
        \item Vergleich zwischen Nowcastern möglich
        \item Noch zu klären: Wie wird $\xdiff$ gewählt?
    \end{itemize}
    \item Forecasting: Pandemie-/Epidemievorhersage, Medikamentenbedarf
\end{itemize}

\subsection{Covid nowcasting}

Data from Covid-Nowcasting-Hub

Besprechen

\begin{itemize}
    \item Was ist x-calligrafisch? aktuell betrachteter Nowcast oder "bester" Nowcast? Aktuell betrachteter Nowcast (keine "Auswahl" nötig mit Maß, Trainingsdatensatz, Rolling Fit, ...)
    \item Welche Lags werden betrachtet? Tage, Woche(n)
    \item Wie geht man mit missing values um? Vor allem auch für missing values für manche issued-times, wenn es aber andere issed-times gibt, für die Werte vorliegen?
    \begin{itemize}
        \item Imputation (2. Wahl)
        \item Nachgereichte Werte mit ursprünglichen Daten (1. Wahl)
        \item Nachgereichte Werte mit späteren Daten $\rightarrow$ nicht benutzen
    \end{itemize}
\end{itemize}

\subsection{Forecasting emergeny department arrivals}

\textcite{Rostami-Tabar2023}

Paper setup:

\begin{itemize}
    \item (Probabilistic) forecast of the hourly number of arrivals in large emergency department (take expectation as point forecast)
    \item Forecasts are produced 48 hours ahead on a 12-hour-rolling basis; background: \enquote{It enables planners to match ED staff to the number of arrivals, redeploy staff, and reconfigure units.}
    \item training data: 2014-04-01 to 2018-2-28
    \item evaluation from 2018-03-01 to 2019-02-28
    \item Evaluation in Paper through RMSE; Pinball Loss and PIT-histograms
    \item consider several lags:
    \begin{itemize}
        \item 72-hours-lag: Is the forecast showing the correct change compared to the most recent observation (problem: strong weekly structure (i.e., Monday and Saturday highest) makes it easy to predict right direction)
        \item 7-days-lag: Correct trend compared to the last shift of same hour and day $\rightarrow$ change policies compared to that
        \item Take first issued forecast for every target time
    \end{itemize}
\end{itemize}

### 2024-02-09 sections/02-notation.tex
Update on Overleaf.

The trending detection of time series, as motivated in the introduction, is fundamentally interesting for the evaluation and comparison of methods from three areas: measurement technology, forecasting and nowcasting. This paper relates to numerical data.\\\\
Measurement aims to obtain accurate and reliable data about the current state of a system. Often the question is how to compare one method of measurement of a variable with another. For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard) for the same variable in the same period of time. The measured values for a measurement method are then available in a time series form, e.g. can be noted with $(x_t)^T_{t=0}$. To detect the systematic difference between two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero\cite{watson2010method}. In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (cf. \cite{lawrence1989concordance,koo2016guideline,}). Furthermore, graphical tools were developed, such as the
Bland-Altman diagram, which visualises the differences between the methods in relation to their mean value (cf. \cite{bland1986statistical}). \\\\
Forecasting is focused on predicting the future based on historical data and its pattern. We note in the following the real realisations of the target value with $(\mathbf{y})^T_{t=0}$ and the prediction data for a target variable  can be organised in such a way that $x_{t,\tau}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$. In practice, a forecasting method is typically evaluated by a forecast accuracy, which is defined by a certain loss function and evaluates the difference between the forecast and the actual outcome of the target variable. A comparison of several forecasters can then be made, for example, by ranking the forecast accuracies of the forecasters. In \cite{hyndman2006another} an overview of various forecasting accuracies used in practice, such as scale-dependent measures like Root Mean Square Error (RMSE) or measures based on percentage errors like Root Mean Square Percentage Error (RMSPE), is presented and discussed. \\\\
Nowcasting (also short-Term-Forecasting), on the other hand, has its origins in meteorology and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (cf. \cite{browning1989nowcasting,schmid2019nowcasting}). Nowcasting focuses on predictions for the present, the immediate future and the recent past, and is now being used in other fields, such as the economy, where important statistics on the current economic situation are available only after a considerable time lag (cf. \cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioural factors (cf. \cite{wu2021nowcasting}). Applications of nowcasting methods for econometrics can be found here in \cite{giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic} or for predictions of epidemic development here in \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}. When predicting values in the present, it is usual that the true values of the recent past are not available, so that only an estimate of these values is provided by the nowcast. Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exiertits for $\tau$ less than or equal $t$. Similar to forecasting, nowcasters are often evaluated and  compared in the literature on the basis of performance measures such as the root mean absolute error (cf. \cite{gunther2021nowcasting}).\\\\
All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods takes into account the local absolute differences in different ways, but does not directly consider the influence of the sign of the differences. To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or prediction with continuous outcomes. Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see chapter \ref{cap: trending}). In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting or nowcasting methods are considered.\\\\
For measurement:
\[\diffxl = (x_{t+l} -x_{t})^T_{t=0}  \]
For forecasting: 
\[\diffxl = (x_{t,t+l} -y_{t})^T_{t=0}  \]
For nowcasting:
\[
\diffxl = 
\begin{cases} 
(x_{t,t+l} -x_{t,t})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
(x_{t,t+l} -y_{t})^T_{t=0}  & \text{else}.
\end{cases}
    \]

%To account for this, we note the nowcast value at time $t$ of the quantity of interest of time $\tau$ as $x^{t}_{\tau}$ (where $\tau$ can be less than $t$)

%Data preparation is similar for measurements and predictions. Let $\mathbf{x} = (x_t)^T_{t=1}$ and $\mathbf{y}=(y_t)^T_{t=1}$ be two time series to be compared (e.g. $\mathbf{x}$, $\mathbf{y}$ are two measurements with two different measuring devices/ measurement methods at the same points in time or predicted values of two predictors in the same time period).

%For the following analysis, the changes between points in time with the same time distance $l\in \mathbb{N}^+ $ in the time series of $\mathbf{x}$ are considered (the same applies for $\mathbf{y}$):

%\[\Delta^lx_t = x_{t+l} - x_{t} \text{ for } t+l\leq T. \]\\
%, then the serie of changes $\mathfrak{X}$, which is later used to assess trendability, is defined as:

%\[\Delta^l \mathfrak{X} _t = x^{t}_{t+l} - x^t_{t} \text{ for } t+l\leq T.\]
%If the true value $y_t$ were known at time $t$, then the following applies: $x^t_t = y_t$.

%\begin{itemize}
%    \item Was ist nowcasting?
%    \item Aus welchem Anwendungsbereichen kommt es? 
%    \item In welchen Bereichen in der Medizin spielt es eine Rolle?
%    \item Warum ist Nowcasting anders als Forecasting/Measurement?
%    \item Wo spielt Trending in der Literatur schon eine Rolle?
%\end{itemize}

### 2024-02-09 sections/04-application.tex
Add first plots for ed arrival application
Update on Overleaf.

@@
 
-Besprechen
+\begin{figure}
+    \centering
+    \includegraphics{plots/covid_nowcast/00_true_data.pdf}
+    \caption{The seven-day-hospitalization of Covid in Germany as reported by the RKI \hl{Add citation}.}
+    \label{fig:app-covid-true}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
+    \caption{Nowcast data without removal of outliers}
+    \label{fig:app-covid-nowcast}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics{plots/covid_nowcast/01_nowcast_data_without_outliers.pdf}
+    \caption{Nowcast data after removal of outliers}
+    \label{fig:app-covid-nowcast-no-outliers}
+\end{figure}
+
+\begin{table}[]
+    \centering
+    \input{plots/covid_nowcast/01_model_scores}
+    \caption{Point evaluation measures for the issued mean of the different models. Note that not for all dates all models issue nowcasts.}
+    \label{tab:app-covid-rmse}
+\end{table}
 
 \begin{itemize}
-    \item Was ist x-calligrafisch? aktuell betrachteter Nowcast oder "bester" Nowcast? Aktuell betrachteter Nowcast (keine "Auswahl" nötig mit Maß, Trainingsdatensatz, Rolling Fit, ...)
-    \item Welche Lags werden betrachtet? Tage, Woche(n)
-    \item Wie geht man mit missing values um? Vor allem auch für missing values für manche issued-times, wenn es aber andere issed-times gibt, für die Werte vorliegen?
-    \begin{itemize}
-        \item Imputation (2. Wahl)
-        \item Nachgereichte Werte mit ursprünglichen Daten (1. Wahl)
-        \item Nachgereichte Werte mit späteren Daten $\rightarrow$ nicht benutzen
+    \item Seven-day hospitalization rate as central steering measure for COVID measures in Germany
+    \item Describe how it is computed
+    \item Outline measures and sources to find methods
+    \item Present overall measures for methods (Table~\ref{tab:app-covid-rmse})
+    \item Present why trending is important, particularly the differentiation between increase and decrease
+    \item Use only a subset of all methods for trending evaluation (best methods)
+    \item Present trending of the subset
 \end{itemize}
-\end{itemize}
-
-\subsection{Forecasting emergeny department arrivals}
+\subsection{Forecasting emergency department arrivals}
 
@@
 \end{itemize}
+
+\begin{table}
+\centering
+\input{plots/ed_arrival/accuracies.tex}
+\caption{Accuracies (without exclusion area) and RMSE for the considered models in \textcite{Rostami-Tabar2023}}	
+\end{table}
+
+\begin{figure}
+	\centering
+	\includegraphics{plots/ed_arrival/Cond_Prob_lag_3.pdf}
+        \caption{Conditional probability plot for ED arrival}
+\end{figure}

### 2024-02-12 sections/02-notation.tex
Add labels to equations; add paragraph breaks

@@
 The trending detection of time series, as motivated in the introduction, is fundamentally interesting for the evaluation and comparison of methods from three areas: measurement technology, forecasting and nowcasting.
-This paper relates to numerical data.\\\\
+This paper relates to numerical data.
+
 Measurement aims to obtain accurate and reliable data about the current state of a system.
@@
 Furthermore, graphical tools were developed, such as the
-Bland-Altman diagram, which visualises the differences between the methods in relation to their mean value (cf. \cite{bland1986statistical}). \\\\
+Bland-Altman diagram, which visualises the differences between the methods in relation to their mean value (cf. \cite{bland1986statistical}).
+
 Forecasting is focused on predicting the future based on historical data and its pattern.
@@
 In \cite{hyndman2006another} an overview of various forecasting accuracies used in practice, such as scale-dependent measures like Root Mean Square Error (RMSE) or measures based on percentage errors like Root Mean Square Percentage Error (RMSPE), is presented and discussed.
-\\\\
+
 Nowcasting (also short-Term-Forecasting), on the other hand, has its origins in meteorology and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (cf.
@@
 Similar to forecasting, nowcasters are often evaluated and  compared in the literature on the basis of performance measures such as the root mean absolute error (cf.
-\cite{gunther2021nowcasting}).\\\\
+\cite{gunther2021nowcasting}).
+
 All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods takes into account the local absolute differences in different ways, but does not directly consider the influence of the sign of the differences.
@@
 Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see chapter \ref{cap: trending}).
-In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting or nowcasting methods are considered.\\\\
+In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting or nowcasting methods are considered.
+
 For measurement:
-\[\diffxl = (x_{t+l} -x_{t})^T_{t=0}  \]
+\begin{equation}
+	\diffxl = (x_{t+l} -x_{t})^T_{t=0} \label{eq:diffxl_measure}
+\end{equation}
 For forecasting: 
-\[\diffxl = (x_{t,t+l} -y_{t})^T_{t=0}  \]
+\begin{equation}
+	\diffxl = (x_{t,t+l} -y_{t})^T_{t=0} \label{eq:diffxl_forecasting}
+\end{equation}
 For nowcasting:
-\[
+\begin{equation}
 \diffxl = 
@@
 (x_{t,t+l} -y_{t})^T_{t=0}  & \text{else}.
-\end{cases}
-    \]
+\end{cases} \label{eq:diffxl_nowcasting}
+\end{equation}
+
+\todo[inline]{Formeln als Tabelle oder in den Text einbinden.}
 

### 2024-02-12 paper.tex
Update on Overleaf.
Add first plots for ed arrival application
Change labels and add math notation

@@
 \usepackage{tabularx}
-\usepackage{tabularray}
+%\usepackage{tabularray}
 
@@
 
-\newcommand{\lag}[1][l]{\Delta_{#1}}
 \DeclarePairedDelimiter{\abs}\lvert\rvert
@@
 \newcommand{\ind}[1]{\mathbbm{1}\{#1\}}
-\newcommand{\ydiff}{D y}
-\newcommand{\ydifft}{Dy^\star}
-\newcommand{\xdiff}{Dx}
-\newcommand{\xdifft}{Dx^\star}
-\newcommand{\Ydiff}{DY}
-\newcommand{\Xdiff}{DX}
 \newcommand{\Prob}[1]{P(#1)}
-\newcommand{\mprob}{\tilde{m}}
-\newcommand{\Ber}{\text{Ber}}
+\newcommand{\cond}{\:\lvert\:}
+
+\newcommand{\R}{\mathbb{R}}
 \newcommand{\SBer}{\text{SBer}}
-\newcommand{\cond}{\:\lvert\:}
+
+\newcommand{\diffxl}{\mathbf{x}^{\Delta,l}}
+\newcommand{\diffyl}{\mathbf{y}^{\Delta,l}}
+\newcommand{\diffx}{\mathbf{x}^{\Delta}}
+\newcommand{\diffy}{\mathbf{y}^{\Delta}}
+\newcommand{\diffxlt}{\mathbf{x}^{\Delta,l}_t}
+\newcommand{\diffylt}{\mathbf{y}^{\Delta,l}_t}
+\newcommand{\diffxt}{x^{\Delta}_t}
+\newcommand{\diffyt}{y^{\Delta}_t}
+
+\newcommand{\acc}{\mu}
+\newcommand{\acceps}{\acc_{\eps}}
+\newcommand{\accpeps}{\acceps^+}
+\newcommand{\accmeps}{\acceps^-}
+\newcommand{\accp}{\acc^+}
+\newcommand{\accm}{\acc^-}
 
@@
 
-\title{Assessment and Optimization of Nowcast Measures for Trend Detection}
+\title{Trending assessment for nowcasts, measurements, and forecasts}
 \author{Oliver Grothe, Bolin Liu, Jonas Rieger}
@@
 
-\section{Introduction}
+Agenda für heute:
+\begin{itemize}
+    \item Aufteilung
+    \begin{itemize}
+        \item Introduction (1.5 pages), Conclusion (0.5 pages), Abstract
+        \item Nowcasting (1 pages): Bolin
+        \item Trending (5-6 pages): Jonas fängt an
+        \item Application (5-6 pages): Jonas
+    \end{itemize}
+    \item Datenquellen
+\end{itemize}
 
 \begin{itemize}    
-    \item 
+    \item Journal/Anwendungsgebiet klären: Angewandtes Medizin Statistikjournal? Dann müsste man die folgenden Abschnitte entsprechend ausrichten
 \end{itemize}
 
-\section{Trending assessment}
+\section{Introduction}\label{sec:introduction}
 
-\begin{itemize}
-    \item Es interessiert uns $P(x y) > 0$
-    \item 
-\end{itemize}
+\input{sections/01-introduction}
+
+\section{Application areas and Notation} \label{sec:notation}
+
+\input{sections/02-notation}
+
+\section{Trending assessment}\label{sec:trending}
+
+\input{sections/03-trending}
+
+%\section{Simulation Studies (?)}
+%
+%Eher nicht, weil angewandtes Paper
+
+\section{Application to real-world data} \label{sec:application}
+
+\input{sections/04-application}
+
+\printbibliography
 

### 2024-02-12 sections/02-notation.tex
Change labels and add math notation

@@
 To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or prediction with continuous outcomes.
-Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see chapter \ref{cap: trending}).
+Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see chapter \ref{sec:trending}).
 In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting or nowcasting methods are considered.
@@
 \todo[inline]{Formeln als Tabelle oder in den Text einbinden.}
+\todo[inline]{$\mathbf{y}$ definieren.}
 

### 2024-02-16 sections/02-notation.tex
Add remarks on section

@@
-The trending detection of time series, as motivated in the introduction, is fundamentally interesting for the evaluation and comparison of methods from three areas: measurement technology, forecasting and nowcasting.
-This paper relates to numerical data.
+The trending detection of time series, as motivated in the introduction, is fundamentally interesting for evaluating and comparing methods from three areas: measurement technology, forecasting, and nowcasting.
+This paper relates to numerical data. \todo{Hängt irgendwie in der Luft}
+\todo{Überleitung, dass jetzt die einzelnen vorgestellt werden}
 
 Measurement aims to obtain accurate and reliable data about the current state of a system.
-Often the question is how to compare one method of measurement of a variable with another.
+Often, the question is how to compare one method of measurement of a variable with another.
 For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard) for the same variable in the same period of time.
-The measured values for a measurement method are then available in a time series form, e.g.
-can be noted with $(x_t)^T_{t=0}$.
-To detect the systematic difference between two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero\cite{watson2010method}.
+The measured values for a measurement method are then available in a time series form, e.g., can be noted with $(x_t)^T_{t=0}$. \todo{Fehlt da nicht noch $y$?}
+To detect the systematic difference between the two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero\cite{watson2010method}.
 In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (cf.
 \cite{lawrence1989concordance,koo2016guideline,}).
-Furthermore, graphical tools were developed, such as the
-Bland-Altman diagram, which visualises the differences between the methods in relation to their mean value (cf. \cite{bland1986statistical}).
+\todo{Andere Zitierfunktion}
+Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizes the differences between the methods in relation to their mean value (cf. \cite{bland1986statistical}).
 
+\todo[inline]{Überleitung}
 Forecasting is focused on predicting the future based on historical data and its pattern.
-We note in the following the real realisations of the target value with $(\mathbf{y})^T_{t=0}$ and the prediction data for a target variable  can be organised in such a way that $x_{t,\tau}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$.
-In practice, a forecasting method is typically evaluated by a forecast accuracy, which is defined by a certain loss function and evaluates the difference between the forecast and the actual outcome of the target variable.
+We note in the following the real realizations of the target value with $(\mathbf{y})^T_{t=0}$ and the prediction data for a target variable can be organized in such a way that $x_{t,\tau}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$.
+\todo{Motivieren, warum beide Indizes nötig}
+In practice, a forecasting method is typically evaluated by forecast accuracy, which is defined by a certain loss function and evaluates the difference between the forecast and the actual outcome of the target variable.
 A comparison of several forecasters can then be made, for example, by ranking the forecast accuracies of the forecasters.
-In \cite{hyndman2006another} an overview of various forecasting accuracies used in practice, such as scale-dependent measures like Root Mean Square Error (RMSE) or measures based on percentage errors like Root Mean Square Percentage Error (RMSPE), is presented and discussed.
+In \cite{hyndman2006another}, an overview of various forecasting accuracies used in practice, such as scale-dependent measures like Root Mean Square Error (RMSE) or measures based on percentage errors like Root Mean Square Percentage Error (RMSPE), is presented and discussed.
+\todo{Herausstellen, dass bisherige Maße auf differenzen zwischen forecast und true herauslaufen; Hyndman eher als Quelle am Ende, sonst erwartet man, dass da noch andere Quellen besprochen werden}
 
-Nowcasting (also short-Term-Forecasting), on the other hand, has its origins in meteorology and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (cf.
+\todo[inline]{Überleitung}
+Nowcasting (also short-term forecasting), on the other hand, has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (cf.
 \cite{browning1989nowcasting,schmid2019nowcasting}).
-Nowcasting focuses on predictions for the present, the immediate future and the recent past, and is now being used in other fields, such as the economy, where important statistics on the current economic situation are available only after a considerable time lag (cf.
-\cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioural factors (cf.
+Nowcasting focuses on predictions for the present, the immediate future, and the recent past and is now being used in other fields, such as the economy\todo{economics?}, where important statistics on the current economic situation are available only after a considerable time lag (cf.
+\cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioral factors (cf.
 \cite{wu2021nowcasting}).
+\todo{sehr langer Satz}
 Applications of nowcasting methods for econometrics can be found here in \cite{giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic} or for predictions of epidemic development here in \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}.
-When predicting values in the present, it is usual that the true values of the recent past are not available, so that only an estimate of these values is provided by the nowcast.
-Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exiertits for $\tau$ less than or equal $t$.
-Similar to forecasting, nowcasters are often evaluated and  compared in the literature on the basis of performance measures such as the root mean absolute error (cf.
-\cite{gunther2021nowcasting}).
+\todo{Satz klingt noch nicht so ganz rund}
+When predicting values in the present, it is usual that the true values of the recent past are not available, so only an estimate of these values is provided by the nowcast.\todo{Vielleicht noch etwas konkreter machen}
+Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exiertits \todo{word?} for $\tau$ less than or equal to $t$.
+Similar to forecasting, nowcasters are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (cf. \cite{gunther2021nowcasting}).
 
-All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods takes into account the local absolute differences in different ways, but does not directly consider the influence of the sign of the differences.
-To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or prediction with continuous outcomes.
-Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see chapter \ref{sec:trending}).
-In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting or nowcasting methods are considered.
+All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods take into account the local absolute differences in different ways but do not directly consider the influence of the sign of the differences.
+\todo{vielleicht: "do not directly consider whether the correct direction of change is issued"}
+To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or predictions with continuous outcomes.
+Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see Chapter~\ref{sec:trending}).
+\todo{Vielleicht nochmal Praxisbezug herausstellen, warum es sinnvoll ist, die Differenzen so zu formulieren?}
+In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting, or nowcasting methods are considered.
 

### 2024-02-16 paper.tex
Add draft for Section 3

@@
 %\usepackage[hidelinks]{hyperref}
-\usepackage[style=authoryear,backend=bibtex]{biblatex}
+\usepackage[style=authoryear,backend=biber]{biblatex}
 \addbibresource{library.bib}
+\addbibresource{library-jonas.bib}
 \usepackage{csquotes}
@@
 
-\usepackage{xargs} % Use optional arguements in commands
+\usepackage{xargs,ifthen} % Use optional arguements in commands
 \newcommandx{\unsure}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
@@
 \let\marginpar\marginnote
+
+%%% --- Acronyms
+\usepackage{acro}
+\DeclareAcronym{kde}{short=KDE, long=kernel density estimation}
+\DeclareAcronym{bca}{short=BCa, long=bias-corrected and accelerated}
 
@@
 \newcommand{\cond}{\:\lvert\:}
+\newcommand{\card}[1]{\abs{#1}}
 
@@
 \newcommand{\diffy}{\mathbf{y}^{\Delta}}
+\newcommand{\diffxrv}{X^{\Delta}}
+\newcommand{\diffyrv}{Y^{\Delta}}
 \newcommand{\diffxlt}{\mathbf{x}^{\Delta,l}_t}
@@
 \newcommand{\acc}{\mu}
-\newcommand{\acceps}{\acc_{\eps}}
-\newcommand{\accpeps}{\acceps^+}
-\newcommand{\accmeps}{\acceps^-}
 \newcommand{\accp}{\acc^+}
 \newcommand{\accm}{\acc^-}
+%\newcommand{\acceps}[1][]{\ifthenelse { \equal {#1} {} }{\acc_{\eps}}{\acc_{#1}}} %https://stackoverflow.com/a/7314951
+\newcommand{\acceps}[1][\varepsilon]{\acc_{#1}} %https://stackoverflow.com/a/7314951
+\newcommand{\accpeps}[1][\varepsilon]{\acceps[#1]^+}
+\newcommand{\accmeps}[1][\varepsilon]{\acceps[#1]^-}
 
@@
 \section{Introduction}\label{sec:introduction}
-
 \input{sections/01-introduction}
@@
 \section{Application areas and Notation} \label{sec:notation}
-
 \input{sections/02-notation}
@@
 \section{Trending assessment}\label{sec:trending}
-
 \input{sections/03-trending}
@@
 \section{Application to real-world data} \label{sec:application}
-
 \input{sections/04-application}
@@
 
+\appendix
+\section{Data generation for Section~\ref{sec:trending}}\label{sec:app-trending-data-generation}
+
+\todo{Write}
+
+Beispiel
+\begin{itemize}
+  \item $\diffyt \sim U(-5, 5) + N(0, 1)$
+  \item $\diffxt \sim \diffyt \cdot \SBer(k) \cdot (\text{Truncated Normal, mean 1, sigma 0.5, trunc 0})$
+\end{itemize}
+
 \end{document}

### 2024-02-16 sections/02-notation.tex
Add draft for Section 3

@@
 To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or predictions with continuous outcomes.
-Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see Chapter~\ref{sec:trending}).
+Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see Section~\ref{sec:trending}).
 \todo{Vielleicht nochmal Praxisbezug herausstellen, warum es sinnvoll ist, die Differenzen so zu formulieren?}

### 2024-02-16 sections/03-trending.tex
Update on Overleaf.
Add draft for Section 3

@@
+As outlined in Section~\ref{sec:introduction}, we refer by trending to the consistency of an observed change with the change issued by another measurement, a nowcast or a forecast.
+In this section, we want to specify trending and lay out different methods of measuring and visualizing trending. 
+In general, trending can be assessed for different lags $l$ separately, and the relevance of considered lags has to be determined based on the question at hand.
+In Section~\ref{sec:notation}, the time series of differences $\diffx$ and $\diffy$ were introduced.
+Whereas the computation of $\diffx$ differs in the different contexts, the trending assessment is similar.
+In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$.
+Similarly, we call a measurement, nowcast, or forecast a \textit{prediction} for the quantity of interest $y$. \unsure{Discuss}
+The remainder of the section is structured as follows.
+Section~\ref{subsec:trending-basics} introduces trending basics and visualizes trending in a synthetical example.
+Section~\ref{subsec:trending-measures} reviews conveyed quantification approaches from other fields.
+The approaches are extended to the trending setting in Section~\ref{subsec:trending-noise}.
+Section~\ref{subsec:trending-cond-prob} introduces a new graphical method of trending assessment.
+Section~\ref{subsec:trending-bootstrap} gives bootstrap methods for computing confidence intervals for the measures and conductes a simulation study on their properties.
+
+\subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-basics}
+
+As noted above, a prediction is trending perfectly if all occurred changes are predicted correctly, that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+Perfect trending is hardly the case in practice, but a weaker requirement is needed.
+Accordingly, when assessing the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
+A simple yet insightful method is the four-quadrant plot.
+For example, it is known in cardiac output measurement analysis~\parencite{Saugel2015}. \todo{Add other citations}
+Thereby, the occurred changes are plotted together with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
+Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays prediction data differences.
+Points in the green upper right and lower left quadrant reflect a correct trending for the respective time step, whereas point in the remaining red quadrants show incorrect predicted changes.
+Figure~\ref{fig:trending_basic_4q} displays a basic four-quadrant plot.
+Points 2, 3, 5, and 6 show trending, whereas points 1, 4, and 7 count for anti-trending behavior.
+Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant plot for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
+The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
+The four-quadrant plot is intuitive to interpret, and the magnitude and direction of change are shown simultaneously.
+It can be extended by including information on the date of difference in the point color.
+In Figure~\ref{fig:trending_basic_4q_sample_color}, blue points refer to small time indices and turn green for larger $t$.
+However, four-quadrant plots become crowded for larger datasets, and the sequential information on the differences is complex to assess thoroughly.
+There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.
+However, they lack the clearness and intuition of the four-quadrant plot while adding no more information on the trending \parencite{Saugel2015}.
+
+\begin{figure}
+\centering
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4Q_without_excl}
+\caption{Basic four-quadrant plot.} \label{fig:trending_basic_4q}
+\end{subfigure}\hspace{0.01\textwidth}
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4q_excl_box}
+\caption{Four-quadrant plot with rectangle exclusion area.}\label{fig:trending_basic_4q_excl_box}
+\end{subfigure}\hspace{0.01\textwidth}
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4q_excl_axis}
+\caption{Four-quadrant plot with horizontal exclusion area.} \label{fig:trending_basic_4q_excl_axis}
+\end{subfigure}\hspace{0.01\textwidth}
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4q_excl_cross}
+\caption{Four-quadrant plot with crossed exclusion area.}\label{fig:trending_basic_4q_excl_cross}
+\end{subfigure}%
+\caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas. }
+\label{fig:trending_4q}
+\end{figure}
+
+\subsection{Trending ratio and other measures}\label{subsec:trending-measures}
+
+Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$.
+Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R$).
+To estimate $P(\diffxrv \diffyrv > 0)$ it lies at hand to use
+\begin{equation}
+    \acc (\diffx, \diffy) \coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0}}{\card{T}}.\label{eq:acc}
+\end{equation}
+We refer to this estimator as the trending ratio of the prediction and set $\mathcal{T} = \{1, \dots, T\}$.
+Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
+Such a $2 \times 2$ table of counts is often called a contingency table and is often used in other scientific areas for evaluation, for example, in dichotomous forecasting or as a confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
+4]{James2021}[Ch.
+3]{Jolliffe2012}.
+There, a wide range of other methods is usually used to analyze further characteristics of contingency tables.
+Two simple measures focusing on specific areas of interest are the positive and negative trending ratio $\accp$ and $\accm$, respectively.
+They are defined as
+\begin{align}
+    \accp (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0}} \label{eq:accp}\\
+    \accm (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0}}\label{eq:accm}
+\end{align}
+In the classification context, these measures are known as positive or negative predictive value and hit rate or detection failure ratio in forecasting.
+The two give estimates of $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$, respectively.
+There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \todo{Citation}.
+Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.
+Unbalanced outcomes of the differences are unlikely in our setting as they are obtained from differencing time series data.
+Nevertheless, if in a four-quadrant plot the number of positive and negative $\diffy$ are observed to differ widely, one should consider another as measures such as Cohen's $\kappa$ or the ones listed in \textcite[Table 3.3]{Jolliffe2012}.
+
+All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
+Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color}.
+Thus, the yearly course of the prediction's trending ratio can be detected.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4Q_sample_without_time}
+\caption{Four-quadrant plot with simulated data.}\label{fig:trending_basic_4q_sample}
+\end{subfigure}%
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4Q_sample_with_time}
+\caption{The data is colored according to the time index $t$, the greener, the later.}\label{fig:trending_basic_4q_sample_color}
+\end{subfigure}%
+\begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/illustrative_examples/trending_ratio_time_series.pdf}
+    \caption{Rolling estimate of the trending ratio over time with window length 91.
+A strong yearly seasonality of the trending ratio becomes visible.
+The overall trending ratios are $\mu = 0.7577$ and with exclusion area $\diffx < \varepsilon$, $\mu_{1.0} = 0.7712$, respectively.
+The green curve $k_t$ shows the probability of $\diffxt$ having the same sign as $\diffyt$ for each time step.
+The rolling estimates hang back as the windows are backward-looking.}\label{fig:trending_ratio_time_series}
+    \end{subfigure}%
+    \caption{We defer information on the data generation process of the data in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{sec:app-trending-data-generation})}
+\end{figure}
+
+Missing values are a common problem, in particular in nowcasting and forecasting.
+This will, for example, become evident in the data study in Section~\ref{sec:application-covid}.
+Data pairs with missing values can be excluded in the computation of the measures.
+Nevertheless, missing values can considerably influence the results, mainly if they appear systematically.
+Predicting only in those times of certainty and omitting prediction in other times would generally improve evaluation measures.
+So, although the estimates can be computed, the time spans with missing values should be inspected visually.
+
+\subsection{Accounting for noise}\label{subsec:trending-noise}
+
+The above measures use only the information on a point's quadrant and neglect further information on its location within the quadrant, as displayed in a four-quadrant plot.
+However, points close to the zero points have intuitively less explanatory power.
+If there is noise or non-systematic effects present in true value or nowcasts, those points are likely to be driven by noise to a large amount.
+Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the above methods.
+Points within that zone are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
+The exclusion area's shape can be chosen according to the noise characteristics in predictions and true values.
+In general, it is likely that the models of predictions have a noise component and should thus be part of the exclusion area.
+We denote the measures accounting for an exclusion area by
+\begin{align}
+    \acceps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{\abs{\diffxt} > \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\abs{\diffxt} > \varepsilon}}\label{eq:acceps}\\
+    \accpeps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > \varepsilon}} \label{eq:accpeps}\\
+    \accmeps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < \varepsilon }}\label{eq:accmeps}
+\end{align}
+The estimators can easily be adapted for other shapes of the exclusion area.
+Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
+A box-shaped exclusion area leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise. \todo{Umformulieren}
+An exclusion area along one axis removes points in which one of the components could change the sign by a small amount of noise.
+This particularly suits prediction models, where small amounts of noise are inevitable.
+A cross-shaped exclusion area along both axes accounts for sign-reversal in both components.
+
+In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
+For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point.
+These shapes are theoretically appealing, but the interpretation of the resulting estimators gets ambiguous.
+Thus, we use the simple exclusion areas above.
+
+In most applications, it is advisable to determine the exact shape and size of the exclusion area based on domain knowledge or expert opinions.
+In addition, the size can also be calculated as a proportion of the total variance or the total range of the data.
+A third approach is to visualize the trending ratio for different values of $\varepsilon$.
+Section~\ref{subsec:trending-example} gives an example of such a plot.
+Instead of determining the exclusion area size beforehand, the plot can be used to choose the appropriate $\varepsilon$.
+\todo[inline]{Abschnitt unelegant}
+
+As a base approach, we advocate using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$\varepsilon$ plot to visualize the course of the trending ratio over different exclusion area sizes.
+
+\subsection{Conditional trending plot}\label{subsec:trending-cond-prob}
+
+The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > \varepsilon)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < \varepsilon)$.
+To assess the trending ability of a prediction, an evaluation on a finder grid is of interest.
+Whereas it would be possible to build analog measures on other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
+In addition, the conditional trending plot facilitates the comparison of various methods in a single plot, and asymmetries in the trending ability can be detected.
+
+The probability $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ cannot be computed by the counting approach, as for almost all $x \in \R$, no data is available.
+Instead, the smoothed approach of a multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$.
+A comprehensive introduction into multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages~\parencite[e.g., for  Python in][]{Seabold2010}.
+The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ for all values of $x \in \R$.
+For kernels with infinite support, for example, the Gaussian kernel, this can convey a false impression of validity for values far beyond the values in $\diffx$.
+Thus, the plot should be limited to the core values of $\diffx$ without outliers.
+
+Figure~\ref{fig:trending_cond_prob} provides two sample conditional trending plots.
+\todo[inline]{Add evaluation of the plot}
+Further examples, including comparisons between methods concerning their trending ability, are available for the applications in Section~\ref{sec:application}.
+
+\begin{figure}
+    \centering
+    \missingfigure{Trending plot for two different simulation examples -> the two data samples of bootstrapping?}
+%    \includegraphics{}
+    \caption{}
+    \label{fig:trending_cond_prob}
+\end{figure}
+
+\subsection{Confidence intervals based on bootstrapping}\label{subsec:trending-bootstrap}
+
+Confidence intervals can be issued to account for the estimation uncertainty of the measures above.
+Bootstrapping confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
+Whereas classical confidence intervals are computed from parametric assumptions on the underlying data, new samples are drawn with replacement from the dataset in bootstrapping.
+The confidence interval is then computed based on the bootstrap samples.
+We examine three methods for bootstrapping here: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
+In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the bootstrap estimators' empirical distribution.
+The \textit{basic} approach computes the confidence interval based on the non-bootstrapped estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
+The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
+Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but needs smaller datasets for reasonable confidence intervals.
+The basic approach balances those two aims.
+We compare the approaches in a small synthetic data study in the following with respect to their small-dataset behavior and their computation time.
+
+In bootstrapping, the sample is assumed to be independent and identically distributed.
+Thus, before applying bootstrapping methods, the strength of sequential dependence should be inspected, for example, by analyzing the autocorrelation and partial autocorrelation.
+Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
+
 \begin{itemize}
-    \item Notation klären (hier oder oben?)
-    \item Trending mathematisch $P(x y) > 0$ / Statistical consistency of $sgn(x)$ mit $sgn(y)$
-    \item Existierende Ansätze aus Literatur: Accuracy, 4Q-Plot (inklusive Agreement-Maße, Kreuztabellen, Measurement Agreement)
-    \item Kleine Beobachtungen interessieren uns weniger als große; verrauschen vor allem Ergebnis
-    \item Daten mit Rauschen: Man kann kleine Beobachtungen (zu klären, was das genau heißt) ausschließen: man schätzt $P(XY > 0 | \abs (X) > \epsilon$ $\rightarrow$ kleiner Exkurs, warum Gewichtung schwierig zu interpretieren ist?
-    \item Grafische Evaluation von $P(XY > 0 | X)$ mit bedingter KDE
-    \begin{itemize}
-        \item Vorteil: Im Gegensatz zum 4Q-Plot kann man mehrere Nowcasts in einem Schaubild vergleichen
-        \item Vorteil: Ermöglicht Analyse von beliebigen Bereichen separat, im Gegensatz zu festgeschriebenen Maßen
-        \item Vorteil: man macht Asymmetrien und lokales Verhalten sichtbar 
-        \item Begründen, warum andere "klassische" Maße nicht funktionieren
+    \item We use the alogrithms provided by Python's \verb|scipy|-package~\parencite{Virtanen2020}.
+    \item Two simulation setups
+    \item Take different $T$
+    \item Compute "true value" based on very large sample.
+    \item Compute share of bootstrap confidence intervals that actually covered the true value for different $B$.
+    \item Results in "double table": columns $T$, main rows $B$, undercategory: bootstrap methods.
 \end{itemize}
-    \item Trending als rolling estimate über die Zeit
-    \item Was ist mit missing values?
-\end{itemize}
-
-Zu diskutieren:
-\begin{itemize}
-    \item Konfidenzintervalle
-    \item Weitere Maße aus dem Kreuztabellen/dichotomes Forecasting: sensitivity, specitivity, ...
-\end{itemize}
-
-\subsection{Beispiele (simple, illustrative Beispiele)}

### 2024-02-18 sections/03-trending.tex
Small corrections
Update 18.2.2024

@@
 Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant plot for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
+The data generation is described in Appendix~\ref{sec:app-trending-data-generation} and will carry through this section.
 The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
@@
 There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.
-However, they lack the clearness and intuition of the four-quadrant plot while adding no more information on the trending \parencite{Saugel2015}.
+However, they lack the clearness and intuition of the four-quadrant plot while adding no more information on the trending~\parencite{Saugel2015}.
 
@@
 \caption{Basic four-quadrant plot.} \label{fig:trending_basic_4q}
-\end{subfigure}\hspace{0.01\textwidth}
+\end{subfigure}\hspace{0.01\textwidth}%
 \begin{subfigure}[t]{.24\textwidth}
@@
 \caption{Four-quadrant plot with rectangle exclusion area.}\label{fig:trending_basic_4q_excl_box}
-\end{subfigure}\hspace{0.01\textwidth}
+\end{subfigure}\hspace{0.01\textwidth}%
 \begin{subfigure}[t]{.24\textwidth}
@@
 \caption{Four-quadrant plot with horizontal exclusion area.} \label{fig:trending_basic_4q_excl_axis}
-\end{subfigure}\hspace{0.01\textwidth}
+\end{subfigure}\hspace{0.01\textwidth}%
 \begin{subfigure}[t]{.24\textwidth}
@@
 Thus, the yearly course of the prediction's trending ratio can be detected.
+This trade cannot be observed in the two four-quadrant plot as it is overlaid by the high number of points.
 
@@
 Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
-A box-shaped exclusion area leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise. \todo{Umformulieren}
+A box-shaped exclusion area leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise.
+Thus, only point 1 is excluded.
 An exclusion area along one axis removes points in which one of the components could change the sign by a small amount of noise.
@@
 A cross-shaped exclusion area along both axes accounts for sign-reversal in both components.
+For those two methods, points 1 and 7 or 1, 6 and 7 are excluded, respectively.
 
@@
 Thus, the plot should be limited to the core values of $\diffx$ without outliers.
+Besides the kernel, a bandwith selector has to be chosen for a multivariate \ac{kde}.
+Various methods are available with different strengths and weaknesses.
+Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross validation maximum likelihood and cross validation least squares.
+\todo{Add analysis.}
 
-Figure~\ref{fig:trending_cond_prob} provides two sample conditional trending plots.
+Figure~\ref{fig:trending_cond_prob} shows the results for the different snythetic datasets again, but only with the best-balanced cross validation least squares estimator.
 \todo[inline]{Add evaluation of the plot}
 Further examples, including comparisons between methods concerning their trending ability, are available for the applications in Section~\ref{sec:application}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}{.48\textwidth}
+        \missingfigure{Influence of bw method; Data set 1}
+    \end{subfigure}
+    \begin{subfigure}{.48\textwidth}
+        \missingfigure{Influence of bw method; Data set 2}
+    \end{subfigure}
+    \caption{Resulting conditional trending plot for different bandwidth selection processes.}\label{fig:trending-cond-prob-bw}
+\end{figure}
 
@@
 We compare the approaches in a small synthetic data study in the following with respect to their small-dataset behavior and their computation time.
+We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
+The considered datasets are again the ones outlined in Appendix~\ref{sec:app-trending-data-generation}.
+The results in terms of coverage of the true values are shown in Table~\ref{tab:trending_bootstrap}.
+The computation times per method and dataset are shown in Table~\ref{tab:trending_bootstrap_time}.
+
+\begin{table}
+    \centering
+    \missingfigure{Bootstrap results}
+    \caption{}
+    \label{tab:trending_bootstrap}
+\end{table}
+
+\begin{table}
+    \centering
+    \missingfigure{Bootstrap results}
+    \caption{}
+    \label{tab:trending_bootstrap_time}
+\end{table}
 

### 2024-02-18 sections/50-appendix-data-generation.tex
Update 18.2.2024

The first dataset is generated by sequentially generating $\diffx$ and $\diffy$.
First, the $\diffxt$ are sampled as a sum of a standard normal random number and a uniform random number on $(-10, 10)$:
\begin{equation*}
    \diffxt \sim N(0, 1) + U(-10, 10) \quad t = 1, \dots, T.
\end{equation*}
Subsequently, the $\diffy$ are simulated for a constant trending ratio $k$ by
\begin{equation*}
    \diffyt = \diffxt \cdot n_t * b_t,
\end{equation*}
where $n_t$ is a truncated normal distribution with mean 1 and standard deviation 0.5, truncated at 0 and a symmetric Bernoulli random variable with parameter $k$.
For a time-varying trending ratio, the parameter $k$ is modified to have a wave-shape over time, that is,
\begin{equation*}
    k_t = 0.75 + \sin(t / T^\star \cdot 2 \pi) / 4,
\end{equation*}
where $T^\star$ is $T$ divided by the number of oscillations; in this case, 4.
For the asymmetric trending ratio, $k$ is a function of $\diffxt$,
\begin{equation*}
    k(x) = 0.5 + \min \left\{ \max \left\{ \frac{x + 5}{10}, 0  \right\} , 1 \right\} / 2.
\end{equation*}

In the second approach, $\diffyt$ and $\diffxt$ are modelled to be multivariate normal with mean 0 and covariance matrix
\begin{equation*}
    \Sigma = \begin{pmatrix} 4 & 3 \\ 3 & 4 \end{pmatrix}.
\end{equation*}
Thus, the conditional probability of trending can be calculated by a conditional normal distribution to
\begin{equation*}
    P(\diffyrv \diffxrv > 0 | \diffxrv = x) = \Phi \left( \frac{3}{4 \sqrt{7}} x \right),
\end{equation*}
where $\Phi$ is a standard normal \ac{cdf}.

\begin{figure}
    \centering
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1}
        \caption{Constant trending ratio.}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_time}
        \caption{Time-varying trending ratio}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_asym}
        \caption{Asymmetric trending ratio}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp2}
        \caption{Second approach}
    \end{subfigure}
    \caption{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{sec:app-trending-data-generation}. Although the first and second plot differ over time, their difference is not discernible in the plots. The third data set's asymmetry is visible in the plot but the decrese of trending ability near 0 is not visible. }
    \label{fig:appendix_dgps}
\end{figure}

### 2024-02-19 sections/03-trending.tex
Update bootstrapping

@@
 Various methods are available with different strengths and weaknesses.
-Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross validation maximum likelihood and cross validation least squares.
-\todo{Add analysis.}
-
-Figure~\ref{fig:trending_cond_prob} shows the results for the different snythetic datasets again, but only with the best-balanced cross validation least squares estimator.
+Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross validation maximum likelihood and cross validation least squares usigng the \verb|statsmodels| python package~\parencite{Seabold2010}.
+The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
+The second method, cross validation least squares, did not converge for the first dataset after a computation time of approximately 300 times the other two methods.
+The rule of thumb yielded slightly oversmoothed results, while the cross validation maximum likelihood yields rather small bandwiths.
 \todo[inline]{Add evaluation of the plot}
@@
     \begin{subfigure}{.48\textwidth}
-        \missingfigure{Influence of bw method; Data set 1}
+        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_asym_butterfly}
+        \caption{First dataset with asymmetric dependence.}
     \end{subfigure}
     \begin{subfigure}{.48\textwidth}
-        \missingfigure{Influence of bw method; Data set 2}
+        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_normal}
+        \caption{Second dataset. }
     \end{subfigure}
-    \caption{Resulting conditional trending plot for different bandwidth selection processes.}\label{fig:trending-cond-prob-bw}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \missingfigure{Trending plot for two different simulation examples -> the two data samples of bootstrapping?}
-%    \includegraphics{}
-    \caption{}
-    \label{fig:trending_cond_prob}
+    \caption{Resulting conditional trending plot for different bandwidth selection processes.
+Cross validation least squares takes a considerably larger computation time while yielding the results with best balance between smoothing and displaying local effects.
+The rule of thumb is the fastest method but tends to oversmooth.
+Cross validation least squares  }\label{fig:trending-cond-prob-bw}
 \end{figure}
@@
 We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
-The considered datasets are again the ones outlined in Appendix~\ref{sec:app-trending-data-generation}.
-The results in terms of coverage of the true values are shown in Table~\ref{tab:trending_bootstrap}.
-The computation times per method and dataset are shown in Table~\ref{tab:trending_bootstrap_time}.
+The considered datasets are again the ones outlined in Appendix~\ref{sec:app-trending-data-generation}, the first dataset with asymmetric dependence.
+In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals were used~\parencite{Virtanen2020}.
+The prescribed confidence level is 90 \% and the number of bootstrap samples is $10,000$.
+The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
+The true values of the accuracy are computed based on a dataset of size $10^9$, yielding 0.7501 and 0.7700 for the two datasets.
+The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.
+\todo[inline]{Add evaluation when computation is finished}
 
@@
     \centering
-    \missingfigure{Bootstrap results}
-    \caption{}
+    \begin{subtable}{.48\textwidth}
+        \input{plots/illustrative_examples/bootstrap_butterfly_cis}
+        \caption{First dataset}
+    \end{subtable}\hspace{0.02\textwidth}
+    \begin{subtable}{.48\textwidth}
+        \input{plots/illustrative_examples/bootstrap_normal_cis}
+        \caption{Second dataset}
+    \end{subtable}
+    \caption{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$.
+The average width of the confidence interval is listed in brackets.}
     \label{tab:trending_bootstrap}
@@
     \centering
-    \missingfigure{Bootstrap results}
-    \caption{}
-    \label{tab:trending_bootstrap_time}
+    \begin{subfigure}{0.48\textwidth}
+        \includegraphics{plots/illustrative_examples/boxplot_comp_time_butterfly}
+        \caption{First dataset}
+    \end{subfigure}
+    \begin{subfigure}{0.48\textwidth}
+        \includegraphics{plots/illustrative_examples/boxplot_comp_time_normal}
+        \caption{Second dataset}
+    \end{subfigure}
+    \caption{Boxplot of the computation time of the different bootstrapping method and data set sizes $T$.
+The computation time refers to bootstrapping one confidence interval based upon $10,000$ values.
+Each boxplot reflects $10,000$ samples.
+The \ac{bca} method takes slighlty longer than the other two but the difference is neglectable.}
+    \label{fig:trending_bootstrap_time}
 \end{table}
@@
 Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
-
-\begin{itemize}
-    \item We use the alogrithms provided by Python's \verb|scipy|-package~\parencite{Virtanen2020}.
-    \item Two simulation setups
-    \item Take different $T$
-    \item Compute "true value" based on very large sample.
-    \item Compute share of bootstrap confidence intervals that actually covered the true value for different $B$.
-    \item Results in "double table": columns $T$, main rows $B$, undercategory: bootstrap methods.
-\end{itemize}

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
-As outlined in Section~\ref{sec:introduction}, we refer by trending to the consistency of an observed change with the change issued by another measurement, a nowcast or a forecast.
+As outlined in Section~\ref{sec:introduction}, we refer by trending to the consistency of an observed change with the change issued by another measurement, a nowcast, or a forecast.
 In this section, we want to specify trending and lay out different methods of measuring and visualizing trending. 

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 Section~\ref{subsec:trending-cond-prob} introduces a new graphical method of trending assessment.
-Section~\ref{subsec:trending-bootstrap} gives bootstrap methods for computing confidence intervals for the measures and conductes a simulation study on their properties.
+Section~\ref{subsec:trending-bootstrap} gives bootstrap methods for computing confidence intervals for the measures and conducts a simulation study on their properties.
 

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays prediction data differences.
-Points in the green upper right and lower left quadrant reflect a correct trending for the respective time step, whereas point in the remaining red quadrants show incorrect predicted changes.
+Points in the green upper right and lower left quadrants reflect a correct trending for the respective time step, whereas points in the remaining red quadrants show incorrect predicted changes.
 Figure~\ref{fig:trending_basic_4q} displays a basic four-quadrant plot.
@@
 
-Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$.
+Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for new elements of $\diffy$ and $\diffx$, respectively.
 Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R$).
@@
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
-Such a $2 \times 2$ table of counts is often called a contingency table and is often used in other scientific areas for evaluation, for example, in dichotomous forecasting or as a confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
+Such a $2 \times 2$ table of counts is often called a contingency table.
+It is often used in other scientific areas for evaluation, for example, in dichotomous forecasting or as a confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
 4]{James2021}[Ch.
@@
 Unbalanced outcomes of the differences are unlikely in our setting as they are obtained from differencing time series data.
-Nevertheless, if in a four-quadrant plot the number of positive and negative $\diffy$ are observed to differ widely, one should consider another as measures such as Cohen's $\kappa$ or the ones listed in \textcite[Table 3.3]{Jolliffe2012}.
+Nevertheless, suppose in a four-quadrant plot, the positive and negative $\diffy$ are observed to differ widely.
+In that case, one should consider another as measures such as Cohen's $\kappa$ or those listed in \textcite[Table 3.3]{Jolliffe2012}.
 
@@
 Thus, the yearly course of the prediction's trending ratio can be detected.
-This trade cannot be observed in the two four-quadrant plot as it is overlaid by the high number of points.
+This seasonal behavior cannot be observed in the two four-quadrant plot as the high number of points overlays it.
 
@@
 This will, for example, become evident in the data study in Section~\ref{sec:application-covid}.
-Data pairs with missing values can be excluded in the computation of the measures.
+Data pairs with missing values can be excluded from the computation of the measures.
 Nevertheless, missing values can considerably influence the results, mainly if they appear systematically.
-Predicting only in those times of certainty and omitting prediction in other times would generally improve evaluation measures.
+Predicting only in those times of certainty and omitting prediction in other times would improve evaluation measures.
 So, although the estimates can be computed, the time spans with missing values should be inspected visually.
@@
 
-The above measures use only the information on a point's quadrant and neglect further information on its location within the quadrant, as displayed in a four-quadrant plot.
-However, points close to the zero points have intuitively less explanatory power.
+The above measures use only the information on a point's quadrant and neglect further details on its location within the quadrant, as displayed in a four-quadrant plot.
+However, points close to the zero point have intuitively less explanatory power.
 If there is noise or non-systematic effects present in true value or nowcasts, those points are likely to be driven by noise to a large amount.
@@
 A cross-shaped exclusion area along both axes accounts for sign-reversal in both components.
-For those two methods, points 1 and 7 or 1, 6 and 7 are excluded, respectively.
+For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
 
@@
 Thus, the plot should be limited to the core values of $\diffx$ without outliers.
-Besides the kernel, a bandwith selector has to be chosen for a multivariate \ac{kde}.
+Besides the kernel, a bandwidth selector has to be chosen for a multivariate \ac{kde}.
 Various methods are available with different strengths and weaknesses.
-Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross validation maximum likelihood and cross validation least squares usigng the \verb|statsmodels| python package~\parencite{Seabold2010}.
+Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| python package~\parencite{Seabold2010}.
 The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
-The second method, cross validation least squares, did not converge for the first dataset after a computation time of approximately 300 times the other two methods.
-The rule of thumb yielded slightly oversmoothed results, while the cross validation maximum likelihood yields rather small bandwiths.
+The second method, cross-validation least squares, did not converge for the first dataset after a computation time of approximately 300 times that of the other two methods.
+The rule of thumb yielded slightly over-smoothed results, while the cross-validation maximum likelihood yields relatively small bandwidths.
 \todo[inline]{Add evaluation of the plot}
@@
     \caption{Resulting conditional trending plot for different bandwidth selection processes.
-Cross validation least squares takes a considerably larger computation time while yielding the results with best balance between smoothing and displaying local effects.
+Cross-validation least squares takes a considerably larger computation time while yielding the results with the best balance between smoothing and displaying local effects.
 The rule of thumb is the fastest method but tends to oversmooth.
-Cross validation least squares  }\label{fig:trending-cond-prob-bw}
+Cross-validation least squares  }\label{fig:trending-cond-prob-bw}
 \end{figure}
@@
 The basic approach balances those two aims.
-We compare the approaches in a small synthetic data study in the following with respect to their small-dataset behavior and their computation time.
+We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time.
 We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
-The considered datasets are again the ones outlined in Appendix~\ref{sec:app-trending-data-generation}, the first dataset with asymmetric dependence.
-In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals were used~\parencite{Virtanen2020}.
-The prescribed confidence level is 90 \% and the number of bootstrap samples is $10,000$.
+The considered datasets are again outlined in Appendix~\ref{sec:app-trending-data-generation}, the first dataset with asymmetric dependence.
+In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\parencite{Virtanen2020}.
+The prescribed confidence level is 90 \%, and the number of bootstrap samples is $10,000$.
 The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
@@
 Each boxplot reflects $10,000$ samples.
-The \ac{bca} method takes slighlty longer than the other two but the difference is neglectable.}
+The \ac{bca} method takes slightly longer than the other two, but the difference is negligible.}
     \label{fig:trending_bootstrap_time}

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 \caption{Four-quadrant plot with simulated data.}\label{fig:trending_basic_4q_sample}
-\end{subfigure}%
+\end{subfigure}\hspace{0.01\textwidth}
 \begin{subfigure}[t]{.24\textwidth}
@@
 \caption{The data is colored according to the time index $t$, the greener, the later.}\label{fig:trending_basic_4q_sample_color}
-\end{subfigure}%
+\end{subfigure}\hspace{0.01\textwidth}
 \begin{subfigure}[t]{.48\textwidth}
     \includegraphics{plots/illustrative_examples/trending_ratio_time_series.pdf}
-    \caption{Rolling estimate of the trending ratio over time with window length 91.
-A strong yearly seasonality of the trending ratio becomes visible.
-The overall trending ratios are $\mu = 0.7577$ and with exclusion area $\diffx < \varepsilon$, $\mu_{1.0} = 0.7712$, respectively.
+    \caption{Rolling estimate of the trending ratio over time with window length 91. }\label{fig:trending_ratio_time_series}
+    \end{subfigure}%
+    \caption{Visualizations for data with a time-varying trending ratio.
+We defer information on the data generation process of the data in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{sec:app-trending-data-generation}).
+The trending ratios on the whole data set are $\mu = 0.7577$ and with exclusion area $\diffx < \varepsilon$, $\mu_{1.0} = 0.7712$, respectively.
+A strong yearly seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
 The green curve $k_t$ shows the probability of $\diffxt$ having the same sign as $\diffyt$ for each time step.
-The rolling estimates hang back as the windows are backward-looking.}\label{fig:trending_ratio_time_series}
-    \end{subfigure}%
-    \caption{We defer information on the data generation process of the data in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{sec:app-trending-data-generation})}
+The rolling estimates hang back as the windows are backward-looking.}
 \end{figure}
@@
 This particularly suits prediction models, where small amounts of noise are inevitable.
-A cross-shaped exclusion area along both axes accounts for sign-reversal in both components.
+A cross-shaped exclusion area along both axes accounts for sign reversal in both components.
 For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 As noted above, a prediction is trending perfectly if all occurred changes are predicted correctly, that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+ \unsure{vielleicht nicht change sondern direction of the change}
 Perfect trending is hardly the case in practice, but a weaker requirement is needed.
@@
 For example, it is known in cardiac output measurement analysis~\parencite{Saugel2015}. \todo{Add other citations} 
+\unsure{\cite{perrino1998intraoperative}}
 Thereby, the occurred changes are plotted together with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
@@
 The trending ratios on the whole data set are $\mu = 0.7577$ and with exclusion area $\diffx < \varepsilon$, $\mu_{1.0} = 0.7712$, respectively.
-A strong yearly seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
+Strong yearly seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
 The green curve $k_t$ shows the probability of $\diffxt$ having the same sign as $\diffyt$ for each time step.

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 However, they lack the clearness and intuition of the four-quadrant plot while adding no more information on the trending~\parencite{Saugel2015}.
-
+\unsure{Vielleicht bei der Einführung des 4q-Plots auch exclusion area schon erwähnen?}
 \begin{figure}
@@
 Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for new elements of $\diffy$ and $\diffx$, respectively.
+\unsure{vielleict genauer: wofür diese zwei ZV stehen}
 Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R$).
+\unsure{der Fall: z1 = z2 = 0?}
 To estimate $P(\diffxrv \diffyrv > 0)$ it lies at hand to use

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 However, they lack the clearness and intuition of the four-quadrant plot while adding no more information on the trending~\parencite{Saugel2015}.
-\unsure{Vielleicht bei der Einführung des 4q-Plots auch exclusion area schon erwähnen?}
+
 \begin{figure}
@@
 In that case, one should consider another as measures such as Cohen's $\kappa$ or those listed in \textcite[Table 3.3]{Jolliffe2012}.
-
+\unsure{was heisst unbalanced konrket? und vielleicht kappa auch irgendwie anders einführen}
 All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
@@
 However, points close to the zero point have intuitively less explanatory power.
-If there is noise or non-systematic effects present in true value or nowcasts, those points are likely to be driven by noise to a large amount.
+If there is noise or non-systematic effects present in true value or nowcasts, those points are likely to be driven by noise to a large amount. \unsure{anstatt nowcasts eher predictions}
 Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the above methods.
@@
 For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
-
+\unsure{kann man vielleicht mu für verschiedene Varianten von Exclusion Area}
 In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
-\unsure{kann man vielleicht mu für verschiedene Varianten von Exclusion Area}
+\unsure{kann man vielleicht mu für verschiedene Varianten von Exclusion Area ausrechnen als Vergleich?}
 In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
-For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point.
+For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point. \unsure{}
 These shapes are theoretically appealing, but the interpretation of the resulting estimators gets ambiguous.

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
-For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point. \unsure{}
-These shapes are theoretically appealing, but the interpretation of the resulting estimators gets ambiguous.
+For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point.
+\unsure{vielleicht genauer bei der Verwendung von Quantilen}
+These shapes are theor etically appealing, but the interpretation of the resulting estimators gets ambiguous.
 Thus, we use the simple exclusion areas above.
@@
 \todo[inline]{Abschnitt unelegant}
-
+\unsure{wollen wir hier den Fass aufmachen?}
 As a base approach, we advocate using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$\varepsilon$ plot to visualize the course of the trending ratio over different exclusion area sizes.

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 \subsection{Conditional trending plot}\label{subsec:trending-cond-prob}
-
+\unsure{Allgemeine Frage: können wir hier von Wahrscheinlichkeit sprechen? }
 The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > \varepsilon)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < \varepsilon)$.

### 2024-02-19 sections/02-notation.tex
Small corrections
Update on Overleaf.

@@
 The trending detection of time series, as motivated in the introduction, is fundamentally interesting for evaluating and comparing methods from three areas: measurement technology, forecasting, and nowcasting.
-This paper relates to numerical data. \todo{Hängt irgendwie in der Luft}
+This paper relates to numerical data. \todo{Hängt irgendwie in der Luft} \unsure{Müsste es nicht auch ausschließlich um wiederholte Messungen gehen?}
 \todo{Überleitung, dass jetzt die einzelnen vorgestellt werden}
@@
 To detect the systematic difference between the two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero\cite{watson2010method}.
-In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (cf.
-\cite{lawrence1989concordance,koo2016guideline,}).
+In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline,}).
 \todo{Andere Zitierfunktion}
-Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizes the differences between the methods in relation to their mean value (cf. \cite{bland1986statistical}).
+Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizes the differences between the methods in relation to their mean value (\cite{bland1986statistical}).
 
@@
 \todo[inline]{Überleitung}
-Nowcasting (also short-term forecasting), on the other hand, has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (cf.
-\cite{browning1989nowcasting,schmid2019nowcasting}).
-Nowcasting focuses on predictions for the present, the immediate future, and the recent past and is now being used in other fields, such as the economy\todo{economics?}, where important statistics on the current economic situation are available only after a considerable time lag (cf.
-\cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioral factors (cf.
-\cite{wu2021nowcasting}).
+Nowcasting (also short-term forecasting), on the other hand, has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
+Nowcasting focuses on predictions for the present, the immediate future, and the recent past and is now being used in other fields, such as the economy\todo{economics?}, where important statistics on the current economic situation are available only after a considerable time lag (\cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioral factors (\cite{wu2021nowcasting}).
 \todo{sehr langer Satz}
@@
 Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exiertits \todo{word?} for $\tau$ less than or equal to $t$.
-Similar to forecasting, nowcasters are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (cf. \cite{gunther2021nowcasting}).
+Similar to forecasting, nowcasters are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}).
 

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 The trending detection of time series, as motivated in the introduction, is fundamentally interesting for evaluating and comparing methods from three areas: measurement technology, forecasting, and nowcasting.
+In the following, the characteristics of these three areas and the respective evaluation and comparison approaches are presented one after the other.
 This paper relates to numerical data. \todo{Hängt irgendwie in der Luft} \unsure{Müsste es nicht auch ausschließlich um wiederholte Messungen gehen?}
-\todo{Überleitung, dass jetzt die einzelnen vorgestellt werden}
 
@@
 Often, the question is how to compare one method of measurement of a variable with another.
-For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard) for the same variable in the same period of time.
+For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard, we note the measured values from the gold standard method with $(y_t)^T_{t=0}$) for the same variable in the same period of time.
 The measured values for a measurement method are then available in a time series form, e.g., can be noted with $(x_t)^T_{t=0}$. \todo{Fehlt da nicht noch $y$?}
-To detect the systematic difference between the two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero\cite{watson2010method}.
+To detect the systematic difference between the two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
 In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline,}).

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline,}).
-\todo{Andere Zitierfunktion}
 Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizes the differences between the methods in relation to their mean value (\cite{bland1986statistical}).
 
-\todo[inline]{Überleitung}
-Forecasting is focused on predicting the future based on historical data and its pattern.
+While measurement series are created by recording data using measuring instruments or direct observation, forecasting focuses on predicting the future based on historical data and its patterns.
 We note in the following the real realizations of the target value with $(\mathbf{y})^T_{t=0}$ and the prediction data for a target variable can be organized in such a way that $x_{t,\tau}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$.

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard, we note the measured values from the gold standard method with $(y_t)^T_{t=0}$) for the same variable in the same period of time.
-The measured values for a measurement method are then available in a time series form, e.g., can be noted with $(x_t)^T_{t=0}$. \todo{Fehlt da nicht noch $y$?}
+The measured values for a measurement method are then available in a time series form, e.g., can be noted with $(x_t)^T_{t=0}$.
 To detect the systematic difference between the two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
@@
 A comparison of several forecasters can then be made, for example, by ranking the forecast accuracies of the forecasters.
-In \cite{hyndman2006another}, an overview of various forecasting accuracies used in practice, such as scale-dependent measures like Root Mean Square Error (RMSE) or measures based on percentage errors like Root Mean Square Percentage Error (RMSPE), is presented and discussed.
-\todo{Herausstellen, dass bisherige Maße auf differenzen zwischen forecast und true herauslaufen; Hyndman eher als Quelle am Ende, sonst erwartet man, dass da noch andere Quellen besprochen werden}
-
-\todo[inline]{Überleitung}
-Nowcasting (also short-term forecasting), on the other hand, has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-Nowcasting focuses on predictions for the present, the immediate future, and the recent past and is now being used in other fields, such as the economy\todo{economics?}, where important statistics on the current economic situation are available only after a considerable time lag (\cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioral factors (\cite{wu2021nowcasting}).
+Compared to conventional forecasting methods, nowcasting focuses on predicting the current status and is therefore also referred to as short-term forecasting.
+Nowcasting  has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
+Nowcasting focuses on predictions for the present, the immediate future, and the recent past and is now being used in other fields, such as the economics, where important statistics on the current economic situation are available only after a considerable time lag (\cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioral factors (\cite{wu2021nowcasting}).
 \todo{sehr langer Satz}

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 A comparison of several forecasters can then be made, for example, by ranking the forecast accuracies of the forecasters.
+In practice, prediction accuracy is measured by scale-dependent measures such as the Root Mean Square Error (RMSE) or measures based on percentage errors such as the Root Mean Square Percentage Error (RMSPE).
+All of these typical measures are based on the absolute difference between forecast and true values, locally or globally.
+For an overview of different measures see \cite{hyndman2006another}.
 Compared to conventional forecasting methods, nowcasting focuses on predicting the current status and is therefore also referred to as short-term forecasting.
@@
 
-All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods take into account the local absolute differences in different ways but do not directly consider the influence of the sign of the differences.
-\todo{vielleicht: "do not directly consider whether the correct direction of change is issued"}
+All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods take into account the local absolute differences in different ways but do not directly consider whether the correct direction of change is issued
 To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or predictions with continuous outcomes.
@@
 \todo{Vielleicht nochmal Praxisbezug herausstellen, warum es sinnvoll ist, die Differenzen so zu formulieren?}
-In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting, or nowcasting methods are considered.
-
-For measurement:
+In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting, or nowcasting methods are considered. For measurement:
 \begin{equation}
@@
 \end{equation}
-For forecasting:
+For forecasting and nowcasting bezeichnen wir mit y die wahren werte der quantity of interesst. For Forcasting ist der Werzum Zeitpu
 \begin{equation}

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 
-All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods take into account the local absolute differences in different ways but do not directly consider whether the correct direction of change is issued
-To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or predictions with continuous outcomes.
+All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods take into account the local absolute differences in different ways but do not directly consider whether the correct direction of change is issued.
 Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see Section~\ref{sec:trending}).
@@
 \end{equation}
-For forecasting and nowcasting bezeichnen wir mit y die wahren werte der quantity of interesst. For Forcasting ist der Werzum Zeitpu
+For forecasts and nowcasts, we use $(\mathbf{y})^T_{t=0}$ to denote the true values of the quantity of interest.
+For forecasts, the current value is known at the time of the forecast, so a forecast for the time $t+l$ means the prediction of a change $x_{t,t+l} -y_{t}$ between $t$ and $t+l$.
+The time series of change can therefore be defined as follows
 \begin{equation}
-    \diffxl = (x_{t,t+l} -y_{t})^T_{t=0} \label{eq:diffxl_forecasting}
+    \diffxl = (x_{t,t+l} -y_{t})^T_{t=0}.\label{eq:diffxl_forecasting}
 \end{equation}
-For nowcasting:
+For nowcasting muss der Wert des aktuellen Zeitpunktes zum Vorhersage-Zeitpunkt nicht beaknnt sein. Dies kann durch folgende Fall Unterscheidung berücjksichtig werden:
 \begin{equation}

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 \todo{Satz klingt noch nicht so ganz rund}
-When predicting values in the present, it is usual that the true values of the recent past are not available, so only an estimate of these values is provided by the nowcast.\todo{Vielleicht noch etwas konkreter machen}
+When predicting values in the present, it is usual that the true values of the recent past are not available, so only an estimate of these values is provided by the nowcast.
 Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exiertits \todo{word?} for $\tau$ less than or equal to $t$.
@@
 Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see Section~\ref{sec:trending}).
-\todo{Vielleicht nochmal Praxisbezug herausstellen, warum es sinnvoll ist, die Differenzen so zu formulieren?}
-In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting, or nowcasting methods are considered. For measurement:
+For this purpose, a time series must be defined for a given time series (of measurements or predictions) that reflects the recording or prediction of changes (with a time delay $l$).
+For series of measurements, the change between two discrete points in time is considered.
 \begin{equation}
-    \diffxl = (x_{t+l} -x_{t})^T_{t=0} \label{eq:diffxl_measure}
+    \diffxl = (x_{t+l} -x_{t})^T_{t=0}. \label{eq:diffxl_measure}
 \end{equation}
@@
 \end{equation}
-For nowcasting muss der Wert des aktuellen Zeitpunktes zum Vorhersage-Zeitpunkt nicht beaknnt sein. Dies kann durch folgende Fall Unterscheidung berücjksichtig werden:
+For nowcasting, the value of the current point in time does not have to be known at the prediction time, often only an estimate of the current value is known.
+This can be taken into account by the following case distinction:
 \begin{equation}
@@
 \end{equation}
-
-\todo[inline]{Formeln als Tabelle oder in den Text einbinden.}
-\todo[inline]{$\mathbf{y}$ definieren.}
 

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 Nowcasting  has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-Nowcasting focuses on predictions for the present, the immediate future, and the recent past and is now being used in other fields, such as the economics, where important statistics on the current economic situation are available only after a considerable time lag (\cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioral factors (\cite{wu2021nowcasting}).
-\todo{sehr langer Satz}
-Applications of nowcasting methods for econometrics can be found here in \cite{giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic} or for predictions of epidemic development here in \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}.
-\todo{Satz klingt noch nicht so ganz rund}
+
 When predicting values in the present, it is usual that the true values of the recent past are not available, so only an estimate of these values is provided by the nowcast.
-Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exiertits \todo{word?} for $\tau$ less than or equal to $t$.
+For example, the current number of cases of an infectious disease may not be immediately available due to delays in data collection or reporting, or the situation may change rapidly so that the reported figures become inaccurate very quickly.
+Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exits for $\tau$ less than or equal to $t$.
 Similar to forecasting, nowcasters are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}).

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 In the following, the characteristics of these three areas and the respective evaluation and comparison approaches are presented one after the other.
-This paper relates to numerical data. \todo{Hängt irgendwie in der Luft} \unsure{Müsste es nicht auch ausschließlich um wiederholte Messungen gehen?}
+\unsure{Müsste es nicht auch ausschließlich um wiederholte Messungen gehen?}
 
@@
 We note in the following the real realizations of the target value with $(\mathbf{y})^T_{t=0}$ and the prediction data for a target variable can be organized in such a way that $x_{t,\tau}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$.
-\todo{Motivieren, warum beide Indizes nötig}
+The forecast is for the future point in time $\tau$ and is based on the information available at time $t$.
 In practice, a forecasting method is typically evaluated by forecast accuracy, which is defined by a certain loss function and evaluates the difference between the forecast and the actual outcome of the target variable.
@@
 Nowcasting  has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-
+Nowcasting focuses on predictions for the present, the immediate future and the recent past, and is now being used in other fields such as economics and medicine.
+Nowcasting can be used to predict important statistics about the current economic situation that are only available with a considerable time lag (\cite{banbura2013now, giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic}).
+In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioural factors (\cite{wu2021nowcasting}).
+For possible applications in this area, see \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}).
 When predicting values in the present, it is usual that the true values of the recent past are not available, so only an estimate of these values is provided by the nowcast.

### 2024-02-19 sections/02-notation.tex
Update on Overleaf.

@@
 In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioural factors (\cite{wu2021nowcasting}).
-For possible applications in this area, see \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}).
+For possible applications in this area, see \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}.
 When predicting values in the present, it is usual that the true values of the recent past are not available, so only an estimate of these values is provided by the nowcast.
@@
 For this purpose, a time series must be defined for a given time series (of measurements or predictions) that reflects the recording or prediction of changes (with a time delay $l$).
-For series of measurements, the change between two discrete points in time is considered.
+For series of measurements, the change between two discrete points in time is considered:
 \begin{equation}

### 2024-02-19 sections/03-trending.tex
Update on Overleaf.

@@
 However, points close to the zero point have intuitively less explanatory power.
-If there is noise or non-systematic effects present in true value or nowcasts, those points are likely to be driven by noise to a large amount. \unsure{anstatt nowcasts eher predictions}
+If there is noise or non-systematic effects present in true value or predictions, those points are likely to be driven by noise to a large amount. 
 Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the above methods.
@@
 \unsure{vielleicht genauer bei der Verwendung von Quantilen}
-These shapes are theor etically appealing, but the interpretation of the resulting estimators gets ambiguous.
+These shapes are theoretically appealing, but the interpretation of the resulting estimators gets ambiguous.
 Thus, we use the simple exclusion areas above.
@@
 In addition, the conditional trending plot facilitates the comparison of various methods in a single plot, and asymmetries in the trending ability can be detected.
+\unsure{Asymmetrie vielleicht irgendwo oben als Begriff einführen}
 
@@
 Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
+
+\unsure{berücsksichtigen die hier angeweneten Verfahren die Abhängigkeit?}

### 2024-02-20 sections/02-notation.tex
Improve manuscript

@@
 For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard, we note the measured values from the gold standard method with $(y_t)^T_{t=0}$) for the same variable in the same period of time.
-The measured values for a measurement method are then available in a time series form, e.g., can be noted with $(x_t)^T_{t=0}$.
+\unsure{Eher keine Klammern}
+The measured values for a measurement method are then available in a time series form, e.g., can be noted with $(x_t)^T_{t=0}$. \unsure{Warum e.g.?}
 To detect the systematic difference between the two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
@@
 While measurement series are created by recording data using measuring instruments or direct observation, forecasting focuses on predicting the future based on historical data and its patterns.
-We note in the following the real realizations of the target value with $(\mathbf{y})^T_{t=0}$ and the prediction data for a target variable can be organized in such a way that $x_{t,\tau}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$.
+We note in the following the real realizations of the target value with $(\mathbf{y})^T_{t=0}$ \unsure{notiert man das so?} and the prediction data for a target variable can be organized in such a way that $x_{t,\tau}$ \unsure{Indices eher andersrum?} the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$.
 The forecast is for the future point in time $\tau$ and is based on the information available at time $t$.
@@
 For an overview of different measures see \cite{hyndman2006another}.
-Compared to conventional forecasting methods, nowcasting focuses on predicting the current status and is therefore also referred to as short-term forecasting.
+Compared \unsure{In contrast} to conventional forecasting methods, nowcasting focuses on predicting the current status and is therefore also referred to as short-term forecasting.
 Nowcasting  has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).

### 2024-02-20 sections/02-notation.tex
Update on Overleaf.

@@
 In the following, the characteristics of these three areas and the respective evaluation and comparison approaches are presented one after the other.
-\unsure{Müsste es nicht auch ausschließlich um wiederholte Messungen gehen?}
+\unsure{Müsste es nicht auch ausschließlich um wiederholte Messungen gehen?; one after the other weglassen}
 
@@
 All of these typical measures are based on the absolute difference between forecast and true values, locally or globally.
-For an overview of different measures see \cite{hyndman2006another}.
+For an overview of different measures, see \cite{hyndman2006another}.
 Compared \unsure{In contrast} to conventional forecasting methods, nowcasting focuses on predicting the current status and is therefore also referred to as short-term forecasting.
-Nowcasting  has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-Nowcasting focuses on predictions for the present, the immediate future and the recent past, and is now being used in other fields such as economics and medicine.
-Nowcasting can be used to predict important statistics about the current economic situation that are only available with a considerable time lag (\cite{banbura2013now, giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic}).
-In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioural factors (\cite{wu2021nowcasting}).
+\unsure[noinline]{Ich finde da wird der Unterschied noch nicht so richtig klar} Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
+Nowcasting focuses on predictions for the present, the immediate future, and the recent past, and is now being used in other fields such as economics and medicine.
+Nowcasting can be used to predict important statistics about the current economic situation that is only available with a considerable time lag (\cite{banbura2013now, giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic}).
+In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
 For possible applications in this area, see \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}.
@@
 Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exits for $\tau$ less than or equal to $t$.
-Similar to forecasting, nowcasters are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}).
+Similar to forecasting, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}).
 
@@
 For this purpose, a time series must be defined for a given time series (of measurements or predictions) that reflects the recording or prediction of changes (with a time delay $l$).
-For series of measurements, the change between two discrete points in time is considered:
+For a series of measurements, the change between two discrete points in time is considered:
 \begin{equation}
@@
 For forecasts, the current value is known at the time of the forecast, so a forecast for the time $t+l$ means the prediction of a change $x_{t,t+l} -y_{t}$ between $t$ and $t+l$.
-The time series of change can therefore be defined as follows
+The time series of change can therefore be defined as follows\improvement{follows weglassen}
 \begin{equation}
@@
 \end{equation}
-For nowcasting, the value of the current point in time does not have to be known at the prediction time, often only an estimate of the current value is known.
+For nowcasting, the value of the current point in time does not have to be known at the prediction time.
+Often, only an estimate of the current value is known.
 This can be taken into account by the following case distinction:

### 2024-02-20 sections/03-trending.tex
Update on Overleaf.

@@
 In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$.
-Similarly, we call a measurement, nowcast, or forecast a \textit{prediction} for the quantity of interest $y$. \unsure{Discuss}
+Similarly, we call a measurement, nowcast, or forecast a \textit{prediction} for the quantity of interest $y$. \discuss{prediction a good word?}
 The remainder of the section is structured as follows.
@@
 
-As noted above, a prediction is trending perfectly if all occurred changes are predicted correctly, that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
- \unsure{vielleicht nicht change sondern direction of the change}
+As noted above, a prediction is trending perfectly if all occurred directions of change are predicted correctly, that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
 Perfect trending is hardly the case in practice, but a weaker requirement is needed.
@@
 A simple yet insightful method is the four-quadrant plot.
-For example, it is known in cardiac output measurement analysis~\parencite{Saugel2015}. \todo{Add other citations} 
-\unsure{\cite{perrino1998intraoperative}}
+For example, it is known in cardiac output measurement analysis~\parencite{Saugel2015,perrino1998intraoperative}. 
 Thereby, the occurred changes are plotted together with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
@@
 Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for new elements of $\diffy$ and $\diffx$, respectively.
-\unsure{vielleict genauer: wofür diese zwei ZV stehen}
-Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R$).
-\unsure{der Fall: z1 = z2 = 0?}
+\discuss{vielleicht genauer: wofür diese zwei ZV stehen}
+Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$).
 To estimate $P(\diffxrv \diffyrv > 0)$ it lies at hand to use
@@
 There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \todo{Citation}.
+\todo{Kappa in einem Satz definieren}
+\todo{Was heißt unbalanced; warum ist das bei uns nicht?}
 Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.
@@
 In that case, one should consider another as measures such as Cohen's $\kappa$ or those listed in \textcite[Table 3.3]{Jolliffe2012}.
-\unsure{was heisst unbalanced konrket? und vielleicht kappa auch irgendwie anders einführen}
 All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
@@
 For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
-\unsure{kann man vielleicht mu für verschiedene Varianten von Exclusion Area ausrechnen als Vergleich?}
+\todo{kann man vielleicht mu für verschiedene Varianten von Exclusion Area ausrechnen als Vergleich? im Anhang?}
 In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
-For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point.
-\unsure{vielleicht genauer bei der Verwendung von Quantilen}
+For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point, for example, exclude the points within the 95\% ...
+.
+\discuss{vielleicht genauer bei der Verwendung von Quantilen}
 These shapes are theoretically appealing, but the interpretation of the resulting estimators gets ambiguous.
@@
 \todo[inline]{Abschnitt unelegant}
-\unsure{wollen wir hier den Fass aufmachen?}
+\discuss{wollen wir hier den Fass aufmachen?}
 As a base approach, we advocate using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$\varepsilon$ plot to visualize the course of the trending ratio over different exclusion area sizes.
@@
 \subsection{Conditional trending plot}\label{subsec:trending-cond-prob}
-\unsure{Allgemeine Frage: können wir hier von Wahrscheinlichkeit sprechen? }
+\discuss{Allgemeine Frage: können wir hier von Wahrscheinlichkeit sprechen? }
 The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > \varepsilon)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < \varepsilon)$.
@@
 Whereas it would be possible to build analog measures on other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
-In addition, the conditional trending plot facilitates the comparison of various methods in a single plot, and asymmetries in the trending ability can be detected.
-\unsure{Asymmetrie vielleicht irgendwo oben als Begriff einführen}
+In addition, the conditional trending plot facilitates the comparison of various methods in a single plot, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
+\done{Asymmetrie vielleicht irgendwo oben als Begriff einführen}
 
@@
 The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
-The second method, cross-validation least squares, did not converge for the first dataset after a computation time of approximately 300 times that of the other two methods.
-The rule of thumb yielded slightly over-smoothed results, while the cross-validation maximum likelihood yields relatively small bandwidths.
-\todo[inline]{Add evaluation of the plot}
+The second method, cross-validation least squares, needs long computation times while yielding small or no bandwidth results, even for the two relatively small datasets.
+The rule-of-thumb and cross-validation maximum likelihood both yield reasonable results at moderate computation times. 
 Further examples, including comparisons between methods concerning their trending ability, are available for the applications in Section~\ref{sec:application}.
@@
     \caption{Resulting conditional trending plot for different bandwidth selection processes.
-Cross-validation least squares takes a considerably larger computation time while yielding the results with the best balance between smoothing and displaying local effects.
+Cross-validation least squares takes a considerably larger computation time.
+It does not converge for the first data set and yields a too small bandwidth for the second data set.
 The rule of thumb is the fastest method but tends to oversmooth.
-Cross-validation least squares  }\label{fig:trending-cond-prob-bw}
+The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time.
+}\label{fig:trending-cond-prob-bw}
 \end{figure}
@@
 
-In bootstrapping, the sample is assumed to be independent and identically distributed.
+In standard bootstrapping, the sample is assumed to be independent and identically distributed.
 Thus, before applying bootstrapping methods, the strength of sequential dependence should be inspected, for example, by analyzing the autocorrelation and partial autocorrelation.
+The simple bootstrapping methods above do not account for serial dependence. 
 Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
 
-\unsure{berücsksichtigen die hier angeweneten Verfahren die Abhängigkeit?}
+\done{berücksichtigen die hier angeweneten Verfahren die Abhängigkeit?}

### 2024-02-20 sections/03-trending.tex
Update on Overleaf.

@@
 Section~\ref{subsec:trending-example} gives an example of such a plot.
-Instead of determining the exclusion area size beforehand, the plot can be used to choose the appropriate $\varepsilon$.
-\todo[inline]{Abschnitt unelegant}
-\discuss{wollen wir hier den Fass aufmachen?}
 As a base approach, we advocate using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$\varepsilon$ plot to visualize the course of the trending ratio over different exclusion area sizes.
@@
 In addition, the conditional trending plot facilitates the comparison of various methods in a single plot, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
-\done{Asymmetrie vielleicht irgendwo oben als Begriff einführen}
 

### 2024-02-20 sections/03-trending.tex
Update on Overleaf.

@@
 In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$.
-Similarly, we call a measurement, nowcast, or forecast a \textit{prediction} for the quantity of interest $y$. \discuss{prediction a good word?}
+Similarly, we call a measurement, nowcast, or forecast a \textit{prediction} for the quantity of interest $y$. \discuss{Is prediction a good word?}
 The remainder of the section is structured as follows.
@@
 Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for new elements of $\diffy$ and $\diffx$, respectively.
-\discuss{vielleicht genauer: wofür diese zwei ZV stehen}
+\discuss{Überlegung: sollte genauer beschrieben sein, wofür die beiden ZV stehen?}
 Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$).
@@
 The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
-The true values of the accuracy are computed based on a dataset of size $10^9$, yielding 0.7501 and 0.7700 for the two datasets.
+The true values of the accuracy are computed based on a dataset of size $10^8$, yielding 0.7501 and 0.7700 for the two datasets.
 The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.

### 2024-02-20 sections/03-trending.tex
Update on Overleaf.

@@
 The two give estimates of $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$, respectively.
-There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \todo{Citation}.
-\todo{Kappa in einem Satz definieren}
-\todo{Was heißt unbalanced; warum ist das bei uns nicht?}
+There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960}.
+Cohen's $\kappa$ is usually used to measure inter-rater agreement and takes into account the ratio of occurred agreement and the probability of agreement by chance.
 Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.
+Balanced outcome hereby refers to approximately equal occurrences of values smaller and greater zero for both $\diffx$ and $\diffy$.
 Unbalanced outcomes of the differences are unlikely in our setting as they are obtained from differencing time series data.
@@
 For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
-\todo{kann man vielleicht mu für verschiedene Varianten von Exclusion Area ausrechnen als Vergleich? im Anhang?}
+\todo{Berechnung $\mu$ für verschiedene Varianten für Beispiel im Anhang?}
 In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
-For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point, for example, exclude the points within the 95\% ...
+For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point, for example, exclude the points within the central 95\% ...
 .

### 2024-02-20 sections/03-trending.tex
Update on Overleaf.

@@
 In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
-For example, one could exclude points likely to be in another quadrant or within a specific quantile of a distribution around the zero point, for example, exclude the points within the central 95\% ...
-.
-\discuss{vielleicht genauer bei der Verwendung von Quantilen}
+For example, one could exclude points likely to be in another quadrant or within a specific quantile range of a distribution around the zero point, for example, exclude the points within the 5 \% quantile and 95 \% quantile of a standard normal distribution.
 These shapes are theoretically appealing, but the interpretation of the resulting estimators gets ambiguous.
@@
 Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
-
-\done{berücksichtigen die hier angeweneten Verfahren die Abhängigkeit?}

### 2024-02-20 sections/02-notation.tex
Update on Overleaf.

@@
 In the following, the characteristics of these three areas and the respective evaluation and comparison approaches are presented one after the other.
-\unsure{Müsste es nicht auch ausschließlich um wiederholte Messungen gehen?; one after the other weglassen}
 
-Measurement aims to obtain accurate and reliable data about the current state of a system.
-Often, the question is how to compare one method of measurement of a variable with another.
-For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard, we note the measured values from the gold standard method with $(y_t)^T_{t=0}$) for the same variable in the same period of time.
-\unsure{Eher keine Klammern}
-The measured values for a measurement method are then available in a time series form, e.g., can be noted with $(x_t)^T_{t=0}$. \unsure{Warum e.g.?}
+Measurement aims to obtain accurate and reliable data about the current state of a system. Often a certain parameter or variable is measured at regular intervals over a certain period of time.
+An important question is how to compare one method of measurement of a variable with another.
+For this purpose, measured values can be recorded with the two measurement methods to be compared.
+One can be the gold standard, we note the measured values from the gold standard method with $(y_t)^T_{t=0}$) for the same variable in the same period of time.
+The measured values for a measurement method are then available in a time series form, can be noted with $(x_t)^T_{t=0}$. 
 To detect the systematic difference between the two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
@@
 While measurement series are created by recording data using measuring instruments or direct observation, forecasting focuses on predicting the future based on historical data and its patterns.
-We note in the following the real realizations of the target value with $(\mathbf{y})^T_{t=0}$ \unsure{notiert man das so?} and the prediction data for a target variable can be organized in such a way that $x_{t,\tau}$ \unsure{Indices eher andersrum?} the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$.
+We note in the following the real realizations of the target value with $(y_t)^T_{t=0}$ and the prediction data for a target variable can be organized in such a way that $x_{\tau|t}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$.
 The forecast is for the future point in time $\tau$ and is based on the information available at time $t$.
@@
 For an overview of different measures, see \cite{hyndman2006another}.
-Compared \unsure{In contrast} to conventional forecasting methods, nowcasting focuses on predicting the current status and is therefore also referred to as short-term forecasting.
+In contrast to conventional forecasting methods, nowcasting focuses on predicting the current status and is therefore sometimes also referred to as short-term forecasting.
 \unsure[noinline]{Ich finde da wird der Unterschied noch nicht so richtig klar} Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
@@
 For example, the current number of cases of an infectious disease may not be immediately available due to delays in data collection or reporting, or the situation may change rapidly so that the reported figures become inaccurate very quickly.
-Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exits for $\tau$ less than or equal to $t$.
+Using the notation for forecasting above, this means that a prediction value $x_{\tau|t}$ also exits for $\tau$ less than or equal to $t$.
 Similar to forecasting, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}).
@@
 For this purpose, a time series must be defined for a given time series (of measurements or predictions) that reflects the recording or prediction of changes (with a time delay $l$).
-For a series of measurements, the change between two discrete points in time is considered:
+For a series of measurements, the change between two discrete points in time is considered
 \begin{equation}
@@
 \end{equation}
-For forecasts and nowcasts, we use $(\mathbf{y})^T_{t=0}$ to denote the true values of the quantity of interest.
-For forecasts, the current value is known at the time of the forecast, so a forecast for the time $t+l$ means the prediction of a change $x_{t,t+l} -y_{t}$ between $t$ and $t+l$.
-The time series of change can therefore be defined as follows\improvement{follows weglassen}
+For forecasts and nowcasts, we use $(y_t)^T_{t=0}$ to denote the true values of the quantity of interest.
+For forecasts, the current value is known at the time of the forecast, so a forecast for the time $t+l$ means the prediction of a change $x_{t+l|t} -y_{t}$ between $t$ and $t+l$.
+The time series of change can therefore be defined as
 \begin{equation}
-    \diffxl = (x_{t,t+l} -y_{t})^T_{t=0}.\label{eq:diffxl_forecasting}
+    \diffxl = (x_{t+l|t} -y_{t})^T_{t=0}.\label{eq:diffxl_forecasting}
 \end{equation}
@@
 Often, only an estimate of the current value is known.
-This can be taken into account by the following case distinction:
+This can be taken into account by the following case distinction
 \begin{equation}
@@
 \begin{cases} 
-(x_{t,t+l} -x_{t,t})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
-(x_{t,t+l} -y_{t})^T_{t=0}  & \text{else}.
+(x_{t+l|t} -x_{t|t})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
+(x_{t+l|t} -y_{t})^T_{t=0}  & \text{else}.
 \end{cases} \label{eq:diffxl_nowcasting}

### 2024-02-20 sections/02-notation.tex
Update on Overleaf.

@@
 Nowcasting focuses on predictions for the present, the immediate future, and the recent past, and is now being used in other fields such as economics and medicine.
-Nowcasting can be used to predict important statistics about the current economic situation that is only available with a considerable time lag (\cite{banbura2013now, giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic}).
+For example, Nowcasting can be used to predict important statistics about the current economic situation that is only available with a considerable time lag (\cite{banbura2013now, giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic}).
 In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).

### 2024-02-20 sections/02-notation.tex
Update on Overleaf.

@@
 In contrast to conventional forecasting methods, nowcasting focuses on predicting the current status and is therefore sometimes also referred to as short-term forecasting.
-\unsure[noinline]{Ich finde da wird der Unterschied noch nicht so richtig klar} Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
+Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
+One difference to the conventional forecast is that nowcasts can also use latent variables, which are often published with higher frequency in the current period, to forecast the target figure for the current point in time (\cite{castle2017forecasting}).
 Nowcasting focuses on predictions for the present, the immediate future, and the recent past, and is now being used in other fields such as economics and medicine.

### 2024-02-20 sections/02-notation.tex
Update on Overleaf.

@@
 For this purpose, measured values can be recorded with the two measurement methods to be compared.
-One can be the gold standard, we note the measured values from the gold standard method with $(y_t)^T_{t=0}$) for the same variable in the same period of time.
+One can be the gold standard, we note the measured values from the gold standard method with $(y_t)^T_{t=0}$ for the same variable in the same period of time.
 The measured values for a measurement method are then available in a time series form, can be noted with $(x_t)^T_{t=0}$. 

### 2024-02-20 sections/03-trending.tex
Update on Overleaf.

@@
 The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.
-\todo[inline]{Add evaluation when computation is finished}
+For the small sample sizes up to $T = 168$, only the \ac{bca} method keeps the confidence interval size and yields slightly wider confidence intervals. 
+The method's results do not differ for the larger sample sizes. 
+The computation time for the \ac{bca} method is slightly larger than for the other methods, but all methods have a moderate computation time.
 

### 2024-02-22 paper.tex
Update 18.2.2024
Update bootstrapping
Improve manuscript
Update on Overleaf.

@@
 %\usepackage{tabularray}
+\usepackage{etoolbox}
+\AtBeginEnvironment{tabular}{\small}
 
@@
 \newcommandx{\unsure}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
+\newcommandx{\discuss}[2][1=]{\todo[linecolor=orange,backgroundcolor=orange!25,bordercolor=orange,#1]{#2}}
 \newcommandx{\missing}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
@@
 \newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
+\newcommandx{\done}[2][1=]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{#2}}
 \newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
@@
 \DeclareAcronym{bca}{short=BCa, long=bias-corrected and accelerated}
+\DeclareAcronym{cdf}{short=CDF, long=cumulative distribution function}
 
@@
 %%% --- ToDo List
-\usepackage{enumitem}
-\newlist{todolist}{itemize}{2}
-\setlist[todolist]{label=$\square$}
-\usepackage{pifont}
-\newcommand{\cmark}{\ding{51}}%
-\newcommand{\xmark}{\ding{55}}%
-\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
-\hspace{-2.5pt}}
-\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}
+%\usepackage{enumitem}
+%\newlist{todolist}{itemize}{2}
+%\setlist[todolist]{label=$\square$}
+%\usepackage{pifont}
+%\newcommand{\cmark}{\ding{51}}%
+%\newcommand{\xmark}{\ding{55}}%
+%\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
+%\hspace{-2.5pt}}
+%\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}
 
@@
 
-Agenda für heute:
 \begin{itemize}
@@
     \begin{itemize}
-        \item Introduction (1.5 pages), Conclusion (0.5 pages), Abstract
-        \item Nowcasting (1 pages): Bolin
+        \item Introduction (1.5 pages): Bolin fängt an, Conclusion (0.5 pages), Abstract
+        \item Application areas (1 pages): Bolin
         \item Trending (5-6 pages): Jonas fängt an
@@
 \begin{itemize}    
-    \item Journal/Anwendungsgebiet klären: Angewandtes Medizin Statistikjournal? Dann müsste man die folgenden Abschnitte entsprechend ausrichten
+    \item Journal/Anwendungsgebiet klären: Angewandtes Medizin Statistikjournal: z.B. BMC Medical Research Methodology
 \end{itemize}
@@
 \section{Data generation for Section~\ref{sec:trending}}\label{sec:app-trending-data-generation}
-
-\todo{Write}
-
-Beispiel
-\begin{itemize}
-  \item $\diffyt \sim U(-5, 5) + N(0, 1)$
-  \item $\diffxt \sim \diffyt \cdot \SBer(k) \cdot (\text{Truncated Normal, mean 1, sigma 0.5, trunc 0})$
-\end{itemize}
+\input{sections/50-appendix-data-generation}
 

### 2024-02-22 sections/01-introduction.tex
Update on Overleaf.

@@
 \end{itemize}
+
+The paper is organized as follows.
+In Section \ref{sec:notation}, we compare the three possible application areas for trending assessment: measurement, forecasting and nowcasting, and define the time series of change according to the application for a given time series.
+In Section \ref{sec:trending}, we formalise the concept of trending independently of the application domain and investigate quantification approaches.
+In particular, we introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries.
+Bootstrap methods for calculating confidence intervals for trending measures are also presented.
+In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting and nowcasting.
+We conclude our paper with the last section.

### 2024-02-23 sections/04-application.tex
Change labels and add math notation
Add draft for Section 3
Add draft of first part on covid nowcasting

@@
-\begin{itemize}
-    \item Drei Anwendungen, jeweils eine aus den Bereichen Messwerte, Nowcasting, Forecasting
-    \item Messwerte: Daten Medizinmesswerte
-    \item Nowcasting: Covid-Nowcasting-Hub
-    \begin{itemize}
-        \item Daten sind schon aufbereitet
-        \item Vergleich zwischen Nowcastern möglich
-        \item Noch zu klären: Wie wird $\xdiff$ gewählt?
-    \end{itemize}
-    \item Forecasting: Pandemie-/Epidemievorhersage, Medikamentenbedarf
-\end{itemize}
+\subsection{Measurement data} \label{sec:application_measurement}
 
-\subsection{Covid nowcasting}
+\newpage
+\subsection{Covid nowcasting} \label{sec:application-covid}
 
-Data from Covid-Nowcasting-Hub
+During the COVID 19 pandemic, the need for reliable and timely nowcasts of pandemic's development has become apparent.
+In Germany, the seven-day hospitalization rate has been established as a central steering measure for the COVID measures in November 2021 and the imposement of severe public restrictions were taken based on it~\parencite{RobertKochInstitute2021}.
+The Robert Koch Institute (RKI) provides daily reports on the number of new hospitalizations.
+However, these reports are subject to delays and revisions rooted in two different sources.
+The first source is technical delays in the reporting process, for example, due to different authorities passing on the data to the RKI~\parencite{RobertKochInstitute2024}.
+The second, more systematic source is the structure of the seven-day hospitalization rate.
+To a given date, all the cases are allocated, whose first positive test result was on that date and who were hospitalized in relation to the disease in the following.
+The average count of cases on the given date and six days before of $100,000$ inhabitants is the seven-day hospitalization rate.
+Thus, the final seven-day hospitalization rate can only be reported with significant delay of up to more than 70 days.
+Nevertheless, as the seven-day hospitalization rate was considered a major indicator for the pandemic's development, many organizations and institutions started to issue nowcasts of the seven-day hospitalization rate, including research teams and newspapers.
+To collect nowcasts of the seven-day hospitalization rate by different nowcast groups, the COVID19-Nowcasting-Hub was established~\parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+The nowcasts contain the predictive mean, median and quantiles for the seven-day hospitalization rate.
 
@@
     \centering
+    \begin{subfigure}[t]{0.48\textwidth}
     \includegraphics{plots/covid_nowcast/00_true_data.pdf}
-    \caption{The seven-day-hospitalization of Covid in Germany as reported by the RKI \hl{Add citation}.}
+    \caption{Realisations.}
     \label{fig:app-covid-true}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
+    \caption{Same-day nowcasts}
+    \label{fig:app-covid-nowcast}
+        \end{subfigure}
+    \caption{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021 to April 29, 2022 \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+    The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
+    \label{fig:app-covid-true-nowcast}
 \end{figure}
 
-\begin{figure}
-    \centering
-    \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
-    \caption{Nowcast data without removal of outliers}
-    \label{fig:app-covid-nowcast}
-\end{figure}
+The data contains 8 nowcasts from scientific and public institutions, nowcast communities and a newspaper, using different input variables, calendar data, and length of training data.
+The model structures are diverse, including bayesian models, generalized additive models, and parametric bootstrapping.
+Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the overall structure of the model.
+Using the nowcasts, two ensemble methods are constructed by using mean or median of the ensembles.
+We denote them by ENS-MEAN and ENS-MED.
+In line with the initial study design, we consider the time span from November 22, 2021 to April 29, 2022 as evaluation period.
+In contrast to~\textcite{Wolffram2023} we use the newest data available on the true values and not the data from August 8, 2022.
+\unsure{Nochmal längere Evaluation?}
+We do not expand on the models' performance on specific regions or age groups in Germany and on the probabilistic nowcast assessment.
+Figure~\ref{fig:app-covid-true-nowcast} displays the true and nowcast data for the evaluation period.
+The time comprises the forth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany lasting until May 28, 2022~\parencite{Tolksdorf2022}.
+Table~\ref{tab:app-covid-rmse} summarizes the point evaluation measures for the issued mean of the different models.
+The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\parencite[see][Tables A2, A3, and A4]{Wolffram2023}.
+The best performing models are the ILM and RKI model in terms of both RMSE and MAE.
+The ensemble methods perform worse than the best models in terms of the mean location.
+The performance of the models is diverse, with more than twice as high RMSE values for the worst models compared to the best models.
+Note that the high values for the EPI model could be driven by a very far off value at the end of the evaluation period.
 
-\begin{figure}
-    \centering
-    \includegraphics{plots/covid_nowcast/01_nowcast_data_without_outliers.pdf}
-    \caption{Nowcast data after removal of outliers}
-    \label{fig:app-covid-nowcast-no-outliers}
-\end{figure}
+In addition to close inspection of the point evaluation measures, assessing trending of the nowcasts is crucial.
+To assess the impact of taken measures and the direction of curve, it is important to distinguish between rising and falling hospitalization rates.
+If hospitalization rates are rising, measures should be tightened, while falling rates might allow for loosening measures.
+Especially, asymmetries are of interest: Are some models better in recognizing a fall than a rise or vice versa?
+
+In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
+We report the trending for the lags 1, 7, and 14 days.
+While lag 1 assesses the short-term trending, lags 7 and 14 assess the medium-term trending.
+For the best models in terms of point evaluation, we also assess the trending with the conditional probability plot.
 
@@
     \input{plots/covid_nowcast/01_model_scores}
-    \caption{Point evaluation measures for the issued mean of the different models. Note that not for all dates all models issue nowcasts.}
+    \caption{Point evaluation measures for the issued mean of the different models. The evaluation period comprises 159 days. }
     \label{tab:app-covid-rmse}
@@
 
-\begin{itemize}
-    \item Seven-day hospitalization rate as central steering measure for COVID measures in Germany
-    \item Describe how it is computed
-    \item Outline measures and sources to find methods
-    \item Present overall measures for methods (Table~\ref{tab:app-covid-rmse})
-    \item Present why trending is important, particularly the differentiation between increase and decrease
-    \item Use only a subset of all methods for trending evaluation (best methods)
-    \item Present trending of the subset
-\end{itemize}
-\subsection{Forecasting emergency department arrivals}
+\begin{table}
+    \centering
+    \missingfigure{TABLE}
+    \caption{Trending ratios for the models with and without exclusion areas.
+The exclusion areas are rectangles centered on the zero points with width and height ...
+.
+The ratios are calculated for the lags 1, 7, and 14 days.}
+    \label{tab:app-covid-ratios}
+\end{table}
+
+\newpage
+\subsection{Forecasting emergency department arrivals}\label{sec:application-eda}
 

### 2024-02-23 sections/55-appendix-application-covid.tex
Add draft of first part on covid nowcasting

\section{Additional material on Section~\ref{sec:application-covid}}\label{sec:appendix-application-covid}

\begin{figure}
    \centering
    \begin{subfigure}[t]{.31\textwidth}
        \missingfigure{KDE for true and nowcasts of lag 1.}
        \caption{\ac{kde} for true values and nowcasts of lag 1.}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}[t]{.31\textwidth}
        \missingfigure{KDE for true and nowcasts of lag 7.}
        \caption{\ac{kde} for true values and nowcasts of lag 7.}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}[t]{.31\textwidth}
        \missingfigure{KDE for true and nowcasts of lag 14.}
        \caption{\ac{kde} for true values and nowcasts of lag 14.}
    \end{subfigure}
    \caption{\Ac{kde} for true values and nowcasts of lags 1, 7, and 14 to assess the distribution of values. \hl{AUSWERTUNG; 10 \% der Werte liegen innerhalb von ... um den Ursprung.}}
    \label{fig:app-covid-kde}
\end{figure}

\begin{table}
    \centering
    \begin{tabular}{l X}
        \toprule
        Abbreviation & Nowcasting hub key \\
        \midrule
        EPI & Epiforecasts-independent \\
        ILM & ILM-prop \\
        KIT & KIT-simple\_nowcast \\
        LMU & LMU\_StaBLab-GAM\_nowcast \\
        RIVM & RIVM-KEW \\
        RKI & RKI-weekly\_report \\
        SU & SU-hier\_bayes \\
        SZ & SZ-hosp\_nowcast\\
        ENS-MEAN & NowcastHub-MeanEnsemble\\
        ENS-MED & NowcastHub-MedianEnsemble\\
        \bottomrule
    \end{tabular}
    \caption{Matching of the abbreviation to the key in the nowcasting hub .Information on the models and references are listed in \textcite[][Table 1]{Wolffram2023}.}
    \label{tab:app-covid-models}
\end{table}

### 2024-02-27 sections/04-application.tex
Update Covid nowcast analysis

@@
     \centering
-    \missingfigure{TABLE}
-    \caption{Trending ratios for the models with and without exclusion areas.
-The exclusion areas are rectangles centered on the zero points with width and height ...
-.
-The ratios are calculated for the lags 1, 7, and 14 days.}
-    \label{tab:app-covid-ratios}
+    \tiny
+    \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
+    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the lag 7 days.
+The exclusion areas are rectangles centered on the zero points with width and height of 10\% of the quantile of the absolute values of nowcast and true values.
+}
+    \label{tab:app-covid-trending-ratios-lag-7}
 \end{table}

### 2024-02-27 sections/55-appendix-application-covid.tex
Update Covid nowcast analysis

@@
     \centering
-    \begin{subfigure}[t]{.31\textwidth}
-        \missingfigure{KDE for true and nowcasts of lag 1.}
+    \begin{subfigure}[t]{.48\textwidth}
+        \includegraphics{plots/covid_nowcast/20_kde_lag_1}
         \caption{\ac{kde} for true values and nowcasts of lag 1.}
     \end{subfigure}\hspace{0.01\textwidth}
-    \begin{subfigure}[t]{.31\textwidth}
-        \missingfigure{KDE for true and nowcasts of lag 7.}
+    \begin{subfigure}[t]{.48\textwidth}
+        \includegraphics{plots/covid_nowcast/20_kde_lag_7}
         \caption{\ac{kde} for true values and nowcasts of lag 7.}
     \end{subfigure}\hspace{0.01\textwidth}
-    \begin{subfigure}[t]{.31\textwidth}
-        \missingfigure{KDE for true and nowcasts of lag 14.}
+    \begin{subfigure}[t]{.48\textwidth}
+        \includegraphics{plots/covid_nowcast/20_kde_lag_14}
         \caption{\ac{kde} for true values and nowcasts of lag 14.}
     \end{subfigure}
-    \caption{\Ac{kde} for true values and nowcasts of lags 1, 7, and 14 to assess the distribution of values. \hl{AUSWERTUNG; 10 \% der Werte liegen innerhalb von ... um den Ursprung.}}
+    \caption{\Ac{kde} for true values and nowcasts of lags 1, 7, and 14 to assess the distribution of values.
+\hl{AUSWERTUNG}.
+Exclusion areas based on the 10\%  quantile of absolute values are listed in Table~\ref{tab:app-covid-marginals}.}
     \label{fig:app-covid-kde}
@@
     \centering
-    \begin{tabular}{l X}
+    \begin{tabular}{l l}
         \toprule
@@
     \end{tabular}
-    \caption{Matching of the abbreviation to the key in the nowcasting hub .Information on the models and references are listed in \textcite[][Table 1]{Wolffram2023}.}
+    \caption{Matching of the abbreviation to the key in the nowcasting hub. Information on the models and references is listed in \textcite[][Table 1]{Wolffram2023}.}
     \label{tab:app-covid-models}
 \end{table}
+
+\begin{table}
+    \centering
+    \input{plots/covid_nowcast/10_marginal_analysis}
+    \caption{Analysis of the nowcast and true differences for the lags 1, 7, and 14 days.}
+    \label{tab:app-covid-marginals}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{subfigure}[t]{\textwidth}
+        \input{plots/covid_nowcast/30_trending_ratios_lag_1.tex}
+    \caption{1 day.}
+    \end{subfigure}
+    \begin{subfigure}[t]{\textwidth}
+        \input{plots/covid_nowcast/30_trending_ratios_lag_1.tex}
+        \caption{14 days.}
+    \end{subfigure}
+    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the lag 1 and 14 days.
+The exclusion areas are rectangles centered on the zero points with width and height of 10\% of the quantile of the absolute values of nowcast and true values.
+}
+    \label{tab:app-covid-trending-ratios-lag-1-14}
+\end{table}
+
+\begin{figure}
+\centering
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_ENS-MEAN_4q_lag_1}
+    \caption{ENS-Mean, lag 1 day}\label{fig:app-covid-4q-ens-mean-1}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_ILM_4q_lag_1}
+    \caption{ILM, lag 1 day}\label{fig:app-covid-4q-ilm-1}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_RIVM_4q_lag_1}
+    \caption{RIVM, lag 1 day}\label{fig:app-covid-4q-rivm-1}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_RKI_4q_lag_1}
+    \caption{RKI, lag 1 day}\label{fig:app-covid-4q-rki-1}
+\end{subfigure}
+% lag 7
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_ENS-MEAN_4q_lag_7}
+    \caption{ENS-Mean, lag 7 days}\label{fig:app-covid-4q-ens-mean-7}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_ILM_4q_lag_7}
+    \caption{ILM, lag 7 days}\label{fig:app-covid-4q-ilm-7}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_RIVM_4q_lag_7}
+    \caption{RIVM, lag 7 days}\label{fig:app-covid-4q-rivm-7}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_RKI_4q_lag_7}
+    \caption{RKI, lag 7 days}\label{fig:app-covid-4q-rki-7}
+\end{subfigure}
+% lag 14
+\begin{subfigure}[t]{0.23\textwidth}%
+    \includegraphics{plots/covid_nowcast/30_ENS-MEAN_4q_lag_14}
+    \caption{ENS-Mean, lag 14 days}\label{fig:app-covid-4q-ens-mean-14}
+\end{subfigure}\hspace{0.01\textwidth}
+\begin{subfigure}[t]{0.23\textwidth}%
+    \includegraphics{plots/covid_nowcast/30_ILM_4q_lag_14}
+    \caption{ILM, lag 14 days}\label{fig:app-covid-4q-ilm-14}
+\end{subfigure}\hspace{0.01\textwidth}
+\begin{subfigure}[t]{0.23\textwidth}%
+    \includegraphics{plots/covid_nowcast/30_RIVM_4q_lag_14}
+    \caption{RIVM, lag 14 days}\label{fig:app-covid-4q-rivm-14}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{0.23\textwidth}
+    \includegraphics{plots/covid_nowcast/30_RKI_4q_lag_14}
+    \caption{RKI, lag 14 days}\label{fig:app-covid-4q-rki-14}
+\end{subfigure}%
+\caption{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the lags of one, seven, and 14 days. }
+\label{fig:app-covid-4q}
+\end{figure}

### 2024-02-29 sections/04-application.tex
First draft for nowcasting and eda sections

@@
 If hospitalization rates are rising, measures should be tightened, while falling rates might allow for loosening measures.
-Especially, asymmetries are of interest: Are some models better in recognizing a fall than a rise or vice versa?
-
-In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
-We report the trending for the lags 1, 7, and 14 days.
-While lag 1 assesses the short-term trending, lags 7 and 14 assess the medium-term trending.
-For the best models in terms of point evaluation, we also assess the trending with the conditional probability plot.
+Especially, asymmetries are of interest to assess whether some models are better in recognizing a fall than a rise or vice versa.
 
@@
 
+\subsubsection*{Results}
+
+In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
+We report the trending for the lags 1, 7, and 14 days.
+While lag 1 assesses the short-term trending, lags 7 and 14 assess the medium-term trending.
+The lags 7 and 14 are particularly interesting as they reflect a usual time span until new policy changes are taken.
+
+Before stating the results, we provide some background information on the marginal distributions of the true values and nowcasts for the different lags.
+Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} provides information on marginal statistics of the nowcasts and true values.
+Overall, the variability and general level of differences grows with the lag.
+The standard deviation of the differences between nowcasts and true values increases from roughly 300 for the lag 1 to 1,200 for the lag 7 and 2,000 for the lag 14.
+Similarly, the 10\%-quantile of differences increases.
+The 10\%-quantile is used for the exclusion areas in the trending assessment.
+The exclusion area is rectangular; a point falls within it, if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the differences.
+Thus, points are still included in the trending assessment if they are large in one dimension, but not in the other; thus ensuring, that large changes in, e.g., $\diffy$ are to be recognized by the nowcaster and vice versa.
+
+Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the lag 7 days.
+The trending ratios without exclusion area range from 0.72 to 0.85 for the lag of 7 days.
+The negative trending ratios are higher than the positive trending ratios for all models.
+For all models, the confidence intervals for the postive and negative trending ratio do not overlap, indicating that the trending ratios are indeed different.
+The 10\%-quantile exclusion areas have at most an influence of 0.03 on the ratios.
+The model with the highest trending ratio is the ILM model, the model with the lowest is the RKI model.
+The confidence intervals between all models overlap, reflecting the evaluation period lasting 159 days.
+The positive trending ratio imply a similar ranking of the models, while the negative ratio is second best for the RKI model.
+For the lags of 1 and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
+
+Figure~\ref{fig:app-covid-cond-prob-trending-ratio-7} shows the conditional trending plots and the trending ratio over exclusion area for the lag 7 days; the respective plots for the lags 1 day and 14 days are shown in Figure~\ref{fig:app-covid-cond-prob-trending-ratio-1-14}.
+Here, only the best four models in terms of point evaluation measures, ILM, RKI, RIVM and ENS-MED, are shown to keep the plots readable.
+If RKI or ILM issue a fall of the hospitalization rate, the probability of a fall is higher than if RIVM or ENS-MED issue a fall.
+The opposite is the case for a nowcasted increase of the hospitalization rate and the margin in probability is higher.
+Similar observations can be made for the lag of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
+For a lag of one day, the models' conditional trending ability difference is less pronounced (see Figure~\ref{fig:app-covid-cond-prob-1}).
+The RKI model is still less conclusive when issuing an increase of the hospitalization rate, while RIVM is most informative in that case.
+The curves cross for an issued fall, with ENS-MED being on top for issued falls above 250.
+
+The trending ratios are plotted for various exclusion areas in Figure~\ref{fig:app-covid-trending-ratio-7}.
+In general, the trending ratio increases with larger exclusion areas.
+While the RIVM and ENS-MED trending ratio evolve similarly, the RKI and ILM trending ratio get closer.
+For the lag of 1 day, the RKI trending ratio decreases with increasing exclusion area size,while the other models evolve increasingly (see Figure~\ref{fig:app-covid-trending-ratio-1}).
+For the lag of 14 days, all trending ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-trending-ratio-14}).
+
 \begin{table}
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the lag 7 days.
-The exclusion areas are rectangles centered on the zero points with width and height of 10\% of the quantile of the absolute values of nowcast and true values.
+    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[l]$, and negative trending ratio $\accml[l]$ for the models with and without exclusion areas for the lag 7 days.
+The exclusion areas are rectangles centered on the zero points with width and height of 10\%-quantile of the absolute values of nowcast and true values.
 }
@@
 
+\begin{figure}
+    \centering
+%    \includegraphics{}
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
+    \caption{Conditional trending plot.}\label{fig:app-covid-cond-prob-7}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
+    \caption{Trending ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-trending-ratio-7}
+    \end{subfigure}
+    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM and ENS-MED for the lag 7 days.}
+    \label{fig:app-covid-cond-prob-trending-ratio-7}
+\end{figure}
+
+\subsubsection*{Discussion}
+
+For all lags, the influence of the exclusion area on the 10\%-quantile level is small.
+The trending ratio is changes at most by 0.03, for example, for the EPI model with $\acc^{-, 14}$.
+The exclusion areas are thus not crucial for the trending assessment in the case of the nowcasts of the seven-day hospitalization rate.
+The lower bound of confidence intervals is at least 0.68 for all models, indicating that they perform better than random guessing the trend.
+
+Trending assessment is unattached to point evaluation measures.
+RKI is among the best in terms of point evaluation measures, but performs worse in terms of trending assessment.
+The assessment of asymmetry in the conditional trending plots is crucial for the interpretation of the trending ratios, with the RKI model being the most prominent example.
+
+In general, a larger training size would be beneficial for the assessment of the models' performance.
+For some models, more data is accessible, but we stick with the study protocol that uses an evaluation period of 159 days.
+
 \newpage
@@
 
-\textcite{Rostami-Tabar2023}
+In a second example, we consider the forecasting of the hourly number of arrivals in a large emergency department.
+Forecasting the number of arrivals in an emergency department is crucial for the planning of staff and resources.
+In \textcite{Rostami-Tabar2023}, various different models are employed on the task of forecasting the hourly number of arrivals in a large emergency department.
+Every 12 hours, the models issue hourly forecasts for the next 48 hours.
+Thus, the management can take measures according to the expected number of arrivals, for example, through redeploying staff and reconfiguring units.
+The models issue probabilistic quantile forecasts which are evaluated through RMSE, pinball loss, pinball skill scores and PIT-histograms.
+The models are trained on data from April 1, 2014 to February 28, 2018 and evaluated on data from March 1, 2018 to February 28, 2019.
+For further notes on the models and the evaluation, we refer to \textcite{Rostami-Tabar2023}.
+From the issued forecasts, we use the mean as point forecast for the trending assessment.
+We use the forecasts issued at the first time point for every target time.
+Thus, the forecasts are issued 36 to 48 hours ahead of the target time and the emergency department management has time to adjust the measures according to the expected number of arrivals.
+As we consider only the forecasts of at least 36 hours ahead, we restrict the evaluation period to March 2, 2018 at 12:00 to February 28, 2019 at 23:00, comprising 8,724 hours.
 
-Paper setup:
+In this setup, trending assessment is a simple and intuitive way to assess the models' performance.
+It is easy for the management to understand and implement, as simple comparisons of the expected workload to an hour in the recent past can be made.
+If, for example, the staff was near capacity in the last shift and an increase in the number of arrivals is expected, the management can take measures to adjust the workload.
 
-\begin{itemize}
-    \item (Probabilistic) forecast of the hourly number of arrivals in large emergency department (take expectation as point forecast)
-    \item Forecasts are produced 48 hours ahead on a 12-hour-rolling basis; background: \enquote{It enables planners to match ED staff to the number of arrivals, redeploy staff, and reconfigure units.}
-    \item training data: 2014-04-01 to 2018-2-28
-    \item evaluation from 2018-03-01 to 2019-02-28
-    \item Evaluation in Paper through RMSE; Pinball Loss and PIT-histograms
-    \item consider several lags:
-    \begin{itemize}
-        \item 72-hours-lag: Is the forecast showing the correct change compared to the most recent observation (problem: strong weekly structure (i.e., Monday and Saturday highest) makes it easy to predict right direction)
-        \item 7-days-lag: Correct trend compared to the last shift of same hour and day $\rightarrow$ change policies compared to that
-        \item Take first issued forecast for every target time
-    \end{itemize}
-\end{itemize}
+The number of arrivals has a strong weekly and daily pattern.
+Thus, we consider the lags of 72 hours, which is the last completely observed shift of the same hour of day, and 7 days, which is the last shift of the same hour and day.
+Table~\ref{tab:app-eda-point-evaluation} lists the point evaluation measures for the models.
+The best performing models are the NBI-2 and Poisson-2 model in terms of both RMSE and MAE.
+More than 8,600 forecasts are available for all of the models, with differences in the number due to missing values on four afternoons in 2018.
+Note that the reported values for the RMSE differ from the values in \textcite{Rostami-Tabar2023}.
+In contrast to their work, we use only the forecast data at least 36 hours ahead and not the entire forecast data for evaluation.
 
@@
 \centering
-\input{plots/ed_arrival/accuracies.tex}
-\caption{Accuracies (without exclusion area) and RMSE for the considered models in \textcite{Rostami-Tabar2023}}	
+\input{plots/ed_arrival/00_point_evaluation_measures}
+\caption{Point evaluation measures for the models.
+The smaller count for some models stems from missing forecast for a few hours in the middle of the evaluation period.}\label{tab:app-eda-point-evaluation}
+\end{table}
+
+\subsubsection*{Results}
+
+Table~\ref{tab:app-eda-marginals} analyzes the differences' marginal distributions for the forecasts and true values for the lags of 3 and 7 days.
+Note that the diffences are in line with Section~\ref{sec:notation} defined as difference between forecasted mean and true value of 3 and 7 days before, as the true value is available when issuing the forecast.
+The fraction of positive differences varies between 0.39 and 0.63 for the lag of 3 days and between 0.37 and 0.63 for the lag of 7 days.
+The variability of differences decreases from lag 3 days to lag 7 days for most models, only for the ETS model it increases.
+The 10\%-quantile of the differences is between 0 and one for all models and lags.
+Thus, we exclude only values smaller than 1 from the trending assessment.
+The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.
+
+\begin{table}
+    \centering
+    \input{plots/ed_arrival/10_marginal_analysis}
+    \caption{Marginal analysis of the nowcast and true differences.
+The column (1), $l=l$ shows the fraction of values greater than zero for lag $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
+    \label{tab:app-eda-marginals}
+\end{table}
+
+Table~\ref{tab:app-eda-trending-ratios} lists the trending ratios for all models for the lags of 72 hours and 7 days.
+The trending ratios range from 0.68 to 0.84 for the lag of three days and from 0.68 to 0.82 for the lag of 7 days.
+The negative and positive trending ratios differ for all models and lags.
+For some models, for example, the GBM-2 model, the positive trending ratio is higher, for some models, for example, the tbats model, the negative trending ratio is higher.
+The confidence intervals width is at most 0.02 for the trending ratios and at most 0.03 for the positive and negative trending ratios.
+The positive trending ratio is higher for other models than the negative trending ratio.
+The models GBM-2, qreg-1 and Benchmark-1 have the highest positive trending ratio for the lag of 3 and seven days, while the models Poisson-2 and NBI-2 have the highest negative trending ratio.
+
+Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags 3 and 7 days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
+The conditional trending plots show similar courses for the two lags, while the curves are shifted downwards for the lag of seven days.
+The models relative trending ability evolves consistently for the two lags, with the NBI-2 and Poisson-2 model being indistinguishable.
+The GBM-2 model outperfoms the qreg-1 model for all $x$.
+The models NBI-2 and Poisson-2 have the highest trending ability for all negative values of $x$ and the lowest trending ability for all positive values of $x$.
+Benchmark-1 lies between the other models for all $x$.
+
+\begin{table}
+    \centering
+    \input{plots/ed_arrival/50_trending_ratio}
+    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the lags 72 hours and 7 days.}
+    \label{tab:app-eda-trending-ratios}
 \end{table}
@@
     \centering
-	\includegraphics{plots/ed_arrival/Cond_Prob_lag_3.pdf}
-        \caption{Conditional probability plot for ED arrival}
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_3}
+    \caption{Lag 3 days}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_7}
+    \caption{Lag 7 days}
+    \end{subfigure}
+    \caption{Conditional trending plots for the lags 3 and 7 days and the models with the best positive or negative trending ability. The plots of NBI-2 and Poisson-2 are indistinguishable.}
+    \label{fig:app-eda-cond-prob}
 \end{figure}
+
+\subsubsection*{Discussion}
+
+Trending ability is consistent for the two lags, with the models' relative trending ability evolving similarly for the two lags.
+The models' trending ability is generally higher for the smaller lag, but the differences are small and confidence intervals overlap.
+
+The trending is for all models different for positive and negative predicted directions of change.
+While some models, such as GBM-2 and qreg-1, have the highest positive trending ratio, others, such as Poisson-2 and NBI-2, have the highest negative trending ratio.
+Thus, the uncertainty of the models predicted change has to be assessed differently based on the direction.
+
+The example shows that trending assessment is detached from standard point evaluation measures.
+While the models with lowest RMSE, NBI-2 and Poisson-2, also have a high trending ability, three models with below average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive trending ability.

### 2024-02-29 sections/55-appendix-application-covid.tex
First draft for nowcasting and eda sections

--- 2024-02-23 sections/55-appendix-application-covid.tex
@@
 
-\begin{figure}
-    \centering
-    \begin{subfigure}[t]{.31\textwidth}
-        \missingfigure{KDE for true and nowcasts of lag 1.}
-        \caption{\ac{kde} for true values and nowcasts of lag 1.}
-    \end{subfigure}\hspace{0.01\textwidth}
-    \begin{subfigure}[t]{.31\textwidth}
-        \missingfigure{KDE for true and nowcasts of lag 7.}
-        \caption{\ac{kde} for true values and nowcasts of lag 7.}
-    \end{subfigure}\hspace{0.01\textwidth}
-    \begin{subfigure}[t]{.31\textwidth}
-        \missingfigure{KDE for true and nowcasts of lag 14.}
-        \caption{\ac{kde} for true values and nowcasts of lag 14.}
-    \end{subfigure}
-    \caption{\Ac{kde} for true values and nowcasts of lags 1, 7, and 14 to assess the distribution of values. \hl{AUSWERTUNG; 10 \% der Werte liegen innerhalb von ... um den Ursprung.}}
-    \label{fig:app-covid-kde}
-\end{figure}
+%\begin{figure}
+%    \centering
+%    \begin{subfigure}[t]{.48\textwidth}
+%        \includegraphics{plots/covid_nowcast/20_kde_lag_1}
+%        \caption{\ac{kde} for true values and nowcasts of lag 1.}
+%    \end{subfigure}\hspace{0.01\textwidth}
+%    \begin{subfigure}[t]{.48\textwidth}
+%        \includegraphics{plots/covid_nowcast/20_kde_lag_7}
+%        \caption{\ac{kde} for true values and nowcasts of lag 7.}
+%    \end{subfigure}\hspace{0.01\textwidth}
+%    \begin{subfigure}[t]{.48\textwidth}
+%        \includegraphics{plots/covid_nowcast/20_kde_lag_14}
+%        \caption{\ac{kde} for true values and nowcasts of lag 14.}
+%    \end{subfigure}
+%    \caption{\Ac{kde} for true values and nowcasts of lags 1, 7, and 14 to assess the distribution of values.
+\hl{AUSWERTUNG}.
+Exclusion areas based on the 10\%  quantile of absolute values are listed in Table~\ref{tab:app-covid-marginals}.}
+%    \label{fig:app-covid-kde}
+%\end{figure}
 
@@
     \end{tabular}
-    \caption{Matching of the abbreviation to the key in the nowcasting hub .Information on the models and references are listed in \textcite[][Table 1]{Wolffram2023}.}
+    \caption{Matching of the abbreviation to the key in the nowcasting hub.
+    Information on the models and references is listed in \textcite[][Table 1]{Wolffram2023}.}
     \label{tab:app-covid-models}
 \end{table}
+
+\begin{table}
+    \centering
+    \input{plots/covid_nowcast/10_marginal_analysis}
+    \caption{Analysis of the nowcast and true differences for the lags 1, 7, and 14 days.
+    The column (1), $l=l$ shows the number of values greater than zero for lag $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
+    \label{tab:app-covid-marginals}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{subfigure}[t]{\textwidth}
+        \input{plots/covid_nowcast/30_trending_ratios_lag_1}
+    \caption{1 day.}
+    \end{subfigure}
+    \begin{subfigure}[t]{\textwidth}
+        \input{plots/covid_nowcast/30_trending_ratios_lag_14}
+        \caption{14 days.}
+    \end{subfigure}
+    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the lag 1 and 14 days.
+The exclusion areas are rectangles centered on the zero points with width and height of 10\% of the quantile of the absolute values of nowcast and true values.
+}
+    \label{tab:app-covid-trending-ratios-lag-1-14}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{plots/covid_nowcast/30_4q_plots}
+\caption{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the lags of one, seven, and 14 days. }
+\label{fig:app-covid-4q}
+\end{figure}
+
+\begin{figure}
+    \centering
+%    \includegraphics{}
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
+    \caption{Conditional trending plot for lag 1.}\label{fig:app-covid-cond-prob-1}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_14}
+    \caption{Conditional trending plot for lag 1.}\label{fig:app-covid-cond-prob-14}
+    \end{subfigure}
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_1}
+    \caption{Trending ratio over exclusion area size in $\diffx$ for lag 1.}\label{fig:app-covid-trending-ratio-1}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_14}
+    \caption{Trending ratio over exclusion area size in $\diffx$ for lag 1.}\label{fig:app-covid-trending-ratio-14}
+    \end{subfigure}
+    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM and ENS-MED for the lag 7 days.}
+    \label{fig:app-covid-cond-prob-trending-ratio-1-14}
+\end{figure}

### 2024-02-29 sections/04-application.tex
Update on Overleaf.

--- 2024-02-27 sections/04-application.tex
@@
 
-During the COVID 19 pandemic, the need for reliable and timely nowcasts of pandemic's development has become apparent.
-In Germany, the seven-day hospitalization rate has been established as a central steering measure for the COVID measures in November 2021 and the imposement of severe public restrictions were taken based on it~\parencite{RobertKochInstitute2021}.
+During the COVID-19 pandemic, the need for reliable and timely nowcasts of the pandemic's development has become apparent.
+In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\parencite{RobertKochInstitute2021}.
 The Robert Koch Institute (RKI) provides daily reports on the number of new hospitalizations.
-However, these reports are subject to delays and revisions rooted in two different sources.
-The first source is technical delays in the reporting process, for example, due to different authorities passing on the data to the RKI~\parencite{RobertKochInstitute2024}.
+However, these reports are subject to delays and revisions in two sources.
+The first source is technical delays in the reporting process, for example, due to different authorities passing the data to the RKI~\parencite{RobertKochInstitute2024}.
 The second, more systematic source is the structure of the seven-day hospitalization rate.
-To a given date, all the cases are allocated, whose first positive test result was on that date and who were hospitalized in relation to the disease in the following.
-The average count of cases on the given date and six days before of $100,000$ inhabitants is the seven-day hospitalization rate.
-Thus, the final seven-day hospitalization rate can only be reported with significant delay of up to more than 70 days.
-Nevertheless, as the seven-day hospitalization rate was considered a major indicator for the pandemic's development, many organizations and institutions started to issue nowcasts of the seven-day hospitalization rate, including research teams and newspapers.
+To a given date, all the cases are allocated whose first positive test result was on that date and who were hospitalized in relation to the disease in the following.
+The seven-day hospitalization rate is the average count of cases per $100,000$ inhabitants on the given date and six days before.
+Thus, the final seven-day hospitalization rate can only be reported with a significant delay of up to more than 70 days.
+Nevertheless, as the seven-day hospitalization rate was considered a major indicator of the pandemic's development, many organizations and institutions started to issue nowcasts, including research teams and newspapers.
 To collect nowcasts of the seven-day hospitalization rate by different nowcast groups, the COVID19-Nowcasting-Hub was established~\parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
-The nowcasts contain the predictive mean, median and quantiles for the seven-day hospitalization rate.
+The nowcasts contain the seven-day hospitalization rate's predictive mean, median, and quantiles.
 
@@
         \end{subfigure}
-    \caption{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021 to April 29, 2022 \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+    \caption{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
     The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
@@
 
-The data contains 8 nowcasts from scientific and public institutions, nowcast communities and a newspaper, using different input variables, calendar data, and length of training data.
-The model structures are diverse, including bayesian models, generalized additive models, and parametric bootstrapping.
-Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the overall structure of the model.
-Using the nowcasts, two ensemble methods are constructed by using mean or median of the ensembles.
+The data contains eight nowcasts from scientific and public institutions, nowcast communities, and a newspaper, using different input variables, calendar data, and length of training data.
+The model structures are diverse, including Bayesian models, generalized additive models, and parametric bootstrapping.
+Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the model's overall structure.
+Using the nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
 We denote them by ENS-MEAN and ENS-MED.
-In line with the initial study design, we consider the time span from November 22, 2021 to April 29, 2022 as evaluation period.
-In contrast to~\textcite{Wolffram2023} we use the newest data available on the true values and not the data from August 8, 2022.
+In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
+In contrast to~\textcite{Wolffram2023}, we use the newest data available on the true values, not the data from August 8, 2022.
 \unsure{Nochmal längere Evaluation?}
-We do not expand on the models' performance on specific regions or age groups in Germany and on the probabilistic nowcast assessment.
+We do not expand on the models' performance on specific regions or age groups in Germany and the probabilistic nowcast assessment.
 Figure~\ref{fig:app-covid-true-nowcast} displays the true and nowcast data for the evaluation period.
-The time comprises the forth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany lasting until May 28, 2022~\parencite{Tolksdorf2022}.
+The time comprises the fourth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany, lasting until May 28, 2022~\parencite{Tolksdorf2022}.
 Table~\ref{tab:app-covid-rmse} summarizes the point evaluation measures for the issued mean of the different models.
 The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\parencite[see][Tables A2, A3, and A4]{Wolffram2023}.
-The best performing models are the ILM and RKI model in terms of both RMSE and MAE.
-The ensemble methods perform worse than the best models in terms of the mean location.
+The best-performing models in terms of RMSE and MAE are the ILM and RKI models.
+The ensemble methods perform worse than the best models regarding the mean location.
 The performance of the models is diverse, with more than twice as high RMSE values for the worst models compared to the best models.
-Note that the high values for the EPI model could be driven by a very far off value at the end of the evaluation period.
+Note that the high values for the EPI model could be driven by an exceptionally far-off value at the end of the evaluation period.
 
-In addition to close inspection of the point evaluation measures, assessing trending of the nowcasts is crucial.
-To assess the impact of taken measures and the direction of curve, it is important to distinguish between rising and falling hospitalization rates.
-If hospitalization rates are rising, measures should be tightened, while falling rates might allow for loosening measures.
-Especially, asymmetries are of interest: Are some models better in recognizing a fall than a rise or vice versa?
-
-In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
-We report the trending for the lags 1, 7, and 14 days.
-While lag 1 assesses the short-term trending, lags 7 and 14 assess the medium-term trending.
-For the best models in terms of point evaluation, we also assess the trending with the conditional probability plot.
+In addition to close inspection of the point evaluation measures, assessing the trending of the nowcasts is crucial.
+To assess the impact of taken measures and the direction of the curve, it is essential to distinguish between rising and falling hospitalization rates.
+If hospitalization rates rise, measures should be tightened, while falling rates might allow for loosening measures.
+Especially, asymmetries are of interest to assess whether some models are better at recognizing a fall than a rise or vice versa.
 
@@
 
+\subsubsection*{Results}
+
+In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
+We report the trending for the lags 1, 7, and 14 days.
+While lag 1 assesses the short-term trending, lags 7 and 14 consider the medium-term trending.
+The lags 7 and 14 are particularly interesting, reflecting a usual period until new policy changes are taken.
+
+Before stating the results, we provide background information on the marginal distributions of the true values and nowcasts for the different lags.
+Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} provides information on marginal statistics of the nowcasts and true values.
+Overall, the variability and general level of differences grow with the lag.
+The standard deviation increases from roughly 300 for lag 1 to 1,200 for lag 7 and 2,000 for lag 14.
+Similarly, the 10\%-quantile of differences increases.
+The 10\%-quantile is used for the exclusion areas in the trending assessment.
+The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the differences.
+Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, e.g., $\diffy$ are to be recognized by the nowcast and vice versa.
+
+Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the lag of 7 days.
+The trending ratios without exclusion area range from 0.72 to 0.85 for the lag of 7 days.
+The negative trending ratios are higher than the positive trending ratios for all models.
+For all models, the confidence intervals for the positive and negative trending ratios do not overlap, indicating that the trending ratios are indeed different.
+The 10\%-quantile exclusion areas have, at most, an influence of 0.03 on the ratios.
+The model with the highest trending ratio is the ILM model, and the model with the lowest is the RKI model.
+The confidence intervals between all models overlap, reflecting the evaluation period lasting 159 days.
+The positive trending ratio implies a similar ranking of the models, while the negative ratio is second best for the RKI model.
+For the lags of 1 and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
+
+Figure~\ref{fig:app-covid-cond-prob-trending-ratio-7} shows the conditional trending plots and the trending ratio over the exclusion area for the lag seven days; the respective plots for the lags one day and 14 days are shown in Figure~\ref{fig:app-covid-cond-prob-trending-ratio-1-14}.
+Here, only the best models in point evaluation measures, ILM, RKI, RIVM, and ENS-MED, are shown to keep the plots readable.
+If RKI or ILM issues a fall in the hospitalization rate, the probability of a fall is higher than if RIVM or ENS-MED issues a fall.
+The opposite is the case for a nowcasted hospitalization rate increase, and the probability margin is higher. \todo{was soll das heißen?}
+Similar observations can be made for the lag of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
+For a lag of one day, the models' conditional trending ability difference is less pronounced (see Figure~\ref{fig:app-covid-cond-prob-1}).
+The RKI model is still less conclusive when issuing an increase in the hospitalization rate, while RIVM is most informative in that case.
+The curves cross for an issued fall, with ENS-MED being on top for issued falls above 250.
+
+The trending ratios are plotted for various exclusion areas in Figure~\ref{fig:app-covid-trending-ratio-7}.
+In general, the trending ratio increases with larger exclusion areas.
+While the RIVM and ENS-MED trending ratios evolve similarly, the RKI and ILM trending ratios get closer.
+For the lag of 1 day, the RKI trending ratio decreases with increasing exclusion area size while the other models evolve increasingly (see Figure~\ref{fig:app-covid-trending-ratio-1}).
+For the lag of 14 days, all trending ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-trending-ratio-14}).
+
 \begin{table}
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the lag 7 days.
-The exclusion areas are rectangles centered on the zero points with width and height of 10\% of the quantile of the absolute values of nowcast and true values.
+    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[l]$, and negative trending ratio $\accml[l]$ for the models with and without exclusion areas for the lag seven days.
+The exclusion areas are rectangles centered on the zero points with a width and height of 10\%-quantile of the absolute values of nowcast and true values.
 }
@@
 
+\begin{figure}
+    \centering
+%    \includegraphics{}
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
+    \caption{Conditional trending plot.}\label{fig:app-covid-cond-prob-7}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
+    \caption{Trending ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-trending-ratio-7}
+    \end{subfigure}
+    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the lag seven days.}
+    \label{fig:app-covid-cond-prob-trending-ratio-7}
+\end{figure}
+
+\subsubsection*{Discussion}
+
+For all lags, the influence of the exclusion area on the 10\%-quantile level is negligible.
+For example, the trending ratio changes at most by 0.03 for the EPI model with $\acc^{-, 14}$.
+The exclusion areas are thus not crucial for the trending assessment in the case of the nowcasts of the seven-day hospitalization rate.
+The lower bound of confidence intervals is at least 0.68 for all models, indicating that they perform better than random guessing the trend.
+
+Trending assessment is unattached to point evaluation measures.
+RKI is among the best in point evaluation measures but performs worse in trending assessment.
+The assessment of asymmetry in the conditional trending plots is crucial for interpreting the trending ratios, with the RKI model being the most prominent example.
+
+A more extensive training size would be beneficial for assessing the models' performance.
+For some models, more data is accessible, but we stick with the study protocol that uses an evaluation period of 159 days.
+
 \newpage
@@
 
-\textcite{Rostami-Tabar2023}
+In a second example, we consider the forecasting of the hourly number of arrivals in a large emergency department.
+Forecasting the number of arrivals in an emergency department is crucial for planning staff and resources.
+In \textcite{Rostami-Tabar2023}, various models are employed to forecast the hourly number of arrivals in a large emergency department.
+Every 12 hours, the models issue hourly forecasts for the next 48 hours.
+Thus, the management can take measures according to the expected number of arrivals, for example, through redeploying staff and reconfiguring units.
+The models issue probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
+The models are trained on data from April 1, 2014, to February 28, 2018, and evaluated on data from March 1, 2018, to February 28, 2019.
+For further notes on the models and the evaluation, we refer to \textcite{Rostami-Tabar2023}.
+From the issued forecasts, we use the mean as a point forecast for the trending assessment.
+We use the forecasts issued at the first time point for every target time.
+Thus, the forecasts are issued 36 to 48 hours ahead of the target time, and the emergency department management has time to adjust the measures according to the expected number of arrivals.
+Considering only the forecasts of at least 36 hours ahead, we restrict the evaluation period to March 2, 2018, at noon, to February 28, 2019, at 23:00, comprising 8,724 hours.
 
-Paper setup:
+In this setup, trending assessment is a simple and intuitive way to assess the models' performance.
+It is easy for the management to understand and implement, as simple comparisons of the expected workload to an hour in the recent past can be made.
+If, for example, the staff was near capacity in the last shift and an increase in the number of arrivals is expected, the management can take measures to adjust the workload.
 
-\begin{itemize}
-    \item (Probabilistic) forecast of the hourly number of arrivals in large emergency department (take expectation as point forecast)
-    \item Forecasts are produced 48 hours ahead on a 12-hour-rolling basis; background: \enquote{It enables planners to match ED staff to the number of arrivals, redeploy staff, and reconfigure units.}
-    \item training data: 2014-04-01 to 2018-2-28
-    \item evaluation from 2018-03-01 to 2019-02-28
-    \item Evaluation in Paper through RMSE; Pinball Loss and PIT-histograms
-    \item consider several lags:
-    \begin{itemize}
-        \item 72-hours-lag: Is the forecast showing the correct change compared to the most recent observation (problem: strong weekly structure (i.e., Monday and Saturday highest) makes it easy to predict right direction)
-        \item 7-days-lag: Correct trend compared to the last shift of same hour and day $\rightarrow$ change policies compared to that
-        \item Take first issued forecast for every target time
-    \end{itemize}
-\end{itemize}
+The number of arrivals has a strong weekly and daily pattern.
+Thus, we consider the lags of 72 hours, the last already observed shift of the same hour of day, and seven days, the previous shift of the same hour and day.
+Table~\ref{tab:app-eda-point-evaluation} lists the point evaluation measures for the models.
+The best-performing models regarding RMSE and MAE are the NBI-2 and Poisson-2 models.
+More than 8,600 forecasts are available for all models, with differences in the number due to missing values on four afternoons in 2018.
+Note that the reported values for the RMSE differ from those in \textcite{Rostami-Tabar2023}.
+In contrast to their work, we use only the forecast data at least 36 hours ahead and not the entire forecast data for evaluation.
 
@@
 \centering
-\input{plots/ed_arrival/accuracies.tex}
-\caption{Accuracies (without exclusion area) and RMSE for the considered models in \textcite{Rostami-Tabar2023}}	
+\input{plots/ed_arrival/00_point_evaluation_measures}
+\caption{Point evaluation measures for the models.
+The smaller count for some models stems from missing forecasts for a few hours in the course of the evaluation period.}\label{tab:app-eda-point-evaluation}
+\end{table}
+
+\subsubsection*{Results}
+
+Table~\ref{tab:app-eda-marginals} analyzes the differences in marginal distributions for the forecasts and true values for the lags of 3 and 7 days.
+Note that the difference definition aligns with Section~\ref{sec:notation}, defined as the difference between the forecasted mean and true value of 3 and 7 days before, as the true value is available when issuing the forecast.
+The fraction of positive differences varies between 0.39 and 0.63 for the lag of 3 days and between 0.37 and 0.63 for the lag of 7 days.
+The variability of differences decreases for the larger lag for most models; only for the ETS model does it increase.
+The 10\%-quantile of the differences is between zero and one for all models and lags.
+Thus, we exclude only values smaller than one from the trending assessment.
+The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.
+
+\begin{table}
+    \centering
+    \input{plots/ed_arrival/10_marginal_analysis}
+    \caption{Marginal analysis of the nowcast and true differences.
+The column (1), $l=l$ shows the fraction of values greater than zero for lag $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
+    \label{tab:app-eda-marginals}
+\end{table}
+
+Table~\ref{tab:app-eda-trending-ratios} lists the trending ratios for all models for three and seven-day lags.
+The trending ratios range from 0.68 to 0.84 for a lag of three days and from 0.68 to 0.82 for seven days.
+The negative and positive trending ratios differ for all models and lags.
+For some models, for example, the GBM-2 model, the positive trending ratio is higher; for some models, for example, the tbats model, the negative trending ratio is higher.
+The confidence interval width is at most 0.02 for the trending ratios and at most 0.03 for the positive and negative trending ratios.
+The positive trending ratio is higher for other models than the negative trending ratio.
+The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for the lag of 3 and seven days, while Poisson-2 and NBI-2 have the highest negative trending ratio.
+
+Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags 3 and 7 days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
+The conditional trending plots show similar courses for the two lags, while the curves are shifted downwards for the lag of seven days.
+The model's relative trending ability evolves consistently for the two lags, with the NBI-2 and Poisson-2 models being indistinguishable.
+The GBM-2 model outperforms the qreg-1 model for all $x$.
+The models NBI-2 and Poisson-2 have the highest trending ability for all negative values of $x$ and the lowest trending ability for all positive values of $x$.
+Benchmark-1 lies between the other models for all $x$.
+
+\begin{table}
+    \centering
+    \input{plots/ed_arrival/50_trending_ratio}
+    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the lags 72 hours and seven days.}
+    \label{tab:app-eda-trending-ratios}
 \end{table}
@@
     \centering
-	\includegraphics{plots/ed_arrival/Cond_Prob_lag_3.pdf}
-        \caption{Conditional probability plot for ED arrival}
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_3}
+    \caption{Lag 3 days}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_7}
+    \caption{Lag 7 days}
+    \end{subfigure}
+    \caption{Conditional trending plots for the lags 3 and 7 days and the models with the best positive or negative trending ability. The plots of NBI-2 and Poisson-2 are indistinguishable.}
+    \label{fig:app-eda-cond-prob}
 \end{figure}
+
+\subsubsection*{Discussion}
+
+Trending ability is consistent for the two lags, with the models' relative trending ability evolving similarly for the two lags.
+The models' trending ability is generally higher for the smaller lag, but the differences are minor, and confidence intervals overlap.
+
+The trending differs for all models for positive and negative predicted change directions.
+While some models, such as GBM-2 and qreg-1, have the highest positive trending ratio, others, such as Poisson-2 and NBI-2, have the highest negative trending ratio.
+Thus, the uncertainty of the model's predicted change has to be assessed differently based on the direction.
+
+The example shows that trending assessment is detached from standard point evaluation measures.
+While the models with the lowest RMSE, NBI-2 and Poisson-2, also have a high trending ability, three models with below-average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive trending ability.

### 2024-03-01 paper.tex
Add draft of first part on covid nowcasting
First draft for nowcasting and eda sections
Add first draft for examples

@@
 %\usepackage{tabularray}
-\usepackage{etoolbox}
-\AtBeginEnvironment{tabular}{\small}
+\usepackage{floatrow}
+\DeclareFloatFont{tiny}{\tiny}% "scriptsize" is defined by floatrow, "tiny" not
+\floatsetup[table]{font=scriptsize} %https://tex.stackexchange.com/a/27100
 
@@
 \DeclareAcronym{cdf}{short=CDF, long=cumulative distribution function}
+\DeclareAcronym{nbp}{short=NBP, long=non-invasive blood pressure}
+\DeclareAcronym{abp}{short=ABP, long=invasive arterial blood pressure}
 
@@
 \newcommand{\accmeps}[1][\varepsilon]{\acceps[#1]^-}
+\newcommand{\accl}[1][l]{\mu^{#1}}
+\newcommand{\accpl}[1][l]{\acc^{+,#1}}
+\newcommand{\accml}[1][l]{\acc^{-,#1}}
 
@@
 \section{Application to real-world data} \label{sec:application}
+\todo{Adjust heading; add introduction to section}
 \input{sections/04-application}
@@
 
+\input{sections/55-appendix-application-covid}
+
 \end{document}

### 2024-03-01 sections/04-application.tex
Add first draft for examples

@@
-\subsection{Measurement data} \label{sec:application_measurement}
-
-\newpage
 \subsection{Covid nowcasting} \label{sec:application-covid}
@@
 
-\newpage
 \subsection{Forecasting emergency department arrivals}\label{sec:application-eda}
@@
 While the models with the lowest RMSE, NBI-2 and Poisson-2, also have a high trending ability, three models with below-average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive trending ability.
+
+\subsection{Invasive and non-invasive blood pressure monitoring} \label{sec:application_measurement}
+
+In the last, briefer example, we consider the trending assessment of measurement data.
+The data is from the MIMIC-III database, including various information on patients in critical care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}).
+The data is publicly available and also includes numerical measurement data such as heart rate, blood pressure, and oxigen saturation in a waveform database \parencites[][]{Moody2017}[available through][]{Goldberger2000}.
+
+For some patients, the data includes \ac{abp} and \ac{nbp} measurements.
+For an overview of blood pressure measurement methods, see \textcite{Saugel2014}.
+While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods.
+For critical patients, changes in the blood pressure can be crucial for the treatment.
+Thus, trending assessment can be performed in addition to standard accuracy analysis~\parencite[see, e.g., ]{Mostafa2020}.
+Thus, we assess the trending ability of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
+Of the 64,168 numerical data records in the database, 2,662 include at least one measurement of systolic \ac{abp} and \ac{nbp} and ...
+include at least measurements of \ac{abp} and \ac{nbp} at the same time.
+The data records generally include measurements for one patient.
+
+We consider the lags of one minute, five minutes, and 15 minutes for the trending assessment, as those are typical intervals of NBP measurements.
+The number of available records with measurements of \ac{abp} and \ac{nbp} for the lags is listed in Table~\ref{tab:app-mimic-data-summary}.
+
+\begin{table}
+    \centering
+    \caption{Data summary of the \ac{abp} and \ac{nbp} measurements for the lags one minute, five minutes, and 15 minutes.}
+    \label{tab:app-mimic-data-summary}
+\end{table}
+
+\subsubsection*{Results}
+
+Again, we exclude the 10\% of absolute differences in trending assessment.
+The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different lags are shown in Figure~\ref{fig:app-mimic-4q}.
+The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} at the same time for two consecutive times with the specified lags.
+For the systolic measurements, 290, 332, and 442 points are available for the lags of one, five, and 15 minutes; for the mean measurements 406, 430, and 542.
+
+The trending ratios including confidence intervals for the different lags are listed in Table~\ref{tab:app-mimic-trending-ratios}.
+For the measurements with lag one minute, the confidence intervals have lower bounds of 0.5 or slightly above.
+For larger lags, the trending ratio increases.
+The difference between positive and negative trending ratio is small for all types and lags, with overlapping confidence intervals in all cases.
+
+Figure~\ref{fig:app-mimic-cond-prob} shows the conditional trending plots for the different lags and the systolic and mean blood pressure measurements.
+It becomes apparent that the systolic measurements have a higher trending ability than the mean measurement, except for small negative predicted changes.
+This is in line the trending ratios, but the confidence intervals overlap.
+
+\begin{figure}
+    \centering
+    \includegraphics{plots/mimic/plot_4q}
+    \caption{Four-quadrant plots for the different lags and the systolic and mean blood pressure measurements.
+The upper row contains systolic measurements, the lower row mean measurements.
+The columns contain the lags one, five, and 15 minutes.}
+    \label{fig:app-mimic-4q}
+\end{figure}
+
+\begin{table}
+    \centering
+    \input{plots/mimic/trending_ratio}
+    \caption{Trending ratios for the different lags and the systolic and mean blood pressure measurements.}
+    \label{tab:app-mimic-trending-ratios}
+\end{table}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{.32\textwidth}
+        \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag1}
+        \caption{Lag one minute.}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \begin{subfigure}[t]{.32\textwidth}
+        \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag5}
+        \caption{Lag five minutes.}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \begin{subfigure}[t]{.32\textwidth}
+        \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag15}
+        \caption{Lag 15 minutes.}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the lags one, five, and 15 minutes. \hl{Short interpretation.}}
+    \label{fig:app-mimic-cond-prob}
+\end{figure}
+
+\subsubsection*{Discussion}
+
+The four-quadrant plots contain a considerable number of extreme points.
+It is not clearly distinguishable whether these points are due to measurement errors or actual extreme values.
+Some authors argue to exclude not only the measurements below 10\%-quantile of the absolute differences but also the points above the 90\%-quantile.
+We do not follow this approach here, as the extreme values are not necessarily measurement errors and could also be particularly relevant.
+
+The differences between positive and negative predicted changes are small in this example.
+The positive and negativ trending ratios have overlapping confidence intervals and the conditional trending plots do not contain prominent deviations in the course.
+This is in line with the four-quadrant plots, where no clear asymmetry is visible.
+
+The bootstrap confidence intervals are wide.
+For the trending ratio the width is around 0.1, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the lag of 1 minute.

### 2024-03-01 sections/04-application.tex
Update on Overleaf.

@@
 
-In the last, briefer example, we consider the trending assessment of measurement data.
+In the last briefer example, we consider the trending assessment of measurement data.
 The data is from the MIMIC-III database, including various information on patients in critical care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}).
-The data is publicly available and also includes numerical measurement data such as heart rate, blood pressure, and oxigen saturation in a waveform database \parencites[][]{Moody2017}[available through][]{Goldberger2000}.
+The data is publicly available and also includes numerical measurement data such as heart rate, blood pressure, and oxygen saturation in a waveform database \parencites[][]{Moody2017}[available through][]{Goldberger2000}.
 
@@
 While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods.
-For critical patients, changes in the blood pressure can be crucial for the treatment.
+For critical patients, changes in blood pressure can be crucial for the treatment.
 Thus, trending assessment can be performed in addition to standard accuracy analysis~\parencite[see, e.g., ]{Mostafa2020}.
@@
 We consider the lags of one minute, five minutes, and 15 minutes for the trending assessment, as those are typical intervals of NBP measurements.
-The number of available records with measurements of \ac{abp} and \ac{nbp} for the lags is listed in Table~\ref{tab:app-mimic-data-summary}.
+The number of available records with measurements of \ac{abp} and \ac{nbp} for the lags are listed in Table~\ref{tab:app-mimic-data-summary}.
 
@@
 The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} at the same time for two consecutive times with the specified lags.
-For the systolic measurements, 290, 332, and 442 points are available for the lags of one, five, and 15 minutes; for the mean measurements 406, 430, and 542.
+For the systolic measurements, 290, 332, and 442 points are available for the lags of one, five, and 15 minutes; for the mean measurements, 406, 430, and 542.
 
-The trending ratios including confidence intervals for the different lags are listed in Table~\ref{tab:app-mimic-trending-ratios}.
-For the measurements with lag one minute, the confidence intervals have lower bounds of 0.5 or slightly above.
+The trending ratios, including confidence intervals for the different lags, are listed in Table~\ref{tab:app-mimic-trending-ratios}.
+For the measurements with a lag of one minute, the confidence intervals have lower bounds of 0.5 or slightly above.
 For larger lags, the trending ratio increases.
-The difference between positive and negative trending ratio is small for all types and lags, with overlapping confidence intervals in all cases.
+The difference between positive and negative trending ratios is small for all types and lags, with overlapping confidence intervals in all cases.
 
@@
 It becomes apparent that the systolic measurements have a higher trending ability than the mean measurement, except for small negative predicted changes.
-This is in line the trending ratios, but the confidence intervals overlap.
+This is in line with the trending ratios, but the confidence intervals overlap.
 
@@
     \caption{Four-quadrant plots for the different lags and the systolic and mean blood pressure measurements.
-The upper row contains systolic measurements, the lower row mean measurements.
+The upper row contains systolic measurements and the lower row mean measurements.
 The columns contain the lags one, five, and 15 minutes.}
@@
 It is not clearly distinguishable whether these points are due to measurement errors or actual extreme values.
-Some authors argue to exclude not only the measurements below 10\%-quantile of the absolute differences but also the points above the 90\%-quantile.
+Some authors argue to exclude not only the measurements below the 10\%-quantile of the absolute differences but also the points above the 90\%-quantile.
 We do not follow this approach here, as the extreme values are not necessarily measurement errors and could also be particularly relevant.
@@
 The differences between positive and negative predicted changes are small in this example.
-The positive and negativ trending ratios have overlapping confidence intervals and the conditional trending plots do not contain prominent deviations in the course.
+The positive and negative trending ratios have overlapping confidence intervals, and the conditional trending plots do not contain prominent deviations in the course.
 This is in line with the four-quadrant plots, where no clear asymmetry is visible.
@@
 The bootstrap confidence intervals are wide.
-For the trending ratio the width is around 0.1, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the lag of 1 minute.
+For the trending ratio, the width is around 0.1, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the lag of 1 minute.

### 2024-03-01 sections/04-application.tex
Update on Overleaf.

@@
 To collect nowcasts of the seven-day hospitalization rate by different nowcast groups, the COVID19-Nowcasting-Hub was established~\parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
-The nowcasts contain the seven-day hospitalization rate's predictive mean, median, and quantiles.
-
+The nowcasts contain the seven-day hospitalization rate's predictive mean, median , and quantiles.
+\unsure{median}
 \begin{figure}
@@
 Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the model's overall structure.
-Using the nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
+Using the nowcasts, two ensemble methods are constructed using the ensembles' mean or for the large .
 We denote them by ENS-MEAN and ENS-MED.
@@
 For the lag of 14 days, all trending ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-trending-ratio-14}).
+
+\unsure{Generell bei diesem Abschnitt: man liest hier eigentlich nur Beobachtungen, kann man auch daraus Schlüsse ziehen? Kann man results und discussion verschmelzen?}
 
@@
 include at least measurements of \ac{abp} and \ac{nbp} at the same time.
+\unsure{... include ...?}
 The data records generally include measurements for one patient.
+\unsure{irgendwie wird die Datenstruktur nicht ganz klar beschrieben. Habe ich hier Datensätze von verschiedenen Patienten und bei jeem Datensatz gibt es dann fehlende Punkte von ABP oder NBP?}
 
@@
 Again, we exclude the 10\% of absolute differences in trending assessment.
+\unsure{vielleicht genauer: we exclude the smallest 10 \% of the absolute differences in the trend assessment.}
 The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different lags are shown in Figure~\ref{fig:app-mimic-4q}.

### 2024-03-03 paper.tex
Improve Chapter on Trending, exclude notation

@@
 \usepackage[short]{optidef} % for aligning linear programs
+\usepackage{nccmath} % for fleqn environment
 
@@
 \newcommand{\diffylt}{\mathbf{y}^{\Delta,l}_t}
-\newcommand{\diffxt}{x^{\Delta}_t}
-\newcommand{\diffyt}{y^{\Delta}_t}
+\newcommand{\diffxt}[1][t]{x^{\Delta}_{#1}}
+\newcommand{\diffyt}[1][t]{y^{\Delta}_{#1}}
 
@@
 
-\section{Application areas and Notation} \label{sec:notation}
-\input{sections/02-notation}
+%\section{Application areas and Notation} \label{sec:notation}
+%\input{sections/02-notation}
 
@@
 
-\section{Application to real-world data} \label{sec:application}
-\todo{Adjust heading; add introduction to section}
+\section{Application to measurement, nowcasting, and forecasting data} \label{sec:application}
 \input{sections/04-application}
+
+\section{Discussion and Conclusion}\label{sec:conclusion}
+
+\todo{Write conclusion.}
 
@@
 \appendix
-\section{Data generation for Section~\ref{sec:trending}}\label{sec:app-trending-data-generation}
+
 \input{sections/50-appendix-data-generation}

### 2024-03-03 sections/03-trending.tex
Update on Overleaf.
Improve Chapter on Trending, exclude notation

--- 2024-02-20 sections/03-trending.tex
@@
 In general, trending can be assessed for different lags $l$ separately, and the relevance of considered lags has to be determined based on the question at hand.
-In Section~\ref{sec:notation}, the time series of differences $\diffx$ and $\diffy$ were introduced.
-Whereas the computation of $\diffx$ differs in the different contexts, the trending assessment is similar.
-In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$.
-Similarly, we call a measurement, nowcast, or forecast a \textit{prediction} for the quantity of interest $y$. \discuss{Is prediction a good word?}
+Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal}.
+Before we introduce measures for assessing trending, we give remarks on computing the predicted change and notation in Section~\ref{subsec:notation}.
 The remainder of the section is structured as follows.
@@
 The approaches are extended to the trending setting in Section~\ref{subsec:trending-noise}.
-Section~\ref{subsec:trending-cond-prob} introduces a new graphical method of trending assessment.
-Section~\ref{subsec:trending-bootstrap} gives bootstrap methods for computing confidence intervals for the measures and conducts a simulation study on their properties.
+Section~\ref{subsec:trending-cond-prob} introduces a new graphical method of trending assessment and gives a brief introduction to bootstrap methods for computing confidence intervals for the measures.
 
+\subsection{Computing the predicted change and notation}\label{subsec:notation}
+
+As noted above, trending refers to observed and predicted changes for a time lag $l$.
+The observed change is straightforward to compute for measurements, nowcasts, and forecasts.
+Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values or gold standard measurements up to time $T$.
+The sequence of observed change is then given by the differences of values in $\mathbf{y}$ with lag $l$, that is
+\begin{equation}\label{eq:diffy}
+    \diffylt = (y_{t+l} - y_t) \quad \text{for}\ t = 1, \dots, T - l.
+\end{equation}
+As noted in the introduction, the definition of predicted change depends on the context.
+Conveyeing the computation for $y$ directly, the predicted change of signal would be the difference of the signal at time $t + l$ and the true value at $t$.
+This approach lags in three practical aspects.
+For nowcasting and forecasting, the true value at time $t$ might not be known with a considerable time lag.
+In the example in Section~\ref{sec:application-covid}, for example, the time lag is more than 80 days.
+Until the true values are made publicly available, the change cannot be computed and no actions can be taken based on it.
+Thus, it is more practically relevant in this setting, to use updated signal values for time $t$ as the reference value.
+
+Second, in the context of nowcasting or forecasting, several signals for the same time $t + l$ are often available with different times of generation.
+Thus, it has to be specified in notation, which signal generated at which time is used.
+We use the term \textit{issue time} for the generation time of a signal and \textit{target time} to denote the time the forecast applies to.
+
+For measurement, usually, the use of a test method is evaluated against a gold standard to test whether the test method could replace the gold standard.
+If the test method is assessed to be consistent with the gold standard, the gold standard measurement values are not available to compute the predicted change.
+Thus, the predicted change is computed based on the test method's values at time $t$ and $t + l$.
+
+Taking the three points together, the predicted change of a signal is based on the type of signal.
+To account for several signals for a forecast time $t$ at different issue times $\tau$, we denote those by $x_{t | \tau}$.
+For nowcasting, $t$ and $\tau$ are usually close and possibly negative, for forecasting, $t$ is usually larger than $\tau$.
+For measurements, this distinction is not necessary.
+Table~\ref{tab:notation} summarizes the notation for the different types of signals and the computation of the predicted change.
+
+\begin{table}
+    \centering
+    \begin{tabularx}{0.75\textwidth}{l X}
+        \toprule
+        Application & Predicted change computation \\
+        \midrule
+        Measurement & $(x_{t + l} - x_t)_{t=0}^T \label{eq:diffxl_measure} $\\
+        Nowcasting & $
+\diffxl =
+\begin{cases}
+(x_{t+l|t+l} -x_{t|t+l})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
+(x_{t+l|t+l} -y_{t})^T_{t=0} & \text{else}.
+\end{cases} \label{eq:diffxl_nowcasting}
+$ \\
+        Forecasting & $
+\diffxl =
+\begin{cases}
+(x_{t+l|t} - x_{t|t})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
+(x_{t+l|t} - y_{t})^T_{t=0}  & \text{else}.
+\end{cases} \label{eq:diffxl_forecasting}
+$\\
+        \bottomrule
+    \end{tabularx}
+    \caption{Computation of the predicted change in the different applications.
+The signals $x_{t | \tau}$ refer to values issued at time $\tau$ with a target time $t$.
+Note the different relations of issue and target time for nowcasts and forecasts.
+}
+    \label{tab:notation}
+\end{table}
+
+\todo{Signal/Prediction vereinheitlichen}
 \subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-basics}
 
-As noted above, a prediction is trending perfectly if all occurred directions of change are predicted correctly, that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+Throughout the remainder of the section, we use the notation of differences, $\diffxl$ and $\diffyl$.
+While the computation of $\diffx$ differs in the different contexts, the trending assessment is similar.
+In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$.
+As noted above, a signal is trending perfectly if all occurred directions of change are predicted correctly, that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
 Perfect trending is hardly the case in practice, but a weaker requirement is needed.
@@
 Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant plot for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
-The data generation is described in Appendix~\ref{sec:app-trending-data-generation} and will carry through this section.
+The data generation is described in Appendix~\ref{subsec:app-trending-data-generation} and will carry through this section.
 The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
@@
 It can be extended by including information on the date of difference in the point color.
-In Figure~\ref{fig:trending_basic_4q_sample_color}, blue points refer to small time indices and turn green for larger $t$.
+In Figure~\ref{fig:trending_basic_4q_sample_color}, blue points refer to earlier measurements and turn green for larger $t$.
+The lag is always one time step.
 However, four-quadrant plots become crowded for larger datasets, and the sequential information on the differences is complex to assess thoroughly.
@@
 
-Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for new elements of $\diffy$ and $\diffx$, respectively.
-\discuss{Überlegung: sollte genauer beschrieben sein, wofür die beiden ZV stehen?}
+Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes, respectively.
 Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$).
@@
 \begin{equation}
-    \acc (\diffx, \diffy) \coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0}}{\card{T}}.\label{eq:acc}
+    \acc (\diffx, \diffy) \coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0}}{\card{T}}.\label{eq:acc}
 \end{equation}
@@
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
-Such a $2 \times 2$ table of counts is often called a contingency table.
+Such a $2 \times 2$ table of counts is called a contingency table.
 It is often used in other scientific areas for evaluation, for example, in dichotomous forecasting or as a confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
@@
 In the classification context, these measures are known as positive or negative predictive value and hit rate or detection failure ratio in forecasting.
-The two give estimates of $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$, respectively.
-There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960}.
+They give the probability for a correct prediction of the direction of change, given that the signal direction is positive or negative, respectively.
+Thus, they measure the trending ability of the prediction for the positive and negative changes separately by cutting the considered data.
+In terms of probability, the two measures give estimates of $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$, respectively.
+
+There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
 Cohen's $\kappa$ is usually used to measure inter-rater agreement and takes into account the ratio of occurred agreement and the probability of agreement by chance.
 Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.
-Balanced outcome hereby refers to approximately equal occurrences of values smaller and greater zero for both $\diffx$ and $\diffy$.
+Balanced outcome hereby refers to approximately equal occurrences of values smaller and greater than zero for both $\diffx$ and $\diffy$.
 Unbalanced outcomes of the differences are unlikely in our setting as they are obtained from differencing time series data.
-Nevertheless, suppose in a four-quadrant plot, the positive and negative $\diffy$ are observed to differ widely.
-In that case, one should consider another as measures such as Cohen's $\kappa$ or those listed in \textcite[Table 3.3]{Jolliffe2012}.
+Nevertheless, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
+
 All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
-Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color}.
+A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
+For the trending ratio a rolling estimate with a backwards-looking window is given by
+\begin{equation*}
+    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^* = t-w + 1}^{t} \ind{\diffxt[t^*] \diffyt[t^*] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
+\end{equation*}
+Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
 Thus, the yearly course of the prediction's trending ratio can be detected.
-This seasonal behavior cannot be observed in the two four-quadrant plot as the high number of points overlays it.
+The trending ability of the signal has a strong yearly, sinodal seasonality with higher trending ability peak after a quarter of a year and a low point after three quarters.
+This seasonal behavior cannot be observed in the two four-quadrant plot as the high number of points overlays it or in the trending ratio.
 
@@
     \caption{Visualizations for data with a time-varying trending ratio.
-We defer information on the data generation process of the data in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{sec:app-trending-data-generation}).
+We defer information on the data generation process of the data in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{subsec:app-trending-data-generation}).
 The trending ratios on the whole data set are $\mu = 0.7577$ and with exclusion area $\diffx < \varepsilon$, $\mu_{1.0} = 0.7712$, respectively.
@@
 
-Missing values are a common problem, in particular in nowcasting and forecasting.
-This will, for example, become evident in the data study in Section~\ref{sec:application-covid}.
-Data pairs with missing values can be excluded from the computation of the measures.
-Nevertheless, missing values can considerably influence the results, mainly if they appear systematically.
-Predicting only in those times of certainty and omitting prediction in other times would improve evaluation measures.
-So, although the estimates can be computed, the time spans with missing values should be inspected visually.
+In applications, signal data is often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}).
+We refer to time steps, for which either signal or true values or both are not available, as missing values.
+Data pairs with missing values can be excluded from the computation of the measures, so that the time indices with missing values are excluded from $\mathcal{T}$.
+In general, the measures can be computed for data with missing values, if the missing values do not systematically occur.
+Systematic missing values would lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
+In this case, the measures should be computed for the time steps with available data and the results should be interpreted with caution.
+Thus, inspection of missing data, for example, visually, should accompany applications with missing data.
+In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}, only few missing values occur and we give information on the missing data by inspecting the missing data as lists.
 
-\subsection{Accounting for noise}\label{subsec:trending-noise}
+\subsection{Accounting for noise and non-informative small changes}\label{subsec:trending-noise}
 
-The above measures use only the information on a point's quadrant and neglect further details on its location within the quadrant, as displayed in a four-quadrant plot.
-However, points close to the zero point have intuitively less explanatory power.
+The above measures use only the information on a point's quadrant and neglect further details on its location within the quadrant.
+However, points close to the zero point have intuitively less explanatory power or are less reliable.
 If there is noise or non-systematic effects present in true value or predictions, those points are likely to be driven by noise to a large amount.
-Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the above methods.
+That means that the quadrant the point's assignment to a quadrant is driven by noise instead of a systematic trending ability of the signal.
+
+Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures}.
 Points within that zone are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
@@
 In general, it is likely that the models of predictions have a noise component and should thus be part of the exclusion area.
-We denote the measures accounting for an exclusion area by
+We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $e$ by
 \begin{align}
-    \acceps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{\abs{\diffxt} > \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\abs{\diffxt} > \varepsilon}}\label{eq:acceps}\\
-    \accpeps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > \varepsilon}} \label{eq:accpeps}\\
-    \accmeps (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < \varepsilon}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < \varepsilon }}\label{eq:accmeps}
+    \acceps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin e}}\label{eq:acceps}\\
+    \accpeps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin e}} \label{eq:accpeps}\\
+    \accmeps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin e}}\label{eq:accmeps}
 \end{align}
-The estimators can easily be adapted for other shapes of the exclusion area.
+
+\todo[inline]{Bsp. für einfache Exclusion area mit rectangle und resulting measures; Bsp. für andere shapes und wie dann $e$ gestaltet wäre}
+
+The measures are then estimators for probability of trending, given that the change is not in the exclusion area $e$, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$.
+The estimators can easily be adapted for various shapes of the exclusion area.
 Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
-A box-shaped exclusion area leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise.
-Thus, only point 1 is excluded.
-An exclusion area along one axis removes points in which one of the components could change the sign by a small amount of noise.
-This particularly suits prediction models, where small amounts of noise are inevitable.
-A cross-shaped exclusion area along both axes accounts for sign reversal in both components.
+A rectangular shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise.
+Points, where at least true value or signal are unlikely to be zero, are not excluded.
+In the example plot, only point 1 is excluded.
+An exclusion area along one axis, for example, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq 0 \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change the sign by a small amount of noise.
+This particularly suits signals, where small amounts of noise are inevitable.
+A cross-shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for likely sign reversal in both components.
 For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
-\todo{Berechnung $\mu$ für verschiedene Varianten für Beispiel im Anhang?}
-In addition to these straight shapes, exclusion areas could also be determined based on an error model of the components.
-For example, one could exclude points likely to be in another quadrant or within a specific quantile range of a distribution around the zero point, for example, exclude the points within the 5 \% quantile and 95 \% quantile of a standard normal distribution.
-These shapes are theoretically appealing, but the interpretation of the resulting estimators gets ambiguous.
+
+In addition to these straigt shapes, exclusion areas could have more difficult shapes, such as ellipsis.
+Such a shape could, for example, be obtained by using a multivariate normal error model around the zero point and excluding points, that exceed a certain likelihood of being obtained by chance from the zero point.
+These shapes are theoretically appealing, but the interpretation of the resulting estimators gets complex.
 Thus, we use the simple exclusion areas above.
@@
 In addition, the size can also be calculated as a proportion of the total variance or the total range of the data.
-A third approach is to visualize the trending ratio for different values of $\varepsilon$.
-Section~\ref{subsec:trending-example} gives an example of such a plot.
-As a base approach, we advocate using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$\varepsilon$ plot to visualize the course of the trending ratio over different exclusion area sizes.
+A third approach is to visualize the trending ratio for different values of $e$.
+In Section~\ref{sec:application}, there are various examples of such a plot.
+As a base approach, we advocate using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.
 
-\subsection{Conditional trending plot}\label{subsec:trending-cond-prob}
-\discuss{Allgemeine Frage: können wir hier von Wahrscheinlichkeit sprechen? }
-The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > \varepsilon)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < \varepsilon)$.
-To assess the trending ability of a prediction, an evaluation on a finder grid is of interest.
+\subsection{The conditional trending plot and confidence intervals based on bootstrapping}\label{subsec:trending-cond-prob}
+The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin e)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin e)$.
+These probabilites point a general picture, but still too coarse for local effects.
+To assess the local trending ability of a prediction, more insights might be gained by considering the conditional distribution.
 Whereas it would be possible to build analog measures on other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
@@
 
-The probability $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ cannot be computed by the counting approach, as for almost all $x \in \R$, no data is available.
-Instead, the smoothed approach of a multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$.
+A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
+\begin{align*}
+P(\diffxrv \diffyrv > 0 | \diffxrv = x) = \begin{cases}
+                                              \int_{-\infty}^0 \frac{f_{\diffxrv, \diffyrv}(x, y)}{f_{\diffxrv}(x)} \: \textrm{d} \: y & \text{if } x < 0, \\
+                                              \int_{0}^{\infty} \frac{f_{\diffxrv, \diffyrv}(x, y)}{f_{\diffxrv}(x)} \: \textrm{d} \: y & \text{if } x > 0, \\
+\end{cases}
+\end{align*}
+for $x \neq 0$ through a \ac{kde}.
 A comprehensive introduction into multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages~\parencite[e.g., for  Python in][]{Seabold2010}.
 The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ for all values of $x \in \R$.
-For kernels with infinite support, for example, the Gaussian kernel, this can convey a false impression of validity for values far beyond the values in $\diffx$.
-Thus, the plot should be limited to the core values of $\diffx$ without outliers.
+Note that for kernels with infinite support, for example, the Gaussian kernel, the plot should be limited to the core values of $\diffx$ without outliers to avoid estimates based on small subsets of the data.
 Besides the kernel, a bandwidth selector has to be chosen for a multivariate \ac{kde}.
@@
 Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| python package~\parencite{Seabold2010}.
+While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.
 The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
@@
 
-\subsection{Confidence intervals based on bootstrapping}\label{subsec:trending-bootstrap}
-
-Confidence intervals can be issued to account for the estimation uncertainty of the measures above.
-Bootstrapping confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
-Whereas classical confidence intervals are computed from parametric assumptions on the underlying data, new samples are drawn with replacement from the dataset in bootstrapping.
+\todo{Übergang zu Bootstrapping}
+Confidence intervals can be used to account for the estimation uncertainty of the measures above.
+Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
+New samples are drawn with replacement from the dataset in bootstrapping and are not based on parametric assumptions as for classical confidence intervals.
 The confidence interval is then computed based on the bootstrap samples.
@@
 The basic approach balances those two aims.
-We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time.
-We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
-The considered datasets are again outlined in Appendix~\ref{sec:app-trending-data-generation}, the first dataset with asymmetric dependence.
-In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\parencite{Virtanen2020}.
-The prescribed confidence level is 90 \%, and the number of bootstrap samples is $10,000$.
-The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
-The true values of the accuracy are computed based on a dataset of size $10^8$, yielding 0.7501 and 0.7700 for the two datasets.
-The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.
-\todo[inline]{Add evaluation when computation is finished}
-
-\begin{table}
-    \centering
-    \begin{subtable}{.48\textwidth}
-        \input{plots/illustrative_examples/bootstrap_butterfly_cis}
-        \caption{First dataset}
-    \end{subtable}\hspace{0.02\textwidth}
-    \begin{subtable}{.48\textwidth}
-        \input{plots/illustrative_examples/bootstrap_normal_cis}
-        \caption{Second dataset}
-    \end{subtable}
-    \caption{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$.
-The average width of the confidence interval is listed in brackets.}
-    \label{tab:trending_bootstrap}
-\end{table}
-
-\begin{table}
-    \centering
-    \begin{subfigure}{0.48\textwidth}
-        \includegraphics{plots/illustrative_examples/boxplot_comp_time_butterfly}
-        \caption{First dataset}
-    \end{subfigure}
-    \begin{subfigure}{0.48\textwidth}
-        \includegraphics{plots/illustrative_examples/boxplot_comp_time_normal}
-        \caption{Second dataset}
-    \end{subfigure}
-    \caption{Boxplot of the computation time of the different bootstrapping method and data set sizes $T$.
-The computation time refers to bootstrapping one confidence interval based upon $10,000$ values.
-Each boxplot reflects $10,000$ samples.
-The \ac{bca} method takes slightly longer than the other two, but the difference is negligible.}
-    \label{fig:trending_bootstrap_time}
-\end{table}
+We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
+For small datasets, the \ac{bca} holds the confidence level also for small data sets with a neglibly higher computation time than the other approaches.
+Thus, we use the \ac{bca} method for the confidence intervals in the applications in Section~\ref{sec:application}.
 
@@
 Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
+
+\subsection{Probabilistic Evaluation}
+
+\todo{Schreiben}

### 2024-03-03 sections/04-application.tex
Improve Chapter on Trending, exclude notation

--- 2024-03-01 sections/04-application.tex
@@
+In this section, we apply the trending assessment to three different applications.
+In the first application in Section~\ref{sec:application-covid}, we consider the nowcasting of the seven-day hospitalization rate for COVID-19 in Germany.
+Section~\ref{sec:application-eda} applies the trending assessment to forecast methods for the number of arrivals in a large emergency department.
+In the last application in Section~\ref{sec:application-measurement}, we consider the trending assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.
+
 \subsection{Covid nowcasting} \label{sec:application-covid}
@@
 To collect nowcasts of the seven-day hospitalization rate by different nowcast groups, the COVID19-Nowcasting-Hub was established~\parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
-The nowcasts contain the seven-day hospitalization rate's predictive mean, median, and quantiles.
+The nowcasts contain the seven-day hospitalization rate's predictive mean, median, and other quantiles.
 
@@
 If RKI or ILM issues a fall in the hospitalization rate, the probability of a fall is higher than if RIVM or ENS-MED issues a fall.
-The opposite is the case for a nowcasted hospitalization rate increase, and the probability margin is higher. \todo{was soll das heißen?}
+The opposite is the case for a nowcasted hospitalization rate increase, and the difference between the models' performance is higher.
 Similar observations can be made for the lag of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
@@
 Thus, we assess the trending ability of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
-Of the 64,168 numerical data records in the database, 2,662 include at least one measurement of systolic \ac{abp} and \ac{nbp} and ...
-include at least measurements of \ac{abp} and \ac{nbp} at the same time.
-The data records generally include measurements for one patient.
+The database contains 64,168 numerical data records.
+One data record includes all numerical measurements for one patient.
+In general, the measured signals vary in length, frequency and type of measurement.
+Thus, only a subset of the data contains measurements of \ac{abp} and \ac{nbp} at the same time.
+2,548 include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one measurement of systolic \ac{abp} and \ac{nbp} at the same time; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
 
@@
 
-Again, we exclude the 10\% of absolute differences in trending assessment.
+Again, we exclude the smallest 10\% of absolute differences in trending assessment.
 The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different lags are shown in Figure~\ref{fig:app-mimic-4q}.

### 2024-03-03 sections/50-appendix-data-generation.tex
Update bootstrapping
Improve Chapter on Trending, exclude notation

@@
+\section{Additional material on Section~\ref{sec:trending}}\label{sec:appendix-trending}
+
+\subsection{Data generation for Section~\ref{sec:trending}}\label{subsec:app-trending-data-generation}
+
 The first dataset is generated by sequentially generating $\diffx$ and $\diffy$.
@@
 \begin{equation*}
-    k_t = 0.75 + \sin(t / T^\star \cdot 2 \pi) / 4,
+    k_t = 0.75 + \sin(t / 365.25 \cdot 2 \pi) / 4,
 \end{equation*}
@@
 where $\Phi$ is a standard normal \ac{cdf}.
+
+The four-quadrant plots for the sample realizations of the data generation schemes are shown in Figure~\ref{fig:appendix_dgps}.
 
@@
     \end{subfigure}
-    \caption{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{sec:app-trending-data-generation}.
+    \caption{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.
 Although the first and second plot differ over time, their difference is not discernible in the plots.
@@
 \end{figure}
+
+\subsection{Simulation study on bootstrapping confidence intervals}\label{subsec:app-trending-bootstrap}
+We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
+The considered datasets are outlined in Appendix~\ref{subsec:app-trending-data-generation}, the first dataset with asymmetric dependence.
+In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\parencite{Virtanen2020}.
+The prescribed confidence level is 90 \%, and the number of bootstrap samples is $10,000$.
+The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
+The true values of the accuracy are computed based on a dataset of size $10^8$, yielding 0.7501 and 0.7700 for the two datasets.
+The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.
+For the small sample sizes up to $T = 168$, only the \ac{bca} method keeps the confidence interval size and yields slightly wider confidence intervals.
+The method's results do not differ for the larger sample sizes.
+The computation time for the \ac{bca} method is slightly larger than for the other methods, but all methods have a moderate computation time.
+
+\begin{table}
+    \centering
+    \begin{subtable}{.48\textwidth}
+        \input{plots/illustrative_examples/bootstrap_butterfly_cis}
+        \caption{First dataset}
+    \end{subtable}\hspace{0.02\textwidth}
+    \begin{subtable}{.48\textwidth}
+        \input{plots/illustrative_examples/bootstrap_normal_cis}
+        \caption{Second dataset}
+    \end{subtable}
+    \caption{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$.
+The average width of the confidence interval is listed in brackets.}
+    \label{tab:trending_bootstrap}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{subfigure}{0.48\textwidth}
+        \includegraphics{plots/illustrative_examples/boxplot_comp_time_butterfly}
+        \caption{First dataset}
+    \end{subfigure}
+    \begin{subfigure}{0.48\textwidth}
+        \includegraphics{plots/illustrative_examples/boxplot_comp_time_normal}
+        \caption{Second dataset}
+    \end{subfigure}
+    \caption{Boxplot of the computation time of the different bootstrapping method and data set sizes $T$.
+The computation time refers to bootstrapping one confidence interval based upon $10,000$ values.
+Each boxplot reflects $10,000$ samples.
+The \ac{bca} method takes slightly longer than the other two, but the difference is negligible.}
+    \label{fig:trending_bootstrap_time}
+\end{table}

### 2024-03-03 sections/03-trending.tex
Update on Overleaf.

@@
 Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal}.
-Before we introduce measures for assessing trending, we give remarks on computing the predicted change and notation in Section~\ref{subsec:notation}.
+Before introducing measures for assessing trending, we give remarks on computing the predicted change and notation in Section~\ref{subsec:notation}.
 The remainder of the section is structured as follows.
@@
 The approaches are extended to the trending setting in Section~\ref{subsec:trending-noise}.
-Section~\ref{subsec:trending-cond-prob} introduces a new graphical method of trending assessment and gives a brief introduction to bootstrap methods for computing confidence intervals for the measures.
+Section~\ref{subsec:trending-cond-prob} introduces a new graphical method of trending assessment and briefly introduces bootstrap methods for computing confidence intervals for the measures.
 
@@
 As noted in the introduction, the definition of predicted change depends on the context.
-Conveyeing the computation for $y$ directly, the predicted change of signal would be the difference of the signal at time $t + l$ and the true value at $t$.
+Conveying the computation for $y$ directly, the signal's predicted change would be the difference between the signal at time $t + l$ and the true value at $t$.
 This approach lags in three practical aspects.
@@
 In the example in Section~\ref{sec:application-covid}, for example, the time lag is more than 80 days.
-Until the true values are made publicly available, the change cannot be computed and no actions can be taken based on it.
-Thus, it is more practically relevant in this setting, to use updated signal values for time $t$ as the reference value.
+Until the true values are publicly available, the change cannot be computed, and no actions can be taken based on it.
+Thus, it is more practically relevant to use updated signal values for time $t$ as the reference value in this setting.
 
-Second, in the context of nowcasting or forecasting, several signals for the same time $t + l$ are often available with different times of generation.
-Thus, it has to be specified in notation, which signal generated at which time is used.
+Second, in the context of nowcasting or forecasting, several signals for the same time $t + l$ are often available at different times of generation.
+Thus, it has to be specified in notation which signal generated at which time is used.
 We use the term \textit{issue time} for the generation time of a signal and \textit{target time} to denote the time the forecast applies to.
 
-For measurement, usually, the use of a test method is evaluated against a gold standard to test whether the test method could replace the gold standard.
+For measurement, usually, a test method is evaluated against a gold standard to test whether the test method could replace the gold standard.
 If the test method is assessed to be consistent with the gold standard, the gold standard measurement values are not available to compute the predicted change.
@@
 
-Taking the three points together, the predicted change of a signal is based on the type of signal.
+Taking the three points together, the signal's predicted change is based on the type of signal.
 To account for several signals for a forecast time $t$ at different issue times $\tau$, we denote those by $x_{t | \tau}$.
-For nowcasting, $t$ and $\tau$ are usually close and possibly negative, for forecasting, $t$ is usually larger than $\tau$.
+For nowcasting, $t$ and $\tau$ are usually close and possibly negative; for forecasting, $t$ is usually larger than $\tau$.
 For measurements, this distinction is not necessary.
-Table~\ref{tab:notation} summarizes the notation for the different types of signals and the computation of the predicted change.
+Table~\ref{tab:notation} summarizes the notation for the different signal types and the predicted change computation.
 
@@
     \caption{Computation of the predicted change in the different applications.
-The signals $x_{t | \tau}$ refer to values issued at time $\tau$ with a target time $t$.
+The signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
 Note the different relations of issue and target time for nowcasts and forecasts.
@@
 Throughout the remainder of the section, we use the notation of differences, $\diffxl$ and $\diffyl$.
-While the computation of $\diffx$ differs in the different contexts, the trending assessment is similar.
+While the computation of $\diffx$ differs in different contexts, the trending assessment is similar.
 In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$.
-As noted above, a signal is trending perfectly if all occurred directions of change are predicted correctly, that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+As noted above, a signal is trending perfectly if all occurred directions of change are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
 Perfect trending is hardly the case in practice, but a weaker requirement is needed.
@@
 Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.
-Balanced outcome hereby refers to approximately equal occurrences of values smaller and greater than zero for both $\diffx$ and $\diffy$.
+Balanced outcome refers to approximately equal occurrences of values smaller and greater than zero for both $\diffx$ and $\diffy$.
 Unbalanced outcomes of the differences are unlikely in our setting as they are obtained from differencing time series data.
@@
 A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
-For the trending ratio a rolling estimate with a backwards-looking window is given by
+For the trending ratio, a rolling estimate with a backward-looking window is given by
 \begin{equation*}
@@
 Thus, the yearly course of the prediction's trending ratio can be detected.
-The trending ability of the signal has a strong yearly, sinodal seasonality with higher trending ability peak after a quarter of a year and a low point after three quarters.
-This seasonal behavior cannot be observed in the two four-quadrant plot as the high number of points overlays it or in the trending ratio.
+The trending ability of the signal has a strong yearly, sinus-shaped seasonality with higher trending ability peak after a quarter of a year and a low point after three quarters.
+This seasonal behavior cannot be observed in the two four-quadrant plots as the high number of points overlays it or in the trending ratio.
 
@@
 In applications, signal data is often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}).
-We refer to time steps, for which either signal or true values or both are not available, as missing values.
-Data pairs with missing values can be excluded from the computation of the measures, so that the time indices with missing values are excluded from $\mathcal{T}$.
-In general, the measures can be computed for data with missing values, if the missing values do not systematically occur.
+We refer to time steps, for which either signal or true values or both are unavailable, as missing values.
+Data pairs with missing values can be excluded from the computation of the measures so that the time indices with missing values are excluded from $\mathcal{T}$.
+In general, the measures can be computed for data with missing values if the missing values do not systematically occur.
 Systematic missing values would lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
-In this case, the measures should be computed for the time steps with available data and the results should be interpreted with caution.
-Thus, inspection of missing data, for example, visually, should accompany applications with missing data.
-In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}, only few missing values occur and we give information on the missing data by inspecting the missing data as lists.
+In this case, the measures should be computed for the time steps with available data, and the results should be interpreted cautiously.
+Thus, inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
+In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}, only a few missing values occur, and we give information on the missing data by inspecting the missing data as lists.
 
@@
 
-The measures are then estimators for probability of trending, given that the change is not in the exclusion area $e$, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$.
+The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $e$, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$.
 The estimators can easily be adapted for various shapes of the exclusion area.
 Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
-A rectangular shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise.
-Points, where at least true value or signal are unlikely to be zero, are not excluded.
+A rectangularly shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise.
+Points where at least the true value or signal is unlikely to be zero are not excluded.
 In the example plot, only point 1 is excluded.
 An exclusion area along one axis, for example, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq 0 \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change the sign by a small amount of noise.
-This particularly suits signals, where small amounts of noise are inevitable.
+This particularly suits signals where small amounts of noise are inevitable.
 A cross-shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for likely sign reversal in both components.
@@
 
-In addition to these straigt shapes, exclusion areas could have more difficult shapes, such as ellipsis.
+In addition to these straight shapes, exclusion areas such as ellipsis could have more difficult shapes.
 Such a shape could, for example, be obtained by using a multivariate normal error model around the zero point and excluding points, that exceed a certain likelihood of being obtained by chance from the zero point.

### 2024-03-03 sections/03-trending.tex
Update on Overleaf.

@@
 
-Second, in the context of nowcasting or forecasting, several signals for the same time $t + l$ are often available at different times of generation.
+Second, in the context of nowcasting or forecasting, several signals for the same time $t + l$ are often available at different generation times.
 Thus, it has to be specified in notation which signal generated at which time is used.
@@
 
-For measurement, usually, a test method is evaluated against a gold standard to test whether the test method could replace the gold standard.
+For measurement, a test method is usually evaluated against a gold standard to test whether the test method could replace the gold standard.
 If the test method is assessed to be consistent with the gold standard, the gold standard measurement values are not available to compute the predicted change.
@@
 
-In addition to these straight shapes, exclusion areas such as ellipsis could have more difficult shapes.
-Such a shape could, for example, be obtained by using a multivariate normal error model around the zero point and excluding points, that exceed a certain likelihood of being obtained by chance from the zero point.
-These shapes are theoretically appealing, but the interpretation of the resulting estimators gets complex.
+In addition to these straight shapes, exclusion areas could have more complex shapes, such as ellipses.
+Such a shape could, for example, be obtained by using a multivariate normal error model around the zero point and excluding points that exceed a certain likelihood of being produced by chance from the zero point.
+These shapes are theoretically appealing, but interpreting the resulting estimators gets complex.
 Thus, we use the simple exclusion areas above.
@@
 A third approach is to visualize the trending ratio for different values of $e$.
-In Section~\ref{sec:application}, there are various examples of such a plot.
+Section~\ref{sec:application} contains various examples of such a plot.
 As a base approach, we advocate using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.
@@
 The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin e)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin e)$.
-These probabilites point a general picture, but still too coarse for local effects.
-To assess the local trending ability of a prediction, more insights might be gained by considering the conditional distribution.
+These probabilities draw a general picture but might still be too coarse for local effects.
+More insights might be gained by considering the conditional distribution to assess the local trending ability of a prediction.
 Whereas it would be possible to build analog measures on other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
-In addition, the conditional trending plot facilitates the comparison of various methods in a single plot, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
+In addition, the conditional trending plot facilitates the comparison of various methods in a single plot and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
 
@@
 The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
-The second method, cross-validation least squares, needs long computation times while yielding small or no bandwidth results, even for the two relatively small datasets.
+The second method, cross-validation least squares, needs long computation times while yielding small or no bandwidth results, even for two relatively small datasets.
 The rule-of-thumb and cross-validation maximum likelihood both yield reasonable results at moderate computation times. 
@@
 Cross-validation least squares takes a considerably larger computation time.
-It does not converge for the first data set and yields a too small bandwidth for the second data set.
+It does not converge for the first data set and yields a bandwidth too small for the second data set.
 The rule of thumb is the fastest method but tends to oversmooth.
@@
 \todo{Übergang zu Bootstrapping}
-Confidence intervals can be used to account for the estimation uncertainty of the measures above.
+Confidence intervals can account for the estimation uncertainty of the measures above.
 Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
@@
 We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
-For small datasets, the \ac{bca} holds the confidence level also for small data sets with a neglibly higher computation time than the other approaches.
+For small datasets, the \ac{bca} holds the confidence level also for small data sets with a negligibly higher computation time than the other approaches.
 Thus, we use the \ac{bca} method for the confidence intervals in the applications in Section~\ref{sec:application}.
 
-In standard bootstrapping, the sample is assumed to be independent and identically distributed.
+The sample is assumed to be independent and identically distributed in standard bootstrapping.
 Thus, before applying bootstrapping methods, the strength of sequential dependence should be inspected, for example, by analyzing the autocorrelation and partial autocorrelation.

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 
-\todo{Signal/Prediction vereinheitlichen}
 \subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-basics}
@@
 Thus, the yearly course of the prediction's trending ratio can be detected.
-The trending ability of the signal has a strong yearly, sinus-shaped seasonality with higher trending ability peak after a quarter of a year and a low point after three quarters.
+The trending ability of the signal has a strong yearly, sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
 This seasonal behavior cannot be observed in the two four-quadrant plots as the high number of points overlays it or in the trending ratio.
@@
 
-\todo{Übergang zu Bootstrapping}
 Confidence intervals can account for the estimation uncertainty of the measures above.

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 In this section, we want to specify trending and lay out different methods of measuring and visualizing trending. 
-In general, trending can be assessed for different lags $l$ separately, and the relevance of considered lags has to be determined based on the question at hand.
-Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal}.
+In general, trending can be assessed for different lags $l$ separately, and the relevance of the considered lags has to be determined based on the question at hand.
+Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
 Before introducing measures for assessing trending, we give remarks on computing the predicted change and notation in Section~\ref{subsec:notation}.
-The remainder of the section is structured as follows.
 Section~\ref{subsec:trending-basics} introduces trending basics and visualizes trending in a synthetical example.
-Section~\ref{subsec:trending-measures} reviews conveyed quantification approaches from other fields.
+In Section~\ref{subsec:trending-measures}, various quantification methods from different disciplines and tasks are examined.
 The approaches are extended to the trending setting in Section~\ref{subsec:trending-noise}.

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 Section~\ref{subsec:trending-basics} introduces trending basics and visualizes trending in a synthetical example.
-In Section~\ref{subsec:trending-measures}, various quantification methods from different disciplines and tasks are examined.
-The approaches are extended to the trending setting in Section~\ref{subsec:trending-noise}.
-Section~\ref{subsec:trending-cond-prob} introduces a new graphical method of trending assessment and briefly introduces bootstrap methods for computing confidence intervals for the measures.
+In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
+The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise}.
+Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment and briefly introduces bootstrap methods for computing confidence intervals for the measures of the prior sections.
+Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:probabilistic}.
 
@@
 As noted above, trending refers to observed and predicted changes for a time lag $l$.
-The observed change is straightforward to compute for measurements, nowcasts, and forecasts.
+The \textit{observed} change is straightforward to compute for all types of signal.
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values or gold standard measurements up to time $T$.
@@
 \end{equation}
-As noted in the introduction, the definition of predicted change depends on the context.
-Conveying the computation for $y$ directly, the signal's predicted change would be the difference between the signal at time $t + l$ and the true value at $t$.
+
+As noted in the introduction, the definition of the predicted change depends on the context.
+Applying the computation for $y$ directly, the signal's predicted change would be the difference between the signal at time $t + l$ and the true value at $t$.
 This approach lags in three practical aspects.
-For nowcasting and forecasting, the true value at time $t$ might not be known with a considerable time lag.
-In the example in Section~\ref{sec:application-covid}, for example, the time lag is more than 80 days.
-Until the true values are publicly available, the change cannot be computed, and no actions can be taken based on it.
+For nowcasting and forecasting, the true value at time $t$ might not be known with a considerable time delay.
+In the nowcasting application of Section~\ref{sec:application-covid}, for example, the time lag is more than 80 days.
+Until the true values are publicly available, the change could not be computed, and no actions could be taken based on it.
 Thus, it is more practically relevant to use updated signal values for time $t$ as the reference value in this setting.
@@
 Taking the three points together, the signal's predicted change is based on the type of signal.
-To account for several signals for a forecast time $t$ at different issue times $\tau$, we denote those by $x_{t | \tau}$.
-For nowcasting, $t$ and $\tau$ are usually close and possibly negative; for forecasting, $t$ is usually larger than $\tau$.
+To account for several signals for a target time $t$ at different issue times $\tau$, we denote those by $x_{t | \tau}$.
+For nowcasting, $\tau$ is usually not considerably larger than $t$; while $t$ is usually at least $\tau$ in forecasting.
 For measurements, this distinction is not necessary.
@@
 
-Throughout the remainder of the section, we use the notation of differences, $\diffxl$ and $\diffyl$.
-While the computation of $\diffx$ differs in different contexts, the trending assessment is similar.
-In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$.
+While the computation of $\diffxl$ differs in different contexts, the trending assessment is similar.
+In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$ as defined in Table~\ref{tab:notation}.
 As noted above, a signal is trending perfectly if all occurred directions of change are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
@@
 \includegraphics{plots/illustrative_examples/4q_excl_box}
-\caption{Four-quadrant plot with rectangle exclusion area.}\label{fig:trending_basic_4q_excl_box}
+\caption{Four-quadrant plot with rectangular exclusion area.}\label{fig:trending_basic_4q_excl_box}
 \end{subfigure}\hspace{0.01\textwidth}%

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 Points in the green upper right and lower left quadrants reflect a correct trending for the respective time step, whereas points in the remaining red quadrants show incorrect predicted changes.
+
 Figure~\ref{fig:trending_basic_4q} displays a basic four-quadrant plot.
@@
 The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
+
 The four-quadrant plot is intuitive to interpret, and the magnitude and direction of change are shown simultaneously.
@@
 \includegraphics{plots/illustrative_examples/4q_excl_cross}
-\caption{Four-quadrant plot with crossed exclusion area.}\label{fig:trending_basic_4q_excl_cross}
+\caption{Four-quadrant plot with cross-shaped exclusion area.}\label{fig:trending_basic_4q_excl_cross}
 \end{subfigure}%
@@
 
-Analyzing the number of points in the green quadrants versus the red quadrants boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes, respectively.
+Analyzing the number of points in the green quadrants versus the red quadrants is a standard approach in trending assessment of measurement data \parencite{Critchley2010,Saugel2015}.
+It boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes, respectively.
 Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$).
@@
 \end{equation}
-We refer to this estimator as the trending ratio of the prediction and set $\mathcal{T} = \{1, \dots, T\}$.
+We refer to this estimator as the trending ratio of the prediction and set $\mathcal{T} = \{1, \dots, T-l\}$.
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
@@
 \includegraphics{plots/illustrative_examples/4Q_sample_with_time}
-\caption{The data is colored according to the time index $t$, the greener, the later.}\label{fig:trending_basic_4q_sample_color}
+\caption{The same data is colored according to the time index $t$, the greener, the later.}\label{fig:trending_basic_4q_sample_color}
 \end{subfigure}\hspace{0.01\textwidth}
@@
 We defer information on the data generation process of the data in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{subsec:app-trending-data-generation}).
-The trending ratios on the whole data set are $\mu = 0.7577$ and with exclusion area $\diffx < \varepsilon$, $\mu_{1.0} = 0.7712$, respectively.
+The trending ratios on the whole data set are $\mu = 0.7577$.
 Strong yearly seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 
+The use of those estimators could encourage the predictors to exploit imbalances in the number of positive and negative changes. 
+A large difference in $\sum_{\mathcal} \ind{\diffyt > 0}$ and $\sum_{\mathcal} \ind{\diffyt > 0}$ is unlikely in the trending setting as $\diffy$ is obtained from differencing time series data.
+Nevertheless, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
 There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
@@
 Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.
-Balanced outcome refers to approximately equal occurrences of values smaller and greater than zero for both $\diffx$ and $\diffy$.
-Unbalanced outcomes of the differences are unlikely in our setting as they are obtained from differencing time series data.
-Nevertheless, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
 
@@
 \begin{equation*}
-    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^* = t-w + 1}^{t} \ind{\diffxt[t^*] \diffyt[t^*] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
+    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
 \end{equation*}
@@
 \end{align}
-
-\todo[inline]{Bsp. für einfache Exclusion area mit rectangle und resulting measures; Bsp. für andere shapes und wie dann $e$ gestaltet wäre}
 

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 The trending ability of the signal has a strong yearly, sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
-This seasonal behavior cannot be observed in the two four-quadrant plots as the high number of points overlays it or in the trending ratio.
+This seasonal behavior cannot be distinguished in the two four-quadrant plots or in the trending ratio.
 
@@
 
-In applications, signal data is often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}).
+In applications, signal data is often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
 We refer to time steps, for which either signal or true values or both are unavailable, as missing values.
-Data pairs with missing values can be excluded from the computation of the measures so that the time indices with missing values are excluded from $\mathcal{T}$.
-In general, the measures can be computed for data with missing values if the missing values do not systematically occur.
-Systematic missing values would lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
-In this case, the measures should be computed for the time steps with available data, and the results should be interpreted cautiously.
-Thus, inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
-In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-forecasting}, only a few missing values occur, and we give information on the missing data by inspecting the missing data as lists.
+Data pairs with missing values can be excluded from the set $\mathcal{T}$ for the computation of the measures, and the measures are computed normally. 
+However, systematical missing values could lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
+In this case, the measures should be interpreted cautiously, and
+inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
+In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}, only a few missing values occur, and we give information on the missing data by inspecting the missing data as lists.
 

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 The use of those estimators could encourage the predictors to exploit imbalances in the number of positive and negative changes. 
-A large difference in $\sum_{\mathcal} \ind{\diffyt > 0}$ and $\sum_{\mathcal} \ind{\diffyt > 0}$ is unlikely in the trending setting as $\diffy$ is obtained from differencing time series data.
+A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t in \mathcal{T}} \ind{\diffyt > 0}$ is unlikely in the trending setting as $\diffy$ is obtained from differencing time series data.
 Nevertheless, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
@@
 However, points close to the zero point have intuitively less explanatory power or are less reliable.
-If there is noise or non-systematic effects present in true value or predictions, those points are likely to be driven by noise to a large amount.
-That means that the quadrant the point's assignment to a quadrant is driven by noise instead of a systematic trending ability of the signal.
+If there are noise or non-systematic effects present in the true values or predictions, a point's assignment to a quadrant can be driven by noise instead of a systematic trending ability of the signal.
+For points with at least one small coordinate, this happens more frequently.
 
 Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures}.
-Points within that zone are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
+Points within that area are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
 The exclusion area's shape can be chosen according to the noise characteristics in predictions and true values.
@@
 \end{align}
-
-The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $e$, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$.
-The estimators can easily be adapted for various shapes of the exclusion area.
+The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $e$, that is, for $\acceps$, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$.
+The estimators accept various shapes of the exclusion area.
 Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
@@
 Points where at least the true value or signal is unlikely to be zero are not excluded.
-In the example plot, only point 1 is excluded.
+In the example plot in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
 An exclusion area along one axis, for example, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq 0 \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change the sign by a small amount of noise.
@@
 A cross-shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for likely sign reversal in both components.
-For those two methods, points 1 and 7 or 1, 6, and 7 are excluded, respectively.
+For those two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
 
@@
 In most applications, it is advisable to determine the exact shape and size of the exclusion area based on domain knowledge or expert opinions.
-In addition, the size can also be calculated as a proportion of the total variance or the total range of the data.
+The size estimation can also be based on a proportion of the total variance or the total range of the data.
 A third approach is to visualize the trending ratio for different values of $e$.
 Section~\ref{sec:application} contains various examples of such a plot.
-As a base approach, we advocate using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.
+As a base approach, we suggest using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.
 

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
-As outlined in Section~\ref{sec:introduction}, we refer by trending to the consistency of an observed change with the change issued by another measurement, a nowcast, or a forecast.
-In this section, we want to specify trending and lay out different methods of measuring and visualizing trending. 
+As described in Section~\ref{sec:introduction}, we refer by trending to the consistency of an observed change with the change issued by another measurement, a nowcast, or a forecast.
+In this section, we want to specify trending and define different methods of measuring and visualizing trending. 
 In general, trending can be assessed for different lags $l$ separately, and the relevance of the considered lags has to be determined based on the question at hand.
 Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
-Before introducing measures for assessing trending, we give remarks on computing the predicted change and notation in Section~\ref{subsec:notation}.
-Section~\ref{subsec:trending-basics} introduces trending basics and visualizes trending in a synthetical example.
+Before introducing measures to evaluate trending, we make remarks on the calculation of predicted change and notation in Section~\ref{subsec:notation}.
+Section~\ref{subsec:trending-basics} introduces the basics of trending and visualizes trending in a synthetic example.
 In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
 The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise}.
-Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment and briefly introduces bootstrap methods for computing confidence intervals for the measures of the prior sections.
+Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment and briefly introduces bootstrap methods for computing confidence intervals for the measures in the previous sections.
 Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:probabilistic}.
@@
 As noted above, trending refers to observed and predicted changes for a time lag $l$.
-The \textit{observed} change is straightforward to compute for all types of signal.
+The \textit{observed} change is straightforward to compute for all types of signals.
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values or gold standard measurements up to time $T$.
@@
 For nowcasting and forecasting, the true value at time $t$ might not be known with a considerable time delay.
-In the nowcasting application of Section~\ref{sec:application-covid}, for example, the time lag is more than 80 days.
+In the nowcasting application of Section~\ref{sec:application-covid}, for example, the time delay is more than 80 days.
 Until the true values are publicly available, the change could not be computed, and no actions could be taken based on it.
@@
 
-\subsection{The conditional trending plot and confidence intervals based on bootstrapping}\label{subsec:trending-cond-prob}
+\subsection{Bootstrap confidence intervals and the conditional trending plot}\label{subsec:trending-cond-prob}
+
+Confidence intervals can account for the estimation uncertainty of the measures above.
+Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
+New samples are drawn with replacement from the dataset and are not based on parametric assumptions as for classical confidence intervals.
+The confidence interval is then computed based on the bootstrap samples.
+We examine three methods for bootstrapping here: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
+In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the bootstrap estimators' empirical distribution.
+The \textit{basic} approach computes the confidence interval based on the non-bootstrapped estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
+The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
+Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but needs smaller datasets for reasonable confidence intervals.
+The basic approach balances those two aims.
+We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
+The \ac{bca} is the only method holding the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
+Thus, we use the \ac{bca} method for the confidence intervals in the applications in Section~\ref{sec:application}.
+
+The sample is assumed to be independent and identically distributed in standard bootstrapping.
+Thus, before applying bootstrapping methods, the strength of sequential dependence should be inspected, for example, by analyzing the autocorrelation and partial autocorrelation.
+The simple bootstrapping methods above do not account for serial dependence. 
+Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
+
 The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin e)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin e)$.
@@
 Whereas it would be possible to build analog measures on other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
-In addition, the conditional trending plot facilitates the comparison of various methods in a single plot and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
+In addition, the plot facilitates the comparison of various methods in a single plot, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
+We refer to the plot as a conditional trending plot.
 
@@
 P(\diffxrv \diffyrv > 0 | \diffxrv = x) = \begin{cases}
-                                              \int_{-\infty}^0 \frac{f_{\diffxrv, \diffyrv}(x, y)}{f_{\diffxrv}(x)} \: \textrm{d} \: y & \text{if } x < 0, \\
-                                              \int_{0}^{\infty} \frac{f_{\diffxrv, \diffyrv}(x, y)}{f_{\diffxrv}(x)} \: \textrm{d} \: y & \text{if } x > 0, \\
+                                              \int_{-\infty}^0 \frac{f_{\diffxrv, \diffyrv}(x, y)}{f_{\diffxrv}(x)} \ \textrm{d} \: y & \text{if } x < 0, \\
+                                              \int_{0}^{\infty} \frac{f_{\diffxrv, \diffyrv}(x, y)}{f_{\diffxrv}(x)} \ \textrm{d} \: y & \text{if } x > 0, \\
 \end{cases}
@@
 The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ for all values of $x \in \R$.
+The multivariate \ac{kde} takes as modeling parameters a kernel and a bandwidth selector. 
 Note that for kernels with infinite support, for example, the Gaussian kernel, the plot should be limited to the core values of $\diffx$ without outliers to avoid estimates based on small subsets of the data.
-Besides the kernel, a bandwidth selector has to be chosen for a multivariate \ac{kde}.
-Various methods are available with different strengths and weaknesses.
+While the kernel is typically Gaussian, various methods for the bandwidth selector are available with different strengths and weaknesses.
+
 Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| python package~\parencite{Seabold2010}.
@@
 
-Confidence intervals can account for the estimation uncertainty of the measures above.
-Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
-New samples are drawn with replacement from the dataset in bootstrapping and are not based on parametric assumptions as for classical confidence intervals.
-The confidence interval is then computed based on the bootstrap samples.
-We examine three methods for bootstrapping here: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
-In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the bootstrap estimators' empirical distribution.
-The \textit{basic} approach computes the confidence interval based on the non-bootstrapped estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
-The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
-Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but needs smaller datasets for reasonable confidence intervals.
-The basic approach balances those two aims.
-We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
-For small datasets, the \ac{bca} holds the confidence level also for small data sets with a negligibly higher computation time than the other approaches.
-Thus, we use the \ac{bca} method for the confidence intervals in the applications in Section~\ref{sec:application}.
-
-The sample is assumed to be independent and identically distributed in standard bootstrapping.
-Thus, before applying bootstrapping methods, the strength of sequential dependence should be inspected, for example, by analyzing the autocorrelation and partial autocorrelation.
-The simple bootstrapping methods above do not account for serial dependence. 
-Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
-
 \subsection{Probabilistic Evaluation}

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 Second, in the context of nowcasting or forecasting, several signals for the same time $t + l$ are often available at different generation times.
-Thus, it has to be specified in notation which signal generated at which time is used.
+Thus, the generation time and forecast time have to be included in the notation of nowcasts and forecasts.
 We use the term \textit{issue time} for the generation time of a signal and \textit{target time} to denote the time the forecast applies to.
@@
 If the test method is assessed to be consistent with the gold standard, the gold standard measurement values are not available to compute the predicted change.
-Thus, the predicted change is computed based on the test method's values at time $t$ and $t + l$.
+Thus, the predicted change is computed on the basis of the test method's values at time $t$ and $t + l$.
 
@@
 For measurements, this distinction is not necessary.
-Table~\ref{tab:notation} summarizes the notation for the different signal types and the predicted change computation.
+Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
 
@@
     \caption{Computation of the predicted change in the different applications.
-The signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
+Signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
 Note the different relations of issue and target time for nowcasts and forecasts.
@@
 
-While the computation of $\diffxl$ differs in different contexts, the trending assessment is similar.
-In the following, we omit the difference's lag $l$ for the ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$ as defined in Table~\ref{tab:notation}.
-As noted above, a signal is trending perfectly if all occurred directions of change are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+Although the calculation of $\diffxl$ differs in different contexts, the trending assessment is similar.
+In the following, we omit the difference lag $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$ as defined in Table~\ref{tab:notation}.
+As noted above, a signal's trending is perfect if all the directions of change that occur are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
 Perfect trending is hardly the case in practice, but a weaker requirement is needed.
-Accordingly, when assessing the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
+Consequently, when evaluating the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
 A simple yet insightful method is the four-quadrant plot.
 For example, it is known in cardiac output measurement analysis~\parencite{Saugel2015,perrino1998intraoperative}. 
-Thereby, the occurred changes are plotted together with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
-Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays prediction data differences.
+Thus, the occured changes are plotted together with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
+Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
 Points in the green upper right and lower left quadrants reflect a correct trending for the respective time step, whereas points in the remaining red quadrants show incorrect predicted changes.
@@
 Points 2, 3, 5, and 6 show trending, whereas points 1, 4, and 7 count for anti-trending behavior.
-Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant plot for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
-The data generation is described in Appendix~\ref{subsec:app-trending-data-generation} and will carry through this section.
+Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant graph for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
+Data generation is described in Appendix~\ref{subsec:app-trending-data-generation} and will continue into this section.
 The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
@@
 The four-quadrant plot is intuitive to interpret, and the magnitude and direction of change are shown simultaneously.
-It can be extended by including information on the date of difference in the point color.
-In Figure~\ref{fig:trending_basic_4q_sample_color}, blue points refer to earlier measurements and turn green for larger $t$.
+It can be extended by including information on the date of difference in the color of the point.
+In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$.
 The lag is always one time step.
-However, four-quadrant plots become crowded for larger datasets, and the sequential information on the differences is complex to assess thoroughly.
+However, four-quadrant plots become crowded for larger datasets, and sequential information on the differences is complex to assess thoroughly.
 There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.
-However, they lack the clearness and intuition of the four-quadrant plot while adding no more information on the trending~\parencite{Saugel2015}.
+However, they lack the clarity and intuition of the four-quadrant plot, without adding any more information on trending~\parencite{Saugel2015}.
 
@@
 Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$).
-To estimate $P(\diffxrv \diffyrv > 0)$ it lies at hand to use
+The standard estimator for $P(\diffxrv \diffyrv > 0)$ is
 \begin{equation}
@@
 3]{Jolliffe2012}.
-There, a wide range of other methods is usually used to analyze further characteristics of contingency tables.
-Two simple measures focusing on specific areas of interest are the positive and negative trending ratio $\accp$ and $\accm$, respectively.
+There, a wide range of other methods are usually used to analyze further characteristics of contingency tables.
+Two simple measures that focus on specific areas of interest are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
 They are defined as
@@
 In the classification context, these measures are known as positive or negative predictive value and hit rate or detection failure ratio in forecasting.
-They give the probability for a correct prediction of the direction of change, given that the signal direction is positive or negative, respectively.
-Thus, they measure the trending ability of the prediction for the positive and negative changes separately by cutting the considered data.
+They give the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, respectively.
+Therefore, they measure the trending ability of the prediction for positive and negative changes separately by cutting the considered data.
 In terms of probability, the two measures give estimates of $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$, respectively.
 
-The use of those estimators could encourage the predictors to exploit imbalances in the number of positive and negative changes. 
-A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t in \mathcal{T}} \ind{\diffyt > 0}$ is unlikely in the trending setting as $\diffy$ is obtained from differencing time series data.
-Nevertheless, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
+The use of these estimators could encourage predictors to exploit imbalances in the number of positive and negative changes. 
+A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t in \mathcal{T}} \ind{\diffyt > 0}$ is unlikely in the trending setting, since $\diffy$ is obtained from differencing time series data.
+However, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
 There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
@@
 Thus, the yearly course of the prediction's trending ratio can be detected.
-The trending ability of the signal has a strong yearly, sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
+The trending ability of the signal has a strong yearly sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
 This seasonal behavior cannot be distinguished in the two four-quadrant plots or in the trending ratio.
@@
 The trending ratios on the whole data set are $\mu = 0.7577$.
-Strong yearly seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
-The green curve $k_t$ shows the probability of $\diffxt$ having the same sign as $\diffyt$ for each time step.
-The rolling estimates hang back as the windows are backward-looking.}
+The strong yearly seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
+The green curve $k_t$ shows the probability that $\diffxt$ has the same sign as $\diffyt$ for each time step.
+The rolling estimates are delayed as the windows look backward.}
 \end{figure}
 
-In applications, signal data is often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
-We refer to time steps, for which either signal or true values or both are unavailable, as missing values.
-Data pairs with missing values can be excluded from the set $\mathcal{T}$ for the computation of the measures, and the measures are computed normally. 
+In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
+We refer to time steps for which either signal or true values or both are unavailable as missing values.
+Pairs of data with missing values can be excluded from the set $\mathcal{T}$ for the computation of the measures, and the measures are normally calculated. 
 However, systematical missing values could lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 We refer to time steps for which either signal or true values or both are unavailable as missing values.
-Pairs of data with missing values can be excluded from the set $\mathcal{T}$ for the computation of the measures, and the measures are normally calculated. 
+Pairs of data with missing values can be excluded from the set $\mathcal{T}$ to calculate the measures. 
 However, systematical missing values could lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
-In this case, the measures should be interpreted cautiously, and
+In this case, measures should be interpreted with caution, and
 inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
-In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}, only a few missing values occur, and we give information on the missing data by inspecting the missing data as lists.
+In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}, only a few missing values occur, and we provide information on the missing data by inspecting the missing data as lists.
 
@@
 
-The above measures use only the information on a point's quadrant and neglect further details on its location within the quadrant.
+The above measures use only information on a point's quadrant and neglect further details on its location within the quadrant.
 However, points close to the zero point have intuitively less explanatory power or are less reliable.
@@
 Points within that area are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
-The exclusion area's shape can be chosen according to the noise characteristics in predictions and true values.
-In general, it is likely that the models of predictions have a noise component and should thus be part of the exclusion area.
+The shape of the exclusion area can be chosen according to the noise characteristics in the predictions and the true values.
+In general, it is likely that the prediction models have a noise component and should thus be part of the exclusion area.
 We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $e$ by
@@
 Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
-A rectangularly shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves points out that are small in both components and are thus likely to be driven away from the zero point by noise.
+A rectangularly shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and are thus likely to be driven away from the zero point by noise.
 Points where at least the true value or signal is unlikely to be zero are not excluded.
-In the example plot in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
-An exclusion area along one axis, for example, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq 0 \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change the sign by a small amount of noise.
+In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
+An exclusion area along one axis, for example, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq 0 \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
 This particularly suits signals where small amounts of noise are inevitable.
-A cross-shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for likely sign reversal in both components.
+A cross-shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for sign reversal in both components.
 For those two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.

### 2024-03-04 sections/03-trending.tex
Update on Overleaf.

@@
 Points within that area are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
-The shape of the exclusion area can be chosen according to the noise characteristics in the predictions and the true values.
+The shape of the exclusion area can be chosen according to the noise characteristics of predictions and true values.
 In general, it is likely that the prediction models have a noise component and should thus be part of the exclusion area.
@@
 Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
-A rectangularly shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and are thus likely to be driven away from the zero point by noise.
+A rectangular exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and therefore likely to be driven by noise.
 Points where at least the true value or signal is unlikely to be zero are not excluded.
@@
 This particularly suits signals where small amounts of noise are inevitable.
-A cross-shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for sign reversal in both components.
-For those two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
+A cross-shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
+For these two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
 
@@
 Such a shape could, for example, be obtained by using a multivariate normal error model around the zero point and excluding points that exceed a certain likelihood of being produced by chance from the zero point.
-These shapes are theoretically appealing, but interpreting the resulting estimators gets complex.
-Thus, we use the simple exclusion areas above.
+These shapes are theoretically appealing, but interpreting the resulting estimators becomes complex.
+Thus, we suggest using the above simple exclusion areas.
 
-In most applications, it is advisable to determine the exact shape and size of the exclusion area based on domain knowledge or expert opinions.
+In most applications, the shape and size of the exclusion area can be chosen based on domain knowledge or expert opinions.
 The size estimation can also be based on a proportion of the total variance or the total range of the data.
@@
 Section~\ref{sec:application} contains various examples of such a plot.
-As a base approach, we suggest using no exclusion area in the four-quadrant plot as the points do not complicate its interpretation and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.
+As a base approach, we suggest using no exclusion area in the four-quadrant plot, as the points do not complicate its interpretation, and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.
 
@@
 The confidence interval is then computed based on the bootstrap samples.
-We examine three methods for bootstrapping here: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
-In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the bootstrap estimators' empirical distribution.
-The \textit{basic} approach computes the confidence interval based on the non-bootstrapped estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
+We examine three methods for bootstrapping: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
+In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
+The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
 The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
-Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but needs smaller datasets for reasonable confidence intervals.
-The basic approach balances those two aims.
-We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
-The \ac{bca} is the only method holding the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
-Thus, we use the \ac{bca} method for the confidence intervals in the applications in Section~\ref{sec:application}.
+Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
+The basic approach balances these two aims.
+We compare the approaches in a small synthetic data study on their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
+The \ac{bca} is the only method that holds the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
+Therefore, we use the \ac{bca} method for confidence intervals in the applications in Section~\ref{sec:application}.
 
@@
 
-The above-defined estimators give information on the probabilites $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin e)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin e)$.
+The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin e)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin e)$.
 These probabilities draw a general picture but might still be too coarse for local effects.
 More insights might be gained by considering the conditional distribution to assess the local trending ability of a prediction.
-Whereas it would be possible to build analog measures on other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
-In addition, the plot facilitates the comparison of various methods in a single plot, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
+While it would be possible to build analog measures for other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
+Furthermore, the graph facilitates the comparison of various methods in a single graph, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
 We refer to the plot as a conditional trending plot.
@@
 for $x \neq 0$ through a \ac{kde}.
-A comprehensive introduction into multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages~\parencite[e.g., for  Python in][]{Seabold2010}.
+A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages~\parencite[e.g., for  Python in][]{Seabold2010}.
 The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ for all values of $x \in \R$.
-The multivariate \ac{kde} takes as modeling parameters a kernel and a bandwidth selector. 
+Multivariate \ac{kde} takes as modeling parameters a kernel and a bandwidth selector. 
+The kernel is usually Gaussian, and there are multiple methods for selecting the bandwidth, each with its own advantages and disadvantages.
 Note that for kernels with infinite support, for example, the Gaussian kernel, the plot should be limited to the core values of $\diffx$ without outliers to avoid estimates based on small subsets of the data.
-While the kernel is typically Gaussian, various methods for the bandwidth selector are available with different strengths and weaknesses.
 
-Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| python package~\parencite{Seabold2010}.
+Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\parencite{Seabold2010}.
 While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.
 The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
-The second method, cross-validation least squares, needs long computation times while yielding small or no bandwidth results, even for two relatively small datasets.
-The rule-of-thumb and cross-validation maximum likelihood both yield reasonable results at moderate computation times. 
-Further examples, including comparisons between methods concerning their trending ability, are available for the applications in Section~\ref{sec:application}.
+The second method, cross-validation least squares, requires long computation times while yielding small or no bandwidth results, even for two relatively small datasets.
+The rule-of-thumb and cross-validation maximum likelihood methods yield reasonable results at moderate computation times. 
+Further examples, including comparisons between methods regarding their trending ability, are available in Section~\ref{sec:application}.
 

### 2024-03-05 sections/01-introduction.tex
Update on Overleaf.

@@
-\begin{itemize}
-    \item Aktualität/Wichtigkeit herausstellen: 
-    \begin{itemize}
-        \item zwei Nowcasts mit gleichem MSE, aber unterschiedlichem Trending $\rightarrow$ Vergleich zwischen Nowcasts
-    \end{itemize}
-    \item Einordnung in Literatur
-    \item Contributions
-    \begin{itemize}
-        \item Argue, why trending is important and why other measures do not \enquote{detect} it
-        \item (Formalisation of Trending Ability)
-        \item Evaluation and review of existing Approaches (both measures and graphical assessment)
-        \item Development of new measure and graphical method to assess trending for measurements, nowcasts and forecasts
-        \item Application to various data examples from practice
-        \item Publish ready-to-use code
-    \end{itemize}
-    \item Überblick über Paper
-\end{itemize}
+The introduction of new medical measurement methods requires comprehensive evaluation and comparison with established standard measurement methods in order to ensure their validity and reliability.
+To detect the systematic difference between one measurement methods with a gold standard method, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
+In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline,}).
+Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizes the differences between the methods in relation to their mean value (\cite{bland1986statistical}).
 
-The paper is organized as follows.
-In Section \ref{sec:notation}, we compare the three possible application areas for trending assessment: measurement, forecasting and nowcasting, and define the time series of change according to the application for a given time series.
-In Section \ref{sec:trending}, we formalise the concept of trending independently of the application domain and investigate quantification approaches.
+One important aspect of testing the consistency of the new method with the gold standard is to adequately validate if the new method is measuring changes in the target variable in the right direction, when the gold standard detects changes during the measurement process.
+Ideally, the signs of the changes between two consecutive points in time (with a defined time interval) measured by the new method and the gold standard method should always be the same.
+In this case, we call the new method trending perfect in line with the golden Stadnard method and refer to this type of agreement as trendability.
+This type of consideration of the trend capability of a measurement method to the gold standard has also recently gained importance in the literature.
+In (\cite{hiraishi2021concordance}), the authors proposed a new concordance rate for the four-quadrant diagram, which is based on the multivariate normal distribution and takes into account the correlation between the data of the individual subjects.
+
+The above-described idea of trending is also of fundamental interest for the evaluation and comparison of forecasting and nowcasting methods.
+While measurement series are created by recording data using measuring instruments or direct observation, forecasting focuses on predicting the future based on historical data and its patterns.
+When forecasting at a specific point in time (forecast origin), the forecast value at the target time corresponds to a prediction of the change in the target variable between the forecast origin and the target time.
+This predicted change results from the difference between the forecast value at the forecast horizon and the current, actual value of the target variable at the forecast origin.
+A forecast is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change.
+In practice, a forecasting method is usually evaluated in terms of its prediction accuracy, which is defined by a certain loss function and measures the difference between the prediction and the actual outcome of the target variable; an overview of the different measures can be found in \cite{hyndman2006another}.
+Prediction accuracy can be measured by scale-dependent measures such as the Root Mean Square Error (RMSE) or measures based on percentage errors such as the Root Mean Square Percentage Error (RMSPE).
+All of these typical measures are based on the absolute difference between the prediction and the true values, locally or globally, and are not capable of assessing the trendability discussed above without further adjustment.
+
+Another area in which the detection of trends can be of interest is nowcasting.
+In contrast to conventional forecasting methods, nowcasting focuses on predicting the current status.
+Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
+Nowcasting methods are now widely used in other fields such as economics and medicine and focus on predictions for the present, the immediate future and the recent past (\cite{banbura2013now, bok2018macroeconomic}).
+Nowcasting can be used, for example, to predict important statistics such as the current economic situation (e.g.
+GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
+Nowcasting methods use indicators that are related to the target variable but can be collected at a higher frequency and published in a timely manner to produce and update early estimates for the target variable (e.g.
+GDP from the first quarter of 2024 ) during the relevant period (i.e.
+from January to March 2024) (\cite{castle2017forecasting}).
+In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
+The nowcasting method can be used, for example, to correct the daily Covid case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}).
+In this context, nowcasting is an estimation procedure in which the value of a target variable (e.g.
+from today) is estimated on the basis of current preliminary measurements, which is only finalised at a later point in time.
+At the time of the forecast, estimates of the target variable can be made for today and possibly for yesterday (if the value for yesterday has not yet been finalised); the predicted change in the target variable is then the difference between the two nowcast values.
+A nowcast is consistent with the development of the target variable if the difference in the actual value between today and yesterday, which is announced at a later date, has the same sign as the predicted change.
+It is obvious that a nowcast with strong trend capability provides crucial information on the development of the target variable (e.g.
+the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.
+Unfortunately, like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is hardly considered in the literature.
+
+As outlined above, the ability to take into account the trendability of a method depending on the area of application, be it measurements, predictions and forecasts, is of high importance.
+However, regardless of the application domain, most conventional method comparisons and evaluations consider local absolute differences in different ways without directly assessing whether the correct direction of change is output.
+In this paper, we aim to formalize trending and present various methods for measuring and visualizing trending that can be applied to the evaluation of trending ability of methods in measurements, nowcasts and forecasts.
+The presented methods will be applied to a variety of real data examples.
+
+The paper is structured as follows.
+In section \ref{sec:trending} we formalize the concept of trending independently of the application domain and investigate quantification approaches for noise-free and noisy data.
 In particular, we introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries.
@@
 We conclude our paper with the last section.
+The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at ...
+available.

### 2024-03-05 sections/03-trending.tex
Update on Overleaf.

@@
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
+\unsure{ist das Wort "observed" wirklich das was wir meinen? Stichwort: Rauschen}
 In general, trending can be assessed for different lags $l$ separately, and the relevance of the considered lags has to be determined based on the question at hand.
@@
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values or gold standard measurements up to time $T$.
+\unsure{Vielleicht schreiben wir hier noch ausfürhlich aus, was dieses y in drei Fällen sein sollte, oder ist das klar nach introduction?}
 The sequence of observed change is then given by the differences of values in $\mathbf{y}$ with lag $l$, that is
@@
 If the test method is assessed to be consistent with the gold standard, the gold standard measurement values are not available to compute the predicted change.
+\unsure{irgendwie für mich nicht klar formuliert: man unteruscht ja bei meausurement nur ob die gemessenen Änderungen von den zwei Methoden konsistent sind oder?}
 Thus, the predicted change is computed on the basis of the test method's values at time $t$ and $t + l$.
@@
 For nowcasting, $\tau$ is usually not considerably larger than $t$; while $t$ is usually at least $\tau$ in forecasting.
+\unsure{Nowcast macht Schätzung i.
+d.
+R.
+so lange, bis der echte Wert bekannt wird oder?}
 For measurements, this distinction is not necessary.
@@
 Analyzing the number of points in the green quadrants versus the red quadrants is a standard approach in trending assessment of measurement data \parencite{Critchley2010,Saugel2015}.
+\unsure{Quelle-Fehler}
 It boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes, respectively.
@@
 The use of these estimators could encourage predictors to exploit imbalances in the number of positive and negative changes. 
-A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t in \mathcal{T}} \ind{\diffyt > 0}$ is unlikely in the trending setting, since $\diffy$ is obtained from differencing time series data.
+A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ is unlikely in the trending setting, since $\diffy$ is obtained from differencing time series data.
+\unsure{hier stimmt was nicht}
 However, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
@@
 \begin{equation*}
-    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w} \quad t = w-1, \dots, T.\label{eq:acc_rolling}
+    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w}, \quad t = w-1, \dots, T.\label{eq:acc_rolling}
 \end{equation*}
@@
 \end{figure} 
-
+\unsure{was ist $\mu_1$ bei dem Plot, hinschreiben}
 In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
@@
 The shape of the exclusion area can be chosen according to the noise characteristics of predictions and true values.
-In general, it is likely that the prediction models have a noise component and should thus be part of the exclusion area.
+In general, it is likely that the prediction models have a noise component and should thus be part of the exclusion area.\unsure{vielleicht eine große Buchstabe für exclusion area?}
 We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $e$ by
@@
     \acceps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin e}}\label{eq:acceps}\\
-    \accpeps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin e}} \label{eq:accpeps}\\
+    \accpeps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin e}} \label{eq:accpeps}\\
     \accmeps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin e}}\label{eq:accmeps}
@@
 Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
-A rectangular exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and therefore likely to be driven by noise.
+A rectangular exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ ,for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and therefore likely to be driven by noise.
 Points where at least the true value or signal is unlikely to be zero are not excluded.
@@
 Section~\ref{sec:application} contains various examples of such a plot.
-As a base approach, we suggest using no exclusion area in the four-quadrant plot, as the points do not complicate its interpretation, and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.
+As a base approach, we suggest using no exclusion area in the four-quadrant plot, as the points do not complicate its interpretation, and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.\unsure{"as the points do not complicate its interpretation" ist das hier die "richtige" Begründung, dass wir kein exlclusion area betrachten?}
 
 \subsection{Bootstrap confidence intervals and the conditional trending plot}\label{subsec:trending-cond-prob}
-
+\unsure{Vielleicht doch zwei subsections?}
 Confidence intervals can account for the estimation uncertainty of the measures above.
@@
 for $x \neq 0$ through a \ac{kde}.
+
 A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages~\parencite[e.g., for  Python in][]{Seabold2010}.

### 2024-03-06 sections/03-trending.tex
Add draft for probabilistic forecasting

@@
         \midrule
-        Measurement & $(x_{t + l} - x_t)_{t=0}^T \label{eq:diffxl_measure} $\\
+        Measurement & $(x_{t + l} - x_t)_{t=1}^T \label{eq:diffxl_measure} $\\
         Nowcasting & $
@@
 \begin{cases}
-(x_{t+l|t+l} -x_{t|t+l})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
-(x_{t+l|t+l} -y_{t})^T_{t=0} & \text{else}.
+(x_{t+l|t+l} -x_{t|t+l})^T_{t=1} & \text{if } y_{t} \text{ is not known at time } t, \\
+(x_{t+l|t+l} -y_{t})^T_{t=1} & \text{else}.
 \end{cases} \label{eq:diffxl_nowcasting}
@@
 \begin{cases}
-(x_{t+l|t} - x_{t|t})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
-(x_{t+l|t} - y_{t})^T_{t=0}  & \text{else}.
+(x_{t+l|t} - x_{t|t})^T_{t=1} & \text{if } y_{t} \text{ is not known at time } t, \\
+(x_{t+l|t} - y_{t})^T_{t=1}  & \text{else}.
 \end{cases} \label{eq:diffxl_forecasting}
@@
 
-\subsection{Probabilistic Evaluation}
+\newpage
+\subsection{Probabilistic Evaluation}\label{subsec:probabilistic}
 
-\todo{Schreiben}
+In nowcasting and forecasting, there is a shift from point predictions to probabilistic predictions (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
+Probabilistic predictions include a point estimate and information on forecast uncertainty and quantiles simultaneously.
+In this section, we extend the trending assessment to probabilistic predictions.
+
+Probabilistic predictions can have the form of a \ac{cdf}, \ac{pdf}, or quantiles.
+The \ac{cdf} is the most general and can be used to derive the others, given they exist.
+Let us first assume that the prediction is a \ac{cdf} and the lag is large enough such that the true value is known when issuing the forecast or nowcast.
+See the end of the section for predictions given as quantiles or for unknown true values.
+
+Let $F_{t | \tau} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $\tau$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
+For the trending assessment, we contrast the probability of positive change to the occurrence of positive changes. 
+The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value. \unsure{Andere Formulierung für: one minus?}
+As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
+For nowcasts, the probability is calculated by
+\begin{equation*}
+    p_t = 1 - F_{t+l | t+l} (y_{t})\quad t = 1, \dots, T-l;
+\end{equation*}
+while for forecasts, it is given by
+\begin{equation*}
+    p_t = 1 - F_{t+l | t} (y_{t})\quad t = 1, \dots, T-l.
+\end{equation*}
+Let $z_t$ denote the indicator that the occurred change at time $t$ is positive, that is,
+\begin{equation*}
+    z_t = \ind{\diffyt > 0} \quad t = 1, \dots, T-l.
+\end{equation*}
+The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=1}^{T-l}$ and the realized changes $\mathbf{z} = (z_t)_{t=1}^{T-l}$.
+The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using the methods of probabilistic dichotomous forecast evaluation.
+Dichotomous forecasts predict a binary outcome, for example, a positive or negative change.
+Numerical measures for forecast quality are scoring rules.
+
+The \ac{bs} is a widely used scoring rule for dichotomous, probabilistic forecasts~\parencite{Brier1950}.
+It is defined as
+\begin{equation*}
+    BS (\mathbf{p}, \mathbf{z}) = \frac{1}{T-l} \sum_{t=1}^{T-l} (p_t - z_t)^2.
+\end{equation*}
+The \ac{bs} assesses the calibration and sharpness of the forecast and observation simultaneously~\parencite{Ranjan2010,Mitchell2011}.
+Calibration refers to the statistical consistency of forecasts and observations, while sharpness measures the spread of the forecast distribution.
+A smaller spread and thus higher sharpness is preferable, as it indicates a greater confidence in the prediction.
+The basic paradigm of probabilistic forecasting is to maximize the sharpness, subject to being calibrated~\parencite{Gneiting2014}.
+These two conflicting aspects cannot be distinguished in the scoring rule.
+
+Graphical models are a standard tool for assessing solely the calibration of probabilistic forecasts.
+In dichotomous forecasting, the reliability diagram is frequently used~\parencite{Ranjan2010}.
+The reliability diagram plots the observed frequency of the positive outcome against the forecasted probability.
+Ideally, the forecasted probability equals the observed frequency, and the reliability diagram is a 45-degree line.
+Local deviations from the 45-degree line indicate miscalibration of the forecast for specific probabilities.
+Thus, the reliability visualizes local and overall calibration simultaneously.
+For an example of a reliability diagram, see Section~\ref{sec:application-eda}.
+
+If the forecasts or nowcasts are given as quantiles, $p_t$ can be computed by interpolations between the quantiles.
+Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/P, \dots, (p-1) / P\}$ ($P \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
+Then, the probability $\bar{p}_t$ of a negative change is between $p^{\star}$ and $p^{\star} + 1/P$ for
+\begin{equation*}
+    p^{\star} = \max \{p \in \{1/P, \dots, (p-1) / P\}: q_p \leq y_t\} , \text{if} q_{1/p} \leq y_t \leq q_{1 - 1/p}.
+\end{equation*}
+The quantiles do not determine the location within the interval.
+Under the assumption of a uniform distribution within the quantile interval, the probability of a negative change is
+\begin{equation*}
+    \bar{p}_t = \frac{y_t - q_{p^\star}}{P (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
+\end{equation*}
+If $y_t$ is smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile, we assume the probability mass to be uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
+This yields
+\begin{equation*}
+\bar{p}_t = \begin{cases}
+    \max \{\frac{1}{P} - \frac{q_{p^\star} - y_t}{P (q_{p2/P} - q_{1/P})}, 0\} &, \text{if } y_t < q_{1/p}, \\
+    \min \{\frac{1}{P} - \frac{y_t - q_{(P-1)/P}}{P (q_{(P-1)/P} - q_{(P-2)/P})}, 1\} &, \text{if } y_t > q_{1 - 1/p}, \\
+    \frac{y_t - q_{p^\star}}{P (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star} &, \text{otherwise.}
+\end{cases}
+\end{equation*}
+The probability of positive change is $p_t = 1 - \bar{p}_t$.
+
+If the true value is unknown at lag $l$, the probabilities $p_t$ can be computed by integration.
+Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
+Then, the probability of a negative change can be computed by
+\begin{align*}
+    \bar{p}_t
+        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \\
+        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \\
+        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1.
+\end{align*}
+Thereby, the distributions are assumed to be independent.
+As a Monte Carlo approximation, the probability can also be computed by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
+For forecasts, the indexes have to be shifted.
+If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or samples can be generated based on the \ac{cdf} or quantiles.

### 2024-03-06 sections/04-application.tex
Add draft for probabilistic forecasting

@@
 
+Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
+The \acf{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the lags 3 and 7 days are shown Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
+The \acp{bs} are smallest and close for NBI-2 and Poisson-2 for both lags, while the Brier scores for the other models are larger and differ more.
+The qreg-1 model has the highest \ac{bs} for both lags.
+The reliability diagrams of GBM-2 and NBI-2 are also close and show overall a too small fraction of increases for the predicted probability.
+For the other models, the reliability diagrams show rather a too large fraction of increases for the corresponding predicted probability.
+
 \begin{table}
@@
 
+\begin{figure}
+    \begin{subfigure}{0.32\textwidth}
+    \tiny
+    \input{plots/ed_arrival/70_brier_scores}
+    \caption{Brier Scores for the different models and lags.}\label{fig:app-eda-prob-brier}
+    \end{subfigure}\hspace{0.01\textwidth}%
+    \begin{subfigure}[t]{0.32\textwidth}
+    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_3}
+    \caption{Reliability diagram for lag 3 days.}\label{fig:app-eda-prob-rel-3}
+    \end{subfigure}\hspace{0.01\textwidth}%
+    \begin{subfigure}[t]{0.32\textwidth}
+    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_7}
+    \caption{Reliability diagram for lag 7 days.}\label{fig:app-eda-prob-rel-7}
+    \end{subfigure}
+    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags 3 and 7 days.}
+    \label{fig:app-eda-prob}
+\end{figure}
+
 \subsubsection*{Discussion}
 
-Trending ability is consistent for the two lags, with the models' relative trending ability evolving similarly for the two lags.
+The trending ability is consistent for the two lags, with the models' relative trending ability evolving similarly for the two lags.
 The models' trending ability is generally higher for the smaller lag, but the differences are minor, and confidence intervals overlap.
@@
 Thus, the uncertainty of the model's predicted change has to be assessed differently based on the direction.
+
+The results of probabilistic trending evaluation \todo{Fertig schreiben}
 
@@
     \centering
+    \missingfigure{MIMIC Data summary}
+%    \input{plots/mimic/}
     \caption{Data summary of the \ac{abp} and \ac{nbp} measurements for the lags one minute, five minutes, and 15 minutes.}

### 2024-03-06 paper.tex
Update on Overleaf.
Add draft for probabilistic forecasting

@@
 \DeclareAcronym{cdf}{short=CDF, long=cumulative distribution function}
+\DeclareAcronym{pdf}{short=PDF, long=probability density function}
 \DeclareAcronym{nbp}{short=NBP, long=non-invasive blood pressure}
 \DeclareAcronym{abp}{short=ABP, long=invasive arterial blood pressure}
+\DeclareAcronym{bs}{short=BS, long=Brier score}
 
@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\N}{\mathbb{N}}
 \newcommand{\SBer}{\text{SBer}}
@@
 %\newcommand{\acceps}[1][]{\ifthenelse { \equal {#1} {} }{\acc_{\eps}}{\acc_{#1}}} %https://stackoverflow.com/a/7314951
-\newcommand{\acceps}[1][\varepsilon]{\acc_{#1}} %https://stackoverflow.com/a/7314951
-\newcommand{\accpeps}[1][\varepsilon]{\acceps[#1]^+}
-\newcommand{\accmeps}[1][\varepsilon]{\acceps[#1]^-}
+\newcommand{\acceps}[1][e]{\acc_{#1}} %https://stackoverflow.com/a/7314951
+\newcommand{\accpeps}[1][e]{\acceps[#1]^+}
+\newcommand{\accmeps}[1][e]{\acceps[#1]^-}
 \newcommand{\accl}[1][l]{\mu^{#1}}
@@
 \newcommand{\accml}[1][l]{\acc^{-,#1}}
+
+\newcommand{\pmax}{\hat{p}} % Notation for the grid resolution in quantiles
 
@@
 
-\begin{itemize}
-    \item Aufteilung
-    \begin{itemize}
-        \item Introduction (1.5 pages): Bolin fängt an, Conclusion (0.5 pages), Abstract
-        \item Application areas (1 pages): Bolin
-        \item Trending (5-6 pages): Jonas fängt an
-        \item Application (5-6 pages): Jonas
-    \end{itemize}
-    \item Datenquellen
-\end{itemize}
-
-\begin{itemize}    
-    \item Journal/Anwendungsgebiet klären: Angewandtes Medizin Statistikjournal: z.B. BMC Medical Research Methodology
-\end{itemize}
-
 \section{Introduction}\label{sec:introduction}
@@
 %\input{sections/02-notation}
+
+\newpage
 

### 2024-03-06 sections/03-trending.tex
Update on Overleaf.

@@
 Calibration refers to the statistical consistency of forecasts and observations, while sharpness measures the spread of the forecast distribution.
-A smaller spread and thus higher sharpness is preferable, as it indicates a greater confidence in the prediction.
+A smaller spread and, thus, higher sharpness is preferable, as it indicates a greater confidence in the prediction.
 The basic paradigm of probabilistic forecasting is to maximize the sharpness, subject to being calibrated~\parencite{Gneiting2014}.
@@
 
-If the forecasts or nowcasts are given as quantiles, $p_t$ can be computed by interpolations between the quantiles.
-Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/P, \dots, (p-1) / P\}$ ($P \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
-Then, the probability $\bar{p}_t$ of a negative change is between $p^{\star}$ and $p^{\star} + 1/P$ for
+If the forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
+Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
+Then, the probability $\bar{p}_t$ of a negative change is between $p^{\star}$ and $p^{\star} + 1/\pmax$ for
 \begin{equation*}
-    p^{\star} = \max \{p \in \{1/P, \dots, (p-1) / P\}: q_p \leq y_t\} , \text{if} q_{1/p} \leq y_t \leq q_{1 - 1/p}.
+    p^{\star} = \max \{p \in \{1/\pmax, \dots, (p-1) / \pmax\}: q_p \leq y_t\} , \text{if} q_{1/p} \leq y_t \leq q_{1 - 1/p}.
 \end{equation*}
-The quantiles do not determine the location within the interval.
+The quantiles do not determine the location within the interval $[p^{\star}, p^{\star} + 1/\pmax]$.
 Under the assumption of a uniform distribution within the quantile interval, the probability of a negative change is
 \begin{equation*}
-    \bar{p}_t = \frac{y_t - q_{p^\star}}{P (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
+    \bar{p}_t = \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
 \end{equation*}
@@
 \bar{p}_t = \begin{cases}
-    \max \{\frac{1}{P} - \frac{q_{p^\star} - y_t}{P (q_{p2/P} - q_{1/P})}, 0\} &, \text{if } y_t < q_{1/p}, \\
-    \min \{\frac{1}{P} - \frac{y_t - q_{(P-1)/P}}{P (q_{(P-1)/P} - q_{(P-2)/P})}, 1\} &, \text{if } y_t > q_{1 - 1/p}, \\
-    \frac{y_t - q_{p^\star}}{P (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star} &, \text{otherwise.}
+    \max \{\frac{1}{\pmax} - \frac{q_{p^\star} - y_t}{\pmax (q_{p2/\pmax} - q_{1/\pmax})}, 0\} &, \text{if } y_t < q_{1/p}, \\
+    \min \{\frac{1}{\pmax} - \frac{y_t - q_{(\pmax-1)/\pmax}}{\pmax (q_{(\pmax-1)/\pmax} - q_{(\pmax-2)/\pmax})}, 1\} &, \text{if } y_t > q_{1 - 1/p}, \\
+    \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star} &, \text{otherwise.}
 \end{cases}

### 2024-03-06 sections/03-trending.tex
Update on Overleaf.

@@
 Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
-Then, the probability $\bar{p}_t$ of a negative change is between $p^{\star}$ and $p^{\star} + 1/\pmax$ for
+The quantiles $q_p$ generally differ for each time step, but we omit an index here for ease of notation.
+The probability $\pc_t$ of a negative change is between $p^{\star}$ and $p^{\star} + 1/\pmax$ for
 \begin{equation*}
-    p^{\star} = \max \{p \in \{1/\pmax, \dots, (p-1) / \pmax\}: q_p \leq y_t\} , \text{if} q_{1/p} \leq y_t \leq q_{1 - 1/p}.
+    p^{\star} = \max \{p \in \{1/\pmax, \dots, (\pmax-1) / \pmax\}: q_p \leq y_t\} , \quad \text{if}\ q_{1/\pmax} \leq y_t \leq q_{1 - 1/\pmax}.
 \end{equation*}
@@
 \end{equation*}
-If $y_t$ is smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile, we assume the probability mass to be uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
+The approach does not assign probabilities for $y_t$ smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile.
+As a simple extension, we assume the probability mass to be uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
 This yields
@@
 
-If the true value is unknown at lag $l$, the probabilities $p_t$ can be computed by integration.
+If the true value is given as a distribution because it is yet unknown, the probabilities $p_t$ can be computed by integration.
 Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
 Then, the probability of a negative change can be computed by
-\begin{align*}
+\begin{align}
     \bar{p}_t
-        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \\
-        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \\
-        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1.
-\end{align*}
+        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
+        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
+        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
+\end{align}
 Thereby, the distributions are assumed to be independent.
-As a Monte Carlo approximation, the probability can also be computed by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
+As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be computed by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
 For forecasts, the indexes have to be shifted.
-If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or samples can be generated based on the \ac{cdf} or quantiles.
+If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to .

### 2024-03-06 sections/03-trending.tex
Update on Overleaf.

@@
-As described in Section~\ref{sec:introduction}, we refer by trending to the consistency of an observed change with the change issued by another measurement, a nowcast, or a forecast.
+As explained in Section~\ref{sec:introduction}, the term "trending" is used to denote the alignment of a detected true development with the development indicated by another measurement, a nowcast, or a forecast.
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
-\unsure{ist das Wort "observed" wirklich das was wir meinen? Stichwort: Rauschen}
 In general, trending can be assessed for different lags $l$ separately, and the relevance of the considered lags has to be determined based on the question at hand.
 Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
-Before introducing measures to evaluate trending, we make remarks on the calculation of predicted change and notation in Section~\ref{subsec:notation}.
+Before introducing measures to evaluate trending, we make remarks on the calculation of the predicted change and notation in Section~\ref{subsec:notation}.
 Section~\ref{subsec:trending-basics} introduces the basics of trending and visualizes trending in a synthetic example.
@@
 The \textit{observed} change is straightforward to compute for all types of signals.
-Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values or gold standard measurements up to time $T$.
-\unsure{Vielleicht schreiben wir hier noch ausfürhlich aus, was dieses y in drei Fällen sein sollte, oder ist das klar nach introduction?}
+Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values for nowcasting or forecasting, or gold standard measurements up to time $T$.
 The sequence of observed change is then given by the differences of values in $\mathbf{y}$ with lag $l$, that is
@@
 
-For measurement, a test method is usually evaluated against a gold standard to test whether the test method could replace the gold standard.
-If the test method is assessed to be consistent with the gold standard, the gold standard measurement values are not available to compute the predicted change.
-\unsure{irgendwie für mich nicht klar formuliert: man unteruscht ja bei meausurement nur ob die gemessenen Änderungen von den zwei Methoden konsistent sind oder?}
-Thus, the predicted change is computed on the basis of the test method's values at time $t$ and $t + l$.
+Third, for measurement, a test method is usually evaluated against a gold standard to test whether the test method could replace the gold standard.
+Thus, the gold standard measurements are not available if the quality of the test method is sufficient.
+\improvement{irgendwie für mich nicht klar formuliert: man unteruscht ja bei meausurement nur ob die gemessenen Änderungen von den zwei Methoden konsistent sind oder?}
+Thus, the predicted change is computed based on the values of the test method at time $t$ and $t + l$.
 
 Taking the three points together, the signal's predicted change is based on the type of signal.
-To account for several signals for a target time $t$ at different issue times $\tau$, we denote those by $x_{t | \tau}$.
+To account for several signals for a target time $t$ at different issue times $\tau$, we denote them by $x_{t | \tau}$.
 For nowcasting, $\tau$ is usually not considerably larger than $t$; while $t$ is usually at least $\tau$ in forecasting.
@@
 The quantiles $q_p$ generally differ for each time step, but we omit an index here for ease of notation.
-The probability $\pc_t$ of a negative change is between $p^{\star}$ and $p^{\star} + 1/\pmax$ for
+The probability $\pc_t$ of a \textit{negative} change is between $p^{\star}$ and $p^{\star} + 1/\pmax$ for
 \begin{equation*}
@@
 For forecasts, the indexes have to be shifted.
-If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to .
+If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to generate samples for the Monte Carlo approximation.

### 2024-03-06 sections/03-trending.tex
Update on Overleaf.

@@
 To account for several signals for a target time $t$ at different issue times $\tau$, we denote them by $x_{t | \tau}$.
-For nowcasting, $\tau$ is usually not considerably larger than $t$; while $t$ is usually at least $\tau$ in forecasting.
-\unsure{Nowcast macht Schätzung i.
-d.
-R.
-so lange, bis der echte Wert bekannt wird oder?}
-For measurements, this distinction is not necessary.
+For nowcasting, target times $t$ are usually not considerably later than issue times $\tau$; while $t$ is usually at least $\tau$ in forecasting. For measurements, this distinction is not necessary.
 Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
@@
 
-Analyzing the number of points in the green quadrants versus the red quadrants is a standard approach in trending assessment of measurement data \parencite{Critchley2010,Saugel2015}.
-\unsure{Quelle-Fehler}
+Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data \parencite{Critchley2010,Saugel2015}. 
 It boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes, respectively.
@@
 The use of these estimators could encourage predictors to exploit imbalances in the number of positive and negative changes. 
-A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ is unlikely in the trending setting, since $\diffy$ is obtained from differencing time series data.
-\unsure{hier stimmt was nicht}
+A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t \in \mathcal{T}} \ind{\diffyt < 0}$ is unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data.
 However, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
@@
 Cohen's $\kappa$ is usually used to measure inter-rater agreement and takes into account the ratio of occurred agreement and the probability of agreement by chance.
-Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$ table and balanced outcomes.
+Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$-table and balanced outcomes.
 
@@
 Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
-Thus, the yearly course of the prediction's trending ratio can be detected.
-The trending ability of the signal has a strong yearly sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
+Thus, the yearly course of the prediction trending ratio can be detected.
+The trending ability of the signal has a strong sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
 This seasonal behavior cannot be distinguished in the two four-quadrant plots or in the trending ratio.
@@
     \caption{Visualizations for data with a time-varying trending ratio.
-We defer information on the data generation process of the data in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{subsec:app-trending-data-generation}).
-The trending ratios on the whole data set are $\mu = 0.7577$.
-The strong yearly seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
+We defer information on the data generation process in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{subsec:app-trending-data-generation}).
+The trending ratio for the entire data set is $\mu = 0.7577$.
+The strong seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
 The green curve $k_t$ shows the probability that $\diffxt$ has the same sign as $\diffyt$ for each time step.
@@
 \end{figure}
-\unsure{was ist $\mu_1$ bei dem Plot, hinschreiben}
+
 In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
@@
 Points within that area are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
-The shape of the exclusion area can be chosen according to the noise characteristics of predictions and true values.
-In general, it is likely that the prediction models have a noise component and should thus be part of the exclusion area.\unsure{vielleicht eine große Buchstabe für exclusion area?}
-We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $e$ by
+The shape of the exclusion area can be chosen according to the noise characteristics of the predictions and the true values.
+In general, it is likely that the prediction models have a noise component and should thus be part of the exclusion area.
+We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $E$ by
 \begin{align}
-    \acceps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin e}}\label{eq:acceps}\\
-    \accpeps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin e}} \label{eq:accpeps}\\
-    \accmeps (\diffx, \diffy, e) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin e}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin e}}\label{eq:accmeps}
+    \acceps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin E}}\label{eq:acceps}\\
+    \accpeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}} \label{eq:accpeps}\\
+    \accmeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}\label{eq:accmeps}
 \end{align}
-The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $e$, that is, for $\acceps$, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$.
+The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $E$, that is, for $\acceps$, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$.
 The estimators accept various shapes of the exclusion area.
 Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
-A rectangular exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ ,for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and therefore likely to be driven by noise.
+A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ 
+ ($\varepsilon_x, \varepsilon_y > 0$), leaves out points that are small in both components and, therefore, likely to be driven by noise.
 Points where at least the true value or signal is unlikely to be zero are not excluded.
 In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
-An exclusion area along one axis, for example, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq 0 \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
+An exclusion area along one axis, for example, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq 0 \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
 This particularly suits signals where small amounts of noise are inevitable.
-A cross-shaped exclusion area, $e = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
+A cross-shaped exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
 For these two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
@@
 The size estimation can also be based on a proportion of the total variance or the total range of the data.
-A third approach is to visualize the trending ratio for different values of $e$.
-Section~\ref{sec:application} contains various examples of such a plot.
-As a base approach, we suggest using no exclusion area in the four-quadrant plot, as the points do not complicate its interpretation, and a $\acceps$-over-$e$ plot to visualize the course of the trending ratio over different exclusion area sizes.\unsure{"as the points do not complicate its interpretation" ist das hier die "richtige" Begründung, dass wir kein exlclusion area betrachten?}
+A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
 

### 2024-03-06 sections/03-trending.tex
Update on Overleaf.

@@
 Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
-New samples are drawn with replacement from the dataset and are not based on parametric assumptions as for classical confidence intervals.
+New samples are drawn with replacement from the dataset and are not based on parametric assumptions as classical confidence intervals are.
 The confidence interval is then computed based on the bootstrap samples.
 We examine three methods for bootstrapping: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
-In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
+In the \textit{percentile} approach, the confidence interval for the level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
 The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
@@
 Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
-The basic approach balances these two aims.
-We compare the approaches in a small synthetic data study on their small-dataset behavior and computation time in Appendix~\ref{sec:appendix}.
-The \ac{bca} is the only method that holds the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
+The basic approach balances these two objectives.
+We compare the approaches in a small synthetic data study on their small-dataset behavior and computation time in Appendix~\ref{sec:appendix-trending}.
+\Ac{bca} is the only method that maintains the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
 Therefore, we use the \ac{bca} method for confidence intervals in the applications in Section~\ref{sec:application}.
@@
 In nowcasting and forecasting, there is a shift from point predictions to probabilistic predictions (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
-Probabilistic predictions include a point estimate and information on forecast uncertainty and quantiles simultaneously.
+Probabilistic predictions include a point estimate and information on the forecast uncertainty and quantiles simultaneously.
 In this section, we extend the trending assessment to probabilistic predictions.
 
-Probabilistic predictions can have the form of a \ac{cdf}, \ac{pdf}, or quantiles.
+Probabilistic predictions can take the form of a \ac{cdf}, \ac{pdf}, or quantiles.
 The \ac{cdf} is the most general and can be used to derive the others, given they exist.
@@
 Let $F_{t | \tau} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $\tau$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
-For the trending assessment, we contrast the probability of positive change to the occurrence of positive changes. 
+For trending assessment, we compare the probability of positive change with the occurrence of positive changes. 
 The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value. \unsure{Andere Formulierung für: one minus?}
@@
 \end{equation*}
-Let $z_t$ denote the indicator that the occurred change at time $t$ is positive, that is,
+Let $z_t$ denote the indicator that the observed change at time $t$ is positive, that is,
 \begin{equation*}
@@
 \end{equation*}
-The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=1}^{T-l}$ and the realized changes $\mathbf{z} = (z_t)_{t=1}^{T-l}$.
-The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using the methods of probabilistic dichotomous forecast evaluation.
+The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=1}^{T-l}$ and the observed changes $\mathbf{z} = (z_t)_{t=1}^{T-l}$.
+The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
 Dichotomous forecasts predict a binary outcome, for example, a positive or negative change.
@@
 
-The \ac{bs} is a widely used scoring rule for dichotomous, probabilistic forecasts~\parencite{Brier1950}.
+The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\parencite{Brier1950}.
 It is defined as
@@
 \end{equation*}
-The \ac{bs} assesses the calibration and sharpness of the forecast and observation simultaneously~\parencite{Ranjan2010,Mitchell2011}.
+The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\parencite{Ranjan2010,Mitchell2011}.
 Calibration refers to the statistical consistency of forecasts and observations, while sharpness measures the spread of the forecast distribution.
-A smaller spread and, thus, higher sharpness is preferable, as it indicates a greater confidence in the prediction.
-The basic paradigm of probabilistic forecasting is to maximize the sharpness, subject to being calibrated~\parencite{Gneiting2014}.
+A smaller spread, and thus, higher sharpness, is preferable, as it indicates greater confidence in the prediction.
+The basic paradigm of probabilistic forecasting is to maximize sharpness, subject to calibration~\parencite{Gneiting2014}.
 These two conflicting aspects cannot be distinguished in the scoring rule.
 
-Graphical models are a standard tool for assessing solely the calibration of probabilistic forecasts.
+Graphical models are a standard tool for evaluating solely the calibration of probabilistic forecasts.
 In dichotomous forecasting, the reliability diagram is frequently used~\parencite{Ranjan2010}.
-The reliability diagram plots the observed frequency of the positive outcome against the forecasted probability.
-Ideally, the forecasted probability equals the observed frequency, and the reliability diagram is a 45-degree line.
-Local deviations from the 45-degree line indicate miscalibration of the forecast for specific probabilities.
-Thus, the reliability visualizes local and overall calibration simultaneously.
+The reliability diagram plots the observed frequency of the positive outcome against the predicted probability.
+Ideally, the predicted probability equals the observed frequency, and the reliability diagram is a 45-degree line.
+Local deviations from the 45-degree line indicate a miscalibration of the forecast for specific probabilities.
+Thus, the reliability visualizes the local and overall calibration simultaneously.
 For an example of a reliability diagram, see Section~\ref{sec:application-eda}.
 
-If the forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
+If forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
 Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
@@
 \end{equation*}
-The quantiles do not determine the location within the interval $[p^{\star}, p^{\star} + 1/\pmax]$.
+Quantiles do not determine the location within the interval $[p^{\star}, p^{\star} + 1/\pmax]$.
 Under the assumption of a uniform distribution within the quantile interval, the probability of a negative change is
@@
 The approach does not assign probabilities for $y_t$ smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile.
-As a simple extension, we assume the probability mass to be uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
+As a simple extension, we assume that the probability mass is uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
 This yields
@@
 
-If the true value is given as a distribution because it is yet unknown, the probabilities $p_t$ can be computed by integration.
+If the true value is given as a distribution because it is still unknown, the probabilities $p_t$ can be computed by integration.
 Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
@@
 Thereby, the distributions are assumed to be independent.
-As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be computed by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
+As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
 For forecasts, the indexes have to be shifted.

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 Section~\ref{sec:application-eda} applies the trending assessment to forecast methods for the number of arrivals in a large emergency department.
-In the last application in Section~\ref{sec:application-measurement}, we consider the trending assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.
+In the last application in Section~\ref{sec:application_measurement}, we consider the trending assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.
 
@@
 Table~\ref{tab:app-eda-marginals} analyzes the differences in marginal distributions for the forecasts and true values for the lags of 3 and 7 days.
-Note that the difference definition aligns with Section~\ref{sec:notation}, defined as the difference between the forecasted mean and true value of 3 and 7 days before, as the true value is available when issuing the forecast.
+Note that the difference definition aligns with Section~\ref{subsec:notation}, defined as the difference between the forecasted mean and true value of 3 and 7 days before, as the true value is available when issuing the forecast.
 The fraction of positive differences varies between 0.39 and 0.63 for the lag of 3 days and between 0.37 and 0.63 for the lag of 7 days.

### 2024-03-06 sections/55-appendix-application-covid.tex
Update on Overleaf.

@@
     \end{tabular}
-    \caption{Matching of the abbreviation to the key in the nowcasting hub.
+    \caption{Matching the abbreviation to the key in the nowcasting hub.
     Information on the models and references is listed in \textcite[][Table 1]{Wolffram2023}.}
@@
     \centering
-    \begin{subfigure}[t]{\textwidth}
+    \begin{subtable}[t]{\textwidth}
         \input{plots/covid_nowcast/30_trending_ratios_lag_1}
     \caption{1 day.}
-    \end{subfigure}
-    \begin{subfigure}[t]{\textwidth}
+    \end{subtable}
+    \begin{subtable}[t]{\textwidth}
         \input{plots/covid_nowcast/30_trending_ratios_lag_14}
         \caption{14 days.}
-    \end{subfigure}
+    \end{subtable}
     \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the lag 1 and 14 days.
-The exclusion areas are rectangles centered on the zero points with width and height of 10\% of the quantile of the absolute values of nowcast and true values.
+The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values.
 }
@@
 \includegraphics{plots/covid_nowcast/30_4q_plots}
-\caption{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the lags of one, seven, and 14 days. }
+\caption{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the lags of one, seven, and 14 days. The spread in both directions increases with the lag. }
 \label{fig:app-covid-4q}
@@
     \end{subfigure}
-    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM and ENS-MED for the lag 7 days.}
+    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the lag seven days.}
     \label{fig:app-covid-cond-prob-trending-ratio-1-14}

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 
-During the COVID-19 pandemic, the need for reliable and timely nowcasts of the pandemic's development has become apparent.
+Amid the COVID-19 pandemic, the importance of accurate and prompt nowcasts of the pandemic's progression has become evident.
+The pandemic's spread and 
 In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\parencite{RobertKochInstitute2021}.
-The Robert Koch Institute (RKI) provides daily reports on the number of new hospitalizations.
-However, these reports are subject to delays and revisions in two sources.
+\todo{Name in Citation RKI!}
+The Robert Koch Institute (RKI) provided preliminary daily reports on the seven-day hospitalization rate.
+However, these reports were subject to delays and revisions in two sources.
 The first source is technical delays in the reporting process, for example, due to different authorities passing the data to the RKI~\parencite{RobertKochInstitute2024}.
@@
 To a given date, all the cases are allocated whose first positive test result was on that date and who were hospitalized in relation to the disease in the following.
-The seven-day hospitalization rate is the average count of cases per $100,000$ inhabitants on the given date and six days before.
-Thus, the final seven-day hospitalization rate can only be reported with a significant delay of up to more than 70 days.
+The seven-day hospitalization rate is the average count of those cases per $100,000$ inhabitants on the given date and six days before.
+Thus, the final seven-day hospitalization rate can only be reported with a significant delay of up to more than 70 days, as the hospitalization of infected inhabitants can occur much later than the first positive test.
 Nevertheless, as the seven-day hospitalization rate was considered a major indicator of the pandemic's development, many organizations and institutions started to issue nowcasts, including research teams and newspapers.
@@
     \includegraphics{plots/covid_nowcast/00_true_data.pdf}
-    \caption{Realisations.}
+    \caption{Realisations.\todo{Capitalize Date}}
     \label{fig:app-covid-true}
@@
     \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
-    \caption{Same-day nowcasts}
+    \caption{Same-day nowcasts. \todo{Capitalize Forecast Date}}
     \label{fig:app-covid-nowcast}

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 Amid the COVID-19 pandemic, the importance of accurate and prompt nowcasts of the pandemic's progression has become evident.
-The pandemic's spread and 
+The pandemic's spread and severity was 
 In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\parencite{RobertKochInstitute2021}.

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

--- 2024-03-06 sections/04-application.tex
@@
 Amid the COVID-19 pandemic, the importance of accurate and prompt nowcasts of the pandemic's progression has become evident.
-The pandemic's spread and 
+The pandemic's spread and severity were measured by various indicators. 
 In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\parencite{RobertKochInstitute2021}.
@@
 The Robert Koch Institute (RKI) provided preliminary daily reports on the seven-day hospitalization rate.
-However, these reports were subject to delays and revisions in two sources.
+However, these reports were subject to severe delays and revisions in two sources.
 The first source is technical delays in the reporting process, for example, due to different authorities passing the data to the RKI~\parencite{RobertKochInstitute2024}.
@@
 The model structures are diverse, including Bayesian models, generalized additive models, and parametric bootstrapping.
-Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the model's overall structure.
-Using the nowcasts, two ensemble methods are constructed using the ensembles' mean or for the large .
+Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations together with the respective names in the COVID19-Nowcasting-Hub.
+For information on the model design, we refer to~\textcite{Wolffram2023}.
+Using the nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
 We denote them by ENS-MEAN and ENS-MED.
+\textcite{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams that was stated in a preregistered study. 
 In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 We denote them by ENS-MEAN and ENS-MED.
-\textcite{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams that was stated in a preregistered study. 
+\textcite{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams that were stated in a preregistered study. 
 In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
 In contrast to~\textcite{Wolffram2023}, we use the newest data available on the true values, not the data from August 8, 2022.
-\unsure{Nochmal längere Evaluation?}
+
 We do not expand on the models' performance on specific regions or age groups in Germany and the probabilistic nowcast assessment.

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 In contrast to~\textcite{Wolffram2023}, we use the newest data available on the true values, not the data from August 8, 2022.
-
+The nowcasts are resolved with regard to all inhabitants but also with regard to certain geographical and age breakdowns.
 We do not expand on the models' performance on specific regions or age groups in Germany and the probabilistic nowcast assessment.
@@
 Table~\ref{tab:app-covid-rmse} summarizes the point evaluation measures for the issued mean of the different models.
-The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\parencite[see][Tables A2, A3, and A4]{Wolffram2023}.
+The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\parencite[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
 The best-performing models in terms of RMSE and MAE are the ILM and RKI models.
@@
 Before stating the results, we provide background information on the marginal distributions of the true values and nowcasts for the different lags.
-Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} provides information on marginal statistics of the nowcasts and true values.
+Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} provides information on marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
 Overall, the variability and general level of differences grow with the lag.
-The standard deviation increases from roughly 300 for lag 1 to 1,200 for lag 7 and 2,000 for lag 14.
+The standard deviation increases from roughly 300 for lag one to 1,200 for lag seven and 2,000 for lag 14 days.
 Similarly, the 10\%-quantile of differences increases.
@@
 The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the differences.
-Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, e.g., $\diffy$ are to be recognized by the nowcast and vice versa.
+Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.
 
 Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the lag of 7 days.
-The trending ratios without exclusion area range from 0.72 to 0.85 for the lag of 7 days.
+The trending ratios without exclusion area range from 0.72 to 0.85 for the lag of seven days.
 The negative trending ratios are higher than the positive trending ratios for all models.
-For all models, the confidence intervals for the positive and negative trending ratios do not overlap, indicating that the trending ratios are indeed different.
+The confidence intervals for the positive and negative trending ratios do not overlap for all models, indicating that the trending ratios are indeed different.
 The 10\%-quantile exclusion areas have, at most, an influence of 0.03 on the ratios.
 The model with the highest trending ratio is the ILM model, and the model with the lowest is the RKI model.
-The confidence intervals between all models overlap, reflecting the evaluation period lasting 159 days.
+The confidence intervals between all models overlap.
 The positive trending ratio implies a similar ranking of the models, while the negative ratio is second best for the RKI model.
-For the lags of 1 and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
+For the lags of one and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
 
@@
 
-The trending ratios are plotted for various exclusion areas in Figure~\ref{fig:app-covid-trending-ratio-7}.
+The trending ratios for various exclusion areas are shown in Figure~\ref{fig:app-covid-trending-ratio-7}.
 In general, the trending ratio increases with larger exclusion areas.
 While the RIVM and ENS-MED trending ratios evolve similarly, the RKI and ILM trending ratios get closer.
-For the lag of 1 day, the RKI trending ratio decreases with increasing exclusion area size while the other models evolve increasingly (see Figure~\ref{fig:app-covid-trending-ratio-1}).
+For the lag of one day, the RKI trending ratio decreases with increasing exclusion area size while the other models rise (see Figure~\ref{fig:app-covid-trending-ratio-1}).
 For the lag of 14 days, all trending ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-trending-ratio-14}).
-
-\unsure{Generell bei diesem Abschnitt: man liest hier eigentlich nur Beobachtungen, kann man auch daraus Schlüsse ziehen? Kann man results und discussion verschmelzen?}
 
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[l]$, and negative trending ratio $\accml[l]$ for the models with and without exclusion areas for the lag seven days.
+    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models with and without exclusion areas for the lag seven days.
 The exclusion areas are rectangles centered on the zero points with a width and height of 10\%-quantile of the absolute values of nowcast and true values.
@@
 For all lags, the influence of the exclusion area on the 10\%-quantile level is negligible.
-For example, the trending ratio changes at most by 0.03 for the EPI model with $\acc^{-, 14}$.
+For example, the trending ratio changes at most by 0.03 for the EPI model with $\accm[14]$.
 The exclusion areas are thus not crucial for the trending assessment in the case of the nowcasts of the seven-day hospitalization rate.
@@
 For critical patients, changes in blood pressure can be crucial for the treatment.
-Thus, trending assessment can be performed in addition to standard accuracy analysis~\parencite[see, e.g., ]{Mostafa2020}.
+Thus, trending assessment can be performed in addition to standard accuracy analysis~\parencite[see, for example, ]{Mostafa2020}.
 Thus, we assess the trending ability of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 Amid the COVID-19 pandemic, the importance of accurate and prompt nowcasts of the pandemic's progression has become evident.
-The pandemic's spread and severity were measured by various indicators. 
+Various indicators measured the pandemic's spread and severity. 
 In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\parencite{RobertKochInstitute2021}.
@@
 To a given date, all the cases are allocated whose first positive test result was on that date and who were hospitalized in relation to the disease in the following.
-The seven-day hospitalization rate is the average count of those cases per $100,000$ inhabitants on the given date and six days before.
+The seven-day hospitalization rate is the average number of those cases per $100,000$ inhabitants on the given date and six days before.
 Thus, the final seven-day hospitalization rate can only be reported with a significant delay of up to more than 70 days, as the hospitalization of infected inhabitants can occur much later than the first positive test.
@@
 The model structures are diverse, including Bayesian models, generalized additive models, and parametric bootstrapping.
-Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations together with the respective names in the COVID19-Nowcasting-Hub.
+Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the respective names in the COVID19-Nowcasting-Hub.
 For information on the model design, we refer to~\textcite{Wolffram2023}.
@@
 We denote them by ENS-MEAN and ENS-MED.
-\textcite{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams that were stated in a preregistered study. 
+\textcite{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams stated in a preregistered study. 
 In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
@@
 
-Trending assessment is unattached to point evaluation measures.
+Trending assessment evaluates the models differently from point evaluation measures.
 RKI is among the best in point evaluation measures but performs worse in trending assessment.
@@
 A more extensive training size would be beneficial for assessing the models' performance.
-For some models, more data is accessible, but we stick with the study protocol that uses an evaluation period of 159 days.
+For the evaluation period of 159 days, the trending ratio confidence intervals overlap, and thus, no conclusions can be drawn from the trending evaluation.
 
@@
 Forecasting the number of arrivals in an emergency department is crucial for planning staff and resources.
-In \textcite{Rostami-Tabar2023}, various models are employed to forecast the hourly number of arrivals in a large emergency department.
+In the study by \textcite{Rostami-Tabar2023}, several models are used for predicting hourly outcomes.
 Every 12 hours, the models issue hourly forecasts for the next 48 hours.
 Thus, the management can take measures according to the expected number of arrivals, for example, through redeploying staff and reconfiguring units.
-The models issue probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
+
+\textcite{Rostami-Tabar2023} publish probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
 The models are trained on data from April 1, 2014, to February 28, 2018, and evaluated on data from March 1, 2018, to February 28, 2019.
 For further notes on the models and the evaluation, we refer to \textcite{Rostami-Tabar2023}.
-From the issued forecasts, we use the mean as a point forecast for the trending assessment.
+From the issued forecasts, we use the mean as a point forecast for the trending assessment and evaluate the probabilistic trending subsequently.
 We use the forecasts issued at the first time point for every target time.
@@
 In this setup, trending assessment is a simple and intuitive way to assess the models' performance.
-It is easy for the management to understand and implement, as simple comparisons of the expected workload to an hour in the recent past can be made.
+The trending perspective is easy for the management to understand and implement, as simple comparisons of the expected workload to a recent shift can be made.
 If, for example, the staff was near capacity in the last shift and an increase in the number of arrivals is expected, the management can take measures to adjust the workload.
@@
 Thus, we consider the lags of 72 hours, the last already observed shift of the same hour of day, and seven days, the previous shift of the same hour and day.
-Table~\ref{tab:app-eda-point-evaluation} lists the point evaluation measures for the models.
+Table~\ref{tab:app-eda-point-evaluation} lists the point evaluation measures and the count of available forecasts.
 The best-performing models regarding RMSE and MAE are the NBI-2 and Poisson-2 models.
@@
 \input{plots/ed_arrival/00_point_evaluation_measures}
-\caption{Point evaluation measures for the models.
-The smaller count for some models stems from missing forecasts for a few hours in the course of the evaluation period.}\label{tab:app-eda-point-evaluation}
+\caption{Point evaluation measures for the models. The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.}\label{tab:app-eda-point-evaluation}
 \end{table}
@@
 
-Table~\ref{tab:app-eda-marginals} analyzes the differences in marginal distributions for the forecasts and true values for the lags of 3 and 7 days.
-Note that the difference definition aligns with Section~\ref{subsec:notation}, defined as the difference between the forecasted mean and true value of 3 and 7 days before, as the true value is available when issuing the forecast.
-The fraction of positive differences varies between 0.39 and 0.63 for the lag of 3 days and between 0.37 and 0.63 for the lag of 7 days.
-The variability of differences decreases for the larger lag for most models; only for the ETS model does it increase.
+Table~\ref{tab:app-eda-marginals} analyzes the differences in marginal distributions for the forecasts and true values for the lags of three and seven days.
+Note that the difference definition aligns with Section~\ref{subsec:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the true value is available when issuing the forecast.
+The fraction of positive differences varies between 0.39 and 0.63 for the lag of three days and between 0.37 and 0.63 for the lag of seven days.
+The variability of differences decreases for the larger lag for most models; only for the ETS model it increases.
 The 10\%-quantile of the differences is between zero and one for all models and lags.

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
-In contrast to~\textcite{Wolffram2023}, we use the newest data available on the true values, not the data from August 8, 2022.
-The nowcasts are resolved with regard to all inhabitants but also with regard to certain geographical and age breakdowns.
+In contrast to~\textcite{Wolffram2023}, we use the newest data available on the true values, not from August 8, 2022.
+The nowcasts are resolved with regard to all inhabitants and specific geographical and age breakdowns.
 We do not expand on the models' performance on specific regions or age groups in Germany and the probabilistic nowcast assessment.
@@
 We report the trending for the lags 1, 7, and 14 days.
-While lag 1 assesses the short-term trending, lags 7 and 14 consider the medium-term trending.
-The lags 7 and 14 are particularly interesting, reflecting a usual period until new policy changes are taken.
+While lag one assesses the short-term trending, lags seven and 14 consider the medium-term trending.
+The lags seven and 14 are particularly interesting, reflecting a usual period until new policy changes are taken.
 
@@
 
-Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the lag of 7 days.
+Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the lag of seven days.
 The trending ratios without exclusion area range from 0.72 to 0.85 for the lag of seven days.
@@
 A more extensive training size would be beneficial for assessing the models' performance.
-For the evaluation period of 159 days, the trending ratio confidence intervals overlap, and thus, no conclusions can be drawn from the trending evaluation.
+For the evaluation period of 159 days, the trending ratio confidence intervals overlap; thus, no conclusions can be drawn from the trending evaluation.
 
@@
 Forecasting the number of arrivals in an emergency department is crucial for planning staff and resources.
-In the study by \textcite{Rostami-Tabar2023}, several models are used for predicting hourly outcomes.
+Several models are used for predicting hourly outcomes in the study by \textcite{Rostami-Tabar2023}.
 Every 12 hours, the models issue hourly forecasts for the next 48 hours.
@@
 The fraction of positive differences varies between 0.39 and 0.63 for the lag of three days and between 0.37 and 0.63 for the lag of seven days.
-The variability of differences decreases for the larger lag for most models; only for the ETS model it increases.
+The variability of differences decreases for the larger lag for most models; only for the ETS model does it increase.
 The 10\%-quantile of the differences is between zero and one for all models and lags.
-Thus, we exclude only values smaller than one from the trending assessment.
+Thus, we exclude only differences smaller than one from the trending assessment.
 The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.
@@
 The confidence interval width is at most 0.02 for the trending ratios and at most 0.03 for the positive and negative trending ratios.
-The positive trending ratio is higher for other models than the negative trending ratio.
-The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for the lag of 3 and seven days, while Poisson-2 and NBI-2 have the highest negative trending ratio.
+Other models exhibit a higher positive trending ratio than the negative trending ratio.
+The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for the lag of three and seven days, while Poisson-2 and NBI-2 have the highest negative trending ratio.
 
-Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags 3 and 7 days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
+Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags three and seven days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
 The conditional trending plots show similar courses for the two lags, while the curves are shifted downwards for the lag of seven days.
@@
 Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
-The \acf{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the lags 3 and 7 days are shown Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
+The \acf{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the lags three and seven days are shown in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
 The \acp{bs} are smallest and close for NBI-2 and Poisson-2 for both lags, while the Brier scores for the other models are larger and differ more.
 The qreg-1 model has the highest \ac{bs} for both lags.
-The reliability diagrams of GBM-2 and NBI-2 are also close and show overall a too small fraction of increases for the predicted probability.
-For the other models, the reliability diagrams show rather a too large fraction of increases for the corresponding predicted probability.
+The reliability diagrams of GBM-2 and NBI-2 are also close and show a too-small fraction of increases for the predicted probability overall.
+For the other models, the reliability diagrams show a fraction of increases that are too large for the corresponding predicted probability.
 
@@
     \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_3}
-    \caption{Lag 3 days}
+    \caption{Lag three days}
     \end{subfigure}\hfill
@@
     \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_7}
-    \caption{Lag 7 days}
+    \caption{Lag seven days}
     \end{subfigure}
-    \caption{Conditional trending plots for the lags 3 and 7 days and the models with the best positive or negative trending ability. The plots of NBI-2 and Poisson-2 are indistinguishable.}
+    \caption{Conditional trending plots for the lags three and seven days and the models with the best positive or negative trending ability. The plots of NBI-2 and Poisson-2 are indistinguishable.}
     \label{fig:app-eda-cond-prob}
@@
     \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_3}
-    \caption{Reliability diagram for lag 3 days.}\label{fig:app-eda-prob-rel-3}
+    \caption{Reliability diagram for lag three days.}\label{fig:app-eda-prob-rel-3}
     \end{subfigure}\hspace{0.01\textwidth}%
@@
     \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_7}
-    \caption{Reliability diagram for lag 7 days.}\label{fig:app-eda-prob-rel-7}
+    \caption{Reliability diagram for lag seven days.}\label{fig:app-eda-prob-rel-7}
     \end{subfigure}
-    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags 3 and 7 days.}
+    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags three and seven days.}
     \label{fig:app-eda-prob}
@@
 
-The results of probabilistic trending evaluation \todo{Fertig schreiben}
+The results of probabilistic trending evaluation \todo{fertig}
 
@@
 For some patients, the data includes \ac{abp} and \ac{nbp} measurements.
+While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods.
 For an overview of blood pressure measurement methods, see \textcite{Saugel2014}.
-While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods.
 For critical patients, changes in blood pressure can be crucial for the treatment.
@@
 One data record includes all numerical measurements for one patient.
-In general, the measured signals vary in length, frequency and type of measurement.
-Thus, only a subset of the data contains measurements of \ac{abp} and \ac{nbp} at the same time.
+In general, the measured signals vary in length, frequency, and type of measurement.
+Thus, only a subset of the data contains measurements of \ac{abp} and \ac{nbp} simultaneously.
 2,548 include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one measurement of systolic \ac{abp} and \ac{nbp} at the same time; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
@@
 The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different lags are shown in Figure~\ref{fig:app-mimic-4q}.
-The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} at the same time for two consecutive times with the specified lags.
+The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} simultaneously for two consecutive times with the specified lags.
 For the systolic measurements, 290, 332, and 442 points are available for the lags of one, five, and 15 minutes; for the mean measurements, 406, 430, and 542.
@@
 For larger lags, the trending ratio increases.
-The difference between positive and negative trending ratios is small for all types and lags, with overlapping confidence intervals in all cases.
+The difference between positive and negative trending ratios is small for all types and lags, with overlapping confidence intervals.
 
@@
 It becomes apparent that the systolic measurements have a higher trending ability than the mean measurement, except for small negative predicted changes.
-This is in line with the trending ratios, but the confidence intervals overlap.
+This aligns with the trending ratios, but the confidence intervals overlap.
 
@@
     \caption{Four-quadrant plots for the different lags and the systolic and mean blood pressure measurements.
-The upper row contains systolic measurements and the lower row mean measurements.
+The upper row contains systolic measurements, and the lower row contains mean measurements.
 The columns contain the lags one, five, and 15 minutes.}
@@
 The four-quadrant plots contain a considerable number of extreme points.
-It is not clearly distinguishable whether these points are due to measurement errors or actual extreme values.
-Some authors argue to exclude not only the measurements below the 10\%-quantile of the absolute differences but also the points above the 90\%-quantile.
-We do not follow this approach here, as the extreme values are not necessarily measurement errors and could also be particularly relevant.
+Whether these points are due to measurement errors or extreme values is not distinguishable.
+Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile.
+We do not follow this approach here, as the extreme values are not necessarily measurement errors and could be particularly relevant.
 
@@
 The positive and negative trending ratios have overlapping confidence intervals, and the conditional trending plots do not contain prominent deviations in the course.
-This is in line with the four-quadrant plots, where no clear asymmetry is visible.
+This aligns with the four-quadrant plots, where no apparent asymmetry is visible.
 
 The bootstrap confidence intervals are wide.
-For the trending ratio, the width is around 0.1, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the lag of 1 minute.
+The width is around 0.1 for the trending ratio, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the lag of one minute.

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
     \caption{Marginal analysis of the nowcast and true differences.
-The column (1), $l=l$ shows the fraction of values greater than zero for lag $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
+The column (1) shows the fraction of values greater than zero for lag $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
     \label{tab:app-eda-marginals}
@@
 
-The results of probabilistic trending evaluation \todo{fertig}
+The results of the probabilistic trending evaluation endorse the point trending assessment and assign the best scores to NBI-2 and Poisson-2. 
+The reliability diagrams show that they underestimate the fraction of increases slightly. 
+This aligns with the trending ratio where the
 
@@
 One data record includes all numerical measurements for one patient.
-In general, the measured signals vary in length, frequency, and type of measurement.
+The measured signals vary in length, frequency, and type of measurement.
 Thus, only a subset of the data contains measurements of \ac{abp} and \ac{nbp} simultaneously.
@@
 We consider the lags of one minute, five minutes, and 15 minutes for the trending assessment, as those are typical intervals of NBP measurements.
-The number of available records with measurements of \ac{abp} and \ac{nbp} for the lags are listed in Table~\ref{tab:app-mimic-data-summary}.
-
-\begin{table}
-    \centering
-    \missingfigure{MIMIC Data summary}
-%    \input{plots/mimic/}
-    \caption{Data summary of the \ac{abp} and \ac{nbp} measurements for the lags one minute, five minutes, and 15 minutes.}
-    \label{tab:app-mimic-data-summary}
-\end{table}
 
@@
 The upper row contains systolic measurements, and the lower row contains mean measurements.
-The columns contain the lags one, five, and 15 minutes.}
+The columns contain the lags one, five, and 15 minutes.\todo{lag ausgeschrieben}}
     \label{fig:app-mimic-4q}
@@
     \end{subfigure}\hspace{0.01\textwidth}
-    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the lags one, five, and 15 minutes. \hl{Short interpretation.}}
+    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the lags one, five, and 15 minutes. \todo{No l = ... in the legend}}
     \label{fig:app-mimic-cond-prob}
@@
 Whether these points are due to measurement errors or extreme values is not distinguishable.
-Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile.
+Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile \parencite[see][]{Critchley2010}.
 We do not follow this approach here, as the extreme values are not necessarily measurement errors and could be particularly relevant.
@@
 The width is around 0.1 for the trending ratio, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the lag of one minute.
+Thus, more measurements would be of interest for a further trending assessment.

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 Other models exhibit a higher positive trending ratio than the negative trending ratio.
-The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for the lag of three and seven days, while Poisson-2 and NBI-2 have the highest negative trending ratio.
+The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for three and seven days lag, while Poisson-2 and NBI-2 have the highest negative trending ratio.
 
@@
 The reliability diagrams show that they underestimate the fraction of increases slightly.
-This aligns with the trending ratio where the
 

### 2024-03-06 sections/04-application.tex
Update on Overleaf.

@@
 In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\parencite{RobertKochInstitute2021}.
-\todo{Name in Citation RKI!}
 The Robert Koch Institute (RKI) provided preliminary daily reports on the seven-day hospitalization rate.
@@
     \includegraphics{plots/covid_nowcast/00_true_data.pdf}
-    \caption{Realisations.\todo{Capitalize Date}}
+    \caption{Realisations.}
     \label{fig:app-covid-true}
@@
     \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
-    \caption{Same-day nowcasts. \todo{Capitalize Forecast Date}}
+    \caption{Same-day nowcasts.}
     \label{fig:app-covid-nowcast}
@@
 The upper row contains systolic measurements, and the lower row contains mean measurements.
-The columns contain the lags one, five, and 15 minutes.\todo{lag ausgeschrieben}}
+The columns contain the lags one, five, and 15 minutes.}
     \label{fig:app-mimic-4q}
@@
     \end{subfigure}\hspace{0.01\textwidth}
-    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the lags one, five, and 15 minutes. \todo{No l = ... in the legend}}
+    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the lags one, five, and 15 minutes. }
     \label{fig:app-mimic-cond-prob}

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
+\unsure[inline]{Der Beginn ist jetzt sehr auf Measurement fixiert.
+Könnte man das vielleicht umdrehen: Analysen von Measurement/Nowcast/Forecast sind im Moment vor allem bezüglich Abstand (RMSE; MAE; aber auch probabilistische Analyse).
+Wichtige Info ist aber auch auch, wie gut ein Anstieg/Fall vorhergesagt wird.
+Dann drauf eingehen, wie das bisher in den Bereichen gemacht wird.}
+
 The introduction of new medical measurement methods requires comprehensive evaluation and comparison with established standard measurement methods in order to ensure their validity and reliability. 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-Nowcasting methods are now widely used in other fields such as economics and medicine and focus on predictions for the present, the immediate future and the recent past (\cite{banbura2013now, bok2018macroeconomic}).
+Nowcasting methods are now widely used in other fields, such as economics and medicine, and focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now, bok2018macroeconomic}).
 Nowcasting can be used, for example, to predict important statistics such as the current economic situation (e.g.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 We conclude our paper with the last section.
-The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at ...
-available.
+The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at [Link will be inserted] available.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 The introduction of new medical measurement methods requires comprehensive evaluation and comparison with established standard measurement methods in order to ensure their validity and reliability. 
-To detect the systematic difference between one measurement methods with a gold standard method, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
-In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline,}).
-Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizes the differences between the methods in relation to their mean value (\cite{bland1986statistical}).
+To detect the systematic difference between one measurement method with a gold standard method, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
+In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
+Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}).
 
@@
 Ideally, the signs of the changes between two consecutive points in time (with a defined time interval) measured by the new method and the gold standard method should always be the same. 
-In this case, we call the new method trending perfect in line with the golden Stadnard method and refer to this type of agreement as trendability. 
-This type of consideration of the trend capability of a measurement method to the gold standard has also recently gained importance in the literature. 
+In this case, we call the new method trending perfect in line with the gold Stadnard method and refer to this type of agreement as trendability. 
+This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature. 
 In (\cite{hiraishi2021concordance}), the authors proposed a new concordance rate for the four-quadrant diagram, which is based on the multivariate normal distribution and takes into account the correlation between the data of the individual subjects.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 
-The above-described idea of trending is also of fundamental interest for the evaluation and comparison of forecasting and nowcasting methods. 
+The above-described idea of trending is also of fundamental interest for the evaluation and comparison of forecasting and methods. 
 While measurement series are created by recording data using measuring instruments or direct observation, forecasting focuses on predicting the future based on historical data and its patterns. 
@@
 the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.
-Unfortunately, like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is hardly considered in the literature.
+Unfortunately, like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge.
 
@@
 The presented methods will be applied to a variety of real data examples.
+The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at [Link will be inserted] available.
 
 The paper is structured as follows. 
-In section \ref{sec:trending} we formalize the concept of trending independently of the application domain and investigate quantification approaches for noise-free and noisy data. 
+In Section \ref{sec:trending}, we formalize the concept of trending independently of the application domain and investigate quantification approaches for noise-free and noisy data. 
 In particular, we introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries. 
@@
 We conclude our paper with the last section.
-The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at [Link will be inserted] available.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 While measurement series are created by recording data using measuring instruments or direct observation, forecasting focuses on predicting the future based on historical data and its patterns. 
-When forecasting at a specific point in time (forecast origin), the forecast value at the target time corresponds to a prediction of the change in the target variable between the forecast origin and the target time.
+When forecasting at a specific point in time, the forecast value at the target time corresponds to a prediction of the change in the target variable between the forecast origin and the target time. 
 This predicted change results from the difference between the forecast value at the forecast horizon and the current, actual value of the target variable at the forecast origin. 
 A forecast is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
-In practice, a forecasting method is usually evaluated in terms of its prediction accuracy, which is defined by a certain loss function and measures the difference between the prediction and the actual outcome of the target variable; an overview of the different measures can be found in \cite{hyndman2006another}.
-Prediction accuracy can be measured by scale-dependent measures such as the Root Mean Square Error (RMSE) or measures based on percentage errors such as the Root Mean Square Percentage Error (RMSPE). 
-All of these typical measures are based on the absolute difference between the prediction and the true values, locally or globally, and are not capable of assessing the trendability discussed above without further adjustment.
+The current practice is that a forecasting method is usually evaluated in terms of its prediction accuracy, which is defined by a certain loss function and measures the difference between the prediction and the actual outcome of the target variable; an overview of the different measures can be found in \cite{hyndman2006another}.
+Popular measures of this type are scale-dependent measures such as the Root Mean Square Error (RMSE) or measures based on percentage errors such as the Root Mean Square Percentage Error (RMSPE). 
+All of these typical measures are based on the absolute difference between the prediction and the true values, locally or globally, and are not capable of assessing the trendability discussed above.
 
-Another area in which the detection of trends can be of interest is nowcasting.
+The evaluation of trending is also interesting for the assessment of nowcasts.
 In contrast to conventional forecasting methods, nowcasting focuses on predicting the current status.
@@
 A nowcast is consistent with the development of the target variable if the difference in the actual value between today and yesterday, which is announced at a later date, has the same sign as the predicted change.
-It is obvious that a nowcast with strong trend capability provides crucial information on the development of the target variable (e.g.
+A nowcast with strong trending ability provides crucial information on the development of the target variable (e.g.
 the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.
-Unfortunately, like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge.
+like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge.
 
@@
 
-The paper is structured as follows. 
+The remainder of this paper is structured as follows. 
 In Section \ref{sec:trending}, we formalize the concept of trending independently of the application domain and investigate quantification approaches for noise-free and noisy data. 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 The above-described idea of trending is also of fundamental interest for the evaluation and comparison of forecasting and methods. 
-While measurement series are created by recording data using measuring instruments or direct observation, forecasting focuses on predicting the future based on historical data and its patterns. 
+While measurement series are created by recording data using measuring instruments or direct observation, Forecasting is about predicting the future based on historical data and its patterns, taking into account exogenous factors.
 When forecasting at a specific point in time, the forecast value at the target time corresponds to a prediction of the change in the target variable between the forecast origin and the target time. 
-This predicted change results from the difference between the forecast value at the forecast horizon and the current, actual value of the target variable at the forecast origin. 
+This predicted change results from the difference between the forecast value at the target time and the current, actual value of the target variable at the time of predicting. 
 A forecast is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
@@
 The evaluation of trending is also interesting for the assessment of nowcasts.
-In contrast to conventional forecasting methods, nowcasting focuses on predicting the current status.
+, nowcasting focuses on predicting the current status.
+
+Nowcasting methods are now widely used in other fields, such as economics and medicine, and focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now, bok2018macroeconomic, wolffram2023collaborative}).
 Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-Nowcasting methods are now widely used in other fields, such as economics and medicine, and focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now, bok2018macroeconomic}).
+
 Nowcasting can be used, for example, to predict important statistics such as the current economic situation (e.g.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 
-Nowcasting methods are now widely used in other fields, such as economics and medicine, and focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now, bok2018macroeconomic, wolffram2023collaborative}).
+In contrast to conventional forecasting methodsNowcasting methods are now widely used in other fields, such as economics and medicine, and focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now, bok2018macroeconomic, wolffram2023collaborative}).
 Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

--- 2024-03-07 sections/01-introduction.tex
@@
 The evaluation of trending is also interesting for the assessment of nowcasts.
-, nowcasting focuses on predicting the current status.
-
-Nowcasting methods are now widely used in other fields, such as economics and medicine, and focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now, bok2018macroeconomic, wolffram2023collaborative}).
+In contrast to conventional forecasting methods, Nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields, such as economics and medicine (\cite{bok2018macroeconomic, wolffram2023collaborative}).
 Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-
-Nowcasting can be used, for example, to predict important statistics such as the current economic situation (e.g.
-GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
-Nowcasting methods use indicators that are related to the target variable but can be collected at a higher frequency and published in a timely manner to produce and update early estimates for the target variable (e.g.
-GDP from the first quarter of 2024 ) during the relevant period (i.e.
-from January to March 2024) (\cite{castle2017forecasting}).
+In economics, Nowcasting can be used, for example, to predict important statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
+Nowcasting methods use indicators that are related to the target variable but can be collected at a higher frequency and published in a timely manner to produce and update early estimates for the target variable during the relevant period (\cite{castle2017forecasting}).
+For example, in the context of forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024, before the official statistics are published.
 In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
@@
 from today) is estimated on the basis of current preliminary measurements, which is only finalised at a later point in time.
-At the time of the forecast, estimates of the target variable can be made for today and possibly for yesterday (if the value for yesterday has not yet been finalised); the predicted change in the target variable is then the difference between the two nowcast values.
-A nowcast is consistent with the development of the target variable if the difference in the actual value between today and yesterday, which is announced at a later date, has the same sign as the predicted change.
+
 A nowcast with strong trending ability provides crucial information on the development of the target variable (e.g.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 In this case, we call the new method trending perfect in line with the gold Stadnard method and refer to this type of agreement as trendability. 
-This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature. 
+This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature \cite{saugel2015tracking, hiraishi2021concordance}. 
 In (\cite{hiraishi2021concordance}), the authors proposed a new concordance rate for the four-quadrant diagram, which is based on the multivariate normal distribution and takes into account the correlation between the data of the individual subjects.
@@
 from today) is estimated on the basis of current preliminary measurements, which is only finalised at a later point in time.
+At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts.
+A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change.
 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 In this case, we call the new method trending perfect in line with the gold Stadnard method and refer to this type of agreement as trendability. 
-This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature \cite{saugel2015tracking, hiraishi2021concordance}. 
-In (\cite{hiraishi2021concordance}), the authors proposed a new concordance rate for the four-quadrant diagram, which is based on the multivariate normal distribution and takes into account the correlation between the data of the individual subjects.
+This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature \cite{saugel2015tracking, saugel2018error, hiraishi2021concordance}.
+
+%In (\cite{hiraishi2021concordance}), the authors proposed a new concordance rate for the four-quadrant diagram, which is based on the multivariate normal distribution and takes into account the correlation between the data of the individual subjects.
 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 To detect the systematic difference between one measurement method with a gold standard method, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
+
 In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
@@
 Ideally, the signs of the changes between two consecutive points in time (with a defined time interval) measured by the new method and the gold standard method should always be the same. 
-In this case, we call the new method trending perfect in line with the gold Stadnard method and refer to this type of agreement as trendability. 
-This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature \cite{saugel2015tracking, saugel2018error, hiraishi2021concordance}.
+In this case, we call the new method trending perfect in line with the gold Standard method and refer to this type of agreement as trendability. 
+This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature (\cite{saugel2015tracking, saugel2018error, hiraishi2021concordance}).
 
@@
 However, regardless of the application domain, most conventional method comparisons and evaluations consider local absolute differences in different ways without directly assessing whether the correct direction of change is output.
-In this paper, we aim to formalize trending and present various methods for measuring and visualizing trending that can be applied to the evaluation of trending ability of methods in measurements, nowcasts and forecasts.
-The presented methods will be applied to a variety of real data examples.
-The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at [Link will be inserted] available.
+
+\subsection*{Contributions.} 
+The main contributions of this paper are summarized below
+\begin{itemize}
+    \item  we formalize the trending and present various methods for measuring and visualizing trending that can be applied to the evaluation of trending ability of methods in measurements, nowcasts and forecasts.
+    \item The presented methods will be applied to a variety of real data examples. 
+\item The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at [Link will be inserted] available.
+\end{itemize}
 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 \begin{itemize}
-    \item  we formalize the trending and present various methods for measuring and visualizing trending that can be applied to the evaluation of trending ability of methods in measurements, nowcasts and forecasts.
-    \item The presented methods will be applied to a variety of real data examples. 
+
+\item We formalise the concept of trending and show how the predicted changes can be meaningfully computed to assess trendability.
+
+\item We present different variants of trend measures that consider noiseless data as well as noisy and small non-informative changes.
+
+\item We introduce the conditional trend plot, a new graphical method for assessing local trends.
+
+\item We introduce bootstrap methods for calculating confidence intervals. Confidence intervals for the measurements from the previous sections.
+
+\item We extend the concept of trending to probabilistic predictions.
+
+\item We apply trend evaluation to three different applications: Measurement, nowcasting and prediction data.
+
 \item The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at [Link will be inserted] available.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 
-\item We present different variants of trend measures that consider noiseless data as well as noisy and small non-informative changes.
+\item We present different variants of trending measures that consider noiseless data as well as noisy and small non-informative changes.
 
-\item We introduce the conditional trend plot, a new graphical method for assessing local trends.
+\item We introduce the conditional trending plot, a new graphical method for assessing local trends.
 
-\item We introduce bootstrap methods for calculating confidence intervals. Confidence intervals for the measurements from the previous sections.
+\item We introduce bootstrap methods for calculating  confidence intervals for the presented treding ratios.
 
@@
 
-\item We apply trend evaluation to three different applications: Measurement, nowcasting and prediction data.
+\item We apply trending evaluation to three different applications: Measurement, nowcasting and prediction data.
 
-\item The source code for the replication of the results of our study as well as a ready-to-use code for trendability detection are available in our public repository at [Link will be inserted] available.
+\item We provide a ready-to-use code for trending detection. The code and the source code to replicate the results of our study are available in our public repository at ... .
 \end{itemize}
@@
 In Section \ref{sec:trending}, we formalize the concept of trending independently of the application domain and investigate quantification approaches for noise-free and noisy data.
+Noisy data refers to data where noise or unsystematic effects are present in the true values or predictions.
 In particular, we introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries. 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 
-The introduction of new medical measurement methods requires comprehensive evaluation and comparison with established standard measurement methods in order to ensure their validity and reliability. 
 To detect the systematic difference between one measurement method with a gold standard method, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
-
 In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
@@
 
-A nowcast with strong trending ability provides crucial information on the development of the target variable (e.g.
-the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.
 like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge.
+However, appropriate measures for assessing trending ability should be constructed, because a nowcast with strong trending ability provides crucial information about the development of the target variable (for example, the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.
 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 Dann drauf eingehen, wie das bisher in den Bereichen gemacht wird.}
+
+ While the introduction of new medical measurement methods requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally important for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
  While the introduction of new medical measurement methods requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally important for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
+Conventionally, the methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account.
+For example, the Continuous Ranked Probability Score (CRPS) can be used to evaluate the accuracy of probabilistic predictions.
+It is also based on the integrated squared differences between the cumulative distribution function (CDF) of the prediction and the Heaviside jump function.
+These measures all fail to take sufficient account of whether the right direction of change is being taken, which can often be highly relevant in practice.
 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 These measures all fail to take sufficient account of whether the right direction of change is being taken, which can often be highly relevant in practice.
+The information about whether an estimator (be it a measurement device, forecast or nowcas) can accurately predict the increase or decrease is important to determine whether a statement can be made about the direction of development of the actual value on the basis of the estimator's prediction results.
+In the following, we first provide an overview of the characteristics and evalutations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.
 
+Measurement aims to obtain accurate and reliable data about the current state of a system.
+Often a certain parameter or variable is measured at regular intervals over a certain period of time.
 To detect the systematic difference between one measurement method with a gold standard method, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 
+Measurement, prediction and forecasting methods are becoming increasingly important and sophisticated in a world of fast developing technologies and increasing data availability.
 While the introduction of new medical measurement methods requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally important for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
@@
 Often a certain parameter or variable is measured at regular intervals over a certain period of time.
-To detect the systematic difference between one measurement method with a gold standard method, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero (\cite{watson2010method}).
-In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
+In addition to the use of a paired t-test to determine systematic differences (\cite{watson2010method}) between two measurement series, various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
 Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}).
-
-One important aspect of testing the consistency of the new method with the gold standard is to adequately validate if the new method is measuring changes in the target variable in the right direction, when the gold standard detects changes during the measurement process.
+As already described above, one important aspect of testing the consistency of the new method with the gold standard is to adequately validate if the new method is measuring changes in the target variable in the right direction, when the gold standard detects changes during the measurement process.
 Ideally, the signs of the changes between two consecutive points in time (with a defined time interval) measured by the new method and the gold standard method should always be the same. 
@@
 A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change.
-
-like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge.
+Like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge.
 However, appropriate measures for assessing trending ability should be constructed, because a nowcast with strong trending ability provides crucial information about the development of the target variable (for example, the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.
@@
 
-\subsection*{Contributions.} 
-The main contributions of this paper are summarized below
+\textbf{Contributions.} The main contributions of this paper are summarized below
 \begin{itemize}

### 2024-03-07 sections/03-trending.tex
Update on Overleaf.

@@
 The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value. \unsure{Andere Formulierung für: one minus?}
+\unsure{vielleicht sowas: The probability $p_t$ of a positive change is given by the probability that the predicted value at the forecasting time is greater than the known value at the issure time. }
 As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.

### 2024-03-07 sections/03-trending.tex
Update on Overleaf.

@@
 Numerical measures for forecast quality are scoring rules.
-
+\unsure{wenn ich hier richtig verstehe, macht man hier quasi aus probabilitsicher Vorhersage zu y eine probabilistische Vorhersage dazu, wie wahrscheinlich die Veränderung positiv ist und dann bewerten.
+Kann man auch hier auch direkt das dem Bewertungsergbnis schießen, wie gut die Vorhersage auf negative Änderung ist?
+Wenn ja, vielleicht irgendwie erwähnen oder angeben.
+}
 The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\parencite{Brier1950}.
@@
 Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
+\unsure{ist das richtig?}
 The quantiles $q_p$ generally differ for each time step, but we omit an index here for ease of notation.
@@
         &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
-\end{align}
+\end{align} \unsure{vielleicht eher d(x,y) wegen einem Integralzeichen?}
 Thereby, the distributions are assumed to be independent.

### 2024-03-07 sections/03-trending.tex
Update on Overleaf.

@@
 \end{align} \unsure{vielleicht eher d(x,y) wegen einem Integralzeichen?}
+
+\unsure{Die Unabhängigkeitsannahme ist nicht zu streng? weil es ist ja eigentlich die Aufgabe von einenm guten Nowcaster diese Abhängigkeit zu modellieren oder?}
 Thereby, the distributions are assumed to be independent.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 These measures all fail to take sufficient account of whether the right direction of change is being taken, which can often be highly relevant in practice.
-The information about whether an estimator (be it a measurement device, forecast or nowcas) can accurately predict the increase or decrease is important to determine whether a statement can be made about the direction of development of the actual value on the basis of the estimator's prediction results.
+The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is important to determine whether a statement can be made about the direction of development of the actual value on the basis of the estimator's prediction results.
 In the following, we first provide an overview of the characteristics and evalutations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.
@@
 Measurement aims to obtain accurate and reliable data about the current state of a system.
-Often a certain parameter or variable is measured at regular intervals over a certain period of time.
+Often, a certain parameter or variable is measured at regular intervals over a certain period of time.
 In addition to the use of a paired t-test to determine systematic differences (\cite{watson2010method}) between two measurement series, various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 
-Measurement, prediction and forecasting methods are becoming increasingly important and sophisticated in a world of fast developing technologies and increasing data availability.
-While the introduction of new medical measurement methods requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally important for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
+Measurement, prediction, and forecasting methods are becoming increasingly important and sophisticated in a world of fast-developing technologies and increasing data availability.
+While introducing new medical measurement methods requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally essential for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
 Conventionally, the methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account.
@@
 It is also based on the integrated squared differences between the cumulative distribution function (CDF) of the prediction and the Heaviside jump function.
-These measures all fail to take sufficient account of whether the right direction of change is being taken, which can often be highly relevant in practice.
-The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is important to determine whether a statement can be made about the direction of development of the actual value on the basis of the estimator's prediction results.
-In the following, we first provide an overview of the characteristics and evalutations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.
+These measures all fail to consider whether the right direction of change is being taken, which can often be highly relevant in practice.
+The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when taking  of development of the actual value on the basis of the estimator's prediction results.
+In the following, we first provide an overview of the characteristics and evaluations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.
 

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 These measures all fail to consider whether the right direction of change is being taken, which can often be highly relevant in practice.
-The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when taking  of development of the actual value on the basis of the estimator's prediction results.
+The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when taking desi of development of the actual value on the basis of the estimator's prediction results.
 In the following, we first provide an overview of the characteristics and evaluations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

--- 2024-03-07 sections/01-introduction.tex
@@
 Conventionally, the methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account.
-For example, the Continuous Ranked Probability Score (CRPS) can be used to evaluate the accuracy of probabilistic predictions.
+For example, the continuously ranked probability score (CRPS) can be used to evaluate the accuracy of probabilistic predictions.
 It is also based on the integrated squared differences between the cumulative distribution function (CDF) of the prediction and the Heaviside jump function.
 These measures all fail to consider whether the right direction of change is being taken, which can often be highly relevant in practice.
-The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when taking  of development of the actual value on the basis of the estimator's prediction results.
+The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when taking decisions based  on  the estimator's prediction results.
 In the following, we first provide an overview of the characteristics and evaluations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.
@@
 Measurement aims to obtain accurate and reliable data about the current state of a system.
-Often, a certain parameter or variable is measured at regular intervals over a certain period of time.
-In addition to the use of a paired t-test to determine systematic differences (\cite{watson2010method}) between two measurement series, various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
-Furthermore, graphical tools were developed, such as the Bland-Altman diagram, which visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}).
-As already described above, one important aspect of testing the consistency of the new method with the gold standard is to adequately validate if the new method is measuring changes in the target variable in the right direction, when the gold standard detects changes during the measurement process.
-Ideally, the signs of the changes between two consecutive points in time (with a defined time interval) measured by the new method and the gold standard method should always be the same. 
-In this case, we call the new method trending perfect in line with the gold Standard method and refer to this type of agreement as trendability. 
-This type of consideration of the trend ability of a measurement method to the gold standard has also recently gained importance in the literature (\cite{saugel2015tracking, saugel2018error, hiraishi2021concordance}).
+Often, a parameter or variable is measured at regular intervals over a certain period.
+In addition to using a paired t-test to determine systematic differences between two measurement series (\cite{watson2010method}), various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
+Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}).
+As already described above, one important aspect of testing the consistency of the new method with the gold standard is to adequately validate if the new method is measuring changes in the target variable in the right direction when the gold standard detects changes during the measurement process.
+Ideally, the signs of the changes between two consecutive points in time with a prescribed lag measured by the new and gold standard methods should always be the same. 
+In this case, we call the new method trending perfect in line with the gold Standard method and refer to this type of agreement as trending ability. 
+This type of consideration of the trending ability of a measurement method to the gold standard has also recently gained importance in the literature (\cite{saugel2015tracking, saugel2018error, hiraishi2021concordance}).
 
@@
 
-The above-described idea of trending is also of fundamental interest for the evaluation and comparison of forecasting and methods. 
-While measurement series are created by recording data using measuring instruments or direct observation, Forecasting is about predicting the future based on historical data and its patterns, taking into account exogenous factors.
-When forecasting at a specific point in time, the forecast value at the target time corresponds to a prediction of the change in the target variable between the forecast origin and the target time. 
-This predicted change results from the difference between the forecast value at the target time and the current, actual value of the target variable at the time of predicting. 
+The above-described idea of trending is also of fundamental interest for evaluating and comparing forecasting and methods. 
+While measurement series are created by recording data using measuring instruments or direct observation, forecasting is about predicting the future based on historical data, its patterns, and exogenous factors.
+When forecasting at a specific time, the forecast value at the target time corresponds to a prediction of the change in the target variable between the forecast origin and the target time. 
+This predicted change results from the difference between the forecast value at the target time and the current actual value of the target variable at the time of prediction. 
 A forecast is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
@@
 Popular measures of this type are scale-dependent measures such as the Root Mean Square Error (RMSE) or measures based on percentage errors such as the Root Mean Square Percentage Error (RMSPE). 
-All of these typical measures are based on the absolute difference between the prediction and the true values, locally or globally, and are not capable of assessing the trendability discussed above.
+All of these typical measures are based on the absolute difference between the prediction and the true values, locally or globally.
+They are not capable of assessing the trending ability discussed above.
 
 The evaluation of trending is also interesting for the assessment of nowcasts.
-In contrast to conventional forecasting methods, Nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields, such as economics and medicine (\cite{bok2018macroeconomic, wolffram2023collaborative}).
-Nowcasting has its origins in meteorology, and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-In economics, Nowcasting can be used, for example, to predict important statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
-Nowcasting methods use indicators that are related to the target variable but can be collected at a higher frequency and published in a timely manner to produce and update early estimates for the target variable during the relevant period (\cite{castle2017forecasting}).
+In contrast to conventional forecasting methods, nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields such as economics and medicine (\cite{bok2018macroeconomic, wolffram2023collaborative}).
+Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
+In economics, nowcasting is used, for example, to predict statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
+Nowcasting methods use indicators related to the target variable but can be collected at a higher frequency and published promptly to produce and update early estimates for the target variable during the relevant period (\cite{castle2017forecasting}).
 For example, in the context of forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024, before the official statistics are published.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
 In economics, nowcasting is used, for example, to predict statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
-Nowcasting methods use indicators related to the target variable but can be collected at a higher frequency and published promptly to produce and update early estimates for the target variable during the relevant period (\cite{castle2017forecasting}).
-For example, in the context of forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024, before the official statistics are published.
-In medicine, epidemic nowcasting is used during an ongoing epidemic to assess the current situation, taking into account the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
-The nowcasting method can be used, for example, to correct the daily Covid case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}).
-In this context, nowcasting is an estimation procedure in which the value of a target variable (e.g.
-from today) is estimated on the basis of current preliminary measurements, which is only finalised at a later point in time.
+Nowcasting methods use indicators that are related to the target variable and collected at a higher frequency.
+Thus, the nowcasts can produce early estimates for the target variable during the relevant period (\cite{castle2017forecasting}).
+For example, in forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024 before the official statistics are published.
+In medicine, epidemic nowcasting is used to assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
+For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}).
+In this context, nowcasting is an estimation procedure in which the value of a target variable (for example, from today) is estimated based on current preliminary measurements, which are only finalized later.
 At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts.
 A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change.
-Like forecasts, nowcasts are often evaluated and compared in the literature on the basis of performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) and the aspect of trending ability is not considered in the literature to the best of our knowledge.
-However, appropriate measures for assessing trending ability should be constructed, because a nowcast with strong trending ability provides crucial information about the development of the target variable (for example, the number of cases of an epidemic) and can be a solid basis for decisions on the introduction or cancellation of policy measures.
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) or probabilistic measures \parencite{Wolffram2023}.
+The aspect of trending ability is not considered in the literature to the best of our knowledge.
+However, appropriate measures for assessing trending ability should be constructed because a nowcast with strong trending ability provides crucial information about the development of the target variable, for example, the number of cases of an epidemic, and can be the foundation of decisions on the introduction or cancellation of policy measures.
 
-As outlined above, the ability to take into account the trendability of a method depending on the area of application, be it measurements, predictions and forecasts, is of high importance. 
+As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
 However, regardless of the application domain, most conventional method comparisons and evaluations consider local absolute differences in different ways without directly assessing whether the correct direction of change is output.

### 2024-03-07 sections/01-introduction.tex
Update on Overleaf.

@@
-\unsure[inline]{Der Beginn ist jetzt sehr auf Measurement fixiert.
-Könnte man das vielleicht umdrehen: Analysen von Measurement/Nowcast/Forecast sind im Moment vor allem bezüglich Abstand (RMSE; MAE; aber auch probabilistische Analyse).
-Wichtige Info ist aber auch auch, wie gut ein Anstieg/Fall vorhergesagt wird.
-Dann drauf eingehen, wie das bisher in den Bereichen gemacht wird.}
-
 Measurement, prediction, and forecasting methods are becoming increasingly important and sophisticated in a world of fast-developing technologies and increasing data availability.
@@
 
-\textbf{Contributions.} The main contributions of this paper are summarized below
+The main contributions of this paper are summarized below
 \begin{itemize}
-
-\item We formalise the concept of trending and show how the predicted changes can be meaningfully computed to assess trendability.
-
-\item We present different variants of trending measures that consider noiseless data as well as noisy and small non-informative changes.
-
+\item We formalize trending and show how the predicted changes can be computed to assess trending for practical relevance. 
+\item We present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
 \item We introduce the conditional trending plot, a new graphical method for assessing local trends.
-
 \item We introduce bootstrap methods for calculating confidence intervals for the presented treding ratios.
-
 \item We extend the concept of trending to probabilistic predictions.
-
-\item We apply trending evaluation to three different applications: Measurement, nowcasting and prediction data.
-
+\item We apply trending evaluation to three applications: Measurement, nowcasting, and forecasting data. 
 \item We provide a ready-to-use code for trending detection. The code and the source code to replicate the results of our study are available in our public repository at ... .
@@
 Bootstrap methods for calculating confidence intervals for trending measures are also presented. 
-In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting and nowcasting. 
+In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting. 
 We conclude our paper with the last section.

### 2024-03-07 sections/03-trending.tex
Update on Overleaf.

@@
-As explained in Section~\ref{sec:introduction}, the term "trending" is used to denote the alignment of a detected true development with the development indicated by another measurement, a nowcast, or a forecast.
+As explained in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by another measurement, a nowcast, or a forecast.
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
@@
 Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
-Before introducing measures to evaluate trending, we make remarks on the calculation of the predicted change and notation in Section~\ref{subsec:notation}.
-Section~\ref{subsec:trending-basics} introduces the basics of trending and visualizes trending in a synthetic example.
+Before introducing measures to evaluate trending, we make remarks on calculating the predicted change and notation in Section~\ref{subsec:notation}.
+Section~\ref{subsec:trending-basics} introduces trending basics and visualizes trending in a synthetic example.
 In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
@@
 In the nowcasting application of Section~\ref{sec:application-covid}, for example, the time delay is more than 80 days.
-Until the true values are publicly available, the change could not be computed, and no actions could be taken based on it.
-Thus, it is more practically relevant to use updated signal values for time $t$ as the reference value in this setting.
+Until the true values are publicly available, the change cannot be computed, and no actions can be taken based on it.
+Thus, using updated signal values for time $t$ as the reference value in this setting is more practically relevant.
 
@@
 To account for several signals for a target time $t$ at different issue times $\tau$, we denote them by $x_{t | \tau}$.
-For nowcasting, target times $t$ are usually not considerably later than issue times $\tau$; while $t$ is usually at least $\tau$ in forecasting. For measurements, this distinction is not necessary.
+For nowcasting, target times $t$ are usually not considerably later than issue times $\tau$, while $t$ is usually at least $\tau$ in forecasting. For measurements, this distinction is not necessary.
 Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
@@
 Signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
-Note the different relations of issue and target time for nowcasts and forecasts.
+Note the different relations between the issue and target time for nowcasts and forecasts.
 }
@@
 In the following, we omit the difference lag $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$ as defined in Table~\ref{tab:notation}.
-As noted above, a signal's trending is perfect if all the directions of change that occur are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+As noted above, a signal's trending is perfect if all the change directions are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
 Perfect trending is hardly the case in practice, but a weaker requirement is needed.
@@
 For example, it is known in cardiac output measurement analysis~\parencite{Saugel2015,perrino1998intraoperative}. 
-Thus, the occured changes are plotted together with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
+Thus, the occured changes are plotted with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
 Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
@@
 The four-quadrant plot is intuitive to interpret, and the magnitude and direction of change are shown simultaneously.
-It can be extended by including information on the date of difference in the color of the point.
+It can be extended by including information on the date in the point color.
 In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$.
@@
 There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.
-However, they lack the clarity and intuition of the four-quadrant plot, without adding any more information on trending~\parencite{Saugel2015}.
+However, they lack the four-quadrant plot's clarity and intuition, without adding more information on trending~\parencite{Saugel2015}.
 
@@
 Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data \parencite{Critchley2010,Saugel2015}. 
-It boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes, respectively.
+It boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
 Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$).
@@
 
-The use of these estimators could encourage predictors to exploit imbalances in the number of positive and negative changes. 
+Using these estimators could encourage predictors to exploit imbalances in the number of positive and negative changes. 
 A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t \in \mathcal{T}} \ind{\diffyt < 0}$ is unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data.
@@
 There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
-Cohen's $\kappa$ is usually used to measure inter-rater agreement and takes into account the ratio of occurred agreement and the probability of agreement by chance.
-Cohen's $\kappa$ reduces to rescaling the trending ratio in the case of a $2\times2$-table and balanced outcomes.
+Cohen's $\kappa$ is usually used to measure inter-rater agreement and considers the ratio of occurred agreement and the probability of agreement by chance.
+Cohen's $\kappa$ reduces to rescaling the trending ratio for a $2\times2$-table and balanced outcomes.
 
@@
 The trending ability of the signal has a strong sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
-This seasonal behavior cannot be distinguished in the two four-quadrant plots or in the trending ratio.
+This seasonal behavior cannot be distinguished in the two four-quadrant plots or the trending ratio.
 
@@
 However, systematical missing values could lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
-In this case, measures should be interpreted with caution, and
+In this case, measures should be interpreted cautiously, and
 inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
@@
 However, points close to the zero point have intuitively less explanatory power or are less reliable.
-If there are noise or non-systematic effects present in the true values or predictions, a point's assignment to a quadrant can be driven by noise instead of a systematic trending ability of the signal.
-For points with at least one small coordinate, this happens more frequently.
+Suppose there are noise or non-systematic effects present in the true values or predictions.
+In that case, a point's assignment to a quadrant can be driven by noise instead of a systematic trending ability of the signal.
+This happens more frequently for points with at least one small coordinate.
 

### 2024-03-07 sections/03-trending.tex
Update on Overleaf.

@@
 While it would be possible to build analog measures for other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
-Furthermore, the graph facilitates the comparison of various methods in a single graph, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
+Furthermore, the graph facilitates the comparison of various methods in a single graph and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
 We refer to the plot as a conditional trending plot.
@@
 The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ for all values of $x \in \R$.
-Multivariate \ac{kde} takes as modeling parameters a kernel and a bandwidth selector. 
-The kernel is usually Gaussian, and there are multiple methods for selecting the bandwidth, each with its own advantages and disadvantages.
+Multivariate \ac{kde} takes a kernel and bandwidth selector as modeling parameters. 
+The kernel is usually Gaussian, and there are multiple methods for selecting the bandwidth, each with its advantages and disadvantages.
 Note that for kernels with infinite support, for example, the Gaussian kernel, the plot should be limited to the core values of $\diffx$ without outliers to avoid estimates based on small subsets of the data.
@@
 The \ac{cdf} is the most general and can be used to derive the others, given they exist.
-Let us first assume that the prediction is a \ac{cdf} and the lag is large enough such that the true value is known when issuing the forecast or nowcast.
+Let us first assume that the prediction is a \ac{cdf} and the lag is large enough to know the true value when issuing the forecast or nowcast.
 See the end of the section for predictions given as quantiles or for unknown true values.
@@
 For trending assessment, we compare the probability of positive change with the occurrence of positive changes. 
-The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value. \unsure{Andere Formulierung für: one minus?}
-\unsure{vielleicht sowas: The probability $p_t$ of a positive change is given by the probability that the predicted value at the forecasting time is greater than the known value at the issure time. }
+The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value at the issue time. 
 As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
@@
 Numerical measures for forecast quality are scoring rules.
-\unsure{wenn ich hier richtig verstehe, macht man hier quasi aus probabilitsicher Vorhersage zu y eine probabilistische Vorhersage dazu, wie wahrscheinlich die Veränderung positiv ist und dann bewerten.
-Kann man auch hier auch direkt das dem Bewertungsergbnis schießen, wie gut die Vorhersage auf negative Änderung ist?
-Wenn ja, vielleicht irgendwie erwähnen oder angeben.
-}
 The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\parencite{Brier1950}.

### 2024-03-07 paper.tex
Update on Overleaf.

@@
 % Language setting
-% Replace `english' with e.g. `spanish' to change the document language
+% Replace `englis' with e.g. `spanish' to change the document language
 \usepackage[english]{babel}
@@
 \newcommand{\pmax}{\hat{p}} % Notation for the grid resolution in quantiles
+\newcommand{\pc}{p^c}
 
@@
 \begin{abstract}
-Existing and new measures for the capability of capturing the trend of nowcasts are presented, evaluated, and compared on synthetic and real-world data.
+Trending refers to the ability of a signal to predict the direction of change for the true variable accurately. 
+An appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers. 
+While the current practice of evaluating signals is based on averaging over time-step-separated measures, we review and introduce trending measures that consider time series differences for certain lags.
+We formalize the concept of trending for all application areas and quantify the trending ability for signals. 
+We review and present new visualization techniques to assess the local trending behaviors and quantify estimation uncertainty by bootstrapping. 
+We apply the methods to three large data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
 \end{abstract}
@@
 
+We formalize the concept of trending independently of the application domain and investigate quantification approaches for noise-free and noisy data.
+We introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries.
+Bootstrap methods for calculating confidence intervals for trending measures are also presented.
+We show the results of applying our method to several practical examples.
+
 \section{Introduction}\label{sec:introduction}
+\todo{State that we want to accompany the current evaluation with a new measure, not replace other measures}
 \input{sections/01-introduction}
@@
 %\input{sections/02-notation}
-
-\newpage
 

### 2024-03-07 paper.tex
Update on Overleaf.

@@
 
-We formalize the concept of trending independently of the application domain and investigate quantification approaches for noise-free and noisy data.
-We introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries.
-Bootstrap methods for calculating confidence intervals for trending measures are also presented.
-We show the results of applying our method to several practical examples.
-
 \section{Introduction}\label{sec:introduction}

### 2024-03-08 sections/03-trending.tex
Update on Overleaf.

@@
 Thereby, the distributions are assumed to be independent.
+If the nowcasts have the form of a multivariate distribution including the dependence of the 
 As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.

### 2024-03-08 sections/03-trending.tex
Update on Overleaf.

--- 2024-03-07 sections/03-trending.tex
@@
 There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.
-However, they lack the four-quadrant plot's clarity and intuition, without adding more information on trending~\parencite{Saugel2015}.
+However, they lack the four-quadrant plot's clarity and intuition without adding more information on trending~\parencite{Saugel2015}.
 
@@
 We refer to time steps for which either signal or true values or both are unavailable as missing values.
-Pairs of data with missing values can be excluded from the set $\mathcal{T}$ to calculate the measures. 
+Data pairs with missing values can be excluded from the set $\mathcal{T}$ to calculate the measures. 
 However, systematical missing values could lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
@@
 However, points close to the zero point have intuitively less explanatory power or are less reliable.
-Suppose there are noise or non-systematic effects present in the true values or predictions.
+Suppose noise or non-systematic effects are present in the true values or predictions.
 In that case, a point's assignment to a quadrant can be driven by noise instead of a systematic trending ability of the signal.
@@
     \bar{p}_t
-        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
+        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: (x_1, x_2) \nonumber\\
         &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
@@
 Thereby, the distributions are assumed to be independent.
+If the nowcasts have the form of a multivariate distribution, including the dependence of the two \acp{pdf}, $f_{t+l|t+l} (x_2)$ has to be replaced by the \ac{pdf} conditional on $x_1$.
 As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.

### 2024-03-08 sections/03-trending.tex
Update on Overleaf.

@@
 Third, for measurement, a test method is usually evaluated against a gold standard to test whether the test method could replace the gold standard.
-Thus, the gold standard measurements are not available if the quality of the test method is sufficient.
-\improvement{irgendwie für mich nicht klar formuliert: man unteruscht ja bei meausurement nur ob die gemessenen Änderungen von den zwei Methoden konsistent sind oder?}
+Thus, the gold standard measurements are not available if the quality of the test method is sufficient. \improvement{Änderung des Testgeräts an sich beurteilen}
 Thus, the predicted change is computed based on the values of the test method at time $t$ and $t + l$.
@@
 
-\subsection{Accounting for noise and non-informative small changes}\label{subsec:trending-noise}
+\subsection{Accounting for noise and non-informative small changes and bootstrap confidence intervals}\label{subsec:trending-noise}
 
@@
 
-\subsection{Bootstrap confidence intervals and the conditional trending plot}\label{subsec:trending-cond-prob}
-\unsure{Vielleicht doch zwei subsections?}
 Confidence intervals can account for the estimation uncertainty of the measures above.
@@
 Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
+
+\subsection{The conditional trending plot}\label{subsec:trending-cond-prob}
 
@@
 The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time.
-}\label{fig:trending-cond-prob-bw}
+\todo{$\varepsilon$ hinzufügen}}\label{fig:trending-cond-prob-bw}
 \end{figure}
@@
         &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
-\end{align} \unsure{vielleicht eher d(x,y) wegen einem Integralzeichen?}
-
+\end{align}
 \unsure{Die Unabhängigkeitsannahme ist nicht zu streng? weil es ist ja eigentlich die Aufgabe von einenm guten Nowcaster diese Abhängigkeit zu modellieren oder?}

### 2024-03-08 sections/01-introduction.tex
Update on Overleaf.

@@
-Measurement, prediction, and forecasting methods are becoming increasingly important and sophisticated in a world of fast-developing technologies and increasing data availability.
-While introducing new medical measurement methods requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally essential for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
+The evaluation of measurement, prediction and forecasting methods become increasingly important and sophisticated, as technologies and data availability enable their application in more and more areas.
+While introducing new medical measurement methods already requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally essential for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
 Conventionally, the methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account.
 For example, the continuously ranked probability score (CRPS) can be used to evaluate the accuracy of probabilistic predictions.
-It is also based on the integrated squared differences between the cumulative distribution function (CDF) of the prediction and the Heaviside jump function.
+%It is also based on the integrated squared differences between the cumulative distribution function (CDF) of the prediction and the Heaviside jump function.
 These measures all fail to consider whether the right direction of change is being taken, which can often be highly relevant in practice.
 The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when taking decisions based  on  the estimator's prediction results.
-In the following, we first provide an overview of the characteristics and evaluations of the three areas: Measurement, Prediction and Nowcasting and their relationship with Trending.
+In the following, we provide an overview of the characteristics and evaluations of measurement, forecasting and nowcasting methods and show their relationships to trending.
 
@@
 A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change.
-Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the root mean absolute error (\cite{gunther2021nowcasting}) or probabilistic measures \parencite{Wolffram2023}.
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the RMSE (\cite{gunther2021nowcasting}) or probabilistic measures \parencite{Wolffram2023}.
 The aspect of trending ability is not considered in the literature to the best of our knowledge.

### 2024-03-08 sections/01-introduction.tex
Update on Overleaf.

@@
 As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
-However, regardless of the application domain, most conventional method comparisons and evaluations consider local absolute differences in different ways without directly assessing whether the correct direction of change is output.
+However, regardless of the area of application, most conventional method comparisons and assessments consider local absolute differences in different ways without directly assessing whether the correct direction of change is being predicted or measured.In this paper, we want to complement the current evaluation measures with trending measurements and propose the evaluation methods that assess the trendability of a method.
 

### 2024-03-08 sections/03-trending.tex
Update on Overleaf.

@@
 Note the different relations between the issue and target time for nowcasts and forecasts.
-}
+\todo{$T-l$}}
     \label{tab:notation}
@@
 We refer to this estimator as the trending ratio of the prediction and set $\mathcal{T} = \{1, \dots, T-l\}$.
+\todo{Durchwalken; nennen, warum ratio; Anteil concordant points an allen, deswegen nennen wir es ratio}
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
@@
 There, a wide range of other methods are usually used to analyze further characteristics of contingency tables.
-Two simple measures that focus on specific areas of interest are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
+Two simple measures that focus on specific areas of interest \todo{welche areas} are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
 They are defined as
@@
 
-All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
+All measures can also be evaluated as a rolling estimate to detect changes in performance over time. \todo{Klarer, warum man das soll. Wir wollen eigentlich trending ratio jetzt wissen. }
 A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
-For the trending ratio, a rolling estimate with a backward-looking window is given by
+For the trending ratio, a rolling estimate with a backward-looking \todo{Klarer, was heißt backward-looking} window is given by
 \begin{equation*}
@@
 \end{equation*}
+\todo{Was heißt w? Großes w weniger fokussiert auf current situation; kleines w lokaler}
 Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
@@
 \end{align}
-\unsure{Die Unabhängigkeitsannahme ist nicht zu streng? weil es ist ja eigentlich die Aufgabe von einenm guten Nowcaster diese Abhängigkeit zu modellieren oder?}
 Thereby, the distributions are assumed to be independent.

### 2024-03-14 sections/03-trending.tex
Update on Overleaf.
Restructure trending section

--- 2024-03-08 sections/03-trending.tex
@@
-As explained in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by another measurement, a nowcast, or a forecast.
+As discussed in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by another measurement, a nowcast, or a forecast.
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
-In general, trending can be assessed for different lags $l$ separately, and the relevance of the considered lags has to be determined based on the question at hand.
+The development is always with respect to a time horizon, that is, the amount of time passed during the development. 
+In general, trending can be assessed for different horizons $l$ separately, and the relevance of the considered horizons has to be determined based on the question at hand.
 Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
@@
 
-\subsection{Computing the predicted change and notation}\label{subsec:notation}
+\subsection{Computing changes and notation}\label{subsec:notation}
 
-As noted above, trending refers to observed and predicted changes for a time lag $l$.
+As noted above, trending refers to observed and predicted changes over a time horizon $l$.
 The \textit{observed} change is straightforward to compute for all types of signals.
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values for nowcasting or forecasting, or gold standard measurements up to time $T$.
-The sequence of observed change is then given by the differences of values in $\mathbf{y}$ with lag $l$, that is
+The sequence of observed changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is
 \begin{equation}\label{eq:diffy}
-    \diffylt = (y_{t+l} - y_t) \quad \text{for}\ t = 1, \dots, T - l.
+    \diffylt = (y_{t} - y_{t-l}) \quad \text{for}\ t = l, \dots, T.
 \end{equation}
-
-As noted in the introduction, the definition of the predicted change depends on the context.
-Applying the computation for $y$ directly, the signal's predicted change would be the difference between the signal at time $t + l$ and the true value at $t$.
-This approach lags in three practical aspects.
-For nowcasting and forecasting, the true value at time $t$ might not be known with a considerable time delay.
-In the nowcasting application of Section~\ref{sec:application-covid}, for example, the time delay is more than 80 days.
-Until the true values are publicly available, the change cannot be computed, and no actions can be taken based on it.
-Thus, using updated signal values for time $t$ as the reference value in this setting is more practically relevant.
-
-Second, in the context of nowcasting or forecasting, several signals for the same time $t + l$ are often available at different generation times.
-Thus, the generation time and forecast time have to be included in the notation of nowcasts and forecasts.
-We use the term \textit{issue time} for the generation time of a signal and \textit{target time} to denote the time the forecast applies to.
-
-Third, for measurement, a test method is usually evaluated against a gold standard to test whether the test method could replace the gold standard.
-Thus, the gold standard measurements are not available if the quality of the test method is sufficient. \improvement{Änderung des Testgeräts an sich beurteilen}
-Thus, the predicted change is computed based on the values of the test method at time $t$ and $t + l$.
-
-Taking the three points together, the signal's predicted change is based on the type of signal.
-To account for several signals for a target time $t$ at different issue times $\tau$, we denote them by $x_{t | \tau}$.
-For nowcasting, target times $t$ are usually not considerably later than issue times $\tau$, while $t$ is usually at least $\tau$ in forecasting. For measurements, this distinction is not necessary.
-Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
 
@@
     \centering
+    \scriptsize
     \begin{tabularx}{0.75\textwidth}{l X}
@@
         \midrule
-        Measurement & $(x_{t + l} - x_t)_{t=1}^T \label{eq:diffxl_measure} $\\
+        Measurement & $(x_{t} - x_{t-l})_{t=l}^T $\\
         Nowcasting & $
@@
 \begin{cases}
-(x_{t+l|t+l} -x_{t|t+l})^T_{t=1} & \text{if } y_{t} \text{ is not known at time } t, \\
-(x_{t+l|t+l} -y_{t})^T_{t=1} & \text{else}.
-\end{cases} \label{eq:diffxl_nowcasting}
+(x_{t|t} - x_{t-l|t})^T_{t=l}, & \text{if } y_{t-l} \text{ is not known at time } t, \\
+(x_{t|t} - y_{t-l})^T_{t=1}, & \text{otherwise}.
+\end{cases} 
 $ \\
-        Forecasting & $
-\diffxl =
-\begin{cases}
-(x_{t+l|t} - x_{t|t})^T_{t=1} & \text{if } y_{t} \text{ is not known at time } t, \\
-(x_{t+l|t} - y_{t})^T_{t=1}  & \text{else}.
-\end{cases} \label{eq:diffxl_forecasting}
-$\\
+        Forecasting & $ \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l}$\\
         \bottomrule
@@
     \caption{Computation of the predicted change in the different applications.
-Signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
-Note the different relations between the issue and target time for nowcasts and forecasts.
-}
+For nowcasting and forecasting, signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
+For measurement, $x_t$ denotes the test device measurement at time $t$.}
     \label{tab:notation}
@@
 
+The definition of the predicted change depends on the context.
+Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
+For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$. 
+We call $t$ the \textit{forecast time} and $\tau$ the \textit{issue time}.
+We compute the predicted change by
+\begin{equation}
+  \diffxl =
+\begin{cases}
+(x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
+(x_{t|t} - y_{t-l})^T_{t=1} & \text{otherwise}.
+\end{cases} \label{eq:diffxl_nowcasting}
+\end{equation}
+When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$ and the true value might not be known yet. 
+Thus, the predicted can be computed with the knowledge of time $t$ and the trending assessment can be used in practice at time $t$.
+
+For forecasting, the notation is similar: Let $x_{t | \tau}$ denote the forecast for forecast time $t$ and issue time $\tau$. 
+The predicted change is computed by 
+\begin{equation} \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l} \label{eq:diffxl_forecasting}
+\end{equation}
+The indices are chosen to be consistent with $\diffyl$. 
+If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification as in Equation~\eqref{eq:diffxl_nowcasting} can be made.
+
+In measurement analysis, the distinction between forecast and issue time is not necessary, as the measurement is typically available at the time. 
+Let, thus, $x_t$ denote the test method measurement for time $t$. 
+The computation
+\begin{equation}
+  (x_{t} - x_{t-l})_{t=l}^T \label{eq:diffxl_measure}
+\end{equation}
+yields the change by the test method. 
+It is computed purely by the test method without the gold standard $y_t$ to analyse whether the changes by gold standard and test method are consistent.
+
 \subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-basics}
 
-Although the calculation of $\diffxl$ differs in different contexts, the trending assessment is similar.
-In the following, we omit the difference lag $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common lag $l$ as defined in Table~\ref{tab:notation}.
+The trending assessment is the same for all three application areas, given the notation for the respective application of Section~\ref{subsec:notation}.
+In the following, we omit the difference horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$ as defined in Table~\ref{tab:notation}.
 As noted above, a signal's trending is perfect if all the change directions are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
-Perfect trending is hardly the case in practice, but a weaker requirement is needed.
 Consequently, when evaluating the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
-A simple yet insightful method is the four-quadrant plot.
-For example, it is known in cardiac output measurement analysis~\parencite{Saugel2015,perrino1998intraoperative}. 
-Thus, the occured changes are plotted with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
+A simple yet insightful method is the four-quadrant plot~\parencite[see, e.g., ][]{Saugel2015,perrino1998intraoperative}. 
+In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
 Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
@@
 In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$.
-The lag is always one time step.
+The horizon is always one time step.
 However, four-quadrant plots become crowded for larger datasets, and sequential information on the differences is complex to assess thoroughly.
@@
 Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data \parencite{Critchley2010,Saugel2015}. 
-It boils down to assessing the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
+With that, we mainly estimate the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
 Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$).
@@
 \end{equation}
-We refer to this estimator as the trending ratio of the prediction and set $\mathcal{T} = \{1, \dots, T-l\}$.
+The enumerator counts the number of same-sign-changes, while the denominator is the number of all points.
+Thus, $\acc$ is the ratio of concordant changes over all changes.
+We refer to this estimator as the trending ratio of the prediction and set $\mathcal{T} = \{l, \dots, T\}$. 
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
@@
 There, a wide range of other methods are usually used to analyze further characteristics of contingency tables.
-Two simple measures that focus on specific areas of interest are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
+Two simple measures that focus on either a positive or negative predicted change are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
 They are defined as
@@
 In the classification context, these measures are known as positive or negative predictive value and hit rate or detection failure ratio in forecasting.
-They give the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, respectively.
-Therefore, they measure the trending ability of the prediction for positive and negative changes separately by cutting the considered data.
-In terms of probability, the two measures give estimates of $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$, respectively.
+They give the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, respectively, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$, respectively.
 
-Using these estimators could encourage predictors to exploit imbalances in the number of positive and negative changes. 
-A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t \in \mathcal{T}} \ind{\diffyt < 0}$ is unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data.
-However, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
-There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
-Cohen's $\kappa$ is usually used to measure inter-rater agreement and considers the ratio of occurred agreement and the probability of agreement by chance.
-Cohen's $\kappa$ reduces to rescaling the trending ratio for a $2\times2$-table and balanced outcomes.
-
-All measures can also be evaluated as a rolling estimate to detect changes in performance over time.
-A rolling estimate is a sequence of estimates, each computed on a length-$w$-window of the data.
-For the trending ratio, a rolling estimate with a backward-looking window is given by
+Rolling estimates of the above measures detect changes in performance over time and can give a sharper estimate of the current trending ability.
+For the trending ratio, a rolling estimate with a backward-looking window of length $w$ at time $t$ is given by
 \begin{equation*}
-    \acc_t (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w}, \quad t = w-1, \dots, T.\label{eq:acc_rolling}
+    \acc_{t; w} (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w}.\label{eq:acc_rolling}
 \end{equation*}
+Backward-looking windows provide an estimate at a time $t$ considering the $w$ time steps before time $t$.
+The window length $w$ controls the smoothing of the estimate, a larger $w$ gives smoother results, while a small $w$ focus on local variations.
+Plotting the rolling estimates for $t = w-1, \dots, T$ yields an estimate of the trending ability over time.
 Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
@@
 However, systematical missing values could lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
-In this case, measures should be interpreted cautiously, and
-inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
+In this case, measures should be interpreted cautiously, and inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
 In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}, only a few missing values occur, and we provide information on the missing data by inspecting the missing data as lists.
@@
 
-The above measures use only information on a point's quadrant and neglect further details on its location within the quadrant.
-However, points close to the zero point have intuitively less explanatory power or are less reliable.
+The above measures can be extended to account for information on the points location within the quadrant.
+For example, points close to the zero point may have less explanatory power or are less reliable than points far away from zero on one of the diagonals.
 Suppose noise or non-systematic effects are present in the true values or predictions. 
@@
 
-Using an exclusion area around the zero point is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures}.
+Using an exclusion area around the zero point as further defined below is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures}.
 Points within that area are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
-The shape of the exclusion area can be chosen according to the noise characteristics of the predictions and the true values.
 In general, it is likely that the prediction models have a noise component and should thus be part of the exclusion area.
@@
 The estimators accept various shapes of the exclusion area.
-Figure~\ref{fig:trending_4q} visualizes different shapes of the exclusion area.
+Figure~\ref{fig:trending_4q} visualizes examples of the exclusion area.
 A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ 
@@
 
-In addition to these straight shapes, exclusion areas could have more complex shapes, such as ellipses.
-Such a shape could, for example, be obtained by using a multivariate normal error model around the zero point and excluding points that exceed a certain likelihood of being produced by chance from the zero point.
-These shapes are theoretically appealing, but interpreting the resulting estimators becomes complex.
-Thus, we suggest using the above simple exclusion areas.
-
 In most applications, the shape and size of the exclusion area can be chosen based on domain knowledge or expert opinions.
-The size estimation can also be based on a proportion of the total variance or the total range of the data.
+The size estimation can also be based on a proportion of the total variance or the total range of the data, for example, the points with at least one coordinate in the smallest 10\% of absolute values are excluded.
 A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
+For examples on those plots, see Section~\ref{sec:application-covid}.
 
@@
 New samples are drawn with replacement from the dataset and are not based on parametric assumptions as classical confidence intervals are.
-The confidence interval is then computed based on the bootstrap samples.
-We examine three methods for bootstrapping: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
-In the \textit{percentile} approach, the confidence interval for the level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
-The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
-The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
-Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
-The basic approach balances these two objectives.
-We compare the approaches in a small synthetic data study on their small-dataset behavior and computation time in Appendix~\ref{sec:appendix-trending}.
-\Ac{bca} is the only method that maintains the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
-Therefore, we use the \ac{bca} method for confidence intervals in the applications in Section~\ref{sec:application}.
-
-The sample is assumed to be independent and identically distributed in standard bootstrapping.
-Thus, before applying bootstrapping methods, the strength of sequential dependence should be inspected, for example, by analyzing the autocorrelation and partial autocorrelation.
-The simple bootstrapping methods above do not account for serial dependence. 
-Bootstrapping focusing on time series data is covered in~\textcite{Hardle2003,Kreiss2012}, for example.
+The confidence interval is then computed based on those bootstrap samples.
+We use the \ac{bca} approach for bootstrapping in the following, as it hold the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-trending-bootstrap}).
 
@@
 
-The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin e)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin e)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin e)$.
-These probabilities draw a general picture but might still be too coarse for local effects.
-More insights might be gained by considering the conditional distribution to assess the local trending ability of a prediction.
-While it would be possible to build analog measures for other intervals, assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
-Furthermore, the graph facilitates the comparison of various methods in a single graph and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
-We refer to the plot as a conditional trending plot.
-
+The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
+A still more fine and local analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ to assess the local trending ability of a prediction.
 A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
@@
 for $x \neq 0$ through a \ac{kde}.
-
 A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages~\parencite[e.g., for  Python in][]{Seabold2010}.
@@
 Multivariate \ac{kde} takes a kernel and bandwidth selector as modeling parameters. 
-The kernel is usually Gaussian, and there are multiple methods for selecting the bandwidth, each with its advantages and disadvantages.
-Note that for kernels with infinite support, for example, the Gaussian kernel, the plot should be limited to the core values of $\diffx$ without outliers to avoid estimates based on small subsets of the data.
+We advise using a Gaussian kernel and the cross-validation maximum likelihood as bandwith selector (see Appendix~\ref{subsec:appendix-kde}).
 
-Figure~\ref{fig:trending-cond-prob-bw} shows the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\parencite{Seabold2010}.
-While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.
-The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
-The second method, cross-validation least squares, requires long computation times while yielding small or no bandwidth results, even for two relatively small datasets.
-The rule-of-thumb and cross-validation maximum likelihood methods yield reasonable results at moderate computation times. 
-Further examples, including comparisons between methods regarding their trending ability, are available in Section~\ref{sec:application}.
+Assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
+Furthermore, the graph facilitates the comparison of various methods in a single graph and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
+We refer to the plot as a conditional trending plot.
 
-\begin{figure}
-    \centering
-    \begin{subfigure}{.48\textwidth}
-        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_asym_butterfly}
-        \caption{First dataset with asymmetric dependence.}
-    \end{subfigure}
-    \begin{subfigure}{.48\textwidth}
-        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_normal}
-        \caption{Second dataset. }
-    \end{subfigure}
-    \caption{Resulting conditional trending plot for different bandwidth selection processes.
-Cross-validation least squares takes a considerably larger computation time.
-It does not converge for the first data set and yields a bandwidth too small for the second data set.
-The rule of thumb is the fastest method but tends to oversmooth.
-The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time.
-\todo{$\varepsilon$ hinzufügen}}\label{fig:trending-cond-prob-bw}
-\end{figure}
-
-\newpage
 \subsection{Probabilistic Evaluation}\label{subsec:probabilistic}
 
-In nowcasting and forecasting, there is a shift from point predictions to probabilistic predictions (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
+In nowcasting and forecasting, probabilistic predictions become more popular in the recent years (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
 Probabilistic predictions include a point estimate and information on the forecast uncertainty and quantiles simultaneously.
+For trending analysis, probabilistic predictions also contain a probability of a positive or negative change.
 In this section, we extend the trending assessment to probabilistic predictions.
@@
 The \ac{cdf} is the most general and can be used to derive the others, given they exist.
-Let us first assume that the prediction is a \ac{cdf} and the lag is large enough to know the true value when issuing the forecast or nowcast.
-See the end of the section for predictions given as quantiles or for unknown true values.
+Let us first assume that the prediction is a \ac{cdf} and the horizon is large enough to know the true value when issuing the forecast or nowcast.
+In the appendix, we extend the analysis for predictions given as quantiles or for unknown true values.
 
-Let $F_{t | \tau} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $\tau$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
+Let $F_{t | t-l} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
 For trending assessment, we compare the probability of positive change with the occurrence of positive changes. 
-The probability $p_t$ of a positive change is given by one minus the \ac{cdf} of the prediction at the true value at the issue time. 
+A positive change occurs if any value larger than the true value $y_{t-l}$ occurs. 
 As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
@@
 \begin{equation*}
-    p_t = 1 - F_{t+l | t+l} (y_{t})\quad t = 1, \dots, T-l;
+    p_t = 1 - F_{t | t} (y_{t-l})\quad t = l, \dots, T;
 \end{equation*}
@@
 \begin{equation*}
-    p_t = 1 - F_{t+l | t} (y_{t})\quad t = 1, \dots, T-l.
+    p_t = 1 - F_{t | t-l} (y_{t-l})\quad t = l, \dots, T.
 \end{equation*}
@@
 \begin{equation*}
-    z_t = \ind{\diffyt > 0} \quad t = 1, \dots, T-l.
+    z_t = \ind{\diffyt > 0} \quad t = l, \dots, T.
 \end{equation*}
-The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=1}^{T-l}$ and the observed changes $\mathbf{z} = (z_t)_{t=1}^{T-l}$.
+The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=l}^{T}$ and the observed changes $\mathbf{z} = (z_t)_{t=l}^{T}$.
 The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
@@
 Numerical measures for forecast quality are scoring rules.
+
 The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\parencite{Brier1950}.
@@
 \begin{equation*}
-    BS (\mathbf{p}, \mathbf{z}) = \frac{1}{T-l} \sum_{t=1}^{T-l} (p_t - z_t)^2.
+    BS (\mathbf{p}, \mathbf{z}) = \frac{1}{T-l} \sum_{t=l}^{T} (p_t - z_t)^2.
 \end{equation*}
@@
 The basic paradigm of probabilistic forecasting is to maximize sharpness, subject to calibration~\parencite{Gneiting2014}.
+
 These two conflicting aspects cannot be distinguished in the scoring rule.
-
-Graphical models are a standard tool for evaluating solely the calibration of probabilistic forecasts.
+Graphical methods are a standard tool for evaluating solely the calibration of probabilistic forecasts.
 In dichotomous forecasting, the reliability diagram is frequently used~\parencite{Ranjan2010}.
@@
 For an example of a reliability diagram, see Section~\ref{sec:application-eda}.
-
-If forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
-Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
-\unsure{ist das richtig?}
-The quantiles $q_p$ generally differ for each time step, but we omit an index here for ease of notation.
-The probability $\pc_t$ of a \textit{negative} change is between $p^{\star}$ and $p^{\star} + 1/\pmax$ for
-\begin{equation*}
-    p^{\star} = \max \{p \in \{1/\pmax, \dots, (\pmax-1) / \pmax\}: q_p \leq y_t\} , \quad \text{if}\ q_{1/\pmax} \leq y_t \leq q_{1 - 1/\pmax}.
-\end{equation*}
-Quantiles do not determine the location within the interval $[p^{\star}, p^{\star} + 1/\pmax]$.
-Under the assumption of a uniform distribution within the quantile interval, the probability of a negative change is
-\begin{equation*}
-    \bar{p}_t = \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
-\end{equation*}
-The approach does not assign probabilities for $y_t$ smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile.
-As a simple extension, we assume that the probability mass is uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
-This yields
-\begin{equation*}
-\bar{p}_t = \begin{cases}
-    \max \{\frac{1}{\pmax} - \frac{q_{p^\star} - y_t}{\pmax (q_{p2/\pmax} - q_{1/\pmax})}, 0\} &, \text{if } y_t < q_{1/p}, \\
-    \min \{\frac{1}{\pmax} - \frac{y_t - q_{(\pmax-1)/\pmax}}{\pmax (q_{(\pmax-1)/\pmax} - q_{(\pmax-2)/\pmax})}, 1\} &, \text{if } y_t > q_{1 - 1/p}, \\
-    \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star} &, \text{otherwise.}
-\end{cases}
-\end{equation*}
-The probability of positive change is $p_t = 1 - \bar{p}_t$.
-
-If the true value is given as a distribution because it is still unknown, the probabilities $p_t$ can be computed by integration.
-Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
-Then, the probability of a negative change can be computed by
-\begin{align}
-    \bar{p}_t
-        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: (x_1, x_2) \nonumber\\
-        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
-        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
-\end{align}
-\unsure{Die Unabhängigkeitsannahme ist nicht zu streng? weil es ist ja eigentlich die Aufgabe von einenm guten Nowcaster diese Abhängigkeit zu modellieren oder?}
-Thereby, the distributions are assumed to be independent.
-If the nowcasts have the form of a multivariate distribution, including the dependence of the two \acp{pdf}, $f_{t+l|t+l} (x_2)$ has to be replaced by the \ac{pdf} conditional on $x_1$.
-As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
-For forecasts, the indexes have to be shifted.
-If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to generate samples for the Monte Carlo approximation.

### 2024-03-14 sections/04-application.tex
Restructure trending section

@@
 In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
-We report the trending for the lags 1, 7, and 14 days.
-While lag one assesses the short-term trending, lags seven and 14 consider the medium-term trending.
-The lags seven and 14 are particularly interesting, reflecting a usual period until new policy changes are taken.
+We report the trending for the horizons 1, 7, and 14 days.
+While horizon one assesses the short-term trending, horizons seven and 14 consider the medium-term trending.
+The horizons seven and 14 are particularly interesting, reflecting a usual period until new policy changes are taken.
 
-Before stating the results, we provide background information on the marginal distributions of the true values and nowcasts for the different lags.
+Before stating the results, we provide background information on the marginal distributions of the true values and nowcasts for the different horizons.
 Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} provides information on marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
-Overall, the variability and general level of differences grow with the lag.
-The standard deviation increases from roughly 300 for lag one to 1,200 for lag seven and 2,000 for lag 14 days.
+Overall, the variability and general level of differences grow with the horizon.
+The standard deviation increases from roughly 300 for horizon one to 1,200 for horizon seven and 2,000 for horizon 14 days.
 Similarly, the 10\%-quantile of differences increases.
@@
 
-Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the lag of seven days.
-The trending ratios without exclusion area range from 0.72 to 0.85 for the lag of seven days.
+Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the horizon of seven days.
+The trending ratios without exclusion area range from 0.72 to 0.85 for the horizon of seven days.
 The negative trending ratios are higher than the positive trending ratios for all models.
@@
 The positive trending ratio implies a similar ranking of the models, while the negative ratio is second best for the RKI model.
-For the lags of one and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
+For the horizons of one and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
 
-Figure~\ref{fig:app-covid-cond-prob-trending-ratio-7} shows the conditional trending plots and the trending ratio over the exclusion area for the lag seven days; the respective plots for the lags one day and 14 days are shown in Figure~\ref{fig:app-covid-cond-prob-trending-ratio-1-14}.
+Figure~\ref{fig:app-covid-cond-prob-trending-ratio-7} shows the conditional trending plots and the trending ratio over the exclusion area for the horizon seven days; the respective plots for the horizons one day and 14 days are shown in Figure~\ref{fig:app-covid-cond-prob-trending-ratio-1-14}.
 Here, only the best models in point evaluation measures, ILM, RKI, RIVM, and ENS-MED, are shown to keep the plots readable.
@@
 The opposite is the case for a nowcasted hospitalization rate increase, and the difference between the models' performance is higher.
-Similar observations can be made for the lag of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
-For a lag of one day, the models' conditional trending ability difference is less pronounced (see Figure~\ref{fig:app-covid-cond-prob-1}).
+Similar observations can be made for the horizon of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
+For a horizon of one day, the models' conditional trending ability difference is less pronounced (see Figure~\ref{fig:app-covid-cond-prob-1}).
 The RKI model is still less conclusive when issuing an increase in the hospitalization rate, while RIVM is most informative in that case.
@@
 While the RIVM and ENS-MED trending ratios evolve similarly, the RKI and ILM trending ratios get closer.
-For the lag of one day, the RKI trending ratio decreases with increasing exclusion area size while the other models rise (see Figure~\ref{fig:app-covid-trending-ratio-1}).
-For the lag of 14 days, all trending ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-trending-ratio-14}).
+For the horizon of one day, the RKI trending ratio decreases with increasing exclusion area size while the other models rise (see Figure~\ref{fig:app-covid-trending-ratio-1}).
+For the horizon of 14 days, all trending ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-trending-ratio-14}).
 
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models with and without exclusion areas for the lag seven days.
+    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models with and without exclusion areas for the horizon seven days.
 The exclusion areas are rectangles centered on the zero points with a width and height of 10\%-quantile of the absolute values of nowcast and true values.
@@
     \end{subfigure}
-    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the lag seven days.}
+    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
     \label{fig:app-covid-cond-prob-trending-ratio-7}
@@
 
-For all lags, the influence of the exclusion area on the 10\%-quantile level is negligible.
+For all horizons, the influence of the exclusion area on the 10\%-quantile level is negligible.
 For example, the trending ratio changes at most by 0.03 for the EPI model with $\accm[14]$.
@@
 The number of arrivals has a strong weekly and daily pattern.
-Thus, we consider the lags of 72 hours, the last already observed shift of the same hour of day, and seven days, the previous shift of the same hour and day.
+Thus, we consider the horizons of 72 hours, the last already observed shift of the same hour of day, and seven days, the previous shift of the same hour and day.
 Table~\ref{tab:app-eda-point-evaluation} lists the point evaluation measures and the count of available forecasts.
@@
 
-Table~\ref{tab:app-eda-marginals} analyzes the differences in marginal distributions for the forecasts and true values for the lags of three and seven days.
+Table~\ref{tab:app-eda-marginals} analyzes the differences in marginal distributions for the forecasts and true values for the horizons of three and seven days.
 Note that the difference definition aligns with Section~\ref{subsec:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the true value is available when issuing the forecast.
-The fraction of positive differences varies between 0.39 and 0.63 for the lag of three days and between 0.37 and 0.63 for the lag of seven days.
-The variability of differences decreases for the larger lag for most models; only for the ETS model does it increase.
-The 10\%-quantile of the differences is between zero and one for all models and lags.
+The fraction of positive differences varies between 0.39 and 0.63 for the horizon of three days and between 0.37 and 0.63 for the horizon of seven days.
+The variability of differences decreases for the larger horizon for most models; only for the ETS model does it increase.
+The 10\%-quantile of the differences is between zero and one for all models and horizons.
 Thus, we exclude only differences smaller than one from the trending assessment.
@@
     \caption{Marginal analysis of the nowcast and true differences.
-The column (1) shows the fraction of values greater than zero for lag $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
+The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
     \label{tab:app-eda-marginals}
@@
 
-Table~\ref{tab:app-eda-trending-ratios} lists the trending ratios for all models for three and seven-day lags.
-The trending ratios range from 0.68 to 0.84 for a lag of three days and from 0.68 to 0.82 for seven days.
-The negative and positive trending ratios differ for all models and lags.
+Table~\ref{tab:app-eda-trending-ratios} lists the trending ratios for all models for three and seven-day horizons.
+The trending ratios range from 0.68 to 0.84 for a horizon of three days and from 0.68 to 0.82 for seven days.
+The negative and positive trending ratios differ for all models and horizons.
 For some models, for example, the GBM-2 model, the positive trending ratio is higher; for some models, for example, the tbats model, the negative trending ratio is higher.
@@
 Other models exhibit a higher positive trending ratio than the negative trending ratio.
-The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for three and seven days lag, while Poisson-2 and NBI-2 have the highest negative trending ratio.
+The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for three and seven days horizon, while Poisson-2 and NBI-2 have the highest negative trending ratio.
 
-Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags three and seven days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
-The conditional trending plots show similar courses for the two lags, while the curves are shifted downwards for the lag of seven days.
-The model's relative trending ability evolves consistently for the two lags, with the NBI-2 and Poisson-2 models being indistinguishable.
+Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
+The conditional trending plots show similar courses for the two horizons, while the curves are shifted downwards for the horizon of seven days.
+The model's relative trending ability evolves consistently for the two horizons, with the NBI-2 and Poisson-2 models being indistinguishable.
 The GBM-2 model outperforms the qreg-1 model for all $x$.
@@
 Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
-The \acf{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the lags three and seven days are shown in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
-The \acp{bs} are smallest and close for NBI-2 and Poisson-2 for both lags, while the Brier scores for the other models are larger and differ more.
-The qreg-1 model has the highest \ac{bs} for both lags.
+The \acf{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days are shown in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
+The \acp{bs} are smallest and close for NBI-2 and Poisson-2 for both horizons, while the Brier scores for the other models are larger and differ more.
+The qreg-1 model has the highest \ac{bs} for both horizons.
 The reliability diagrams of GBM-2 and NBI-2 are also close and show a too-small fraction of increases for the predicted probability overall.
@@
     \input{plots/ed_arrival/50_trending_ratio}
-    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the lags 72 hours and seven days.}
+    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days.}
     \label{tab:app-eda-trending-ratios}
@@
     \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_3}
-    \caption{Lag three days}
+    \caption{Horizon three days}
     \end{subfigure}\hfill
@@
     \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_7}
-    \caption{Lag seven days}
+    \caption{Horizon seven days}
     \end{subfigure}
-    \caption{Conditional trending plots for the lags three and seven days and the models with the best positive or negative trending ability. The plots of NBI-2 and Poisson-2 are indistinguishable.}
+    \caption{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability.
+The plots of NBI-2 and Poisson-2 are indistinguishable.}
     \label{fig:app-eda-cond-prob}
@@
     \input{plots/ed_arrival/70_brier_scores}
-    \caption{Brier Scores for the different models and lags.}\label{fig:app-eda-prob-brier}
+    \caption{Brier Scores for the different models and horizons.}\label{fig:app-eda-prob-brier}
     \end{subfigure}\hspace{0.01\textwidth}%
@@
     \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_3}
-    \caption{Reliability diagram for lag three days.}\label{fig:app-eda-prob-rel-3}
+    \caption{Reliability diagram for horizon three days.}\label{fig:app-eda-prob-rel-3}
     \end{subfigure}\hspace{0.01\textwidth}%
@@
     \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_7}
-    \caption{Reliability diagram for lag seven days.}\label{fig:app-eda-prob-rel-7}
+    \caption{Reliability diagram for horizon seven days.}\label{fig:app-eda-prob-rel-7}
     \end{subfigure}
-    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the lags three and seven days.}
+    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days.}
     \label{fig:app-eda-prob}
@@
 
-The trending ability is consistent for the two lags, with the models' relative trending ability evolving similarly for the two lags.
-The models' trending ability is generally higher for the smaller lag, but the differences are minor, and confidence intervals overlap.
+The trending ability is consistent for the two horizons, with the models' relative trending ability evolving similarly for the two horizons.
+The models' trending ability is generally higher for the smaller horizon, but the differences are minor, and confidence intervals overlap.
 
@@
 
-We consider the lags of one minute, five minutes, and 15 minutes for the trending assessment, as those are typical intervals of NBP measurements.
+We consider the horizons of one minute, five minutes, and 15 minutes for the trending assessment, as those are typical intervals of NBP measurements.
 
@@
 Again, we exclude the smallest 10\% of absolute differences in trending assessment.
-The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different lags are shown in Figure~\ref{fig:app-mimic-4q}.
-The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} simultaneously for two consecutive times with the specified lags.
-For the systolic measurements, 290, 332, and 442 points are available for the lags of one, five, and 15 minutes; for the mean measurements, 406, 430, and 542.
+The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different horizons are shown in Figure~\ref{fig:app-mimic-4q}.
+The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} simultaneously for two consecutive times with the specified horizons.
+For the systolic measurements, 290, 332, and 442 points are available for the horizons of one, five, and 15 minutes; for the mean measurements, 406, 430, and 542.
 
-The trending ratios, including confidence intervals for the different lags, are listed in Table~\ref{tab:app-mimic-trending-ratios}.
-For the measurements with a lag of one minute, the confidence intervals have lower bounds of 0.5 or slightly above.
-For larger lags, the trending ratio increases.
-The difference between positive and negative trending ratios is small for all types and lags, with overlapping confidence intervals.
+The trending ratios, including confidence intervals for the different horizons, are listed in Table~\ref{tab:app-mimic-trending-ratios}.
+For the measurements with a horizon of one minute, the confidence intervals have lower bounds of 0.5 or slightly above.
+For larger horizons, the trending ratio increases.
+The difference between positive and negative trending ratios is small for all types and horizons, with overlapping confidence intervals.
 
-Figure~\ref{fig:app-mimic-cond-prob} shows the conditional trending plots for the different lags and the systolic and mean blood pressure measurements.
+Figure~\ref{fig:app-mimic-cond-prob} shows the conditional trending plots for the different horizons and the systolic and mean blood pressure measurements.
 It becomes apparent that the systolic measurements have a higher trending ability than the mean measurement, except for small negative predicted changes.
@@
     \includegraphics{plots/mimic/plot_4q}
-    \caption{Four-quadrant plots for the different lags and the systolic and mean blood pressure measurements.
+    \caption{Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements.
 The upper row contains systolic measurements, and the lower row contains mean measurements.
-The columns contain the lags one, five, and 15 minutes.}
+The columns contain the horizons one, five, and 15 minutes.}
     \label{fig:app-mimic-4q}
@@
     \input{plots/mimic/trending_ratio}
-    \caption{Trending ratios for the different lags and the systolic and mean blood pressure measurements.}
+    \caption{Trending ratios for the different horizons and the systolic and mean blood pressure measurements.}
     \label{tab:app-mimic-trending-ratios}
@@
         \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag1}
-        \caption{Lag one minute.}
+        \caption{Horizon one minute.}
     \end{subfigure}\hspace{0.01\textwidth}
@@
         \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag5}
-        \caption{Lag five minutes.}
+        \caption{Horizon five minutes.}
     \end{subfigure}\hspace{0.01\textwidth}
@@
         \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag15}
-        \caption{Lag 15 minutes.}
+        \caption{Horizon 15 minutes.}
     \end{subfigure}\hspace{0.01\textwidth}
-    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the lags one, five, and 15 minutes. }
+    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the horizons one, five, and 15 minutes. }
     \label{fig:app-mimic-cond-prob}
@@
 The bootstrap confidence intervals are wide.
-The width is around 0.1 for the trending ratio, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the lag of one minute.
+The width is around 0.1 for the trending ratio, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the horizon of one minute.
 Thus, more measurements would be of interest for a further trending assessment.

### 2024-03-14 sections/05-conclusion.tex
Restructure trending section

\begin{itemize}
  \item Summary of work
  \item highlight important general remarks of section trending / application
  \item points for further research/other methods: imbalances in the number of positive and negative changes; sequential dependence of the differences
\end{itemize}

\subsection{Modeling remarks}

Using these estimators could encourage predictors to exploit imbalances in the number of positive and negative changes. 
A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t \in \mathcal{T}} \ind{\diffyt < 0}$ is unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data. 
However, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
Cohen's $\kappa$ is usually used to measure inter-rater agreement and considers the ratio of occurred agreement and the probability of agreement by chance.
Cohen's $\kappa$ reduces to rescaling the trending ratio for a $2\times2$-table and balanced outcomes.

### 2024-03-14 sections/50-appendix-data-generation.tex
Update on Overleaf.
Restructure trending section

@@
 
-\subsection{Simulation study on bootstrapping confidence intervals}\label{subsec:app-trending-bootstrap}
+\subsection{Simulation study on bootstrapping confidence intervals}\label{subsec:appendix-trending-bootstrap}
+
+We examine three methods for bootstrapping: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
+In the \textit{percentile} approach, the confidence interval for the level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
+The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
+The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
+Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
+The basic approach balances these two objectives.
+We compare the approaches in a small synthetic data study on their small-dataset behavior and computation time.
+
 We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
@@
 The computation time for the \ac{bca} method is slightly larger than for the other methods, but all methods have a moderate computation time.
+\Ac{bca} is the only method that maintains the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
+Therefore, we use the \ac{bca} method for confidence intervals in the applications in Section~\ref{sec:application}.
 
@@
 
-\begin{table}
+\subsection{Visualization of different bandwidth selectors in multivariate \ac{kde}}\label{subsec:appendix-kde}
+
+\begin{figure}
     \centering
@@
     \label{fig:trending_bootstrap_time}
-\end{table}
+\end{figure}
+
+\subsection{Probabilistic trending evaluation}
+
+If forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
+Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
+\unsure{ist das richtig?}
+The quantiles $q_p$ generally differ for each time step, but we omit an index here for ease of notation.
+The probability $\pc_t$ of a \textit{negative} change is between $p^{\star}$ and $p^{\star} + 1/\pmax$ for
+\begin{equation*}
+    p^{\star} = \max \{p \in \{1/\pmax, \dots, (\pmax-1) / \pmax\}: q_p \leq y_t\} , \quad \text{if}\ q_{1/\pmax} \leq y_t \leq q_{1 - 1/\pmax}.
+\end{equation*}
+Quantiles do not determine the location within the interval $[p^{\star}, p^{\star} + 1/\pmax]$.
+Under the assumption of a uniform distribution within the quantile interval, the probability of a negative change is
+\begin{equation*}
+    \bar{p}_t = \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
+\end{equation*}
+The approach does not assign probabilities for $y_t$ smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile.
+As a simple extension, we assume that the probability mass is uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
+This yields
+\begin{equation*}
+\bar{p}_t = \begin{cases}
+    \max \{\frac{1}{\pmax} - \frac{q_{p^\star} - y_t}{\pmax (q_{p2/\pmax} - q_{1/\pmax})}, 0\} &, \text{if } y_t < q_{1/p}, \\
+    \min \{\frac{1}{\pmax} - \frac{y_t - q_{(\pmax-1)/\pmax}}{\pmax (q_{(\pmax-1)/\pmax} - q_{(\pmax-2)/\pmax})}, 1\} &, \text{if } y_t > q_{1 - 1/p}, \\
+    \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star} &, \text{otherwise.}
+\end{cases}
+\end{equation*}
+The probability of positive change is $p_t = 1 - \bar{p}_t$.
+
+If the true value is given as a distribution because it is still unknown, the probabilities $p_t$ can be computed by integration.
+Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
+Then, the probability of a negative change can be computed by
+\begin{align}
+    \bar{p}_t
+        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: (x_1, x_2) \nonumber\\
+        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
+        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
+\end{align}
+Thereby, the distributions are assumed to be independent.
+If the nowcasts have the form of a multivariate distribution, including the dependence of the two \acp{pdf}, $f_{t+l|t+l} (x_2)$ has to be replaced by the \ac{pdf} conditional on $x_1$.
+As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
+For forecasts, the indexes have to be shifted.
+If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to generate samples for the Monte Carlo approximation.

### 2024-03-14 sections/03-trending.tex
Update on Overleaf.

@@
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
-The development is always with respect to a time horizon, that is, the amount of time passed during the development. 
-In general, trending can be assessed for different horizons $l$ separately, and the relevance of the considered horizons has to be determined based on the question at hand.
+Development always has respect to a time horizon, that is, the amount of time passed during development. 
+In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the question at hand.
 Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
-Before introducing measures to evaluate trending, we make remarks on calculating the predicted change and notation in Section~\ref{subsec:notation}.
+Before introducing measures to evaluate trending, we remark on calculating the predicted change and notation in Section~\ref{subsec:notation}.
 Section~\ref{subsec:trending-basics} introduces trending basics and visualizes trending in a synthetic example.
@@
 \end{equation}
-When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$ and the true value might not be known yet. 
-Thus, the predicted can be computed with the knowledge of time $t$ and the trending assessment can be used in practice at time $t$.
+When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet. 
+Thus, the predicted can be computed with the knowledge of time $t$, and the trending assessment can be used in practice at time $t$.
 
-For forecasting, the notation is similar: Let $x_{t | \tau}$ denote the forecast for forecast time $t$ and issue time $\tau$. 
+The notation is similar for forecasting: Let $x_{t | \tau}$ denote the forecast for forecast time $t$ and issue time $\tau$. 
 The predicted change is computed by 
@@
 The indices are chosen to be consistent with $\diffyl$. 
-If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification as in Equation~\eqref{eq:diffxl_nowcasting} can be made.
+If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification can be made as in Equation~\eqref{eq:diffxl_nowcasting}.
 
-In measurement analysis, the distinction between forecast and issue time is not necessary, as the measurement is typically available at the time. 
-Let, thus, $x_t$ denote the test method measurement for time $t$. 
+The distinction between forecast and issue time is unnecessary in measurement analysis, as the measurement is typically available with a very short time lag. 
+Thus, $x_t$ denotes the test method measurement for time $t$. 
 The computation
@@
 yields the change by the test method. 
-It is computed purely by the test method without the gold standard $y_t$ to analyse whether the changes by gold standard and test method are consistent.
+It is computed purely by the test method without the gold standard $y_t$ to analyze whether the gold standard and test method changes are consistent.
 
@@
 \end{equation}
-The enumerator counts the number of same-sign-changes, while the denominator is the number of all points.
+The numerator counts the number of same-sign-changes, while the denominator is the number of all points.
 Thus, $\acc$ is the ratio of concordant changes over all changes.

### 2024-03-14 sections/03-trending.tex
Update on Overleaf.

@@
 Backward-looking windows provide an estimate at a time $t$ considering the $w$ time steps before time $t$.
-The window length $w$ controls the smoothing of the estimate, a larger $w$ gives smoother results, while a small $w$ focus on local variations.
+The window length $w$ controls the smoothing of the estimate; a larger $w$ gives smoother results, while a small $w$ focuses on local variations.
 Plotting the rolling estimates for $t = w-1, \dots, T$ yields an estimate of the trending ability over time.
@@
 In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
-We refer to time steps for which either signal or true values or both are unavailable as missing values.
+We refer to time steps for either signal or true values, or both are unavailable as missing values.
 Data pairs with missing values can be excluded from the set $\mathcal{T}$ to calculate the measures. 
@@
 
-The above measures can be extended to account for information on the points location within the quadrant.
+The above measures can be extended to account for information on the point's location within the quadrant.
 For example, points close to the zero point may have less explanatory power or are less reliable than points far away from zero on one of the diagonals.
@@
 
-Using an exclusion area around the zero point as further defined below is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures}.
+Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures}.
 Points within that area are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
@@
 In most applications, the shape and size of the exclusion area can be chosen based on domain knowledge or expert opinions.
-The size estimation can also be based on a proportion of the total variance or the total range of the data, for example, the points with at least one coordinate in the smallest 10\% of absolute values are excluded.
+The size estimation can also be based on a proportion of the total variance or the total range of the data; for example, the points with at least one coordinate in the smallest 10\% of absolute values are excluded.
 A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
@@
 The confidence interval is then computed based on those bootstrap samples.
-We use the \ac{bca} approach for bootstrapping in the following, as it hold the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-trending-bootstrap}).
+We use the \ac{bca} approach for bootstrapping in the following, as it holds the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-trending-bootstrap}).
 
@@
 Multivariate \ac{kde} takes a kernel and bandwidth selector as modeling parameters. 
-We advise using a Gaussian kernel and the cross-validation maximum likelihood as bandwith selector (see Appendix~\ref{subsec:appendix-kde}).
+We advise using a Gaussian kernel and the cross-validation maximum likelihood as bandwidth selector (see Appendix~\ref{subsec:appendix-kde}).
 
@@
 
-In nowcasting and forecasting, probabilistic predictions become more popular in the recent years (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
+In nowcasting and forecasting, probabilistic predictions have become more prevalent in recent years (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
 Probabilistic predictions include a point estimate and information on the forecast uncertainty and quantiles simultaneously.
@@
 The \ac{cdf} is the most general and can be used to derive the others, given they exist.
-Let us first assume that the prediction is a \ac{cdf} and the horizon is large enough to know the true value when issuing the forecast or nowcast.
-In the appendix, we extend the analysis for predictions given as quantiles or for unknown true values.
+Let us first assume that the prediction is a \ac{cdf}, and the horizon is large enough to know the true value when issuing the forecast or nowcast.
+In the appendix, we extend the analysis for predictions given as quantiles or unknown true values.
 
@@
 Calibration refers to the statistical consistency of forecasts and observations, while sharpness measures the spread of the forecast distribution.
-A smaller spread, and thus, higher sharpness, is preferable, as it indicates greater confidence in the prediction.
+A smaller spread, thus, higher sharpness, is preferable, as it indicates greater confidence in the prediction.
 The basic paradigm of probabilistic forecasting is to maximize sharpness, subject to calibration~\parencite{Gneiting2014}.

### 2024-03-14 sections/01-introduction.tex
Update on Overleaf.

--- 2024-03-08 sections/01-introduction.tex
@@
-The evaluation of measurement, prediction and forecasting methods become increasingly important and sophisticated, as technologies and data availability enable their application in more and more areas.
+The evaluation of measurement, prediction, and forecasting methods become increasingly important and sophisticated as technologies and data availability enable their application in more and more areas.
 While introducing new medical measurement methods already requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally essential for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
@@
 These measures all fail to consider whether the right direction of change is being taken, which can often be highly relevant in practice.
-The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when taking decisions based  on  the estimator's prediction results.
-In the following, we provide an overview of the characteristics and evaluations of measurement, forecasting and nowcasting methods and show their relationships to trending.
+The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when making decisions based on the estimator's prediction results.
+In the following, we provide an overview of the characteristics and evaluations of measurement, forecasting, and nowcasting methods and show their relationships to trending.
 
@@
 As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
-However, regardless of the application domain, most conventional method comparisons and evaluations consider local absolute differences in different ways without directly assessing whether the correct direction of change is output.
+However, regardless of the area of application, most conventional method comparisons and assessments consider local absolute differences in different ways without directly assessing whether the correct direction of change is being predicted or measured.
+In this paper, we want to complement the current evaluation measures with trending measurements and propose evaluation methods that assess the trending ability of a method.
 

### 2024-03-14 sections/01-introduction.tex
Update on Overleaf.

@@
 For example, the continuously ranked probability score (CRPS) can be used to evaluate the accuracy of probabilistic predictions.
-%It is also based on the integrated squared differences between the cumulative distribution function (CDF) of the prediction and the Heaviside jump function.
-These measures all fail to consider whether the right direction of change is being taken, which can often be highly relevant in practice.
+%It is also based on the integrated squared differences between the prediction's cumulative distribution function (CDF) and the Heaviside jump function.
+These measures fail to consider whether the right direction of change is being taken, which can often be highly relevant in practice.
 The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when making decisions based on the estimator's prediction results.
@@
 Measurement aims to obtain accurate and reliable data about the current state of a system.
-Often, a parameter or variable is measured at regular intervals over a certain period.
+Often, a parameter or variable is measured regularly over a certain period.
 In addition to using a paired t-test to determine systematic differences between two measurement series (\cite{watson2010method}), various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
@@
 
-The above-described idea of trending is also of fundamental interest for evaluating and comparing forecasting and methods. 
+The above-described trending idea is also of fundamental interest for evaluating and comparing forecasting and methods. 
 While measurement series are created by recording data using measuring instruments or direct observation, forecasting is about predicting the future based on historical data, its patterns, and exogenous factors.
@@
 For example, in forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024 before the official statistics are published.
-In medicine, epidemic nowcasting is used to assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
+In medicine, epidemic nowcasting assesses the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
 For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}).
@@
 As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
-However, regardless of the area of application, most conventional method comparisons and assessments consider local absolute differences in different ways without directly assessing whether the correct direction of change is being predicted or measured.
+However, regardless of the application area, most conventional method comparisons and assessments consider local absolute differences differently without directly assessing whether the correct direction of change is being predicted or measured.
 In this paper, we want to complement the current evaluation measures with trending measurements and propose evaluation methods that assess the trending ability of a method.

### 2024-03-15 paper.tex
Restructure trending section
Update paper

@@
 %\usepackage[hidelinks]{hyperref}
-\usepackage[style=authoryear,backend=biber]{biblatex}
+\usepackage[style=authoryear,backend=bibtex]{biblatex}
 \addbibresource{library.bib}
@@
 \DeclareAcronym{bs}{short=BS, long=Brier score}
+\DeclareAcronym{rmse}{short=RMSE, long=root mean square error}
 
@@
 An appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers. 
-While the current practice of evaluating signals is based on averaging over time-step-separated measures, we review and introduce trending measures that consider time series differences for certain lags.
+While the current practice of evaluating signals is based on averaging over time-step-separated measures, we review and introduce trending measures that consider whether the direction of development fits the quantity of interest.
 We formalize the concept of trending for all application areas and quantify the trending ability for signals. 
@@
 \section{Introduction}\label{sec:introduction}
-\todo{State that we want to accompany the current evaluation with a new measure, not replace other measures}
 \input{sections/01-introduction}
@@
 
-\todo{Write conclusion.}
+\input{sections/05-conclusion.tex}
 

### 2024-03-15 sections/01-introduction.tex
Update paper

@@
 The evaluation of measurement, prediction, and forecasting methods become increasingly important and sophisticated as technologies and data availability enable their application in more and more areas.
-While introducing new medical measurement methods already requires a comprehensive evaluation and comparison with established standard measurement methods to ensure their validity and reliability, it is generally essential for forecasts and nowcasting methods to assess the extent to which the predictions correspond to the development of reality.
-Conventionally, the methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account.
-For example, the continuously ranked probability score (CRPS) can be used to evaluate the accuracy of probabilistic predictions.
-%It is also based on the integrated squared differences between the prediction's cumulative distribution function (CDF) and the Heaviside jump function.
-These measures fail to consider whether the right direction of change is being taken, which can often be highly relevant in practice.
-The information about whether an estimator, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when making decisions based on the estimator's prediction results.
-In the following, we provide an overview of the characteristics and evaluations of measurement, forecasting, and nowcasting methods and show their relationships to trending.
+Conventionally, methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account. 
+These measures fail to consider whether the right direction of change is predicted or measured correctly.
+The information about whether a signal, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when making decisions based on the estimator's prediction results.
+In the following, we provide an overview over the characteristics and current evaluation schemes of the three application areas, measurement, forecasting, and nowcasting methods.
+We highlight why trending evaluation is highly relevant in the respective fields.
 
 Measurement aims to obtain accurate and reliable data about the current state of a system. 
-Often, a parameter or variable is measured regularly over a certain period.
+A parameter or variable is can be measured regularly over a certain period to evaluate the systems development.
+When introducing new measurement methods they are evaluated against current measurement techniques, called gold standard, by measuring the same quantity in different settings, for example, different individuals or environments.
 In addition to using a paired t-test to determine systematic differences between two measurement series (\cite{watson2010method}), various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
 Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}). 
-As already described above, one important aspect of testing the consistency of the new method with the gold standard is to adequately validate if the new method is measuring changes in the target variable in the right direction when the gold standard detects changes during the measurement process.
-Ideally, the signs of the changes between two consecutive points in time with a prescribed lag measured by the new and gold standard methods should always be the same. 
-In this case, we call the new method trending perfect in line with the gold Standard method and refer to this type of agreement as trending ability. 
-This type of consideration of the trending ability of a measurement method to the gold standard has also recently gained importance in the literature (\cite{saugel2015tracking, saugel2018error, hiraishi2021concordance}).
+Trending considerations, that is the consistency of the test methods development with the gold standard, have already gained importance in literature (\cite{Saugel2015,saugel2018error,hiraishi2021concordance}).
+In this paper, we build upon the existing methods and extend them with new functionality.
 
-%In (\cite{hiraishi2021concordance}), the authors proposed a new concordance rate for the four-quadrant diagram, which is based on the multivariate normal distribution and takes into account the correlation between the data of the individual subjects.
-
-The above-described trending idea is also of fundamental interest for evaluating and comparing forecasting and methods. 
-While measurement series are created by recording data using measuring instruments or direct observation, forecasting is about predicting the future based on historical data, its patterns, and exogenous factors.
-When forecasting at a specific time, the forecast value at the target time corresponds to a prediction of the change in the target variable between the forecast origin and the target time. 
-This predicted change results from the difference between the forecast value at the target time and the current actual value of the target variable at the time of prediction. 
-A forecast is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
-The current practice is that a forecasting method is usually evaluated in terms of its prediction accuracy, which is defined by a certain loss function and measures the difference between the prediction and the actual outcome of the target variable; an overview of the different measures can be found in \cite{hyndman2006another}.
-Popular measures of this type are scale-dependent measures such as the Root Mean Square Error (RMSE) or measures based on percentage errors such as the Root Mean Square Percentage Error (RMSPE). 
+The above-described trending idea is also of fundamental interest for evaluating and comparing forecasting methods. 
+While measurement series are created by recording data using measuring instruments or direct observation, forecasting is concerned with predicting the future based on historical data, its patterns, and exogenous factors.
+The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time is made.
+A forecast's trending is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
+In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the difference between the prediction and the actual outcome of the target variable; different measures are, for example, reviewed in \cite{hyndman2006another}.
+Popular measures of this type are scale-dependent measures such as the \ac{rmse} or measures based on percentage errors such as the root mean squared percentage error. 
 All of these typical measures are based on the absolute difference between the prediction and the true values, locally or globally. 
@@
 The evaluation of trending is also interesting for the assessment of nowcasts.
-In contrast to conventional forecasting methods, nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields such as economics and medicine (\cite{bok2018macroeconomic, wolffram2023collaborative}).
+In contrast to conventional forecasting methods, nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields such as economics and medicine (\cite{bok2018macroeconomic, Wolffram2023}).
 Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
@@
 Nowcasting methods use indicators that are related to the target variable and collected at a higher frequency.
-Thus, the nowcasts can produce early estimates for the target variable during the relevant period (\cite{castle2017forecasting}).
-For example, in forecasting GDP, we use economic indicators collected at a higher frequency from January to March 2024 to forecast GDP from the first quarter of 2024 before the official statistics are published.
+Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period (\cite{castle2017forecasting}). 
 In medicine, epidemic nowcasting assesses the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
 For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}).
-In this context, nowcasting is an estimation procedure in which the value of a target variable (for example, from today) is estimated based on current preliminary measurements, which are only finalized later.
+In this context, nowcasting is an estimation procedure in which the value of a target variable for a certain time, for example, today, is estimated based on current preliminary measurements, which are only finalized later.
 At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts.
 A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change.
-Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the RMSE (\cite{gunther2021nowcasting}) or probabilistic measures \parencite{Wolffram2023}.
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} (\cite{gunther2021nowcasting}) or probabilistic measures \parencite{Wolffram2023}.
 The aspect of trending ability is not considered in the literature to the best of our knowledge.
@@
 As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
-However, regardless of the application area, most conventional method comparisons and assessments consider local absolute differences differently without directly assessing whether the correct direction of change is being predicted or measured.
+However, regardless of the application area, most conventional method comparisons and assessments consider local absolute deviations without assessing whether the correct direction of change is being predicted or measured.
 In this paper, we want to complement the current evaluation measures with trending measurements and propose evaluation methods that assess the trending ability of a method.
@@
 \item We present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
-\item We introduce the conditional trending plot, a new graphical method for assessing local trends.
+\item We introduce the conditional trending plot, a new graphical method for assessing local trending behavior.
 \item We introduce bootstrap methods for calculating confidence intervals for the presented treding ratios.
@@
 \item We apply trending evaluation to three applications: Measurement, nowcasting, and forecasting data. 
-\item We provide a ready-to-use code for trending detection. The code and the source code to replicate the results of our study are available in our public repository at ... .
+\item We provide a ready-to-use code for trending evaluation. The code and the source code to replicate the results of our study are available in our public repository at ... .
 \end{itemize}
@@
 In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting. 
-We conclude our paper with the last section.
+We conclude our paper in Section~\ref{sec:conclusion}.

### 2024-03-15 sections/03-trending.tex
Update paper

@@
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
-Development always has respect to a time horizon, that is, the amount of time passed during development. 
+Development always has respect to a time horizon, that is, the amount of time passed. 
 In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the question at hand.
@@
 Before introducing measures to evaluate trending, we remark on calculating the predicted change and notation in Section~\ref{subsec:notation}.
-Section~\ref{subsec:trending-basics} introduces trending basics and visualizes trending in a synthetic example.
+Section~\ref{subsec:trending-four-quadrant-plot} introduces trending basics and visualizes trending in a synthetic example.
 In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
-The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise}.
-Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment and briefly introduces bootstrap methods for computing confidence intervals for the measures in the previous sections.
-Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:probabilistic}.
+The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise} and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
+Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment.
+Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:trending-probabilistic}.
 
@@
 (x_{t|t} - x_{t-l|t})^T_{t=l}, & \text{if } y_{t-l} \text{ is not known at time } t, \\
-(x_{t|t} - y_{t-l})^T_{t=1}, & \text{otherwise}.
+(x_{t|t} - y_{t-l})^T_{t=l}, & \text{otherwise}.
 \end{cases} 
@@
 (x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
-(x_{t|t} - y_{t-l})^T_{t=1} & \text{otherwise}.
+(x_{t|t} - y_{t-l})^T_{t=l} & \text{otherwise}.
 \end{cases} \label{eq:diffxl_nowcasting}
@@
 
-\subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-basics}
+\subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-four-quadrant-plot}
 
@@
 A simple yet insightful method is the four-quadrant plot~\parencite[see, e.g., ][]{Saugel2015,perrino1998intraoperative}. 
-In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffy_t, \diffx_t)$ for $t = 1, \dots, T$.
+In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffyt, \diffxt)$ for $t = 1, \dots, T$.
 Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
@@
 Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant graph for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
-Data generation is described in Appendix~\ref{subsec:app-trending-data-generation} and will continue into this section.
+Data generation is described in Appendix~\ref{subsec:app-trending-data-generation}.
 The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
@@
 \end{equation}
-The numerator counts the number of same-sign-changes, while the denominator is the number of all points.
+The numerator counts the number of same-sign-changes, while the denominator is the number of considered pairs $(\diffyt, \diffxt)$.
 Thus, $\acc$ is the ratio of concordant changes over all changes.
@@
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
-Such a $2 \times 2$ table of counts is called a contingency table.
-It is often used in other scientific areas for evaluation, for example, in dichotomous forecasting or as a confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
+Similar evaluations are used in other scientific areas, for example, in dichotomous forecasting or as a confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
 4]{James2021}[Ch.
@@
 \begin{align}
-    \accp (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0}} \label{eq:accp}\\
+    \accp (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0}},\ \text{and} \label{eq:accp}\\
     \accm (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0}}\label{eq:accm}
 \end{align}
-In the classification context, these measures are known as positive or negative predictive value and hit rate or detection failure ratio in forecasting.
-They give the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, respectively, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$, respectively.
+They give the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, respectively, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$.
 
@@
     \caption{Visualizations for data with a time-varying trending ratio.
-We defer information on the data generation process in \ref{fig:trending_basic_4q_sample} and \ref{fig:trending_basic_4q_sample_color} to the appendix (see \ref{subsec:app-trending-data-generation}).
+We defer information on the data generation process to the appendix (see \ref{subsec:app-trending-data-generation}).
 The trending ratio for the entire data set is $\mu = 0.7577$.
@@
 
-Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures}.
-Points within that area are neither plotted in the four-quadrant plot nor included in the calculation of the measures.
-In general, it is likely that the prediction models have a noise component and should thus be part of the exclusion area.
+Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} \parencite[see, e.g.,][]{Saugel2015,Critchley2010}.
+Points within that area are ommited in the four-quadrant plot and in the calculation of the measures.
+In particular the prediction models are likely to have a noise component and, thus, $\diffx$ should have an exclusion area.
 We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $E$ by
@@
 \end{align}
+The additional argument $(\diffyt, \diffxt) \notin E$ in the indicator ensures that only points not lying in the exclusion are evaluated when computing the measures.
 The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $E$, that is, for $\acceps$, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$.
 The estimators accept various shapes of the exclusion area.
-Figure~\ref{fig:trending_4q} visualizes examples of the exclusion area.
+Figure~\ref{fig:trending_4q} visualizes examples of exclusion area shapes.
 A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ 
@@
 In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
-An exclusion area along one axis, for example, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq 0 \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
+An exclusion area along one axis, for example, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
 This particularly suits signals where small amounts of noise are inevitable.
@@
 In most applications, the shape and size of the exclusion area can be chosen based on domain knowledge or expert opinions.
-The size estimation can also be based on a proportion of the total variance or the total range of the data; for example, the points with at least one coordinate in the smallest 10\% of absolute values are excluded.
+The size estimation can also be based on a proportion of the total variance or the total range of the data; for example, the size is twice the 10\% quantile of absolute values in each component and thus excludes the smallest 10 \% of points.
 A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
@@
 A still more fine and local analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ to assess the local trending ability of a prediction.
+$P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ denotes the probability of a correct trending given a predicted change of $x$. 
 A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
@@
 
-Assessing $P(\diffxrv \diffyrv > 0 | \diffx = x)$ graphically eases the simultaneous evaluation of various intervals.
+Assessing $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ graphically eases the simultaneous evaluation of various $x$.
 Furthermore, the graph facilitates the comparison of various methods in a single graph and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
-We refer to the plot as a conditional trending plot.
+We refer to the plot as a \textit{conditional trending plot}.
 
-\subsection{Probabilistic Evaluation}\label{subsec:probabilistic}
+\subsection{Probabilistic Evaluation}\label{subsec:trending-probabilistic}
 
@@
 Probabilistic predictions include a point estimate and information on the forecast uncertainty and quantiles simultaneously.
-For trending analysis, probabilistic predictions also contain a probability of a positive or negative change.
+For trending analysis, probabilistic predictions thus also contain a probability of a positive or negative change.
 In this section, we extend the trending assessment to probabilistic predictions.
@@
 Let $F_{t | t-l} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
-For trending assessment, we compare the probability of positive change with the occurrence of positive changes. 
-A positive change occurs if any value larger than the true value $y_{t-l}$ occurs. 
+For trending assessment, we compare the probability of positive change with the observed occurrence of positive changes. 
+For the prediction, a positive change is predicted for any value at $t$ larger than the true value $y_{t-l}$. 
 As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
@@
 The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
-Dichotomous forecasts predict a binary outcome, for example, a positive or negative change.
+Dichotomous forecasts predict a binary outcome, as here, a positive or negative change.
 Numerical measures for forecast quality are scoring rules.
 
-The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\parencite{Brier1950}.
-It is defined as
+The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\parencite{Brier1950} and is defined as
 \begin{equation*}
-    BS (\mathbf{p}, \mathbf{z}) = \frac{1}{T-l} \sum_{t=l}^{T} (p_t - z_t)^2.
+    BS (\mathbf{p}, \mathbf{z}) = \frac{1}{T-l+1} \sum_{t=l}^{T} (p_t - z_t)^2.
 \end{equation*}
@@
 
-These two conflicting aspects cannot be distinguished in the scoring rule.
+These two aspects cannot be distinguished in the scoring rule.
 Graphical methods are a standard tool for evaluating solely the calibration of probabilistic forecasts.
@@
 The reliability diagram plots the observed frequency of the positive outcome against the predicted probability.
+Thus, it visualizes the observed frequency of a positive change given a probability for positive change of, for example, 0.6, is predicted.
 Ideally, the predicted probability equals the observed frequency, and the reliability diagram is a 45-degree line.

### 2024-03-15 sections/04-application.tex
Update paper

@@
-In this section, we apply the trending assessment to three different applications.
+In this section, we evalute the trending in three different applications.
 In the first application in Section~\ref{sec:application-covid}, we consider the nowcasting of the seven-day hospitalization rate for COVID-19 in Germany.
@@
 The seven-day hospitalization rate is the average number of those cases per $100,000$ inhabitants on the given date and six days before.
-Thus, the final seven-day hospitalization rate can only be reported with a significant delay of up to more than 70 days, as the hospitalization of infected inhabitants can occur much later than the first positive test.
+Thus, the final seven-day hospitalization rate can only be reported with a significant delay of more than 70 days, as the hospitalization of infected inhabitants can occur much later than the first positive test.
 Nevertheless, as the seven-day hospitalization rate was considered a major indicator of the pandemic's development, many organizations and institutions started to issue nowcasts, including research teams and newspapers.
@@
 In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
-In contrast to~\textcite{Wolffram2023}, we use the newest data available on the true values, not from August 8, 2022.
+For the true values, we use the data from February 8, 2024.
 The nowcasts are resolved with regard to all inhabitants and specific geographical and age breakdowns.
@@
 The time comprises the fourth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany, lasting until May 28, 2022~\parencite{Tolksdorf2022}.
+
 Table~\ref{tab:app-covid-rmse} summarizes the point evaluation measures for the issued mean of the different models.
@@
 Before stating the results, we provide background information on the marginal distributions of the true values and nowcasts for the different horizons.
-Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} provides information on marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
+Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} presents marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
 Overall, the variability and general level of differences grow with the horizon.
@@
 The 10\%-quantile is used for the exclusion areas in the trending assessment.
-The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the differences.
+The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the absolute differences.
 Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.
@@
 Figure~\ref{fig:app-covid-cond-prob-trending-ratio-7} shows the conditional trending plots and the trending ratio over the exclusion area for the horizon seven days; the respective plots for the horizons one day and 14 days are shown in Figure~\ref{fig:app-covid-cond-prob-trending-ratio-1-14}.
-Here, only the best models in point evaluation measures, ILM, RKI, RIVM, and ENS-MED, are shown to keep the plots readable.
+Here, only the best models in point evaluation measures, ILM, RKI, RIVM, and ENS-MED, are shown to keep the plots easily readable.
 If RKI or ILM issues a fall in the hospitalization rate, the probability of a fall is higher than if RIVM or ENS-MED issues a fall.
@@
     \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models with and without exclusion areas for the horizon seven days.
-The exclusion areas are rectangles centered on the zero points with a width and height of 10\%-quantile of the absolute values of nowcast and true values.
+The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
 }
@@
 For all horizons, the influence of the exclusion area on the 10\%-quantile level is negligible.
-For example, the trending ratio changes at most by 0.03 for the EPI model with $\accm[14]$.
+For example, the trending ratio changes at most by 0.03 for the EPI model with $\accml[14]$.
 The exclusion areas are thus not crucial for the trending assessment in the case of the nowcasts of the seven-day hospitalization rate.
@@
 
+Figure~\ref{fig:app-covid-trending-ratio-7} shows that the trending ratio increases with larger exclusion areas. 
+This indicates that if the model predicts a large change, the direction is indeed better than when a small change is predicted.
+
 A more extensive training size would be beneficial for assessing the models' performance.
@@
 
-In a second example, we consider the forecasting of the hourly number of arrivals in a large emergency department.
-Forecasting the number of arrivals in an emergency department is crucial for planning staff and resources.
-Several models are used for predicting hourly outcomes in the study by \textcite{Rostami-Tabar2023}.
+In a second example, we consider forecasting the hourly number of arrivals in a large emergency department.
+Good forecasts are crucial for planning staff and resources.
+Several models are used for predicting hourly outcomes in a study by \textcite{Rostami-Tabar2023}.
 Every 12 hours, the models issue hourly forecasts for the next 48 hours.
@@
 
-\textcite{Rostami-Tabar2023} publish probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
+\textcite{Rostami-Tabar2023} publish means and probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
 The models are trained on data from April 1, 2014, to February 28, 2018, and evaluated on data from March 1, 2018, to February 28, 2019.
@@
 The confidence interval width is at most 0.02 for the trending ratios and at most 0.03 for the positive and negative trending ratios.
-Other models exhibit a higher positive trending ratio than the negative trending ratio.
 The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for three and seven days horizon, while Poisson-2 and NBI-2 have the highest negative trending ratio.
@@
 Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
-The \acf{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days are shown in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
-The \acp{bs} are smallest and close for NBI-2 and Poisson-2 for both horizons, while the Brier scores for the other models are larger and differ more.
+The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days are shown in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
+The \acp{bs} are smallest for NBI-2 and Poisson-2 for both horizons, while the \acp{bs} for the other models are larger and differ more.
 The qreg-1 model has the highest \ac{bs} for both horizons.
@@
 For critical patients, changes in blood pressure can be crucial for the treatment.
-Thus, trending assessment can be performed in addition to standard accuracy analysis~\parencite[see, for example, ]{Mostafa2020}.
+Thus, trending assessment can be performed in addition to standard accuracy analysis~\parencite[see, for example, ][]{Mostafa2020}.
 Thus, we assess the trending ability of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
@@
 The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} simultaneously for two consecutive times with the specified horizons.
+Thus, we use the \ac{nbp} measurements as test method and the \ac{abp} measurements as gold standard.
 For the systolic measurements, 290, 332, and 442 points are available for the horizons of one, five, and 15 minutes; for the mean measurements, 406, 430, and 542.

### 2024-03-15 sections/05-conclusion.tex
Update paper

@@
-\begin{itemize}
-  \item Summary of work
-  \item highlight important general remarks of section trending / application
-  \item points for further research/other methods: imbalances in the number of positive and negative changes; sequential dependence of the differences
-\end{itemize}
+In this paper, we examine various methods to analyze trending, the statistical consistency of observed and predicted change direction.
+While the computation of predicted change varies between the application areas measurement (see Section~\ref{subsec:notation}), nowcasting and forecasting, the assessment of trending using the computed predicted and observed change is similar.
+The evaluation of trending can accompany other evaluation techniques such as measures of deviation or probabilistic scoring rules.
 
-\subsection{Modeling remarks}
+Four-quadrant plots facilitate the visual inspection of the trending ability for a signal (see Section~\ref{subsec:trending-four-quadrant-plot}).
+The trending ratio, that is, the ratio of change directions predicted correctly over all changes, numerically evaluates trending.
+Visually, it is the ratio of concordant points over all points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
+The positive and negative trending ratio analyze the trending ratio given the predicted change is positive or negative, respectively.
+Thus, they quantify the credibility of the respective predictions. 
+The applications of Section~\ref{sec:application_measurement} show, that models, in general, indeed have different positive and negative trending abilities and that they add valuable information to the trending ratio.
+In the applications, the bootstrap condfidence intervals of Section~\ref{subsec:trending-measures} are used to quantify the estimation uncertainty of the trending measures.
+The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
+For reasonably well trending models, 100 samples are thus sufficient to differentiate from random guessing or models with high trending differences, 1000 samples enable a more precise differentiation.
 
-Using these estimators could encourage predictors to exploit imbalances in the number of positive and negative changes. 
-A large difference in $\sum_{t \in \mathcal{T}} \ind{\diffyt > 0}$ and $\sum_{t \in \mathcal{T}} \ind{\diffyt < 0}$ is unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data.
-However, if the number of positive and negative $\diffy$ differs widely, the use of unbalanced-data-aware measures should be considered.
+A conditional trending plot visualizes the probability of correct trending over the predicted change of the signal (see Section\ref{subsec:trending-cond-prob}).
+It is based on a multivariate \acf{kde} of predicted and observed change.
+In the application, it gives reasonable insights into local effects of the trending ability.
+Section~\ref{subsec:trending-probabilistic} adapts measures of probabilistic forecast evaluation to the trending evaluation of probabilistic forecasts and nowcasts.
+The \acf{bs} as numerical assessment of probabilistic trending is introduced and reliability diagrams are used to visualize the local trending ability of probabilistic forecasts.
+
+The methods of trending assessment are applied to COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements in Section~\ref{sec:application}.
+While trending evaluation is not the only aspect of evaluation, it is a valuable addition to the evaluation of nowcasts, forecasts, and measurements.
+For models with very different accuracies, these are usually replicated in trending evaluation, but the trending evaluation can differentiate between models with similar accuracies.
+As in the application in Section~\ref{sec:application-covid}, models with average point forecast evaluation measures can have the most meaningful positive trending ability.
+
+Throughout this paper, we did not expand on two modeling aspects, which we leave for further research.
+In the estimation, we did not consider sequential correlation.
+The computation of differences is a standard procedure in time series analysis to remove sequential dependence, but, in general, the trending measures could be adapted to consider sequential correlation.
+Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation by using time series bootstrap methods~\parencite{Hardle2003,Kreiss2012}.
+
+The estimators of Section~\ref{subsec:trending-measures} could encourage predictors to exploit imbalances in the number of observed positive and negative changes.
+A large difference in the number of observed positive and negative changes is unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data.
+In addition, we do not advise assessing signals solely on their trending ability and, thus, provide an incentive for the exploitation of estimator characteristics.
+However, if the number of positive and negative observed changes differs widely, the use of unbalanced-data-aware measures should be considered.
 There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
-Cohen's $\kappa$ is usually used to measure inter-rater agreement and considers the ratio of occurred agreement and the probability of agreement by chance.
-Cohen's $\kappa$ reduces to rescaling the trending ratio for a $2\times2$-table and balanced outcomes.

### 2024-03-15 sections/50-appendix-data-generation.tex
Update paper

@@
 \begin{equation*}
-    \diffyt = \diffxt \cdot n_t * b_t,
+    \diffyt = \diffxt \cdot n_t \cdot b_t,
 \end{equation*}
@@
 \begin{equation*}
-    k_t = 0.75 + \sin(t / 365.25 \cdot 2 \pi) / 4,
+    k_t = 0.75 + \sin(t / 365.25 \cdot 2 \pi) / 4.
 \end{equation*}
-where $T^\star$ is $T$ divided by the number of oscillations; in this case, 4.
 For the asymmetric trending ratio, $k$ is a function of $\diffxt$,
@@
 \begin{equation*}
-    P(\diffyrv \diffxrv > 0 | \diffxrv = x) = \Phi \left( \frac{3}{4 \sqrt{7}} x \right),
+    P(\diffyrv \diffxrv > 0 | \diffxrv = x) = \Phi \left( \frac{3}{2 \sqrt{7}} x \right),
 \end{equation*}
@@
 
-\subsection{Visualization of different bandwidth selectors in multivariate \ac{kde}}\label{subsec:appendix-kde}
-
 \begin{figure}
@@
 
+\subsection{Visualization of different bandwidth selectors in multivariate \ac{kde}}\label{subsec:appendix-kde}
+
+We examine the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\parencite{Seabold2010} in Figure~\ref{fig:trending-cond-prob-bw}.
+While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.
+The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
+The second method, cross-validation least squares, requires long computation times while yielding small or no bandwidth results, even for two relatively small datasets.
+The rule-of-thumb and cross-validation maximum likelihood methods yield reasonable results at moderate computation times. 
+Further examples, including comparisons between methods regarding their trending ability, are available in Section~\ref{sec:application}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}{.48\textwidth}
+        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_asym_butterfly}
+        \caption{First dataset with asymmetric dependence.}
+    \end{subfigure}
+    \begin{subfigure}{.48\textwidth}
+        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_normal}
+        \caption{Second dataset. }
+    \end{subfigure}
+    \caption{Resulting conditional trending plot for different bandwidth selection processes.
+Cross-validation least squares takes a considerably larger computation time.
+It does not converge for the first data set and yields a bandwidth too small for the second data set.
+The rule of thumb is the fastest method but tends to oversmooth.
+The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time.
+$\varepsilon$ specifies an exclusion area $E = \{(x, y) \in \R^2: (-\varepsilon \leq x \leq \varepsilon)\}$ in $\diffx$-direction.}\label{fig:trending-cond-prob-bw}
+\end{figure}
+
 \subsection{Probabilistic trending evaluation}
@@
 \begin{equation*}
-    \bar{p}_t = \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
+    \pc_t = \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
 \end{equation*}
@@
 \begin{equation*}
-\bar{p}_t = \begin{cases}
+\pc_t = \begin{cases}
     \max \{\frac{1}{\pmax} - \frac{q_{p^\star} - y_t}{\pmax (q_{p2/\pmax} - q_{1/\pmax})}, 0\} &, \text{if } y_t < q_{1/p}, \\
@@
 \end{equation*}
-The probability of positive change is $p_t = 1 - \bar{p}_t$.
+The probability of positive change is $p_t = 1 - \pc_t$.
 
@@
 \begin{align}
-    \bar{p}_t
+    \pc_t
         &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: (x_1, x_2) \nonumber\\

### 2024-03-18 paper.tex
Update on Overleaf.

@@
 %\usepackage[hidelinks]{hyperref}
-\usepackage[style=authoryear,backend=bibtex]{biblatex}
+\usepackage[style=authoryear,backend=biber]{biblatex}
 \addbibresource{library.bib}
@@
 \begin{abstract}
-Trending refers to the ability of a signal to predict the direction of change for the true variable accurately. 
-An appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers. 
+An intuitive and easily accessible information on a quantity measured over time is whether it increases or decreases.
+Trending refers to the ability to predict the direction of change for a true variable accurately. 
+Thus, an appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers. 
 While the current practice of evaluating signals is based on averaging over time-step-separated measures, we review and introduce trending measures that consider whether the direction of development fits the quantity of interest.

### 2024-03-18 paper.tex
Update on Overleaf.

@@
 Thus, an appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers. 
-While the current practice of evaluating signals is based on averaging over time-step-separated measures, we review and introduce trending measures that consider whether the direction of development fits the quantity of interest.
-We formalize the concept of trending for all application areas and quantify the trending ability for signals. 
-We review and present new visualization techniques to assess the local trending behaviors and quantify estimation uncertainty by bootstrapping. 
+While the current practice of evaluating signals is based on averaging over local deviation measures, we review and introduce trending measures that consider whether the direction of development fits the quantity of interest.
+We formalize the concept of trending for all application areas and present quantitative measures. 
+Using bootstrap confidence intervals and exclusion areas, we provide extensions for noisy data and estimation uncertainty.
+We review and present new visualization techniques to assess the local trending behaviors and enlarge the scope to probabilistic nowcasts and forecasts. 
 We apply the methods to three large data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.

### 2024-03-18 sections/01-introduction.tex
Update on Overleaf.

@@
-The evaluation of measurement, prediction, and forecasting methods become increasingly important and sophisticated as technologies and data availability enable their application in more and more areas.
-Conventionally, methods are evaluated using distance measures that take aggregated local absolute differences between predictions (or measurements) and target values into account. 
+The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
+Conventionally, methods are evaluated using distance measures of local differences between predictions (or measurements) and target values into account. 
 These measures fail to consider whether the right direction of change is predicted or measured correctly.
-The information about whether a signal, be it a measurement device, forecast, or nowcast, can accurately predict an increase or decrease is crucial when making decisions based on the estimator's prediction results.
-In the following, we provide an overview over the characteristics and current evaluation schemes of the three application areas, measurement, forecasting, and nowcasting methods.
+The information about whether an increase or decrease is predicted correctly is crucial when making decisions based on the estimator's prediction results. 
+In the following, we provide an overview of the characteristics and current evaluation schemes of the three application areas: measurement, forecasting, and nowcasting.
 We highlight why trending evaluation is highly relevant in the respective fields.
@@
 Measurement aims to obtain accurate and reliable data about the current state of a system. 
-A parameter or variable is can be measured regularly over a certain period to evaluate the systems development.
-When introducing new measurement methods they are evaluated against current measurement techniques, called gold standard, by measuring the same quantity in different settings, for example, different individuals or environments.
+A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
+When introducing new measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
 In addition to using a paired t-test to determine systematic differences between two measurement series (\cite{watson2010method}), various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
 Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}). 
-Trending considerations, that is the consistency of the test methods development with the gold standard, have already gained importance in literature (\cite{Saugel2015,saugel2018error,hiraishi2021concordance}).
+Trending considerations, that is, the consistency of the test methods development with the gold standard, have already gained importance in literature (\cite{Saugel2015,saugel2018error,hiraishi2021concordance}).
 In this paper, we build upon the existing methods and extend them with new functionality.

### 2024-03-18 sections/01-introduction.tex
Update on Overleaf.

@@
 When introducing new measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
-In addition to using a paired t-test to determine systematic differences between two measurement series (\cite{watson2010method}), various indices such as the interclass correlation coefficient or Lin's concordance correlation have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}).
+Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}) in addition to paired t-tests to determine systematic differences between two measurement series (\cite{watson2010method}).
 Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}). 
-Trending considerations, that is, the consistency of the test methods development with the gold standard, have already gained importance in literature (\cite{Saugel2015,saugel2018error,hiraishi2021concordance}).
+Trending considerations, that is, the consistency of the test methods development with the gold standard, is a field of active research in the last years (\cite{Saugel2015,saugel2018error,hiraishi2021concordance}).
 In this paper, we build upon the existing methods and extend them with new functionality.
@@
 The above-described trending idea is also of fundamental interest for evaluating and comparing forecasting methods. 
-While measurement series are created by recording data using measuring instruments or direct observation, forecasting is concerned with predicting the future based on historical data, its patterns, and exogenous factors.
-The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time is made.
+While measurement series are created by recording data using measuring instruments or direct observation, forecasting predicts the future based on historical data, patterns, and exogenous factors. 
+The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
 A forecast's trending is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
-In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the difference between the prediction and the actual outcome of the target variable; different measures are, for example, reviewed in \cite{hyndman2006another}.
+In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable; different measures are, for example, reviewed in \cite{hyndman2006another}.
 Popular measures of this type are scale-dependent measures such as the \ac{rmse} or measures based on percentage errors such as the root mean squared percentage error. 
-All of these typical measures are based on the absolute difference between the prediction and the true values, locally or globally. 
+These typical measures are based on the absolute difference between the prediction and the true values, locally or globally. 
 They are not capable of assessing the trending ability discussed above.
 
-The evaluation of trending is also interesting for the assessment of nowcasts.
-In contrast to conventional forecasting methods, nowcasting methods focus on predictions for the present, the immediate future, and the recent past (\cite{banbura2013now}) and are now widely used in fields such as economics and medicine (\cite{bok2018macroeconomic, Wolffram2023}).
-Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (\cite{browning1989nowcasting,schmid2019nowcasting}).
-In economics, nowcasting is used, for example, to predict statistics such as the current economic situation (for example, GDP), which are collected with low frequency and are available with a considerable time delay (\cite{banbura2013now}).
-Nowcasting methods use indicators that are related to the target variable and collected at a higher frequency.
-Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period (\cite{castle2017forecasting}). 
-In medicine, epidemic nowcasting assesses the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors (\cite{wu2021nowcasting}).
-For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported before this data has been fully recorded at a central location (\cite{gunther2021nowcasting}).
-In this context, nowcasting is an estimation procedure in which the value of a target variable for a certain time, for example, today, is estimated based on current preliminary measurements, which are only finalized later.
+In contrast to conventional forecasting methods, nowcasting methods focus on predictions for the present, the immediate future, and the recent past \parencite{banbura2013now} and are now widely used in fields such as economics and medicine \parencite{bok2018macroeconomic, Wolffram2023}.
+Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours \parencite{browning1989nowcasting,schmid2019nowcasting}.
+In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \parencite{banbura2013now}.
+Nowcasting methods use indicators related to the target variable and collected at a higher frequency.
+Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period \parencite{castle2017forecasting}. 
+In medicine, epidemic nowcasting assesses the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors \parencite{wu2021nowcasting}.
+For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported (\cite{gunther2021nowcasting}). 
+In this context, nowcasting is an estimation procedure in which the value of a target variable for a specific time is estimated based on current preliminary measurements, which are only finalized later.
 At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts.

### 2024-03-18 sections/01-introduction.tex
Update on Overleaf.

@@
 A forecast's trending is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
-In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable; different measures are, for example, reviewed in \cite{hyndman2006another}.
+In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable; different measures are, for example, reviewed in \parencite{hyndman2006another}.
 Popular measures of this type are scale-dependent measures such as the \ac{rmse} or measures based on percentage errors such as the root mean squared percentage error. 

### 2024-03-18 sections/01-introduction.tex
Update on Overleaf.

--- 2024-03-18 sections/01-introduction.tex
@@
 When introducing new measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
-Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods (\cite{lawrence1989concordance,koo2016guideline}) in addition to paired t-tests to determine systematic differences between two measurement series (\cite{watson2010method}).
-Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value (\cite{bland1986statistical}). 
-Trending considerations, that is, the consistency of the test methods development with the gold standard, is a field of active research in the last years (\cite{Saugel2015,saugel2018error,hiraishi2021concordance}).
+Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods \parencite{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series \parencite{watson2010method}.
+Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value \parencite{bland1986statistical}. 
+Trending considerations, that is, the consistency of the test methods development with the gold standard, is a field of active research in the last years \parencite{Saugel2015,saugel2018error,hiraishi2021concordance}.
 In this paper, we build upon the existing methods and extend them with new functionality.
@@
 A forecast's trending is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
-In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable; different measures are, for example, reviewed in \cite{hyndman2006another}.
+In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable; different measures are, for example, reviewed in \textcite{hyndman2006another}.
 Popular measures of this type are scale-dependent measures such as the \ac{rmse} or measures based on percentage errors such as the root mean squared percentage error. 
@@
 In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \parencite{banbura2013now}.
-Nowcasting methods use indicators related to the target variable and collected at a higher frequency.
+Nowcasting methods use high-frequency indicators related to the target variable.
 Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period \parencite{castle2017forecasting}. 
 In medicine, epidemic nowcasting assesses the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors \parencite{wu2021nowcasting}.
-For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported (\cite{gunther2021nowcasting}). 
-In this context, nowcasting is an estimation procedure in which the value of a target variable for a specific time is estimated based on current preliminary measurements, which are only finalized later.
+For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported \parencite{gunther2021nowcasting}. 
+In the medical context, nowcasting is an estimation procedure in which the value of a target variable for a specific time is estimated based on current preliminary measurements, which are only finalized later.
 At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts.
 A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change.
-Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} (\cite{gunther2021nowcasting}) or probabilistic measures \parencite{Wolffram2023}.
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \parencite{gunther2021nowcasting} or probabilistic measures \parencite{Wolffram2023}.
 The aspect of trending ability is not considered in the literature to the best of our knowledge. 
-However, appropriate measures for assessing trending ability should be constructed because a nowcast with strong trending ability provides crucial information about the development of the target variable, for example, the number of cases of an epidemic, and can be the foundation of decisions on the introduction or cancellation of policy measures.
+However, trending evaluation reveals a method's capability of predicting the development of the target variable, for example, the number of cases of an epidemic, that can be the foundation of decisions on the introduction or cancellation of policy measures.
+\todo{Satz ist noch nicht gut}
 

### 2024-03-18 sections/01-introduction.tex
Update on Overleaf.

@@
 In the medical context, nowcasting is an estimation procedure in which the value of a target variable for a specific time is estimated based on current preliminary measurements, which are only finalized later.
-At the time of nowcasting, estimates of the target variable can be made for the current time and for a past time; the predicted change in the target variable is then the difference between the two nowcasts.
-A nowcast is consistent with the evolution of the target variable if the difference between the actual values at the current time and the actual values at a past time has the same sign as the predicted change.
 Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \parencite{gunther2021nowcasting} or probabilistic measures \parencite{Wolffram2023}.
 The aspect of trending ability is not considered in the literature to the best of our knowledge. 
-However, trending evaluation reveals a method's capability of predicting the development of the target variable, for example, the number of cases of an epidemic, that can be the foundation of decisions on the introduction or cancellation of policy measures.
+However, trending evaluation reveals a method's capability of predicting the development of the target variable, for example, the number of cases of an epidemic, that can be the foundation of decisions on introducing or canceling policy measures.
 \todo{Satz ist noch nicht gut}
@@
 However, regardless of the application area, most conventional method comparisons and assessments consider local absolute deviations without assessing whether the correct direction of change is being predicted or measured.
-In this paper, we want to complement the current evaluation measures with trending measurements and propose evaluation methods that assess the trending ability of a method.
+In this paper, we develop trending evaluation methods that complement current evaluation measures with trending measurements.
 
-The main contributions of this paper are summarized below
+The main contributions of this paper are manifold.
 \begin{itemize}
-\item We formalize trending and show how the predicted changes can be computed to assess trending for practical relevance. 
-\item We present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
-\item We introduce the conditional trending plot, a new graphical method for assessing local trending behavior.
-\item We introduce bootstrap methods for calculating confidence intervals for the presented treding ratios.
+\item We formalize trending and present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
+\item We introduce the conditional trending plot, a new graphical method for assessing local trending behavior, and review bootstrap methods for calculating confidence intervals.
 \item We extend the concept of trending to probabilistic predictions.
 \item We apply trending evaluation to three applications: Measurement, nowcasting, and forecasting data. 
-\item We provide a ready-to-use code for trending evaluation. The code and the source code to replicate the results of our study are available in our public repository at ... .
+\item We provide a ready-to-use code for trending evaluation. The code and the source code to replicate the results of our study are available at ... .
 \end{itemize}
@@
 The remainder of this paper is structured as follows. 
-In Section \ref{sec:trending}, we formalize the concept of trending independently of the application domain and investigate quantification approaches for noise-free and noisy data.
-Noisy data refers to data where noise or unsystematic effects are present in the true values or predictions.
+Section \ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts. 
 In particular, we introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries. 
-Bootstrap methods for calculating confidence intervals for trending measures are also presented. 
 In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting. 
-We conclude our paper in Section~\ref{sec:conclusion}.
+We conclude the paper in Section~\ref{sec:conclusion}.

### 2024-03-18 sections/03-trending.tex
Update on Overleaf.

@@
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
-Development always has respect to a time horizon, that is, the amount of time passed. 
+Development always has respect to a time horizon, that is, the amount of time that has passed. 
 In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the question at hand.
@@
 In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
-The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise} and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
+The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise}, and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
 Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment.
@@
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values for nowcasting or forecasting, or gold standard measurements up to time $T$.
-The sequence of observed changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is
+The sequence of observed changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is,
 \begin{equation}\label{eq:diffy}
@@
 
-The definition of the predicted change depends on the context.
+The definition of the \textit{predicted} change depends on the context.
 Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
@@
 When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet. 
-Thus, the predicted can be computed with the knowledge of time $t$, and the trending assessment can be used in practice at time $t$.
+Thus, the predicted change can be computed with the knowledge of time $t$, and the trending assessment can be used in practice at time $t$.
 
@@
 The indices are chosen to be consistent with $\diffyl$. 
-If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification can be made as in Equation~\eqref{eq:diffxl_nowcasting}.
+If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification as in Equation~\eqref{eq:diffxl_nowcasting} can be made.
 
@@
 Consequently, when evaluating the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
-A simple yet insightful method is the four-quadrant plot~\parencite[see, e.g., ][]{Saugel2015,perrino1998intraoperative}. 
-In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffyt, \diffxt)$ for $t = 1, \dots, T$.
+A simple yet insightful method is the four-quadrant plot~\parencite[see, e.g., ][]{perrino1998intraoperative,Saugel2015}. 
+In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
 Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
@@
 It can be extended by including information on the date in the point color.
-In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$.
-The horizon is always one time step.
+In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, $(\diffyt[l], \diffxt)$
 However, four-quadrant plots become crowded for larger datasets, and sequential information on the differences is complex to assess thoroughly.
+
 There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.

### 2024-03-18 sections/03-trending.tex
Update on Overleaf.

@@
 It can be extended by including information on the date in the point color.
-In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, $(\diffyt[l], \diffxt)$
+In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, $(\diffyt[l], \diffxt[l])$ is blue and turns green until $(\diffyt[T], \diffxt[T])$.
 However, four-quadrant plots become crowded for larger datasets, and sequential information on the differences is complex to assess thoroughly.
@@
 With that, we mainly estimate the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
-Note that $z_1 z_2 > 0$ if and only if $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$).
-The standard estimator for $P(\diffxrv \diffyrv > 0)$ is
+Since $z_1 z_2 > 0$ imposes the same condition as $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$), the standard estimator for $P(\diffxrv \diffyrv > 0)$ is
 \begin{equation}
@@
 3]{Jolliffe2012}.
-There, a wide range of other methods are usually used to analyze further characteristics of contingency tables.
+There, a wide range of other methods are usually used to analyze further characteristics of such count tables.
 Two simple measures that focus on either a positive or negative predicted change are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
@@
     \accp (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0}},\ \text{and} \label{eq:accp}\\
-    \accm (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0}}\label{eq:accm}
+    \accm (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0}}.\label{eq:accm}
 \end{align}
@@
 \end{equation*}
-Backward-looking windows provide an estimate at a time $t$ considering the $w$ time steps before time $t$.
+Backward-looking windows estimate the trending ratio at a time $t$ considering the $w$ time steps before time $t$.
 The window length $w$ controls the smoothing of the estimate; a larger $w$ gives smoother results, while a small $w$ focuses on local variations.
@@
 The trending ability of the signal has a strong sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
-This seasonal behavior cannot be distinguished in the two four-quadrant plots or the trending ratio.
+This seasonal behavior cannot be distinguished in the two four-quadrant plots or the overall trending ratio.
 
@@
 In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
-We refer to time steps for either signal or true values, or both are unavailable as missing values.
+We refer to time steps for which either signal or true values, or both, are unavailable as missing values.
 Data pairs with missing values can be excluded from the set $\mathcal{T}$ to calculate the measures. 
-However, systematical missing values could lead to a biased estimate of the trending ratio, for example, if signal data is omitted in times of high change and thus uncertainty.
-In this case, measures should be interpreted cautiously, and inspection of missing data should accompany applications with missing data, for example, through a visual assessment.
+However, systematical missing values could lead to a biased estimate of the trending ratio and missing data should be inspected for any underlying patterns.
 In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}, only a few missing values occur, and we provide information on the missing data by inspecting the missing data as lists.
@@
 Points within that area are ommited in the four-quadrant plot and in the calculation of the measures.
-In particular the prediction models are likely to have a noise component and, thus, $\diffx$ should have an exclusion area.
+In particular, the prediction models are likely to have a noise component and, thus, $\diffx$ should be subject to an exclusion area.
 We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $E$ by
 \begin{align}
-    \acceps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin E}}\label{eq:acceps}\\
-    \accpeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}} \label{eq:accpeps}\\
-    \accmeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}\label{eq:accmeps}
+    \acceps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin E}},\label{eq:acceps}\\
+    \accpeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}, \ \text{and} \label{eq:accpeps}\\
+    \accmeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}.
+\label{eq:accmeps}
 \end{align}
 The additional argument $(\diffyt, \diffxt) \notin E$ in the indicator ensures that only points not lying in the exclusion are evaluated when computing the measures.
-The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $E$, that is, for $\acceps$, $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$.
+The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $E$, that is, $\acceps$ is an estimator for $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$.
+
 The estimators accept various shapes of the exclusion area.
@@
  ($\varepsilon_x, \varepsilon_y > 0$), leaves out points that are small in both components and, therefore, likely to be driven by noise.
-Points where at least the true value or signal is unlikely to be zero are not excluded.
+Points where the true value or signal is unlikely to be noise-driven, are included.
 In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
@@
 A cross-shaped exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
+A point is only included if no component is likely to be driven by noise.
 For these two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
@@
 In most applications, the shape and size of the exclusion area can be chosen based on domain knowledge or expert opinions.
-The size estimation can also be based on a proportion of the total variance or the total range of the data; for example, the size is twice the 10\% quantile of absolute values in each component and thus excludes the smallest 10 \% of points.
+The size estimation can also be based on a proportion of the total variance or the total range of the data; for example, the size is twice the 10\% quantile of absolute values in each component and thus excludes approximately the smallest 10 \% of points.
 A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.

### 2024-03-18 sections/03-trending.tex
Update on Overleaf.

@@
 In most applications, the shape and size of the exclusion area can be chosen based on domain knowledge or expert opinions.
-The size estimation can also be based on a proportion of the total variance or the total range of the data; for example, the size is twice the 10\% quantile of absolute values in each component and thus excludes approximately the smallest 10 \% of points.
+The size estimation can also be based on a proportion of the total variance or the total range of the data; for example, the 10\% smallest absolute values in each component determine the exclusion area size.
 A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
-For examples on those plots, see Section~\ref{sec:application-covid}.
+For examples on such plots, see Section~\ref{sec:application-covid}.
 
 Confidence intervals can account for the estimation uncertainty of the measures above.
-Bootstrap confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
+Bootstrap confidence intervals are based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
 New samples are drawn with replacement from the dataset and are not based on parametric assumptions as classical confidence intervals are.
@@
 
-Let $F_{t | t-l} (x)$ denote the predicted \ac{cdf} of forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
+Let $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
 For trending assessment, we compare the probability of positive change with the observed occurrence of positive changes. 
@@
 
-These two aspects cannot be distinguished in the scoring rule.
 Graphical methods are a standard tool for evaluating solely the calibration of probabilistic forecasts.
@@
 The reliability diagram plots the observed frequency of the positive outcome against the predicted probability.
-Thus, it visualizes the observed frequency of a positive change given a probability for positive change of, for example, 0.6, is predicted.
+Thus, it visualizes the observed frequency of a positive change given a predicted probability for positive change of, for example, approximately 0.6.
 Ideally, the predicted probability equals the observed frequency, and the reliability diagram is a 45-degree line.

### 2024-03-18 sections/05-conclusion.tex
Update on Overleaf.

@@
-In this paper, we examine various methods to analyze trending, the statistical consistency of observed and predicted change direction.
-While the computation of predicted change varies between the application areas measurement (see Section~\ref{subsec:notation}), nowcasting and forecasting, the assessment of trending using the computed predicted and observed change is similar.
-The evaluation of trending can accompany other evaluation techniques such as measures of deviation or probabilistic scoring rules.
+In this paper, we examine various methods to analyze trending, that is, the statistical consistency of observed and predicted change direction.
+While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting (see Section~\ref{subsec:notation}), the trending assessment using the computed predicted and observed change is similar.
+The evaluation of trending can accompany other evaluation techniques, such as measures of deviation or probabilistic scoring rules.
 
 Four-quadrant plots facilitate the visual inspection of the trending ability for a signal (see Section~\ref{subsec:trending-four-quadrant-plot}).
-The trending ratio, that is, the ratio of change directions predicted correctly over all changes, numerically evaluates trending.
+The trending ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates trending.
 Visually, it is the ratio of concordant points over all points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
@@
 Thus, they quantify the credibility of the respective predictions. 
-The applications of Section~\ref{sec:application_measurement} show, that models, in general, indeed have different positive and negative trending abilities and that they add valuable information to the trending ratio.
-In the applications, the bootstrap condfidence intervals of Section~\ref{subsec:trending-measures} are used to quantify the estimation uncertainty of the trending measures.
+The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative trending abilities and that they add valuable information to the trending ratio.
+In the applications, the bootstrap confidence intervals of Section~\ref{subsec:trending-measures} are used to quantify the estimation uncertainty of the trending measures.
 The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
-For reasonably well trending models, 100 samples are thus sufficient to differentiate from random guessing or models with high trending differences, 1000 samples enable a more precise differentiation.
+For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or assess models with high trending differences, and 1000 samples enable a more precise differentiation.
 
@@
 It is based on a multivariate \acf{kde} of predicted and observed change.
-In the application, it gives reasonable insights into local effects of the trending ability.
-Section~\ref{subsec:trending-probabilistic} adapts measures of probabilistic forecast evaluation to the trending evaluation of probabilistic forecasts and nowcasts.
-The \acf{bs} as numerical assessment of probabilistic trending is introduced and reliability diagrams are used to visualize the local trending ability of probabilistic forecasts.
+In the application, the conditional trending plot gives reasonable insights into the local effects of the trending ability.
+Section~\ref{subsec:trending-probabilistic} adapts measures of probabilistic forecast evaluation to the trending assessment of probabilistic forecasts and nowcasts.
+The \acf{bs} as numerical assessment of probabilistic trending is introduced, and reliability diagrams are used to visualize the local trending ability of probabilistic forecasts.
 
 The methods of trending assessment are applied to COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements in Section~\ref{sec:application}.
-While trending evaluation is not the only aspect of evaluation, it is a valuable addition to the evaluation of nowcasts, forecasts, and measurements.
-For models with very different accuracies, these are usually replicated in trending evaluation, but the trending evaluation can differentiate between models with similar accuracies.
+While trending evaluation should not be the only aspect, it is a valuable addition to evaluating nowcasts, forecasts, and measurements.
+Models with very different accuracies are usually scored similiar in trending evaluation, but trending evaluation can differentiate between models with similar accuracies.
 As in the application in Section~\ref{sec:application-covid}, models with average point forecast evaluation measures can have the most meaningful positive trending ability.
 
-Throughout this paper, we did not expand on two modeling aspects, which we leave for further research.
+We did not expand on two modeling aspects throughout this paper, which we leave for further research.
 In the estimation, we did not consider sequential correlation.
-The computation of differences is a standard procedure in time series analysis to remove sequential dependence, but, in general, the trending measures could be adapted to consider sequential correlation.
+The computation of differences is a standard procedure in time series analysis to remove sequential dependence, but, in general, some could r.
 Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation by using time series bootstrap methods~\parencite{Hardle2003,Kreiss2012}.

### 2024-03-18 sections/05-conclusion.tex
Update on Overleaf.

@@
 In the estimation, we did not consider sequential correlation.
-The computation of differences is a standard procedure in time series analysis to remove sequential dependence, but, in general, some could r.
-Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation by using time series bootstrap methods~\parencite{Hardle2003,Kreiss2012}.
+The computation of differences is a standard procedure in time series analysis to remove sequential dependence, but, in general, some could remain, and the estimators could account for it.
+Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation using time-series bootstrap methods~\parencite{Hardle2003,Kreiss2012}.
 
-The estimators of Section~\ref{subsec:trending-measures} could encourage predictors to exploit imbalances in the number of observed positive and negative changes.
-A large difference in the number of observed positive and negative changes is unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data.
-In addition, we do not advise assessing signals solely on their trending ability and, thus, provide an incentive for the exploitation of estimator characteristics.
-However, if the number of positive and negative observed changes differs widely, the use of unbalanced-data-aware measures should be considered.
+The estimators of Section~\ref{subsec:trending-measures} do not account for imbalances in the number of observed positive and negative changes. \todo{Jolliffe hinzufügen}
+Large differences in the number of observed positive and negative changes are unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.
+However, if the number of positive and negative observed changes differs widely, unbalanced-data-aware measures should be considered.
 There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.

### 2024-03-18 sections/01-introduction.tex
Update on Overleaf.

@@
 In the medical context, nowcasting is an estimation procedure in which the value of a target variable for a specific time is estimated based on current preliminary measurements, which are only finalized later.
-Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \parencite{gunther2021nowcasting} or probabilistic measures \parencite{Wolffram2023}.
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \parencite{gunther2021nowcasting} or probabilistic scoring rules \parencite{Wolffram2023}.
 The aspect of trending ability is not considered in the literature to the best of our knowledge. 
-However, trending evaluation reveals a method's capability of predicting the development of the target variable, for example, the number of cases of an epidemic, that can be the foundation of decisions on introducing or canceling policy measures.
-\todo{Satz ist noch nicht gut}
+However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic.
+The epidemic's development, in turn, can be the foundation of decisions on introducing or canceling policy measures.
 

### 2024-03-18 sections/03-trending.tex
Update on Overleaf.

@@
 The numerator counts the number of same-sign-changes, while the denominator is the number of considered pairs $(\diffyt, \diffxt)$.
-Thus, $\acc$ is the ratio of concordant changes over all changes.
+Thus, $\acc$ is the proportion of concordant changes on all changes.
 We refer to this estimator as the trending ratio of the prediction and set $\mathcal{T} = \{l, \dots, T\}$. 
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
-Similar evaluations are used in other scientific areas, for example, in dichotomous forecasting or as a confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
+Similar evaluations are used in other scientific areas, for example, with contingency tables in dichotomous forecasting or with confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
 4]{James2021}[Ch.
 3]{Jolliffe2012}.
-There, a wide range of other methods are usually used to analyze further characteristics of such count tables.
+Many other measures can be adapted from those fields to deepen the trending analysis.
 Two simple measures that focus on either a positive or negative predicted change are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
@@
 Data pairs with missing values can be excluded from the set $\mathcal{T}$ to calculate the measures. 
-However, systematical missing values could lead to a biased estimate of the trending ratio and missing data should be inspected for any underlying patterns.
+However, systematical missing values could lead to a biased estimate of the trending ratio, and missing data should be inspected for any underlying patterns.
 In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}, only a few missing values occur, and we provide information on the missing data by inspecting the missing data as lists.
@@
 Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} \parencite[see, e.g.,][]{Saugel2015,Critchley2010}.
-Points within that area are ommited in the four-quadrant plot and in the calculation of the measures.
-In particular, the prediction models are likely to have a noise component and, thus, $\diffx$ should be subject to an exclusion area.
+Points within that area are omitted in the four-quadrant plot and the calculation of the measures.
+In particular, the prediction models are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
 We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $E$ by
@@
 Figure~\ref{fig:trending_4q} visualizes examples of exclusion area shapes.
-A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ 
- ($\varepsilon_x, \varepsilon_y > 0$), leaves out points that are small in both components and, therefore, likely to be driven by noise.
-Points where the true value or signal is unlikely to be noise-driven, are included.
+A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ ($\varepsilon_x, \varepsilon_y > 0$), leaves out points that are small in both components and, therefore, likely to be driven by noise.
+Points where the true value or signal is unlikely to be noise-driven are included.
 In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
@@
 A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
-For examples on such plots, see Section~\ref{sec:application-covid}.
+For examples of such plots, see Section~\ref{sec:application-covid}.
 
@@
 Let $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
-For trending assessment, we compare the probability of positive change with the observed occurrence of positive changes. 
+For trending assessment, we compare the predicted probability of positive change with the occurrence of positive changes. 
 For the prediction, a positive change is predicted for any value at $t$ larger than the true value $y_{t-l}$. 
@@
 The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
-Dichotomous forecasts predict a binary outcome, as here, a positive or negative change.
+Dichotomous forecasts predict a binary outcome, such as a positive or negative change.
 Numerical measures for forecast quality are scoring rules.
@@
 
-Graphical methods are a standard tool for evaluating solely the calibration of probabilistic forecasts.
+Graphical methods are a standard tool for evaluating the calibration of probabilistic forecasts in detail.
 In dichotomous forecasting, the reliability diagram is frequently used~\parencite{Ranjan2010}.

### 2024-03-18 sections/05-conclusion.tex
Update on Overleaf.

@@
 The trending ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates trending.
-Visually, it is the ratio of concordant points over all points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
+Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
 The positive and negative trending ratio analyze the trending ratio given the predicted change is positive or negative, respectively.
@@
 While trending evaluation should not be the only aspect, it is a valuable addition to evaluating nowcasts, forecasts, and measurements.
-Models with very different accuracies are usually scored similiar in trending evaluation, but trending evaluation can differentiate between models with similar accuracies.
+Models with very different accuracies are usually scored similarly in trending evaluation, but trending evaluation can differentiate between models with similar accuracies.
 As in the application in Section~\ref{sec:application-covid}, models with average point forecast evaluation measures can have the most meaningful positive trending ability.
@@
 
-The estimators of Section~\ref{subsec:trending-measures} do not account for imbalances in the number of observed positive and negative changes. \todo{Jolliffe hinzufügen}
-Large differences in the number of observed positive and negative changes are unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.
+The estimators of Section~\ref{subsec:trending-measures} do not account for imbalances in the number of observed positive and negative changes \parencite[for a theoretical analysis, see][Chapter 3]{Jolliffe2012}.
+Significant differences in the number of observed positive and negative changes are unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.
 However, if the number of positive and negative observed changes differs widely, unbalanced-data-aware measures should be considered.

### 2024-03-19 paper.tex
Update on Overleaf.

@@
 \begin{abstract}
-An intuitive and easily accessible information on a quantity measured over time is whether it increases or decreases.
-Trending refers to the ability to predict the direction of change for a true variable accurately. 
-Thus, an appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers. 
+An intuitive and easily interpretable information on a quantity measured over time is whether it increases or decreases.
+Trending refers to the ability to predict or measure the direction of change for a true variable accurately. 
+An appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers, as it enables them to assess the method's credibility accurately.
 While the current practice of evaluating signals is based on averaging over local deviation measures, we review and introduce trending measures that consider whether the direction of development fits the quantity of interest.
@@
 Using bootstrap confidence intervals and exclusion areas, we provide extensions for noisy data and estimation uncertainty.
-We review and present new visualization techniques to assess the local trending behaviors and enlarge the scope to probabilistic nowcasts and forecasts. 
+We review and present new visualization techniques to assess the local trending behavior and enlarge the scope to probabilistic nowcasts and forecasts. 
 We apply the methods to three large data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
@@
 
-\listoftodos
+% \listoftodos
 

### 2024-03-19 sections/01-introduction.tex
Update on Overleaf.

@@
 The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
-Conventionally, methods are evaluated using distance measures of local differences between predictions (or measurements) and target values into account. 
-These measures fail to consider whether the right direction of change is predicted or measured correctly.
+Conventionally, methods are evaluated using distance measures of local differences between predictions (or measurements) and target values. 
+These measures do not consider whether the right direction of change is predicted or measured.
 The information about whether an increase or decrease is predicted correctly is crucial when making decisions based on the estimator's prediction results. 

### 2024-03-19 sections/01-introduction.tex
Update on Overleaf.

@@
 The information about whether an increase or decrease is predicted correctly is crucial when making decisions based on the estimator's prediction results. 
-In the following, we provide an overview of the characteristics and current evaluation schemes of the three application areas: measurement, forecasting, and nowcasting.
-We highlight why trending evaluation is highly relevant in the respective fields.
+In the following, we provide a more specific overview of the characteristics and current evaluation schemes of the three application fields, forecasting, nowcasting, and measurement, and highlight why trending evaluation is highly relevant in the respective fields.
+
+The above-described trending idea is of fundamental interest for evaluating and comparing forecasting methods. 
+Forecasting methods predict the future based on historical data, patterns, and exogenous factors. 
+The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
+A forecast's trending is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
+In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable. 
+Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules \textcite[see the review in][]{hyndman2006another}.
+These typical measures are based on the absolute difference between the prediction and the true values, locally or globally. 
+They are not capable of assessing the trending ability discussed above.
+
+Methodologically evolved from forecasting, nowcasting methods focus on predictions for the present, the immediate future, and the recent past \parencite{banbura2013now} and are now widely used in fields such as economics and medicine \parencite{bok2018macroeconomic, Wolffram2023}.
+Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours \parencite{browning1989nowcasting,schmid2019nowcasting}.
+In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \parencite{banbura2013now}.
+In medicine, epidemic nowcasting assesses the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors \parencite{wu2021nowcasting}.
+Nowcasting methods use high-frequency indicators related to the target variable and estimate the value of a target variable for a specific time based on current preliminary measurements, which are finalized with a considerable time delay.
+Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period \parencite{castle2017forecasting}. 
+For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported \parencite{gunther2021nowcasting}. 
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \parencite{gunther2021nowcasting} or probabilistic scoring rules \parencite{Wolffram2023}.
+The aspect of trending ability is not considered in the literature to the best of our knowledge. 
+However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic.
+The epidemic's development, in turn, can be the foundation of decisions on introducing or canceling policy measures.
 
@@
 In this paper, we build upon the existing methods and extend them with new functionality.
-
-The above-described trending idea is also of fundamental interest for evaluating and comparing forecasting methods. 
-While measurement series are created by recording data using measuring instruments or direct observation, forecasting predicts the future based on historical data, patterns, and exogenous factors. 
-The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
-A forecast's trending is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
-In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable; different measures are, for example, reviewed in \textcite{hyndman2006another}.
-Popular measures of this type are scale-dependent measures such as the \ac{rmse} or measures based on percentage errors such as the root mean squared percentage error. 
-These typical measures are based on the absolute difference between the prediction and the true values, locally or globally. 
-They are not capable of assessing the trending ability discussed above.
-
-In contrast to conventional forecasting methods, nowcasting methods focus on predictions for the present, the immediate future, and the recent past \parencite{banbura2013now} and are now widely used in fields such as economics and medicine \parencite{bok2018macroeconomic, Wolffram2023}.
-Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours \parencite{browning1989nowcasting,schmid2019nowcasting}.
-In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \parencite{banbura2013now}.
-Nowcasting methods use high-frequency indicators related to the target variable.
-Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period \parencite{castle2017forecasting}. 
-In medicine, epidemic nowcasting assesses the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors \parencite{wu2021nowcasting}.
-For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported \parencite{gunther2021nowcasting}. 
-In the medical context, nowcasting is an estimation procedure in which the value of a target variable for a specific time is estimated based on current preliminary measurements, which are only finalized later.
-Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \parencite{gunther2021nowcasting} or probabilistic scoring rules \parencite{Wolffram2023}.
-The aspect of trending ability is not considered in the literature to the best of our knowledge. 
-However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic.
-The epidemic's development, in turn, can be the foundation of decisions on introducing or canceling policy measures.
 

### 2024-03-19 sections/01-introduction.tex
Update on Overleaf.

@@
 
-Measurement aims to obtain accurate and reliable data about the current state of a system. 
+Measurement is more technical and aims to obtain accurate and reliable data about the current state of a system based on sensors. 
 A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
-When introducing new measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
+When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
 Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods \parencite{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series \parencite{watson2010method}.
-Furthermore, graphical tools were developed, such as the Bland-Altman diagram, visualizing the differences between the methods in relation to their mean value \parencite{bland1986statistical}. 
-Trending considerations, that is, the consistency of the test methods development with the gold standard, is a field of active research in the last years \parencite{Saugel2015,saugel2018error,hiraishi2021concordance}.
-In this paper, we build upon the existing methods and extend them with new functionality.
+Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \parencite{bland1986statistical}. 
+Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years \parencite{Saugel2015,saugel2018error,hiraishi2021concordance}.
 
 As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
-However, regardless of the application area, most conventional method comparisons and assessments consider local absolute deviations without assessing whether the correct direction of change is being predicted or measured.
-In this paper, we develop trending evaluation methods that complement current evaluation measures with trending measurements.
+We apply the existing trending methods from measurement evaluation to the other fields and extend their functionality.
+In this paper, we develop trending evaluation methods that complement the current evaluation with trending measurements.
 
@@
 Section \ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts. 
-In particular, we introduce a new graphical method for evaluating trending, the conditional trending plot, which can represent the simultaneous evaluation of different intervals and asymmetries. 
+We introduce a new graphical method for evaluating trending, the conditional trending plot, which can simultaneously assess different intervals and asymmetries. 
 In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting. 

### 2024-03-19 sections/03-trending.tex
Update on Overleaf.

@@
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
-Development always has respect to a time horizon, that is, the amount of time that has passed. 
+The development in trending assessment always has respect to a time horizon, that is, the amount of time that has passed. 
 In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the question at hand.
@@
 
-The definition of the \textit{predicted} change depends on the context.
+The definition of the \textit{predicted} change depends on the context;
 Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
@@
 We call $t$ the \textit{forecast time} and $\tau$ the \textit{issue time}.
-We compute the predicted change by
+The predicted change is computed by
 \begin{equation}
@@
 When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet. 
-Thus, the predicted change can be computed with the knowledge of time $t$, and the trending assessment can be used in practice at time $t$.
+Thus, the predicted change can be computed with the knowledge of time $t$. Therefore, the trending information of the following section can be used in practice at time $t$.
 
@@
 The predicted change is computed by 
-\begin{equation} \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l} \label{eq:diffxl_forecasting}
+\begin{equation} \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l}\label{eq:diffxl_forecasting}
 \end{equation}
+with the same structure as in the nowcasting case. 
 The indices are chosen to be consistent with $\diffyl$. 
@@
 yields the change by the test method. 
-It is computed purely by the test method without the gold standard $y_t$ to analyze whether the gold standard and test method changes are consistent.
+It is computed purely by the test method without the gold standard $y_t$ to analyze whether the gold standard and test method changes are consistent within themselves.
 

### 2024-03-19 sections/03-trending.tex
Update on Overleaf.

@@
 The trending assessment is the same for all three application areas, given the notation for the respective application of Section~\ref{subsec:notation}.
-In the following, we omit the difference horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$ as defined in Table~\ref{tab:notation}.
+In the following, we omit the difference horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
 As noted above, a signal's trending is perfect if all the change directions are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
@@
 The four-quadrant plot is intuitive to interpret, and the magnitude and direction of change are shown simultaneously.
-It can be extended by including information on the date in the point color.
+It can be extended by including information on the time index in the point color to reveal effects over time.
 In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, $(\diffyt[l], \diffxt[l])$ is blue and turns green until $(\diffyt[T], \diffxt[T])$.
@@
 Thus, $\acc$ is the proportion of concordant changes on all changes.
-We refer to this estimator as the trending ratio of the prediction and set $\mathcal{T} = \{l, \dots, T\}$. 
+We refer to this estimator as the \textit{trending ratio} of the prediction and set $\mathcal{T} = \{l, \dots, T\}$. 
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
@@
 \end{align}
-They give the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, respectively, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$.
+They give the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$.
 
@@
 However, systematical missing values could lead to a biased estimate of the trending ratio, and missing data should be inspected for any underlying patterns.
-In the applications in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}, only a few missing values occur, and we provide information on the missing data by inspecting the missing data as lists.
 
@@
 The above measures can be extended to account for information on the point's location within the quadrant.
-For example, points close to the zero point may have less explanatory power or are less reliable than points far away from zero on one of the diagonals.
+For example, points close to the zero point may have less explanatory power or may be less reliable than points far away from zero on one of the diagonals.
 Suppose noise or non-systematic effects are present in the true values or predictions. 
 In that case, a point's assignment to a quadrant can be driven by noise instead of a systematic trending ability of the signal.
-This happens more frequently for points with at least one small coordinate.
+This is more likely for points with at least one small coordinate.
 
@@
 A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ ($\varepsilon_x, \varepsilon_y > 0$), leaves out points that are small in both components and, therefore, likely to be driven by noise.
-Points where the true value or signal is unlikely to be noise-driven are included.
+Points are included if the true value or signal is unlikely to be noise-driven.
 In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
@@
 Confidence intervals can account for the estimation uncertainty of the measures above.
-Bootstrap confidence intervals are based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
-New samples are drawn with replacement from the dataset and are not based on parametric assumptions as classical confidence intervals are.
-The confidence interval is then computed based on those bootstrap samples.
+Bootstrap confidence intervals are based on resampling and not on parametric assumptions as classical confidence intervals are~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
+New samples are drawn with replacement from the dataset and the confidence interval is then computed based on those bootstrap samples.
 We use the \ac{bca} approach for bootstrapping in the following, as it holds the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-trending-bootstrap}).
@@
 for $x \neq 0$ through a \ac{kde}.
-A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages~\parencite[e.g., for  Python in][]{Seabold2010}.
+A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages, for example in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
 The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ for all values of $x \in \R$.
@@
 In nowcasting and forecasting, probabilistic predictions have become more prevalent in recent years (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
-Probabilistic predictions include a point estimate and information on the forecast uncertainty and quantiles simultaneously.
+Probabilistic predictions include a point estimate and information on the prediction uncertainty and quantiles simultaneously.
 For trending analysis, probabilistic predictions thus also contain a probability of a positive or negative change.
@@
 The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\parencite{Ranjan2010,Mitchell2011}.
-Calibration refers to the statistical consistency of forecasts and observations, while sharpness measures the spread of the forecast distribution.
-A smaller spread, thus, higher sharpness, is preferable, as it indicates greater confidence in the prediction.
+Calibration refers to the statistical consistency of forecasts and observations; that is, the event occurs with the issued probability.
+Sharpness refers to the spread of the forecast; that is, higher probabilities are preferable as they convey a higher certainty.
 The basic paradigm of probabilistic forecasting is to maximize sharpness, subject to calibration~\parencite{Gneiting2014}.
@@
 In dichotomous forecasting, the reliability diagram is frequently used~\parencite{Ranjan2010}.
-The reliability diagram plots the observed frequency of the positive outcome against the predicted probability.
-Thus, it visualizes the observed frequency of a positive change given a predicted probability for positive change of, for example, approximately 0.6.
+The reliability diagram plots the observed frequency of the positive outcome against the (binned) predicted probability.
+For example, it shows the proportion of observed increases, given the predicted probability of increase was approximately $0.7$.
 Ideally, the predicted probability equals the observed frequency, and the reliability diagram is a 45-degree line.

### 2024-03-19 sections/03-trending.tex
Update on Overleaf.

@@
 When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet. 
-Thus, the predicted change can be computed with the knowledge of time $t$. Therefore, the trending information of the following section can be used in practice at time $t$.
+Thus, the predicted change can be computed with the knowledge of time $t$. Therefore, the trending information in the following section can be used in practice at time $t$.
 
@@
 Bootstrap confidence intervals are based on resampling and not on parametric assumptions as classical confidence intervals are~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
-New samples are drawn with replacement from the dataset and the confidence interval is then computed based on those bootstrap samples.
+New samples are drawn with replacement from the dataset, and the confidence interval is then computed based on those bootstrap samples.
 We use the \ac{bca} approach for bootstrapping in the following, as it holds the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-trending-bootstrap}).
@@
 for $x \neq 0$ through a \ac{kde}.
-A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages, for example in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
+A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages, for example, in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
 The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ for all values of $x \in \R$.
@@
 Calibration refers to the statistical consistency of forecasts and observations; that is, the event occurs with the issued probability.
-Sharpness refers to the spread of the forecast; that is, higher probabilities are preferable as they convey a higher certainty.
+Sharpness refers to the spread of the forecast; higher probabilities are preferable as they convey a higher certainty.
 The basic paradigm of probabilistic forecasting is to maximize sharpness, subject to calibration~\parencite{Gneiting2014}.

### 2024-03-19 sections/05-conclusion.tex
Update on Overleaf.

@@
 The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
-For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or assess models with high trending differences, and 1000 samples enable a more precise differentiation.
+For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or assess models with high trending differences.
 
-A conditional trending plot visualizes the probability of correct trending over the predicted change of the signal (see Section\ref{subsec:trending-cond-prob}).
+A conditional trending plot visualizes the probability of correct trending over the predicted change of the signal (see Section~\ref{subsec:trending-cond-prob}).
 It is based on a multivariate \acf{kde} of predicted and observed change.

### 2024-03-19 sections/50-appendix-data-generation.tex
Update on Overleaf.

@@
 \end{equation*}
-where $n_t$ is a truncated normal distribution with mean 1 and standard deviation 0.5, truncated at 0 and a symmetric Bernoulli random variable with parameter $k$.
+where $n_t$ is a truncated normal distribution with mean one and standard deviation 0.5, truncated at 0, and $b_t$ is a symmetric Bernoulli random variable with parameter $k$.
 For a time-varying trending ratio, the parameter $k$ is modified to have a wave-shape over time, that is,
@@
 
-In the second approach, $\diffyt$ and $\diffxt$ are modelled to be multivariate normal with mean 0 and covariance matrix
+In the second approach, $\diffyt$ and $\diffxt$ are modeled to be multivariate normal with mean 0 and covariance matrix
 \begin{equation*}
@@
     \caption{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.
-Although the first and second plot differ over time, their difference is not discernible in the plots.
-The third data set's asymmetry is visible in the plot but the decrese of trending ability near 0 is not visible.
+Although the first and second plots differ over time, their difference is not discernible in the plots.
+The third data set's asymmetry is visible in the plot, but the decrease in the trending ability near 0 is not visible.
 }
@@
 We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
-The considered datasets are outlined in Appendix~\ref{subsec:app-trending-data-generation}, the first dataset with asymmetric dependence.
+The considered datasets are the first dataset with asymmetric dependence and the second dataset outlined in Appendix~\ref{subsec:app-trending-data-generation}.
 In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\parencite{Virtanen2020}.
@@
 The computation time for the \ac{bca} method is slightly larger than for the other methods, but all methods have a moderate computation time.
-\Ac{bca} is the only method that maintains the confidence level also for small datasets while increasing the computation time only moderately for larger datasets.
+\Ac{bca} is the only method that maintains the confidence level for small datasets while increasing the computation time only moderately for larger datasets.
 Therefore, we use the \ac{bca} method for confidence intervals in the applications in Section~\ref{sec:application}.
@@
 Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
-\unsure{ist das richtig?}
 The quantiles $q_p$ generally differ for each time step, but we omit an index here for ease of notation.

### 2024-03-21 sections/03-trending.tex
Merge changes; add bmc template

@@
 Consequently, when evaluating the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
-A simple yet insightful method is the four-quadrant plot~\parencite[see, e.g., ][]{perrino1998intraoperative,Saugel2015}. 
+A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{perrino1998intraoperative,Saugel2015}. 
 In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
@@
 There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.
-However, they lack the four-quadrant plot's clarity and intuition without adding more information on trending~\parencite{Saugel2015}.
+However, they lack the four-quadrant plot's clarity and intuition without adding more information on trending~\citep{Saugel2015}.
 
@@
 
-Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data \parencite{Critchley2010,Saugel2015}. 
+Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data \citep{Critchley2010,Saugel2015}. 
 With that, we mainly estimate the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
@@
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
-Similar evaluations are used in other scientific areas, for example, with contingency tables in dichotomous forecasting or with confusion matrix in classification analysis~\parencites(see, e.g., the introductions in)()[Ch.
-4]{James2021}[Ch.
-3]{Jolliffe2012}.
+Similar evaluations are used in other scientific areas, for example, with contingency tables in dichotomous forecasting or with confusion matrix in classification analysis~\citetext{see, e.g., the introductions in \citealt{James2021}, Ch.
+4, and \citealp{Jolliffe2012}, Ch.
+3}.
 Many other measures can be adapted from those fields to deepen the trending analysis.
@@
 
-Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} \parencite[see, e.g.,][]{Saugel2015,Critchley2010}.
+Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} \citep[see, e.g.,][]{Saugel2015,Critchley2010}.
 Points within that area are omitted in the four-quadrant plot and the calculation of the measures.
@@
 
-The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\parencite{Brier1950} and is defined as
+The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\citep{Brier1950} and is defined as
 \begin{equation*}
@@
 \end{equation*}
-The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\parencite{Ranjan2010,Mitchell2011}.
+The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\citep{Ranjan2010,Mitchell2011}.
 Calibration refers to the statistical consistency of forecasts and observations; that is, the event occurs with the issued probability.
 Sharpness refers to the spread of the forecast; higher probabilities are preferable as they convey a higher certainty.
-The basic paradigm of probabilistic forecasting is to maximize sharpness, subject to calibration~\parencite{Gneiting2014}.
+The basic paradigm of probabilistic forecasting is to maximize sharpness, subject to calibration~\citep{Gneiting2014}.
 
 Graphical methods are a standard tool for evaluating the calibration of probabilistic forecasts in detail.
-In dichotomous forecasting, the reliability diagram is frequently used~\parencite{Ranjan2010}.
+In dichotomous forecasting, the reliability diagram is frequently used~\citep{Ranjan2010}.
 The reliability diagram plots the observed frequency of the positive outcome against the (binned) predicted probability.

### 2024-03-21 sections/04-application.tex
Merge changes; add bmc template

@@
 Various indicators measured the pandemic's spread and severity. 
-In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\parencite{RobertKochInstitute2021}.
+In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\citep{RobertKochInstitute2021}.
 The Robert Koch Institute (RKI) provided preliminary daily reports on the seven-day hospitalization rate.
 However, these reports were subject to severe delays and revisions in two sources.
-The first source is technical delays in the reporting process, for example, due to different authorities passing the data to the RKI~\parencite{RobertKochInstitute2024}.
+The first source is technical delays in the reporting process, for example, due to different authorities passing the data to the RKI~\citep{RobertKochInstitute2024}.
 The second, more systematic source is the structure of the seven-day hospitalization rate.
@@
 Nevertheless, as the seven-day hospitalization rate was considered a major indicator of the pandemic's development, many organizations and institutions started to issue nowcasts, including research teams and newspapers.
-To collect nowcasts of the seven-day hospitalization rate by different nowcast groups, the COVID19-Nowcasting-Hub was established~\parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+To collect nowcasts of the seven-day hospitalization rate by different nowcast groups, the COVID19-Nowcasting-Hub was established~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
 The nowcasts contain the seven-day hospitalization rate's predictive mean, median, and other quantiles.
@@
         \end{subfigure}
-    \caption{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+    \caption{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
     The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
@@
 Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the respective names in the COVID19-Nowcasting-Hub.
-For information on the model design, we refer to~\textcite{Wolffram2023}.
+For information on the model design, we refer to~\citet{Wolffram2023}.
 Using the nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
 We denote them by ENS-MEAN and ENS-MED.
-\textcite{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams stated in a preregistered study. 
+\citet{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams stated in a preregistered study. 
 In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
@@
 Figure~\ref{fig:app-covid-true-nowcast} displays the true and nowcast data for the evaluation period.
-The time comprises the fourth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany, lasting until May 28, 2022~\parencite{Tolksdorf2022}.
+The time comprises the fourth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany, lasting until May 28, 2022~\citep{Tolksdorf2022}.
 
 Table~\ref{tab:app-covid-rmse} summarizes the point evaluation measures for the issued mean of the different models.
-The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\parencite[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
+The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\citep[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
 The best-performing models in terms of RMSE and MAE are the ILM and RKI models.
@@
 Good forecasts are crucial for planning staff and resources.
-Several models are used for predicting hourly outcomes in a study by \textcite{Rostami-Tabar2023}.
+Several models are used for predicting hourly outcomes in a study by \citet{Rostami-Tabar2023}.
 Every 12 hours, the models issue hourly forecasts for the next 48 hours.
@@
 
-\textcite{Rostami-Tabar2023} publish means and probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
+\citet{Rostami-Tabar2023} publish means and probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
 The models are trained on data from April 1, 2014, to February 28, 2018, and evaluated on data from March 1, 2018, to February 28, 2019.
-For further notes on the models and the evaluation, we refer to \textcite{Rostami-Tabar2023}.
+For further notes on the models and the evaluation, we refer to \citet{Rostami-Tabar2023}.
 From the issued forecasts, we use the mean as a point forecast for the trending assessment and evaluate the probabilistic trending subsequently.
@@
 More than 8,600 forecasts are available for all models, with differences in the number due to missing values on four afternoons in 2018.
-Note that the reported values for the RMSE differ from those in \textcite{Rostami-Tabar2023}.
+Note that the reported values for the RMSE differ from those in \citet{Rostami-Tabar2023}.
 In contrast to their work, we use only the forecast data at least 36 hours ahead and not the entire forecast data for evaluation.
@@
 The data is from the MIMIC-III database, including various information on patients in critical care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}).
-The data is publicly available and also includes numerical measurement data such as heart rate, blood pressure, and oxygen saturation in a waveform database \parencites[][]{Moody2017}[available through][]{Goldberger2000}.
+The data is publicly available and also includes numerical measurement data such as heart rate, blood pressure, and oxygen saturation in a waveform database \citetext{\citealp{Moody2017}; available through \citealp{Goldberger2000}}.
 
@@
 While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods.
-For an overview of blood pressure measurement methods, see \textcite{Saugel2014}.
+For an overview of blood pressure measurement methods, see \citet{Saugel2014}.
 For critical patients, changes in blood pressure can be crucial for the treatment.
-Thus, trending assessment can be performed in addition to standard accuracy analysis~\parencite[see, for example, ][]{Mostafa2020}.
+Thus, trending assessment can be performed in addition to standard accuracy analysis~\citep[see, for example, ][]{Mostafa2020}.
 Thus, we assess the trending ability of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
@@
 Whether these points are due to measurement errors or extreme values is not distinguishable.
-Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile \parencite[see][]{Critchley2010}.
+Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile \citep[see][]{Critchley2010}.
 We do not follow this approach here, as the extreme values are not necessarily measurement errors and could be particularly relevant.

### 2024-03-21 sections/05-conclusion.tex
Merge changes; add bmc template

@@
 The computation of differences is a standard procedure in time series analysis to remove sequential dependence, but, in general, some could remain, and the estimators could account for it.
-Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation using time-series bootstrap methods~\parencite{Hardle2003,Kreiss2012}.
+Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation using time-series bootstrap methods~\citep{Hardle2003,Kreiss2012}.
 
-The estimators of Section~\ref{subsec:trending-measures} do not account for imbalances in the number of observed positive and negative changes \parencite[for a theoretical analysis, see][Chapter 3]{Jolliffe2012}.
+The estimators of Section~\ref{subsec:trending-measures} do not account for imbalances in the number of observed positive and negative changes \citep[for a theoretical analysis, see][Chapter 3]{Jolliffe2012}.
 Significant differences in the number of observed positive and negative changes are unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.
 However, if the number of positive and negative observed changes differs widely, unbalanced-data-aware measures should be considered.
-There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \parencite{Cohen1960} or those listed in \textcite[Table 3.3]{Jolliffe2012}.
+There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \citep{Cohen1960} or those listed in \citet[Table 3.3]{Jolliffe2012}.

### 2024-03-21 sections/50-appendix-data-generation.tex
Merge changes; add bmc template

@@
 In the \textit{percentile} approach, the confidence interval for the level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
-The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
-The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
+The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\citep{Davison1997}.
+The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\citep{Efron1987}.
 Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
@@
 The considered datasets are the first dataset with asymmetric dependence and the second dataset outlined in Appendix~\ref{subsec:app-trending-data-generation}.
-In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\parencite{Virtanen2020}.
+In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\citep{Virtanen2020}.
 The prescribed confidence level is 90 \%, and the number of bootstrap samples is $10,000$.
@@
 
-We examine the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\parencite{Seabold2010} in Figure~\ref{fig:trending-cond-prob-bw}.
+We examine the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\citep{Seabold2010} in Figure~\ref{fig:trending-cond-prob-bw}.
 While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.

### 2024-03-21 paper.tex
Use bmc template

@@
-\documentclass[oneside]{article}
-
+%\documentclass[referee,sn-mathphys-ay,pdflatex]{sn-jnl}
+%\documentclass[sn-mathphys-ay,pdflatex]{sn-jnl}
+\documentclass[pdflatex]{sn-jnl}
 % Language setting
@@
 \usepackage[english]{babel}
-\usepackage[utf8]{inputenc} % für Umlaute
+\usepackage[utf8]{inputenc} % fuer Umlaute
 \usepackage{fontenc}
@@
 % Replace `letterpaper' with`a4paper' for UK/EU standard size
-\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+%\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
 
@@
 \usepackage[short]{optidef} % for aligning linear programs
-\usepackage{nccmath} % for fleqn environment
+%\usepackage{nccmath} % for fleqn environment
 
 %%% --- Other
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
+%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 %\usepackage[hidelinks]{hyperref}
-\usepackage[style=authoryear,backend=biber]{biblatex}
+\usepackage[style=authoryear,backend=biber,natbib=true]{biblatex}
+%\usepackage{natbib}
+%\bibliographystyle{
 \addbibresource{library.bib}
@@
 
+\begin{document}
+
 \title{Trending assessment for nowcasts, measurements, and forecasts}
-\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
+%\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
+\author[1]{Oliver Grothe}
+\equalcont{These authors contributed equally to this work.}
+\author[1]{Bolin Liu}
+\equalcont{These authors contributed equally to this work.}
+\author*[1]{Jonas Rieger}
+\equalcont{These authors contributed equally to this work.}
 
-\begin{document}
-\maketitle
+\affil*[1]{\orgdiv{Institute for Operations Research}, \orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\street{Kaiserstr. 12}, \city{Karlsruhe}, \postcode{76131}, \country{Germany}}}
 
-\begin{abstract}
+\abstract{
 An intuitive and easily interpretable information on a quantity measured over time is whether it increases or decreases.
@@
 We apply the methods to three large data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
-\end{abstract}
+}
 
+\keywords{}
+
+\maketitle
 % \listoftodos
@@
 \printbibliography
+%\bibliography{library,library-jonas}
 

### 2024-03-21 paper.tex
Improve paper

@@
 %\usepackage[hidelinks]{hyperref}
-\usepackage[style=authoryear,backend=biber,natbib=true]{biblatex}
+\usepackage[style=authoryear,backend=bibtex,natbib=true]{biblatex}
 %\usepackage{natbib}
@@
 %%% --- Commenting, Todonotes
-\usepackage[colorinlistoftodos,prependcaption]{todonotes}
-\usepackage{comment, soul}
-\newcommand{\ar}{$\rightarrow$}
-\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
+%\usepackage[colorinlistoftodos,prependcaption]{todonotes}
+%\usepackage{comment, soul}
+%\newcommand{\ar}{$\rightarrow$}
+%\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
 %\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}
 
-\usepackage{xargs,ifthen} % Use optional arguements in commands
-\newcommandx{\unsure}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
-\newcommandx{\discuss}[2][1=]{\todo[linecolor=orange,backgroundcolor=orange!25,bordercolor=orange,#1]{#2}}
-\newcommandx{\missing}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-\newcommandx{\change}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
-\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
-\newcommandx{\done}[2][1=]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{#2}}
-\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
-\usepackage{marginnote}
-\let\marginpar\marginnote
+%\usepackage{xargs,ifthen} % Use optional arguements in commands
+%\newcommandx{\unsure}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
+%\newcommandx{\discuss}[2][1=]{\todo[linecolor=orange,backgroundcolor=orange!25,bordercolor=orange,#1]{#2}}
+%\newcommandx{\missing}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+%\newcommandx{\change}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
+%\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
+%\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
+%\newcommandx{\done}[2][1=]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{#2}}
+%\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
+%\usepackage{marginnote}
+%\let\marginpar\marginnote
 
@@
 \DeclareAcronym{rmse}{short=RMSE, long=root mean square error}
+\DeclareAcronym{mae}{short=MAE, long=mean absolute error}
+\DeclareAcronym{tadda}{short=TADDA, long=targeted absolute deviation with direction augmentation}
 
@@
 An intuitive and easily interpretable information on a quantity measured over time is whether it increases or decreases.
-Trending refers to the ability to predict or measure the direction of change for a true variable accurately. 
-An appropriate assessment of the trending ability is crucial for measurements, nowcasts, and forecasts for decision-makers, as it enables them to assess the method's credibility accurately.
-While the current practice of evaluating signals is based on averaging over local deviation measures, we review and introduce trending measures that consider whether the direction of development fits the quantity of interest.
+If the quantity of interest is not yet accessible, we refer to the ability to accurately predict, nowcast, or measure the direction of change for a true variable as trending. 
+An appropriate assessment of the trending ability is crucial for decision-makers, as it enables them to assess the method's credibility and changes in the underlying process early.
+While the current practice of evaluating signals is focused on averaging over local deviation measures, we present trending evaluation techniques as accompanying assessments that consider whether the development direction fits the quantity of interest.
 We formalize the concept of trending for all application areas and present quantitative measures. 
@@
 We review and present new visualization techniques to assess the local trending behavior and enlarge the scope to probabilistic nowcasts and forecasts. 
-We apply the methods to three large data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+We apply the methods to three large medical data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
 }

### 2024-03-21 sections/01-introduction.tex
Improve paper

@@
 The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
-Conventionally, methods are evaluated using distance measures of local differences between predictions (or measurements) and target values. 
+Conventionally, methods are evaluated using differences between predictions (or measurements) and target values, such as the \ac{rmse} or \ac{mae}. 
 These measures do not consider whether the right direction of change is predicted or measured.
-The information about whether an increase or decrease is predicted correctly is crucial when making decisions based on the estimator's prediction results. 
+However, the information about an increase or decrease is crucial to detect changes in the underlying process early. 
 In the following, we provide a more specific overview of the characteristics and current evaluation schemes of the three application fields, forecasting, nowcasting, and measurement, and highlight why trending evaluation is highly relevant in the respective fields.
 
-The above-described trending idea is of fundamental interest for evaluating and comparing forecasting methods. 
 Forecasting methods predict the future based on historical data, patterns, and exogenous factors. 
+In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008,Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
 The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
-A forecast's trending is perfectly consistent with the actual development of the target value if the actual change in the target value over this period matches the forecast change. 
 In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable. 
-Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules \textcite[see the review in][]{hyndman2006another}.
-These typical measures are based on the absolute difference between the prediction and the true values, locally or globally. 
+Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules \citep[see the review in][]{hyndman2006another}.
 They are not capable of assessing the trending ability discussed above.
+In a prediction competition on armed conflicts, trending recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with a trending component for evaluation. 
+However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
+This paper focuses on pure trending evaluation that accompanies traditional scores.
 
-Methodologically evolved from forecasting, nowcasting methods focus on predictions for the present, the immediate future, and the recent past \parencite{banbura2013now} and are now widely used in fields such as economics and medicine \parencite{bok2018macroeconomic, Wolffram2023}.
-Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours \parencite{browning1989nowcasting,schmid2019nowcasting}.
-In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \parencite{banbura2013now}.
-In medicine, epidemic nowcasting assesses the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors \parencite{wu2021nowcasting}.
+Methodologically evolved from forecasting, nowcasting methods focus on predictions for the present, the immediate future, and the recent past \citep{banbura2013now}.
+Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours \citep{browning1989nowcasting,schmid2019nowcasting}.
+%In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \citep{banbura2013now}.
+In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors \citep{wu2021nowcasting} or correct the daily COVID case numbers for events that have occurred but have not yet been reported \citep{gunther2021nowcasting}.
 Nowcasting methods use high-frequency indicators related to the target variable and estimate the value of a target variable for a specific time based on current preliminary measurements, which are finalized with a considerable time delay.
-Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period \parencite{castle2017forecasting}. 
-For example, the nowcasting method can correct the daily COVID case numbers for events that have occurred but have not yet been reported \parencite{gunther2021nowcasting}. 
-Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \parencite{gunther2021nowcasting} or probabilistic scoring rules \parencite{Wolffram2023}.
+Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period \citep{castle2017forecasting}. 
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \citep{gunther2021nowcasting} or probabilistic scoring rules \citep{Wolffram2023}.
 The aspect of trending ability is not considered in the literature to the best of our knowledge. 
-However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic.
+However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
 The epidemic's development, in turn, can be the foundation of decisions on introducing or canceling policy measures.
@@
 A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
-When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as individuals or environments.
-Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods \parencite{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series \parencite{watson2010method}.
-Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \parencite{bland1986statistical}. 
-Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years \parencite{Saugel2015,saugel2018error,hiraishi2021concordance}.
+When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
+Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods \citep{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series \citep{watson2010method}.
+Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \citep{bland1986statistical}. 
+Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years \citep{Saugel2015,saugel2018error,hiraishi2021concordance}.
 
-As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts. 
+As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts, as research interests in various disciplines reveal. 
 We apply the existing trending methods from measurement evaluation to the other fields and extend their functionality.
-In this paper, we develop trending evaluation methods that complement the current evaluation with trending measurements.
+The trending evaluation methods are meant to complement the current evaluation to reveal additional information on the considered methods.
 
 The main contributions of this paper are manifold.
-\begin{itemize}
-\item We formalize trending and present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
-\item We introduce the conditional trending plot, a new graphical method for assessing local trending behavior, and review bootstrap methods for calculating confidence intervals.
-\item We extend the concept of trending to probabilistic predictions.
-\item We apply trending evaluation to three applications: Measurement, nowcasting, and forecasting data. 
-\item We provide a ready-to-use code for trending evaluation. The code and the source code to replicate the results of our study are available at ... .
-\end{itemize}
+We formalize trending and present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
+We introduce the conditional trending plot, a new graphical method for assessing local trending behavior, and review bootstrap methods for calculating confidence intervals.
+We extend the concept of trending to probabilistic predictions.
+The methods are applied to three applications from the abovementioned fields: COVID-19-nowcasting, emergency department arrival forecasting, and blood pressure measurement.
+The ready-to-use code for trending evaluation from the applications is available on ... .
 

### 2024-03-22 sections/03-trending.tex
Fix layout and renew some plots

@@
     \scriptsize
-    \begin{tabularx}{0.75\textwidth}{l X}
+    \begin{tabular}{l l}
         \toprule
@@
         \bottomrule
-    \end{tabularx}
+    \end{tabular}
     \caption{Computation of the predicted change in the different applications.
@@
 \end{equation*}
+Lower values indicate a higher probabilistic trending ability of the method.
 The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\citep{Ranjan2010,Mitchell2011}.

### 2024-03-27 paper.tex
Update with comments by Grothe

@@
 %\usepackage[hidelinks]{hyperref}
-\usepackage[style=authoryear,backend=bibtex,natbib=true]{biblatex}
+\usepackage[style=authoryear,backend=biber,natbib=true]{biblatex}
 %\usepackage{natbib}
@@
 %%% --- Commenting, Todonotes
-%\usepackage[colorinlistoftodos,prependcaption]{todonotes}
+\usepackage[colorinlistoftodos,prependcaption]{todonotes}
 %\usepackage{comment, soul}
 %\newcommand{\ar}{$\rightarrow$}
-%\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
+\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
 %\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}
@@
 
+\newcommand{\xcond}{\chi}
+
 %%% --- Other Commands
@@
 
-\affil*[1]{\orgdiv{Institute for Operations Research}, \orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\street{Kaiserstr. 12}, \city{Karlsruhe}, \postcode{76131}, \country{Germany}}}
+\affil[1]{\orgdiv{Institute for Operations Research}, \orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\street{Kaiserstr. 12}, \city{Karlsruhe}, \postcode{76131}, \country{Germany}}}
 
@@
 \maketitle
-% \listoftodos
+ \listoftodos
 

### 2024-03-27 sections/01-introduction.tex
Update with comments by Grothe

@@
 The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
-Conventionally, methods are evaluated using differences between predictions (or measurements) and target values, such as the \ac{rmse} or \ac{mae}. 
-These measures do not consider whether the right direction of change is predicted or measured.
+Conventionally, differences between predictions (or measurements) and target values, such as the \ac{rmse} or \ac{mae}, form the basis of method evaluation. 
+These measures do not consider trending, that is, whether the right direction of change is predicted or measured.
 However, the information about an increase or decrease is crucial to detect changes in the underlying process early. 
@@
 
-Methodologically evolved from forecasting, nowcasting methods focus on predictions for the present, the immediate future, and the recent past \citep{banbura2013now}.
+Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past \citep{banbura2013now}.
 Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours \citep{browning1989nowcasting,schmid2019nowcasting}.
@@
 
-As outlined above, trending evaluation is crucial for measurements, predictions, and forecasts, as research interests in various disciplines reveal. 
-We apply the existing trending methods from measurement evaluation to the other fields and extend their functionality.
-The trending evaluation methods are meant to complement the current evaluation to reveal additional information on the considered methods.
-
 The main contributions of this paper are manifold.
@@
 The ready-to-use code for trending evaluation from the applications is available on ... .
+The trending evaluation methods are intended to complement the current evaluation and reveal additional information on the methods considered.
 

### 2024-03-27 sections/03-trending.tex
Update with comments by Grothe

@@
-As discussed in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by another measurement, a nowcast, or a forecast.
+As discussed in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by a measurement, a nowcast, or a forecast.
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
-The development in trending assessment always has respect to a time horizon, that is, the amount of time that has passed. 
-In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the question at hand.
-Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is insignificant.
+In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the task.
+Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is irrelevant.
 Before introducing measures to evaluate trending, we remark on calculating the predicted change and notation in Section~\ref{subsec:notation}.
@@
 When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet. 
-Thus, the predicted change can be computed with the knowledge of time $t$. Therefore, the trending information in the following section can be used in practice at time $t$.
+Through the computation in Equation~\eqref{eq:diffxl_nowcasting}, the predicted change can be computed with the knowledge of the nowcaster at time $t$.
 
@@
 yields the change by the test method. 
-It is computed purely by the test method without the gold standard $y_t$ to analyze whether the gold standard and test method changes are consistent within themselves.
+It is computed purely by the test method without the gold standard $y_t$ to analyze whether the gold standard and test method changes are consistent.
+Accordingly, $y_{t-l}$ is not used in the computation even if known at time $t$ in contrast to forecasting and nowcasting.
 
@@
 
-The trending assessment is the same for all three application areas, given the notation for the respective application of Section~\ref{subsec:notation}.
+Formally, the trending assessment is the same for all three application areas, given the notation for the respective application of Section~\ref{subsec:notation}.
 In the following, we omit the difference horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
@@
 It can be extended by including information on the time index in the point color to reveal effects over time.
-In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, $(\diffyt[l], \diffxt[l])$ is blue and turns green until $(\diffyt[T], \diffxt[T])$.
+In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, more recent values; $(\diffyt[l], \diffxt[l])$ is blue and turns green until $(\diffyt[T], \diffxt[T])$.
 However, four-quadrant plots become crowded for larger datasets, and sequential information on the differences is complex to assess thoroughly.
@@
 \end{equation}
-The numerator counts the number of same-sign-changes, while the denominator is the number of considered pairs $(\diffyt, \diffxt)$.
+Here, the numerator counts the number of same-sign-changes, while the denominator is the number of considered pairs $(\diffyt, \diffxt)$.
 Thus, $\acc$ is the proportion of concordant changes on all changes.
@@
 \end{align}
-They give the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$.
+They estimate the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$.
 
@@
 The trending ability of the signal has a strong sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
-This seasonal behavior cannot be distinguished in the two four-quadrant plots or the overall trending ratio.
+While colored four-quadrant plots, as in Figure~\ref{fig:trending_basic_4q_sample_color}, illustrate ongoing overall drifts in the trending ability, seasonal aspects are only revealed in rolling window estimates.
 
@@
 The strong seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
-The green curve $k_t$ shows the probability that $\diffxt$ has the same sign as $\diffyt$ for each time step.
-The rolling estimates are delayed as the windows look backward.}
+The green curve $k_t$ shows the theoretical probability that $\diffxt$ has the same sign as $\diffyt$ for each time step.
+The rolling estimates are shifted to the right compared to $k_t$ as the windows look backward.}
 \end{figure}
@@
 
-\subsection{Accounting for noise and non-informative small changes and bootstrap confidence intervals}\label{subsec:trending-noise}
+\subsection{Accounting for noise and non-informative small changes and bootstrapping confidence intervals}\label{subsec:trending-noise}
 
@@
 
-Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} \citep[see, e.g.,][]{Saugel2015,Critchley2010}.
+Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} accounting for such effects\citep[see, e.g.,][]{Saugel2015,Critchley2010}.
 Points within that area are omitted in the four-quadrant plot and the calculation of the measures.
@@
     \acceps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin E}},\label{eq:acceps}\\
-    \accpeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, , (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}, \ \text{and} \label{eq:accpeps}\\
+    \accpeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}, \ \text{and} \label{eq:accpeps}\\
     \accmeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}.
@@
 The additional argument $(\diffyt, \diffxt) \notin E$ in the indicator ensures that only points not lying in the exclusion are evaluated when computing the measures.
-The measures are then estimators for the probability of trending, given that the change is not in the exclusion area $E$, that is, $\acceps$ is an estimator for $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$.
+The measures are then estimators for the probability of trending, given that the point's location is not driven by noise or non-informative changes.
 
@@
 A point is only included if no component is likely to be driven by noise.
-For these two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
+For the latter two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
 
 In most applications, the shape and size of the exclusion area can be chosen based on domain knowledge or expert opinions.
-The size estimation can also be based on a proportion of the total variance or the total range of the data; for example, the 10\% smallest absolute values in each component determine the exclusion area size.
+The size determination can also be based on a proportion of the total variance or the total range of the data; for example, the 10\% smallest absolute values in each component determine the exclusion area size.
 A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
@@
 Bootstrap confidence intervals are based on resampling and not on parametric assumptions as classical confidence intervals are~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
-New samples are drawn with replacement from the dataset, and the confidence interval is then computed based on those bootstrap samples.
+Many new samples are drawn with replacement from the dataset, and the statistic of interest is computed for each sample, yielding an estimate for the distribution of the statistic of interest. 
+Based on the derived \enquote{new} samples of the statistic, the confidence intervals can be derived through differently sophisticated bootstrapping methods.
 We use the \ac{bca} approach for bootstrapping in the following, as it holds the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-trending-bootstrap}).
@@
 The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
-A still more fine and local analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ to assess the local trending ability of a prediction.
-$P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ denotes the probability of a correct trending given a predicted change of $x$. 
-A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
+A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the trending ability of a prediction for a certain change size $\diffxrv = \xcond$ of the signal.
+Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct trending given a predicted change of $\xcond$. 
+Thus, if a change of $\xcond$ is observed in practice, one can directly assess the credibility of this specific signal.
+A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
 \begin{align*}
-P(\diffxrv \diffyrv > 0 | \diffxrv = x) = \begin{cases}
-                                              \int_{-\infty}^0 \frac{f_{\diffxrv, \diffyrv}(x, y)}{f_{\diffxrv}(x)} \ \textrm{d} \: y & \text{if } x < 0, \\
-                                              \int_{0}^{\infty} \frac{f_{\diffxrv, \diffyrv}(x, y)}{f_{\diffxrv}(x)} \ \textrm{d} \: y & \text{if } x > 0, \\
+P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond) = \begin{cases}
+                                              \int_{-\infty}^0 \frac{f_{\diffxrv, \diffyrv}(\xcond, y)}{f_{\diffxrv}(\xcond)} \ \textrm{d} \: y & \text{if } \xcond < 0, \\
+                                              \int_{0}^{\infty} \frac{f_{\diffxrv, \diffyrv}(\xcond, y)}{f_{\diffxrv}(\xcond)} \ \textrm{d} \: y & \text{if } \xcond > 0, \\
 \end{cases}
 \end{align*}
-for $x \neq 0$ through a \ac{kde}.
+for $\xcond \neq 0$ through a \ac{kde}.
 A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages, for example, in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
-The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ for all values of $x \in \R$.
+The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ for all values of $\xcond \in \R$.
 Multivariate \ac{kde} takes a kernel and bandwidth selector as modeling parameters. 
@@
 
-Assessing $P(\diffxrv \diffyrv > 0 | \diffxrv = x)$ graphically eases the simultaneous evaluation of various $x$.
-Furthermore, the graph facilitates the comparison of various methods in a single graph and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = x)$ with respect to $x$ in the trending ability can be detected.
+Assessing $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ graphically by drawing $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ against $\xcond$ eases the simultaneous evaluation of various $\xcond$.
+Furthermore, the graph facilitates the comparison of various methods in a single graph, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = \xcond)$ with respect to $x$ in the trending ability can be detected.
 We refer to the plot as a \textit{conditional trending plot}.
@@
 In nowcasting and forecasting, probabilistic predictions have become more prevalent in recent years (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
-Probabilistic predictions include a point estimate and information on the prediction uncertainty and quantiles simultaneously.
+In this section, we develop trending measures for probabilistic signals.
+Probabilistic predictions issue a probability distribution for the quantity of interest based on their available information and, thus, include a point estimate and information on the prediction uncertainty and quantiles simultaneously.
 For trending analysis, probabilistic predictions thus also contain a probability of a positive or negative change.
@@
 The \ac{cdf} is the most general and can be used to derive the others, given they exist.
-Let us first assume that the prediction is a \ac{cdf}, and the horizon is large enough to know the true value when issuing the forecast or nowcast.
+Let us first assume that the prediction is a \ac{cdf}, and that $y_{t-l}$ is known at time $t$ (see Table~\ref{tab:notation}).
 In the appendix, we extend the analysis for predictions given as quantiles or unknown true values.
 
-Let $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
-For trending assessment, we compare the predicted probability of positive change with the occurrence of positive changes. 
-For the prediction, a positive change is predicted for any value at $t$ larger than the true value $y_{t-l}$. 
+Let for a forecast $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
+Thereby, $F_{t | t-l} (x)$ yields the probability that the quantity of interest is at most $x$.
+For trending assessment, we compare the predicted probability of positive change, denoted by $p_t$, with the occurrence of positive changes. 
+A positive change occurs for any value at $t$ larger than the true value $y_{t-l}$, as then the change compared to $y_{t-l}$ is positive.
+The \ac{cdf} $F_{t | t-l} (y_{t-l})$ yields the predicted probability of any value at most $y_{t-l}$, a negative change.
+Thus, $1 - F_{t | t-l} (y_{t-l})$ is the probability of a positive change.
 As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
@@
 \begin{equation*}
-    p_t = 1 - F_{t | t} (y_{t-l})\quad t = l, \dots, T;
+    p_t = 1 - F_{t | t-l} (y_{t-l})\quad t = l, \dots, T;
 \end{equation*}
-while for forecasts, it is given by
+while for nowcasts it is given by
 \begin{equation*}
-    p_t = 1 - F_{t | t-l} (y_{t-l})\quad t = l, \dots, T.
+    p_t = 1 - F_{t | t} (y_{t-l})\quad t = l, \dots, T,
 \end{equation*}
+with analogous derivations as above.
 Let $z_t$ denote the indicator that the observed change at time $t$ is positive, that is,
@@
 
-The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\citep{Brier1950} and is defined as
+The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\citep{Brier1950}.
+In our context, it is
 \begin{equation*}
@@
 \end{equation*}
-Lower values indicate a higher probabilistic trending ability of the method.
+Thereby, lower values indicate a higher probabilistic trending ability of the considered method.
 The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\citep{Ranjan2010,Mitchell2011}.
@@
 Sharpness refers to the spread of the forecast; higher probabilities are preferable as they convey a higher certainty.
-The basic paradigm of probabilistic forecasting is to maximize sharpness, subject to calibration~\citep{Gneiting2014}.
 

### 2024-03-27 sections/04-application.tex
Update plots and tables for new layout
Update with comments by Grothe

@@
-In this section, we evalute the trending in three different applications.
+In this section, we evaluate the trending in three different applications.
 In the first application in Section~\ref{sec:application-covid}, we consider the nowcasting of the seven-day hospitalization rate for COVID-19 in Germany.
@@
 In the last application in Section~\ref{sec:application_measurement}, we consider the trending assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.
+
+The examples illustrate the use of the methods of Section~\ref{sec:trending} and yield further insights into the measures and plots.  
+All code used to generate the analysis is available online.
 
@@
 In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
-For the true values, we use the data from February 8, 2024.
+We use the data from February 8, 2024, for the true values.
 The nowcasts are resolved with regard to all inhabitants and specific geographical and age breakdowns.
@@
 If hospitalization rates rise, measures should be tightened, while falling rates might allow for loosening measures.
-Especially, asymmetries are of interest to assess whether some models are better at recognizing a fall than a rise or vice versa.
+Asymmetries are especially of interest in assessing whether some models are better at recognizing a fall than a rise or vice versa.
 
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models with and without exclusion areas for the horizon seven days.
+    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days.
 The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
+The measures with the exclusion area are denoted with the subscript $q_{0.1}$.
 }
@@
 
-A more extensive training size would be beneficial for assessing the models' performance.
-For the evaluation period of 159 days, the trending ratio confidence intervals overlap; thus, no conclusions can be drawn from the trending evaluation.
+A more extensive data size would be beneficial for assessing the models' performance.
+For the evaluation period of 159 days, the trending ratio confidence intervals overlap; thus, no conclusions can be drawn from the trending evaluation comparing the models.
 
@@
 
-In a second example, we consider forecasting the hourly number of arrivals in a large emergency department.
+In a second example, we consider forecasting the number of arrivals in a large emergency department per hour.
 Good forecasts are crucial for planning staff and resources.
@@
 Every 12 hours, the models issue hourly forecasts for the next 48 hours.
-Thus, the management can take measures according to the expected number of arrivals, for example, through redeploying staff and reconfiguring units.
+Thus, management can take measures according to the expected number of arrivals, for example, by redeploying staff and reconfiguring units.
 
@@
 The trending perspective is easy for the management to understand and implement, as simple comparisons of the expected workload to a recent shift can be made.
-If, for example, the staff was near capacity in the last shift and an increase in the number of arrivals is expected, the management can take measures to adjust the workload.
+If, for example, the staff was near capacity during the last shift and an increase in the number of arrivals is expected, management can take measures to adjust the workload.
 
@@
 Note that the reported values for the RMSE differ from those in \citet{Rostami-Tabar2023}.
-In contrast to their work, we use only the forecast data at least 36 hours ahead and not the entire forecast data for evaluation.
+In contrast to their work, we use only the forecast data at least 36 hours ahead rather than the entire forecast data for evaluation.
 
@@
 
-Table~\ref{tab:app-eda-marginals} analyzes the differences in marginal distributions for the forecasts and true values for the horizons of three and seven days.
-Note that the difference definition aligns with Section~\ref{subsec:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the true value is available when issuing the forecast.
+We start by analyzing the marginal distributions for the changes of forecasts and true values for the horizons of three and seven days in Table~\ref{tab:app-eda-marginals}, again.
+Thereby, the computed difference aligns with Section~\ref{subset:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the true value is available when issuing the forecast.
 The fraction of positive differences varies between 0.39 and 0.63 for the horizon of three days and between 0.37 and 0.63 for the horizon of seven days.
@@
 The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.
+\todo{Schwer verstaendlich (Auch Abbildung~\ref{tab:app-eda-marginals})}
 
@@
     \centering
+    \tiny
     \input{plots/ed_arrival/10_marginal_analysis}
     \caption{Marginal analysis of the nowcast and true differences.
-The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
+The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.
+Column (2) shows the fraction of values not in the exclusion area of size one.}
     \label{tab:app-eda-marginals}
@@
 The confidence interval width is at most 0.02 for the trending ratios and at most 0.03 for the positive and negative trending ratios.
-The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for three and seven days horizon, while Poisson-2 and NBI-2 have the highest negative trending ratio.
+The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for the three and seven-day horizon, while Poisson-2 and NBI-2 have the highest negative trending ratio.
 
@@
 Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
-The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days are shown in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
+The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
+Remember that low a \ac{bs} and a plot along the diagonal are signs for good trending.
 The \acp{bs} are smallest for NBI-2 and Poisson-2 for both horizons, while the \acp{bs} for the other models are larger and differ more.
-The qreg-1 model has the highest \ac{bs} for both horizons.
+The qreg-1 model has both horizons' highest \ac{bs}.
 The reliability diagrams of GBM-2 and NBI-2 are also close and show a too-small fraction of increases for the predicted probability overall.
@@
     \centering
+    \tiny
     \input{plots/ed_arrival/50_trending_ratio}
@@
     \input{plots/ed_arrival/70_brier_scores}
-    \caption{Brier Scores for the different models and horizons.}\label{fig:app-eda-prob-brier}
+    \caption{BS for the different models and horizons.}\label{fig:app-eda-prob-brier}
     \end{subfigure}\hspace{0.01\textwidth}%
@@
     \end{subfigure}
-    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days.}
+    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days.
+The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic trending simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically.
+Probabilistic trending by the \ac{bs} is best for the models NBI-2 and Poisson-2 for both lags.}
     \label{fig:app-eda-prob}
@@
 The example shows that trending assessment is detached from standard point evaluation measures.
-While the models with the lowest RMSE, NBI-2 and Poisson-2, also have a high trending ability, three models with below-average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive trending ability.
+While the models with the lowest RMSE, NBI-2, and Poisson-2, also have a high trending ability, three models with below-average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive trending ability.
 
@@
 The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} simultaneously for two consecutive times with the specified horizons.
-Thus, we use the \ac{nbp} measurements as test method and the \ac{abp} measurements as gold standard.
+Thus, we use the \ac{nbp} measurements as the test method and the \ac{abp} measurements as the gold standard.
 For the systolic measurements, 290, 332, and 442 points are available for the horizons of one, five, and 15 minutes; for the mean measurements, 406, 430, and 542.
@@
 The trending ratios, including confidence intervals for the different horizons, are listed in Table~\ref{tab:app-mimic-trending-ratios}.
-For the measurements with a horizon of one minute, the confidence intervals have lower bounds of 0.5 or slightly above.
+The confidence intervals have lower bounds of 0.5 or slightly above for the measurements with a horizon of one minute.
 For larger horizons, the trending ratio increases.
@@
 The width is around 0.1 for the trending ratio, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the horizon of one minute.
-Thus, more measurements would be of interest for a further trending assessment.
+Thus, the confidence intervals cover 0.5 for systolic measurement and the horizon of one minute, and the equality to random guessing cannot be excluded.

### 2024-03-27 sections/05-conclusion.tex
Update with comments by Grothe

@@
 Models with very different accuracies are usually scored similarly in trending evaluation, but trending evaluation can differentiate between models with similar accuracies.
-As in the application in Section~\ref{sec:application-covid}, models with average point forecast evaluation measures can have the most meaningful positive trending ability.
+As in the application in Section~\ref{sec:application-covid}, models with medial point forecast evaluation measures can have the most meaningful positive trending ability.
 

### 2024-03-28 sections/01-introduction.tex
Correct some typos and add covid prob assessment

@@
 In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable. 
-Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules \citep[see the review in][]{hyndman2006another}.
+Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules~\citep[see the review in][]{hyndman2006another}.
 They are not capable of assessing the trending ability discussed above.
@@
 Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past \citep{banbura2013now}.
-Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours \citep{browning1989nowcasting,schmid2019nowcasting}.
+Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours~\citep{browning1989nowcasting,schmid2019nowcasting}.
 %In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \citep{banbura2013now}.
-In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors \citep{wu2021nowcasting} or correct the daily COVID case numbers for events that have occurred but have not yet been reported \citep{gunther2021nowcasting}.
+In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{wu2021nowcasting} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{gunther2021nowcasting}.
 Nowcasting methods use high-frequency indicators related to the target variable and estimate the value of a target variable for a specific time based on current preliminary measurements, which are finalized with a considerable time delay.
-Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period \citep{castle2017forecasting}. 
-Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse} \citep{gunther2021nowcasting} or probabilistic scoring rules \citep{Wolffram2023}.
+Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period~\citep{castle2017forecasting}.
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse}~\citep{gunther2021nowcasting} or probabilistic scoring rules~\citep{Wolffram2023}.
 The aspect of trending ability is not considered in the literature to the best of our knowledge. 
@@
 When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
-Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods \citep{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series \citep{watson2010method}.
+Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods~\citep{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series~\citep{watson2010method}.
 Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \citep{bland1986statistical}. 
@@
 The remainder of this paper is structured as follows.
-Section \ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts. 
+Section~\ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts.
 We introduce a new graphical method for evaluating trending, the conditional trending plot, which can simultaneously assess different intervals and asymmetries.
-In Section \ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting. 
+In Section~\ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting.
+Hence, we provide blueprints for applying trending evaluation in practice and make remarks on interpreting the measures.
 We conclude the paper in Section~\ref{sec:conclusion}.

### 2024-03-28 sections/03-trending.tex
Correct some typos and add covid prob assessment

@@
 When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet.
+If the true value $y_{t-l}$ is known at time $t$, the predicted change is computed by the difference between the nowcast and the true value, as $y_{t-l}$ is also known by the nowcaster and incorporated into the nowcast.
 Through the computation in Equation~\eqref{eq:diffxl_nowcasting}, the predicted change can be computed with the knowledge of the nowcaster at time $t$.
@@
 Accordingly, $y_{t-l}$ is not used in the computation even if known at time $t$ in contrast to forecasting and nowcasting.
+
+In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
+We refer to time steps for which either signal or true values, or both, are unavailable as missing values.
+Systematical missing values could lead to a biased estimate of the trending ratio, and missing data should be inspected for any underlying patterns.
+If the missing values are not systematic, random, and occur scarcely, data pairs with missing values can be excluded from the data to calculate the measures~\parencite[see][Section 1.3]{VanBuuren2018}.
+Note that in the case of measurement data, one missing value in the time series leads to two undefined differences in the change series; that is if $x_t$ is missing, $\diffxt$ and $\diffxt[t+l]$ are undefined; if an observation $y_t$ is missing, $\diffyt$ and $\diffyt[t+l]$ are undefined.
+The data pair is excluded even if the corresponding nowcast or forecast is available.
 
@@
 
-In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
-We refer to time steps for which either signal or true values, or both, are unavailable as missing values.
-Data pairs with missing values can be excluded from the set $\mathcal{T}$ to calculate the measures. 
-However, systematical missing values could lead to a biased estimate of the trending ratio, and missing data should be inspected for any underlying patterns.
-
 \subsection{Accounting for noise and non-informative small changes and bootstrapping confidence intervals}\label{subsec:trending-noise}
@@
 Let us first assume that the prediction is a \ac{cdf}, and that $y_{t-l}$ is known at time $t$ (see Table~\ref{tab:notation}).
-In the appendix, we extend the analysis for predictions given as quantiles or unknown true values.
+In Appendix~\ref{subsec:app-probabilistic-trending-evaluation}, we extend the analysis for predictions given as quantiles or unknown true values.
 

### 2024-03-28 sections/04-application.tex
Correct some typos and add covid prob assessment

@@
 For information on the model design, we refer to~\citet{Wolffram2023}.
-Using the nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
+In addition to those nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
 We denote them by ENS-MEAN and ENS-MED.
@@
 If hospitalization rates rise, measures should be tightened, while falling rates might allow for loosening measures.
-Asymmetries are especially of interest in assessing whether some models are better at recognizing a fall than a rise or vice versa.
+Asymmetries are especially relevant for assessing whether some models are better at recognizing a fall than a rise or vice versa.
 
@@
 
-Before stating the results, we provide background information on the marginal distributions of the true values and nowcasts for the different horizons.
+Before stating the results, we provide background information on the marginal distributions of the actual values and nowcasts for the different horizons.
 Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} presents marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
@@
 Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.
+
+\begin{table}
+    \centering
+    \tiny
+    \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
+    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days.
+The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
+The subscript $q_{0.1}$ denotes the measures with exclusion area.
+}
+    \label{tab:app-covid-trending-ratios-lag-7}
+\end{table}
 
@@
 For the horizons of one and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
+
+\begin{figure}
+    \centering
+%    \includegraphics{}
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
+    \caption{Conditional trending plot.}\label{fig:app-covid-cond-prob-7}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
+    \caption{Trending ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-trending-ratio-7}
+    \end{subfigure}
+    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
+    \label{fig:app-covid-cond-prob-trending-ratio-7}
+\end{figure}
 
@@
 
-\begin{table}
+Figure~\ref{fig:app-covid-prob} shows the \acf{bs} and reliability diagrams for the same subset of models, the ILM, RKI, RIVM, and ENS-MED.
+The probabilities of increase for the different models are computed using the nowcast quantiles.
+For each horizon $l$, $10,000$ samples of the forecast date $t$ and the forecast date $t-l$ based on the nowcasts of issue date $t$ are generated, and the proportion of positive changes is computed (see Appendix~\ref{subsec:app-probabilistic-trending-evaluation}).
+Remember that low a \ac{bs} and a plot along the diagonal are signs of good trending.
+The \ac{bs} is the lowest for the RIVM model, while the ENS-MED model has the best \ac{bs} for the horizon of seven and 14 days.
+The RKI model yields the highest \ac{bs} for all horizons.
+Note that the \ac{bs} is bounded by 0.25 for random guessing; thus, all models perform better than random guessing.
+The reliability diagrams show that the models are not well calibrated for the horizon of one day.
+While for predicted probabilities below 0.5, the observed ratio of increases is smaller, it is higher than predicted for probabilities above 0.5.
+Figure~\ref{fig:app-covid-prob-hist} shows the observed predicted increase probabilities for all horizons.
+For the horizons of seven and 14 days, the nowcasters issue only a few moderate probabilities.
+
+\begin{figure}
+    \begin{subfigure}{0.48\textwidth}
         \centering
         \tiny
-    \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days.
-The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
-The measures with the exclusion area are denoted with the subscript $q_{0.1}$.
-}
-    \label{tab:app-covid-trending-ratios-lag-7}
-\end{table}
-
-\begin{figure}
-    \centering
-%    \includegraphics{}
-    \begin{subfigure}[t]{.48\textwidth}
-    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
-    \caption{Conditional trending plot.}\label{fig:app-covid-cond-prob-7}
-    \end{subfigure}\hfill
-    \begin{subfigure}[t]{.48\textwidth}
-    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
-    \caption{Trending ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-trending-ratio-7}
-    \end{subfigure}
-    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
-    \label{fig:app-covid-cond-prob-trending-ratio-7}
+        \input{plots/covid_nowcast/70_brier_scores}
+    \caption{BS for the different models and horizons.}\label{fig:app-covid-prob-brier}
+    \end{subfigure}\hspace{0.01\textwidth}%
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_1}
+    \caption{Reliability diagram for horizon one day.}\label{fig:app-covid-prob-rel-1}
+    \end{subfigure}\hspace{0.01\textwidth}%
+    \caption[Brier Scores and reliability diagrams for ILM, RKI, RIVM, and ENS-MED.]{\aclp{bs} and reliability diagrams for ILM, RKI, RIVM, and ENS-MED.
+    The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
+    In the computation of BS and reliability diagram, missing values are excluded.
+    The reliability diagram for the horizon seven and 14 days is in the appendix (see Figure~\ref{fig:app-covid-rel}).}
+    \label{fig:app-covid-prob}
 \end{figure}
@@
 Figure~\ref{fig:app-covid-trending-ratio-7} shows that the trending ratio increases with larger exclusion areas.
-This indicates that if the model predicts a large change, the direction is indeed better than when a small change is predicted.
+This indicates that if the model predicts a large change, the direction is more accurate than when a small change is predicted.
+
+The probabilistic trending assessment shows that the models are better than random guessing.
+In general, the reliability diagram cannot provide information if specific probabilities are issued scarcely.
+Thus, the reliability diagrams for the horizons of 7 and 14 days do not contain information on moderate probabilities.
+The \ac{bs} values, however, work well for those examples and provide a good measure for the trending ability of the models.
 
@@
 Good forecasts are crucial for planning staff and resources.
-Several models are used for predicting hourly outcomes in a study by \citet{Rostami-Tabar2023}.
+In a study by \citet{Rostami-Tabar2023}, several models predict hourly outcomes.
 Every 12 hours, the models issue hourly forecasts for the next 48 hours.
@@
 
-We start by analyzing the marginal distributions for the changes of forecasts and true values for the horizons of three and seven days in Table~\ref{tab:app-eda-marginals}, again.
-Thereby, the computed difference aligns with Section~\ref{subset:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the true value is available when issuing the forecast.
+We start by analyzing the marginal distributions for the changes of forecasts and actual values for the horizons of three and seven days in Table~\ref{tab:app-eda-marginals}, again.
+Thereby, the computed difference aligns with Section~\ref{subset:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the actual value is available when issuing the forecast.
 The fraction of positive differences varies between 0.39 and 0.63 for the horizon of three days and between 0.37 and 0.63 for the horizon of seven days.
@@
 The 10\%-quantile of the differences is between zero and one for all models and horizons.
-Thus, we exclude only differences smaller than one from the trending assessment.
+Thus, we exclude differences smaller than one from the trending assessment.
 The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.
-\todo{Schwer verstaendlich (Auch Abbildung~\ref{tab:app-eda-marginals})}
 
@@
 Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
-The conditional trending plots show similar courses for the two horizons, while the curves are shifted downwards for the horizon of seven days.
+The conditional trending plots show similar courses for the two horizons, though the curves are shifted downwards for the horizon of seven days.
 The model's relative trending ability evolves consistently for the two horizons, with the NBI-2 and Poisson-2 models being indistinguishable.
@@
 The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
-Remember that low a \ac{bs} and a plot along the diagonal are signs for good trending.
 The \acp{bs} are smallest for NBI-2 and Poisson-2 for both horizons, while the \acp{bs} for the other models are larger and differ more.
@@
 Whether these points are due to measurement errors or extreme values is not distinguishable.
-Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile \citep[see][]{Critchley2010}.
+Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile~\citep[see][]{Critchley2010}.
 We do not follow this approach here, as the extreme values are not necessarily measurement errors and could be particularly relevant.

### 2024-03-28 sections/05-conclusion.tex
Correct some typos and add covid prob assessment

@@
 Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
-The positive and negative trending ratio analyze the trending ratio given the predicted change is positive or negative, respectively.
+The positive and negative trending ratios analyze the trending ratio given whether the predicted change is positive or negative, respectively.
 Thus, they quantify the credibility of the respective predictions.
@@
 The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
-For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or assess models with high trending differences.
+For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or to assess models with high trending differences.
 

### 2024-03-28 sections/50-appendix-data-generation.tex
Fix layout and renew some plots
Correct some typos and add covid prob assessment

@@
     \centering
+    \tiny
     \begin{subtable}{.48\textwidth}
@@
     \begin{subfigure}{.48\textwidth}
-        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_asym_butterfly}
+        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_butterfly}
         \caption{First dataset with asymmetric dependence.}
@@
 
-\subsection{Probabilistic trending evaluation}
+\subsection{Probabilistic trending evaluation}\label{subsec:app-probabilistic-trending-evaluation}
 
@@
 If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to generate samples for the Monte Carlo approximation.
+This approach is applied in Section~\ref{sec:application-covid}, as the true values are published with a delay of more than 80 days, and the nowcasts are given as quantiles.

### 2024-03-28 sections/55-appendix-application-covid.tex
Merge changes; add bmc template
Fix layout and renew some plots
Correct some typos and add covid prob assessment

@@
     \caption{Matching the abbreviation to the key in the nowcasting hub.
-    Information on the models and references is listed in \textcite[][Table 1]{Wolffram2023}.}
+    Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
     \label{tab:app-covid-models}
@@
     \centering
+    \tiny
     \input{plots/covid_nowcast/10_marginal_analysis}
@@
 \end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{.48\textwidth}
+        \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_7}
+        \caption{Reliability diagram for horizon seven days.} \label{fig:app-covid-reliability-7}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+        \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_14}
+        \caption{Reliability diagram for horizon 14 days.} \label{fig:app-covid-reliability-14}
+    \end{subfigure}
+    \begin{subfigure}{\textwidth}
+        \includegraphics{plots/covid_nowcast/70_prob_hist}
+        \caption{Count histogram of the predicted probabilities for the horizon one, seven, and 14 days.} \label{fig:app-covid-prob-hist}
+    \end{subfigure}
+    \caption{The reliability diagram for the models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
+    Additionally, the count of predicted probabilities for the horizons is shown.
+        The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
+    As the models issue small or large probabilities of increase for the higher horizons, little information on the accuracy of moderate probability predictions is available.}
+    \label{fig:}
+\end{figure}

### 2024-03-28 paper.tex
Add declarations

@@
 
+\section{Declarations}
+\subsection{Availability of data and materials}
+
+The datasets analysed during the study are available in the corresponding databases or repositories \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024,Rostami-Tabar2023, Moody2017}.
+All code is available in the repository [GITHUB LINK WILL BE PROVIDED AFTER REVIEW].
+
+\subsection{Competing interests}
+The authors declare that they have no competing interests.
+
+\subsection{Funding}
+
+JR greatfully acknowledges funding by the Bischoefliche Studienstiftung Cusanuswerk (Bonn, Germany).
+
+\subsection{Authors' contributions}
+
+OG, BL and JR developed the theoretical formalism.
+BL and JR implemented the methods.
+JR conducted the simulations and analyzed the data.
+BL and JR wrote the manuscript.
+All authors discussed the results, critically reviewed and contributed to the final manuscript.
+
+\subsection{Acknowledgments}
+
+Not applicable.
+
 \printbibliography

### 2024-03-29 paper.tex
Move everything to one file

@@
 \author[1]{Oliver Grothe}
-\equalcont{These authors contributed equally to this work.}
 \author[1]{Bolin Liu}
-\equalcont{These authors contributed equally to this work.}
 \author*[1]{Jonas Rieger}
-\equalcont{These authors contributed equally to this work.}
 
@@
 \section{Introduction}\label{sec:introduction}
-\input{sections/01-introduction}
+The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
+Conventionally, differences between predictions (or measurements) and target values, such as the \ac{rmse} or \ac{mae}, form the basis of method evaluation.
+These measures do not consider trending, that is, whether the right direction of change is predicted or measured.
+However, the information about an increase or decrease is crucial to detect changes in the underlying process early.
+In the following, we provide a more specific overview of the characteristics and current evaluation schemes of the three application fields, forecasting, nowcasting, and measurement, and highlight why trending evaluation is highly relevant in the respective fields.
 
-%\section{Application areas and Notation} \label{sec:notation}
-%\input{sections/02-notation}
+Forecasting methods predict the future based on historical data, patterns, and exogenous factors.
+In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008,Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
+The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
+In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable.
+Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules~\citep[see the review in][]{hyndman2006another}.
+They are not capable of assessing the trending ability discussed above.
+In a prediction competition on armed conflicts, trending recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with a trending component for evaluation.
+However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
+This paper focuses on pure trending evaluation that accompanies traditional scores.
+
+Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past \citep{banbura2013now}.
+Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours~\citep{browning1989nowcasting,schmid2019nowcasting}.
+%In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \citep{banbura2013now}.
+In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{wu2021nowcasting} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{gunther2021nowcasting}.
+Nowcasting methods use high-frequency indicators related to the target variable and estimate the value of a target variable for a specific time based on current preliminary measurements, which are finalized with a considerable time delay.
+Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period~\citep{castle2017forecasting}.
+Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse}~\citep{gunther2021nowcasting} or probabilistic scoring rules~\citep{Wolffram2023}.
+The aspect of trending ability is not considered in the literature to the best of our knowledge.
+However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
+The epidemic's development, in turn, can be the foundation of decisions on introducing or canceling policy measures.
+
+Measurement is more technical and aims to obtain accurate and reliable data about the current state of a system based on sensors.
+A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
+When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
+Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods~\citep{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series~\citep{watson2010method}.
+Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \citep{bland1986statistical}.
+Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years \citep{Saugel2015,saugel2018error,hiraishi2021concordance}.
+
+The main contributions of this paper are manifold.
+We formalize trending and present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
+We introduce the conditional trending plot, a new graphical method for assessing local trending behavior, and review bootstrap methods for calculating confidence intervals.
+We extend the concept of trending to probabilistic predictions.
+The methods are applied to three applications from the abovementioned fields: COVID-19-nowcasting, emergency department arrival forecasting, and blood pressure measurement.
+The ready-to-use code for trending evaluation from the applications is available on ... .
+The trending evaluation methods are intended to complement the current evaluation and reveal additional information on the methods considered.
+
+The remainder of this paper is structured as follows.
+Section~\ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts.
+We introduce a new graphical method for evaluating trending, the conditional trending plot, which can simultaneously assess different intervals and asymmetries.
+In Section~\ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting.
+Hence, we provide blueprints for applying trending evaluation in practice and make remarks on interpreting the measures.
+We conclude the paper in Section~\ref{sec:conclusion}.
 
 \section{Trending assessment}\label{sec:trending}
-\input{sections/03-trending}
 
-%\section{Simulation Studies (?)}
-%
-%Eher nicht, weil angewandtes Paper
+As discussed in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by a measurement, a nowcast, or a forecast.
+In this section, we want to specify trending and define different methods of measuring and visualizing trending.
+In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the task.
+Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is irrelevant.
+Before introducing measures to evaluate trending, we remark on calculating the predicted change and notation in Section~\ref{subsec:notation}.
+Section~\ref{subsec:trending-four-quadrant-plot} introduces trending basics and visualizes trending in a synthetic example.
+In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
+The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise}, and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
+Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment.
+Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:trending-probabilistic}.
+
+\subsection{Computing changes and notation}\label{subsec:notation}
+
+As noted above, trending refers to observed and predicted changes over a time horizon $l$.
+The \textit{observed} change is straightforward to compute for all types of signals.
+Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values for nowcasting or forecasting, or gold standard measurements up to time $T$.
+The sequence of observed changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is,
+\begin{equation}\label{eq:diffy}
+    \diffylt = (y_{t} - y_{t-l}) \quad \text{for}\ t = l, \dots, T.
+\end{equation}
+
+\begin{table}
+    \centering
+    \scriptsize
+    \begin{tabular}{l l}
+        \toprule
+        Application & Predicted change computation \\
+        \midrule
+        Measurement & $(x_{t} - x_{t-l})_{t=l}^T $\\
+        Nowcasting & $
+\diffxl =
+\begin{cases}
+(x_{t|t} - x_{t-l|t})^T_{t=l}, & \text{if } y_{t-l} \text{ is not known at time } t, \\
+(x_{t|t} - y_{t-l})^T_{t=l}, & \text{otherwise}.
+\end{cases}
+$ \\
+        Forecasting & $ \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l}$\\
+        \bottomrule
+    \end{tabular}
+    \caption{Computation of the predicted change in the different applications.
+For nowcasting and forecasting, signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
+For measurement, $x_t$ denotes the test device measurement at time $t$.}
+    \label{tab:notation}
+\end{table}
+
+The definition of the \textit{predicted} change depends on the context;
+Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
+For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$.
+We call $t$ the \textit{forecast time} and $\tau$ the \textit{issue time}.
+The predicted change is computed by
+\begin{equation}
+  \diffxl =
+\begin{cases}
+(x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
+(x_{t|t} - y_{t-l})^T_{t=l} & \text{otherwise}.
+\end{cases} \label{eq:diffxl_nowcasting}
+\end{equation}
+When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet.
+If the true value $y_{t-l}$ is known at time $t$, the predicted change is computed by the difference between the nowcast and the true value, as $y_{t-l}$ is also known by the nowcaster and incorporated into the nowcast.
+Through the computation in Equation~\eqref{eq:diffxl_nowcasting}, the predicted change can be computed with the knowledge of the nowcaster at time $t$.
+
+The notation is similar for forecasting: Let $x_{t | \tau}$ denote the forecast for forecast time $t$ and issue time $\tau$.
+The predicted change is computed by
+\begin{equation} \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l}\label{eq:diffxl_forecasting}
+\end{equation}
+with the same structure as in the nowcasting case.
+The indices are chosen to be consistent with $\diffyl$.
+If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification as in Equation~\eqref{eq:diffxl_nowcasting} can be made.
+
+The distinction between forecast and issue time is unnecessary in measurement analysis, as the measurement is typically available with a very short time lag.
+Thus, $x_t$ denotes the test method measurement for time $t$.
+The computation
+\begin{equation}
+  (x_{t} - x_{t-l})_{t=l}^T \label{eq:diffxl_measure}
+\end{equation}
+yields the change by the test method.
+It is computed purely by the test method without the gold standard $y_t$ to analyze whether the gold standard and test method changes are consistent.
+Accordingly, $y_{t-l}$ is not used in the computation even if known at time $t$ in contrast to forecasting and nowcasting.
+
+In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
+We refer to time steps for which either signal or true values, or both, are unavailable as missing values.
+Systematical missing values could lead to a biased estimate of the trending ratio, and missing data should be inspected for any underlying patterns.
+If the missing values are not systematic, random, and occur scarcely, data pairs with missing values can be excluded from the data to calculate the measures~\parencite[see][Section 1.3]{VanBuuren2018}.
+Note that in the case of measurement data, one missing value in the time series leads to two undefined differences in the change series; that is if $x_t$ is missing, $\diffxt$ and $\diffxt[t+l]$ are undefined; if an observation $y_t$ is missing, $\diffyt$ and $\diffyt[t+l]$ are undefined.
+The data pair is excluded even if the corresponding nowcast or forecast is available.
+
+\subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-four-quadrant-plot}
+
+Formally, the trending assessment is the same for all three application areas, given the notation for the respective application of Section~\ref{subsec:notation}.
+In the following, we omit the difference horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
+As noted above, a signal's trending is perfect if all the change directions are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+Consequently, when evaluating the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
+A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{perrino1998intraoperative,Saugel2015}.
+In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
+Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
+Points in the green upper right and lower left quadrants reflect a correct trending for the respective time step, whereas points in the remaining red quadrants show incorrect predicted changes.
+
+Figure~\ref{fig:trending_basic_4q} displays a basic four-quadrant plot.
+Points 2, 3, 5, and 6 show trending, whereas points 1, 4, and 7 count for anti-trending behavior.
+Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant graph for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
+Data generation is described in Appendix~\ref{subsec:app-trending-data-generation}.
+The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
+
+The four-quadrant plot is intuitive to interpret, and the magnitude and direction of change are shown simultaneously.
+It can be extended by including information on the time index in the point color to reveal effects over time.
+In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, more recent values; $(\diffyt[l], \diffxt[l])$ is blue and turns green until $(\diffyt[T], \diffxt[T])$.
+However, four-quadrant plots become crowded for larger datasets, and sequential information on the differences is complex to assess thoroughly.
+
+There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.
+However, they lack the four-quadrant plot's clarity and intuition without adding more information on trending~\citep{Saugel2015}.
+
+\begin{figure}
+\centering
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4Q_without_excl}
+\caption{Basic four-quadrant plot.} \label{fig:trending_basic_4q}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4q_excl_box}
+\caption{Four-quadrant plot with rectangular exclusion area.}\label{fig:trending_basic_4q_excl_box}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4q_excl_axis}
+\caption{Four-quadrant plot with horizontal exclusion area.} \label{fig:trending_basic_4q_excl_axis}
+\end{subfigure}\hspace{0.01\textwidth}%
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4q_excl_cross}
+\caption{Four-quadrant plot with cross-shaped exclusion area.}\label{fig:trending_basic_4q_excl_cross}
+\end{subfigure}%
+\caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas. }
+\label{fig:trending_4q}
+\end{figure}
+
+\subsection{Trending ratio and other measures}\label{subsec:trending-measures}
+
+Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data \citep{Critchley2010,Saugel2015}.
+With that, we mainly estimate the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
+Since $z_1 z_2 > 0$ imposes the same condition as $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$), the standard estimator for $P(\diffxrv \diffyrv > 0)$ is
+\begin{equation}
+    \acc (\diffx, \diffy) \coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0}}{\card{T}}.\label{eq:acc}
+\end{equation}
+Here, the numerator counts the number of same-sign-changes, while the denominator is the number of considered pairs $(\diffyt, \diffxt)$.
+Thus, $\acc$ is the proportion of concordant changes on all changes.
+We refer to this estimator as the \textit{trending ratio} of the prediction and set $\mathcal{T} = \{l, \dots, T\}$.
+Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
+Similar evaluations are used in other scientific areas, for example, with contingency tables in dichotomous forecasting or with confusion matrix in classification analysis~\citetext{see, e.g., the introductions in \citealt{James2021}, Ch.
+4, and \citealp{Jolliffe2012}, Ch.
+3}.
+Many other measures can be adapted from those fields to deepen the trending analysis.
+Two simple measures that focus on either a positive or negative predicted change are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
+They are defined as
+\begin{align}
+    \accp (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0}},\ \text{and} \label{eq:accp}\\
+    \accm (\diffx, \diffy) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0}}.\label{eq:accm}
+\end{align}
+They estimate the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$.
+
+Rolling estimates of the above measures detect changes in performance over time and can give a sharper estimate of the current trending ability.
+For the trending ratio, a rolling estimate with a backward-looking window of length $w$ at time $t$ is given by
+\begin{equation*}
+    \acc_{t; w} (\diffx, \diffy) \coloneqq \frac{\sum_{t^\star = t-w + 1}^{t} \ind{\diffxt[t^\star] \diffyt[t^\star] > 0}}{w}.\label{eq:acc_rolling}
+\end{equation*}
+Backward-looking windows estimate the trending ratio at a time $t$ considering the $w$ time steps before time $t$.
+The window length $w$ controls the smoothing of the estimate; a larger $w$ gives smoother results, while a small $w$ focuses on local variations.
+Plotting the rolling estimates for $t = w-1, \dots, T$ yields an estimate of the trending ability over time.
+Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
+Thus, the yearly course of the prediction trending ratio can be detected.
+The trending ability of the signal has a strong sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
+While colored four-quadrant plots, as in Figure~\ref{fig:trending_basic_4q_sample_color}, illustrate ongoing overall drifts in the trending ability, seasonal aspects are only revealed in rolling window estimates.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4Q_sample_without_time}
+\caption{Four-quadrant plot with simulated data.}\label{fig:trending_basic_4q_sample}
+\end{subfigure}\hspace{0.01\textwidth}
+\begin{subfigure}[t]{.24\textwidth}
+\includegraphics{plots/illustrative_examples/4Q_sample_with_time}
+\caption{The same data is colored according to the time index $t$, the greener, the later.}\label{fig:trending_basic_4q_sample_color}
+\end{subfigure}\hspace{0.01\textwidth}
+\begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/illustrative_examples/trending_ratio_time_series.pdf}
+    \caption{Rolling estimate of the trending ratio over time with window length 91. }\label{fig:trending_ratio_time_series}
+    \end{subfigure}%
+    \caption{Visualizations for data with a time-varying trending ratio.
+We defer information on the data generation process to the appendix (see \ref{subsec:app-trending-data-generation}).
+The trending ratio for the entire data set is $\mu = 0.7577$.
+The strong seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
+The green curve $k_t$ shows the theoretical probability that $\diffxt$ has the same sign as $\diffyt$ for each time step.
+The rolling estimates are shifted to the right compared to $k_t$ as the windows look backward.}
+\end{figure}
+
+\subsection{Accounting for noise and non-informative small changes and bootstrapping confidence intervals}\label{subsec:trending-noise}
+
+The above measures can be extended to account for information on the point's location within the quadrant.
+For example, points close to the zero point may have less explanatory power or may be less reliable than points far away from zero on one of the diagonals.
+Suppose noise or non-systematic effects are present in the true values or predictions.
+In that case, a point's assignment to a quadrant can be driven by noise instead of a systematic trending ability of the signal.
+This is more likely for points with at least one small coordinate.
+
+Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} accounting for such effects\citep[see, e.g.,][]{Saugel2015,Critchley2010}.
+Points within that area are omitted in the four-quadrant plot and the calculation of the measures.
+In particular, the prediction models are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
+We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $E$ by
+\begin{align}
+    \acceps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin E}},\label{eq:acceps}\\
+    \accpeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}, \ \text{and} \label{eq:accpeps}\\
+    \accmeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}.
+\label{eq:accmeps}
+\end{align}
+The additional argument $(\diffyt, \diffxt) \notin E$ in the indicator ensures that only points not lying in the exclusion are evaluated when computing the measures.
+The measures are then estimators for the probability of trending, given that the point's location is not driven by noise or non-informative changes.
+
+The estimators accept various shapes of the exclusion area.
+Figure~\ref{fig:trending_4q} visualizes examples of exclusion area shapes.
+A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ ($\varepsilon_x, \varepsilon_y > 0$), leaves out points that are small in both components and, therefore, likely to be driven by noise.
+Points are included if the true value or signal is unlikely to be noise-driven.
+In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
+An exclusion area along one axis, for example, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
+This particularly suits signals where small amounts of noise are inevitable.
+A cross-shaped exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
+A point is only included if no component is likely to be driven by noise.
+For the latter two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
+
+In most applications, the shape and size of the exclusion area can be chosen based on domain knowledge or expert opinions.
+The size determination can also be based on a proportion of the total variance or the total range of the data; for example, the 10\% smallest absolute values in each component determine the exclusion area size.
+A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
+For examples of such plots, see Section~\ref{sec:application-covid}.
+
+Confidence intervals can account for the estimation uncertainty of the measures above.
+Bootstrap confidence intervals are based on resampling and not on parametric assumptions as classical confidence intervals are~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
+Many new samples are drawn with replacement from the dataset, and the statistic of interest is computed for each sample, yielding an estimate for the distribution of the statistic of interest.
+Based on the derived \enquote{new} samples of the statistic, the confidence intervals can be derived through differently sophisticated bootstrapping methods.
+We use the \ac{bca} approach for bootstrapping in the following, as it holds the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-trending-bootstrap}).
+
+\subsection{The conditional trending plot}\label{subsec:trending-cond-prob}
+
+The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
+A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the trending ability of a prediction for a certain change size $\diffxrv = \xcond$ of the signal.
+Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct trending given a predicted change of $\xcond$.
+Thus, if a change of $\xcond$ is observed in practice, one can directly assess the credibility of this specific signal.
+A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
+\begin{align*}
+P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond) = \begin{cases}
+                                              \int_{-\infty}^0 \frac{f_{\diffxrv, \diffyrv}(\xcond, y)}{f_{\diffxrv}(\xcond)} \ \textrm{d} \: y & \text{if } \xcond < 0, \\
+                                              \int_{0}^{\infty} \frac{f_{\diffxrv, \diffyrv}(\xcond, y)}{f_{\diffxrv}(\xcond)} \ \textrm{d} \: y & \text{if } \xcond > 0, \\
+\end{cases}
+\end{align*}
+for $\xcond \neq 0$ through a \ac{kde}.
+A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages, for example, in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
+The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ for all values of $\xcond \in \R$.
+Multivariate \ac{kde} takes a kernel and bandwidth selector as modeling parameters.
+We advise using a Gaussian kernel and the cross-validation maximum likelihood as bandwidth selector (see Appendix~\ref{subsec:appendix-kde}).
+
+Assessing $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ graphically by drawing $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ against $\xcond$ eases the simultaneous evaluation of various $\xcond$.
+Furthermore, the graph facilitates the comparison of various methods in a single graph, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = \xcond)$ with respect to $x$ in the trending ability can be detected.
+We refer to the plot as a \textit{conditional trending plot}.
+
+\subsection{Probabilistic Evaluation}\label{subsec:trending-probabilistic}
+
+In nowcasting and forecasting, probabilistic predictions have become more prevalent in recent years (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
+In this section, we develop trending measures for probabilistic signals.
+Probabilistic predictions issue a probability distribution for the quantity of interest based on their available information and, thus, include a point estimate and information on the prediction uncertainty and quantiles simultaneously.
+For trending analysis, probabilistic predictions thus also contain a probability of a positive or negative change.
+In this section, we extend the trending assessment to probabilistic predictions.
+
+Probabilistic predictions can take the form of a \ac{cdf}, \ac{pdf}, or quantiles.
+The \ac{cdf} is the most general and can be used to derive the others, given they exist.
+Let us first assume that the prediction is a \ac{cdf}, and that $y_{t-l}$ is known at time $t$ (see Table~\ref{tab:notation}).
+In Appendix~\ref{subsec:app-probabilistic-trending-evaluation}, we extend the analysis for predictions given as quantiles or unknown true values.
+
+Let for a forecast $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
+Thereby, $F_{t | t-l} (x)$ yields the probability that the quantity of interest is at most $x$.
+For trending assessment, we compare the predicted probability of positive change, denoted by $p_t$, with the occurrence of positive changes.
+A positive change occurs for any value at $t$ larger than the true value $y_{t-l}$, as then the change compared to $y_{t-l}$ is positive.
+The \ac{cdf} $F_{t | t-l} (y_{t-l})$ yields the predicted probability of any value at most $y_{t-l}$, a negative change.
+Thus, $1 - F_{t | t-l} (y_{t-l})$ is the probability of a positive change.
+As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
+For nowcasts, the probability is calculated by
+\begin{equation*}
+    p_t = 1 - F_{t | t-l} (y_{t-l})\quad t = l, \dots, T;
+\end{equation*}
+while for nowcasts it is given by
+\begin{equation*}
+    p_t = 1 - F_{t | t} (y_{t-l})\quad t = l, \dots, T,
+\end{equation*}
+with analogous derivations as above.
+Let $z_t$ denote the indicator that the observed change at time $t$ is positive, that is,
+\begin{equation*}
+    z_t = \ind{\diffyt > 0} \quad t = l, \dots, T.
+\end{equation*}
+The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=l}^{T}$ and the observed changes $\mathbf{z} = (z_t)_{t=l}^{T}$.
+The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
+Dichotomous forecasts predict a binary outcome, such as a positive or negative change.
+Numerical measures for forecast quality are scoring rules.
+
+The \ac{bs} is a widely used scoring rule for dichotomous probabilistic forecasts~\citep{Brier1950}.
+In our context, it is
+\begin{equation*}
+    BS (\mathbf{p}, \mathbf{z}) = \frac{1}{T-l+1} \sum_{t=l}^{T} (p_t - z_t)^2.
+\end{equation*}
+Thereby, lower values indicate a higher probabilistic trending ability of the considered method.
+The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\citep{Ranjan2010,Mitchell2011}.
+Calibration refers to the statistical consistency of forecasts and observations; that is, the event occurs with the issued probability.
+Sharpness refers to the spread of the forecast; higher probabilities are preferable as they convey a higher certainty.
+
+Graphical methods are a standard tool for evaluating the calibration of probabilistic forecasts in detail.
+In dichotomous forecasting, the reliability diagram is frequently used~\citep{Ranjan2010}.
+The reliability diagram plots the observed frequency of the positive outcome against the (binned) predicted probability.
+For example, it shows the proportion of observed increases, given the predicted probability of increase was approximately $0.7$.
+Ideally, the predicted probability equals the observed frequency, and the reliability diagram is a 45-degree line.
+Local deviations from the 45-degree line indicate a miscalibration of the forecast for specific probabilities.
+Thus, the reliability visualizes the local and overall calibration simultaneously.
+For an example of a reliability diagram, see Section~\ref{sec:application-eda}.
 
 \section{Application to measurement, nowcasting, and forecasting data} \label{sec:application}
-\input{sections/04-application}
+In this section, we evaluate the trending in three different applications.
+In the first application in Section~\ref{sec:application-covid}, we consider the nowcasting of the seven-day hospitalization rate for COVID-19 in Germany.
+Section~\ref{sec:application-eda} applies the trending assessment to forecast methods for the number of arrivals in a large emergency department.
+In the last application in Section~\ref{sec:application_measurement}, we consider the trending assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.
+
+The examples illustrate the use of the methods of Section~\ref{sec:trending} and yield further insights into the measures and plots.
+All code used to generate the analysis is available online.
+
+\subsection{Covid nowcasting} \label{sec:application-covid}
+
+Amid the COVID-19 pandemic, the importance of accurate and prompt nowcasts of the pandemic's progression has become evident.
+Various indicators measured the pandemic's spread and severity.
+In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\citep{RobertKochInstitute2021}.
+The Robert Koch Institute (RKI) provided preliminary daily reports on the seven-day hospitalization rate.
+However, these reports were subject to severe delays and revisions in two sources.
+The first source is technical delays in the reporting process, for example, due to different authorities passing the data to the RKI~\citep{RobertKochInstitute2024}.
+The second, more systematic source is the structure of the seven-day hospitalization rate.
+To a given date, all the cases are allocated whose first positive test result was on that date and who were hospitalized in relation to the disease in the following.
+The seven-day hospitalization rate is the average number of those cases per $100,000$ inhabitants on the given date and six days before.
+Thus, the final seven-day hospitalization rate can only be reported with a significant delay of more than 70 days, as the hospitalization of infected inhabitants can occur much later than the first positive test.
+Nevertheless, as the seven-day hospitalization rate was considered a major indicator of the pandemic's development, many organizations and institutions started to issue nowcasts, including research teams and newspapers.
+To collect nowcasts of the seven-day hospitalization rate by different nowcast groups, the COVID19-Nowcasting-Hub was established~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+The nowcasts contain the seven-day hospitalization rate's predictive mean, median, and other quantiles.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/covid_nowcast/00_true_data.pdf}
+    \caption{Realisations.}
+    \label{fig:app-covid-true}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
+    \caption{Same-day nowcasts.}
+    \label{fig:app-covid-nowcast}
+        \end{subfigure}
+    \caption{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+    The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
+    \label{fig:app-covid-true-nowcast}
+\end{figure}
+
+The data contains eight nowcasts from scientific and public institutions, nowcast communities, and a newspaper, using different input variables, calendar data, and length of training data.
+The model structures are diverse, including Bayesian models, generalized additive models, and parametric bootstrapping.
+Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the respective names in the COVID19-Nowcasting-Hub.
+For information on the model design, we refer to~\citet{Wolffram2023}.
+In addition to those nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
+We denote them by ENS-MEAN and ENS-MED.
+\citet{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams stated in a preregistered study.
+In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
+We use the data from February 8, 2024, for the true values.
+The nowcasts are resolved with regard to all inhabitants and specific geographical and age breakdowns.
+We do not expand on the models' performance on specific regions or age groups in Germany and the probabilistic nowcast assessment.
+Figure~\ref{fig:app-covid-true-nowcast} displays the true and nowcast data for the evaluation period.
+The time comprises the fourth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany, lasting until May 28, 2022~\citep{Tolksdorf2022}.
+
+Table~\ref{tab:app-covid-rmse} summarizes the point evaluation measures for the issued mean of the different models.
+The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\citep[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
+The best-performing models in terms of RMSE and MAE are the ILM and RKI models.
+The ensemble methods perform worse than the best models regarding the mean location.
+The performance of the models is diverse, with more than twice as high RMSE values for the worst models compared to the best models.
+Note that the high values for the EPI model could be driven by an exceptionally far-off value at the end of the evaluation period.
+
+In addition to close inspection of the point evaluation measures, assessing the trending of the nowcasts is crucial.
+To assess the impact of taken measures and the direction of the curve, it is essential to distinguish between rising and falling hospitalization rates.
+If hospitalization rates rise, measures should be tightened, while falling rates might allow for loosening measures.
+Asymmetries are especially relevant for assessing whether some models are better at recognizing a fall than a rise or vice versa.
+
+\begin{table}[]
+    \centering
+    \input{plots/covid_nowcast/01_model_scores}
+    \caption{Point evaluation measures for the issued mean of the different models. The evaluation period comprises 159 days. }
+    \label{tab:app-covid-rmse}
+\end{table}
+
+\subsubsection*{Results}
+
+In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
+We report the trending for the horizons 1, 7, and 14 days.
+While horizon one assesses the short-term trending, horizons seven and 14 consider the medium-term trending.
+The horizons seven and 14 are particularly interesting, reflecting a usual period until new policy changes are taken.
+
+Before stating the results, we provide background information on the marginal distributions of the actual values and nowcasts for the different horizons.
+Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} presents marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
+Overall, the variability and general level of differences grow with the horizon.
+The standard deviation increases from roughly 300 for horizon one to 1,200 for horizon seven and 2,000 for horizon 14 days.
+Similarly, the 10\%-quantile of differences increases.
+The 10\%-quantile is used for the exclusion areas in the trending assessment.
+The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the absolute differences.
+Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.
+
+\begin{table}
+    \centering
+    \tiny
+    \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
+    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days.
+The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
+The subscript $q_{0.1}$ denotes the measures with exclusion area.
+}
+    \label{tab:app-covid-trending-ratios-lag-7}
+\end{table}
+
+Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the horizon of seven days.
+The trending ratios without exclusion area range from 0.72 to 0.85 for the horizon of seven days.
+The negative trending ratios are higher than the positive trending ratios for all models.
+The confidence intervals for the positive and negative trending ratios do not overlap for all models, indicating that the trending ratios are indeed different.
+The 10\%-quantile exclusion areas have, at most, an influence of 0.03 on the ratios.
+The model with the highest trending ratio is the ILM model, and the model with the lowest is the RKI model.
+The confidence intervals between all models overlap.
+The positive trending ratio implies a similar ranking of the models, while the negative ratio is second best for the RKI model.
+For the horizons of one and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
+
+\begin{figure}
+    \centering
+%    \includegraphics{}
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
+    \caption{Conditional trending plot.}\label{fig:app-covid-cond-prob-7}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
+    \caption{Trending ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-trending-ratio-7}
+    \end{subfigure}
+    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
+    \label{fig:app-covid-cond-prob-trending-ratio-7}
+\end{figure}
+
+Figure~\ref{fig:app-covid-cond-prob-trending-ratio-7} shows the conditional trending plots and the trending ratio over the exclusion area for the horizon seven days; the respective plots for the horizons one day and 14 days are shown in Figure~\ref{fig:app-covid-cond-prob-trending-ratio-1-14}.
+Here, only the best models in point evaluation measures, ILM, RKI, RIVM, and ENS-MED, are shown to keep the plots easily readable.
+If RKI or ILM issues a fall in the hospitalization rate, the probability of a fall is higher than if RIVM or ENS-MED issues a fall.
+The opposite is the case for a nowcasted hospitalization rate increase, and the difference between the models' performance is higher.
+Similar observations can be made for the horizon of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
+For a horizon of one day, the models' conditional trending ability difference is less pronounced (see Figure~\ref{fig:app-covid-cond-prob-1}).
+The RKI model is still less conclusive when issuing an increase in the hospitalization rate, while RIVM is most informative in that case.
+The curves cross for an issued fall, with ENS-MED being on top for issued falls above 250.
+
+The trending ratios for various exclusion areas are shown in Figure~\ref{fig:app-covid-trending-ratio-7}.
+In general, the trending ratio increases with larger exclusion areas.
+While the RIVM and ENS-MED trending ratios evolve similarly, the RKI and ILM trending ratios get closer.
+For the horizon of one day, the RKI trending ratio decreases with increasing exclusion area size while the other models rise (see Figure~\ref{fig:app-covid-trending-ratio-1}).
+For the horizon of 14 days, all trending ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-trending-ratio-14}).
+
+Figure~\ref{fig:app-covid-prob} shows the \acf{bs} and reliability diagrams for the same subset of models, the ILM, RKI, RIVM, and ENS-MED.
+The probabilities of increase for the different models are computed using the nowcast quantiles.
+For each horizon $l$, $10,000$ samples of the forecast date $t$ and the forecast date $t-l$ based on the nowcasts of issue date $t$ are generated, and the proportion of positive changes is computed (see Appendix~\ref{subsec:app-probabilistic-trending-evaluation}).
+Remember that low a \ac{bs} and a plot along the diagonal are signs of good trending.
+The \ac{bs} is the lowest for the RIVM model, while the ENS-MED model has the best \ac{bs} for the horizon of seven and 14 days.
+The RKI model yields the highest \ac{bs} for all horizons.
+Note that the \ac{bs} is bounded by 0.25 for random guessing; thus, all models perform better than random guessing.
+The reliability diagrams show that the models are not well calibrated for the horizon of one day.
+While for predicted probabilities below 0.5, the observed ratio of increases is smaller, it is higher than predicted for probabilities above 0.5.
+Figure~\ref{fig:app-covid-prob-hist} shows the observed predicted increase probabilities for all horizons.
+For the horizons of seven and 14 days, the nowcasters issue only a few moderate probabilities.
+
+\begin{figure}
+    \begin{subfigure}{0.48\textwidth}
+        \centering
+        \tiny
+        \input{plots/covid_nowcast/70_brier_scores}
+    \caption{BS for the different models and horizons.}\label{fig:app-covid-prob-brier}
+    \end{subfigure}\hspace{0.01\textwidth}%
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_1}
+    \caption{Reliability diagram for horizon one day.}\label{fig:app-covid-prob-rel-1}
+    \end{subfigure}\hspace{0.01\textwidth}%
+    \caption[Brier Scores and reliability diagrams for ILM, RKI, RIVM, and ENS-MED.]{\aclp{bs} and reliability diagrams for ILM, RKI, RIVM, and ENS-MED.
+    The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
+    In the computation of BS and reliability diagram, missing values are excluded.
+    The reliability diagram for the horizon seven and 14 days is in the appendix (see Figure~\ref{fig:app-covid-rel}).}
+    \label{fig:app-covid-prob}
+\end{figure}
+
+\subsubsection*{Discussion}
+
+For all horizons, the influence of the exclusion area on the 10\%-quantile level is negligible.
+For example, the trending ratio changes at most by 0.03 for the EPI model with $\accml[14]$.
+The exclusion areas are thus not crucial for the trending assessment in the case of the nowcasts of the seven-day hospitalization rate.
+The lower bound of confidence intervals is at least 0.68 for all models, indicating that they perform better than random guessing the trend.
+
+Trending assessment evaluates the models differently from point evaluation measures.
+RKI is among the best in point evaluation measures but performs worse in trending assessment.
+The assessment of asymmetry in the conditional trending plots is crucial for interpreting the trending ratios, with the RKI model being the most prominent example.
+
+Figure~\ref{fig:app-covid-trending-ratio-7} shows that the trending ratio increases with larger exclusion areas.
+This indicates that if the model predicts a large change, the direction is more accurate than when a small change is predicted.
+
+The probabilistic trending assessment shows that the models are better than random guessing.
+In general, the reliability diagram cannot provide information if specific probabilities are issued scarcely.
+Thus, the reliability diagrams for the horizons of 7 and 14 days do not contain information on moderate probabilities.
+The \ac{bs} values, however, work well for those examples and provide a good measure for the trending ability of the models.
+
+A more extensive data size would be beneficial for assessing the models' performance.
+For the evaluation period of 159 days, the trending ratio confidence intervals overlap; thus, no conclusions can be drawn from the trending evaluation comparing the models.
+
+\subsection{Forecasting emergency department arrivals}\label{sec:application-eda}
+
+In a second example, we consider forecasting the number of arrivals in a large emergency department per hour.
+Good forecasts are crucial for planning staff and resources.
+In a study by \citet{Rostami-Tabar2023}, several models predict hourly outcomes.
+Every 12 hours, the models issue hourly forecasts for the next 48 hours.
+Thus, management can take measures according to the expected number of arrivals, for example, by redeploying staff and reconfiguring units.
+
+\citet{Rostami-Tabar2023} publish means and probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
+The models are trained on data from April 1, 2014, to February 28, 2018, and evaluated on data from March 1, 2018, to February 28, 2019.
+For further notes on the models and the evaluation, we refer to \citet{Rostami-Tabar2023}.
+From the issued forecasts, we use the mean as a point forecast for the trending assessment and evaluate the probabilistic trending subsequently.
+We use the forecasts issued at the first time point for every target time.
+Thus, the forecasts are issued 36 to 48 hours ahead of the target time, and the emergency department management has time to adjust the measures according to the expected number of arrivals.
+Considering only the forecasts of at least 36 hours ahead, we restrict the evaluation period to March 2, 2018, at noon, to February 28, 2019, at 23:00, comprising 8,724 hours.
+
+In this setup, trending assessment is a simple and intuitive way to assess the models' performance.
+The trending perspective is easy for the management to understand and implement, as simple comparisons of the expected workload to a recent shift can be made.
+If, for example, the staff was near capacity during the last shift and an increase in the number of arrivals is expected, management can take measures to adjust the workload.
+
+The number of arrivals has a strong weekly and daily pattern.
+Thus, we consider the horizons of 72 hours, the last already observed shift of the same hour of day, and seven days, the previous shift of the same hour and day.
+Table~\ref{tab:app-eda-point-evaluation} lists the point evaluation measures and the count of available forecasts.
+The best-performing models regarding RMSE and MAE are the NBI-2 and Poisson-2 models.
+More than 8,600 forecasts are available for all models, with differences in the number due to missing values on four afternoons in 2018.
+Note that the reported values for the RMSE differ from those in \citet{Rostami-Tabar2023}.
+In contrast to their work, we use only the forecast data at least 36 hours ahead rather than the entire forecast data for evaluation.
+
+\begin{table}
+\centering
+\input{plots/ed_arrival/00_point_evaluation_measures}
+\caption{Point evaluation measures for the models. The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.}\label{tab:app-eda-point-evaluation}
+\end{table}
+
+\subsubsection*{Results}
+
+We start by analyzing the marginal distributions for the changes of forecasts and actual values for the horizons of three and seven days in Table~\ref{tab:app-eda-marginals}, again.
+Thereby, the computed difference aligns with Section~\ref{subsec:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the actual value is available when issuing the forecast.
+The fraction of positive differences varies between 0.39 and 0.63 for the horizon of three days and between 0.37 and 0.63 for the horizon of seven days.
+The variability of differences decreases for the larger horizon for most models; only for the ETS model does it increase.
+The 10\%-quantile of the differences is between zero and one for all models and horizons.
+Thus, we exclude differences smaller than one from the trending assessment.
+The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.
+
+\begin{table}
+    \centering
+    \tiny
+    \input{plots/ed_arrival/10_marginal_analysis}
+    \caption{Marginal analysis of the forecast and true differences.
+The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.
+Column (2) shows the fraction of values not in the exclusion area of size one.}
+    \label{tab:app-eda-marginals}
+\end{table}
+
+Table~\ref{tab:app-eda-trending-ratios} lists the trending ratios for all models for three and seven-day horizons.
+The trending ratios range from 0.68 to 0.84 for a horizon of three days and from 0.68 to 0.82 for seven days.
+The negative and positive trending ratios differ for all models and horizons.
+For some models, for example, the GBM-2 model, the positive trending ratio is higher; for some models, for example, the tbats model, the negative trending ratio is higher.
+The confidence interval width is at most 0.02 for the trending ratios and at most 0.03 for the positive and negative trending ratios.
+The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for the three and seven-day horizon, while Poisson-2 and NBI-2 have the highest negative trending ratio.
+
+Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
+The conditional trending plots show similar courses for the two horizons, though the curves are shifted downwards for the horizon of seven days.
+The model's relative trending ability evolves consistently for the two horizons, with the NBI-2 and Poisson-2 models being indistinguishable.
+The GBM-2 model outperforms the qreg-1 model for all $x$.
+The models NBI-2 and Poisson-2 have the highest trending ability for all negative values of $x$ and the lowest trending ability for all positive values of $x$.
+Benchmark-1 lies between the other models for all $x$.
+
+Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
+The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
+The \acp{bs} are smallest for NBI-2 and Poisson-2 for both horizons, while the \acp{bs} for the other models are larger and differ more.
+The qreg-1 model has both horizons' highest \ac{bs}.
+The reliability diagrams of GBM-2 and NBI-2 are also close and show a too-small fraction of increases for the predicted probability overall.
+For the other models, the reliability diagrams show a fraction of increases that are too large for the corresponding predicted probability.
+
+\begin{table}
+    \centering
+    \tiny
+    \input{plots/ed_arrival/50_trending_ratio}
+    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days.}
+    \label{tab:app-eda-trending-ratios}
+\end{table}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_3}
+    \caption{Horizon three days}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_7}
+    \caption{Horizon seven days}
+    \end{subfigure}
+    \caption{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability.
+The plots of NBI-2 and Poisson-2 are indistinguishable.}
+    \label{fig:app-eda-cond-prob}
+\end{figure}
+
+\begin{figure}
+    \begin{subfigure}{0.32\textwidth}
+    \tiny
+    \input{plots/ed_arrival/70_brier_scores}
+    \caption{BS for the different models and horizons.}\label{fig:app-eda-prob-brier}
+    \end{subfigure}\hspace{0.01\textwidth}%
+    \begin{subfigure}[t]{0.32\textwidth}
+    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_3}
+    \caption{Reliability diagram for horizon three days.}\label{fig:app-eda-prob-rel-3}
+    \end{subfigure}\hspace{0.01\textwidth}%
+    \begin{subfigure}[t]{0.32\textwidth}
+    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_7}
+    \caption{Reliability diagram for horizon seven days.}\label{fig:app-eda-prob-rel-7}
+    \end{subfigure}
+    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days.
+The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic trending simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically.
+Probabilistic trending by the \ac{bs} is best for the models NBI-2 and Poisson-2 for both lags.}
+    \label{fig:app-eda-prob}
+\end{figure}
+
+\subsubsection*{Discussion}
+
+The trending ability is consistent for the two horizons, with the models' relative trending ability evolving similarly for the two horizons.
+The models' trending ability is generally higher for the smaller horizon, but the differences are minor, and confidence intervals overlap.
+
+The trending differs for all models for positive and negative predicted change directions.
+While some models, such as GBM-2 and qreg-1, have the highest positive trending ratio, others, such as Poisson-2 and NBI-2, have the highest negative trending ratio.
+Thus, the uncertainty of the model's predicted change has to be assessed differently based on the direction.
+
+The results of the probabilistic trending evaluation endorse the point trending assessment and assign the best scores to NBI-2 and Poisson-2.
+The reliability diagrams show that they underestimate the fraction of increases slightly.
+
+The example shows that trending assessment is detached from standard point evaluation measures.
+While the models with the lowest RMSE, NBI-2, and Poisson-2, also have a high trending ability, three models with below-average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive trending ability.
+
+\subsection{Invasive and non-invasive blood pressure monitoring} \label{sec:application_measurement}
+
+In the last briefer example, we consider the trending assessment of measurement data.
+The data is from the MIMIC-III database, including various information on patients in critical care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}).
+The data is publicly available and also includes numerical measurement data such as heart rate, blood pressure, and oxygen saturation in a waveform database \citetext{\citealp{Moody2017}; available through \citealp{Goldberger2000}}.
+
+For some patients, the data includes \ac{abp} and \ac{nbp} measurements.
+While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods.
+For an overview of blood pressure measurement methods, see \citet{Saugel2014}.
+For critical patients, changes in blood pressure can be crucial for the treatment.
+Thus, trending assessment can be performed in addition to standard accuracy analysis~\citep[see, for example, ][]{Mostafa2020}.
+Thus, we assess the trending ability of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
+The database contains 64,168 numerical data records.
+One data record includes all numerical measurements for one patient.
+The measured signals vary in length, frequency, and type of measurement.
+Thus, only a subset of the data contains measurements of \ac{abp} and \ac{nbp} simultaneously.
+2,548 include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one measurement of systolic \ac{abp} and \ac{nbp} at the same time; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
+
+We consider the horizons of one minute, five minutes, and 15 minutes for the trending assessment, as those are typical intervals of NBP measurements.
+
+\subsubsection*{Results}
+
+Again, we exclude the smallest 10\% of absolute differences in trending assessment.
+The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different horizons are shown in Figure~\ref{fig:app-mimic-4q}.
+The number of points in the four-quadrant plot is smaller due to the restriction to data records with measurements of mean or systolic \ac{abp} and \ac{nbp} simultaneously for two consecutive times with the specified horizons.
+Thus, we use the \ac{nbp} measurements as the test method and the \ac{abp} measurements as the gold standard.
+For the systolic measurements, 290, 332, and 442 points are available for the horizons of one, five, and 15 minutes; for the mean measurements, 406, 430, and 542.
+
+The trending ratios, including confidence intervals for the different horizons, are listed in Table~\ref{tab:app-mimic-trending-ratios}.
+The confidence intervals have lower bounds of 0.5 or slightly above for the measurements with a horizon of one minute.
+For larger horizons, the trending ratio increases.
+The difference between positive and negative trending ratios is small for all types and horizons, with overlapping confidence intervals.
+
+Figure~\ref{fig:app-mimic-cond-prob} shows the conditional trending plots for the different horizons and the systolic and mean blood pressure measurements.
+It becomes apparent that the systolic measurements have a higher trending ability than the mean measurement, except for small negative predicted changes.
+This aligns with the trending ratios, but the confidence intervals overlap.
+
+\begin{figure}
+    \centering
+    \includegraphics{plots/mimic/plot_4q}
+    \caption{Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements.
+The upper row contains systolic measurements, and the lower row contains mean measurements.
+The columns contain the horizons one, five, and 15 minutes.}
+    \label{fig:app-mimic-4q}
+\end{figure}
+
+\begin{table}
+    \centering
+    \input{plots/mimic/trending_ratio}
+    \caption{Trending ratios for the different horizons and the systolic and mean blood pressure measurements.}
+    \label{tab:app-mimic-trending-ratios}
+\end{table}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{.32\textwidth}
+        \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag1}
+        \caption{Horizon one minute.}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \begin{subfigure}[t]{.32\textwidth}
+        \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag5}
+        \caption{Horizon five minutes.}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \begin{subfigure}[t]{.32\textwidth}
+        \includegraphics{plots/mimic/cond_prob_diff_nbp_abp_lag15}
+        \caption{Horizon 15 minutes.}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the horizons one, five, and 15 minutes. }
+    \label{fig:app-mimic-cond-prob}
+\end{figure}
+
+\subsubsection*{Discussion}
+
+The four-quadrant plots contain a considerable number of extreme points.
+Whether these points are due to measurement errors or extreme values is not distinguishable.
+Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile~\citep[see][]{Critchley2010}.
+We do not follow this approach here, as the extreme values are not necessarily measurement errors and could be particularly relevant.
+
+The differences between positive and negative predicted changes are small in this example.
+The positive and negative trending ratios have overlapping confidence intervals, and the conditional trending plots do not contain prominent deviations in the course.
+This aligns with the four-quadrant plots, where no apparent asymmetry is visible.
+
+The bootstrap confidence intervals are wide.
+The width is around 0.1 for the trending ratio, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the horizon of one minute.
+Thus, the confidence intervals cover 0.5 for systolic measurement and the horizon of one minute, and the equality to random guessing cannot be excluded.
 
@@
 
-\input{sections/05-conclusion.tex}
+In this paper, we examine various methods to analyze trending, that is, the statistical consistency of observed and predicted change direction.
+While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting (see Section~\ref{subsec:notation}), the trending assessment using the computed predicted and observed change is similar.
+The evaluation of trending can accompany other evaluation techniques, such as measures of deviation or probabilistic scoring rules.
+
+Four-quadrant plots facilitate the visual inspection of the trending ability for a signal (see Section~\ref{subsec:trending-four-quadrant-plot}).
+The trending ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates trending.
+Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
+The positive and negative trending ratios analyze the trending ratio given whether the predicted change is positive or negative, respectively.
+Thus, they quantify the credibility of the respective predictions.
+The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative trending abilities and that they add valuable information to the trending ratio.
+In the applications, the bootstrap confidence intervals of Section~\ref{subsec:trending-measures} are used to quantify the estimation uncertainty of the trending measures.
+The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
+For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or to assess models with high trending differences.
+
+A conditional trending plot visualizes the probability of correct trending over the predicted change of the signal (see Section~\ref{subsec:trending-cond-prob}).
+It is based on a multivariate \acf{kde} of predicted and observed change.
+In the application, the conditional trending plot gives reasonable insights into the local effects of the trending ability.
+Section~\ref{subsec:trending-probabilistic} adapts measures of probabilistic forecast evaluation to the trending assessment of probabilistic forecasts and nowcasts.
+The \acf{bs} as numerical assessment of probabilistic trending is introduced, and reliability diagrams are used to visualize the local trending ability of probabilistic forecasts.
+
+The methods of trending assessment are applied to COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements in Section~\ref{sec:application}.
+While trending evaluation should not be the only aspect, it is a valuable addition to evaluating nowcasts, forecasts, and measurements.
+Models with very different accuracies are usually scored similarly in trending evaluation, but trending evaluation can differentiate between models with similar accuracies.
+As in the application in Section~\ref{sec:application-covid}, models with medial point forecast evaluation measures can have the most meaningful positive trending ability.
+
+We did not expand on two modeling aspects throughout this paper, which we leave for further research.
+In the estimation, we did not consider sequential correlation.
+The computation of differences is a standard procedure in time series analysis to remove sequential dependence, but, in general, some could remain, and the estimators could account for it.
+Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation using time-series bootstrap methods~\citep{Hardle2003,Kreiss2012}.
+
+The estimators of Section~\ref{subsec:trending-measures} do not account for imbalances in the number of observed positive and negative changes \citep[for a theoretical analysis, see][Chapter 3]{Jolliffe2012}.
+Significant differences in the number of observed positive and negative changes are unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.
+However, if the number of positive and negative observed changes differs widely, unbalanced-data-aware measures should be considered.
+There are various adapted measures for unbalanced outcomes, for example, Cohen's $\kappa$ \citep{Cohen1960} or those listed in \citet[Table 3.3]{Jolliffe2012}.
 
@@
 
-\input{sections/50-appendix-data-generation}
+\section{Additional material on Section~\ref{sec:trending}}\label{sec:appendix-trending}
 
-\input{sections/55-appendix-application-covid}
+\subsection{Data generation for Section~\ref{sec:trending}}\label{subsec:app-trending-data-generation}
+
+The first dataset is generated by sequentially generating $\diffx$ and $\diffy$.
+First, the $\diffxt$ are sampled as a sum of a standard normal random number and a uniform random number on $(-10, 10)$:
+\begin{equation*}
+    \diffxt \sim N(0, 1) + U(-10, 10) \quad t = 1, \dots, T.
+\end{equation*}
+Subsequently, the $\diffy$ are simulated for a constant trending ratio $k$ by
+\begin{equation*}
+    \diffyt = \diffxt \cdot n_t \cdot b_t,
+\end{equation*}
+where $n_t$ is a truncated normal distribution with mean one and standard deviation 0.5, truncated at 0, and $b_t$ is a symmetric Bernoulli random variable with parameter $k$.
+For a time-varying trending ratio, the parameter $k$ is modified to have a wave-shape over time, that is,
+\begin{equation*}
+    k_t = 0.75 + \sin(t / 365.25 \cdot 2 \pi) / 4.
+\end{equation*}
+For the asymmetric trending ratio, $k$ is a function of $\diffxt$,
+\begin{equation*}
+    k(x) = 0.5 + \min \left\{ \max \left\{ \frac{x + 5}{10}, 0  \right\} , 1 \right\} / 2.
+\end{equation*}
+
+In the second approach, $\diffyt$ and $\diffxt$ are modeled to be multivariate normal with mean 0 and covariance matrix
+\begin{equation*}
+    \Sigma = \begin{pmatrix} 4 & 3 \\ 3 & 4 \end{pmatrix}.
+\end{equation*}
+Thus, the conditional probability of trending can be calculated by a conditional normal distribution to
+\begin{equation*}
+    P(\diffyrv \diffxrv > 0 | \diffxrv = x) = \Phi \left( \frac{3}{2 \sqrt{7}} x \right),
+\end{equation*}
+where $\Phi$ is a standard normal \ac{cdf}.
+
+The four-quadrant plots for the sample realizations of the data generation schemes are shown in Figure~\ref{fig:appendix_dgps}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}{0.24\textwidth}
+        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1}
+        \caption{Constant trending ratio.}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \begin{subfigure}{0.24\textwidth}
+        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_time}
+        \caption{Time-varying trending ratio}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \begin{subfigure}{0.24\textwidth}
+        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_asym}
+        \caption{Asymmetric trending ratio}
+    \end{subfigure}\hspace{0.01\textwidth}
+    \begin{subfigure}{0.24\textwidth}
+        \includegraphics{plots/illustrative_examples/appendix_4q_dgp2}
+        \caption{Second approach}
+    \end{subfigure}
+    \caption{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.
+Although the first and second plots differ over time, their difference is not discernible in the plots.
+The third data set's asymmetry is visible in the plot, but the decrease in the trending ability near 0 is not visible.
+}
+    \label{fig:appendix_dgps}
+\end{figure}
+
+\subsection{Simulation study on bootstrapping confidence intervals}\label{subsec:appendix-trending-bootstrap}
+
+We examine three methods for bootstrapping: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
+In the \textit{percentile} approach, the confidence interval for the level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
+The \textit{basic} approach computes the confidence interval based on the non-bootstrap estimate using the bootstrapped quantile deviations~\citep{Davison1997}.
+The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\citep{Efron1987}.
+Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
+The basic approach balances these two objectives.
+We compare the approaches in a small synthetic data study on their small-dataset behavior and computation time.
+
+We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
+The considered datasets are the first dataset with asymmetric dependence and the second dataset outlined in Appendix~\ref{subsec:app-trending-data-generation}.
+In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\citep{Virtanen2020}.
+The prescribed confidence level is 90 \%, and the number of bootstrap samples is $10,000$.
+The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
+The true values of the accuracy are computed based on a dataset of size $10^8$, yielding 0.7501 and 0.7700 for the two datasets.
+The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.
+For the small sample sizes up to $T = 168$, only the \ac{bca} method keeps the confidence interval size and yields slightly wider confidence intervals.
+The method's results do not differ for the larger sample sizes.
+The computation time for the \ac{bca} method is slightly larger than for the other methods, but all methods have a moderate computation time.
+\Ac{bca} is the only method that maintains the confidence level for small datasets while increasing the computation time only moderately for larger datasets.
+Therefore, we use the \ac{bca} method for confidence intervals in the applications in Section~\ref{sec:application}.
+
+\begin{table}
+    \centering
+    \tiny
+    \begin{subtable}{.48\textwidth}
+        \input{plots/illustrative_examples/bootstrap_butterfly_cis}
+        \caption{First dataset}
+    \end{subtable}\hspace{0.02\textwidth}
+    \begin{subtable}{.48\textwidth}
+        \input{plots/illustrative_examples/bootstrap_normal_cis}
+        \caption{Second dataset}
+    \end{subtable}
+    \caption{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$.
+The average width of the confidence interval is listed in brackets.}
+    \label{tab:trending_bootstrap}
+\end{table}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}{0.48\textwidth}
+\includegraphics{plots/illustrative_examples/boxplot_comp_time_butterfly}
+        \caption{First dataset}
+    \end{subfigure}
+    \begin{subfigure}{0.48\textwidth}
+    \includegraphics{plots/illustrative_examples/boxplot_comp_time_normal}
+        \caption{Second dataset}
+    \end{subfigure}
+    \caption{Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$.
+The computation time refers to bootstrapping one confidence interval based upon $10,000$ values.
+Each boxplot reflects $10,000$ samples.
+The \ac{bca} method takes slightly longer than the other two, but the difference is negligible.}
+    \label{fig:trending_bootstrap_time}
+\end{figure}
+
+\subsection{Visualization of different bandwidth selectors in multivariate \ac{kde}}\label{subsec:appendix-kde}
+
+We examine the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\citep{Seabold2010} in Figure~\ref{fig:trending-cond-prob-bw}.
+While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.
+The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
+The second method, cross-validation least squares, requires long computation times while yielding small or no bandwidth results, even for two relatively small datasets.
+The rule-of-thumb and cross-validation maximum likelihood methods yield reasonable results at moderate computation times.
+Further examples, including comparisons between methods regarding their trending ability, are available in Section~\ref{sec:application}.
+
+\begin{figure}
+    \centering
+    \begin{subfigure}{.48\textwidth}
+        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_butterfly}
+        \caption{First dataset with asymmetric dependence.}
+    \end{subfigure}
+    \begin{subfigure}{.48\textwidth}
+        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_normal}
+        \caption{Second dataset. }
+    \end{subfigure}
+    \caption{Conditional trending plot for different bandwidth selection processes.
+Cross-validation least squares takes a considerably larger computation time.
+It does not converge for the first data set and yields a bandwidth too small for the second data set.
+The rule of thumb is the fastest method but tends to oversmooth.
+The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time.
+$\varepsilon$ specifies an exclusion area $E = \{(x, y) \in \R^2: (-\varepsilon \leq x \leq \varepsilon)\}$ in $\diffx$-direction.}\label{fig:trending-cond-prob-bw}
+\end{figure}
+
+\subsection{Probabilistic trending evaluation}\label{subsec:app-probabilistic-trending-evaluation}
+
+If forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
+Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
+The quantiles $q_p$ generally differ for each time step, but we omit an index here for ease of notation.
+The probability $\pc_t$ of a \textit{negative} change is between $p^{\star}$ and $p^{\star} + 1/\pmax$ for
+\begin{equation*}
+    p^{\star} = \max \{p \in \{1/\pmax, \dots, (\pmax-1) / \pmax\}: q_p \leq y_t\} , \quad \text{if}\ q_{1/\pmax} \leq y_t \leq q_{1 - 1/\pmax}.
+\end{equation*}
+Quantiles do not determine the location within the interval $[p^{\star}, p^{\star} + 1/\pmax]$.
+Under the assumption of a uniform distribution within the quantile interval, the probability of a negative change is
+\begin{equation*}
+    \pc_t = \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star}.
+\end{equation*}
+The approach does not assign probabilities for $y_t$ smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile.
+As a simple extension, we assume that the probability mass is uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
+This yields
+\begin{equation*}
+\pc_t = \begin{cases}
+    \max \{\frac{1}{\pmax} - \frac{q_{p^\star} - y_t}{\pmax (q_{p2/\pmax} - q_{1/\pmax})}, 0\} &, \text{if } y_t < q_{1/p}, \\
+    \min \{\frac{1}{\pmax} - \frac{y_t - q_{(\pmax-1)/\pmax}}{\pmax (q_{(\pmax-1)/\pmax} - q_{(\pmax-2)/\pmax})}, 1\} &, \text{if } y_t > q_{1 - 1/p}, \\
+    \frac{y_t - q_{p^\star}}{\pmax (q_{p^{\star} + 1} - q_{p^{\star}})} + p^{\star} &, \text{otherwise.}
+\end{cases}
+\end{equation*}
+The probability of positive change is $p_t = 1 - \pc_t$.
+
+If the true value is given as a distribution because it is still unknown, the probabilities $p_t$ can be computed by integration.
+Let for two nowcasts the distributions be given by \acp{pdf} $f_{t+l|t+l}$ and $f_{t|t+l}$ with \acp{cdf}  $F_{t+l|t+l}$ and $F_{t|t+l}$.
+Then, the probability of a negative change can be computed by
+\begin{align}
+    \pc_t
+        &= \int_{\begin{subarray}{l}x_1, x_2 \in \R: \\ x_2 < x_1\end{subarray}} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: (x_1, x_2) \nonumber\\
+        &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
+        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
+\end{align}
+Thereby, the distributions are assumed to be independent.
+If the nowcasts have the form of a multivariate distribution, including the dependence of the two \acp{pdf}, $f_{t+l|t+l} (x_2)$ has to be replaced by the \ac{pdf} conditional on $x_1$.
+As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
+For forecasts, the indexes have to be shifted.
+If no \acp{pdf} are available, they can be estimated from the \ac{cdf} or quantiles, or the \ac{cdf} or quantiles can be used to generate samples for the Monte Carlo approximation.
+This approach is applied in Section~\ref{sec:application-covid}, as the true values are published with a delay of more than 80 days, and the nowcasts are given as quantiles.
+
+ \section{Additional material on Section~\ref{sec:application-covid}}\label{sec:appendix-application-covid}
+
+\begin{table}
+    \centering
+    \begin{tabular}{l l}
+        \toprule
+        Abbreviation & Nowcasting hub key \\
+        \midrule
+        EPI & Epiforecasts-independent \\
+        ILM & ILM-prop \\
+        KIT & KIT-simple\_nowcast \\
+        LMU & LMU\_StaBLab-GAM\_nowcast \\
+        RIVM & RIVM-KEW \\
+        RKI & RKI-weekly\_report \\
+        SU & SU-hier\_bayes \\
+        SZ & SZ-hosp\_nowcast\\
+        ENS-MEAN & NowcastHub-MeanEnsemble\\
+        ENS-MED & NowcastHub-MedianEnsemble\\
+        \bottomrule
+    \end{tabular}
+    \caption{Matching the abbreviation to the key in the nowcasting hub.
+    Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
+    \label{tab:app-covid-models}
+\end{table}
+
+\begin{table}
+    \centering
+    \tiny
+    \input{plots/covid_nowcast/10_marginal_analysis}
+    \caption{Analysis of the nowcast and true differences for the horizons 1, 7, and 14 days.
+    The column (1), $l=l$ shows the number of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
+    \label{tab:app-covid-marginals}
+\end{table}
+
+\begin{table}
+    \centering
+    \begin{subtable}[t]{\textwidth}
+        \input{plots/covid_nowcast/30_trending_ratios_lag_1}
+    \caption{1 day.}
+    \end{subtable}
+    \begin{subtable}[t]{\textwidth}
+        \input{plots/covid_nowcast/30_trending_ratios_lag_14}
+        \caption{14 days.}
+    \end{subtable}
+    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the horizon 1 and 14 days.
+The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values.
+}
+    \label{tab:app-covid-trending-ratios-lag-1-14}
+\end{table}
+
+\begin{figure}
+\centering
+\includegraphics{plots/covid_nowcast/30_4q_plots}
+\caption{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days. The spread in both directions increases with the horizon. }
+\label{fig:app-covid-4q}
+\end{figure}
+
+\begin{figure}
+    \centering
+%    \includegraphics{}
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
+    \caption{Conditional trending plot for horzion 1.}\label{fig:app-covid-cond-prob-1}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_14}
+    \caption{Conditional trending plot for horizon 1.}\label{fig:app-covid-cond-prob-14}
+    \end{subfigure}
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_1}
+    \caption{Trending ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-trending-ratio-1}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_14}
+    \caption{Trending ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-trending-ratio-14}
+    \end{subfigure}
+    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the lag seven days.}
+    \label{fig:app-covid-cond-prob-trending-ratio-1-14}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{.48\textwidth}
+        \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_7}
+        \caption{Reliability diagram for horizon seven days.} \label{fig:app-covid-reliability-7}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{.48\textwidth}
+        \includegraphics{plots/covid_nowcast/60_reliability_diagram_lag_14}
+        \caption{Reliability diagram for horizon 14 days.} \label{fig:app-covid-reliability-14}
+    \end{subfigure}
+    \begin{subfigure}{\textwidth}
+        \includegraphics{plots/covid_nowcast/70_prob_hist}
+        \caption{Count histogram of the predicted probabilities for the horizon one, seven, and 14 days.} \label{fig:app-covid-prob-hist}
+    \end{subfigure}
+    \caption{The reliability diagram for the models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
+    Additionally, the count of predicted probabilities for the horizons is shown.
+        The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
+    As the models issue small or large probabilities of increase for the higher horizons, little information on the accuracy of moderate probability predictions is available.}
+    \label{fig:}
+\end{figure}
 

### 2024-03-29 paper.tex
Some corrections; add short captions

@@
 %\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
+\author*[1]{Jonas Rieger}
+\author[1]{Bolin Liu}
 \author[1]{Oliver Grothe}
-\author[1]{Bolin Liu}
-\author*[1]{Jonas Rieger}
 
@@
 \maketitle
- \listoftodos
-
-\section{Introduction}\label{sec:introduction}
+% \listoftodos
+
+\section{Introduction}\label{sec:trending-introduction}
+
 The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
@@
     \end{tabular}
-    \caption{Computation of the predicted change in the different applications.
+    \caption[Computation of the predicted change in the different applications.]{Computation of the predicted change in the different applications.
 For nowcasting and forecasting, signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
@@
     \end{subfigure}%
-    \caption{Visualizations for data with a time-varying trending ratio.
+    \caption[Visualizations of the four-quadrant plot for data with a time-varying trending ratio.]{Visualizations for data with a time-varying trending ratio.
 We defer information on the data generation process to the appendix (see \ref{subsec:app-trending-data-generation}).
@@
 \section{Application to measurement, nowcasting, and forecasting data} \label{sec:application}
+
 In this section, we evaluate the trending in three different applications.
@@
         \end{subfigure}
-    \caption{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+    \caption[True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022.]{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
     The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
@@
     \input{plots/covid_nowcast/01_model_scores}
-    \caption{Point evaluation measures for the issued mean of the different models. The evaluation period comprises 159 days. }
+    \caption[The trending ratios for Covid nowcasting.]{Point evaluation measures for the issued mean of the different models. The evaluation period comprises 159 days. }
     \label{tab:app-covid-rmse}
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days.
+    \caption[The trending ratios for Covid nowcasting.]{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days.
 The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
@@
     \end{subfigure}
-    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
+    \caption[Conditional trending plot and trending ratio over exclusion area for the Covid nowcasts.]{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
     \label{fig:app-covid-cond-prob-trending-ratio-7}
@@
 \input{plots/ed_arrival/00_point_evaluation_measures}
-\caption{Point evaluation measures for the models. The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.}\label{tab:app-eda-point-evaluation}
+\caption[Point evaluation measures for the emergency department arrival forecasting models.]{Point evaluation measures for the models.
+The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.}\label{tab:app-eda-point-evaluation}
 \end{table}
@@
     \input{plots/ed_arrival/10_marginal_analysis}
-    \caption{Marginal analysis of the forecast and true differences.
+    \caption[Marginal analysis of the forecast and true differences in ermergency department arrival forecasting.]{Marginal analysis of the forecast and true differences.
 The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.
@@
     \input{plots/ed_arrival/50_trending_ratio}
-    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days.}
+    \caption[Trending ratios for the emergency deparment arrival forecasts.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days.}
     \label{tab:app-eda-trending-ratios}
@@
     \end{subfigure}
-    \caption{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability.
+    \caption[Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability in emergency department arrival forecasting.]{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability.
 The plots of NBI-2 and Poisson-2 are indistinguishable.}
@@
     \end{subfigure}
-    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days.
+    \caption[Probabilistic trending evaluation in emergency department arrival forecasting.]{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days.
 The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic trending simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically.
@@
     \includegraphics{plots/mimic/plot_4q}
-    \caption{Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements.
+    \caption[Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements.
+]{Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements.
 The upper row contains systolic measurements, and the lower row contains mean measurements.
@@
 
-\section{Discussion and Conclusion}\label{sec:conclusion}
+\section{Discussion and Conclusion}\label{sec:trending-conclusion}
 
@@
     \end{subfigure}
-    \caption{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.
+    \caption[Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.]{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.
 Although the first and second plots differ over time, their difference is not discernible in the plots.
@@
     \end{subtable}
-    \caption{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$.
+    \caption[Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$ in the bootstrap simulation study.]{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$.
 The average width of the confidence interval is listed in brackets.}
@@
     \end{subfigure}
-    \caption{Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$.
+    \caption[Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$.]{Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$.
 The computation time refers to bootstrapping one confidence interval based upon $10,000$ values.
@@
     \end{subfigure}
-    \caption{Conditional trending plot for different bandwidth selection processes.
+    \caption[Conditional trending plots for different bandwidth selection processes.]{Conditional trending plot for different bandwidth selection processes.
 Cross-validation least squares takes a considerably larger computation time.
@@
     \end{tabular}
-    \caption{Matching the abbreviation to the key in the nowcasting hub.
+    \caption[Matching the abbreviation to the key in the nowcasting hub.]{Matching the abbreviation to the key in the nowcasting hub.
     Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
@@
     \input{plots/covid_nowcast/10_marginal_analysis}
-    \caption{Analysis of the nowcast and true differences for the horizons 1, 7, and 14 days.
+    \caption[Marginal analysis of the Covid nowcast and true differences for the horizons 1, 7, and 14 days.]{Marginal analysis of the nowcast and true differences for the horizons 1, 7, and 14 days.
     The column (1), $l=l$ shows the number of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
@@
     \end{subtable}
-    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the horizon 1 and 14 days.
+    \caption[Trending ratios for the models with and without exclusion areas for the horizon 1 and 14 days in Covid nowcasting.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the horizon 1 and 14 days.
 The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values.
@@
 \includegraphics{plots/covid_nowcast/30_4q_plots}
-\caption{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days. The spread in both directions increases with the horizon. }
+\caption[Four-quadrant plots for the Covid nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.]{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.
+The spread in both directions increases with the horizon.
+}
 \label{fig:app-covid-4q}
@@
     \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
-    \caption{Conditional trending plot for horzion 1.}\label{fig:app-covid-cond-prob-1}
+    \caption{Conditional trending plot for horizon 1.}\label{fig:app-covid-cond-prob-1}
     \end{subfigure}\hfill
@@
     \end{subfigure}
-    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the lag seven days.}
+    \caption[Conditional trending plot and trending ratio over exclusion area in Covid nowcasting.]{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
     \label{fig:app-covid-cond-prob-trending-ratio-1-14}
@@
     \end{subfigure}
-    \caption{The reliability diagram for the models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
+    \caption[Reliability diagrams for Covid nowcasting models and the horizon seven and 14 days.]{The reliability diagram for the models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
     Additionally, the count of predicted probabilities for the horizons is shown.
@@
     As the models issue small or large probabilities of increase for the higher horizons, little information on the accuracy of moderate probability predictions is available.}
-    \label{fig:}
+    \label{fig:app-covid-rel}
 \end{figure}

### 2024-04-02 paper.tex
Small corrections and plot updates

@@
 Forecasting methods predict the future based on historical data, patterns, and exogenous factors.
-In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008,Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
+In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in][]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008,Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
 The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
@@
 
-Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past \citep{banbura2013now}.
+Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past~\citep{banbura2013now}.
 Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours~\citep{browning1989nowcasting,schmid2019nowcasting}.
-%In economics, nowcasting is used to predict statistics on the current economic situation, for example, the gross domestic product, which is collected with low frequency and is available with a considerable time delay \citep{banbura2013now}.
 In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{wu2021nowcasting} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{gunther2021nowcasting}.
@@
 Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods~\citep{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series~\citep{watson2010method}.
-Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value \citep{bland1986statistical}.
-Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years \citep{Saugel2015,saugel2018error,hiraishi2021concordance}.
+Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value~\citep{bland1986statistical}.
+Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years~\citep{Saugel2015,saugel2018error,hiraishi2021concordance}.
 
@@
 Hence, we provide blueprints for applying trending evaluation in practice and make remarks on interpreting the measures.
-We conclude the paper in Section~\ref{sec:conclusion}.
+We conclude the paper in Section~\ref{sec:trending-conclusion}.
 
@@
 
-As discussed in Section~\ref{sec:introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by a measurement, a nowcast, or a forecast.
+As discussed in Section~\ref{sec:trending-introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by a measurement, a nowcast, or a forecast.
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
@@
 
-Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data \citep{Critchley2010,Saugel2015}.
+Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data~\citep{Critchley2010,Saugel2015}.
 With that, we mainly estimate the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
@@
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
-Similar evaluations are used in other scientific areas, for example, with contingency tables in dichotomous forecasting or with confusion matrix in classification analysis~\citetext{see, e.g., the introductions in \citealt{James2021}, Ch.
+Similar evaluations are used in other scientific areas, for example, with contingency tables in dichotomous forecasting or with confusion matrices in classification analysis~\citetext{see, e.g., the introductions in \citealt{James2021}, Ch.
 4, and \citealp{Jolliffe2012}, Ch.
@@
 
-Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} accounting for such effects\citep[see, e.g.,][]{Saugel2015,Critchley2010}.
+Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} accounting for such effects~\citep[see, e.g.,][]{Saugel2015,Critchley2010}.
 Points within that area are omitted in the four-quadrant plot and the calculation of the measures.
@@
 Figure~\ref{fig:trending_4q} visualizes examples of exclusion area shapes.
-A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ ($\varepsilon_x, \varepsilon_y > 0$), leaves out points that are small in both components and, therefore, likely to be driven by noise.
-Points are included if the true value or signal is unlikely to be noise-driven.
+A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and, therefore, likely to be driven by noise.
+Points are included if at least one of true value or signal is unlikely to be noise-driven.
 In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
@@
 Let for a forecast $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
-Thereby, $F_{t | t-l} (x)$ yields the probability that the quantity of interest is at most $x$.
+Thereby, $F_{t | t-l} (x)$ specifies the forecastes probability that the quantity of interest is at most $x$.
 For trending assessment, we compare the predicted probability of positive change, denoted by $p_t$, with the occurrence of positive changes.
 A positive change occurs for any value at $t$ larger than the true value $y_{t-l}$, as then the change compared to $y_{t-l}$ is positive.
-The \ac{cdf} $F_{t | t-l} (y_{t-l})$ yields the predicted probability of any value at most $y_{t-l}$, a negative change.
-Thus, $1 - F_{t | t-l} (y_{t-l})$ is the probability of a positive change.
+The \ac{cdf} $F_{t | t-l} (y_{t-l})$ yields the predicted probability of any value at most $y_{t-l}$, and, thus, a negative change.
+Accordingly, $1 - F_{t | t-l} (y_{t-l})$ is the probability of a positive change.
 As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
-For nowcasts, the probability is calculated by
+For forecasts, the probability is calculated by
 \begin{equation*}
@@
     \begin{subfigure}[t]{0.48\textwidth}
-    \includegraphics{plots/covid_nowcast/00_true_data.pdf}
+    \includegraphics{plots/covid_nowcast/00_true_data}
     \caption{Realisations.}
@@
     \begin{subfigure}[t]{0.48\textwidth}
-    \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
+    \includegraphics{plots/covid_nowcast/00_nowcast_data}
     \caption{Same-day nowcasts.}
@@
 For information on the model design, we refer to~\citet{Wolffram2023}.
-In addition to those nowcasts, two ensemble methods are constructed using the ensembles' mean or median.
+In addition to those nowcasts, \citet{Wolffram2023} construct two ensemble methods using the ensembles' mean or median.
 We denote them by ENS-MEAN and ENS-MED.
@@
     \input{plots/covid_nowcast/01_model_scores}
-    \caption[The trending ratios for Covid nowcasting.]{Point evaluation measures for the issued mean of the different models. The evaluation period comprises 159 days. }
+    \caption[Point evaluation measures for the issued mean of the different models in Covid nowcasting.]{Point evaluation measures for the issued mean of the different models in Covid nowcasting.
+The evaluation period comprises 159 days.
+}
     \label{tab:app-covid-rmse}
@@
 In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
-We report the trending for the horizons 1, 7, and 14 days.
+We report the trending for the horizons one, seven, and 14 days.
 While horizon one assesses the short-term trending, horizons seven and 14 consider the medium-term trending.
@@
 Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} presents marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
-Overall, the variability and general level of differences grow with the horizon.
+The variability and general level of differences grow with the horizon.
 The standard deviation increases from roughly 300 for horizon one to 1,200 for horizon seven and 2,000 for horizon 14 days.
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption[The trending ratios for Covid nowcasting.]{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days.
+    \caption[The trending ratios for Covid nowcasting.]{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days in Covid nowcasting.
 The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
@@
     \end{subfigure}
-    \caption[Conditional trending plot and trending ratio over exclusion area for the Covid nowcasts.]{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
+    \caption[Conditional trending plot and trending ratio over exclusion area for the Covid nowcasts.]{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days in Covid nowcasting.}
     \label{fig:app-covid-cond-prob-trending-ratio-7}
@@
 If RKI or ILM issues a fall in the hospitalization rate, the probability of a fall is higher than if RIVM or ENS-MED issues a fall.
-The opposite is the case for a nowcasted hospitalization rate increase, and the difference between the models' performance is higher.
+The opposite is the case for a nowcasted hospitalization rate increase, and the difference between the models' performance is larger than for a fall.
 Similar observations can be made for the horizon of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
@@
 For each horizon $l$, $10,000$ samples of the forecast date $t$ and the forecast date $t-l$ based on the nowcasts of issue date $t$ are generated, and the proportion of positive changes is computed (see Appendix~\ref{subsec:app-probabilistic-trending-evaluation}).
-Remember that low a \ac{bs} and a plot along the diagonal are signs of good trending.
-The \ac{bs} is the lowest for the RIVM model, while the ENS-MED model has the best \ac{bs} for the horizon of seven and 14 days.
+Remember that a low \ac{bs} and a reliability diagram along the diagonal are signs of good trending.
+The \ac{bs} is the lowest for the RIVM model for a one-day-horizon, while the ENS-MED model has the best \ac{bs} for the horizon of seven and 14 days.
 The RKI model yields the highest \ac{bs} for all horizons.
-Note that the \ac{bs} is bounded by 0.25 for random guessing; thus, all models perform better than random guessing.
+Note that the \ac{bs} is 0.25 for random guessing; thus, all models perform better than random guessing.
 The reliability diagrams show that the models are not well calibrated for the horizon of one day.
@@
 Figure~\ref{fig:app-covid-prob-hist} shows the observed predicted increase probabilities for all horizons.
-For the horizons of seven and 14 days, the nowcasters issue only a few moderate probabilities.
+For the horizons of seven and 14 days, the nowcasters issue only a few moderate probabilities, and most probabilities are near zero and one.
 
@@
     \end{subfigure}\hspace{0.01\textwidth}%
-    \caption[Brier Scores and reliability diagrams for ILM, RKI, RIVM, and ENS-MED.]{\aclp{bs} and reliability diagrams for ILM, RKI, RIVM, and ENS-MED.
+    \caption[Brier Scores and reliability diagrams for the Covid nowcasting models ILM, RKI, RIVM, and ENS-MED.]{\aclp{bs} and reliability diagrams for the Covid nowcasting models ILM, RKI, RIVM, and ENS-MED.
     The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
@@
 In general, the reliability diagram cannot provide information if specific probabilities are issued scarcely.
-Thus, the reliability diagrams for the horizons of 7 and 14 days do not contain information on moderate probabilities.
+Thus, the reliability diagrams for the horizons of seven and 14 days do not contain information on moderate probabilities.
 The \ac{bs} values, however, work well for those examples and provide a good measure for the trending ability of the models.
@@
 \input{plots/ed_arrival/00_point_evaluation_measures}
-\caption[Point evaluation measures for the emergency department arrival forecasting models.]{Point evaluation measures for the models.
+\caption[Point evaluation measures for the emergency department arrival forecasting models.]{Point evaluation measures for the emergency department arrival forecasting models.
 The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.}\label{tab:app-eda-point-evaluation}
@@
 The 10\%-quantile of the differences is between zero and one for all models and horizons.
-Thus, we exclude differences smaller than one from the trending assessment.
+Thus, we use an exclusion area of size 1.
 The resulting fraction of included values in the computation is also listed in Table~\ref{tab:app-eda-marginals} and is at least 79\% of the values.
@@
     \input{plots/ed_arrival/10_marginal_analysis}
-    \caption[Marginal analysis of the forecast and true differences in ermergency department arrival forecasting.]{Marginal analysis of the forecast and true differences.
+    \caption[Marginal analysis of the forecast and true differences in emergency department arrival forecasting.]{Marginal analysis of the forecast and true differences in emergency department arrival forecasting.
 The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.
@@
 The model's relative trending ability evolves consistently for the two horizons, with the NBI-2 and Poisson-2 models being indistinguishable.
-The GBM-2 model outperforms the qreg-1 model for all $x$.
-The models NBI-2 and Poisson-2 have the highest trending ability for all negative values of $x$ and the lowest trending ability for all positive values of $x$.
-Benchmark-1 lies between the other models for all $x$.
+The GBM-2 model outperforms the qreg-1 model for all predicted changes.
+The models NBI-2 and Poisson-2 have the highest trending ability for all negative predicted changes and the lowest for all positive predicted changes.
+Benchmark-1 lies between the other models for all predicted changes.
 
 Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
-The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days in Figures~\ref{fig:app-eda-prob-rel-3} and \ref{fig:app-eda-prob-rel-7}.
+The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days in Figures~\ref{fig:app-eda-prob-rel-3} and~\ref{fig:app-eda-prob-rel-7}.
 The \acp{bs} are smallest for NBI-2 and Poisson-2 for both horizons, while the \acp{bs} for the other models are larger and differ more.
@@
     \input{plots/ed_arrival/50_trending_ratio}
-    \caption[Trending ratios for the emergency deparment arrival forecasts.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days.}
+    \caption[Trending ratios for the emergency department arrival forecasts.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days in emergency department arrival forecasting.}
     \label{tab:app-eda-trending-ratios}
@@
     \end{subfigure}
-    \caption[Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability in emergency department arrival forecasting.]{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability.
+    \caption[Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability in emergency department arrival forecasting.]{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability in emergency department arrival forecasting.
 The plots of NBI-2 and Poisson-2 are indistinguishable.}
@@
     \end{subfigure}
-    \caption[Probabilistic trending evaluation in emergency department arrival forecasting.]{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days.
+    \caption[Probabilistic trending evaluation in emergency department arrival forecasting.]{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days in emergency department arrival forecasting.
 The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic trending simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically.
@@
 
-The example shows that trending assessment is detached from standard point evaluation measures.
+Overall, the example provides performance assessments that are different from standard point evaluation measures and thus provide further insights into the strengths and weaknesses of the models.
 While the models with the lowest RMSE, NBI-2, and Poisson-2, also have a high trending ability, three models with below-average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive trending ability.
@@
     \caption[Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements.
-]{Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements.
+]{Four-quadrant plots for the different horizons $l$ and the systolic and mean blood pressure measurements.
 The upper row contains systolic measurements, and the lower row contains mean measurements.
@@
     \input{plots/mimic/trending_ratio}
-    \caption{Trending ratios for the different horizons and the systolic and mean blood pressure measurements.}
+    \caption[Trending ratios for the different horizons $l$ and the systolic and mean blood pressure measurements.]{Trending ratios for the different horizons $l$ and the systolic and mean blood pressure measurements.}
     \label{tab:app-mimic-trending-ratios}
@@
 
-\section{Declarations}
-\subsection{Availability of data and materials}
+\section*{Declarations}
+\subsection*{Availability of data and materials}
 
@@
 
-\subsection{Competing interests}
+\subsection*{Competing interests}
 The authors declare that they have no competing interests.
 
-\subsection{Funding}
+\subsection*{Funding}
 
@@
 
-\subsection{Authors' contributions}
+\subsection*{Authors' contributions}
 
@@
 
-\subsection{Acknowledgments}
+\subsection*{Acknowledgments}
 
@@
     \begin{subtable}{.48\textwidth}
+		\centering
         \input{plots/illustrative_examples/bootstrap_butterfly_cis}
@@
     \begin{subtable}{.48\textwidth}
+		\centering        
 		\input{plots/illustrative_examples/bootstrap_normal_cis}
@@
 While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.
-The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = x)$.
+The dashed line shows the theoretical $P(\diffyrv \diffxrv > 0 | \diffxrv = \xcond)$.
 The second method, cross-validation least squares, requires long computation times while yielding small or no bandwidth results, even for two relatively small datasets.
 The rule-of-thumb and cross-validation maximum likelihood methods yield reasonable results at moderate computation times.
-Further examples, including comparisons between methods regarding their trending ability, are available in Section~\ref{sec:application}.
 
@@
 Cross-validation least squares takes a considerably larger computation time.
-It does not converge for the first data set and yields a bandwidth too small for the second data set.
+It does not converge for the first data set and the second data set with exclusion area and yields a bandwidth too small for the second data set.
 The rule of thumb is the fastest method but tends to oversmooth.
@@
 
+Section~\ref{subsec:trending-probabilistic} outlines the probabilistic evaluation of trending for nowcasts and forecasts and specifies the computation for predictions in terms of a \ac{cdf} and known true values.
+Here, we outline the computation for quantile forecasts and probabilistic true values.
 If forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
@@
     \end{tabular}
-    \caption[Matching the abbreviation to the key in the nowcasting hub.]{Matching the abbreviation to the key in the nowcasting hub.
+    \caption[Matching the abbreviation to the key in the Covid nowcasting hub.]{Matching the abbreviation to the key in the Covid nowcasting hub.
     Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
@@
     \input{plots/covid_nowcast/10_marginal_analysis}
-    \caption[Marginal analysis of the Covid nowcast and true differences for the horizons 1, 7, and 14 days.]{Marginal analysis of the nowcast and true differences for the horizons 1, 7, and 14 days.
+    \caption[Marginal analysis of the Covid nowcast and true differences for the horizons one, seven, and 14 days.]{Marginal analysis of the Covid nowcast and true differences for the horizons one, seven, and 14 days.
     The column (1), $l=l$ shows the number of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
@@
     \end{subtable}
-    \caption[Trending ratios for the models with and without exclusion areas for the horizon 1 and 14 days in Covid nowcasting.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the horizon 1 and 14 days.
+    \caption[Trending ratios for the models with and without exclusion areas for the horizon one and 14 days in Covid nowcasting.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the horizon 1 and 14 days in Covid nowcasting.
 The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values.
@@
 \includegraphics{plots/covid_nowcast/30_4q_plots}
-\caption[Four-quadrant plots for the Covid nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.]{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.
+\caption[Four-quadrant plots for the Covid nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.]{Four-quadrant plots for the Covid nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.
 The spread in both directions increases with the horizon.
@@
     \end{subfigure}
-    \caption[Conditional trending plot and trending ratio over exclusion area in Covid nowcasting.]{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
+    \caption[Conditional trending plot and trending ratio over exclusion area in Covid nowcasting.]{Conditional trending plot and trending ratio over exclusion area for the Covid nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
     \label{fig:app-covid-cond-prob-trending-ratio-1-14}
@@
     \end{subfigure}
-    \caption[Reliability diagrams for Covid nowcasting models and the horizon seven and 14 days.]{The reliability diagram for the models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
+    \caption[Reliability diagrams for Covid nowcasting models and the horizon seven and 14 days.]{The reliability diagram for the Covid nowcasting models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
     Additionally, the count of predicted probabilities for the horizons is shown.

### 2024-04-02 paper.tex
Add mail

@@
 %\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
-\author*[1]{Jonas Rieger}
+\author*[1]{Jonas Rieger}\email{jonas.rieger@kit.edu}
 \author[1]{Bolin Liu}
@@
 An intuitive and easily interpretable information on a quantity measured over time is whether it increases or decreases.
-If the quantity of interest is not yet accessible, we refer to the ability to accurately predict, nowcast, or measure the direction of change for a true variable as trending. 
+We refer to the ability to accurately predict, nowcast, or measure the direction of change of a true variable as trending. 
 An appropriate assessment of the trending ability is crucial for decision-makers, as it enables them to assess the method's credibility and changes in the underlying process early.

### 2024-04-08 paper.tex
Update on Overleaf.

@@
 An intuitive and easily interpretable information on a quantity measured over time is whether it increases or decreases.
-We refer to the ability to accurately predict, nowcast, or measure the direction of change of a true variable as trending. 
-An appropriate assessment of the trending ability is crucial for decision-makers, as it enables them to assess the method's credibility and changes in the underlying process early.
-While the current practice of evaluating signals is focused on averaging over local deviation measures, we present trending evaluation techniques as accompanying assessments that consider whether the development direction fits the quantity of interest.
+We refer to the ability to accurately predict, nowcast, or measure the direction of change of a true variable (value???) as "trending". 
+Trending describes the ability of a method to measure or predict the right direction [and magnitude???] of changes in a true variable (value???).
+Currently, evaluating signals to assess trending  is focused on averaging over local deviation measures.
+We here present techniques to assess trending as accompanying assessments that consider whether the development direction [direction of change???] fits the quantity of interest.
 We formalize the concept of trending for all application areas and present quantitative measures. 
@@
 We review and present new visualization techniques to assess the local trending behavior and enlarge the scope to probabilistic nowcasts and forecasts. 
-We apply the methods to three large medical data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+We exemplarily apply the methods to three medical or healthcare problems: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
 }
@@
 The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
-Conventionally, differences between predictions (or measurements) and target values, such as the \ac{rmse} or \ac{mae}, form the basis of method evaluation.
-These measures do not consider trending, that is, whether the right direction of change is predicted or measured.
-However, the information about an increase or decrease is crucial to detect changes in the underlying process early.
+Conventionally, differences between predictions (or measurements) and target variables (values???), such as the \ac{rmse} or \ac{mae}, form the basis of method evaluation.
+These measures do not consider trending, that is, the ability of a method to measure or predict the right direction [and magnitude???] of changes in a true variable (value???). 
+However, the information about an increase or decrease in a true variable (value???) is crucial to detect changes in underlying processes.
 In the following, we provide a more specific overview of the characteristics and current evaluation schemes of the three application fields, forecasting, nowcasting, and measurement, and highlight why trending evaluation is highly relevant in the respective fields.
@@
 
-Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past~\citep{banbura2013now}.
+Methodologically evolved from forecasting, nowcasting methods focus on predictions for the present, the immediate future, and the recent past~\citep{banbura2013now}.
 Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours~\citep{browning1989nowcasting,schmid2019nowcasting}.
@@
 Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse}~\citep{gunther2021nowcasting} or probabilistic scoring rules~\citep{Wolffram2023}.
-The aspect of trending ability is not considered in the literature to the best of our knowledge.
+The aspect of trending is not considered in the literature to the best of our knowledge.
 However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
@@
 
-The main contributions of this paper are manifold.
-We formalize trending and present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
+In this article, we formalize trending and present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
 We introduce the conditional trending plot, a new graphical method for assessing local trending behavior, and review bootstrap methods for calculating confidence intervals.

### 2024-04-10 paper.tex
Update on Overleaf.
Change wording and shorten

--- 2024-04-02 paper.tex
@@
 %\bibliographystyle{
-\addbibresource{library.bib}
+%\addbibresource{library.bib}
 \addbibresource{library-jonas.bib}
@@
 \DeclareAcronym{tadda}{short=TADDA, long=targeted absolute deviation with direction augmentation}
+\DeclareAcronym{atc}{short=ATC, long=ability to track changes}
+\DeclareAcronym{aatc}{short=AATC, long=assessment of the ability to track changes}
 
@@
 
-\title{Trending assessment for nowcasts, measurements, and forecasts}
+\title{Assessment of the ability to track changes for nowcasts, measurements, and forecasts}
 %\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
@@
 \abstract{
-An intuitive and easily interpretable information on a quantity measured over time is whether it increases or decreases.
-We refer to the ability to accurately predict, nowcast, or measure the direction of change of a true variable as trending. 
-An appropriate assessment of the trending ability is crucial for decision-makers, as it enables them to assess the method's credibility and changes in the underlying process early.
-While the current practice of evaluating signals is focused on averaging over local deviation measures, we present trending evaluation techniques as accompanying assessments that consider whether the development direction fits the quantity of interest.
-We formalize the concept of trending for all application areas and present quantitative measures. 
-Using bootstrap confidence intervals and exclusion areas, we provide extensions for noisy data and estimation uncertainty.
-We review and present new visualization techniques to assess the local trending behavior and enlarge the scope to probabilistic nowcasts and forecasts. 
-We apply the methods to three large medical data applications: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+Measurements, nowcasts, and forecasts aim to indicate a value of interest's change through direction and magnitude.
+The direction of change alone is easily and intuitively interpretable information on the development and can be an early sign of regime switching.
+We refer to a measurement's, nowcast's, or forecast's ability to predict the right direction of development as the ability to track changes.
+We present techniques to assess this ability as accompanying evaluation in all application areas, including quantitative measures.
+We provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
+We review and present new visualization techniques to assess the local ability to track changes and enlarge the scope to probabilistic nowcasts and forecasts.
+We exemplify the methods for three medical or healthcare problems: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
 }
@@
 
-\section{Introduction}\label{sec:trending-introduction}
+\section{Introduction}\label{sec:atc-introduction}
 
 The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
-Conventionally, differences between predictions (or measurements) and target values, such as the \ac{rmse} or \ac{mae}, form the basis of method evaluation.
-These measures do not consider trending, that is, whether the right direction of change is predicted or measured.
-However, the information about an increase or decrease is crucial to detect changes in the underlying process early.
-In the following, we provide a more specific overview of the characteristics and current evaluation schemes of the three application fields, forecasting, nowcasting, and measurement, and highlight why trending evaluation is highly relevant in the respective fields.
+Although the three areas differ in the framework and methods used, they share the goal of predicting an unknown value, and thus, similar techniques can be used in the evaluation.
 
 Forecasting methods predict the future based on historical data, patterns, and exogenous factors.
-In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in][]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008,Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
+In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in][]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008, Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
 The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
-In the current practice, a forecasting method is usually evaluated in terms of its prediction accuracy, measuring the deviation of the prediction from the actual outcome of the target variable.
-Popular measures are scale-dependent measures such as the \ac{rmse}, measures based on percentage errors such as the root mean squared percentage error, or probabilistic scoring rules~\citep[see the review in][]{hyndman2006another}.
-They are not capable of assessing the trending ability discussed above.
-In a prediction competition on armed conflicts, trending recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with a trending component for evaluation.
+Methodologically evolved from forecasting~\citep{Browning1989}, nowcasting methods focus on predictions for the present, the immediate future, and the recent past~\citep{Banbura2013, WorldMeteorologicalOrganizationWMO2017}.
+Nowcasting methods use high-frequency indicators related to the target value and estimate the value of a target value for a specific time based on current preliminary measurements, which are finalized with a considerable delay.
+Thus, it focuses on ongoing predictions due to updating information for the same time~\citep{Castle2017}.
+In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{Wu2021} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{Gunther2021, Wolffram2023}.
+In forecasting and nowcasting, the evaluation is based on accuracy measures such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
+None of them assesses the method's \ac{atc}.
+In a prediction competition on armed conflicts, the \ac{aatc} recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-This paper focuses on pure trending evaluation that accompanies traditional scores.
-
-Methodologically evolved from forecasting; nowcasting methods focus on predictions for the present, the immediate future, and the recent past~\citep{banbura2013now}.
-Nowcasting has its origins in meteorology, and the methods were initially developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours~\citep{browning1989nowcasting,schmid2019nowcasting}.
-In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{wu2021nowcasting} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{gunther2021nowcasting}.
-Nowcasting methods use high-frequency indicators related to the target variable and estimate the value of a target variable for a specific time based on current preliminary measurements, which are finalized with a considerable time delay.
-Thus, the nowcasts can produce early and ongoing estimates for the target variable during the relevant period~\citep{castle2017forecasting}.
-Like forecasts, nowcasts are often evaluated and compared in the literature based on performance measures such as the \ac{rmse}~\citep{gunther2021nowcasting} or probabilistic scoring rules~\citep{Wolffram2023}.
-The aspect of trending ability is not considered in the literature to the best of our knowledge.
-However, trending evaluation adds valuable information on the methods' capability of predicting the development of the target variable, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
-The epidemic's development, in turn, can be the foundation of decisions on introducing or canceling policy measures.
+This paper focuses on pure \ac{aatc} accompanying traditional scores.
+The \ac{aatc} adds valuable information on the methods' capability of predicting the development of the target value, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
 
 Measurement is more technical and aims to obtain accurate and reliable data about the current state of a system based on sensors.
-A parameter or variable can be measured regularly over a certain period to evaluate the system's development.
+A parameter or value can be measured regularly over a certain period to evaluate the system's development.
 When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
-Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods~\citep{lawrence1989concordance,koo2016guideline} in addition to paired t-tests to determine systematic differences between two measurement series~\citep{watson2010method}.
-Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value~\citep{bland1986statistical}.
-Trending considerations, that is, the consistency of the test method's development with the gold standard, is a field of active research in the last years~\citep{Saugel2015,saugel2018error,hiraishi2021concordance}.
-
-The main contributions of this paper are manifold.
-We formalize trending and present different variants of trending measures that consider either noiseless data or data with noise and small non-informative changes.
-We introduce the conditional trending plot, a new graphical method for assessing local trending behavior, and review bootstrap methods for calculating confidence intervals.
-We extend the concept of trending to probabilistic predictions.
-The methods are applied to three applications from the abovementioned fields: COVID-19-nowcasting, emergency department arrival forecasting, and blood pressure measurement.
-The ready-to-use code for trending evaluation from the applications is available on ... .
-The trending evaluation methods are intended to complement the current evaluation and reveal additional information on the methods considered.
-
-The remainder of this paper is structured as follows.
-Section~\ref{sec:trending} formalizes trending and investigates several extensions, such as noise-aware methods, confidence intervals, and probabilistic forecasts and nowcasts.
-We introduce a new graphical method for evaluating trending, the conditional trending plot, which can simultaneously assess different intervals and asymmetries.
-In Section~\ref{sec:application}, we show the results of applying our method to several practical examples from measurement, forecasting, and nowcasting.
-Hence, we provide blueprints for applying trending evaluation in practice and make remarks on interpreting the measures.
-We conclude the paper in Section~\ref{sec:trending-conclusion}.
-
-\section{Trending assessment}\label{sec:trending}
-
-As discussed in Section~\ref{sec:trending-introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by a measurement, a nowcast, or a forecast.
+Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods~\citep{Lin1989, Koo2016} in addition to paired t-tests to determine systematic differences between two measurement series~\citep{Watson2010}.
+Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value~\citep{Bland1986}.
+The \ac{aatc} is a field of active research in recent years~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
+
+In this article, we formalize \ac{atc} and present different variants of \acs{aatc} measures that consider either noiseless data or data with noise and small non-informative changes (see Section~\ref{sec:aatc}).
+We introduce the conditional \ac{atc} plot, a new graphical method for assessing local \ac{atc} behavior, and review bootstrap methods for calculating confidence intervals (see Section~\ref{subsec:aatc-noise} and ~\ref{subsec:aatc-cond-prob}).
+We extend the concept of \ac{aatc} to probabilistic predictions in Section~\ref{subsec:aatc-probabilistic}.
+The methods are applied to three applications from the abovementioned fields: COVID-19-nowcasting, emergency department arrival forecasting, and blood pressure measurement, and, thus, provide blueprints for future assessments.
+The ready-to-use code for \ac{aatc} from the applications is available on [LINK WILL BE INSERTED].
+
+\section{\ac{aatc}}\label{sec:aatc}
+
+% Einleitung weglassen; näher an Definition von oben.
+As discussed in Section~\ref{sec:atc-introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by a measurement, a nowcast, or a forecast.
 In this section, we want to specify trending and define different methods of measuring and visualizing trending.
@@
 In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
-The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:trending-noise}, and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
-Section~\ref{subsec:trending-cond-prob} introduces a new graphical method for local trending assessment.
-Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:trending-probabilistic}.
+The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:aatc-noise}, and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
+Section~\ref{subsec:aatc-cond-prob} introduces a new graphical method for local trending assessment.
+Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:aatc-probabilistic}.
 
@@
 Consequently, when evaluating the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
-A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{perrino1998intraoperative,Saugel2015}.
+A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{Perrino1998,Saugel2015}.
 In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
@@
 
-\subsection{Accounting for noise and non-informative small changes and bootstrapping confidence intervals}\label{subsec:trending-noise}
+\subsection{Accounting for noise and non-informative small changes and bootstrapping confidence intervals}\label{subsec:aatc-noise}
 
@@
 
-\subsection{The conditional trending plot}\label{subsec:trending-cond-prob}
+\subsection{The conditional trending plot}\label{subsec:aatc-cond-prob}
 
@@
 
-\subsection{Probabilistic Evaluation}\label{subsec:trending-probabilistic}
+\subsection{Probabilistic evaluation}\label{subsec:aatc-probabilistic}
 
@@
 
-The examples illustrate the use of the methods of Section~\ref{sec:trending} and yield further insights into the measures and plots.
+The examples illustrate the use of the methods of Section~\ref{sec:aatc} and yield further insights into the measures and plots.
 All code used to generate the analysis is available online.
@@
 
-In the following, we apply the trending assessment of Section~\ref{sec:trending} to the nowcasts of the seven-day hospitalization rate.
+In the following, we apply the trending assessment of Section~\ref{sec:aatc} to the nowcasts of the seven-day hospitalization rate.
 We report the trending for the horizons one, seven, and 14 days.
@@
 
-\section{Discussion and Conclusion}\label{sec:trending-conclusion}
+\section{Discussion and conclusion}\label{sec:trending-conclusion}
 
@@
 
-A conditional trending plot visualizes the probability of correct trending over the predicted change of the signal (see Section~\ref{subsec:trending-cond-prob}).
+A conditional trending plot visualizes the probability of correct trending over the predicted change of the signal (see Section~\ref{subsec:aatc-cond-prob}).
 It is based on a multivariate \acf{kde} of predicted and observed change.
 In the application, the conditional trending plot gives reasonable insights into the local effects of the trending ability.
-Section~\ref{subsec:trending-probabilistic} adapts measures of probabilistic forecast evaluation to the trending assessment of probabilistic forecasts and nowcasts.
+Section~\ref{subsec:aatc-probabilistic} adapts measures of probabilistic forecast evaluation to the trending assessment of probabilistic forecasts and nowcasts.
 The \acf{bs} as numerical assessment of probabilistic trending is introduced, and reliability diagrams are used to visualize the local trending ability of probabilistic forecasts.
@@
 
-\section{Additional material on Section~\ref{sec:trending}}\label{sec:appendix-trending}
-
-\subsection{Data generation for Section~\ref{sec:trending}}\label{subsec:app-trending-data-generation}
+\section{Additional material on Section~\ref{sec:aatc}}\label{sec:appendix-trending}
+
+\subsection{Data generation for Section~\ref{sec:aatc}}\label{subsec:app-trending-data-generation}
 
@@
     \end{subfigure}
-    \caption[Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$.]{Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$.
+    \caption[Boxplot of the computation time for different bootstrapping methods and data set sizes.]{Boxplot of the computation time for different bootstrapping methods and data set sizes $T$.
 The computation time refers to bootstrapping one confidence interval based upon $10,000$ values.
@@
 
-Section~\ref{subsec:trending-probabilistic} outlines the probabilistic evaluation of trending for nowcasts and forecasts and specifies the computation for predictions in terms of a \ac{cdf} and known true values.
+Section~\ref{subsec:aatc-probabilistic} outlines the probabilistic evaluation of trending for nowcasts and forecasts and specifies the computation for predictions in terms of a \ac{cdf} and known true values.
 Here, we outline the computation for quantile forecasts and probabilistic true values.
@@
 
-\section{Additional material on Section~\ref{sec:application-covid}}\label{sec:appendix-application-covid}
+\section{Additional results for Section~\ref{sec:application-covid}}\label{sec:appendix-application-covid}
 
@@
     \end{subtable}
-    \caption[Trending ratios for the models with and without exclusion areas for the horizon one and 14 days in Covid nowcasting.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the horizon 1 and 14 days in Covid nowcasting.
+    \caption[Trending ratios for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.
 The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values.

### 2024-04-10 paper.tex
Update on Overleaf.

@@
 
-\abstract{
-Measurements, nowcasts, and forecasts aim to indicate a value of interest's change through direction and magnitude.
-The direction of change alone is easily and intuitively interpretable information on the development and can be an early sign of regime switching.
-We refer to a measurement's, nowcast's, or forecast's ability to predict the right direction of development as the ability to track changes.
+\abstract{ % Weniger We; Erste zwei Sätze kombinieren
+In addition to the numerical precision, the direction of change alone is easily and intuitively interpretable information on the development.
+We consider whether measurements, nowcasts, and forecasts are able to correctly indicate the direction of change of a value of interest.
+We refer to the ability to predict the right direction of development as the ability to track changes.
+
 We present techniques to assess this ability as accompanying evaluation in all application areas, including quantitative measures.

### 2024-04-10 paper.tex
Update on Overleaf.

@@
 \abstract{ % Weniger We; Erste zwei Sätze kombinieren
-In addition to the numerical precision, the direction of change alone is easily and intuitively interpretable information on the development.
+In addition to the numerical precision of measurements, nowcasts, and forecasts, their ability to reflect changes in the underlying values can be crucial in applications.
+We refer to the ability to predict the right direction of development as the ability to track changes and present visual techniques and quantitative measures to assess this ability.
+We provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
+Three medical or healthcare problems illustrate the assessment: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+
+In addition to the numerical precision, the direction of change alone is easily and intuitively interpretable information on the development of a value of interest.
 We consider whether measurements, nowcasts, and forecasts are able to correctly indicate the direction of change of a value of interest.
-We refer to the ability to predict the right direction of development as the ability to track changes.
 

### 2024-04-10 paper.tex
Update on Overleaf.

--- 2024-04-10 paper.tex
@@
 
-\abstract{
-Measurements, nowcasts, and forecasts aim to indicate a value of interest's change through direction and magnitude.
-The direction of change alone is easily and intuitively interpretable information on the development and can be an early sign of regime switching.
-We refer to a measurement's, nowcast's, or forecast's ability to predict the right direction of development as the ability to track changes.
-We present techniques to assess this ability as accompanying evaluation in all application areas, including quantitative measures.
-We provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
-We review and present new visualization techniques to assess the local ability to track changes and enlarge the scope to probabilistic nowcasts and forecasts.
-We exemplify the methods for three medical or healthcare problems: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+\abstract{ % Weniger We; Erste zwei Sätze kombinieren
+In addition to the numerical precision of measurements, nowcasts, and forecasts,
+
+it can be crucial in applications
+
+it is of interest
+
+how well these reflect changes in the underlying values.
+In this article, we focus on the ability to predict the correct direction of changes and refer to it as the ability to track changes.
+We present visual techniques and quantitative measures to assess this ability and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
+Three medical or healthcare problems illustrate the assessment: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
 }

### 2024-04-10 paper.tex
Update on Overleaf.

@@
 \abstract{ % Weniger We; Erste zwei Sätze kombinieren
-In addition to the numerical precision of measurements, nowcasts, and forecasts,
-
-it can be crucial in applications
-
-it is of interest
-
-how well these reflect changes in the underlying values.
-In this article, we focus on the ability to predict the correct direction of changes and refer to it as the ability to track changes.
+In addition to measurements, nowcasts, or forecasts being numerically precise, it is essential how well they reflect changes in the underlying values.
+In this article, we focus on predicting the correct direction of changes and refer to it as the ability to track changes.
+
+In this article, we focus on whether they predict the correct direction of changes and refer to it as the ability to track changes.
+
 We present visual techniques and quantitative measures to assess this ability and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
-Three medical or healthcare problems illustrate the assessment: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+Three medical or healthcare problems illustrate the methods: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
 }

### 2024-04-10 paper.tex
Update on Overleaf.

@@
 In addition to measurements, nowcasts, or forecasts being numerically precise, it is essential how well they reflect changes in the underlying values.
-In this article, we focus on predicting the correct direction of changes and refer to it as the ability to track changes.
-
-In this article, we focus on whether they predict the correct direction of changes and refer to it as the ability to track changes.
-
+In this article, we focus on evaluating the ability to predict the correct direction of changes, which we refer to as the ability to track changes. 
 We present visual techniques and quantitative measures to assess this ability and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
-Three medical or healthcare problems illustrate the methods: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+Worked through examples illustrate the assessment using the proposed methods: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
 }
@@
 The \ac{aatc} adds valuable information on the methods' capability of predicting the development of the target value, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
+% Measurement noch integrieren
 

### 2024-04-10 paper.tex
Update on Overleaf.

--- 2024-04-10 paper.tex
@@
 \author[1]{Bolin Liu}
+\author[2]{Bernd Saugel}
 \author[1]{Oliver Grothe}
@@
 
-\abstract{ % Weniger We; Erste zwei Sätze kombinieren
-In addition to the numerical precision of measurements, nowcasts, and forecasts,
-
-it can be crucial in applications
-
-it is of interest
-
-how well these reflect changes in the underlying values.
-In this article, we focus on the ability to predict the correct direction of changes and refer to it as the ability to track changes.
+\abstract{ 
+In addition to measurements, nowcasts, or forecasts being numerically precise, it is essential how well they reflect changes in the underlying values.
+In this article, we focus on evaluating the ability to predict the correct direction of changes, which we refer to as the ability to track changes (ATC). 
 We present visual techniques and quantitative measures to assess this ability and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
-Three medical or healthcare problems illustrate the assessment: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+Worked-through examples illustrate the assessment using the proposed methods: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
 }
@@
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-This paper focuses on pure \ac{aatc} accompanying traditional scores.
-The \ac{aatc} adds valuable information on the methods' capability of predicting the development of the target value, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
+
+% Measurement weniger abgetrennt
 
@@
 The \ac{aatc} is a field of active research in recent years~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
+
+% AATC weg als Abkürzung
+
+This paper focuses on pure \ac{aatc} accompanying traditional scores.
+The \ac{aatc} adds valuable information on the methods' capability of predicting the development of the target value, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
 

### 2024-04-10 paper.tex
Update on Overleaf.

@@
 This paper focuses on pure \ac{aatc} accompanying traditional scores.
-The \ac{aatc} adds valuable information on the methods' capability of predicting the development of the target value, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
-
-In this article, we formalize \ac{atc} and present different variants of \acs{aatc} measures that consider either noiseless data or data with noise and small non-informative changes (see Section~\ref{sec:aatc}).
+The assessment of \ac{atc} adds valuable information on the methods' capability of predicting the development of the target value, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
+
+In this article, we formalize \ac{atc} and present different variants of assessment of \ac{atc} measures that consider either noiseless data or data with noise and small non-informative changes (see Section~\ref{sec:aatc}).
 We introduce the conditional \ac{atc} plot, a new graphical method for assessing local \ac{atc} behavior, and review bootstrap methods for calculating confidence intervals (see Section~\ref{subsec:aatc-noise} and ~\ref{subsec:aatc-cond-prob}).
-We extend the concept of \ac{aatc} to probabilistic predictions in Section~\ref{subsec:aatc-probabilistic}.
+We extend the concept of assessment of \ac{atc} to probabilistic predictions in Section~\ref{subsec:aatc-probabilistic}.
 The methods are applied to three applications from the abovementioned fields: COVID-19-nowcasting, emergency department arrival forecasting, and blood pressure measurement, and, thus, provide blueprints for future assessments.
-The ready-to-use code for \ac{aatc} from the applications is available on [LINK WILL BE INSERTED].
-
-\section{\ac{aatc}}\label{sec:aatc}
+The ready-to-use code for the assessment of \ac{atc} from the applications is available on [LINK WILL BE INSERTED].
+
+\section{Assessment of the \acf{atc}}\label{sec:aatc}
 

### 2024-04-10 paper.tex
Update on Overleaf.

@@
 \DeclareAcronym{atc}{short=ATC, long=ability to track changes}
-\DeclareAcronym{aatc}{short=AATC, long=assessment of the ability to track changes}
+\DeclareAcronym{mnf}{short=MNF,}
 
@@
 In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{Wu2021} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{Gunther2021, Wolffram2023}.
+Measurement aims to obtain accurate and reliable data about the current state of a system based on sensors.
+A parameter or value can be measured regularly over a certain period to evaluate the system's development.
+When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
 In forecasting and nowcasting, the evaluation is based on accuracy measures such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
 None of them assesses the method's \ac{atc}.
-In a prediction competition on armed conflicts, the \ac{aatc} recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
+In a prediction competition on armed conflicts, the assessment of \ac{atc} recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-
-% Measurement weniger abgetrennt
-
-Measurement is more technical and aims to obtain accurate and reliable data about the current state of a system based on sensors.
-A parameter or value can be measured regularly over a certain period to evaluate the system's development.
-When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
-Various indices, such as the interclass correlation coefficient or Lin's concordance correlation, have been proposed to check the reproducibility of the measurement or to compare different measurement methods~\citep{Lin1989, Koo2016} in addition to paired t-tests to determine systematic differences between two measurement series~\citep{Watson2010}.
-Graphical tools, such as the Bland-Altman diagram, are used to visualize the differences between the methods in relation to their mean value~\citep{Bland1986}.
-The \ac{aatc} is a field of active research in recent years~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
-
-% AATC weg als Abkürzung
-
-This paper focuses on pure \ac{aatc} accompanying traditional scores.
-The assessment of \ac{atc} adds valuable information on the methods' capability of predicting the development of the target value, for example, the number of cases of an epidemic, where increases can be an early sign of the beginning of a wave.
-
-In this article, we formalize \ac{atc} and present different variants of assessment of \ac{atc} measures that consider either noiseless data or data with noise and small non-informative changes (see Section~\ref{sec:aatc}).
+In measurement agreement analysis, a typical analysis includes the Bland-Altman analysis and a percentage error~\citep{Bland1986}.
+The assessment of \ac{atc} is a field of active research in recent years~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
+
+This article focuses on pure assessment of \ac{atc} accompanying traditional scores.
+ we formalize \ac{atc} and present different variants of assessment of \ac{atc} measures that consider either noiseless data or data with noise and small non-informative changes (see Section~\ref{sec:aatc}).
 We introduce the conditional \ac{atc} plot, a new graphical method for assessing local \ac{atc} behavior, and review bootstrap methods for calculating confidence intervals (see Section~\ref{subsec:aatc-noise} and ~\ref{subsec:aatc-cond-prob}).
@@
 
-% Einleitung weglassen; näher an Definition von oben.
-As discussed in Section~\ref{sec:atc-introduction}, the term \enquote{trending} denotes the alignment of a detected true development with the development indicated by a measurement, a nowcast, or a forecast.
-In this section, we want to specify trending and define different methods of measuring and visualizing trending.
-In general, trending can be assessed for different horizons $l$ separately, and the considered horizons' relevance must be determined based on the task.
-Throughout the section, we call a measurement, nowcast, or forecast a \textit{prediction} or \textit{signal} whenever the application area is irrelevant.
-Before introducing measures to evaluate trending, we remark on calculating the predicted change and notation in Section~\ref{subsec:notation}.
-Section~\ref{subsec:trending-four-quadrant-plot} introduces trending basics and visualizes trending in a synthetic example.
-In Section~\ref{subsec:trending-measures}, various methods from other disciplines and tasks are applied and examined.
-The approaches are extended to account for noise and non-systematic effects in Section~\ref{subsec:aatc-noise}, and bootstrap methods for computing confidence intervals for the measures in the previous sections are briefly introduced.
-Section~\ref{subsec:aatc-cond-prob} introduces a new graphical method for local trending assessment.
-Trending assessment for probabilistic nowcasts and forecasts is treated in Section~\ref{subsec:aatc-probabilistic}.
-
-\subsection{Computing changes and notation}\label{subsec:notation}
-
-As noted above, trending refers to observed and predicted changes over a time horizon $l$.
-The \textit{observed} change is straightforward to compute for all types of signals.
+% \subsection{Computing changes and notation}\label{subsec:notation}
+
+As noted above, the assessment of \ac{atc} is based on the observed and predicted changes over a time horizon $l$.
+The \textit{observed} change is straightforward to compute for all types of \ac{mnf}.
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values for nowcasting or forecasting, or gold standard measurements up to time $T$.

### 2024-04-10 paper.tex
Update on Overleaf.

@@
 \DeclareAcronym{atc}{short=ATC, long=ability to track changes}
-\DeclareAcronym{mnf}{short=MNF,}
+\DeclareAcronym{mnf}{short=MNF, long={measurement, nowcast, or forecast}}
 
@@
 
-\title{Assessment of the ability to track changes for nowcasts, measurements, and forecasts}
+\title{Assessment of the ability to track changes for measurements, nowcasts, and forecasts}
 %\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
@@
 
-\section{Assessment of the \acf{atc}}\label{sec:aatc}
+\section{Assessment of the ability to track changes (ATC)}\label{sec:aatc}
 
@@
     \caption[Computation of the predicted change in the different applications.]{Computation of the predicted change in the different applications.
-For nowcasting and forecasting, signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
+For nowcasting and forecasting, $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$.
 For measurement, $x_t$ denotes the test device measurement at time $t$.}
@@
 For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$.
-We call $t$ the \textit{forecast time} and $\tau$ the \textit{issue time}.
+We call $t$ the \textit{target time} and $\tau$ the \textit{issue time}.
 The predicted change is computed by
@@
 
-The notation is similar for forecasting: Let $x_{t | \tau}$ denote the forecast for forecast time $t$ and issue time $\tau$.
+The notation is similar for forecasting: Let $x_{t | \tau}$ denote the forecast for target time $t$ and issue time $\tau$.
 The predicted change is computed by
@@
 
-\subsection{Basics of trending and four-quadrant plots}\label{subsec:trending-four-quadrant-plot}
-
-Formally, the trending assessment is the same for all three application areas, given the notation for the respective application of Section~\ref{subsec:notation}.
-In the following, we omit the difference horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
-As noted above, a signal's trending is perfect if all the change directions are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
-Consequently, when evaluating the trending, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
+\subsection{The four-quadrant plot}\label{subsec:aatc-four-quadrant-plot}
+
+Formally, the assessment of \ac{atc} is the same for \ac{mnf}, given the notation for the respective application of Section~\ref{subsec:notation}.
+In the following, we omit the horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
+The \ac{atc} is maximal if all the change directions are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+Consequently, when assessing the \ac{atc}, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
 A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{Perrino1998,Saugel2015}.
@@
 Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant graph for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
-Data generation is described in Appendix~\ref{subsec:app-trending-data-generation}.
 The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
@@
 \end{subfigure}%
-\caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas. }
+\caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas.}
 \label{fig:trending_4q}
@@
 
-Let for a forecast $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for forecast time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
+Let for a forecast $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for target time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
 Thereby, $F_{t | t-l} (x)$ specifies the forecastes probability that the quantity of interest is at most $x$.
@@
 
-Four-quadrant plots facilitate the visual inspection of the trending ability for a signal (see Section~\ref{subsec:trending-four-quadrant-plot}).
+Four-quadrant plots facilitate the visual inspection of the trending ability for a signal (see Section~\ref{subsec:aatc-four-quadrant-plot}).
 The trending ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates trending.
@@
 If forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
-Let $q_p$ denote the quantiles for forecast time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
+Let $q_p$ denote the quantiles for target time $t+l$ for even-spaced probabilities $p \in \{1/\pmax, \dots, (p-1) / \pmax\}$ ($\pmax \in \N \setminus \{1, 2\}$) and $y_t$ the true value at time $t$.
 The quantiles $q_p$ generally differ for each time step, but we omit an index here for ease of notation.

### 2024-04-10 paper.tex
Update on Overleaf.

@@
 \abstract{ 
-In addition to measurements, nowcasts, or forecasts being numerically precise, it is essential how well they reflect changes in the underlying values.
+In addition to measurements, nowcasts, or forecasts being numerically precise, it is essential [to know] how well they reflect changes in the underlying values.
 In this article, we focus on evaluating the ability to predict the correct direction of changes, which we refer to as the ability to track changes (ATC). 
@@
 This article focuses on pure assessment of \ac{atc} accompanying traditional scores.
- we formalize \ac{atc} and present different variants of assessment of \ac{atc} measures that consider either noiseless data or data with noise and small non-informative changes (see Section~\ref{sec:aatc}).
+ We formalize \ac{atc} and present different variants of assessment of \ac{atc} measures that consider either noiseless data or data with noise and small non-informative changes (see Section~\ref{sec:aatc}).
 We introduce the conditional \ac{atc} plot, a new graphical method for assessing local \ac{atc} behavior, and review bootstrap methods for calculating confidence intervals (see Section~\ref{subsec:aatc-noise} and ~\ref{subsec:aatc-cond-prob}).
@@
 Points in the green upper right and lower left quadrants reflect a correct trending for the respective time step, whereas points in the remaining red quadrants show incorrect predicted changes.
-
-Figure~\ref{fig:trending_basic_4q} displays a basic four-quadrant plot.
-Points 2, 3, 5, and 6 show trending, whereas points 1, 4, and 7 count for anti-trending behavior.
-Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant graph for a more extensive simulated data set with $T=1461$, for example, four years of daily data.
-The four-quadrant plot resembles a butterfly-like shape, with more data in the upper right and lower left quadrants than in the remaining quadrants.
-
-The four-quadrant plot is intuitive to interpret, and the magnitude and direction of change are shown simultaneously.
+Figure~\ref{fig:trending_basic_4q} displays a basic four-quadrant plot and
+Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data.
+
 It can be extended by including information on the time index in the point color to reveal effects over time.
@@
 
-There are other visualization techniques, such as polar plots or the Bland-Altmann analysis.
-However, they lack the four-quadrant plot's clarity and intuition without adding more information on trending~\citep{Saugel2015}.
+The four-quadrant plot is intuitive to interpret, and the magnitude and direction of change are shown simultaneously.
+Other visualization techniques, such as polar plots, lack the four-quadrant plot's clarity and intuition without adding more information on the \ac{atc}~\citep{Saugel2015}.
 
@@
 \end{subfigure}%
-\caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas.}
+\caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas.
+In Figure~\ref{fig:trending_basic_4q}, points 2, 3, 5, and 6 show trending, whereas points 1, 4, and 7 count for anti-trending behavior.}
 \label{fig:trending_4q}
@@
 
-\subsection{Trending ratio and other measures}\label{subsec:trending-measures}
-
-Analyzing the number of points in the green versus red quadrants is a standard approach in the trending assessment of measurement data~\citep{Critchley2010,Saugel2015}.
-With that, we mainly estimate the probability of trending $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
+\subsection{ATC ratio and other measures}\label{subsec:atc-measures}
+
+Analyzing the number of points in the green versus red quadrants is a standard approach in the \ac{atc} assessment of measurement data~\citep{Critchley2010,Saugel2015}.
+With that, we mainly estimate the probability of a correctly predicted change direction, $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
 Since $z_1 z_2 > 0$ imposes the same condition as $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$), the standard estimator for $P(\diffxrv \diffyrv > 0)$ is
@@
 Thus, $\acc$ is the proportion of concordant changes on all changes.
-We refer to this estimator as the \textit{trending ratio} of the prediction and set $\mathcal{T} = \{l, \dots, T\}$.
+We refer to this estimator as the \textit{ATC ratio} of the prediction and set $\mathcal{T} = \{l, \dots, T\}$.
 Visually, the measure computes the fraction of points in the upper right or lower left quadrant.
@@
 
-Rolling estimates of the above measures detect changes in performance over time and can give a sharper estimate of the current trending ability.
-For the trending ratio, a rolling estimate with a backward-looking window of length $w$ at time $t$ is given by
+Rolling estimates of the above measures detect changes in performance over time and can give a sharper estimate of the current \ac{atc}.
+For the ATC ratio, a rolling estimate with a backward-looking window of length $w$ at time $t$ is given by
 \begin{equation*}
@@
 \end{equation*}
-Backward-looking windows estimate the trending ratio at a time $t$ considering the $w$ time steps before time $t$.
+Backward-looking windows estimate the ATC ratio at a time $t$ considering the $w$ time steps before time $t$.
 The window length $w$ controls the smoothing of the estimate; a larger $w$ gives smoother results, while a small $w$ focuses on local variations.
-Plotting the rolling estimates for $t = w-1, \dots, T$ yields an estimate of the trending ability over time.
-Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the trending ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
-Thus, the yearly course of the prediction trending ratio can be detected.
-The trending ability of the signal has a strong sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
-While colored four-quadrant plots, as in Figure~\ref{fig:trending_basic_4q_sample_color}, illustrate ongoing overall drifts in the trending ability, seasonal aspects are only revealed in rolling window estimates.
+Plotting the rolling estimates for $t = w-1, \dots, T$ yields an estimate of the ATC ratio over time.
+Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the ATC ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
+While colored four-quadrant plots, as in Figure~\ref{fig:trending_basic_4q_sample_color}, illustrate ongoing overall drifts in the \ac{atc}, seasonal aspects are only revealed in rolling window estimates.
 
@@
 The green curve $k_t$ shows the theoretical probability that $\diffxt$ has the same sign as $\diffyt$ for each time step.
-The rolling estimates are shifted to the right compared to $k_t$ as the windows look backward.}
+The rolling estimates are shifted to the right compared to $k_t$ as the windows look backward.
+Thus, the yearly course of the ATC ratio can be detected.
+The trending ability of the signal has a strong sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
+}
 \end{figure}
@@
 
-Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:trending-measures} accounting for such effects~\citep[see, e.g.,][]{Saugel2015,Critchley2010}.
+Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:atc-measures} accounting for such effects~\citep[see, e.g.,][]{Saugel2015,Critchley2010}.
 Points within that area are omitted in the four-quadrant plot and the calculation of the measures.
@@
 The trending ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates trending.
-Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:trending-measures}).
+Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:atc-measures}).
 The positive and negative trending ratios analyze the trending ratio given whether the predicted change is positive or negative, respectively.
@@
 The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative trending abilities and that they add valuable information to the trending ratio.
-In the applications, the bootstrap confidence intervals of Section~\ref{subsec:trending-measures} are used to quantify the estimation uncertainty of the trending measures.
+In the applications, the bootstrap confidence intervals of Section~\ref{subsec:atc-measures} are used to quantify the estimation uncertainty of the trending measures.
 The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
@@
 
-The estimators of Section~\ref{subsec:trending-measures} do not account for imbalances in the number of observed positive and negative changes \citep[for a theoretical analysis, see][Chapter 3]{Jolliffe2012}.
+The estimators of Section~\ref{subsec:atc-measures} do not account for imbalances in the number of observed positive and negative changes \citep[for a theoretical analysis, see][Chapter 3]{Jolliffe2012}.
 Significant differences in the number of observed positive and negative changes are unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.

### 2024-04-10 paper.tex
Remove trending and signal from text

@@
 The definition of the \textit{predicted} change depends on the context;
-Table~\ref{tab:notation} summarizes the notation for the different signal types and the computation of the predicted change.
+Table~\ref{tab:notation} summarizes the notation for \ac{mnf} and the computation of the predicted change.
 For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$.
@@
 
-In applications, signal data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
-We refer to time steps for which either signal or true values, or both, are unavailable as missing values.
-Systematical missing values could lead to a biased estimate of the trending ratio, and missing data should be inspected for any underlying patterns.
+In applications, data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
+We refer to time steps for which either \ac{mnf} or true values, or both, are unavailable as missing values.
+Systematical missing values could lead to a biased assessments, and missing data should be inspected for any underlying patterns.
 If the missing values are not systematic, random, and occur scarcely, data pairs with missing values can be excluded from the data to calculate the measures~\parencite[see][Section 1.3]{VanBuuren2018}.
@@
 Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
-Points in the green upper right and lower left quadrants reflect a correct trending for the respective time step, whereas points in the remaining red quadrants show incorrect predicted changes.
-Figure~\ref{fig:trending_basic_4q} displays a basic four-quadrant plot and
-Figure~\ref{fig:trending_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data.
+Points in the green upper right and lower left quadrants reflect a correct change direction for the respective time step, whereas points in the remaining red quadrants show incorrect predicted changes.
+Figure~\ref{fig:aatc_basic_4q} displays a basic four-quadrant plot and
+Figure~\ref{fig:aatc_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data.
 
 It can be extended by including information on the time index in the point color to reveal effects over time.
-In Figure~\ref{fig:trending_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, more recent values; $(\diffyt[l], \diffxt[l])$ is blue and turns green until $(\diffyt[T], \diffxt[T])$.
+In Figure~\ref{fig:aatc_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, more recent values; $(\diffyt[l], \diffxt[l])$ is blue and turns green until $(\diffyt[T], \diffxt[T])$.
 However, four-quadrant plots become crowded for larger datasets, and sequential information on the differences is complex to assess thoroughly.
@@
 \includegraphics{plots/illustrative_examples/4Q_without_excl}
-\caption{Basic four-quadrant plot.} \label{fig:trending_basic_4q}
+\caption{Basic four-quadrant plot.} \label{fig:aatc_basic_4q}
 \end{subfigure}\hspace{0.01\textwidth}%
@@
 \includegraphics{plots/illustrative_examples/4q_excl_box}
-\caption{Four-quadrant plot with rectangular exclusion area.}\label{fig:trending_basic_4q_excl_box}
+\caption{Four-quadrant plot with rectangular exclusion area.}\label{fig:aatc_basic_4q_excl_box}
 \end{subfigure}\hspace{0.01\textwidth}%
@@
 \includegraphics{plots/illustrative_examples/4q_excl_axis}
-\caption{Four-quadrant plot with horizontal exclusion area.} \label{fig:trending_basic_4q_excl_axis}
+\caption{Four-quadrant plot with horizontal exclusion area.} \label{fig:aatc_basic_4q_excl_axis}
 \end{subfigure}\hspace{0.01\textwidth}%
@@
 \includegraphics{plots/illustrative_examples/4q_excl_cross}
-\caption{Four-quadrant plot with cross-shaped exclusion area.}\label{fig:trending_basic_4q_excl_cross}
+\caption{Four-quadrant plot with cross-shaped exclusion area.}\label{fig:aatc_basic_4q_excl_cross}
 \end{subfigure}%
 \caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas.
-In Figure~\ref{fig:trending_basic_4q}, points 2, 3, 5, and 6 show trending, whereas points 1, 4, and 7 count for anti-trending behavior.}
-\label{fig:trending_4q}
+In Figure~\ref{fig:aatc_basic_4q}, points 2, 3, 5, and 6 correctly track change, whereas points 1, 4, and 7 do not.}
+\label{fig:aatc_4q}
 \end{figure}
@@
 3}.
-Many other measures can be adapted from those fields to deepen the trending analysis.
-Two simple measures that focus on either a positive or negative predicted change are the positive and negative trending ratios $\accp$ and $\accm$, respectively.
+Many other measures can be adapted from those fields to deepen the analysis.
+Two simple measures that focus on either a positive or negative predicted change are the positive and negative \ac{atc} ratios $\accp$ and $\accm$, respectively.
 They are defined as
@@
 \end{align}
-They estimate the probability of a correct prediction of the direction of change, given that the signal direction is positive or negative, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$.
+They estimate the probability of a correct prediction of the direction of change, given that the predicted direction is positive or negative, that is, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0)$.
 
@@
 Plotting the rolling estimates for $t = w-1, \dots, T$ yields an estimate of the ATC ratio over time.
-Figure~\ref{fig:trending_ratio_time_series} depicts a rolling window estimate of the ATC ratio for the simulated data of Figures~\ref{fig:trending_basic_4q_sample} and~\ref{fig:trending_basic_4q_sample_color}.
-While colored four-quadrant plots, as in Figure~\ref{fig:trending_basic_4q_sample_color}, illustrate ongoing overall drifts in the \ac{atc}, seasonal aspects are only revealed in rolling window estimates.
+Figure~\ref{fig:atc_ratio_time_series} depicts a rolling window estimate of the ATC ratio for the simulated data of Figures~\ref{fig:aatc_basic_4q_sample} and~\ref{fig:aatc_basic_4q_sample_color}.
+While colored four-quadrant plots, as in Figure~\ref{fig:aatc_basic_4q_sample_color}, illustrate ongoing overall drifts in the \ac{atc}, seasonal aspects are only revealed in rolling window estimates.
 
@@
 \includegraphics{plots/illustrative_examples/4Q_sample_without_time}
-\caption{Four-quadrant plot with simulated data.}\label{fig:trending_basic_4q_sample}
+\caption{Four-quadrant plot with simulated data.}\label{fig:aatc_basic_4q_sample}
 \end{subfigure}\hspace{0.01\textwidth}
@@
 \includegraphics{plots/illustrative_examples/4Q_sample_with_time}
-\caption{The same data is colored according to the time index $t$, the greener, the later.}\label{fig:trending_basic_4q_sample_color}
+\caption{The same data is colored according to the time index $t$, the greener, the later.}\label{fig:aatc_basic_4q_sample_color}
 \end{subfigure}\hspace{0.01\textwidth}
 \begin{subfigure}[t]{.48\textwidth}
-    \includegraphics{plots/illustrative_examples/trending_ratio_time_series.pdf}
-    \caption{Rolling estimate of the trending ratio over time with window length 91. }\label{fig:trending_ratio_time_series}
+    \includegraphics{plots/illustrative_examples/trending_ratio_time_series}
+    \caption{Rolling estimate of the \ac{atc} ratio over time with window length 91. }\label{fig:atc_ratio_time_series}
     \end{subfigure}%
-    \caption[Visualizations of the four-quadrant plot for data with a time-varying trending ratio.]{Visualizations for data with a time-varying trending ratio.
-We defer information on the data generation process to the appendix (see \ref{subsec:app-trending-data-generation}).
-The trending ratio for the entire data set is $\mu = 0.7577$.
-The strong seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}.
+    \caption[Visualizations of the four-quadrant plot for data with a time-varying ATC ratio.]{Visualizations for data with a time-varying ATC ratio.
+We defer information on the data generation process to the appendix (see \ref{subsec:app-ATC-data-generation}).
+The ATC ratio for the entire data set is $\mu = 0.7577$.
+The strong seasonality of the ATC ratio becomes visible in Figure~\ref{fig:atc_ratio_time_series}.
 The green curve $k_t$ shows the theoretical probability that $\diffxt$ has the same sign as $\diffyt$ for each time step.
@@
 Thus, the yearly course of the ATC ratio can be detected.
-The trending ability of the signal has a strong sinus-shaped seasonality with a trending ability peak after a quarter of a year and a low point after three quarters.
+The ATC ratio has a strong sinus-shaped seasonality with a peak after a quarter of a year and a low point after three quarters.
 }
@@
 Suppose noise or non-systematic effects are present in the true values or predictions.
-In that case, a point's assignment to a quadrant can be driven by noise instead of a systematic trending ability of the signal.
+In that case, a point's assignment to a quadrant can be driven by noise instead of a systematic \ac{atc}.
 This is more likely for points with at least one small coordinate.
@@
 The additional argument $(\diffyt, \diffxt) \notin E$ in the indicator ensures that only points not lying in the exclusion are evaluated when computing the measures.
-The measures are then estimators for the probability of trending, given that the point's location is not driven by noise or non-informative changes.
+The measures are then estimators for the probability of predicting the correct direction, given that the point's location is not driven by noise or non-informative changes.
 
 The estimators accept various shapes of the exclusion area.
-Figure~\ref{fig:trending_4q} visualizes examples of exclusion area shapes.
+Figure~\ref{fig:aatc_4q} visualizes examples of exclusion area shapes.
 A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and, therefore, likely to be driven by noise.
-Points are included if at least one of true value or signal is unlikely to be noise-driven.
-In the example graph in Figure~\ref{fig:trending_basic_4q_excl_box}, only point 1 is excluded.
+Points are included if at least one of true value or prediction is unlikely to be noise-driven.
+In the example graph in Figure~\ref{fig:aatc_basic_4q_excl_box}, only point 1 is excluded.
 An exclusion area along one axis, for example, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
-This particularly suits signals where small amounts of noise are inevitable.
+This particularly suits \ac{mnf} where small amounts of noise are inevitable.
 A cross-shaped exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
 A point is only included if no component is likely to be driven by noise.
-For the latter two methods, points 1 and 7 in Figure~\ref{fig:trending_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:trending_basic_4q_excl_cross}, respectively.
+For the latter two methods, points 1 and 7 in Figure~\ref{fig:aatc_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:aatc_basic_4q_excl_cross}, respectively.
 
@@
 The size determination can also be based on a proportion of the total variance or the total range of the data; for example, the 10\% smallest absolute values in each component determine the exclusion area size.
-A third approach is to visualize the trending ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
+A third approach is to visualize the \ac{atc} ratio for different sizes of $E$ and thus inspect the effects of the exclusion area size on the estimates.
 For examples of such plots, see Section~\ref{sec:application-covid}.
@@
 Based on the derived \enquote{new} samples of the statistic, the confidence intervals can be derived through differently sophisticated bootstrapping methods.
-We use the \ac{bca} approach for bootstrapping in the following, as it holds the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-trending-bootstrap}).
-
-\subsection{The conditional trending plot}\label{subsec:aatc-cond-prob}
+We use the \ac{bca} approach for bootstrapping in the following, as it holds the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-aatc-bootstrap}).
+
+\subsection{The conditional ATC plot}\label{subsec:aatc-cond-prob}
 
 The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
-A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the trending ability of a prediction for a certain change size $\diffxrv = \xcond$ of the signal.
-Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct trending given a predicted change of $\xcond$.
-Thus, if a change of $\xcond$ is observed in practice, one can directly assess the credibility of this specific signal.
+A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the \ac{atc} ability of a prediction for a certain change $\diffxrv = \xcond$ of the \ac{mnf}.
+Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct direction given a predicted change of $\xcond$.
+Thus, if a change of $\xcond$ is observed in practice, one can directly assess its credibility.
 A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
@@
 Assessing $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ graphically by drawing $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ against $\xcond$ eases the simultaneous evaluation of various $\xcond$.
-Furthermore, the graph facilitates the comparison of various methods in a single graph, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = \xcond)$ with respect to $x$ in the trending ability can be detected.
-We refer to the plot as a \textit{conditional trending plot}.
+Furthermore, the graph facilitates the comparison of various methods in a single graph, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = \xcond)$ with respect to $x$ in the \ac{atc} can be detected.
+We refer to the plot as a \textit{conditional ATC plot}.
 
@@
 In nowcasting and forecasting, probabilistic predictions have become more prevalent in recent years (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
-In this section, we develop trending measures for probabilistic signals.
+In this section, we develop \ac{atc} assessments for probabilistic measurements and nowcasts.
 Probabilistic predictions issue a probability distribution for the quantity of interest based on their available information and, thus, include a point estimate and information on the prediction uncertainty and quantiles simultaneously.
-For trending analysis, probabilistic predictions thus also contain a probability of a positive or negative change.
-In this section, we extend the trending assessment to probabilistic predictions.
+For \ac{atc} analysis, probabilistic predictions thus also contain a probability of a positive or negative change.
+In this section, we extend the \ac{atc} assessment to probabilistic predictions.
 
@@
 Let us first assume that the prediction is a \ac{cdf}, and that $y_{t-l}$ is known at time $t$ (see Table~\ref{tab:notation}).
-In Appendix~\ref{subsec:app-probabilistic-trending-evaluation}, we extend the analysis for predictions given as quantiles or unknown true values.
+In Appendix~\ref{subsec:app-probabilistic-atc-evaluation}, we extend the analysis for predictions given as quantiles or unknown true values.
 
@@
 Thereby, $F_{t | t-l} (x)$ specifies the forecastes probability that the quantity of interest is at most $x$.
-For trending assessment, we compare the predicted probability of positive change, denoted by $p_t$, with the occurrence of positive changes.
+For \ac{atc} assessment, we compare the predicted probability of positive change, denoted by $p_t$, with the occurrence of positive changes.
 A positive change occurs for any value at $t$ larger than the true value $y_{t-l}$, as then the change compared to $y_{t-l}$ is positive.
@@
 \end{equation*}
-The probabilistic trending evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=l}^{T}$ and the observed changes $\mathbf{z} = (z_t)_{t=l}^{T}$.
+The probabilistic \ac{atc} evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=l}^{T}$ and the observed changes $\mathbf{z} = (z_t)_{t=l}^{T}$.
 The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
@@
 \end{equation*}
-Thereby, lower values indicate a higher probabilistic trending ability of the considered method.
+Thereby, lower values indicate a higher probabilistic \ac{atc} ability of the considered method.
 The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\citep{Ranjan2010,Mitchell2011}.
@@
 
-In this section, we evaluate the trending in three different applications.
+In this section, we evaluate the \ac{atc} in three different applications.
 In the first application in Section~\ref{sec:application-covid}, we consider the nowcasting of the seven-day hospitalization rate for COVID-19 in Germany.
-Section~\ref{sec:application-eda} applies the trending assessment to forecast methods for the number of arrivals in a large emergency department.
-In the last application in Section~\ref{sec:application_measurement}, we consider the trending assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.
+Section~\ref{sec:application-eda} applies the \ac{atc} assessment to forecast methods for the number of arrivals in a large emergency department.
+In the last application in Section~\ref{sec:application_measurement}, we consider the \ac{atc} assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.
 
@@
 
-In addition to close inspection of the point evaluation measures, assessing the trending of the nowcasts is crucial.
+In addition to close inspection of the point evaluation measures, assessing the \ac{atc} of the nowcasts is crucial.
 To assess the impact of taken measures and the direction of the curve, it is essential to distinguish between rising and falling hospitalization rates.
@@
 
-In the following, we apply the trending assessment of Section~\ref{sec:aatc} to the nowcasts of the seven-day hospitalization rate.
-We report the trending for the horizons one, seven, and 14 days.
-While horizon one assesses the short-term trending, horizons seven and 14 consider the medium-term trending.
+In the following, we apply the \ac{atc} assessment of Section~\ref{sec:aatc} to the nowcasts of the seven-day hospitalization rate.
+We report the \ac{atc} for the horizons one, seven, and 14 days.
+While horizon one assesses the short-term \ac{atc}, horizons seven and 14 consider the medium-term \ac{atc}.
 The horizons seven and 14 are particularly interesting, reflecting a usual period until new policy changes are taken.
@@
 Similarly, the 10\%-quantile of differences increases.
-The 10\%-quantile is used for the exclusion areas in the trending assessment.
+The 10\%-quantile is used for the exclusion areas in the \ac{atc} assessment.
 The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the absolute differences.
-Thus, points are still included in the trending assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.
+Thus, points are still included in the \ac{atc} assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.
 
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption[The trending ratios for Covid nowcasting.]{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days in Covid nowcasting.
+    \caption[The ATC ratios for Covid nowcasting.]{The \ac{atc} ratio $\accl[7]$, positive \ac{atc} ratio $\accpl[7]$, and negative \ac{atc} ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days in Covid nowcasting.
 The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
@@
 }
-    \label{tab:app-covid-trending-ratios-lag-7}
+    \label{tab:app-covid-atc-ratios-lag-7}
 \end{table}
 
-Table~\ref{tab:app-covid-trending-ratios-lag-7} lists the trending ratios for all models with and without exclusion areas for the horizon of seven days.
-The trending ratios without exclusion area range from 0.72 to 0.85 for the horizon of seven days.
-The negative trending ratios are higher than the positive trending ratios for all models.
-The confidence intervals for the positive and negative trending ratios do not overlap for all models, indicating that the trending ratios are indeed different.
+Table~\ref{tab:app-covid-atc-ratios-lag-7} lists the \ac{atc} ratios for all models with and without exclusion areas for the horizon of seven days.
+The \ac{atc} ratios without exclusion area range from 0.72 to 0.85 for the horizon of seven days.
+The negative \ac{atc} ratios are higher than the positive \ac{atc} ratios for all models.
+The confidence intervals for the positive and negative \ac{atc} ratios do not overlap for all models, indicating that the \ac{atc} ratios are indeed different.
 The 10\%-quantile exclusion areas have, at most, an influence of 0.03 on the ratios.
-The model with the highest trending ratio is the ILM model, and the model with the lowest is the RKI model.
+The model with the highest \ac{atc} ratio is the ILM model, and the model with the lowest is the RKI model.
 The confidence intervals between all models overlap.
-The positive trending ratio implies a similar ranking of the models, while the negative ratio is second best for the RKI model.
-For the horizons of one and 14 days, we refer to Table~\ref{tab:app-covid-trending-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
+The positive \ac{atc} ratio implies a similar ranking of the models, while the negative ratio is second best for the RKI model.
+For the horizons of one and 14 days, we refer to Table~\ref{tab:app-covid-atc-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
 
@@
     \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
-    \caption{Conditional trending plot.}\label{fig:app-covid-cond-prob-7}
+    \caption{Conditional ATC plot.}\label{fig:app-covid-cond-prob-7}
     \end{subfigure}\hfill
@@
     \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
-    \caption{Trending ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-trending-ratio-7}
+    \caption{ATC ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-atc-ratio-7}
     \end{subfigure}
-    \caption[Conditional trending plot and trending ratio over exclusion area for the Covid nowcasts.]{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days in Covid nowcasting.}
-    \label{fig:app-covid-cond-prob-trending-ratio-7}
-\end{figure}
-
-Figure~\ref{fig:app-covid-cond-prob-trending-ratio-7} shows the conditional trending plots and the trending ratio over the exclusion area for the horizon seven days; the respective plots for the horizons one day and 14 days are shown in Figure~\ref{fig:app-covid-cond-prob-trending-ratio-1-14}.
+    \caption[Conditional ATC plot and ATC ratio over exclusion area for the Covid nowcasts.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days in Covid nowcasting.}
+    \label{fig:app-covid-cond-prob-atc-ratio-7}
+\end{figure}
+
+Figure~\ref{fig:app-covid-cond-prob-atc-ratio-7} shows the conditional \ac{atc} plots and the \ac{atc} ratio over the exclusion area for the horizon seven days; the respective plots for the horizons one day and 14 days are shown in Figure~\ref{fig:app-covid-cond-prob-atc-ratio-1-14}.
 Here, only the best models in point evaluation measures, ILM, RKI, RIVM, and ENS-MED, are shown to keep the plots easily readable.
@@
 Similar observations can be made for the horizon of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
-For a horizon of one day, the models' conditional trending ability difference is less pronounced (see Figure~\ref{fig:app-covid-cond-prob-1}).
+For a horizon of one day, the models' conditional ATC difference is less pronounced (see Figure~\ref{fig:app-covid-cond-prob-1}).
 The RKI model is still less conclusive when issuing an increase in the hospitalization rate, while RIVM is most informative in that case.
@@
 
-The trending ratios for various exclusion areas are shown in Figure~\ref{fig:app-covid-trending-ratio-7}.
-In general, the trending ratio increases with larger exclusion areas.
-While the RIVM and ENS-MED trending ratios evolve similarly, the RKI and ILM trending ratios get closer.
-For the horizon of one day, the RKI trending ratio decreases with increasing exclusion area size while the other models rise (see Figure~\ref{fig:app-covid-trending-ratio-1}).
-For the horizon of 14 days, all trending ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-trending-ratio-14}).
+The \ac{atc} ratios for various exclusion areas are shown in Figure~\ref{fig:app-covid-atc-ratio-7}.
+In general, the \ac{atc} ratio increases with larger exclusion areas.
+While the RIVM and ENS-MED \ac{atc} ratios evolve similarly, the RKI and ILM \ac{atc} ratios get closer.
+For the horizon of one day, the RKI \ac{atc} ratio decreases with increasing exclusion area size while the other models rise (see Figure~\ref{fig:app-covid-atc-ratio-1}).
+For the horizon of 14 days, all \ac{atc} ratio curves increase with the exclusion area size (see Figure~\ref{fig:app-covid-atc-ratio-14}).
 
@@
 The probabilities of increase for the different models are computed using the nowcast quantiles.
-For each horizon $l$, $10,000$ samples of the forecast date $t$ and the forecast date $t-l$ based on the nowcasts of issue date $t$ are generated, and the proportion of positive changes is computed (see Appendix~\ref{subsec:app-probabilistic-trending-evaluation}).
-Remember that a low \ac{bs} and a reliability diagram along the diagonal are signs of good trending.
+For each horizon $l$, $10,000$ samples of the forecast date $t$ and the forecast date $t-l$ based on the nowcasts of issue date $t$ are generated, and the proportion of positive changes is computed (see Appendix~\ref{subsec:app-probabilistic-atc-evaluation}).
+Remember that a low \ac{bs} and a reliability diagram along the diagonal are signs of a high \ac{atc}.
 The \ac{bs} is the lowest for the RIVM model for a one-day-horizon, while the ENS-MED model has the best \ac{bs} for the horizon of seven and 14 days.
@@
 For all horizons, the influence of the exclusion area on the 10\%-quantile level is negligible.
-For example, the trending ratio changes at most by 0.03 for the EPI model with $\accml[14]$.
-The exclusion areas are thus not crucial for the trending assessment in the case of the nowcasts of the seven-day hospitalization rate.
+For example, the \ac{atc} ratio changes at most by 0.03 for the EPI model with $\accml[14]$.
+The exclusion areas are thus not crucial for the \ac{atc} assessment in the case of the nowcasts of the seven-day hospitalization rate.
 The lower bound of confidence intervals is at least 0.68 for all models, indicating that they perform better than random guessing the trend.
 
-Trending assessment evaluates the models differently from point evaluation measures.
-RKI is among the best in point evaluation measures but performs worse in trending assessment.
-The assessment of asymmetry in the conditional trending plots is crucial for interpreting the trending ratios, with the RKI model being the most prominent example.
-
-Figure~\ref{fig:app-covid-trending-ratio-7} shows that the trending ratio increases with larger exclusion areas.
+\ac{atc} assessment evaluates the models differently from point evaluation measures.
+RKI is among the best in point evaluation measures but performs worse in \ac{atc} assessment.
+The assessment of asymmetry in the conditional \ac{atc} plots is crucial for interpreting the \ac{atc} ratios, with the RKI model being the most prominent example.
+
+Figure~\ref{fig:app-covid-atc-ratio-7} shows that the \ac{atc} ratio increases with larger exclusion areas.
 This indicates that if the model predicts a large change, the direction is more accurate than when a small change is predicted.
 
-The probabilistic trending assessment shows that the models are better than random guessing.
+The probabilistic \ac{atc} assessment shows that the models are better than random guessing.
 In general, the reliability diagram cannot provide information if specific probabilities are issued scarcely.
 Thus, the reliability diagrams for the horizons of seven and 14 days do not contain information on moderate probabilities.
-The \ac{bs} values, however, work well for those examples and provide a good measure for the trending ability of the models.
+The \ac{bs} values, however, work well for those examples and provide a good measure for the \ac{atc} of the models.
 
 A more extensive data size would be beneficial for assessing the models' performance.
-For the evaluation period of 159 days, the trending ratio confidence intervals overlap; thus, no conclusions can be drawn from the trending evaluation comparing the models.
+For the evaluation period of 159 days, the \ac{atc} ratio confidence intervals overlap; thus, no conclusions can be drawn from the \ac{atc} assessment comparing the models.
 
@@
 For further notes on the models and the evaluation, we refer to \citet{Rostami-Tabar2023}.
-From the issued forecasts, we use the mean as a point forecast for the trending assessment and evaluate the probabilistic trending subsequently.
+From the issued forecasts, we use the mean as a point forecast for the \ac{atc} assessment and evaluate the probabilistic \ac{atc} subsequently.
 We use the forecasts issued at the first time point for every target time.
@@
 
-In this setup, trending assessment is a simple and intuitive way to assess the models' performance.
-The trending perspective is easy for the management to understand and implement, as simple comparisons of the expected workload to a recent shift can be made.
+In this setup, \ac{atc} assessment is a simple and intuitive way to assess the models' performance.
+The \ac{atc} perspective is easy for the management to understand and implement, as simple comparisons of the expected workload to a recent shift can be made.
 If, for example, the staff was near capacity during the last shift and an increase in the number of arrivals is expected, management can take measures to adjust the workload.
@@
 
-Table~\ref{tab:app-eda-trending-ratios} lists the trending ratios for all models for three and seven-day horizons.
-The trending ratios range from 0.68 to 0.84 for a horizon of three days and from 0.68 to 0.82 for seven days.
-The negative and positive trending ratios differ for all models and horizons.
-For some models, for example, the GBM-2 model, the positive trending ratio is higher; for some models, for example, the tbats model, the negative trending ratio is higher.
-The confidence interval width is at most 0.02 for the trending ratios and at most 0.03 for the positive and negative trending ratios.
-The models GBM-2, qreg-1, and Benchmark-1 have the highest positive trending ratio for the three and seven-day horizon, while Poisson-2 and NBI-2 have the highest negative trending ratio.
-
-Figure~\ref{fig:app-eda-cond-prob} shows the conditional trending plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days and thus inspects the local trending ability of the models with highest positive and negative trending ratio.
-The conditional trending plots show similar courses for the two horizons, though the curves are shifted downwards for the horizon of seven days.
-The model's relative trending ability evolves consistently for the two horizons, with the NBI-2 and Poisson-2 models being indistinguishable.
+Table~\ref{tab:app-eda-atc-ratios} lists the \ac{atc} ratios for all models for three and seven-day horizons.
+The \ac{atc} ratios range from 0.68 to 0.84 for a horizon of three days and from 0.68 to 0.82 for seven days.
+The negative and positive \ac{atc} ratios differ for all models and horizons.
+For some models, for example, the GBM-2 model, the positive \ac{atc} ratio is higher; for some models, for example, the tbats model, the negative \ac{atc} ratio is higher.
+The confidence interval width is at most 0.02 for the \ac{atc} ratios and at most 0.03 for the positive and negative \ac{atc} ratios.
+The models GBM-2, qreg-1, and Benchmark-1 have the highest positive \ac{atc} ratio for the three and seven-day horizon, while Poisson-2 and NBI-2 have the highest negative \ac{atc} ratio.
+
+Figure~\ref{fig:app-eda-cond-prob} shows the conditional \ac{atc} plots for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days and thus inspects the local \ac{atc} of the models with highest positive and negative \ac{atc} ratio.
+The conditional \ac{atc} plots show similar courses for the two horizons, though the curves are shifted downwards for the horizon of seven days.
+The model's relative \ac{atc} evolves consistently for the two horizons, with the NBI-2 and Poisson-2 models being indistinguishable.
 The GBM-2 model outperforms the qreg-1 model for all predicted changes.
-The models NBI-2 and Poisson-2 have the highest trending ability for all negative predicted changes and the lowest for all positive predicted changes.
+The models NBI-2 and Poisson-2 have the highest \ac{atc} for all negative predicted changes and the lowest for all positive predicted changes.
 Benchmark-1 lies between the other models for all predicted changes.
 
-Figure~\ref{fig:app-eda-prob} visualizes the probabilistic trending evaluation for the same subset of models.
+Figure~\ref{fig:app-eda-prob} visualizes the probabilistic \ac{atc} assessment for the same subset of models.
 The \acfp{bs} are shown in Figure~\ref{fig:app-eda-prob-brier}, and the reliability diagrams for the horizons three and seven days in Figures~\ref{fig:app-eda-prob-rel-3} and~\ref{fig:app-eda-prob-rel-7}.
@@
     \input{plots/ed_arrival/50_trending_ratio}
-    \caption[Trending ratios for the emergency department arrival forecasts.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days in emergency department arrival forecasting.}
-    \label{tab:app-eda-trending-ratios}
+    \caption[ATC ratios for the emergency department arrival forecasts.]{\Ac{atc} ratio $\acc$, positive \ac{atc} ratio $\accp$, and negative \ac{atc} ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days in emergency department arrival forecasting.}
+    \label{tab:app-eda-atc-ratios}
 \end{table}
@@
     \end{subfigure}
-    \caption[Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability in emergency department arrival forecasting.]{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability in emergency department arrival forecasting.
+    \caption[Conditional ATC plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} ability in emergency department arrival forecasting.]{Conditional \ac{atc} plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} ability in emergency department arrival forecasting.
 The plots of NBI-2 and Poisson-2 are indistinguishable.}
@@
     \end{subfigure}
-    \caption[Probabilistic trending evaluation in emergency department arrival forecasting.]{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days in emergency department arrival forecasting.
-The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic trending simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically.
-Probabilistic trending by the \ac{bs} is best for the models NBI-2 and Poisson-2 for both lags.}
+    \caption[Probabilistic \ac{atc} assessment in emergency department arrival forecasting.]{Probabilistic \ac{atc} assessment for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days in emergency department arrival forecasting.
+The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic \ac{atc} simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically.
+Probabilistic \ac{atc} by the \ac{bs} is best for the models NBI-2 and Poisson-2 for both lags.}
     \label{fig:app-eda-prob}
@@
 
-The trending ability is consistent for the two horizons, with the models' relative trending ability evolving similarly for the two horizons.
-The models' trending ability is generally higher for the smaller horizon, but the differences are minor, and confidence intervals overlap.
-
-The trending differs for all models for positive and negative predicted change directions.
-While some models, such as GBM-2 and qreg-1, have the highest positive trending ratio, others, such as Poisson-2 and NBI-2, have the highest negative trending ratio.
+The \ac{atc} is consistent for the two horizons, with the models' relative \ac{atc} evolving similarly for the two horizons.
+The models' \ac{atc} is generally higher for the smaller horizon, but the differences are minor, and confidence intervals overlap.
+
+The \ac{atc} differs for all models for positive and negative predicted change directions.
+While some models, such as GBM-2 and qreg-1, have the highest positive \ac{atc} ratio, others, such as Poisson-2 and NBI-2, have the highest negative \ac{atc} ratio.
 Thus, the uncertainty of the model's predicted change has to be assessed differently based on the direction.
 
-The results of the probabilistic trending evaluation endorse the point trending assessment and assign the best scores to NBI-2 and Poisson-2.
+The results of the probabilistic \ac{atc} assessment endorse the point \ac{atc} assessment and assign the best scores to NBI-2 and Poisson-2.
 The reliability diagrams show that they underestimate the fraction of increases slightly.
@@
 Overall, the example provides performance assessments that are different from standard point evaluation measures and thus provide further insights into the strengths and weaknesses of the models.
-While the models with the lowest RMSE, NBI-2, and Poisson-2, also have a high trending ability, three models with below-average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive trending ability.
+While the models with the lowest RMSE, NBI-2, and Poisson-2, also have a high \ac{atc}, three models with below-average point evaluation measures, Benchmark-1, GBM-2, and qreg-1, have a high positive \ac{atc}.
 
@@
 
-In the last briefer example, we consider the trending assessment of measurement data.
+In the last briefer example, we consider the \ac{atc} assessment of measurement data.
 The data is from the MIMIC-III database, including various information on patients in critical care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}).
@@
 For critical patients, changes in blood pressure can be crucial for the treatment.
-Thus, trending assessment can be performed in addition to standard accuracy analysis~\citep[see, for example, ][]{Mostafa2020}.
-Thus, we assess the trending ability of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
+Thus, \ac{atc} assessment can be performed in addition to standard accuracy analysis~\citep[see, for example, ][]{Mostafa2020}.
+Thus, we assess the \ac{atc} of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
 The database contains 64,168 numerical data records.
@@
 
-We consider the horizons of one minute, five minutes, and 15 minutes for the trending assessment, as those are typical intervals of NBP measurements.
+We consider the horizons of one minute, five minutes, and 15 minutes for the \ac{atc} assessment, as those are typical intervals of NBP measurements.
 
@@
 
-Again, we exclude the smallest 10\% of absolute differences in trending assessment.
+Again, we exclude the smallest 10\% of absolute differences in \ac{atc} assessment.
 The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different horizons are shown in Figure~\ref{fig:app-mimic-4q}.
@@
 
-The trending ratios, including confidence intervals for the different horizons, are listed in Table~\ref{tab:app-mimic-trending-ratios}.
+The \ac{atc} ratios, including confidence intervals for the different horizons, are listed in Table~\ref{tab:app-mimic-atc-ratios}.
 The confidence intervals have lower bounds of 0.5 or slightly above for the measurements with a horizon of one minute.
-For larger horizons, the trending ratio increases.
-The difference between positive and negative trending ratios is small for all types and horizons, with overlapping confidence intervals.
-
-Figure~\ref{fig:app-mimic-cond-prob} shows the conditional trending plots for the different horizons and the systolic and mean blood pressure measurements.
-It becomes apparent that the systolic measurements have a higher trending ability than the mean measurement, except for small negative predicted changes.
-This aligns with the trending ratios, but the confidence intervals overlap.
+For larger horizons, the \ac{atc} ratio increases.
+The difference between positive and negative \ac{atc} ratios is small for all types and horizons, with overlapping confidence intervals.
+
+Figure~\ref{fig:app-mimic-cond-prob} shows the conditional \ac{atc} plots for the different horizons and the systolic and mean blood pressure measurements.
+It becomes apparent that the systolic measurements have a higher \ac{atc} ability than the mean measurement, except for small negative predicted changes.
+This aligns with the \ac{atc} ratios, but the confidence intervals overlap.
 
@@
     \input{plots/mimic/trending_ratio}
-    \caption[Trending ratios for the different horizons $l$ and the systolic and mean blood pressure measurements.]{Trending ratios for the different horizons $l$ and the systolic and mean blood pressure measurements.}
-    \label{tab:app-mimic-trending-ratios}
+    \caption[ATC ratios for the different horizons $l$ and the systolic and mean blood pressure measurements.]{\Ac{atc} ratios for the different horizons $l$ and the systolic and mean blood pressure measurements.}
+    \label{tab:app-mimic-atc-ratios}
 \end{table}
@@
     \end{subfigure}\hspace{0.01\textwidth}
-    \caption{Conditional trending plot for the systolic and mean blood pressure measurements and the horizons one, five, and 15 minutes. }
+    \caption{Conditional ATC plot for the systolic and mean blood pressure measurements and the horizons one, five, and 15 minutes. }
     \label{fig:app-mimic-cond-prob}
@@
 The differences between positive and negative predicted changes are small in this example.
-The positive and negative trending ratios have overlapping confidence intervals, and the conditional trending plots do not contain prominent deviations in the course.
+The positive and negative \ac{atc} ratios have overlapping confidence intervals, and the conditional \ac{atc} plots do not contain prominent deviations in the course.
 This aligns with the four-quadrant plots, where no apparent asymmetry is visible.
@@
 The bootstrap confidence intervals are wide.
-The width is around 0.1 for the trending ratio, while it gets up to 0.16 for the negative trending ratio for systolic measurement and the horizon of one minute.
+The width is around 0.1 for the \ac{atc} ratio, while it gets up to 0.16 for the negative \ac{atc} ratio for systolic measurement and the horizon of one minute.
 Thus, the confidence intervals cover 0.5 for systolic measurement and the horizon of one minute, and the equality to random guessing cannot be excluded.
 
-\section{Discussion and conclusion}\label{sec:trending-conclusion}
-
-In this paper, we examine various methods to analyze trending, that is, the statistical consistency of observed and predicted change direction.
-While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting (see Section~\ref{subsec:notation}), the trending assessment using the computed predicted and observed change is similar.
-The evaluation of trending can accompany other evaluation techniques, such as measures of deviation or probabilistic scoring rules.
-
-Four-quadrant plots facilitate the visual inspection of the trending ability for a signal (see Section~\ref{subsec:aatc-four-quadrant-plot}).
-The trending ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates trending.
+\section{Discussion and conclusion}\label{sec:atc-conclusion}
+
+In this paper, we examine various methods to assess the \acf{atc} for \acf{mnf}, that is, whether predict the correct direction of changes.
+While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting, the predict the correct direction of changes assessment using the computed predicted and observed change is similar.
+The \ac{atc} assessment can accompany other evaluation techniques, such as measures of deviation or probabilistic scoring rules.
+
+Four-quadrant plots facilitate the visual inspection of the \ac{atc} for a \ac{mnf} (see Section~\ref{subsec:aatc-four-quadrant-plot}).
+The \ac{atc} ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates \ac{atc}.
 Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:atc-measures}).
-The positive and negative trending ratios analyze the trending ratio given whether the predicted change is positive or negative, respectively.
+The positive and negative \ac{atc} ratios analyze the \ac{atc} ratio given whether the predicted change is positive or negative, respectively.
 Thus, they quantify the credibility of the respective predictions.
-The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative trending abilities and that they add valuable information to the trending ratio.
-In the applications, the bootstrap confidence intervals of Section~\ref{subsec:atc-measures} are used to quantify the estimation uncertainty of the trending measures.
+The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative \ac{atc} and that they add valuable information to the \ac{atc} ratio.
+In the applications, the bootstrap confidence intervals of Section~\ref{subsec:atc-measures} are used to quantify the estimation uncertainty of the \ac{atc} measures.
 The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
-For reasonably well-trending models, 100 samples are thus sufficient to differentiate from random guessing or to assess models with high trending differences.
-
-A conditional trending plot visualizes the probability of correct trending over the predicted change of the signal (see Section~\ref{subsec:aatc-cond-prob}).
+For models with reasonably high \ac{atc}, 100 samples are thus sufficient to differentiate from random guessing or to assess models with high \ac{atc} differences.
+
+A conditional \ac{atc} plot visualizes the probability of correct change direction prediction over the predicted change of the \ac{mnf} (see Section~\ref{subsec:aatc-cond-prob}).
 It is based on a multivariate \acf{kde} of predicted and observed change.
-In the application, the conditional trending plot gives reasonable insights into the local effects of the trending ability.
-Section~\ref{subsec:aatc-probabilistic} adapts measures of probabilistic forecast evaluation to the trending assessment of probabilistic forecasts and nowcasts.
-The \acf{bs} as numerical assessment of probabilistic trending is introduced, and reliability diagrams are used to visualize the local trending ability of probabilistic forecasts.
-
-The methods of trending assessment are applied to COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements in Section~\ref{sec:application}.
-While trending evaluation should not be the only aspect, it is a valuable addition to evaluating nowcasts, forecasts, and measurements.
-Models with very different accuracies are usually scored similarly in trending evaluation, but trending evaluation can differentiate between models with similar accuracies.
-As in the application in Section~\ref{sec:application-covid}, models with medial point forecast evaluation measures can have the most meaningful positive trending ability.
+In the application, the conditional \ac{atc} plot gives reasonable insights into the local effects of the \ac{atc}.
+Section~\ref{subsec:aatc-probabilistic} adapts measures of probabilistic forecast evaluation to the \ac{atc} assessment of probabilistic forecasts and nowcasts.
+The \acf{bs} as numerical assessment of probabilistic \ac{atc} is introduced, and reliability diagrams are used to visualize the local \ac{atc} of probabilistic forecasts.
+
+The methods of \ac{atc} assessment are applied to COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements in Section~\ref{sec:application}.
+While \ac{atc} assessment should not be the only aspect, it is a valuable addition to evaluating nowcasts, forecasts, and measurements.
+Models with highly different accuracies are usually scored similarly in \ac{atc} assessment, but \ac{atc} assessment can differentiate between models with similar accuracies.
+As in the application in Section~\ref{sec:application-covid}, models with medial point forecast evaluation measures can have the most meaningful positive \ac{atc}.
 
@@
 The estimators of Section~\ref{subsec:atc-measures} do not account for imbalances in the number of observed positive and negative changes \citep[for a theoretical analysis, see][Chapter 3]{Jolliffe2012}.
-Significant differences in the number of observed positive and negative changes are unlikely in the trending setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.
+Significant differences in the number of observed positive and negative changes are unlikely in the \ac{atc} setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.
 However, if the number of positive and negative observed changes differs widely, unbalanced-data-aware measures should be considered.
@@
 
-\section{Additional material on Section~\ref{sec:aatc}}\label{sec:appendix-trending}
-
-\subsection{Data generation for Section~\ref{sec:aatc}}\label{subsec:app-trending-data-generation}
+\section{Additional material on Section~\ref{sec:aatc}}\label{sec:appendix-atc}
+
+\subsection{Data generation for Section~\ref{sec:aatc}}\label{subsec:app-ATC-data-generation}
 
@@
 \end{equation*}
-Subsequently, the $\diffy$ are simulated for a constant trending ratio $k$ by
+Subsequently, the $\diffy$ are simulated for a constant \ac{atc} ratio $k$ by
 \begin{equation*}
@@
 where $n_t$ is a truncated normal distribution with mean one and standard deviation 0.5, truncated at 0, and $b_t$ is a symmetric Bernoulli random variable with parameter $k$.
-For a time-varying trending ratio, the parameter $k$ is modified to have a wave-shape over time, that is,
+For a time-varying \ac{atc} ratio, the parameter $k$ is modified to have a wave-shape over time, that is,
 \begin{equation*}
@@
 \end{equation*}
-For the asymmetric trending ratio, $k$ is a function of $\diffxt$,
+For the asymmetric \ac{atc} ratio, $k$ is a function of $\diffxt$,
 \begin{equation*}
@@
 \end{equation*}
-Thus, the conditional probability of trending can be calculated by a conditional normal distribution to
+Thus, the conditional probability of correct direction prediction can be calculated by a conditional normal distribution to
 \begin{equation*}
@@
         \includegraphics{plots/illustrative_examples/appendix_4q_dgp1}
-        \caption{Constant trending ratio.}
+        \caption{Constant \ac{atc} ratio.}
     \end{subfigure}\hspace{0.01\textwidth}
@@
         \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_time}
-        \caption{Time-varying trending ratio}
+        \caption{Time-varying \ac{atc} ratio}
     \end{subfigure}\hspace{0.01\textwidth}
@@
         \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_asym}
-        \caption{Asymmetric trending ratio}
+        \caption{Asymmetric \ac{atc} ratio}
     \end{subfigure}\hspace{0.01\textwidth}
@@
     \end{subfigure}
-    \caption[Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.]{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.
+    \caption[Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-ATC-data-generation}.]{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-ATC-data-generation}.
 Although the first and second plots differ over time, their difference is not discernible in the plots.
-The third data set's asymmetry is visible in the plot, but the decrease in the trending ability near 0 is not visible.
+The third data set's asymmetry is visible in the plot, but the decrease in the \ac{atc} near 0 is not visible.
 }
@@
 
-\subsection{Simulation study on bootstrapping confidence intervals}\label{subsec:appendix-trending-bootstrap}
+\subsection{Simulation study on bootstrapping confidence intervals}\label{subsec:appendix-aatc-bootstrap}
 
@@
 We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
-The considered datasets are the first dataset with asymmetric dependence and the second dataset outlined in Appendix~\ref{subsec:app-trending-data-generation}.
+The considered datasets are the first dataset with asymmetric dependence and the second dataset outlined in Appendix~\ref{subsec:app-ATC-data-generation}.
 In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\citep{Virtanen2020}.
 The prescribed confidence level is 90 \%, and the number of bootstrap samples is $10,000$.
-The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
+The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:atc_bootstrap}.
 The true values of the accuracy are computed based on a dataset of size $10^8$, yielding 0.7501 and 0.7700 for the two datasets.
-The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.
+The computation times per method and dataset are shown in Figure~\ref{fig:atc_bootstrap_time}.
 For the small sample sizes up to $T = 168$, only the \ac{bca} method keeps the confidence interval size and yields slightly wider confidence intervals.
@@
     \end{subtable}
-    \caption[Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$ in the bootstrap simulation study.]{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$.
+    \caption[Proportion of bootstrapping confidence intervals covering the true value of \ac{atc} ratio per method and sample size $T$ in the bootstrap simulation study.]{Proportion of bootstrapping confidence intervals covering the true value of \ac{atc} ratio per method and sample size $T$.
 The average width of the confidence interval is listed in brackets.}
-    \label{tab:trending_bootstrap}
+    \label{tab:atc_bootstrap}
 \end{table}
@@
 The \ac{bca} method takes slightly longer than the other two, but the difference is negligible.}
-    \label{fig:trending_bootstrap_time}
+    \label{fig:atc_bootstrap_time}
 \end{figure}
@@
 
-We examine the resulting conditional trending plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\citep{Seabold2010} in Figure~\ref{fig:trending-cond-prob-bw}.
+We examine the resulting conditional \ac{atc} plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\citep{Seabold2010} in Figure~\ref{fig:atc-cond-prob-bw}.
 While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.
@@
     \end{subfigure}
-    \caption[Conditional trending plots for different bandwidth selection processes.]{Conditional trending plot for different bandwidth selection processes.
+    \caption[Conditional ATC plots for different bandwidth selection processes.]{Conditional \ac{atc} plot for different bandwidth selection processes.
 Cross-validation least squares takes a considerably larger computation time.
@@
 The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time.
-$\varepsilon$ specifies an exclusion area $E = \{(x, y) \in \R^2: (-\varepsilon \leq x \leq \varepsilon)\}$ in $\diffx$-direction.}\label{fig:trending-cond-prob-bw}
-\end{figure}
-
-\subsection{Probabilistic trending evaluation}\label{subsec:app-probabilistic-trending-evaluation}
-
-Section~\ref{subsec:aatc-probabilistic} outlines the probabilistic evaluation of trending for nowcasts and forecasts and specifies the computation for predictions in terms of a \ac{cdf} and known true values.
+$\varepsilon$ specifies an exclusion area $E = \{(x, y) \in \R^2: (-\varepsilon \leq x \leq \varepsilon)\}$ in $\diffx$-direction.}\label{fig:atc-cond-prob-bw}
+\end{figure}
+
+\subsection{Probabilistic ATC evaluation}\label{subsec:app-probabilistic-atc-evaluation}
+
+Section~\ref{subsec:aatc-probabilistic} outlines the assessment of probabilistic \ac{atc} for nowcasts and forecasts and specifies the computation for predictions in terms of a \ac{cdf} and known true values.
 Here, we outline the computation for quantile forecasts and probabilistic true values.
@@
         &= \int_{x_1 \in \R} \int_{-\infty}^{x_1} f_{t|t+l} (x_1) f_{t+l|t+l} (x_2)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \nonumber\\
-        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:trending-probabilistic-pdf}.
+        &= \int_{x_1 \in \R} f_{t|t+l} (x_1) F_{t+l|t+l} (x_1)  \ \textrm{d} \: x_2 \ \textrm{d} \: x_1 \label{eq:atc-probabilistic-pdf}.
 \end{align}
@@
 If the nowcasts have the form of a multivariate distribution, including the dependence of the two \acp{pdf}, $f_{t+l|t+l} (x_2)$ has to be replaced by the \ac{pdf} conditional on $x_1$.
-As a Monte Carlo approximation of Equation~\eqref{eq:trending-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
+As a Monte Carlo approximation of Equation~\eqref{eq:atc-probabilistic-pdf}, the probability can also be calculated by sampling from $f_{t+l|t+l}$ and $f_{t|t+l}$ and calculating the fraction of negative changes.
 For forecasts, the indexes have to be shifted.
@@
     \end{subtable}
-    \caption[Trending ratios for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.
+    \caption[ATC ratios for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.]{\Ac{atc} ratio $\acc$, positive \ac{atc} ratio $\accp$, and negative \ac{atc} ratio $\accm$ for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.
 The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values.
 }
-    \label{tab:app-covid-trending-ratios-lag-1-14}
+    \label{tab:app-covid-atc-ratios-lag-1-14}
 \end{table}
@@
     \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
-    \caption{Conditional trending plot for horizon 1.}\label{fig:app-covid-cond-prob-1}
+    \caption{Conditional \ac{atc} plot for horizon 1.}\label{fig:app-covid-cond-prob-1}
     \end{subfigure}\hfill
@@
     \includegraphics{plots/covid_nowcast/40_cond_prob_lag_14}
-    \caption{Conditional trending plot for horizon 1.}\label{fig:app-covid-cond-prob-14}
+    \caption{Conditional \ac{atc} plot for horizon 1.}\label{fig:app-covid-cond-prob-14}
     \end{subfigure}
@@
     \includegraphics{plots/covid_nowcast/40_acc_eps_lag_1}
-    \caption{Trending ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-trending-ratio-1}
+    \caption{\Ac{atc} ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-atc-ratio-1}
     \end{subfigure}\hfill
@@
     \includegraphics{plots/covid_nowcast/40_acc_eps_lag_14}
-    \caption{Trending ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-trending-ratio-14}
+    \caption{\ac{atc} ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-atc-ratio-14}
     \end{subfigure}
-    \caption[Conditional trending plot and trending ratio over exclusion area in Covid nowcasting.]{Conditional trending plot and trending ratio over exclusion area for the Covid nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
-    \label{fig:app-covid-cond-prob-trending-ratio-1-14}
+    \caption[Conditional ATC plot and ATC ratio over exclusion area in Covid nowcasting.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the Covid nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
+    \label{fig:app-covid-cond-prob-atc-ratio-1-14}
 \end{figure}

### 2024-04-11 paper.tex
Update on Overleaf.
Shorten Paper

@@
 \renewcommand{\arraystretch}{1.2}
+
+% (Approximate) Word Count
+% Abstract
 
@@
 
-% \subsection{Computing changes and notation}\label{subsec:notation}
+\subsection{Computing changes and notation}\label{subsec:notation}
 
@@
 The \textit{observed} change is straightforward to compute for all types of \ac{mnf}.
-Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the true values for nowcasting or forecasting, or gold standard measurements up to time $T$.
+Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the actual values for nowcasting or forecasting, or gold standard measurements up to time $T$.
 The sequence of observed changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is,
@@
 \end{equation}
-with the same structure as in the nowcasting case.
-The indices are chosen to be consistent with $\diffyl$.
-If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification as in Equation~\eqref{eq:diffxl_nowcasting} can be made.
+with the same structure as in the nowcasting case and consistent indices with $\diffyl$.
+If the true value $y_{t-l}$ is not known at time $t-l$, a similar modification can be made as in Equation~\eqref{eq:diffxl_nowcasting}.
 
@@
 In applications, data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
-We refer to time steps for which either \ac{mnf} or true values, or both, are unavailable as missing values.
-Systematical missing values could lead to a biased assessments, and missing data should be inspected for any underlying patterns.
+We refer to time steps for which either \ac{mnf} or true values are unavailable as missing values.
+Systematical missing values could lead to a biased assessment, and missing data should be inspected for any underlying patterns.
 If the missing values are not systematic, random, and occur scarcely, data pairs with missing values can be excluded from the data to calculate the measures~\parencite[see][Section 1.3]{VanBuuren2018}.
@@
 In the following, we omit the horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
-The \ac{atc} is maximal if all the change directions are predicted correctly; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
+The \ac{atc} is maximal if all predicted change directions are correct; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
 Consequently, when assessing the \ac{atc}, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
-A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{Perrino1998,Saugel2015}.
-In a four-quadrant plot, the occured changes are plotted with the predicted changes, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
+A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{Perrino1998, Saugel2015}.
+In a four-quadrant plot, the occured changes and the predicted changes are plotted together, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
 Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
-Points in the green upper right and lower left quadrants reflect a correct change direction for the respective time step, whereas points in the remaining red quadrants show incorrect predicted changes.
-Figure~\ref{fig:aatc_basic_4q} displays a basic four-quadrant plot and
-Figure~\ref{fig:aatc_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data.
-
-It can be extended by including information on the time index in the point color to reveal effects over time.
+Points in the green upper right and lower left quadrants reflect a correct change direction for the respective time step, whereas points in the remaining red quadrants show incorrectly predicted changes.
+Figure~\ref{fig:aatc_basic_4q} displays a basic four-quadrant plot and Figure~\ref{fig:aatc_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data.
+
+The four-quadrant plot can be extended by including information on the time index in the point color to reveal effects over time.
 In Figure~\ref{fig:aatc_basic_4q_sample_color}, the point colors turn from blue to green for higher time indices $t$, that is, more recent values; $(\diffyt[l], \diffxt[l])$ is blue and turns green until $(\diffyt[T], \diffxt[T])$.
@@
 \includegraphics{plots/illustrative_examples/4q_excl_box}
-\caption{Four-quadrant plot with rectangular exclusion area.}\label{fig:aatc_basic_4q_excl_box}
+\caption{Four-quadrant plot with rectangular exclusion area. Point 1 is excluded.}\label{fig:aatc_basic_4q_excl_box}
 \end{subfigure}\hspace{0.01\textwidth}%
@@
 \includegraphics{plots/illustrative_examples/4q_excl_axis}
-\caption{Four-quadrant plot with horizontal exclusion area.} \label{fig:aatc_basic_4q_excl_axis}
+\caption{Four-quadrant plot with horizontal exclusion area. Points 1 and 7 are excluded.} \label{fig:aatc_basic_4q_excl_axis}
 \end{subfigure}\hspace{0.01\textwidth}%
@@
 \includegraphics{plots/illustrative_examples/4q_excl_cross}
-\caption{Four-quadrant plot with cross-shaped exclusion area.}\label{fig:aatc_basic_4q_excl_cross}
+\caption{Four-quadrant plot with cross-shaped exclusion area. Points 1, 6 and 7 are excluded.}\label{fig:aatc_basic_4q_excl_cross}
 \end{subfigure}%
 \caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas.
-In Figure~\ref{fig:aatc_basic_4q}, points 2, 3, 5, and 6 correctly track change, whereas points 1, 4, and 7 do not.}
+The rectangular exclusion area in Figure~\ref{fig:aatc_basic_4q_excl_box} excludes only points where both components are likely to be noise-driven, while the exclusion areas in Figures~\ref{fig:aatc_basic_4q_excl_axis} and~\ref{fig:aatc_basic_4q_excl_cross} excludes points where at least one component is noise-driven.
+}
 \label{fig:aatc_4q}
@@
 
-Analyzing the number of points in the green versus red quadrants is a standard approach in the \ac{atc} assessment of measurement data~\citep{Critchley2010,Saugel2015}.
-With that, we mainly estimate the probability of a correctly predicted change direction, $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
+Analyzing the number of points in the green versus red quadrants is a standard approach in the \ac{atc} assessment of measurement data~\citep{Critchley2010, Saugel2015}.
+With that, we estimate the probability of a correctly predicted change direction, $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
 Since $z_1 z_2 > 0$ imposes the same condition as $\sign(z_1) = \sign(z_2)$ ($z_1, z_2 \in \R \setminus \{ 0 \}$), the standard estimator for $P(\diffxrv \diffyrv > 0)$ is
@@
 Many other measures can be adapted from those fields to deepen the analysis.
-Two simple measures that focus on either a positive or negative predicted change are the positive and negative \ac{atc} ratios $\accp$ and $\accm$, respectively.
+Two simple measures that focus on a positive or negative predicted change are the positive and negative \ac{atc} ratios $\accp$ and $\accm$, respectively.
 They are defined as
@@
 Thus, the yearly course of the ATC ratio can be detected.
-The ATC ratio has a strong sinus-shaped seasonality with a peak after a quarter of a year and a low point after three quarters.
+The ATC ratio has a pronounced sinus-shaped seasonality with a peak after a quarter of a year and a low point after three quarters.
 }
@@
 Suppose noise or non-systematic effects are present in the true values or predictions.
-In that case, a point's assignment to a quadrant can be driven by noise instead of a systematic \ac{atc}.
+In that case, noise can drive a point's assignment to a quadrant instead of a systematic \ac{atc}.
 This is more likely for points with at least one small coordinate.
 
-Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:atc-measures} accounting for such effects~\citep[see, e.g.,][]{Saugel2015,Critchley2010}.
-Points within that area are omitted in the four-quadrant plot and the calculation of the measures.
-In particular, the prediction models are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
-We denote the measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} accounting for an exclusion area $E$ by
+Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:atc-measures} accounting for such effects~\citep[see, e.g.,][]{Saugel2015, Critchley2010}.
+Points within that area are omitted in the calculation of the measures.
+In particular, the \ac{mnf} are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
+The measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} without points in the exclusion area $E$ are
 \begin{align}
@@
 \end{align}
-The additional argument $(\diffyt, \diffxt) \notin E$ in the indicator ensures that only points not lying in the exclusion are evaluated when computing the measures.
 The measures are then estimators for the probability of predicting the correct direction, given that the point's location is not driven by noise or non-informative changes.
 
-The estimators accept various shapes of the exclusion area.
-Figure~\ref{fig:aatc_4q} visualizes examples of exclusion area shapes.
-A rectangular exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components and, therefore, likely to be driven by noise.
-Points are included if at least one of true value or prediction is unlikely to be noise-driven.
-In the example graph in Figure~\ref{fig:aatc_basic_4q_excl_box}, only point 1 is excluded.
-An exclusion area along one axis, for example, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
-This particularly suits \ac{mnf} where small amounts of noise are inevitable.
-A cross-shaped exclusion area, $E = \{(x, y) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
-A point is only included if no component is likely to be driven by noise.
-For the latter two methods, points 1 and 7 in Figure~\ref{fig:aatc_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:aatc_basic_4q_excl_cross}, respectively.
+The estimators accept various shapes of the exclusion area (see Figure~\ref{fig:aatc_4q}).
+A rectangular exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components.
+%Points are included if the true value or prediction is unlikely to be noise-driven.
+%In the example graph in Figure~\ref{fig:aatc_basic_4q_excl_box}, only point 1 is excluded.
+An exclusion area along one axis, for example, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
+%At least for \ac{mnf}, small amounts of noise are inevitable.
+A cross-shaped exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
+%A point is only included if no component is likely to be driven by noise.
+%For the latter two methods, points 1 and 7 in Figure~\ref{fig:aatc_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:aatc_basic_4q_excl_cross}, respectively.
 
@@
 Confidence intervals can account for the estimation uncertainty of the measures above.
-Bootstrap confidence intervals are based on resampling and not on parametric assumptions as classical confidence intervals are~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
+Bootstrap confidence intervals are based on resampling and not on parametric assumptions as classical confidence intervals are~\parencite[for introductions see][]{Hesterberg2011, Bittmann2021}.
 Many new samples are drawn with replacement from the dataset, and the statistic of interest is computed for each sample, yielding an estimate for the distribution of the statistic of interest.
-Based on the derived \enquote{new} samples of the statistic, the confidence intervals can be derived through differently sophisticated bootstrapping methods.
+Based on the derived \enquote{new} samples of the statistic, the confidence intervals can be derived through different bootstrapping methods.
 We use the \ac{bca} approach for bootstrapping in the following, as it holds the confidence level for small and large samples and has a moderate computation time (see the simulation study in Appendix~\ref{subsec:appendix-aatc-bootstrap}).
@@
 The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
-A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the \ac{atc} ability of a prediction for a certain change $\diffxrv = \xcond$ of the \ac{mnf}.
+A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the \ac{atc} of a prediction for a specific change $\diffxrv = \xcond$ of the \ac{mnf}.
 Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct direction given a predicted change of $\xcond$.
@@
 for $\xcond \neq 0$ through a \ac{kde}.
-A comprehensive introduction to multivariate \ac{kde} can be found in \textcite{Gramacki2018}, and implementations are available in many programming languages, for example, in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
+\textcite{Gramacki2018} provides a comprehensive introduction to multivariate \ac{kde}, and implementations are available in many programming languages, for example, in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
 The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ for all values of $\xcond \in \R$.
@@
 Assessing $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ graphically by drawing $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ against $\xcond$ eases the simultaneous evaluation of various $\xcond$.
-Furthermore, the graph facilitates the comparison of various methods in a single graph, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = \xcond)$ with respect to $x$ in the \ac{atc} can be detected.
+Furthermore, the graph facilitates the comparison of various methods in a single graph, and asymmetries of $P(\diffxrv \diffyrv > 0 | \diffx = \xcond)$ with respect to $\xcond$ in the \ac{atc} can be detected.
 We refer to the plot as a \textit{conditional ATC plot}.
@@
 
-In nowcasting and forecasting, probabilistic predictions have become more prevalent in recent years (see, for example, Sections~\ref{sec:application-covid} and~\ref{sec:application-eda} or the review in~\cite{Gneiting2014}).
+In nowcasting and forecasting, probabilistic predictions have become more prevalent in recent years (see Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
 In this section, we develop \ac{atc} assessments for probabilistic measurements and nowcasts.
 Probabilistic predictions issue a probability distribution for the quantity of interest based on their available information and, thus, include a point estimate and information on the prediction uncertainty and quantiles simultaneously.
-For \ac{atc} analysis, probabilistic predictions thus also contain a probability of a positive or negative change.
-In this section, we extend the \ac{atc} assessment to probabilistic predictions.
-
-Probabilistic predictions can take the form of a \ac{cdf}, \ac{pdf}, or quantiles.
-The \ac{cdf} is the most general and can be used to derive the others, given they exist.
+Probabilistic predictions thus also contain a probability of a positive or negative change.
+For \ac{atc} assessment, we compare the predicted probability of positive change, denoted by $p_t$, with the occurrence of positive changes.
+
+Probabilistic predictions can be a \ac{cdf}, \ac{pdf}, or quantiles.
+The \ac{cdf} is the most general and can be used to derive the others, given that they exist.
 Let us first assume that the prediction is a \ac{cdf}, and that $y_{t-l}$ is known at time $t$ (see Table~\ref{tab:notation}).
-In Appendix~\ref{subsec:app-probabilistic-atc-evaluation}, we extend the analysis for predictions given as quantiles or unknown true values.
-
-Let for a forecast $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for target time $t$ and issue time $t - l$, where the index is analogous to the point prediction notation of Section~\ref{subsec:notation}.
-Thereby, $F_{t | t-l} (x)$ specifies the forecastes probability that the quantity of interest is at most $x$.
-For \ac{atc} assessment, we compare the predicted probability of positive change, denoted by $p_t$, with the occurrence of positive changes.
-A positive change occurs for any value at $t$ larger than the true value $y_{t-l}$, as then the change compared to $y_{t-l}$ is positive.
-The \ac{cdf} $F_{t | t-l} (y_{t-l})$ yields the predicted probability of any value at most $y_{t-l}$, and, thus, a negative change.
-Accordingly, $1 - F_{t | t-l} (y_{t-l})$ is the probability of a positive change.
-As for the differences in Section~\ref{subsec:notation}, the computation differs slightly for nowcasts and forecasts.
-For forecasts, the probability is calculated by
-\begin{equation*}
-    p_t = 1 - F_{t | t-l} (y_{t-l})\quad t = l, \dots, T;
-\end{equation*}
-while for nowcasts it is given by
+Appendix~\ref{subsec:app-probabilistic-atc-evaluation} extends the analysis to quantile predictions or unknown true values.
+
+Let for a forecast $F_{t | t-l} (x)$ denote the predicted \ac{cdf} for target time $t$ and issue time $t - l$, where the index is analogous to the point notation of Section~\ref{subsec:notation}.
+The \ac{cdf} $F_{t | t-l} (x)$ specifies the forecasted probability that the quantity of interest is at most $x$.
+A positive change occurs for any value at $t$ larger than the true value $y_{t-l}$ and the \ac{cdf} $F_{t | t-l} (y_{t-l})$ yields the predicted probability of any value at most $y_{t-l}$, and, thus, a negative change.
+Accordingly, the forecasted probability of a positive change is
+\begin{equation*}
+    p_t = 1 - F_{t | t-l} (y_{t-l})\quad t = l, \dots, T.
+\end{equation*}
+The computation differs slightly for nowcasts, that is,
 \begin{equation*}
@@
 \end{equation*}
-The probabilistic \ac{atc} evaluation is then based on the probabilities $\mathbf{p} = (p_t)_{t=l}^{T}$ and the observed changes $\mathbf{z} = (z_t)_{t=l}^{T}$.
-The predictive power of $\mathbf{p}$ for $\mathbf{z}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
-Dichotomous forecasts predict a binary outcome, such as a positive or negative change.
-Numerical measures for forecast quality are scoring rules.
+The predictive power of $\mathbf{p} = (p_t)_{t=l}^{T}$ for $\mathbf{z} = (z_t)_{t=l}^{T}$ can be assessed using probabilistic dichotomous forecast evaluation methods.
+Dichotomous forecasts predict a binary outcome, such as a positive or negative change, and are evaluated numerically using scoring rules or visually through reliability diagrams.
 
@@
 \end{equation*}
-Thereby, lower values indicate a higher probabilistic \ac{atc} ability of the considered method.
-The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\citep{Ranjan2010,Mitchell2011}.
-Calibration refers to the statistical consistency of forecasts and observations; that is, the event occurs with the issued probability.
-Sharpness refers to the spread of the forecast; higher probabilities are preferable as they convey a higher certainty.
+Lower values indicate the considered method's higher probabilistic \ac{atc}.
+The \ac{bs} assesses the calibration and sharpness of the forecast and the observation simultaneously~\citep{Ranjan2010, Mitchell2011}.
+Calibration refers to the statistical consistency of forecasts and observations; that is, the event occurs with the issued probability and is considered the more fundamental quality~\citep{Gneiting2007}.
+Sharpness refers to the spread of the forecast; probabilities close to zero and one are preferable as they convey a higher certainty.
 
@@
 The reliability diagram plots the observed frequency of the positive outcome against the (binned) predicted probability.
-For example, it shows the proportion of observed increases, given the predicted probability of increase was approximately $0.7$.
+For example, it shows the proportion of observed increases, given that the predicted probability of increase was approximately $0.7$.
 Ideally, the predicted probability equals the observed frequency, and the reliability diagram is a 45-degree line.
-Local deviations from the 45-degree line indicate a miscalibration of the forecast for specific probabilities.
+Local deviations from the 45-degree line indicate a miscalibration for specific forecast probabilities.
 Thus, the reliability visualizes the local and overall calibration simultaneously.
@@
 
-In this section, we evaluate the \ac{atc} in three different applications.
-In the first application in Section~\ref{sec:application-covid}, we consider the nowcasting of the seven-day hospitalization rate for COVID-19 in Germany.
-Section~\ref{sec:application-eda} applies the \ac{atc} assessment to forecast methods for the number of arrivals in a large emergency department.
-In the last application in Section~\ref{sec:application_measurement}, we consider the \ac{atc} assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.
-
-The examples illustrate the use of the methods of Section~\ref{sec:aatc} and yield further insights into the measures and plots.
-All code used to generate the analysis is available online.
-
 \subsection{Covid nowcasting} \label{sec:application-covid}
 
-Amid the COVID-19 pandemic, the importance of accurate and prompt nowcasts of the pandemic's progression has become evident.
-Various indicators measured the pandemic's spread and severity.
-In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 measures in November 2021, and the imposition of severe public restrictions was based on it~\citep{RobertKochInstitute2021}.
-The Robert Koch Institute (RKI) provided preliminary daily reports on the seven-day hospitalization rate.
-However, these reports were subject to severe delays and revisions in two sources.
+In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 in November 2021, and the imposition of severe public restrictions was based on it~\citep{RobertKochInstitute2021}.
+However, the publication of the definite rate was subject to severe delays and revisions in two sources.
 The first source is technical delays in the reporting process, for example, due to different authorities passing the data to the RKI~\citep{RobertKochInstitute2024}.
-The second, more systematic source is the structure of the seven-day hospitalization rate.
-To a given date, all the cases are allocated whose first positive test result was on that date and who were hospitalized in relation to the disease in the following.
-The seven-day hospitalization rate is the average number of those cases per $100,000$ inhabitants on the given date and six days before.
-Thus, the final seven-day hospitalization rate can only be reported with a significant delay of more than 70 days, as the hospitalization of infected inhabitants can occur much later than the first positive test.
-Nevertheless, as the seven-day hospitalization rate was considered a major indicator of the pandemic's development, many organizations and institutions started to issue nowcasts, including research teams and newspapers.
-To collect nowcasts of the seven-day hospitalization rate by different nowcast groups, the COVID19-Nowcasting-Hub was established~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
-The nowcasts contain the seven-day hospitalization rate's predictive mean, median, and other quantiles.
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[t]{0.48\textwidth}
-    \includegraphics{plots/covid_nowcast/00_true_data}
-    \caption{Realisations.}
-    \label{fig:app-covid-true}
-    \end{subfigure}\hfill
-    \begin{subfigure}[t]{0.48\textwidth}
-    \includegraphics{plots/covid_nowcast/00_nowcast_data}
-    \caption{Same-day nowcasts.}
-    \label{fig:app-covid-nowcast}
-        \end{subfigure}
-    \caption[True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022.]{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
-    The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
-    \label{fig:app-covid-true-nowcast}
-\end{figure}
-
-The data contains eight nowcasts from scientific and public institutions, nowcast communities, and a newspaper, using different input variables, calendar data, and length of training data.
-The model structures are diverse, including Bayesian models, generalized additive models, and parametric bootstrapping.
+The second, more systematic source is the structure of the seven-day hospitalization rate, allocating all Covid-related hospitalization to the date of the first positive test~\parencite[for a detailed description, see][]{Wolffram2023}.
+Nevertheless, as the seven-day hospitalization rate was considered a central indicator of the pandemic's development, various organizations and institutions started to issue nowcasts.
+The COVID19-Nowcasting-Hub collected various nowcasts in a predefined setup, including the seven-day hospitalization rate's predictive mean, median, and other quantiles~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+For information on the model design, nowcast tasks, and the data submission guidelines, we refer to~\citet{Wolffram2023}.
 Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the respective names in the COVID19-Nowcasting-Hub.
-For information on the model design, we refer to~\citet{Wolffram2023}.
 In addition to those nowcasts, \citet{Wolffram2023} construct two ensemble methods using the ensembles' mean or median.
 We denote them by ENS-MEAN and ENS-MED.
-\citet{Wolffram2023} describes the design of the nowcast tasks and the data submission guidelines for the teams stated in a preregistered study.
 In line with the initial study design, we consider the period from November 22, 2021, to April 29, 2022, as the evaluation period.
-We use the data from February 8, 2024, for the true values.
-The nowcasts are resolved with regard to all inhabitants and specific geographical and age breakdowns.
-We do not expand on the models' performance on specific regions or age groups in Germany and the probabilistic nowcast assessment.
-Figure~\ref{fig:app-covid-true-nowcast} displays the true and nowcast data for the evaluation period.
+We use the data from February 8, 2024, for the true values and focus on nowcasts for all inhabitants of Germany.
+Figure~\ref{fig:app-covid-true-nowcast} in Appendix~\ref{sec:appendix-application-covid} displays the true and nowcast data for the evaluation period.
 The time comprises the fourth wave's end in December 2021 and nearly the entire fifth wave of the pandemic in Germany, lasting until May 28, 2022~\citep{Tolksdorf2022}.
 
-Table~\ref{tab:app-covid-rmse} summarizes the point evaluation measures for the issued mean of the different models.
-The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\citep[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
-The best-performing models in terms of RMSE and MAE are the ILM and RKI models.
-The ensemble methods perform worse than the best models regarding the mean location.
-The performance of the models is diverse, with more than twice as high RMSE values for the worst models compared to the best models.
-Note that the high values for the EPI model could be driven by an exceptionally far-off value at the end of the evaluation period.
-
-In addition to close inspection of the point evaluation measures, assessing the \ac{atc} of the nowcasts is crucial.
-To assess the impact of taken measures and the direction of the curve, it is essential to distinguish between rising and falling hospitalization rates.
+To assess the impact of taken measures and the direction of the curve, the knowledge of whether hospitalization rates rise or fall is essential.
+Thus, the \ac{atc} assessment is particularly relevant for the nowcasts of the seven-day hospitalization rate.
 If hospitalization rates rise, measures should be tightened, while falling rates might allow for loosening measures.
 Asymmetries are especially relevant for assessing whether some models are better at recognizing a fall than a rise or vice versa.
+
+\subsubsection*{Results}
 
@@
 The evaluation period comprises 159 days.
-}
+\enquote{RMSE} and \enquote{MAE} are accuracy measures, while \enquote{Count} lists the number of non-missing values.
+The RMSE orders the models.}
     \label{tab:app-covid-rmse}
@@
 
-\subsubsection*{Results}
-
-In the following, we apply the \ac{atc} assessment of Section~\ref{sec:aatc} to the nowcasts of the seven-day hospitalization rate.
-We report the \ac{atc} for the horizons one, seven, and 14 days.
-While horizon one assesses the short-term \ac{atc}, horizons seven and 14 consider the medium-term \ac{atc}.
-The horizons seven and 14 are particularly interesting, reflecting a usual period until new policy changes are taken.
-
-Before stating the results, we provide background information on the marginal distributions of the actual values and nowcasts for the different horizons.
-Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} presents marginal statistics such as standard deviation and quantiles of the nowcasts and true values.
-The variability and general level of differences grow with the horizon.
-The standard deviation increases from roughly 300 for horizon one to 1,200 for horizon seven and 2,000 for horizon 14 days.
-Similarly, the 10\%-quantile of differences increases.
-The 10\%-quantile is used for the exclusion areas in the \ac{atc} assessment.
-The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the absolute differences.
+Table~\ref{tab:app-covid-rmse} summarizes the non-\ac{ATC}-aware point evaluation measures for the issued mean of the different models.
+The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\citep[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
+The best-performing models in terms of RMSE and MAE are the ILM and RKI models.
+The ensemble methods ENS-MED and ENS-MEAN perform worse than the best models regarding the mean location.
+The performance of the models is diverse, with more than twice as high RMSE values for the worst models compared to the best models.
+Note that the high values for the EPI model could be driven by an exceptionally far-off value at the end of the evaluation period.
+
+In the following, we apply the \ac{atc} assessment for the short-term horizons one and medium-term horizons seven and 14 days.
+The horizons seven and 14 reflect a typical period until new policy changes are taken.
+We start by providing background information on the marginal distributions of the actual value and nowcast changes for the different horizons in Table~\ref{tab:app-covid-marginals} in Appendix~\ref{sec:appendix-application-covid} such as standard deviation and quantiles of the nowcasts and true values.
+The variability and general level of changes grow with the horizon: The standard deviation increases from roughly 300 for horizon one to 1,200 for horizon seven and 2,000 for horizon 14 days.
+Similarly, the 10\%-quantile of changes, the basis for the exclusion area size, increases.
+The exclusion area is rectangular; a point falls within it if both $\diffy$ and $\diffx$ are below the respective 10\%-quantile of the absolute changes.
 Thus, points are still included in the \ac{atc} assessment if they are large in one dimension but not in the other, thus ensuring that substantial changes in, for example, $\diffy$ are to be recognized by the nowcast and vice versa.
@@
 
-Table~\ref{tab:app-covid-atc-ratios-lag-7} lists the \ac{atc} ratios for all models with and without exclusion areas for the horizon of seven days.
+Table~\ref{tab:app-covid-atc-ratios-lag-7} lists the \ac{atc} ratios for all models without and with exclusion areas for the horizon of seven days.
 The \ac{atc} ratios without exclusion area range from 0.72 to 0.85 for the horizon of seven days.
@@
 The model with the highest \ac{atc} ratio is the ILM model, and the model with the lowest is the RKI model.
-The confidence intervals between all models overlap.
-The positive \ac{atc} ratio implies a similar ranking of the models, while the negative ratio is second best for the RKI model.
+The confidence intervals between all models without and with exclusion areas overlap.
+The positive \ac{atc} ratio implies a similar ranking of the models than the overall \ac{atc}, while the negative ratio provides a different ranking, for example, for the RKI model.
 For the horizons of one and 14 days, we refer to Table~\ref{tab:app-covid-atc-ratios-lag-1-14} in Appendix~\ref{sec:appendix-application-covid}.
@@
 If RKI or ILM issues a fall in the hospitalization rate, the probability of a fall is higher than if RIVM or ENS-MED issues a fall.
-The opposite is the case for a nowcasted hospitalization rate increase, and the difference between the models' performance is larger than for a fall.
+The opposite is the case for a nowcasted hospitalization rate increase, and the difference between the models' performance is more prominent than for a fall.
 Similar observations can be made for the horizon of 14 days in Figure~\ref{fig:app-covid-cond-prob-14}.
@@
 The \ac{atc} ratios for various exclusion areas are shown in Figure~\ref{fig:app-covid-atc-ratio-7}.
-In general, the \ac{atc} ratio increases with larger exclusion areas.
+The \ac{atc} ratio generally increases with larger exclusion areas.
 While the RIVM and ENS-MED \ac{atc} ratios evolve similarly, the RKI and ILM \ac{atc} ratios get closer.
@@
 
-Figure~\ref{fig:app-covid-atc-ratio-7} shows that the \ac{atc} ratio increases with larger exclusion areas.
-This indicates that if the model predicts a large change, the direction is more accurate than when a small change is predicted.
+Figure~\ref{fig:app-covid-atc-ratio-7} shows that larger exclusion areas increase the \ac{atc} ratio, indicating that the predicted direction is more accurate for large predicted changes.
 
 The probabilistic \ac{atc} assessment shows that the models are better than random guessing.
-In general, the reliability diagram cannot provide information if specific probabilities are issued scarcely.
+The reliability diagram cannot provide information if specific probabilities are issued scarcely.
 Thus, the reliability diagrams for the horizons of seven and 14 days do not contain information on moderate probabilities.
@@
 
-In a second example, we consider forecasting the number of arrivals in a large emergency department per hour.
-Good forecasts are crucial for planning staff and resources.
-In a study by \citet{Rostami-Tabar2023}, several models predict hourly outcomes.
+In a second example, we consider forecasting the number of arrivals in a large emergency department per hour with data and models by \citet{Rostami-Tabar2023}.
 Every 12 hours, the models issue hourly forecasts for the next 48 hours.
-Thus, management can take measures according to the expected number of arrivals, for example, by redeploying staff and reconfiguring units.
-
-\citet{Rostami-Tabar2023} publish means and probabilistic quantile forecasts, which are evaluated through RMSE, pinball loss, pinball skill scores, and PIT-histograms.
-The models are trained on data from April 1, 2014, to February 28, 2018, and evaluated on data from March 1, 2018, to February 28, 2019.
-For further notes on the models and the evaluation, we refer to \citet{Rostami-Tabar2023}.
-From the issued forecasts, we use the mean as a point forecast for the \ac{atc} assessment and evaluate the probabilistic \ac{atc} subsequently.
-We use the forecasts issued at the first time point for every target time.
-Thus, the forecasts are issued 36 to 48 hours ahead of the target time, and the emergency department management has time to adjust the measures according to the expected number of arrivals.
+
+\citet{Rostami-Tabar2023} publish means and probabilistic quantile forecasts for various models and input data.
+We use the published mean as a point forecast for the \ac{atc} assessment and evaluate the probabilistic \ac{atc} based on the quantile forecasts subsequently.
 Considering only the forecasts of at least 36 hours ahead, we restrict the evaluation period to March 2, 2018, at noon, to February 28, 2019, at 23:00, comprising 8,724 hours.
@@
 The number of arrivals has a strong weekly and daily pattern.
-Thus, we consider the horizons of 72 hours, the last already observed shift of the same hour of day, and seven days, the previous shift of the same hour and day.
+Thus, we consider the horizons of 72 hours, the last already observed shift of the same hour of day, and seven days, the previous shift of the same hour and day, in \ac{atc} assessment.
+
+\subsubsection*{Results}
+
 Table~\ref{tab:app-eda-point-evaluation} lists the point evaluation measures and the count of available forecasts.
 The best-performing models regarding RMSE and MAE are the NBI-2 and Poisson-2 models.
-More than 8,600 forecasts are available for all models, with differences in the number due to missing values on four afternoons in 2018.
-Note that the reported values for the RMSE differ from those in \citet{Rostami-Tabar2023}.
-In contrast to their work, we use only the forecast data at least 36 hours ahead rather than the entire forecast data for evaluation.
+More than 8,600 forecasts are available for all models, with changes in the number due to missing values on four afternoons in 2018.
 
@@
 \caption[Point evaluation measures for the emergency department arrival forecasting models.]{Point evaluation measures for the emergency department arrival forecasting models.
-The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.}\label{tab:app-eda-point-evaluation}
+The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.
+Note that the reported values for the RMSE differ from those in \citet{Rostami-Tabar2023} due to differences in the evaluation period.}\label{tab:app-eda-point-evaluation}
 \end{table}
 
-\subsubsection*{Results}
-
-We start by analyzing the marginal distributions for the changes of forecasts and actual values for the horizons of three and seven days in Table~\ref{tab:app-eda-marginals}, again.
-Thereby, the computed difference aligns with Section~\ref{subsec:notation}, defined as the difference between the forecasted mean and true value of three and seven days before, as the actual value is available when issuing the forecast.
-The fraction of positive differences varies between 0.39 and 0.63 for the horizon of three days and between 0.37 and 0.63 for the horizon of seven days.
-The variability of differences decreases for the larger horizon for most models; only for the ETS model does it increase.
-The 10\%-quantile of the differences is between zero and one for all models and horizons.
+We start by analyzing the marginal distributions for the predicted and observed changes for the three- and seven-day horizons in Table~\ref{tab:app-eda-marginals}, again.
+The computed difference aligns with Section~\ref{subsec:notation}, that is, the difference between the forecasted mean and true value of three and seven days before, as the actual value is available when issuing the forecast.
+The positive change fraction varies between 0.39 and 0.63 for the horizon of three days and between 0.37 and 0.63 for the horizon of seven days.
+The variability of changes decreases for the larger horizon for most models; only for the ETS model does it increase.
+The 10\%-quantile of the changes is between zero and one for all models and horizons.
 Thus, we use an exclusion area of size 1.
@@
     \input{plots/ed_arrival/10_marginal_analysis}
-    \caption[Marginal analysis of the forecast and true differences in emergency department arrival forecasting.]{Marginal analysis of the forecast and true differences in emergency department arrival forecasting.
-The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.
+    \caption[Marginal analysis of the forecast and true changes in emergency department arrival forecasting.]{Marginal analysis of the forecast and true changes in emergency department arrival forecasting.
+The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the changes' absolute values.
 Column (2) shows the fraction of values not in the exclusion area of size one.}
@@
 The negative and positive \ac{atc} ratios differ for all models and horizons.
-For some models, for example, the GBM-2 model, the positive \ac{atc} ratio is higher; for some models, for example, the tbats model, the negative \ac{atc} ratio is higher.
+For some models, for example, the GBM-2 model, the positive \ac{atc} ratio is higher than the negative \ac{atc} ratio, and for some models, for example, the tbats model, vice versa.
 The confidence interval width is at most 0.02 for the \ac{atc} ratios and at most 0.03 for the positive and negative \ac{atc} ratios.
@@
     \end{subfigure}
-    \caption[Conditional ATC plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} ability in emergency department arrival forecasting.]{Conditional \ac{atc} plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} ability in emergency department arrival forecasting.
+    \caption[Conditional ATC plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} in emergency department arrival forecasting.]{Conditional \ac{atc} plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} in emergency department arrival forecasting.
 The plots of NBI-2 and Poisson-2 are indistinguishable.}
@@
 The \ac{atc} is consistent for the two horizons, with the models' relative \ac{atc} evolving similarly for the two horizons.
-The models' \ac{atc} is generally higher for the smaller horizon, but the differences are minor, and confidence intervals overlap.
-
-The \ac{atc} differs for all models for positive and negative predicted change directions.
+The models' \ac{atc} is generally higher for the smaller horizon, but the changes are minor, and confidence intervals overlap.
+
+The positive and negative \ac{atc} ratios differ for all models.
 While some models, such as GBM-2 and qreg-1, have the highest positive \ac{atc} ratio, others, such as Poisson-2 and NBI-2, have the highest negative \ac{atc} ratio.
@@
 
-The results of the probabilistic \ac{atc} assessment endorse the point \ac{atc} assessment and assign the best scores to NBI-2 and Poisson-2.
+The probabilistic \ac{atc} assessment results endorse the point \ac{atc} assessment and assign the best scores to NBI-2 and Poisson-2.
 The reliability diagrams show that they underestimate the fraction of increases slightly.
@@
 In the last briefer example, we consider the \ac{atc} assessment of measurement data.
-The data is from the MIMIC-III database, including various information on patients in critical care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}).
-The data is publicly available and also includes numerical measurement data such as heart rate, blood pressure, and oxygen saturation in a waveform database \citetext{\citealp{Moody2017}; available through \citealp{Goldberger2000}}.
-
-For some patients, the data includes \ac{abp} and \ac{nbp} measurements.
-While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods.
-For an overview of blood pressure measurement methods, see \citet{Saugel2014}.
+The data is from the MIMIC-III database, including various numerical measurement data, such as heart rate, blood pressure, and oxygen saturation, on patients in critical care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}, \citealp{Moody2017}; available through \citealp{Goldberger2000}).
+We focus on \acf{abp} and \acf{nbp} measurements.
+While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods~\citet[see][]{Saugel2014}.
 For critical patients, changes in blood pressure can be crucial for the treatment.
@@
 Thus, we assess the \ac{atc} of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
-The database contains 64,168 numerical data records.
-One data record includes all numerical measurements for one patient.
+The database contains 64,168 numerical data records, comprising all numerical measurements for one patient.
 The measured signals vary in length, frequency, and type of measurement.
@@
 2,548 include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one measurement of systolic \ac{abp} and \ac{nbp} at the same time; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
-
 We consider the horizons of one minute, five minutes, and 15 minutes for the \ac{atc} assessment, as those are typical intervals of NBP measurements.
@@
 
-Again, we exclude the smallest 10\% of absolute differences in \ac{atc} assessment.
+Again, we exclude the smallest 10\% of absolute changes in \ac{atc} assessment.
 The resulting four-quadrant plots of the mean and systolic blood pressure measurements for the different horizons are shown in Figure~\ref{fig:app-mimic-4q}.
@@
 Figure~\ref{fig:app-mimic-cond-prob} shows the conditional \ac{atc} plots for the different horizons and the systolic and mean blood pressure measurements.
-It becomes apparent that the systolic measurements have a higher \ac{atc} ability than the mean measurement, except for small negative predicted changes.
-This aligns with the \ac{atc} ratios, but the confidence intervals overlap.
+It becomes apparent that the systolic measurements have a higher \ac{atc} than the mean measurements, except for small negative predicted changes.
 
@@
 Whether these points are due to measurement errors or extreme values is not distinguishable.
-Some authors argue to exclude the measurements below the 10\%-quantile of the absolute differences and the points above the 90\%-quantile~\citep[see][]{Critchley2010}.
+Some authors argue to exclude the measurements below the 10\%-quantile of the absolute changes and the points above the 90\%-quantile~\citep[see][]{Critchley2010}.
 We do not follow this approach here, as the extreme values are not necessarily measurement errors and could be particularly relevant.
 
-The differences between positive and negative predicted changes are small in this example.
-The positive and negative \ac{atc} ratios have overlapping confidence intervals, and the conditional \ac{atc} plots do not contain prominent deviations in the course.
-This aligns with the four-quadrant plots, where no apparent asymmetry is visible.
+The difference between positive and negative \ac{atc} ratios is small in this example.
+The positive and negative \ac{atc} ratios have overlapping confidence intervals, the conditional \ac{atc} plots do not contain prominent deviations in the course, and the four-quadrant plots do not display asymmetries.
 
@@
 
-In this paper, we examine various methods to assess the \acf{atc} for \acf{mnf}, that is, whether predict the correct direction of changes.
+In this paper, we examine various methods to assess the \acf{atc} for \acf{mnf}, that is, whether it predicts the correct direction of changes.
 While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting, the predict the correct direction of changes assessment using the computed predicted and observed change is similar.
@@
 The computation of differences is a standard procedure in time series analysis to remove sequential dependence, but, in general, some could remain, and the estimators could account for it.
-Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation using time-series bootstrap methods~\citep{Hardle2003,Kreiss2012}.
-
-The estimators of Section~\ref{subsec:atc-measures} do not account for imbalances in the number of observed positive and negative changes \citep[for a theoretical analysis, see][Chapter 3]{Jolliffe2012}.
+Similarly, the bootstrap confidence intervals could be adapted to consider sequential correlation using time-series bootstrap methods~\citep{Hardle2003, Kreiss2012}.
+
+The estimators of Section~\ref{subsec:atc-measures} do not account for imbalances in the number of observed positive and negative changes \citep[for theoretical analysis, see][Chapter 3]{Jolliffe2012}.
 Significant differences in the number of observed positive and negative changes are unlikely in the \ac{atc} setting, as $\diffy$ is obtained from differencing time series data and occur, for example, if the true value contains a few high jumps in one direction and many smaller jumps in the other.
@@
 
+%TC:ignore
 \section*{Declarations}
@@
 
-The datasets analysed during the study are available in the corresponding databases or repositories \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024,Rostami-Tabar2023, Moody2017}.
+The datasets analyzed during the study are available in the corresponding databases or repositories \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024, Rostami-Tabar2023, Moody2017}.
 All code is available in the repository [GITHUB LINK WILL BE PROVIDED AFTER REVIEW].
@@
 
-JR greatfully acknowledges funding by the Bischoefliche Studienstiftung Cusanuswerk (Bonn, Germany).
+JR gratefully acknowledges funding from the Bischoefliche Studienstiftung Cusanuswerk (Bonn, Germany).
 
@@
 
-OG, BL and JR developed the theoretical formalism.
+OG, BL, and JR developed the theoretical formalism.
 BL and JR implemented the methods.
@@
 BL and JR wrote the manuscript.
-All authors discussed the results, critically reviewed and contributed to the final manuscript.
+All authors discussed the results, critically reviewed them, and contributed to the final manuscript.
 
@@
 For the small sample sizes up to $T = 168$, only the \ac{bca} method keeps the confidence interval size and yields slightly wider confidence intervals.
-The method's results do not differ for the larger sample sizes.
+The method's results are similar for the larger sample sizes.
 The computation time for the \ac{bca} method is slightly larger than for the other methods, but all methods have a moderate computation time.
@@
 Cross-validation least squares takes a considerably larger computation time.
-It does not converge for the first data set and the second data set with exclusion area and yields a bandwidth too small for the second data set.
+It converges neither for the first nor the second data set with an exclusion area and yields a bandwidth too small for the second data set.
 The rule of thumb is the fastest method but tends to oversmooth.
@@
 \section{Additional results for Section~\ref{sec:application-covid}}\label{sec:appendix-application-covid}
+
+\begin{figure}
+    \centering
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/covid_nowcast/00_true_data}
+    \caption{Realisations.}
+    \label{fig:app-covid-true}
+    \end{subfigure}\hfill
+    \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics{plots/covid_nowcast/00_nowcast_data}
+    \caption{Same-day nowcasts.}
+    \label{fig:app-covid-nowcast}
+        \end{subfigure}
+    \caption[True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022.]{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+    The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
+    \label{fig:app-covid-true-nowcast}
+\end{figure}
 
@@
     \input{plots/covid_nowcast/10_marginal_analysis}
-    \caption[Marginal analysis of the Covid nowcast and true differences for the horizons one, seven, and 14 days.]{Marginal analysis of the Covid nowcast and true differences for the horizons one, seven, and 14 days.
-    The column (1), $l=l$ shows the number of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
+    \caption[Marginal analysis of the Covid nowcast and true changes for the horizons one, seven, and 14 days.]{Marginal analysis of the Covid nowcast and true changes for the horizons one, seven, and 14 days.
+    The column (1), $l=l$ shows the number of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the changes' absolute values.}
     \label{tab:app-covid-marginals}
@@
 \end{figure}
+%TC:endignore
 

### 2024-04-11 paper.tex
Update paper

@@
 
-The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
+The appropriate evaluation of \ac{mnf} methods is fundamental for assessing the method's suitability for the task.
+If the development of the underlying values is of interest, the ability to predict the correct direction of changes, we call \acf{atc}, is crucial.
 Although the three areas differ in the framework and methods used, they share the goal of predicting an unknown value, and thus, similar techniques can be used in the evaluation.
@@
 Forecasting methods predict the future based on historical data, patterns, and exogenous factors.
-In medicine, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in][]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008, Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
+In medicine and healthcare, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in][]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008, Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
 The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
 Methodologically evolved from forecasting~\citep{Browning1989}, nowcasting methods focus on predictions for the present, the immediate future, and the recent past~\citep{Banbura2013, WorldMeteorologicalOrganizationWMO2017}.
-Nowcasting methods use high-frequency indicators related to the target value and estimate the value of a target value for a specific time based on current preliminary measurements, which are finalized with a considerable delay.
-Thus, it focuses on ongoing predictions due to updating information for the same time~\citep{Castle2017}.
-In medicine, epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{Wu2021} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{Gunther2021, Wolffram2023}.
+Nowcasting methods use high-frequency indicators or preliminary measurements related to the target value and focuses on ongoing predictions due to updating information for the same time~\citep{Castle2017}.
+Epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{Wu2021} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{Gunther2021, Wolffram2023}.
 Measurement aims to obtain accurate and reliable data about the current state of a system based on sensors.
-A parameter or value can be measured regularly over a certain period to evaluate the system's development.
+Through repeated measurements a system's development is tracked.
 When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
+
 In forecasting and nowcasting, the evaluation is based on accuracy measures such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
@@
 The second, more systematic source is the structure of the seven-day hospitalization rate, allocating all Covid-related hospitalization to the date of the first positive test~\parencite[for a detailed description, see][]{Wolffram2023}.
-Nevertheless, as the seven-day hospitalization rate was considered a central indicator of the pandemic's development, various organizations and institutions started to issue nowcasts.
+Nevertheless, as the seven-day hospitalization rate was considered a central indicator of the pandemic's development, various organizations and institutions issued nowcasts.
 The COVID19-Nowcasting-Hub collected various nowcasts in a predefined setup, including the seven-day hospitalization rate's predictive mean, median, and other quantiles~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
 For information on the model design, nowcast tasks, and the data submission guidelines, we refer to~\citet{Wolffram2023}.
-Table~\ref{tab:app-covid-models} in the appendix lists the different abbreviations and the respective names in the COVID19-Nowcasting-Hub.
+Table~\ref{tab:app-covid-models} matches the abbreviations here to the respective names in the COVID19-Nowcasting-Hub.
 In addition to those nowcasts, \citet{Wolffram2023} construct two ensemble methods using the ensembles' mean or median.
@@
 
-\begin{table}[]
+\begin{table}
     \centering
@@
     \caption[Point evaluation measures for the issued mean of the different models in Covid nowcasting.]{Point evaluation measures for the issued mean of the different models in Covid nowcasting.
-The evaluation period comprises 159 days.
 \enquote{RMSE} and \enquote{MAE} are accuracy measures, while \enquote{Count} lists the number of non-missing values.
-The RMSE orders the models.}
+The RMSE orders the models.
+The evaluation period comprises 159 days and only few nowcasts are missing~\citep[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
+Note that the high values for the EPI model could be driven by an exceptionally far-off value at the end of the evaluation period (see Figure~\ref{fig:app-covid-true-nowcast}).}
     \label{tab:app-covid-rmse}
@@
 
-Table~\ref{tab:app-covid-rmse} summarizes the non-\ac{ATC}-aware point evaluation measures for the issued mean of the different models.
-The models issue same-day nowcasts for nearly all 159 days of the evaluation period~\citep[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
+Table~\ref{tab:app-covid-rmse} summarizes the non-\ac{atc}-aware point evaluation measures for the issued mean of the different models.
 The best-performing models in terms of RMSE and MAE are the ILM and RKI models.
@@
 The performance of the models is diverse, with more than twice as high RMSE values for the worst models compared to the best models.
-Note that the high values for the EPI model could be driven by an exceptionally far-off value at the end of the evaluation period.
 

### 2024-04-11 paper.tex
Update on Overleaf.

@@
 \affil[1]{\orgdiv{Institute for Operations Research}, \orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\street{Kaiserstr. 12}, \city{Karlsruhe}, \postcode{76131}, \country{Germany}}}
-
+%TC:ignore
 \abstract{ 
@@
 }
-
+%TC:endignore
 \keywords{}
@@
 where $n_t$ is a truncated normal distribution with mean one and standard deviation 0.5, truncated at 0, and $b_t$ is a symmetric Bernoulli random variable with parameter $k$.
-For a time-varying \ac{atc} ratio, the parameter $k$ is modified to have a wave-shape over time, that is,
+For a time-varying \ac{atc} ratio, the parameter $k$ is modified to have a wave-shape function over time, that is,
 \begin{equation*}
@@
 
-We examine three methods for bootstrapping: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
+We examine three methods for bootstrapping for computing confidence intervals for the \ac{atc} ratio: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
 In the \textit{percentile} approach, the confidence interval for the level $\alpha$ is built directly from the empirical distribution of the bootstrap estimators.
@@
 
-We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
+We vary the number of time steps $T$ to take typical time-series values, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
 The considered datasets are the first dataset with asymmetric dependence and the second dataset outlined in Appendix~\ref{subsec:app-ATC-data-generation}.
@@
 
-We examine the resulting conditional \ac{atc} plots for the three well-known selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\citep{Seabold2010} in Figure~\ref{fig:atc-cond-prob-bw}.
+We examine the resulting conditional \ac{atc} plots for the three well-known \ac{kde} bandwidth selectors, rule-of-thumb, cross-validation maximum likelihood, and cross-validation least squares using the \verb|statsmodels| Python package~\citep{Seabold2010} in Figure~\ref{fig:atc-cond-prob-bw}.
 While the rule-of-thumb is based only on the covariance matrix, the other two numerically optimize the bandwidth with a hold-one-out least squares or likelihood objective function.

### 2024-04-11 paper.tex
Update on Overleaf.

@@
 Section~\ref{subsec:aatc-probabilistic} outlines the assessment of probabilistic \ac{atc} for nowcasts and forecasts and specifies the computation for predictions in terms of a \ac{cdf} and known true values.
-Here, we outline the computation for quantile forecasts and probabilistic true values.
+Here, we outline the computation for quantile forecasts and yet unknown, probabilistic true values.
+
 If forecasts or nowcasts are given as quantiles, $p_t$ can be determined by interpolations among the quantiles.
@@
 \end{equation*}
-The approach does not assign probabilities for $y_t$ smaller than the smallest quantile $q_{1/p}$ or larger than the largest quantile.
+The approach does not yet assign probabilities for $y_t$ smaller than the smallest quantile $q_{1/p}$ or greater than the largest quantile.
 As a simple extension, we assume that the probability mass is uniformly distributed on an interval of the same length as the nearest interval specified by the quantiles.
@@
     \caption[ATC ratios for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.]{\Ac{atc} ratio $\acc$, positive \ac{atc} ratio $\accp$, and negative \ac{atc} ratio $\accm$ for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.
-The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values.
+The exclusion areas are rectangles centered on the zero points with a width and height to exclude the 10\%-quantile of the absolute values of nowcast or true values.
 }

### 2024-04-14 paper.tex
Update on Overleaf.

@@
 
-% (Approximate) Word Count
-% Abstract
+% Ungefähre Wortanzahl des Hauptteils (von Introduction bis Conclusion): oben links auf 'Menu' und dann 'Word Count'
+% Ungefähre Aufteilung auf Abschnitte (nicht direkt in Overleaf sichtbar):
+% Introduction: 500
+% Assessment of ATC: 2225
+% Application: 3045
+% Discussion and Conclusion: 587
+
+% Abstract: 99
 
@@
 
-\title{Assessment of the ability to track changes for measurements, nowcasts, and forecasts}
+\title{The assessment of the ability of measurements, nowcasts, and forecasts to track changes}
 %\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
@@
 \abstract{ 
-In addition to measurements, nowcasts, or forecasts being numerically precise, it is essential [to know] how well they reflect changes in the underlying values.
-In this article, we focus on evaluating the ability to predict the correct direction of changes, which we refer to as the ability to track changes (ATC). 
-We present visual techniques and quantitative measures to assess this ability and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
-Worked-through examples illustrate the assessment using the proposed methods: COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements.
+Measurements, nowcasts, or forecasts ideally should correctly reflect changes in the values of interest.
+In this article, we focus on how to assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the ability to track changes.
+We present visual techniques and quantitative measures to assess the ability to track changes and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
+We exemplarily illustrate the proposed methods to assess the ability to track changes for nowcasting during the COVID-19 pandemic, patient admissions to an emergency department, and non-invasive blood pressure measurements.
 }
@@
 \section{Introduction}\label{sec:atc-introduction}
-
-The appropriate evaluation of \ac{mnf} methods is fundamental for assessing the method's suitability for the task.
-If the development of the underlying values is of interest, the ability to predict the correct direction of changes, we call \acf{atc}, is crucial.
-Although the three areas differ in the framework and methods used, they share the goal of predicting an unknown value, and thus, similar techniques can be used in the evaluation.
-
-Forecasting methods predict the future based on historical data, patterns, and exogenous factors.
-In medicine and healthcare, forecasting is used in various applications, such as population and personal health~\parencite[see, e.g., the review in][]{Soyiri2013}, emergency department demand forecasting~\parencite{Jones2008, Rostami-Tabar2023}, and emergency medical services~\parencite{HaugsboHermansen2021}.
-The forecast is computed based on the current value of the quantity of interest and an estimate of the development until the target time.
+BCS: We need to avoid non-standard abbreviations that confuse readers such as MNF. Please delete the abbreviation MNF throughout the text.
+
+Measurements, nowcasts, or forecasts ideally should correctly reflect changes in the values of interest.
+It is thus important to meticulously assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the ability to track changes (\ac{atc}).
+Although measurements, nowcasts, and forecasts fundamentally differ as they either measure or predict a value, similar methods can be used to assess their \ac{atc}.
+
+Forecasting methods predict the future based on historical data, patterns, or exogenous factors.
+A forecast is computed based on the current value of interest and an estimate of its future development.
+In medicine and healthcare, forecasting – for example – is used to predict patient volumes in emergency departments~\parencite{Jones2008, Rostami-Tabar2023} or the demand of emergency medical services~\parencite{HaugsboHermansen2021}.
+
 Methodologically evolved from forecasting~\citep{Browning1989}, nowcasting methods focus on predictions for the present, the immediate future, and the recent past~\citep{Banbura2013, WorldMeteorologicalOrganizationWMO2017}.
-Nowcasting methods use high-frequency indicators or preliminary measurements related to the target value and focuses on ongoing predictions due to updating information for the same time~\citep{Castle2017}.
-Epidemic nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{Wu2021} or correct the daily COVID case numbers for events that have occurred but have not yet been reported~\citep{Gunther2021, Wolffram2023}.
-Measurement aims to obtain accurate and reliable data about the current state of a system based on sensors.
-Through repeated measurements a system's development is tracked.
-When introducing new, cheaper, simpler, or less invasive measurement methods, they are evaluated against current measurement techniques, called the gold standard, by measuring the same quantity in different settings, such as various individuals or environments.
-
-In forecasting and nowcasting, the evaluation is based on accuracy measures such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
+Nowcasting methods use high-frequency indicators or preliminary measurements related to the value of interest and focus on updating predictions using currently available information~\citep{Castle2017}.
+Nowcasting, for example, can assess the current situation during an ongoing epidemic, considering the main pathogenic, epidemiological, clinical, and socio-behavioral factors~\citep{Wu2021} or provide daily numbers of COVID-19 cases for events that have occurred but have not yet been reported~\citep{Gunther2021, Wolffram2023}.
+
+Measurements aim to obtain accurate and precise values of a measureable quantity (measureand).
+BCS: Please cite https://pubmed.ncbi.nlm.nih.gov/32185615/ and https://pubmed.ncbi.nlm.nih.gov/32185616/
+Repeated measurements can be used to track changes in a value over time.  
+When introducing new measurement methods, they are evaluated against current reference methods, often called the "gold standard", by simultaneously measuring the same quantity with the new method and the reference method – often in various individuals or different clinical settings.
+
+In forecasting and nowcasting, the evaluation of the measurement performance is based on statistical methods quantifiying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
 None of them assesses the method's \ac{atc}.
-In a prediction competition on armed conflicts, the assessment of \ac{atc} recently gained attention as \citet{Vesco2022} used the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
+
+In a prediction competition on armed conflicts, the assessment of the \ac{atc} recently gained attention as \citet{Vesco2022} proposed the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-In measurement agreement analysis, a typical analysis includes the Bland-Altman analysis and a percentage error~\citep{Bland1986}.
-The assessment of \ac{atc} is a field of active research in recent years~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
-
-This article focuses on pure assessment of \ac{atc} accompanying traditional scores.
- We formalize \ac{atc} and present different variants of assessment of \ac{atc} measures that consider either noiseless data or data with noise and small non-informative changes (see Section~\ref{sec:aatc}).
-We introduce the conditional \ac{atc} plot, a new graphical method for assessing local \ac{atc} behavior, and review bootstrap methods for calculating confidence intervals (see Section~\ref{subsec:aatc-noise} and ~\ref{subsec:aatc-cond-prob}).
-We extend the concept of assessment of \ac{atc} to probabilistic predictions in Section~\ref{subsec:aatc-probabilistic}.
-The methods are applied to three applications from the abovementioned fields: COVID-19-nowcasting, emergency department arrival forecasting, and blood pressure measurement, and, thus, provide blueprints for future assessments.
+When evaluating the measurment performance of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error [BCS: Please cite: https://pubmed.ncbi.nlm.nih.gov/12578081/] are commonly used.
+How to best assess the \ac{atc} of measurement methods is a field of ongoing research~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
+
+In this article, we focus on how to assess the \ac{atc} of measurements, nowcasts, or forecasts to track changes.
+We formalize the concept of \ac{atc} and present visual techniques and quantitative measures to assess the \ac{atc} – considering both noiseless data and data with noise and small non-informative changes.
+We introduce the conditional \ac{atc} plot, a new graphical method for assessing local \ac{atc} behavior, and review bootstrap methods for calculating confidence intervals.
+We extend the concept of assessment of \ac{atc} to probabilistic predictions.
+We exemplarily illustrate the proposed methods to assess the \ac{atc} for nowcasting during the COVID-19 pandemic, patient admissions to an emergency department, and non-invasive blood pressure measurements – and thus provide blueprints for future assessments.
 The ready-to-use code for the assessment of \ac{atc} from the applications is available on [LINK WILL BE INSERTED].
@@
 
-As noted above, the assessment of \ac{atc} is based on the observed and predicted changes over a time horizon $l$.
-The \textit{observed} change is straightforward to compute for all types of \ac{mnf}.
+The assessment of \ac{atc} is based on the measured and predicted changes in a value of interest over a time horizon $l$.
+The \textit{measured} change is straightforward to compute for all types of \ac{mnf}.
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the actual values for nowcasting or forecasting, or gold standard measurements up to time $T$.
@@
 
-\section{Application to measurement, nowcasting, and forecasting data} \label{sec:application}
-
-\subsection{Covid nowcasting} \label{sec:application-covid}
-
-In Germany, the seven-day hospitalization rate was established as a central steering measure for COVID-19 in November 2021, and the imposition of severe public restrictions was based on it~\citep{RobertKochInstitute2021}.
-However, the publication of the definite rate was subject to severe delays and revisions in two sources.
-The first source is technical delays in the reporting process, for example, due to different authorities passing the data to the RKI~\citep{RobertKochInstitute2024}.
-The second, more systematic source is the structure of the seven-day hospitalization rate, allocating all Covid-related hospitalization to the date of the first positive test~\parencite[for a detailed description, see][]{Wolffram2023}.
-Nevertheless, as the seven-day hospitalization rate was considered a central indicator of the pandemic's development, various organizations and institutions issued nowcasts.
-The COVID19-Nowcasting-Hub collected various nowcasts in a predefined setup, including the seven-day hospitalization rate's predictive mean, median, and other quantiles~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
+\section{Application to medical/healthcare nowcasting, forecasting, and measurement data} \label{sec:application}
+
+\subsection{Nowcasting during the COVID-19 pandemic} \label{sec:application-covid}
+
+In Germany, the seven-day hospitalization rate was established as a central steering measure in November 2021 during the COVID-19 pandemic, and the imposition of severe public restrictions was based on it~\citep{RobertKochInstitute2021}.
+However, the publication of the definite hospitalization rate was substantially delayed and partially flawed for two main reasons.
+First, the reporting process was delayed because – among other reasons – different authorities reported the data to the RKI~\citep{RobertKochInstitute2024}.
+Second, the seven-day hospitalization rate allocated all Covid-19-related hospitalizations to the date of the first positive test~\parencite[for a detailed description, see][]{Wolffram2023}.
+Nevertheless, as the seven-day hospitalization rate was considered an important indicator of how the pandemic developed, various organizations and institutions issued nowcasts.
+The COVID-19-Nowcasting-Hub collected various nowcasts in a predefined setup, including the mean, median, and other quantiles of the predicted seven-day hospitalization rate~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
 For information on the model design, nowcast tasks, and the data submission guidelines, we refer to~\citet{Wolffram2023}.
-Table~\ref{tab:app-covid-models} matches the abbreviations here to the respective names in the COVID19-Nowcasting-Hub.
+Table~\ref{tab:app-covid-models} matches the abbreviations here to the respective names in the COVID-19-Nowcasting-Hub.
 In addition to those nowcasts, \citet{Wolffram2023} construct two ensemble methods using the ensembles' mean or median.
@@
 
-\subsection{Forecasting emergency department arrivals}\label{sec:application-eda}
-
-In a second example, we consider forecasting the number of arrivals in a large emergency department per hour with data and models by \citet{Rostami-Tabar2023}.
+\subsection{Forecasting patient admissions to an emergency department}\label{sec:application-eda}
+
+In a second example, we consider forecasting patient admissions to an emergency department per hour with data and models by \citet{Rostami-Tabar2023}.
 Every 12 hours, the models issue hourly forecasts for the next 48 hours.
@@
 The \ac{atc} perspective is easy for the management to understand and implement, as simple comparisons of the expected workload to a recent shift can be made.
-If, for example, the staff was near capacity during the last shift and an increase in the number of arrivals is expected, management can take measures to adjust the workload.
-
-The number of arrivals has a strong weekly and daily pattern.
+If, for example, the staff was near capacity during the last shift and an increase in the number of patient admissions is expected, management can take measures to adjust the workload.
+
+The number of patient admissions has a strong weekly and daily pattern.
 Thus, we consider the horizons of 72 hours, the last already observed shift of the same hour of day, and seven days, the previous shift of the same hour and day, in \ac{atc} assessment.
@@
 \input{plots/ed_arrival/00_point_evaluation_measures}
-\caption[Point evaluation measures for the emergency department arrival forecasting models.]{Point evaluation measures for the emergency department arrival forecasting models.
+\caption[Point evaluation measures for forecasting models of patient admissions to an emergency department.]{Point evaluation measures for the forecasting models for patient admissions to an emergency department.
 The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.
@@
     \input{plots/ed_arrival/10_marginal_analysis}
-    \caption[Marginal analysis of the forecast and true changes in emergency department arrival forecasting.]{Marginal analysis of the forecast and true changes in emergency department arrival forecasting.
+    \caption[Marginal analysis of the forecast and true changes in patient admissions to an emergency department.]{Marginal analysis of the forecast and true changes in patient admissions to an emergency department.
 The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the changes' absolute values.
@@
 
-\subsection{Invasive and non-invasive blood pressure monitoring} \label{sec:application_measurement}
-
-In the last briefer example, we consider the \ac{atc} assessment of measurement data.
-The data is from the MIMIC-III database, including various numerical measurement data, such as heart rate, blood pressure, and oxygen saturation, on patients in critical care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}, \citealp{Moody2017}; available through \citealp{Goldberger2000}).
-We focus on \acf{abp} and \acf{nbp} measurements.
-While non-invasive blood pressure measurement methods are relatively gentle, they are less accurate than invasive methods~\citet[see][]{Saugel2014}.
-For critical patients, changes in blood pressure can be crucial for the treatment.
-Thus, \ac{atc} assessment can be performed in addition to standard accuracy analysis~\citep[see, for example, ][]{Mostafa2020}.
-Thus, we assess the \ac{atc} of the non-invasive blood pressure measurements compared to the invasive blood pressure measurements.
-The database contains 64,168 numerical data records, comprising all numerical measurements for one patient.
-The measured signals vary in length, frequency, and type of measurement.
-Thus, only a subset of the data contains measurements of \ac{abp} and \ac{nbp} simultaneously.
-2,548 include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one measurement of systolic \ac{abp} and \ac{nbp} at the same time; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
+\subsection{Non-invasive blood pressure monitoring} \label{sec:application_measurement}
+
+We here consider the \ac{atc} of non-invasive blood pressure measurements from the MIMIC-III database that includes of critically ill patients treated in intensive care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}, \citealp{Moody2017}; available through \citealp{Goldberger2000}).
+We focus on \acf{abp} and \acf{nbp} measurements and thus limit our analysis to datasets containing simultaneous measurements of \ac{abp} and \ac{nbp} simultaneously.
+2,548 (BCS: what?
+datasets?
+patients?) include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one pair of simultaneously measured systolic \ac{abp} and \ac{nbp}; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
+We assess the \ac{atc} of non-invasive blood pressure measurements (test emthod) compared to intraarterial blood pressure measurements (referecne method, gold standard).
 We consider the horizons of one minute, five minutes, and 15 minutes for the \ac{atc} assessment, as those are typical intervals of NBP measurements.
@@
 In this paper, we examine various methods to assess the \acf{atc} for \acf{mnf}, that is, whether it predicts the correct direction of changes.
+BCS: Hier einfach wieder gleiches Wording wie in Einleitung etc. Sonst verwirren wir die Leser
 While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting, the predict the correct direction of changes assessment using the computed predicted and observed change is similar.
@@
     \centering
+    \tiny
     \begin{subtable}[t]{\textwidth}

### 2024-04-15 paper.tex
Work on BCS todos

@@
 %%% --- Commenting, Todonotes
-\usepackage[colorinlistoftodos,prependcaption]{todonotes}
-%\usepackage{comment, soul}
-%\newcommand{\ar}{$\rightarrow$}
-\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
+%\usepackage[colorinlistoftodos,prependcaption]{todonotes}
+%%\usepackage{comment, soul}
+%%\newcommand{\ar}{$\rightarrow$}
+%\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
 %\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}
@@
 \DeclareAcronym{atc}{short=ATC, long=ability to track changes}
-\DeclareAcronym{mnf}{short=MNF, long={measurement, nowcast, or forecast}}
 
@@
 \section{Introduction}\label{sec:atc-introduction}
-BCS: We need to avoid non-standard abbreviations that confuse readers such as MNF. Please delete the abbreviation MNF throughout the text.
 
 Measurements, nowcasts, or forecasts ideally should correctly reflect changes in the values of interest.
-It is thus important to meticulously assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the ability to track changes (\ac{atc}).
+It is thus important to meticulously assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the \acf{atc}.
 Although measurements, nowcasts, and forecasts fundamentally differ as they either measure or predict a value, similar methods can be used to assess their \ac{atc}.
@@
 
-Measurements aim to obtain accurate and precise values of a measureable quantity (measureand).
-BCS: Please cite https://pubmed.ncbi.nlm.nih.gov/32185615/ and https://pubmed.ncbi.nlm.nih.gov/32185616/
+Measurements aim to obtain accurate and precise values of a measureable quantity (measureand; \citealp{Squara2021,Squara2021a}).
 Repeated measurements can be used to track changes in a value over time.
-When introducing new measurement methods, they are evaluated against current reference methods, often called the "gold standard", by simultaneously measuring the same quantity with the new method and the reference method – often in various individuals or different clinical settings.
+When introducing new measurement methods, they are evaluated against current reference methods, often called the \enquote{gold standard}, by simultaneously measuring the same quantity with the new method and the reference method – often in various individuals or different clinical settings.
 
@@
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-When evaluating the measurment performance of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error [BCS: Please cite: https://pubmed.ncbi.nlm.nih.gov/12578081/] are commonly used.
+When evaluating the measurment performance of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
 How to best assess the \ac{atc} of measurement methods is a field of ongoing research~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
@@
 The assessment of \ac{atc} is based on the measured and predicted changes in a value of interest over a time horizon $l$.
-The \textit{measured} change is straightforward to compute for all types of \ac{mnf}.
+The \textit{measured} change is straightforward to compute for all types of measurement, nowcast, or forecast.
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the actual values for nowcasting or forecasting, or gold standard measurements up to time $T$.
@@
 The definition of the \textit{predicted} change depends on the context;
-Table~\ref{tab:notation} summarizes the notation for \ac{mnf} and the computation of the predicted change.
+Table~\ref{tab:notation} summarizes the notation for measurements, nowcasts, or forecasts and the computation of the predicted change.
 For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$.
@@
 In applications, data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
-We refer to time steps for which either \ac{mnf} or true values are unavailable as missing values.
+We refer to time steps for which either measurement, nowcast, or forecast or true values are unavailable as missing values.
 Systematical missing values could lead to a biased assessment, and missing data should be inspected for any underlying patterns.
@@
 
-Formally, the assessment of \ac{atc} is the same for \ac{mnf}, given the notation for the respective application of Section~\ref{subsec:notation}.
+Formally, the assessment of \ac{atc} is the same for measurement, nowcast, or forecast, given the notation for the respective application of Section~\ref{subsec:notation}.
 In the following, we omit the horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
@@
 Points within that area are omitted in the calculation of the measures.
-In particular, the \ac{mnf} are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
+In particular, the measurement, nowcast, or forecast are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
 The measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} without points in the exclusion area $E$ are
@@
 An exclusion area along one axis, for example, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
-%At least for \ac{mnf}, small amounts of noise are inevitable.
+%At least for measurement, nowcast, or forecast, small amounts of noise are inevitable.
 A cross-shaped exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
@@
 The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
-A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the \ac{atc} of a prediction for a specific change $\diffxrv = \xcond$ of the \ac{mnf}.
+A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the \ac{atc} of a prediction for a specific change $\diffxrv = \xcond$ of the measurement, nowcast, or forecast.
 Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct direction given a predicted change of $\xcond$.
@@
 We focus on \acf{abp} and \acf{nbp} measurements and thus limit our analysis to datasets containing simultaneous measurements of \ac{abp} and \ac{nbp} simultaneously.
-2,548 (BCS: what?
-datasets?
-patients?) include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one pair of simultaneously measured systolic \ac{abp} and \ac{nbp}; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
+2,548 datasets include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one pair of simultaneously measured systolic \ac{abp} and \ac{nbp}; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
 We assess the \ac{atc} of non-invasive blood pressure measurements (test emthod) compared to intraarterial blood pressure measurements (referecne method, gold standard).
@@
 
-In this paper, we examine various methods to assess the \acf{atc} for \acf{mnf}, that is, whether it predicts the correct direction of changes.
-BCS: Hier einfach wieder gleiches Wording wie in Einleitung etc. Sonst verwirren wir die Leser
+In this paper, we examine various methods to assess the \acf{atc} for measurements, nowcasts, or forecasts, that is, whether they correctly predict the direction of changes in values.
 While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting, the predict the correct direction of changes assessment using the computed predicted and observed change is similar.
@@
 
-Four-quadrant plots facilitate the visual inspection of the \ac{atc} for a \ac{mnf} (see Section~\ref{subsec:aatc-four-quadrant-plot}).
+Four-quadrant plots facilitate the visual inspection of the \ac{atc} for a measurement, nowcast, or forecast (see Section~\ref{subsec:aatc-four-quadrant-plot}).
 The \ac{atc} ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates \ac{atc}.
@@
 
-A conditional \ac{atc} plot visualizes the probability of correct change direction prediction over the predicted change of the \ac{mnf} (see Section~\ref{subsec:aatc-cond-prob}).
+A conditional \ac{atc} plot visualizes the probability of correct change direction prediction over the predicted change of the measurement, nowcast, or forecast (see Section~\ref{subsec:aatc-cond-prob}).
 It is based on a multivariate \acf{kde} of predicted and observed change.

### 2024-04-15 paper.tex
Fix typos

@@
 
-Measurements aim to obtain accurate and precise values of a measureable quantity (measureand; \citealp{Squara2021,Squara2021a}).
+Measurements aim to obtain accurate and precise values of a measurable quantity (measurand; \citealp{Squara2021,Squara2021a}).
 Repeated measurements can be used to track changes in a value over time.
@@
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-When evaluating the measurment performance of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
+When evaluating the measurement performance of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
 How to best assess the \ac{atc} of measurement methods is a field of ongoing research~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
@@
 
-Formally, the assessment of \ac{atc} is the same for measurement, nowcast, or forecast, given the notation for the respective application of Section~\ref{subsec:notation}.
+Formally, the assessment of \ac{atc} is the same for measurements, nowcasts, and forecasts, given the notation for the respective application of Section~\ref{subsec:notation}.
 In the following, we omit the horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
@@
 Points within that area are omitted in the calculation of the measures.
-In particular, the measurement, nowcast, or forecast are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
+In particular, the measurement, nowcast, or forecast is likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
 The measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} without points in the exclusion area $E$ are
@@
 First, the reporting process was delayed because – among other reasons – different authorities reported the data to the RKI~\citep{RobertKochInstitute2024}.
-Second, the seven-day hospitalization rate allocated all Covid-19-related hospitalizations to the date of the first positive test~\parencite[for a detailed description, see][]{Wolffram2023}.
+Second, the seven-day hospitalization rate allocated all COVID-19-related hospitalizations to the date of the first positive test~\parencite[for a detailed description, see][]{Wolffram2023}.
 Nevertheless, as the seven-day hospitalization rate was considered an important indicator of how the pandemic developed, various organizations and institutions issued nowcasts.
@@
     \input{plots/covid_nowcast/01_model_scores}
-    \caption[Point evaluation measures for the issued mean of the different models in Covid nowcasting.]{Point evaluation measures for the issued mean of the different models in Covid nowcasting.
+    \caption[Point evaluation measures for the issued mean of the different models in COVID-19 nowcasting.]{Point evaluation measures for the issued mean of the different models in COVID-19 nowcasting.
 \enquote{RMSE} and \enquote{MAE} are accuracy measures, while \enquote{Count} lists the number of non-missing values.
@@
     \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
-    \caption[The ATC ratios for Covid nowcasting.]{The \ac{atc} ratio $\accl[7]$, positive \ac{atc} ratio $\accpl[7]$, and negative \ac{atc} ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days in Covid nowcasting.
+    \caption[The ATC ratios for COVID-19 nowcasting.]{The \ac{atc} ratio $\accl[7]$, positive \ac{atc} ratio $\accpl[7]$, and negative \ac{atc} ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days in COVID-19 nowcasting.
 The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values.
@@
     \end{subfigure}
-    \caption[Conditional ATC plot and ATC ratio over exclusion area for the Covid nowcasts.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days in Covid nowcasting.}
+    \caption[Conditional ATC plot and ATC ratio over exclusion area for the COVID-19 nowcasts.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days in COVID-19 nowcasting.}
     \label{fig:app-covid-cond-prob-atc-ratio-7}
@@
     \end{subfigure}\hspace{0.01\textwidth}%
-    \caption[Brier Scores and reliability diagrams for the Covid nowcasting models ILM, RKI, RIVM, and ENS-MED.]{\aclp{bs} and reliability diagrams for the Covid nowcasting models ILM, RKI, RIVM, and ENS-MED.
+    \caption[Brier Scores and reliability diagrams for the COVID-19 nowcasting models ILM, RKI, RIVM, and ENS-MED.]{\aclp{bs} and reliability diagrams for the COVID-19 nowcasting models ILM, RKI, RIVM, and ENS-MED.
     The reliability diagram bins are chosen according to the empirical quantiles of the predicted probabilities.
@@
 
-We here consider the \ac{atc} of non-invasive blood pressure measurements from the MIMIC-III database that includes of critically ill patients treated in intensive care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}, \citealp{Moody2017}; available through \citealp{Goldberger2000}).
+We here consider the \ac{atc} of non-invasive blood pressure measurements from the MIMIC-III database that includes data of critically ill patients treated in intensive care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}, \citealp{Moody2017}; available through \citealp{Goldberger2000}).
 We focus on \acf{abp} and \acf{nbp} measurements and thus limit our analysis to datasets containing simultaneous measurements of \ac{abp} and \ac{nbp} simultaneously.
-2,548 datasets include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one pair of simultaneously measured systolic \ac{abp} and \ac{nbp}; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
-We assess the \ac{atc} of non-invasive blood pressure measurements (test emthod) compared to intraarterial blood pressure measurements (referecne method, gold standard).
+2,548 datasets include at least one measurement of systolic \ac{abp} and \ac{nbp}, and 1,327 include at least one pair of simultaneously measured systolic \ac{abp} and \ac{nbp}; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
+We assess the \ac{atc} of non-invasive blood pressure measurements (test method) compared to intraarterial blood pressure measurements (reference method, gold standard).
 We consider the horizons of one minute, five minutes, and 15 minutes for the \ac{atc} assessment, as those are typical intervals of NBP measurements.
@@
 In this paper, we examine various methods to assess the \acf{atc} for measurements, nowcasts, or forecasts, that is, whether they correctly predict the direction of changes in values.
-While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting, the predict the correct direction of changes assessment using the computed predicted and observed change is similar.
+While the computation of predicted change varies between the application areas of measurement, nowcasting, and forecasting, the assessment can be based on the same methods.
 The \ac{atc} assessment can accompany other evaluation techniques, such as measures of deviation or probabilistic scoring rules.
@@
     \end{tabular}
-    \caption[Matching the abbreviation to the key in the Covid nowcasting hub.]{Matching the abbreviation to the key in the Covid nowcasting hub.
+    \caption[Matching the abbreviation to the key in the COVID-19 nowcasting hub.]{Matching the abbreviation to the key in the COVID-19 nowcasting hub.
     Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
@@
     \input{plots/covid_nowcast/10_marginal_analysis}
-    \caption[Marginal analysis of the Covid nowcast and true changes for the horizons one, seven, and 14 days.]{Marginal analysis of the Covid nowcast and true changes for the horizons one, seven, and 14 days.
+    \caption[Marginal analysis of the COVID-19 nowcast and true changes for the horizons one, seven, and 14 days.]{Marginal analysis of the COVID-19 nowcast and true changes for the horizons one, seven, and 14 days.
     The column (1), $l=l$ shows the number of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the changes' absolute values.}
@@
     \end{subtable}
-    \caption[ATC ratios for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.]{\Ac{atc} ratio $\acc$, positive \ac{atc} ratio $\accp$, and negative \ac{atc} ratio $\accm$ for the models without and with exclusion areas for the horizon one and 14 days in Covid nowcasting.
+    \caption[ATC ratios for the models without and with exclusion areas for the horizon one and 14 days in COVID-19 nowcasting.]{\Ac{atc} ratio $\acc$, positive \ac{atc} ratio $\accp$, and negative \ac{atc} ratio $\accm$ for the models without and with exclusion areas for the horizon one and 14 days in COVID-19 nowcasting.
 The exclusion areas are rectangles centered on the zero points with a width and height to exclude the 10\%-quantile of the absolute values of nowcast or true values.
@@
 \includegraphics{plots/covid_nowcast/30_4q_plots}
-\caption[Four-quadrant plots for the Covid nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.]{Four-quadrant plots for the Covid nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.
+\caption[Four-quadrant plots for the COVID-19 nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.]{Four-quadrant plots for the COVID-19 nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.
 The spread in both directions increases with the horizon.
@@
     \end{subfigure}
-    \caption[Conditional ATC plot and ATC ratio over exclusion area in Covid nowcasting.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the Covid nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
+    \caption[Conditional ATC plot and ATC ratio over exclusion area in COVID-19 nowcasting.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the COVID-19 nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
     \label{fig:app-covid-cond-prob-atc-ratio-1-14}
@@
     \end{subfigure}
-    \caption[Reliability diagrams for Covid nowcasting models and the horizon seven and 14 days.]{The reliability diagram for the Covid nowcasting models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
+    \caption[Reliability diagrams for COVID-19 nowcasting models and the horizon seven and 14 days.]{The reliability diagram for the COVID-19 nowcasting models ILM, RIVM, RKI, and ENS-MED for the horizon seven and 14 days.
     Additionally, the count of predicted probabilities for the horizons is shown.

### 2024-04-15 paper.tex
Update on Overleaf.

@@
 
-In this article, we focus on how to assess the \ac{atc} of measurements, nowcasts, or forecasts to track changes.
-We formalize the concept of \ac{atc} and present visual techniques and quantitative measures to assess the \ac{atc} – considering both noiseless data and data with noise and small non-informative changes.
-We introduce the conditional \ac{atc} plot, a new graphical method for assessing local \ac{atc} behavior, and review bootstrap methods for calculating confidence intervals.
-We extend the concept of assessment of \ac{atc} to probabilistic predictions.
+In this article, we focus on how to assess the ability of measurements, nowcasts, or forecasts to track changes.
+We formalize the concept of \ac{atc} and present visual techniques and quantitative measures to assess it – considering both noiseless data and data with noise and small non-informative changes.
+We introduce the conditional \ac{atc} plot, a new graphical method for assessing the local ability, and review bootstrap methods for calculating confidence intervals.
+We extend the concept of assessment to probabilistic predictions.
 We exemplarily illustrate the proposed methods to assess the \ac{atc} for nowcasting during the COVID-19 pandemic, patient admissions to an emergency department, and non-invasive blood pressure measurements – and thus provide blueprints for future assessments.
-The ready-to-use code for the assessment of \ac{atc} from the applications is available on [LINK WILL BE INSERTED].
+Ready-to-use code is available on [LINK WILL BE INSERTED].
 
@@
 
-The assessment of \ac{atc} is based on the measured and predicted changes in a value of interest over a time horizon $l$.
+The assessment of \ac{atc} is based on the measured, observed, or true and the predicted changes in a value of interest over a time horizon $l$.
 The \textit{measured} change is straightforward to compute for all types of measurement, nowcast, or forecast.
@@
 However, the publication of the definite hospitalization rate was substantially delayed and partially flawed for two main reasons.
-First, the reporting process was delayed because – among other reasons – different authorities reported the data to the RKI~\citep{RobertKochInstitute2024}.
+First, the reporting process was delayed because – among other reasons – different authorities were involved in passing the data to the RKI~\citep{RobertKochInstitute2024}.
 Second, the seven-day hospitalization rate allocated all COVID-19-related hospitalizations to the date of the first positive test~\parencite[for a detailed description, see][]{Wolffram2023}.

### 2024-04-16 paper.tex
Shorten intro C19-Nowcasting

@@
 
-The assessment of \ac{atc} is based on the measured, observed, or true and the predicted changes in a value of interest over a time horizon $l$.
+The assessment of \ac{atc} is based on the measured/observed/true and the predicted changes in a value of interest over a time horizon $l$.
 The \textit{measured} change is straightforward to compute for all types of measurement, nowcast, or forecast.
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the actual values for nowcasting or forecasting, or gold standard measurements up to time $T$.
-The sequence of observed changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is,
+The sequence of changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is,
 \begin{equation}\label{eq:diffy}
@@
 Second, the seven-day hospitalization rate allocated all COVID-19-related hospitalizations to the date of the first positive test~\parencite[for a detailed description, see][]{Wolffram2023}.
-Nevertheless, as the seven-day hospitalization rate was considered an important indicator of how the pandemic developed, various organizations and institutions issued nowcasts.
-The COVID-19-Nowcasting-Hub collected various nowcasts in a predefined setup, including the mean, median, and other quantiles of the predicted seven-day hospitalization rate~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
-For information on the model design, nowcast tasks, and the data submission guidelines, we refer to~\citet{Wolffram2023}.
-Table~\ref{tab:app-covid-models} matches the abbreviations here to the respective names in the COVID-19-Nowcasting-Hub.
+The COVID-19-Nowcasting-Hub~\citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024} collected various nowcasts in a predefined setup, including the mean, median and other quantiles of the predicted seven-day hospitalization rate~\parentext{for further information see~\citealp{Wolffram2023} and Table~\ref{tab:app-covid-models} for the abbreviations used}.
 In addition to those nowcasts, \citet{Wolffram2023} construct two ensemble methods using the ensembles' mean or median.

### 2024-04-21 paper.tex
Import changes from diss

@@
 
-In forecasting and nowcasting, the evaluation of the measurement performance is based on statistical methods quantifiying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
+In forecasting and nowcasting, the evaluation of the measurement performance is based on statistical methods quantifying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
 None of them assesses the method's \ac{atc}.
@@
 Points in the green upper right and lower left quadrants reflect a correct change direction for the respective time step, whereas points in the remaining red quadrants show incorrectly predicted changes.
-Figure~\ref{fig:aatc_basic_4q} displays a basic four-quadrant plot and Figure~\ref{fig:aatc_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data.
+Figure~\ref{fig:aatc_basic_4q} displays a basic four-quadrant plot, and Figure~\ref{fig:aatc_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data (for the data generation, see Appendix~\ref{subsec:app-ATC-data-generation}).
 
@@
 \end{subfigure}%
-\caption{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas.
-The rectangular exclusion area in Figure~\ref{fig:aatc_basic_4q_excl_box} excludes only points where both components are likely to be noise-driven, while the exclusion areas in Figures~\ref{fig:aatc_basic_4q_excl_axis} and~\ref{fig:aatc_basic_4q_excl_cross} excludes points where at least one component is noise-driven.
+\caption[Illustrations of the four-quadrant plot with sample points and with and without exclusion areas.]{Illustrations of the four-quadrant plot with sample points and with and without exclusion areas.
+The rectangular exclusion area in Figure~\ref{fig:aatc_basic_4q_excl_box} excludes only points where both components are likely to be noise-driven, while the exclusion areas in Figures~\ref{fig:aatc_basic_4q_excl_axis} and~\ref{fig:aatc_basic_4q_excl_cross} exclude points where at least one component is noise-driven.
 }
@@
 The RMSE orders the models.
-The evaluation period comprises 159 days and only few nowcasts are missing~\citep[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
+The evaluation period comprises 159 days, and only a few nowcasts are missing~\citep[for explanations of the missing values, see][Tables A2, A3, and A4]{Wolffram2023}.
 Note that the high values for the EPI model could be driven by an exceptionally far-off value at the end of the evaluation period (see Figure~\ref{fig:app-covid-true-nowcast}).}
@@
     In the computation of BS and reliability diagram, missing values are excluded.
-    The reliability diagram for the horizon seven and 14 days is in the appendix (see Figure~\ref{fig:app-covid-rel}).}
+    The reliability diagram for the horizons seven and 14 days is in the appendix (see Figure~\ref{fig:app-covid-rel}).}
     \label{fig:app-covid-prob}
@@
 The confidence interval width is at most 0.02 for the \ac{atc} ratios and at most 0.03 for the positive and negative \ac{atc} ratios.
-The models GBM-2, qreg-1, and Benchmark-1 have the highest positive \ac{atc} ratio for the three and seven-day horizon, while Poisson-2 and NBI-2 have the highest negative \ac{atc} ratio.
+The models GBM-2, qreg-1, and Benchmark-1 have the highest positive \ac{atc} ratio for the three and seven-day horizons, while Poisson-2 and NBI-2 have the highest negative \ac{atc} ratio.
 
@@
     \input{plots/ed_arrival/50_trending_ratio}
-    \caption[ATC ratios for the emergency department arrival forecasts.]{\Ac{atc} ratio $\acc$, positive \ac{atc} ratio $\accp$, and negative \ac{atc} ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days in emergency department arrival forecasting.}
+    \caption[ATC ratios for the forecasts of patient admission to an emergency department.]{\Ac{atc} ratio $\acc$, positive \ac{atc} ratio $\accp$, and negative \ac{atc} ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days in the forecasting of patient admissions to an emergency department.}
     \label{tab:app-eda-atc-ratios}
@@
     \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_3}
-    \caption{Horizon three days}
+    \caption{Horizon three days.}
     \end{subfigure}\hfill
@@
     \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_7}
-    \caption{Horizon seven days}
+    \caption{Horizon seven days.}
     \end{subfigure}
-    \caption[Conditional ATC plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} in emergency department arrival forecasting.]{Conditional \ac{atc} plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} in emergency department arrival forecasting.
+    \caption[Conditional ATC plots for the horizons three and seven days and the models with the best positive or negative \acs{atc} in forecasting the patient admissions to an emergency department.]{Conditional \ac{atc} plots for the horizons three and seven days and the models with the best positive or negative \ac{atc} in forecasting the patient admissions to an emergency department.
 The plots of NBI-2 and Poisson-2 are indistinguishable.}
@@
     \end{subfigure}
-    \caption[Probabilistic \ac{atc} assessment in emergency department arrival forecasting.]{Probabilistic \ac{atc} assessment for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days in emergency department arrival forecasting.
+    \caption[Probabilistic \acs{atc} assessment in forecasting patient admissions to an emergency department.]{Probabilistic \ac{atc} assessment for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days in forecasting patient admissions to an emergency department.
 The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic \ac{atc} simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically.
@@
 
-The methods of \ac{atc} assessment are applied to COVID-19-nowcasting, emergency department arrival forecasting, and invasive and non-invasive blood pressure measurements in Section~\ref{sec:application}.
+The methods of \ac{atc} assessment are applied to COVID-19-nowcasting, forecasting the patient admissions to an emergency department, and invasive and non-invasive blood pressure measurements in Section~\ref{sec:application}.
 While \ac{atc} assessment should not be the only aspect, it is a valuable addition to evaluating nowcasts, forecasts, and measurements.
@@
         \includegraphics{plots/illustrative_examples/appendix_4q_dgp2}
-        \caption{Second approach}
+        \caption{Second approach.}
     \end{subfigure}
@@
         \input{plots/illustrative_examples/bootstrap_butterfly_cis}
-        \caption{First dataset}
+        \caption{First dataset.}
     \end{subtable}\hspace{0.02\textwidth}
@@
        \input{plots/illustrative_examples/bootstrap_normal_cis}
-        \caption{Second dataset}
+        \caption{Second dataset.}
     \end{subtable}
-    \caption[Proportion of bootstrapping confidence intervals covering the true value of \ac{atc} ratio per method and sample size $T$ in the bootstrap simulation study.]{Proportion of bootstrapping confidence intervals covering the true value of \ac{atc} ratio per method and sample size $T$.
+    \caption[Proportion of bootstrap confidence intervals covering the true value of \ac{atc} ratio per method and sample size $T$ in the bootstrap simulation study.]{Proportion of bootstrap confidence intervals covering the true value of \ac{atc} ratio per method and sample size $T$.
 The average width of the confidence interval is listed in brackets.}
@@
 \includegraphics{plots/illustrative_examples/boxplot_comp_time_butterfly}
-        \caption{First dataset}
+        \caption{First dataset with asymmetry.}
     \end{subfigure}
@@
     \includegraphics{plots/illustrative_examples/boxplot_comp_time_normal}
-        \caption{Second dataset}
+        \caption{Second dataset.}
     \end{subfigure}
@@
     \caption[Matching the abbreviation to the key in the COVID-19 nowcasting hub.]{Matching the abbreviation to the key in the COVID-19 nowcasting hub.
-    Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
+Information on the models and references are listed in \citet[][Table 1]{Wolffram2023}.}
     \label{tab:app-covid-models}
@@
         \input{plots/covid_nowcast/30_trending_ratios_lag_1}
-    \caption{1 day.}
+    \caption{One day.}
     \end{subtable}
@@
     \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
-    \caption{Conditional \ac{atc} plot for horizon 1.}\label{fig:app-covid-cond-prob-1}
+    \caption{Conditional \ac{atc} plot for horizon one.}\label{fig:app-covid-cond-prob-1}
     \end{subfigure}\hfill
@@
     \includegraphics{plots/covid_nowcast/40_cond_prob_lag_14}
-    \caption{Conditional \ac{atc} plot for horizon 1.}\label{fig:app-covid-cond-prob-14}
+    \caption{Conditional \ac{atc} plot for horizon 14.}\label{fig:app-covid-cond-prob-14}
     \end{subfigure}
@@
     \includegraphics{plots/covid_nowcast/40_acc_eps_lag_1}
-    \caption{\Ac{atc} ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-atc-ratio-1}
+    \caption{\Ac{atc} ratio over exclusion area size in $\diffx$ for horizon one.}\label{fig:app-covid-atc-ratio-1}
     \end{subfigure}\hfill
@@
     \includegraphics{plots/covid_nowcast/40_acc_eps_lag_14}
-    \caption{\ac{atc} ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-atc-ratio-14}
+    \caption{\ac{atc} ratio over exclusion area size in $\diffx$ for horizon 14.}\label{fig:app-covid-atc-ratio-14}
     \end{subfigure}
-    \caption[Conditional ATC plot and ATC ratio over exclusion area in COVID-19 nowcasting.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the COVID-19 nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
+    \caption[Conditional ATC plot and ATC ratio over exclusion area in COVID-19 nowcasting.]{Conditional \ac{atc} plot and \ac{atc} ratio over exclusion area for the COVID-19 nowcast models ILM, RKI, RIVM, and ENS-MED for the horizon seven days.
+Note the different axis scalings of the ATC ratios for the two horizons.
+While the ENS-MED model has a rather low \ac{atc} ratio for small exclusion areas, it is increasing fast for larger exclusion areas.
+The RKI model is among the worst, except for the small negative predicted changes.
+}
     \label{fig:app-covid-cond-prob-atc-ratio-1-14}

### 2024-04-21 paper.tex
Change affiliations / contributions

@@
 \affil[1]{\orgdiv{Institute for Operations Research}, \orgname{Karlsruhe Institute of Technology (KIT)}, \orgaddress{\street{Kaiserstr. 12}, \city{Karlsruhe}, \postcode{76131}, \country{Germany}}}
+\affil[2]{\orgdiv{Department of Anesthesiology, Center of Anesthesiology and Intensive Care Medicine}, \orgname{University Medical Center Hamburg-Eppendorf}, \orgaddress{\street{Martinistrasse 52}, \city{Hamburg}, \postcode{20246}, \country{Germany}}}
 %TC:ignore
@@
 
-OG, BL, and JR developed the theoretical formalism.
-BL and JR implemented the methods.
+JR, BL, and OG developed the theoretical formalism.
+JR and BL implemented the methods.
 JR conducted the simulations and analyzed the data.
-BL and JR wrote the manuscript.
-All authors discussed the results, critically reviewed them, and contributed to the final manuscript.
+All authors interpreted the data.
+BS and OG substantively revised the manuscript.
+JR and BL drafted the manuscript.
+All authors have approved the submitted version and have agreed both to be personally accountable for the author's own contributions and to ensure that questions related to the accuracy or integrity of any part of the work, even ones in which the author was not personally involved, are appropriately investigated, resolved, and the resolution documented in the literature.
 

### 2024-05-03 paper.tex
Update on Overleaf.

@@
 
-\title{The assessment of the ability of measurements, nowcasts, and forecasts to track changes}
+\title{On the assessment of the ability of measurements, nowcasts, and forecasts to track changes}
 %\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
@@
 
-In forecasting and nowcasting, the evaluation of the measurement performance is based on statistical methods quantifying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
+In forecasting and nowcasting the evaluation of performance or goodness is usually based on statistical methods quantifying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
 None of them assesses the method's \ac{atc}.
-
-In a prediction competition on armed conflicts, the assessment of the \ac{atc} recently gained attention as \citet{Vesco2022} proposed the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
+In a forecasting competition on armed conflicts, the assessment of the \ac{atc} recently gained attention as \citet{Vesco2022} proposed the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-When evaluating the measurement performance of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
+When evaluating the performance or goodness of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
 How to best assess the \ac{atc} of measurement methods is a field of ongoing research~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
@@
 The assessment of \ac{atc} is based on the measured/observed/true and the predicted changes in a value of interest over a time horizon $l$.
-The \textit{measured} change is straightforward to compute for all types of measurement, nowcast, or forecast.
+The \textit{true} change is straightforward to compute for all types of measurement, nowcast, or forecast.
 Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the actual values for nowcasting or forecasting, or gold standard measurements up to time $T$.
@@
 
-\subsection{ATC ratio and other measures}\label{subsec:atc-measures}
+\subsection{The ATC ratio and other measures}\label{subsec:atc-measures}
 
@@
 The green curve $k_t$ shows the theoretical probability that $\diffxt$ has the same sign as $\diffyt$ for each time step.
-The rolling estimates are shifted to the right compared to $k_t$ as the windows look backward.
-Thus, the yearly course of the ATC ratio can be detected.
 The ATC ratio has a pronounced sinus-shaped seasonality with a peak after a quarter of a year and a low point after three quarters.
+The rolling estimates detect the yearly course of the ATC ration.
+Naturally, they are shifted to the right compared to $k_t$ as the windows look backward.
+
 }
@@
 Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct direction given a predicted change of $\xcond$.
-Thus, if a change of $\xcond$ is observed in practice, one can directly assess its credibility.
-A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
+Thus, if a change of $\xcond$ is observed in practice, one can directly assess its credibility regarding the direction.
+Multivariate \acf{kde} facilitates the continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
 \begin{align*}
@@
 We start by analyzing the marginal distributions for the predicted and observed changes for the three- and seven-day horizons in Table~\ref{tab:app-eda-marginals}, again.
+% XXX should we delegate that table to the appendix as before?
 The computed difference aligns with Section~\ref{subsec:notation}, that is, the difference between the forecasted mean and true value of three and seven days before, as the actual value is available when issuing the forecast.
@@
 \end{table}
+
+% XXX in the figure above: comment on the exclusion area; notation in columns is not consistent to the example anbove (with respect to using an exclusion area a or not), I think; i.e., \mu_q vs \mu
 
@@
 \end{table}
+% XXX hier wieder die Sache mit der Notation und Exclusion Zone
 
@@
 \subsubsection*{Discussion}
+
+% XXX should we exclude the 1 min horizon, i.e., does it make sense to look at such a small horizon?
+(especially also because of the large CIs?
+Though that might be interesting for didactic reasons)?
+The NBP will take some time, so I am not sure how the exact time of the measurement is chosen
 

### 2024-07-25 paper.tex
Update on Overleaf.

@@
 We start by analyzing the marginal distributions for the predicted and observed changes for the three- and seven-day horizons in Table~\ref{tab:app-eda-marginals}, again.
-% XXX should we delegate that table to the appendix as before?
 The computed difference aligns with Section~\ref{subsec:notation}, that is, the difference between the forecasted mean and true value of three and seven days before, as the actual value is available when issuing the forecast.
@@
     \input{plots/ed_arrival/50_trending_ratio}
-    \caption[ATC ratios for the forecasts of patient admission to an emergency department.]{\Ac{atc} ratio $\acc$, positive \ac{atc} ratio $\accp$, and negative \ac{atc} ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days in the forecasting of patient admissions to an emergency department.}
+    \caption[ATC ratios for the forecasts of patient admission to an emergency department.]{\Ac{atc} ratio $\acc_+$, positive \ac{atc} ratio $\accp_+$, and negative \ac{atc} ratio $\accm_+$ for the models for the horizons 72 hours and seven days in the forecasting of patient admissions to an emergency department with an exclusion zone equal to the axis indicated by the subscript $+$.}
     \label{tab:app-eda-atc-ratios}
 \end{table}
-
-% XXX in the figure above: comment on the exclusion area; notation in columns is not consistent to the example anbove (with respect to using an exclusion area a or not), I think; i.e., \mu_q vs \mu
 
@@
 \end{table}
-% XXX hier wieder die Sache mit der Notation und Exclusion Zone
 
@@
 \subsubsection*{Discussion}
-
-% XXX should we exclude the 1 min horizon, i.e., does it make sense to look at such a small horizon?
-(especially also because of the large CIs?
-Though that might be interesting for didactic reasons)?
-The NBP will take some time, so I am not sure how the exact time of the measurement is chosen
 

### 2024-07-28 CoverLetter/letter-to-the-editor.tex
Inital commit cover letter

\documentclass[%
 KITletter, %the letter class option
 %bwKITlogo, %--- activate to have the KIT logo black/white instead of coloured
 %officeprintableoutline, %--- activate for a layout printable on office printers
 %nooutlineframe, %--- activate tp print the letter on the pre-printed form of KIT
 %class options that do not change the layout can be added here, e.g. twoside,
fromemail,fromurl, %--- scrlttr2 options; activate as needed
 USenglish,%UKenglish%,french%,ngerman (already default) ... %--- activate one (or more) language options as needed
 %helvet, %heros, %--- use one of these to have a fallback font as default
 %cursor %--- activate to have the fallback typewriter font as default
  backaddress=off,
foldmarks=off,
 ]{scrlttr2}
\usepackage{hyperref}%for detailed information see: texdoc scrguide

%--- the back address (placed above the adressee)
%\setkomavar{backaddress}{KIT \textbar{} Dep. of Economics and Management \textbar{} Kaiserstr. 89 \textbar{} Karlsruhe, Germany}
\setkomavar{backaddress}{}

%--- the possible fields for the sender information (right of the addressee):
     %-- remove the fields not needed or set them to "{}",
     %-- use the optional arguments to adapt the respective descriptions or even to divert fields, e.g.
     %   \setkomavar{myref}[Position]{Project leader}
\setkomavar{fromname}{Jonas Rieger}
%\setkomavar{headofOU}{Prof. Vorname Surename}
\setkomavar{fromaddress}{Karlsruhe Institute of Technology\\Institute of Operations Research\\Chair Prof. Dr. O. Grothe – Analytics and Statistics\\76131 Karlsruhe, Germany}
%\setkomavar{fromphone}{0721-608-xxxxx}          %(to be activated by class option)
%\setkomavar{fromfax}{0721-608-xxxxx}            %          "
\setkomavar{fromemail}[Email:]{jonas.rieger@kit.edu}%          "
\setkomavar{fromurl}[Web:]{https://as.ior.kit.edu/}      %          "
%\setkomavar{contact}{Vorname Surename}
%\setkomavar{myref}{xxxx}
%\setkomavar{fromlogo}{\includegraphics[width=17mm]{info}}%(to be activated by class option)
%(the info graphic comes from the standard package "notes"
% and must in case of need b replaced by your individual logo)

%--- set-up for the main part and the closing of the letter:
%\setkomavar{title}{Testtitel} %--- rarely needed, placed before the subject line
\setkomavar{subject}{Submission to BMC Medical Research Methodology} %--- placed before the opening phrase
%\setkomavar{signature}{\includegraphics[width=2.5cm]{unterschrift}\\Jonas Rieger} %--- used by \closing{...}; overwrites the fromname variable
\setkomavar{signature}{[signature to be inserted]\\Jonas Rieger} %--- used by \closing{...}; overwrites the fromname variable
                                             %(you can provide a hand-written signature with \includegraphics)
\setplength{sigbeforevskip}{\baselineskip} %--- skip between closing phrase and signature

\setkomavar{firstfoot}{}%activate to suppress the footer of the letterhead page

\begin{document}

%\justifying %--- activate to have justified instead of ragged-right text

\def\pagename{} %--- activate to suppress "Page", "Seite" etc.  before the page number on pages 2 onwards

\begin{letter}{%
To the Editor of \\ BMC Medical Research Methodology
%\\LAND
}

\opening{Dear Dr. Lo Monaco,}

We hereby submit our manuscript

\begin{quote}
\textit{The assessment of the ability of measurements, nowcasts, and forecasts to track changes}
\end{quote}

by Bolin Liu, Bernd Saugel, Oliver Grothe, and myself for your kind consideration for publication in BMC Medical Research Methodology.

The manuscript introduces novel statistical methods and visualization techniques for assessing the ability of measurements, nowcasts, and forecasts to predict the correct direction of change.
These methods are exemplified using three distinct medical and healthcare datasets.
Hence, the manuscript aligns directly with the scope of BMC Medical Research Methodology.

We declare that there are no potential competing interests related to this research or manuscript.
All authors have reviewed and approved this manuscript for submission.
We confirm that the content of this manuscript has not been published, nor has it been submitted for publication in another journal.
This work is original and has not been considered by any other journal.
Please note that the material is also part of my doctoral thesis.

We respectfully suggest Prof. Dr. Alexander Hapfelmeier (\url{alexander.hapfelmeier@mri.tum.de}) from the Institute of AI and Informatics in Medicine at the Technical University of Munich, and Dr. Simon Vistisen (\url{vistisen@clin.au.dk}), Associate Professor and Business Engagement Partner in the Department of Clinical Medicine at Aarhus University and the Department of Anaesthesia \& Intensive Care at Aarhus University Hospital, as potential reviewers for our manuscript.

We are confident this paper represents a valuable and novel contribution to the evaluation of forecasting, nowcasting, and measurement methods in healthcare research.

\closing{\iflanguage{ngerman}{Mit freundlichen Grüßen}{Yours sincerely}}
%Jonas Rieger
%another idea for the closing:
%\setkomavar{signature}{Vorname Nachname\\Position}
%\setplength{sigbeforevskip}{0mm}
%\closing{\rule{12em}{0.5bp}}

\end{letter}

\end{document}

### 2024-07-29 CoverLetter/letter-to-the-editor.tex
Add medical to summary

@@
 \setkomavar{subject}{Submission to BMC Medical Research Methodology} %--- placed before the opening phrase
-%\setkomavar{signature}{\includegraphics[width=2.5cm]{unterschrift}\\Jonas Rieger} %--- used by \closing{...}; overwrites the fromname variable
-\setkomavar{signature}{[signature to be inserted]\\Jonas Rieger} %--- used by \closing{...}; overwrites the fromname variable
+\setkomavar{signature}{\includegraphics[width=2.5cm]{unterschrift}\\Jonas Rieger} %--- used by \closing{...}; overwrites the fromname variable
+%\setkomavar{signature}{[signature to be inserted]\\Jonas Rieger} %--- used by \closing{...}; overwrites the fromname variable
                                              %(you can provide a hand-written signature with \includegraphics)
@@
 
-The manuscript introduces novel statistical methods and visualization techniques for assessing the ability of measurements, nowcasts, and forecasts to predict the correct direction of change.
+The manuscript introduces novel statistical methods and visualization techniques for assessing the ability of medical measurements, nowcasts, and forecasts to predict the correct direction of change.
 These methods are exemplified using three distinct medical and healthcare datasets.

### 2024-08-01 paper.tex
Remove comments
Submitted version
Adjust paper formatting according to journal instructions

@@
-%\documentclass[referee,sn-mathphys-ay,pdflatex]{sn-jnl}
-%\documentclass[sn-mathphys-ay,pdflatex]{sn-jnl}
 \documentclass[pdflatex]{sn-jnl}
-% Language setting
-% Replace `englis' with e.g. `spanish' to change the document language
 \usepackage[english]{babel}
@@
 \usepackage{fontenc}
-
-% Set page size and margins
-% Replace `letterpaper' with`a4paper' for UK/EU standard size
-%\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
-
-%%%% ---- Useful packages
 
@@
 \usepackage{tabularx}
-%\usepackage{tabularray}
 \usepackage{floatrow}
@@
 \usepackage[short]{optidef} % for aligning linear programs
-%\usepackage{nccmath} % for fleqn environment
 
 %%% --- Other
-%\usepackage[colorlinks=true, allcolors=blue]{hyperref}
-%\usepackage[hidelinks]{hyperref}
-\usepackage[style=authoryear,backend=biber,natbib=true]{biblatex}
-%\usepackage{natbib}
-%\bibliographystyle{
-%\addbibresource{library.bib}
-\addbibresource{library-jonas.bib}
+\usepackage[style=numeric,backend=biber,natbib=true]{biblatex}
+\addbibresource{paper_biber.bib}
 \usepackage{csquotes}
 \usepackage[shortcuts]{extdash} % For Hyphens in English
-%\usepackage{acro}
-% \DeclareAcronym{pdf}{short=PDF, long={probability distribution function}}
-
-%%% --- Commenting, Todonotes
-%\usepackage[colorinlistoftodos,prependcaption]{todonotes}
-%%\usepackage{comment, soul}
-%%\newcommand{\ar}{$\rightarrow$}
-%\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
-%\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}
-
-%\usepackage{xargs,ifthen} % Use optional arguements in commands
-%\newcommandx{\unsure}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
-%\newcommandx{\discuss}[2][1=]{\todo[linecolor=orange,backgroundcolor=orange!25,bordercolor=orange,#1]{#2}}
-%\newcommandx{\missing}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-%\newcommandx{\change}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
-%\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
-%\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}
-%\newcommandx{\done}[2][1=]{\todo[linecolor=gray,backgroundcolor=gray!25,bordercolor=gray,#1]{#2}}
-%\newcommandx{\thiswillnotshow}[2][1=]{\todo[disable,#1]{#2}}
-%\usepackage{marginnote}
-%\let\marginpar\marginnote
 
@@
 \newcommand{\accm}{\acc^-}
-%\newcommand{\acceps}[1][]{\ifthenelse { \equal {#1} {} }{\acc_{\eps}}{\acc_{#1}}} %https://stackoverflow.com/a/7314951
 \newcommand{\acceps}[1][e]{\acc_{#1}} %https://stackoverflow.com/a/7314951
@@
 
-% Ungefähre Wortanzahl des Hauptteils (von Introduction bis Conclusion): oben links auf 'Menu' und dann 'Word Count'
-% Ungefähre Aufteilung auf Abschnitte (nicht direkt in Overleaf sichtbar):
-% Introduction: 500
-% Assessment of ATC: 2225
-% Application: 3045
-% Discussion and Conclusion: 587
-
-% Abstract: 99
-
-%%% --- ToDo List
-%\usepackage{enumitem}
-%\newlist{todolist}{itemize}{2}
-%\setlist[todolist]{label=$\square$}
-%\usepackage{pifont}
-%\newcommand{\cmark}{\ding{51}}%
-%\newcommand{\xmark}{\ding{55}}%
-%\newcommand{\done}{\rlap{$\square$}{\raisebox{2pt}{\large\hspace{1pt}\cmark}}%
-%\hspace{-2.5pt}}
-%\newcommand{\wontfix}{\rlap{$\square$}{\large\hspace{1pt}\xmark}}
-
 \begin{document}
@@
 \title{On the assessment of the ability of measurements, nowcasts, and forecasts to track changes}
-%\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
 \author*[1]{Jonas Rieger}\email{jonas.rieger@kit.edu}
@@
 \affil[2]{\orgdiv{Department of Anesthesiology, Center of Anesthesiology and Intensive Care Medicine}, \orgname{University Medical Center Hamburg-Eppendorf}, \orgaddress{\street{Martinistrasse 52}, \city{Hamburg}, \postcode{20246}, \country{Germany}}}
-%TC:ignore
 \abstract{
@@
 }
-%TC:endignore
 \keywords{}
@@
 \maketitle
-% \listoftodos
 
@@
 A rectangular exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components.
-%Points are included if the true value or prediction is unlikely to be noise-driven.
-%In the example graph in Figure~\ref{fig:aatc_basic_4q_excl_box}, only point 1 is excluded.
 An exclusion area along one axis, for example, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
-%At least for measurement, nowcast, or forecast, small amounts of noise are inevitable.
 A cross-shaped exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
-%A point is only included if no component is likely to be driven by noise.
-%For the latter two methods, points 1 and 7 in Figure~\ref{fig:aatc_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:aatc_basic_4q_excl_cross}, respectively.
 
@@
     \centering
-%    \includegraphics{}
     \begin{subfigure}[t]{.48\textwidth}
@@
 
-%TC:ignore
 \section*{Declarations}
+\subsection*{Ethics approval and consent to participate}
+
+Not applicable.
+
+\subsection*{Consent for publication}
+
+Not applicable.
+
 \subsection*{Availability of data and materials}
 
-The datasets analyzed during the study are available in the corresponding databases or repositories \parencite{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024, Rostami-Tabar2023, Moody2017}.
-All code is available in the repository [GITHUB LINK WILL BE PROVIDED AFTER REVIEW].
+The datasets analyzed during the study are available in the corresponding databases or repositories (\url{https://covid19nowcasthub.de/}, \url{https://github.com/bahmanrostamitabar/hourly-emergency-care}, and \url{https://physionet.org/content/mimic3wdb/1.0/}).
+All code is available in the repository \url{https://github.com/jo-rie/aatc}.
 
@@
 \printbibliography
-%\bibliography{library,library-jonas}
 
@@
 \end{figure}
-%TC:endignore
 

### 2024-10-14 202410-rebuttal/main.tex
First draft for rebuttal

%% journalrebuttal.tex
%% Copyright 2020 Pranav Hosangadi
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License, either version 1.3
% of this license or (at your option) any later version.
% The latest version of this license is in
%   http://www.latex-project.org/lppl.txt
% and version 1.3 or later is part of all distributions of LaTeX
% version 2005/12/01 or later.
%
% This work has the LPPL maintenance status `maintained'.
% 
% The Current Maintainer of this work is Pranav Hosangadi.
%
% This work consists of the file journalrebuttal.cls and the demo
% file journalrebuttal.tex
%% 
%%
%% journalrebuttal.tex
%% Example usage for journalrebuttal.cls: 
%%    A LaTeX class to create rebuttal documents for journal
%%    reviews
%% Created: 2020-06-28
%% Author: Pranav Hosangadi (pranav.hosangadi@gmail.com)
%% https://github.com/pranavh/JournalRebuttal_LaTeX
%% Last Modified: 2020-06-29
%% Version: 1.0
%%
\documentclass[12pt]{journalrebuttal}
\title{On the assessment of the ability of measurements, nowcasts, and forecasts to track changes}
\author{Jonas Rieger*, Bolin Liu, Bernd Saugel, Oliver Grothe}
% TODO Authors anpassen
\journal{BMC Medical Research Methodology}
\manuscriptid{304c3895-aec8-4c97-867a-23905ca7f85d}

\usepackage{verbatim, csquotes}
\usepackage{amsmath, amssymb, mathtools,bm}

\definecolor{green}{rgb}{0.0, 0.5, 0.0}
\definecolor{blue}{rgb}{0.0, 0.5, 1.0}

%% You can define Note commands using the \ColorNote command 
%% provided in the class. 
\newcommand{\TODO}[1]{\ColorNote{green}{TODO}{#1}}
\newcommand{\IDEA}[1]{\ColorNote{blue}{Idea}{#1}}

\begin{document}
%\pagenumbering{arabic}
\maketitle

We thank the two reviewers for their comments and ideas for improving the manuscript. 
Please find our answers to the issues raised by the referees below.

\nextreviewer

\begin{revcomment}
	The reported ATC Ratio results would be improved if you include the raw number of positive and negative concordance pairs in your tables or table captions. You reference were an interested person could look at the raw data, but it would be helpful to have these summaries in the paper.
\end{revcomment}
\begin{response}
Thank you for your helpful suggestion. We have added the raw numbers of positive and negative concordance pairs to the table captions, providing readers with direct access to these important summaries within the paper itself.
\end{response}

\begin{revcomment}
I could see this paper being useful for forecasters and public health officials better communicate and understand the limitations of a particular model. I think it would be valuable to expand your discussion on how these metrics can be used in practice, particularly for forecasters and public health officials.
\end{revcomment}
\begin{response}
    Thank you for this valuable insight. We appreciate your perspective on the practical applications of our work. In response, we have expanded the discussion section to include more detailed commentary on how these metrics can be utilized in practice, particularly focusing on their relevance for forecasters and public health officials.
\end{response}

\begin{revcomment}
The literature review seems a little thin on references to concordance metrics for forecasts. It seems that references 21, 34, and 35 are the only one, but I am sure there are more out there.
\end{revcomment}
\begin{response}
    Thank you for bringing this to our attention. We agree that a more comprehensive literature review would ease further review for the readers. In response, we have expanded the number of references pertaining to concordance metrics for forecasts, providing a more thorough overview of the current state of research in this area.
\end{response}

\begin{revcomment}
    Related to the previous point, you could more clearly identify the new methods you are proposing versus what has already been explored in the literature. I think just changing some of the wording in Section 2 would help differentiate your work from other people's work in this area.
\end{revcomment}
\begin{response}
    We appreciate your suggestion for improving the clarity of our contribution. We have revised the wording in Section 2 to more distinctly highlight our proposed new methods and differentiate them from existing work in the literature.
\end{response}

\begin{revcomment}
On page 27, there is the line \enquote{the BCa is computationally expensive but requires smaller datasets for reasonable confidence intervals.} I don't believe this correct. The BCa method does not require smaller sample. It is more robust to small sample size and able to provide reasonable confidence intervals when the sample size is small.
\end{revcomment}
\begin{response}
    Thank you for this important clarification. You are correct, and we appreciate you pointing out this inaccuracy. We have revised the statement to more accurately reflect that the BCa method is more robust to small sample sizes and can provide reasonable confidence intervals even when the sample size is small, rather than requiring smaller datasets. The formulation has been made clearer as you suggested.
\end{response}

\nextreviewer

\begin{revcomment}
    Secondly, I suggest the authors elaborate on the importance of the ATC ratio, particularly its relevance to the specific data being analyzed. ATC radio is essentially one specific property of a distribution, and having better prediction accuracy for ATC radio does not automatically make it the superior model. The objects of interest could be mean, variance, quantile, expected shortfall, or even the entire distribution. It would be beneficial to explain why ATC is crucial for the three empirical studies. For instance, with Covid data, one could argue that tail performance (high infection rate or death toll) is more important for decision-makers.
\end{revcomment}
\begin{response}
    Thank you for this insightful comment. We agree that the ATC ratio is indeed one specific property of a distribution and not always the most critical measure. As we stated in the paper, we view it as an additional measure of interest that focuses on a particular aspect of performance. To address your suggestion, we have elaborated on why the ATC ratio adds value to the evaluation in each example's introduction, providing clearer context for its relevance alongside other important metrics such as tail performance.
\end{response}

\begin{revcomment}
    Finally, I recommend comparing the ATC-based evaluation methods with more traditional methods based on other statistical properties, like the log score, continuous ranked probability score, or quantile scores. It would be useful to discuss what additional insights the ATC-based approaches can offer compared to these traditional ones.
\end{revcomment}
\begin{response}
    We greatly appreciate your thoughtful recommendation. We agree that comparing ATC-based evaluation methods with traditional statistical approaches would provide valuable context. In response, we have added the log score, continuous ranked probability score, and quantile scores to the tables in the appendix, making these traditional metrics more explicit for comparison. While we have not included a theoretical assessment of the differences between these scores and our methods to maintain the practical focus of this paper, we recognize the importance of this point and plan to expand on it in a future, separate publication dedicated to this comparison.
\end{response}

\end{document}

### 2024-10-14 202410-rebuttal/main.tex
Modifications after review

@@
 As we stated in the paper, we view it as an additional measure of interest that focuses on a particular aspect of performance.
-To address your suggestion, we have elaborated on why the ATC ratio adds value to the evaluation in each example's introduction, providing clearer context for its relevance alongside other important metrics such as tail performance.
+To address your suggestion, we have elaborated on why the ATC ratio adds value to the evaluation in each example's introduction, providing clearer context for its relevance alongside other important metrics.
 \end{response}

### 2024-10-14 paper.tex
Modifications after review

@@
+% !TeX spellcheck = en_US
 \documentclass[pdflatex]{sn-jnl}
@@
 \abstract{
+Background: 
 Measurements, nowcasts, or forecasts ideally should correctly reflect changes in the values of interest.
 In this article, we focus on how to assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the ability to track changes.
-We present visual techniques and quantitative measures to assess the ability to track changes and provide extensions for noisy data and estimation uncertainty using bootstrap confidence intervals and exclusion areas.
+
+Methods: We developed visual techniques and quantitative measures to assess ATC. 
+Extensions for noisy data and estimation uncertainty were implemented using bootstrap confidence intervals and exclusion areas.
+
+Results: 
 We exemplarily illustrate the proposed methods to assess the ability to track changes for nowcasting during the COVID-19 pandemic, patient admissions to an emergency department, and non-invasive blood pressure measurements.
+The proposed methods effectively evaluated ATC across different applications.
+
+Conclusions: 
+The developed ATC assessment methods offer a comprehensive toolkit for evaluating the ATC of measurements, nowcasts, and forecasts. 
+These techniques provide valuable insights into model performance, complementing traditional accuracy measures and enabling more informed decision-making in various fields, including public health, healthcare management, and medical diagnostics.
 }
@@
 We exemplarily illustrate the proposed methods to assess the \ac{atc} for nowcasting during the COVID-19 pandemic, patient admissions to an emergency department, and non-invasive blood pressure measurements – and thus provide blueprints for future assessments.
-Ready-to-use code is available on [LINK WILL BE INSERTED].
+We discuss practical implementation and interpretation of the measures, thus providing the basis for communicating model limitations to forecasters, nowcasters, and public health officials.
+Ready-to-use code is available on \url{https://github.com/jo-rie/aatc}.
 
@@
 
-The assessment of \ac{atc} is based on the measured/observed/true and the predicted changes in a value of interest over a time horizon $l$.
+We base the assessment of \ac{atc} on the measured/observed/true and the predicted changes in a value of interest over a time horizon $l$.
 The \textit{true} change is straightforward to compute for all types of measurement, nowcast, or forecast.
@@
 Table~\ref{tab:notation} summarizes the notation for measurements, nowcasts, or forecasts and the computation of the predicted change.
+While the computation for measurements is straightforward and well-established, we develop the framework for nowcasts and forecasts in the following sections.
 For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$.
@@
 Consequently, when assessing the \ac{atc}, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
-A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{Perrino1998, Saugel2015}.
+A simple yet insightful method is the four-quadrant plot, which is well-established in measurement analysis and can be extended to nowcasts and forecasts~\citep[see, e.g., ][]{Perrino1998, Saugel2015}.
 In a four-quadrant plot, the occured changes and the predicted changes are plotted together, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
@@
 
-Analyzing the number of points in the green versus red quadrants is a standard approach in the \ac{atc} assessment of measurement data~\citep{Critchley2010, Saugel2015}.
+Analyzing the number of points in the green versus red quadrants is a standard approach in the \ac{atc} assessment of measurement data~\citep{Critchley2010, Saugel2015}, which we extend here to forecasts and nowcasts.
 With that, we estimate the probability of a correctly predicted change direction, $P(\diffxrv \diffyrv > 0)$, where $\diffyrv$ and $\diffxrv$ denote random variables for future incremental changes.
@@
 
-Confidence intervals can account for the estimation uncertainty of the measures above.
+Confidence intervals can account for the estimation uncertainty of the measures above, an approach not yet applied in the literature.
 Bootstrap confidence intervals are based on resampling and not on parametric assumptions as classical confidence intervals are~\parencite[for introductions see][]{Hesterberg2011, Bittmann2021}.
@@
 
-The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
+The estimators described above provide information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$, and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
+Notably, the visual analysis of these probabilities has not yet been addressed in the literature.
 A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the \ac{atc} of a prediction for a specific change $\diffxrv = \xcond$ of the measurement, nowcast, or forecast.
@@
 In nowcasting and forecasting, probabilistic predictions have become more prevalent in recent years (see Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
-In this section, we develop \ac{atc} assessments for probabilistic measurements and nowcasts.
+In this section, we develop \ac{atc} assessments for probabilistic measurements and nowcasts, an approach not yet explored in the literature.
 Probabilistic predictions issue a probability distribution for the quantity of interest based on their available information and, thus, include a point estimate and information on the prediction uncertainty and quantiles simultaneously.
@@
 
-To assess the impact of taken measures and the direction of the curve, the knowledge of whether hospitalization rates rise or fall is essential.
-Thus, the \ac{atc} assessment is particularly relevant for the nowcasts of the seven-day hospitalization rate.
-If hospitalization rates rise, measures should be tightened, while falling rates might allow for loosening measures.
-Asymmetries are especially relevant for assessing whether some models are better at recognizing a fall than a rise or vice versa.
+Traditional methods often focus on point or distributional accuracy of hospitalization rates, but understanding the direction and reliability of changes is crucial for effective decision-making.
+The \ac{atc} assessment offers easy-to-interpret insights and helps determine not only if rates are rising or falling, but also how confidently we can make this determination.
+For instance, if hospitalization rates are rising, public health measures may need to be tightened.
+Conversely, falling rates might justify loosening restrictions.
+The \ac{atc} assessment's ability to reveal asymmetries is especially valuable, as it can show whether certain models are more adept at recognizing decreases than increases, or vice versa.
+This information, which is not readily apparent from traditional methods, can significantly impact the interpretation of nowcasts and subsequent policy decisions for public health officials.
 
@@
 
-In this setup, \ac{atc} assessment is a simple and intuitive way to assess the models' performance.
-The \ac{atc} perspective is easy for the management to understand and implement, as simple comparisons of the expected workload to a recent shift can be made.
-If, for example, the staff was near capacity during the last shift and an increase in the number of patient admissions is expected, management can take measures to adjust the workload.
+While traditional methods provide point estimates, \ac{atc} assessment offers a simple, intuitive way to evaluate model performance and directional accuracy.
+This approach is particularly valuable for management, as it facilitates easy comparisons between expected workload and recent shifts.
+For instance, if staff was near capacity during the last shift, \ac{atc} assessment can clearly indicate whether an increase in patient admissions is likely, allowing management to proactively adjust resources.
+This directional insight, often hidden in conventional methods, enables more informed and timely decision-making.
 
@@
 
+While \ac{atc} assessment is well-established in measurement analysis, our new methods extend its application, revealing insights such as asymmetries. 
+This extension enhances the utility of \ac{atc} assessment beyond its conventional use, offering a more comprehensive evaluation of model performance.
+
 \subsubsection*{Results}
@@
 As in the application in Section~\ref{sec:application-covid}, models with medial point forecast evaluation measures can have the most meaningful positive \ac{atc}.
+
+For public health officials and forecasters analyzing methods, we recommend presenting four-quadrant plots, \ac{atc} ratios and the conditional \ac{atc} plots as a fundamental approach. 
+This presentation makes a model's \ac{atc} easily interpretable and and reveals performance asymmetries.
+An exclusion area can be incorporated to address for non-relevant changes.
 
@@
 The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\citep{Efron1987}.
-Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but requires smaller datasets for reasonable confidence intervals.
+Typically, the percentile approach requires larger datasets and provides an easy and fast estimate, while the \ac{bca}, though computationally expensive, can produce reasonable confidence intervals with smaller datasets.
 The basic approach balances these two objectives.

### 2024-10-14 202410-rebuttal/main.tex
Updates concerning programm tasks

@@
+% !TeX spellcheck = en_US
 %% journalrebuttal.tex
@@
 \begin{response}
-Thank you for your helpful suggestion.
-We have added the raw numbers of positive and negative concordance pairs to the table captions, providing readers with direct access to these important summaries within the paper itself.
+    We appreciate your thoughtful suggestion.
+To improve clarity while maintaining readability, we have added the numbers of positive and negative true changes to the table captions in the first two applications.
+This provides essential context without overloading the already dense tables.
+We believe this strikes a balance between offering key information and preserving the tables' accessibility, as including all raw numbers would significantly complicate their presentation.
 \end{response}
@@
     Thank you for this insightful comment.
-We agree that the ATC ratio is indeed one specific property of a distribution and not always the most critical measure.
+We agree that the ATC ratio is indeed one specific property of a distribution and not necessarily always the most critical measure.
 As we stated in the paper, we view it as an additional measure of interest that focuses on a particular aspect of performance.
@@
 \begin{response}
-    We greatly appreciate your thoughtful recommendation.
-We agree that comparing ATC-based evaluation methods with traditional statistical approaches would provide valuable context.
-In response, we have added the log score, continuous ranked probability score, and quantile scores to the tables in the appendix, making these traditional metrics more explicit for comparison.
-While we have not included a theoretical assessment of the differences between these scores and our methods to maintain the practical focus of this paper, we recognize the importance of this point and plan to expand on it in a future, separate publication dedicated to this comparison.
+Your suggestion highlights an important avenue for further research.
+While we agree that comparing ATC-based evaluation methods with traditional statistical approaches would be valuable, we've chosen to maintain the practical focus of this paper.
+The current empirical study is already comprehensive, and including a full analysis of all these scores would significantly extend the length and shift the paper's emphasis.
+However, we recognize the importance of this comparison and are planning a separate publication dedicated to a theoretical analysis of ATC assessment in relation to traditional scores.
+This future work will allow us to explore these connections in the depth they require.
 \end{response}

### 2024-10-14 paper.tex
Updates concerning programm tasks

@@
 The subscript $q_{0.1}$ denotes the measures with exclusion area.
-}
+There are 66 positive and 93 negative actual changes in the considered time period.}
     \label{tab:app-covid-atc-ratios-lag-7}
@@
     \input{plots/ed_arrival/50_trending_ratio}
-    \caption[ATC ratios for the forecasts of patient admission to an emergency department.]{\Ac{atc} ratio $\acc_+$, positive \ac{atc} ratio $\accp_+$, and negative \ac{atc} ratio $\accm_+$ for the models for the horizons 72 hours and seven days in the forecasting of patient admissions to an emergency department with an exclusion zone equal to the axis indicated by the subscript $+$.}
+    \caption[ATC ratios for the forecasts of patient admission to an emergency department.]{\Ac{atc} ratio $\acc_+$, positive \ac{atc} ratio $\accp_+$, and negative \ac{atc} ratio $\accm_+$ for the models for the horizons 72 hours and seven days in the forecasting of patient admissions to an emergency department with an exclusion zone equal to the axis indicated by the subscript $+$.
+There are 4030 positive changes and 4051 negative changes for the horizon of 72 hours and 4042 and 3961 for the horizon of 7 days.}
     \label{tab:app-eda-atc-ratios}
@@
 The exclusion areas are rectangles centered on the zero points with a width and height to exclude the 10\%-quantile of the absolute values of nowcast or true values.
-}
+There are 75 positive and 84 negative actual changes in the considered time period for horizon one day and 73 and 86 for the horizon of 14 days.}
     \label{tab:app-covid-atc-ratios-lag-1-14}

### 2024-10-19 202410-rebuttal/main.tex
Update on Overleaf.

@@
 % !TeX spellcheck = en_US
-%% journalrebuttal.tex
-%% Copyright 2020 Pranav Hosangadi
-%
-% This work may be distributed and/or modified under the
-% conditions of the LaTeX Project Public License, either version 1.3
-% of this license or (at your option) any later version.
-% The latest version of this license is in
-%   http://www.latex-project.org/lppl.txt
-% and version 1.3 or later is part of all distributions of LaTeX
-% version 2005/12/01 or later.
-%
-% This work has the LPPL maintenance status `maintained'.
-% 
-% The Current Maintainer of this work is Pranav Hosangadi.
-%
-% This work consists of the file journalrebuttal.cls and the demo
-% file journalrebuttal.tex
-%% 
-%%
-%% journalrebuttal.tex
-%% Example usage for journalrebuttal.cls: 
-%%    A LaTeX class to create rebuttal documents for journal
-%%    reviews
-%% Created: 2020-06-28
-%% Author: Pranav Hosangadi (pranav.hosangadi@gmail.com)
-%% https://github.com/pranavh/JournalRebuttal_LaTeX
-%% Last Modified: 2020-06-29
-%% Version: 1.0
-%%
+
 \documentclass[12pt]{journalrebuttal}
@@
 We thank the two reviewers for their comments and ideas for improving the manuscript. 
-Please find our answers to the issues raised by the referees below.
+Please find our answers to the issues raised by the referees in the following.
 
@@
 \begin{response}
-    Thank you for this valuable insight.
-We appreciate your perspective on the practical applications of our work.
+    Thank you for this valuable suggestion.
 In response, we have expanded the discussion section to include more detailed commentary on how these metrics can be utilized in practice, particularly focusing on their relevance for forecasters and public health officials.
@@
 \begin{response}
-    Thank you for bringing this to our attention.
-We agree that a more comprehensive literature review would ease further review for the readers.
-In response, we have expanded the number of references pertaining to concordance metrics for forecasts, providing a more thorough overview of the current state of research in this area.
+    Thank you for sharing this observation.
+In response, we have expanded the number of references related to concordance metrics for forecasts, providing a more thorough overview of the current state of research in this area.
 \end{response}
@@
 We have revised the statement to more accurately reflect that the BCa method is more robust to small sample sizes and can provide reasonable confidence intervals even when the sample size is small, rather than requiring smaller datasets.
-The formulation has been made clearer as you suggested.
 \end{response}
@@
 As we stated in the paper, we view it as an additional measure of interest that focuses on a particular aspect of performance.
-To address your suggestion, we have elaborated on why the ATC ratio adds value to the evaluation in each example's introduction, providing clearer context for its relevance alongside other important metrics.
+To address your suggestion, we have elaborated on why the ATC ratio adds value to the evaluation in each example's introduction.
 \end{response}
@@
 While we agree that comparing ATC-based evaluation methods with traditional statistical approaches would be valuable, we've chosen to maintain the practical focus of this paper.
-The current empirical study is already comprehensive, and including a full analysis of all these scores would significantly extend the length and shift the paper's emphasis.
-However, we recognize the importance of this comparison and are planning a separate publication dedicated to a theoretical analysis of ATC assessment in relation to traditional scores.
+The current empirical study is already comprehensive, and including a full analysis of all these scores would significantly extend the length of the paper and shift its emphasis.
+However, we recognize the importance of this comparison and plan a separate publication dedicated to a theoretical analysis of ATC assessment in relation to traditional scores.
 This future work will allow us to explore these connections in the depth they require.

### 2024-10-19 paper.tex
Update on Overleaf.

@@
 
-Methods: We developed visual techniques and quantitative measures to assess ATC. 
+Methods: We reviewed and developed visual techniques and quantitative measures to assess ATC. 
 Extensions for noisy data and estimation uncertainty were implemented using bootstrap confidence intervals and exclusion areas.
@@
 
-In forecasting and nowcasting the evaluation of performance or goodness is usually based on statistical methods quantifying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}.
-None of them assesses the method's \ac{atc}.
+In forecasting and nowcasting the evaluation of performance or goodness is usually based on statistical methods quantifying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}, see also~\citep{SteyerbergEwoutW.2010AtPo} for an overview in the medical context together with less known measures.
+However, although techniques like ROC curves for binary direction models or more general concordance measures (e.g., \citep{Goenen2005}) are used, there is no general measure directly framed to assesses the method's \ac{atc}.
 In a forecasting competition on armed conflicts, the assessment of the \ac{atc} recently gained attention as \citet{Vesco2022} proposed the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-When evaluating the performance or goodness of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
-How to best assess the \ac{atc} of measurement methods is a field of ongoing research~\citep{Saugel2015, Saugel2018, Hiraishi2021}.
+When evaluating the performance or goodness of a measurement method like in classical mathod comparation studies, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
+In this strand of literature,  the best way to assess the \ac{atc} of measurement methods is a field of active ongoing research~\citep{Critchley20101180, CRITCHLEY2011536, Saugel2015, Saugel2018, Hiraishi2021}.
 

### 2024-10-20 202410-rebuttal/main.tex
Fix typos, update rebuttal

@@
 \author{Jonas Rieger*, Bolin Liu, Bernd Saugel, Oliver Grothe}
-% TODO Authors anpassen
 \journal{BMC Medical Research Methodology}
@@
     Thank you for sharing this observation.
-In response, we have expanded the number of references related to concordance metrics for forecasts, providing a more thorough overview of the current state of research in this area.
+In response, we have expanded the number of references related to concordance metrics for forecasts in the introduction, providing a more thorough overview of the current state of research in this area.
 \end{response}

### 2024-10-20 paper.tex
Fix typos, update rebuttal

@@
 
-In forecasting and nowcasting the evaluation of performance or goodness is usually based on statistical methods quantifying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}, see also~\citep{SteyerbergEwoutW.2010AtPo} for an overview in the medical context together with less known measures.
-However, although techniques like ROC curves for binary direction models or more general concordance measures (e.g., \citep{Goenen2005}) are used, there is no general measure directly framed to assesses the method's \ac{atc}.
+In forecasting and nowcasting the evaluation of performance or goodness is usually based on statistical methods quantifying the accuracy such as the \ac{rmse}, probabilistic scoring rules, and calibration measures~\citep{Gneiting2007, Gunther2021, Wolffram2023}, see also~\citet{SteyerbergEwoutW.2010AtPo} for an overview in the medical context together with less known measures.
+However, although techniques like ROC curves for binary direction models or more general concordance measures \citep[see, e.g., ][]{Goenen2005} are used, there is no general measure directly framed to assess the method's \ac{atc}.
 In a forecasting competition on armed conflicts, the assessment of the \ac{atc} recently gained attention as \citet{Vesco2022} proposed the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
 However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
-When evaluating the performance or goodness of a measurement method like in classical mathod comparation studies, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
+When evaluating the performance or goodness of a measurement method like in classical method comparison studies, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
 In this strand of literature, the best way to assess the \ac{atc} of measurement methods is a field of active ongoing research~\citep{Critchley20101180, CRITCHLEY2011536, Saugel2015, Saugel2018, Hiraishi2021}.
